### 2023-04-02 paper.tex
move repo

% interactcadsample.tex
% v1.03 - April 2017

\documentclass[]{interact}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage{subfigure}% Support for small, `sub' figures and tables
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required

\usepackage{natbib}% Citation support using natbib.sty
\bibpunct[, ]{(}{)}{;}{a}{}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{lscape}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}

\articletype{ARTICLE TEMPLATE}

\title{A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics
with the Lineup Protocol}

\author{\name{Weihao Li$^{a}$, Dianne Cook$^{a}$, Emi
Tanaka$^{a}$, Susan VanderPlas$^{b}$}
\affil{$^{a}$Department of Econometrics and Business Statistics, Monash
University, Clayton, VIC, Australia; $^{b}$Department of Statistics,
University of Nebraska, Lincoln, Nebraska, USA}
}

\thanks{CONTACT Weihao
Li. Email: \href{mailto:weihao.li@monash.edu}{\nolinkurl{weihao.li@monash.edu}}, Dianne
Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}, Emi
Tanaka. Email: \href{mailto:emi.tanaka@monash.edu}{\nolinkurl{emi.tanaka@monash.edu}}, Susan
VanderPlas. Email: \href{mailto:susan.vanderplas@unl.edu}{\nolinkurl{susan.vanderplas@unl.edu}}}

\maketitle

\begin{abstract}
Abstract to fill.
\end{abstract}

\begin{keywords}
statistical graphics; data visualization; visual inference; hypothesis
testing; reression analysis; cognitive perception; simulation; practical
significance; effect size
\end{keywords}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\begin{quote}
\emph{``Since all models are wrong the scientist must be alert to what
is importantly wrong.''} \citep{box1976science}
\end{quote}

Diagnostics are the key to determining whether there is anything
importantly wrong with a model. In linear regression analysis, residuals
from the model fit are commonly used. Residuals summarise what is not
captured by the model, and thus provide the capacity to identify what
might be wrong.

We can assess residuals in multiple ways. Residuals may be plotted as a
histogram or quantile-quantile plot to examine the distribution. Using
the classical normal linear regression model as an example, if the
distribution is symmetric and unimodal, we consider it to be well
behaved. However, if the distribution is skewed, bimodal, multimodal, or
contains outliers, there is a cause for concern. One could also inspect
the distribution by conducting a goodness of fit test, such as the
Shapiro-Wilk Normality test \citep{shapiro1965analysis}.

More typically, residuals will be plotted, as a scatter plot against the
predicted values and each of the explanatory variables to scrutinize
their relationships. If there are any visually discoverable patterns,
the model is potentially misspecified. In general, one looks for
noticeable departures from the model like non-linear dependency or
heteroskedasticity. However, correctly judging a residual plot where no
pattern exists can be a painstakingly difficult task for humans. It is
especially common, particularly among students, to misinterpret patterns
that are random noise and random deviation from a model
\citep{loy2021bringing}. It is also possible to conduct hypothesis tests
for non-linear dependence \citep{ramsey_tests_1969}, and use a
Breusch-Pagan test \citep{breusch_simple_1979} for heteroskedasticity.

Abundance of literature describe appropriate diagnostic methods for
linear regression, e.g. \citet{draper1998applied},
\citet{montgomery1982introduction}, \citet{belsley_regression_1980},
\citet{cook_applied_1999} and \citet{cook1982residuals}. All these
writings consider plotting residuals to be a standard technique that
should be examined routinely in all regression modelling problems. In
addition, \citet{draper1998applied} and \citet{belsley_regression_1980}
believe that residual plots are usually revealing when the assumptions
are violated. \citet{cook_applied_1999} thinks formal tests and
graphical procedures are complementary and both have a place in residual
analysis, but they focus on graphical methods rather than on formal
testing, as they are easier to use. \citet{montgomery1982introduction}
even suggests that residual plots are more informative in most practical
situations than the corresponding formal tests, and statistical tests on
regression model residuals are not widely used based on their
experience.

A common guidance by experts is that optimal method for diagnosing model
fits is by plotting the data. The persistence of this advice to check
the plots is curious, and investigating why this might be common advice
is the subject of this paper. The paper is structured as follows. The
next section describes the background on the types of departures that
one expects to detect, and outlines a formal statistical process for
reading residual plots, called visual inference. Section
\ref{experimental-design} details the experimental design to compare the
decisions made by formal hypothesis testing, and how humans would read
diagnostic plots. The results are reported in Section \ref{results}. We
conclude with a discussion of future work, in particular, how the
responsibility for residual plot reading might be relieved.

\hypertarget{background}{%
\section{Background}\label{background}}

\hypertarget{departures-from-good-residual-plots}{%
\subsection{Departures from good residual
plots}\label{departures-from-good-residual-plots}}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/residual-plot-common-departures-1}

}

\caption{Example residual vs fitted value plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0. The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted value plot.}\label{fig:residual-plot-common-departures}
\end{figure}

Graphical summaries in which residuals are plotted against fitted values
or other functions of the predictors that are approximately orthogonal
to the residuals are referred to as standard residual plots in
\citet{cook1982residuals}. Figure
\ref{fig:residual-plot-common-departures}A shows an ideal residual plot
where the residuals are evenly distributed at both sides of the
horizontal zero line, with no noticeable patterns.

There are various types of departures from an ideal residual plot.
Non-linearity, heteroskedasticity and non-normality are perhaps the
three mostly checked departures.

Non-linearity is a type of model misspecification caused by failing to
include higher order terms of the predictors in the regression equation.
Any non-linear functional form of residuals on fitted values in the
residual plot could be indicative of non-linearity. An example residual
plot containing visual pattern of non-linearity is shown in Figure
\ref{fig:residual-plot-common-departures}B. One can clearly observe the
``S-shape'' from the residual plot as the cubic term is not captured by
the misspecified model.

Heteroskedasticity refers to the presence of nonconstant error variance
in a regression model. It is mostly due to the strict but false
assumptions on the variance-covariance matrix of the error term. The
usual pattern of heteroskedasticity on a residual plot is the
inconsistent spread of the residuals across the horizontal axis.
Visually, it sometimes results in the so-called ``butterfly'' shape as
shown in Figure \ref{fig:residual-plot-common-departures}C, or the
``left-triangle'' and ``right-triangle'' shape where the smallest
variance occurs at one side of the horizontal axis.

Compared to non-linearity and heteroskedasticity, non-normality is
usually harder to detect from a residual plot since a scatter plot do
not readily reveal the marginal distribution. A favourable graphical
summary for this task is the quantile-quantile plot. As we mainly
discuss residual plots, non-normality will not be the focus of this
paper. For a consistent comparison, the residual plot of this departure
is still presented in Figure \ref{fig:residual-plot-common-departures}D.
When the number residuals below and above the horizontal axis are uneven
across the local regions along the \(x\)-axis, we expect that the
normality assumption is violated. For example, given a skewed error
distribution, there will be fewer data points and more outliers on one
side of the horizontal axis as shown in Figure
\ref{fig:residual-plot-common-departures}D.

\hypertarget{conventionally-testing-for-departures}{%
\subsection{Conventionally testing for
departures}\label{conventionally-testing-for-departures}}

Other than checking diagnostic plots, analysts may perform formal
hypothesis testing to detect model defects. A variety of tests can be
applied; each testing for a specific violation of the null hypothesis.
For example, the presence of heteroskedasticity can usually be tested by
applying the White test \citep{white_heteroskedasticity-consistent_1980}
or the Breusch-Pagan test \citep{breusch_simple_1979}, which are both
derived from the Lagrange multiplier test \citep{silvey1959lagrangian}
principle that relies on the asymptotic properties of the null
distribution. To test specific forms of non-linearity, one may apply the
F-test as a model structural test to examine the significance of
specific polynomial and non-linear forms of the predictors, or the
significance of proxy variables as in the Ramsey Regression Equation
Specification Error Test (RESET) \citep{ramsey_tests_1969}. The
Shapiro-Wilk test \citep{shapiro1965analysis} is the most widely used
test of non-normality included by many of the statistical software
programs. The Jarque--Bera test \citep{jarque1980efficient} is also used
to directly check whether the sample skewness and kurtosis match a
normal distribution.

We apply the RESET test, Breusch-Pagan test and Shapiro-Wilk test to the
residual plots shown in Figure
\ref{fig:residual-plot-common-departures}; the results of this is shown
in Table \ref{tab:example-residual-plot-table}. The Breusch-Pagan test
and the Shapiro-Wilk test both reject \(H_0\) for residual plots that
display non-linearity and heteroscedasticity, even though these are not
the original intention of the test. As discussed in
\citet{cook1982residuals}, most residual-based tests for a particular
type of departure from model assumptions are also sensitive to other
types of departures. It is likely \(H_0\) is correctly rejected but for
the wrong reason, a phenomenon known as the ``Type III error''.
Additionally, outliers will often incorrectly trigger the rejection of
\(H_0\) despite when majority of the residuals are well-behaved
\citep{cook_applied_1999}. Furthermore, with a sufficiently large sample
size, residual-based tests may reject \(H_0\) due to a slight departure
that is of little practical significance. These can be largely avoided
in diagnostic plots as experienced analysts can evaluate the
acceptability of assumptions flexibly, even in the presence of outliers
and slight departures.

\begin{table}

\caption{\label{tab:example-residual-plot-table}Statistical significance testing for departures from good residuals for plots in Figure \ref{fig:residual-plot-common-departures}. Shown are the $p$-values calculated for the RESET, the Breusch-Pagan and the Shapiro–Wilk tests. The good residual plot (A) is judged a good residual plot, as expected, by all tests. The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.}
\centering
\begin{tabular}[t]{llrrr}
\toprule
Plot & Departures & RESET & Breusch-Pagan & Shapiro–Wilk\\
\midrule
A & None & 0.779 & 0.133 & 0.728\\
B & Non-linearity & \em{0.000} & \em{0.000} & \em{0.039}\\
C & Heteroskedasticity & 0.658 & \em{0.000} & \em{0.000}\\
D & Non-normality & 0.863 & 0.736 & \em{0.000}\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{visual-test-procedure-based-on-lineups}{%
\subsection{Visual test procedure based on
lineups}\label{visual-test-procedure-based-on-lineups}}

One may argue that reading diagnostic plots is to some extent subjective
and indecisive compared to those rigorous statistical procedures as it
relies on graphical perception - human ability to interpret and decode
the information embedded in graph \citep{cleveland_graphical_1984}.
Further, the degree of the presence of the visual features typically can
not be measured quantitatively and objectively, which may lead to over
or under-interpretations of the data. For instance, people
over-interpret the separation between gene groups in a two-dimensional
projection from a linear discriminant analysis when in fact there are no
differences in the expression levels between the gene groups and
separation is not an uncommon occurrence
\citep{roy_chowdhury_using_2015}.

Visual inference was first introduced in a 1999 Joint Statistical
Meetings (JSM) talk with the title ``Inference for Data Visualization''
by \citet{buja_inference_1999} as an idea to address the issue of valid
inference for visual discoveries of data plots. Later,
\citet{buja_statistical_2009} proposed the lineup protocol as a visual
test inspired by the ``police lineup'' or ``identity parade'' which is
the act of asking the eyewitness to identify criminal suspect from a
group of irrelevant people. The protocol consists of \(m\) randomly
placed plots, where one plot is the data plot, and the remaining
\(m - 1\) null plots have the identical graphical procedure except the
data has been replaced with data consistent with \(H_0\), that the model
is correctly specified. Then, an observer who have not seen the data
plot will be asked to point out the most different plot from the lineup.
Under \(H_0\), it is expected that the data plot would have no
distinguishable difference from the null plots, and the probability that
the observer correctly picks the data plot is \(1/m\). If one rejects
\(H_0\) as the observer correctly picks the data plot, then the Type I
error of this test is \(1/m\).

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/first-example-lineup-1}

}

\caption{Visual testing is conducted using a lineup, as in the example here. The residual plot computed from the observed data (plot $2^2 + 2$, exhibiting non-linearity) is embedded among 19 null plots, where the residuals are simulated from a standard error model. Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot. A small $p$-value would result from a substantial number selecting plot $2^2 + 2$.}\label{fig:first-example-lineup}
\end{figure}

Figure \ref{fig:first-example-lineup} is an example of a lineup
protocol. If the data plot at position \(2^2 + 2\) is identifiable, then
it is evidence for the rejection of \(H_0\). In fact, the actual
residual plot is obtained from a misspecified regression model with
missing non-linear terms.

Data used in the \(m - 1\) null plots needs to be simulated. In
regression diagnostics, sampling data consistent with \(H_0\) is
equivalent to sampling data from the assumed model. As
\citet{buja_statistical_2009} suggested, \(H_0\) is usually a composite
hypothesis controlled by nuisance parameters. Since regression models
can have various forms, there is no general solution to this problem,
but it sometimes can be reduced to a so called ``reference
distribution'' by applying one of the three methods: (i) sampling from a
conditional distribution given a minimal sufficient statistic under
\(H_0\), (ii) parametric bootstrap sampling with nuisance parameters
estimated under \(H_0\), and (iii) Bayesian posterior predictive
sampling. The conditional distribution given a minimal sufficient
statistic is the best justified reference distribution among the three
\citep{buja_statistical_2009}. Essentially, null residuals can be
simulated by regressing \(N\) i.i.d standard normal random draws on the
predictors, then rescaling it by the ratio of residual sum of square in
two regressions.

The effectiveness of lineup protocol for regression analysis is
validated by \citet{majumder_validation_2013} under relatively simple
settings with up to two predictors. Their results suggest that visual
tests are capable of testing the significance of a single predictor with
a similar power as a t-test, though they express that in general it is
unnecessary to use visual inference if there exists a conventional test,
and they do not expect the visual test to perform equally well as the
conventional test. In their third experiment, where there is not a
conventional test, visual test outperforms the conventional test for a
large margin. This is encouraging, as it promotes the use of visual
inference in situations where there are no existing statistical testing
procedures. Visual inference have also been integrated into diagnostic
of hierarchical linear models by \citet{loy2013diagnostic},
\citet{loy2014hlmdiag} and \citet{loy2015you}. They use lineup protocols
to judge the assumption of linearity, normality and constant error
variance for both the level-1 and level-2 residuals.

\hypertarget{calculation-of-statistical-significance-and-test-power}{%
\section{Calculation of statistical significance and test
power}\label{calculation-of-statistical-significance-and-test-power}}

\hypertarget{what-is-being-tested}{%
\subsection{What is being tested?}\label{what-is-being-tested}}

In diagnosing a model fit from residuals, we are generally interested in
\(H_0:\) \emph{The regression model is correctly specified} against the
broad alternative \(H_a:\) \emph{The regression model is misspecified}.

However, it is practically impossible to test this specific \(H_0\) with
conventional tests, which are constructed to measure specific
departures. For example, the RESET test is formulated as
\(H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0\) against
\(H_a: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0\),
from
\(y = \tau_0 + \sum_{i=1}^{p}\tau_px_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)\).
Similarly, the Breusch-Pagan test is designed to specifically test
\(H_0:\) \emph{error variances are all equal}
(\(\zeta_i=0 \text{ for } i=1,..,p\)) versus the alternative \(H_a:\)
\emph{that the error variances are a multiplicative function of one or
more variables} (\(\text{at least one } \zeta_i\neq 0\)) from
\(e^2 = \zeta_0 + \sum_{i=1}^{p}\zeta_i x_i + u, ~ u\sim N(0,\sigma_u^2)\).

One of the potential benefits of the visual test, based on the lineup
protocol, is that it should be able to detect a range of departures from
good residuals.

\hypertarget{statistical-significance}{%
\subsection{Statistical significance}\label{statistical-significance}}

In hypothesis testing, a \(p\)-value is defined as the probability of
observing test results as least as extreme as the observed result given
\(H_0\) is true. Conventional hypothesis tests usually have an existing
method to derive or compute \(p\)-value based on the null distribution.
What we need to discuss in the following is the method to estimate
\(p\)-value for a visual test.

Within the context of visual inference, by involving \(k\) independent
observers, the visual \(p\)-value can be interpreted as the probability
of having as many or more subjects detect the data plot than the
observed result.

Let \(X_j = \{0,1\}\) be a Bernoulli random variable denoting whether
subject \(j\) correctly detecting the data plot, and
\(X = \sum_{j=1}^{K}X_j\) be the number of observers correctly picking
the data plot. Then, by imposing a relatively strong assumption on the
visual test that all \(K\) evaluations are fully independent, under
\(H_0\), \(X \sim \mathrm{Binom}_{K,1/m}\). Therefore, the \(p\)-value
of a lineup of size \(m\) evaluated by \(K\) observer is given as
\(P(X \geq x) = 1 - F(x) + f(x)\), where \(F(.)\) is the binomial
cumulative distribution function, \(f(.)\) is the binomial probability
mass function and \(x\) is the realization of number of observers
correctly picking the data plot \citep{majumder_validation_2013}.

As pointed out by \citet{vanderplas2021statistical}, this basic binomial
model does not take into account the possible dependencies in the visual
test due to repeated evaluations of the same lineup. And it is
inapplicable to visual test where subjects are asked to select one or
more ``most different'' plots from the lineup.
\citet{vanderplas2021statistical} summarises three common scenarios in
visual inference: (1) \(K\) different lineups are shown to \(K\)
subjects, (2) \(K\) lineups with different null plots but the same data
plot are shown to \(K\) subjects, and (3) the same lineup is shown to
\(K\) subjects. Out of these three scenarios, Scenario 3 is the most
common in previous studies as it puts the least constraints on the
experiment design. For Scenario 3, \citet{vanderplas2021statistical}
models the probability of a plot \(i\) being selected from a lineup as
\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
\(i=1,...,m\) and \(\alpha > 0\). The number of times plot \(i\) being
selected in \(K\) evaluations is denoted as \(c_i\). In case subject
\(j\) makes multiple selections, \(1/s_j\) will be added to \(c_i\)
instead of one, where \(s_j\) is the number of plots subject \(j\)
selected for \(j=1,...K\). This ensures \(\sum_{i}c_i=K\). Since we are
only interested in the selections of the data plot \(i\), the marginal
model can be simplified to a beta-binomial model and thus the visual
\(p\)-value is given as

\begin{equation} \label{eq:pvalue-beta-binomial}
P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+
\end{equation}

\noindent where \(B(.)\) is the beta function defined as

\begin{equation} \label{eq:betafunction}
B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0.
\end{equation}

Note that Equation \ref{eq:pvalue-beta-binomial} given in
\citet{vanderplas2021statistical} only works with non-negative integer
\(c_i\). We extend the equation to non-negative real number \(c_i\) by
applying a linear approximation

\begin{equation} \label{eq:pvalue-beta-binomial-approx}
P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+,
\end{equation}

where \(P(C \geq \lceil c_i \rceil)\) is calculated using Equation
\ref{eq:pvalue-beta-binomial} and \(P(C = \lfloor c_i \rfloor)\) is
calculated by

\begin{equation} \label{eq:pmf-beta-binomial}
P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+.
\end{equation}

Besides, the parameter \(\alpha\) used in Equation
\ref{eq:pvalue-beta-binomial} and \ref{eq:pmf-beta-binomial} is usually
unknown and hence needs to be estimated from the survey data. For low
values of \(\alpha\), only a few plots are attractive to the observers
and tend to be selected. For higher values of \(\alpha\), the
distribution of the probability of each plot being selected is more
even. \citet{vanderplas2021statistical} defines that a plot is
\(c\)-interesting if \(c\) or more participants select the plot as the
most different. Given the definition, The expected number of plots
selected at least \(c\) times, \(E[Z_c]\), is calculated as

\begin{equation} \label{eq:c-interesting-expectation}
E[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}

With Equation \ref{eq:c-interesting-expectation}, \(\alpha\) can be
estimated using maximum likelihood estimation. But for precise estimate
of \(\alpha\), additional human responses to Rorschach lineups, which is
a type of lineup that consists of plots constructed from the same null
data generating mechanism, are required.

\hypertarget{power-of-the-tests}{%
\subsection{Power of the tests}\label{power-of-the-tests}}

The power of a model misspecification test is the probability that
\(H_0\) is rejected given the regression model is misspecified in a
specific way. It is an important indicator when one is concerned about
whether model assumptions have been violated. Although in practice, one
might be more interested in knowing how much the residuals deviate from
the model assumptions, and whether this deviation is of practical
significance.

The power of a conventional hypothesis test is affected by both the true
parameter \(\boldsymbol{\theta}\) and the sample size \(n\). These two
can be quantified in terms of effect size \(E\) to measure the strength
of the residual departures from the model assumptions. Details about the
effect size is provided in Section \ref{effect-size} after the
introduction of the simulation model used in our human subject
experiment. The theoretical power of a test is sometimes not a trival
solution, but it can be estimated if the data generating process is
known. We use a predefined model to generate a large set of simulated
data under different effect sizes, and record if the conventional test
rejects \(H_0\). The probability of the conventional test rejects
\(H_0\) is then fitted by a logistic regression formulated as

\begin{equation} \label{eq:logistic-regression-1-1}
Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),
\end{equation}

\noindent where \(\Lambda(.)\) is the standard logistic function given
as \(\Lambda(z) = exp(z)/(1+exp(z))\). The effect size \(E\) is the only
predictor and the intercept is fixed to \(log(0.05/0.95)\) so that
\(\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05\), which is the desired
significance level.

The power of a visual test on the other hand, may additionally depend on
the ability of the particular subject, as the skill of the individual
may affect the number of observers who identify the data plot from the
lineup \citep{majumder_validation_2013}. To address this issue,
\citet{majumder_validation_2013} models the probability of a subject
\(j\) correctly picking the data plot from a lineup \(l\) using a
mixed-effect logistic regression, with subjects treated as random
effect. Then, the estimated power of a visual test evaluated by a single
subject is the predicted value obtained from the mix-effect model.
However, this mix-effect model does not work with scenario where
subjects are asked to select one or more most different plots. In this
scenario, having the probability of a subject \(j\) correctly picking
the data plot from a lineup \(l\) is insufficient to determine the power
of a visual test because it does not provide information about the
number of selections made by the subject for the calculation of the
\(p\)-value (See Equation \ref{eq:pvalue-beta-binomial-approx}).
Therefore, we directly estimate the probability of a lineup being
rejected by assuming that individual skill has negligible effect on the
variation of the power. This assumption is not necessary true, but it
helps simplifying the model structure, thereby obviate a costly
large-scale experiment to estimate complex covariance matrices. The same
model given in Equation \ref{eq:logistic-regression-1-1} is applied to
model the power of a visual test.

To study various factors contributing to the power of both tests, the
same logistic regression model is fit on different subsets of the
collated data grouped by levels of factors. These include the
distribution of the fitted values, type of the simulation model and the
shape of the residual departures.

\hypertarget{experimental-design}{%
\section{Experimental design}\label{experimental-design}}

An experiment is conducted in three data collection periods to
investigate the difference between conventional hypothesis testing and
visual inference in the application of linear regression diagnostics.
Two types of departures, non-linearity and heteroskedasticity, are
collected during data collection periods I and II. The data collection
period III was designed primarily to measure human responses to null
lineups so that the parameter \(\alpha\) in Equation
\ref{eq:pvalue-beta-binomial} can be estimated. Additional lineups for
both non-linearity and heteroskedasticity, using uniform fitted value
distribution, were included so that the participants were evaluating
some lineups with signal also. It would be too frustrating for
participants to only be assigned lineups with all null plots. Overall,
we collected 7974 evaluations on 1152 unique lineups performed by 443
subjects throughout three data collection periods.

\hypertarget{simulating-departures-from-good-residuals}{%
\subsection{Simulating departures from good
residuals}\label{simulating-departures-from-good-residuals}}

\hypertarget{non-linearity}{%
\subsubsection{Non-linearity}\label{non-linearity}}

Data collection period I is designed to study the ability of human
subjects to detect the effect of a non-linear term \(\boldsymbol{z}\)
constructed using Hermite polynomials on random vector
\(\boldsymbol{x}\) formulated as

\begin{align} \label{eq:nonlinearity-model}
\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
\end{align}

\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
\(\boldsymbol{\varepsilon}\), \(\boldsymbol{x}_{raw}\),
\(\boldsymbol{z}_{raw}\) are vectors of size \(n\), \(He_{j}(.)\) is the
\(j\)th-order probabilist's Hermite polynomials,
\(\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\), and
\(g(\boldsymbol{x}, k)\) is a scaling function to enforce the support of
the random vector to be \([-k, k]^n\) defined as

\begin{equation} \label{eq:scaling-function}
g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x}))2k - k, \quad \text{for} \quad k > 0. 
\end{equation}

According to \citet{abramowitz1964handbook}, Hermite polynomials were
initially defined by \citet{de1820theorie}, but named after Hermite
\citep{hermite1864nouveau} because of the unrecognisable form of
Laplace's work. When simulating \(\boldsymbol{z}_{raw}\), function
\texttt{hermite} from the R package \texttt{mpoly} \citep{mpoly} is used
to generate Hermite polynomials.

The null regression model used to fit the realizations generated by the
above model is formulated as

\begin{equation} \label{eq:null-model}
\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
\end{equation}

\noindent where
\(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\).

Since \(z = O(x^j)\), for \(j > 1\), \(z\) is a higher order term leaves
out by the null regression, which will lead to model misspecification.

Visual patterns of non-linearity are simulated using four different
orders of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)). (A
summary of the factors is given in Table \ref{tab:model-factor-table}.)
The values of \(j\) is chosen so that distinct shapes of non-linearity
are included in the residual plot. These include ``U'', ``S'', ``M'' and
``triple-U'' shape as shown in Figure
\ref{fig:different-shape-of-herimite}. A greater value of \(j\) will
result in a curve with more turning points. It is expected that the
``U'' shape will be the easiest one to detect because complex shape
tends to be concealed by cluster of data points.

Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups
used in non-linearity detection. This lineup is produced by the
non-linearity model with \(j = 6\). The data plot location is
\(2^3 - 4\). All five subjects correctly identify the data plot from
this lineup.

\begin{table}

\caption{\label{tab:model-factor-table}Levels of the factors used in data collection periods I, II, III.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{rr|rr|rcrr|rr|rcrr|rr|rcrr|rr|rcrr|rr|rcrr|rr|rc}
\toprule
\multicolumn{2}{c}{Non-linearity} & \multicolumn{2}{c}{Heteroskedasticity} & \multicolumn{2}{c}{Common} \\
\cmidrule(l{3pt}r{3pt}){1-2} \cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6}
\multicolumn{1}{c}{Poly Order ($j$)} & \multicolumn{1}{c}{SD ($\sigma$)} & \multicolumn{1}{c}{Shape ($a$)} & \multicolumn{1}{c}{Ratio ($b$)} & \multicolumn{1}{c}{Size ($n$)} & \multicolumn{1}{c}{Distribution of fitted values} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
2 & 0.25 & -1 & 0.25 & 50 & Uniform\\
3 & 1.00 & 0 & 1.00 & 100 & Normal\\
6 & 2.00 & 1 & 4.00 & 300 & Skewed\\
18 & 4.00 &  & 16.00 &  & Discrete uniform\\
 &  &  & 64.00 &  & \\
\bottomrule
\end{tabular}}
\end{table}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-shape-of-herimite-1}

}

\caption{Polynomial forms generated for the residual plots used to assess detecting non-linearity. The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$.}\label{fig:different-shape-of-herimite}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-sigma-1}

}

\caption{Examining the effect of $\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform fitted value distribution and the "U" shape. As $\sigma$ increases the signal strength decreases, to the point that the "U" is almost unrecognisable when $\sigma=4$.}\label{fig:different-sigma}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/example-poly-lineup-1}

}

\caption{One of the lineups containing non-linearity patterns used in data collection period I. Can you spot the most different plot? The data plot is positioned at $2^3 + 1$.}\label{fig:example-poly-lineup}
\end{figure}

\hypertarget{heteroskedasticity}{%
\subsubsection{Heteroskedasticity}\label{heteroskedasticity}}

Data collection period II is designed to study the ability of human
subjects to detect the appearance of a heteroskedasticity pattern under
a simple linear regression model setting:

\begin{align} \label{eq:heter-model}
\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}), 
\end{align}

\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
\(\boldsymbol{\varepsilon}\) are vectors of size \(n\) and \(g(.)\) is
the scaling function defined in Equation \ref{eq:scaling-function}.

The null regression model used to fit the realizations generated by the
above model is formulated exactly the same as Equation
\ref{eq:null-model}.

For \(b \neq 0\), the variance-covariance matrix of the error term
\(\boldsymbol{\varepsilon}\) is correlated with the predictor
\(\boldsymbol{x}\), which will lead to the presence of
heteroskedasticity. Visual patterns of heteroskedasticity are simulated
using three different shapes (\(a\) = -1, 0, 1). (A summary of the
factors can be found in Table \ref{tab:model-factor-table}.)

Since \(supp(X) = [-1, 1]\), choosing \(a\) to be \(-1\), \(0\) and
\(1\) can generate ``left-triangle'', ``butterfly'' and
``right-triangle'' shape as displayed in Figure
\ref{fig:different-shape-of-heter}. The term \((2 - |a|)\) maintains the
magnitude of residuals across different values of \(a\).

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-shape-of-heter-1}

}

\caption{Heteroskedasticity forms used in the experiment. Three different shapes ($a = -1, 0, 1$) are used in the experiment to create left-triangle, "butterfly" and "right-triangle" shapes, respectively.}\label{fig:different-shape-of-heter}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-b-1}

}

\caption{Five different values of $b$ are used in heteroskedasticity simulation to control the strength of the signal. Larger values of $b$ yield a bigger difference in variation, and stus stronger heteroskedasticity signal.}\label{fig:different-b}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/example-heter-lineup-1}

}

\caption{One of the lineups containing heteroskedasticity pattern used in data collection period II. Can you spot the most different plot? The data plot is positioned at $3^3 - 3^2$}\label{fig:example-heter-lineup}
\end{figure}

An example lineup of this model used in data collection period II is
shown in Figure \ref{fig:example-heter-lineup} with \(a = -1\). The data
plot location is \(2^4 + 2\). Nine out of 11 subjects correctly identify
the data plot from this lineup.

\hypertarget{factors-common-to-both-data-collection-periods}{%
\subsubsection{Factors common to both data collection
periods}\label{factors-common-to-both-data-collection-periods}}

Fitted values are a function of the independent variables, and the
distribution of the observed values affects the distribution of the
fitted values. In the best case scenario the fitted values will have a
uniform distribution, which means that there is even coverage of
possible observed values across all of the predictors. This is not
always present in the collected data. Sometimes the fitted values are
discrete because one or more predictors were measured discretely. The
distribution may be relatively Gaussian, reflecting a linear combinatio
of many predictors, adhering to the Central Limit Theorem. It is also
common to see a skewed distribution of fitted values, if one or more of
the predictors has a skewed distribution. This latter problem is usually
corrected before modelling using a variable transformation. Our
simulation assess this by using four different distributions to
represent fitted values: (1) uniform, (2) normal, (3) skewed and (4)
discrete uniform. This is constructed by defining the raw predictor
\(X_{raw}\) in four corresponding distributions: (1) \(U(-1, 1)\), (2)
\(N(0, 0.3^2)\), (3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). We
would expect that the best reading of residual plots occurs when the
fitted values are uniformly distributed.

Three different sample sizes are used, \(n=50, 100, 300\) across the
experiments. We would expect considerable variation in the signal
strength in the simulated data plots with smaller \(n\). A sample size
of 300 is typically enough for structure to be visible in a scatter plot
reliably.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-dist-1}

}

\caption{Variations in fitted values, that might affect perception of residual plots. Four different distributions are used.}\label{fig:different-dist}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-n-1}

}

\caption{Examining the effect of signal strength for the three different values of $n$ used in the experiment, for non-linear structure with fixed $\sigma = 1.5$, uniform fitted value distribution, and "S" shape. For these factor levels, only when $n = 300$ is the "S" shape clearly visible.}\label{fig:different-n}
\end{figure}

\hypertarget{experimental-setup}{%
\subsection{Experimental setup}\label{experimental-setup}}

\hypertarget{controlling-the-strength-of-the-signal}{%
\subsubsection{Controlling the strength of the
signal}\label{controlling-the-strength-of-the-signal}}

As summarised in Table \ref{tab:model-factor-table}, three additional
parameters \(n\), \(\sigma\) and \(b\) are used to control the strength
of the signal so that different difficulty levels of lineups are
generated, and therefore, the estimated power curve will be smooth and
continuous. Parameter \(\sigma \in \{0.5, 1, 2, 4\}\) and
\(b \in \{0.25, 1, 4, 16, 64\}\) are used in data collection periods I
and II respectively. Figure \ref{fig:different-sigma} and
\ref{fig:different-b} demonstrate the impact of these two parameters. A
large value of \(\sigma\) will increase the variation of the error of
the non-linearity model and decrease the visibility of the visual
pattern. The parameter \(b\) controls the standard deviation of the
error across the support of the predictor. Given \(x \neq a\), a larger
value of \(b\) will lead to a larger ratio of the variance at \(x\) to
the variance at \(x - a = 0\), making the visual pattern more obvious.

Three different sample sizes are used (n = 50, 100, 300) in all three
data collection periods. It can be observed from Figure
\ref{fig:different-n} that with fewer data points drawn in a residual
plot, the visual pattern is more difficult to be detected.

\hypertarget{effect-size}{%
\subsubsection{Effect size}\label{effect-size}}

Effect size in statistics measures the strength of the signal relative
to the noise. It is surprisingly difficult to quantify in general, even
for simulated data as used in this experiment.

For the non-linearity model, the key items defining effect size are
sample size (\(n\)) and variance of the error term (\(\sigma^2\)), and
so effect size would be roughly calculated as \(\sqrt{n}/{\sigma}\). As
sample size increases the effect size would increase, but as variance
increases the effect size decreases. However, it is not clear how the
additional parameter for the model polynomial order, \(k\), should be
incorporated. Intuitively, the large \(k\) means more complex pattern,
which likely means effect size would decrease. For the purposes of our
calculations we have chosen to use an approach based on Kullback-Leibler
divergence \citep{kullback1951information}, coupled with simulation.
This formulation defines effect size to be:

\[E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right)\]
\noindent where \(diag(.)\) is the diagonal matrix constructed from the
diagonal elements of a matrix,
\(\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a\) is the
residual operator,
\(\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'\)
is the hat matrix,
\(\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z\)
is the expected values of residuals with \(\boldsymbol{Z}\) be any
higher order terms of \(\boldsymbol{X}\) leave out by the regression
equation and \(\boldsymbol{\beta}_z\) be the corresponding coefficients,
and \(\sigma^2\boldsymbol{I}\) is the assumed covariance matrix of the
error term when \(H_0\) is true.

In the heteroskedasticity model, the key elements for measuring effect
size are sample size, \(n\), and the ratio of the biggest variance to
smallest variance, \(b\). Larger values of both would produce higher
effect size. However, it is not clear how to incorporate the additional
shape parameter, \(a\). Thus the same approach is used here, where the
formula can be written as:

\[E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)\]
\noindent where \(\boldsymbol{V}\) is the actual covariance matrix of
the error term.

Derivations for these equations are provided in the Appendix.

To compute the effect size for each lineup we simulate a sufficient
large number of samples from the same model, in each sample, the number
of observations \(n\) is fixed. We then compute the effect size for each
sample and take the average as the final value. This ensures lineups
constructed with the same experimental factors will share the same
effect size.

\hypertarget{subject-allocation}{%
\subsubsection{Subject allocation}\label{subject-allocation}}

As shown in Table \ref{tab:model-factor-table}, there are a total of
\(4 \times 4 \times 3 \times 4 = 192\) and
\(3 \times 5 \times 3 \times 4 = 180\) number of combinations of
parameter values for non-linearity model and heteroskedasticity model
respectively. Three replications are made for each of the combination
results in \(192 \times 3 = 576\) and \(180 \times 3 = 540\) lineups. In
addition, each lineup is designed to be evaluated by five different
subjects. After attempting some pilot studies internally, we decide to
present a block of 20 lineups to every subject. And to ensure the
quality of the survey data, two lineups with obvious visual patterns are
included as attention checks. Thus, \(576 \times 5 / (20-2) = 160\) and
\(540 \times 5 / (20-2) = 150\) subjects are recruited to satisfy the
design of the data collection period I and II respectively.

As mentioned in Section \ref{power-of-the-tests}, \(\alpha\) used in
Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using null
lineups. Three replications are made for \(3 \times 4 = 12\)
combinations of common factors \(n\) and fitted value distribution,
results in \(12 \times 3 = 36\) lineups included in data collection
period III. In these lineups, the data of the data plot is generated
from a model with zero effect size, while the data of the 19 null plots
are generated using the same simulation method discussed in Section
\ref{visual-test-procedure-based-on-lineups}. This generation procedure
differs from the canonical Rorschach lineup procedure, which requires
that all 20 plots are generated directly from the null model. However,
these lineups serve the same fundamental purpose: to assess the number
of visually interesting plots generated under \(H_0\).

To account for the fact that our simulation method for these lineups is
not the Rorschach procedure, we use the method suggested in
\citet{vanderplas2021statistical} for typical lineups containing a data
plot to estimate \(\alpha\). We have included a sensitivity analysis in
the Appendix to examine the impact of the variance of the \(\alpha\)
estimate on our findings.

All lineups consist of only null plots are planned to be evaluated by 20
subjects. However, presenting only these lineups to subjects are
considered to be bad practices as subjects will lose interest quickly.
Therefore, we plan to collect 6 more evaluations on the 279 lineups with
uniform fitted value distribution, result in
\((36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133\) subjects
recruited for data collection period III.

\hypertarget{collecting-results}{%
\subsubsection{Collecting results}\label{collecting-results}}

Subjects for all three data collection periods are recruited from an
crowdsourcing platform called Prolific \citep{palan2018prolific}.
Prescreening procedure is applied during the recruitment, subjects are
required to be fluent in English, with \(98\%\) minimum approval rate
and 10 minimum submissions in other studies.

During the experiment, every subject is presented with a block of 20
lineups. A lineup consists of a randomly placed data plot and 19 null
plots, which are all residual plots drawn with raw residuals on the
y-axis and fitted values on the x-axis. An additional horizontal red
line is added at \(y = 0\) as a helping line.

The data of the data plot is simulated from one of two models described
in Section \ref{simulating-departures-from-good-residuals}, while the
data of the remaining 19 null plots are generated by the residual
rotation technique discussed in Section
\ref{visual-test-procedure-based-on-lineups}.

In every lineup evaluation, the subject is asked to select one or more
plots that are most different from others, provide a reason for their
selections, and evaluate how different they think the selected plots are
from others. If there is no noticeable difference between plots in a
lineup, subjects are permitted to select zero plots without providing
the reason. No subject are shown the same lineup twice. Information
about preferred pronoun, age group, education, and previous experience
in visual experiments are also collected. A subject's submission is only
accepted if the data plot is identified for at least one attention
check. Data of rejected submissions are discarded automatically to
maintain the overall data quality.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{overview}{%
\subsection{Overview}\label{overview}}

There are 2880, 2700 and 1674 lineups evaluation made by 160, 150 and
133 subjects recruited for data collection periods I, II and III
respectively. In the total of 7974 lineup evaluations, 3744 use lineups
produced by the non-linearity model, and 4230 use lineups produced by
the heteroskedasticity model. Besides, there are 886 attention checks
and 720 evaluations on null lineups needed for the estimate of
\(\alpha\) not included in the analysis. The collated dataset is
provided in \texttt{vi\_survey} of the \texttt{visage} \texttt{R}
package.

In the following analysis, lineups with uniform fitted values will be
the focus. Visual patterns are more likely to be revealed under a
uniform distribution. Additionally, we have collected extra evaluations
on these lineups, which will result in a more reliable analysis.
Analysis of lineups with other fitted value distributions can be found
in Section \ref{effect-of-fitted-value-distributions}.

\hypertarget{power-comparison-of-different-tests}{%
\subsection{Power comparison of different
tests}\label{power-comparison-of-different-tests}}

Figure \ref{fig:polypower} shows the estimated power of visual test on
lineups produced by the non-linearity model with uniform fitted values,
against the natural logarithm of the effect \(log_e(E)\), with a 5\%
significance level. At the bottom of the figure \ref{fig:polypower},
there are a sequence of example residual plots with increasing levels of
\(log_e(E)\). Readers can evaluate them from left to right and determine
at which level the departure from a good residual plot becomes
detectable.

As discussed in Section \ref{conventionally-testing-for-departures},
many conventionally tests are available for detecting residual
departures. Implementation-wise, the built-in R package \texttt{stats}
provides some commonly used residual-based tests, such as Shapiro-Wilk
test. A more comprehensive collection of regression diagnostics tests
can be found in the R package \texttt{lmtest} \citep{lmtest}. In terms
of heteroskedasticity diagnostics, the R package \texttt{skedastic}
\citep{skedastic} collects and implements 25 existing conventional tests
published since 1961.

We pick RESET test (\texttt{resettest}) and Breusch-Pagan test
(\texttt{bptest}) from the R package \texttt{lmtest}, and Shapiro-Wilk
test (\texttt{shapiro.test}) from the built-in R package \texttt{stats}.
Among them, RESET test is the only exact and appropriate test in this
scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are
approximate and inappropriate tests. Their estimated power is shown in
Figure \ref{fig:polypower}. To set up the RESET test, we include
different powers of fitted values as proxies. According to
\citet{ramsey_tests_1969}, there are no general rules for the power of
the fitted values needed by the RESET test, but it finds power up to
four is usually sufficient. Thus, we follow this guideline to conduct
the RESET test. For the Breusch-Pagan test, the choice of predictors in
the auxiliary regression is left to the user
\citep{breusch_simple_1979}. But as \citet{waldman1983note} suggested,
it is a good choice for the set of auxiliary predictors in the
Breusch-Pagan test be the same as the White test. Thus, we include both
\(x\) and \(x^2\) in the auxiliary regression.

Figure \ref{fig:heterpower} is similar to Figure \ref{fig:polypower},
but shows corresponding information on lineups produced by the
heteroskedasticity model. In this scenario, the visual test is compared
to an approximate test - Breusch-Pagan test, and two other inappropriate
tests - RESET test and Shapiro-Wilk test.

For non-linearity patterns, the power curve of RESET test climbs
aggressively from 7\% to 84\% as \(log_e(E)\) increases from 0 to 2.5,
while power of other tests respond inactively to the change of effect,
showing that RESET test is much more sensitive to weak non-linear
structure. Meanwhile, no noticeable residual departures can be spotted
from the example residual plots.

In terms of heteroskedasticity patterns, the power of Breusch-Pagan test
is constantly greater than the power of visual test. At
\(log_e(E) \approx 2.5\), where the power curve of the visual test
remains at a low level, the Breusch-Pagan test has around 50\% chance of
rejecting \(H_0\). Similarly, the visual feature is nearly unobservable
from the example residual plots at this level of effect size.

The power of visual test arises steadily as \(log_e(E)\) increases from
3 to 5 for both non-linearity patterns and heteroskedasticity patterns,
suggesting that the effect starts to make significant impact on the
degree of the presence of the designed visual features. This can also be
observed from the example residual plots that when \(log_e(E) = 3.5\), a
weak ``S-shape'' and a weak ``right-triangle'' shape are presented in
Figure \ref{fig:polypower} and Figure \ref{fig:heterpower} respectively.
The visual pattern becomes clearer as \(log_e(E)\) increases. At
\(log_e(E) = 5\), the power of visual tests for both patterns reaches
almost 100\%.

Power curves of inappropriate tests show improvement as the effect
increases but at a lower rate than the visual test in both scenarios.
This coincides the point made by \citet{cook1982residuals} that
residual-based tests for a specific type of model defect may be
sensitive to other types of model defects. The power curve of RESET test
remains at around 5\% in Figure \ref{fig:heterpower} since there are no
non-linear terms leave out in the heteroskedasticity model and \(H_0\)
of the test is always satisfied.

Overall, the power comparison suggests that conventional tests differs
significantly from visual tests in two regression diagnostics scenarios
designed by us. Visual test have much higher tolerance of the residual
departures than the conventional test. Since fail to reject \(H_0\) in a
visual test literally means that there are no obvious visual discoveries
found in the residual plot, analysts and the general public as the
consumers of the output may not be convinced of the existence of
significant residual departures in spite of the rejection of \(H_0\)
given by the conventional test. Even if the rejection is accepted, the
model violation may be considered as impactless due to the fact that
they are not clearly visible. Besides, the sensitivity of the
conventional test to weak residual departures could also distract and
discourage analysts from finding simple but good linear approximation to
the data. The rejection of \(H_0\) because of human acceptable and
negligible residual departures is not practically meaningful and useful
to analysts and decision makers. In contrast, if the strict correctness
of the model assumption is of particular interest, conducting a
conventional test is still the recommended choice based on the findings.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/polypower-1}

}

\caption{Comparison of power between different tests for non-linear patterns (uniform fitted values only). Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups. Surrounding lines of the visual test show the estimated power of 500 bootstrap samples. Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by vertical lines in main plot. Where would you draw the line of too much non-linearity in the residuals? For the RESET test this is around log effect size 2, but for the visual test it is around 3.5.}\label{fig:polypower}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/heterpower-1}

}

\caption{Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only). Main plot shows the power curves, with dots indicating human evaluations of lineups. Surrounding lines of the visual test show the estimated power of 500 bootstrap samples. Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by vertical lines in main plot. Where would you draw the line of too much heteroskedasticity in the residuals? For the BP test this is around log effect size 2.5, but for the visual test it is around 3.5.}\label{fig:heterpower}
\end{figure}

\hypertarget{comparison-of-test-decisions-based-on-p-values}{%
\subsection{\texorpdfstring{Comparison of test decisions based on
\(p\)-values}{Comparison of test decisions based on p-values}}\label{comparison-of-test-decisions-based-on-p-values}}

The power comparison illustrates that appropriate conventional tests
will reject \(H_0\) more aggressively than visual tests. In this
section, we explore how often they agree with each other by
investigating the rejections for the two model designs based on
\(p\)-values for each lineup.

Figure \ref{fig:p-value-comparison} provides a mosaic plot showing the
rejection rate of visual tests and conventional tests for both
non-linearity patterns and heteroskedasticity patterns.

For lineups containing non-linearity patterns, conventional tests reject
69\% and visual tests reject 32\% of the time. Of the lineups rejected
by the conventional test, 46\% are rejected by the visual test, that is,
approximately half as many as the conventional test. There are no
lineups that are rejected by the visual test but not by the conventional
test.

In terms of lineups containing heteroskedasticity patterns, 76\% are
rejected by conventional tests, while 56\% are rejected by visual tests.
When the conventional test rejects a lineup, there is a great chance
(73\%) the visual test will also reject it.

Surprisingly, the visual test rejects 1 of the 33 (3\%) of lineups where
the conventional test does not reject. Figure \ref{fig:heter-example}
shows this lineup. The data plot in position seventeen displays a
relatively strong heteroskedasticity pattern, and has a strong effect
size (\(log_e(E)=4.02\)). This is reflected by the visual test
\(p\text{-value} = 0.026\). But the Breusch-Pagan test
\(p\text{-value} = 0.056\), is slightly above the significance cutoff of
\(0.05\). This lineup was evaluated by 11 subjects, it has experimental
factors \(a = 0\) (``butterfly'' shape), \(b = 64\) (large variance
ratio), \(n = 50\) (small sample size), and a uniform distribution for
the fitted values. It must be the small sample size that may have
resulted in the lack of detection.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/p-value-comparison-1}

}

\caption{Rejection rate ($p$-value $\leq0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot. The visual test rejects less frequently than the conventional test. We would generally expect the visual test to only reject when the conventional test does. Surprisingly, one lineup containing a heteroskedasticity pattern does not follow this rule. }\label{fig:p-value-comparison}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/heter-example-1}

}

\caption{The single heteroskedasticity lineup that is rejected by the visual test but not by the BP test. The data plot at panel 17 contains a "butterfly" shape. It has effect size $ = 3.76$, somewhat surprising that it is not detected by the BP test.}\label{fig:heter-example}
\end{figure}

\hypertarget{effect-of-shape-of-non-linearity}{%
\subsection{Effect of shape of
non-linearity}\label{effect-of-shape-of-non-linearity}}

A primary factor contributes to the non-linearity model is the shape of
the non-linearity pattern. According to Figure
\ref{fig:poly-power-uniform-j}, conventional tests have higher power in
testing shapes constructed from lower orders of Hermite polynomials. But
the chance of detecting the ``triple-U'' shape drops significantly
compared to other shapes. To understand why this is, one needs to return
to the way the RESET test is applied. It requires a parameter indicating
degree of fitted values to test for, and the recommendation is to
generically use four \citep{ramsey_tests_1969}. However, the
``triple-U'' shape constructed from the Hermite polynomials use power up
to 18. If the RESET test had been applied using a higher power no less
than six, the power curve of ``triple-U'' shape will be only slight
lower than the power curve of ``M'' shape. The recommendation of the
polynomial power for the RESET test should be revised, perhaps. This
illustrates the sensitivity of conventional testing to the parameters,
and it also points to a limitation that one needs to know the data
generating process in order to set the parameters for the test.

For visual tests, based on the orders of the Hermite polynomials, we
expect the ``U'' shape will be the easiest one to be detected by
subjects followed by the ``S'', ``M'' and ``triple-U'' shape. From
Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power
curves are mostly aligned with our expectation, except for the ``M''
shape, which is as easy to be detected as the ``S'' shape. This implies,
unlike the conventional test, the visibility of the shape do not
strictly follow the degree of the polynomials.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/poly-power-uniform-j-1}

}

\caption{Power of conventional tests and visual tests on lineups containing four different non-linearity shapes. Power curves with higher order of non-linearity are drawn with deeper colours. The default RESET tests under-perform significantly in detecting "triple-U" shape. To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test. But this means the order needs to be known prior to testing.}\label{fig:poly-power-uniform-j}
\end{figure}

\hypertarget{effect-of-shape-of-heteroskedasticity}{%
\subsection{Effect of shape of
heteroskedasticity}\label{effect-of-shape-of-heteroskedasticity}}

We have also investigated the impact of different heteroskedasticity
shapes on power of conventional tests and visual tests. In theory, the
``left-triangle'' and the ``right-triangle'' shapes are functionally
identical from the point of view of a Breusch-Pagan test. As shown in
Figure \ref{fig:heter-power-uniform-a}, this is indeed the case where
little difference between the power curves can be perceived. Similarly,
visual tests should have the same power in detecting these two shapes if
they are equally likely to be identified. However, it can be observed
from Figure \ref{fig:heter-power-uniform-a} that the power curve of the
``left-triangle'' shape is higher than the one of the ``right-triangle''
shape, indicating a potential favour of orientation by human, which is
worth to be explored in future studies. Besides, the chance of detecting
the ``butterfly'' shape are higher than other two shapes in both
conventional and visual testing.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/heter-power-uniform-a-1}

}

\caption{Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$. The visual test has an asymmetric power in testing the "left-triangle" shape and "right-triangle" shape, while these two shapes are equivalent in conventional testing.}\label{fig:heter-power-uniform-a}
\end{figure}

\hypertarget{effect-of-fitted-value-distributions}{%
\subsection{Effect of fitted value
distributions}\label{effect-of-fitted-value-distributions}}

The prediction in a regression model is \(E(Y|X)\), that is, it is
conditional on observed values of the predictors. The distribution of
\(X\), or consequently \(\hat{Y}\), may however affect the ability to
read any patterns in the residuals. The four distributions of fitted
values, used in this experiment, were designed to examine this.

Figure \ref{fig:different-x-dist-poly-power} illustrates the impact of
fitted value distribution on the power of conventional tests and visual
tests. For conventional tests, only the power curves of appropriate
tests which are RESET tests and Breusch-Pagan tests for non-linearity
pattern and heteroskedasticity pattern respectively are shown. For
visual tests, note that we have collected more evaluations on lineup
with uniform fitted value distribution, and the number of evaluations on
a lineup will affect the power of the visual test. To have a fair
comparison, for lineups with uniform fitted value distribution, we
randomly sample five evaluations from the total eleven evaluations to
estimate the power curves. The sampling process is repeated 500 times to
produce an indication of the variation of the power.

In terms of visual tests, we do not observe significant difference
between the power curve of the uniform distribution and the power curve
of the discrete uniform distribution. Normal and lognormal distributions
produce lower power than uniform and discrete uniform distributions in
regards of residual departure patterns, but the gap is larger when the
residual plot contains heteroskedasticity pattern. This suggests that
residual departures displayed with evenly distributed fitted values are
visually more attractive and effective. Besides, the skewness of the
fitted value distribution plays a role in the degree of presence of
visual features as lognormal distribution produce the lowest power for
both non-linearity and heteroskedasticity patterns. A symmetric fitted
value distribution can better reveal the underlying shape.

Although it is unusual to consider fitted value distribution or
predictor distribution in power analysis, it is not so surprised that
the power of conventional tests vary because of fitted value
distribution. For instance, the power of the RESET test which is
effectively a F test depends on the realizations of the additional
predictors included in the auxiliary regression equation
\citep{jamshidian2007study, olvera2019relationship, zhang2018practical}.
And those additional predictors are linear transformation of random
vectors following certain predictor distributions. What is truly
unexpected is the huge variability in power of conventional tests
compared to visual tests. The discrete uniform distribution constantly
produce the highest power while the normal distribution and the
lognormal distribution produce the lowest power for non-linearity and
heteroskedasticity pattern respectively. The difference in power
corresponding to different distributions peak at around 90\% when the
effect size is at a moderate level. This huge difference is undesirable
as is not uncommon for analysts to omit the factor of predictor
distribution in power calculation and sample size calculation.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-x-dist-poly-power-1}

}

\caption{Comparison of power on lineups with different fitted value distributions for conventional tests and visual tests. Power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and Breusch-Pagan tests respectively. Power curves of visual tests are estimated using five evaluations on each lineup. For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations. The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines. The fitted value distribution has greater impact on the power of conventional tests than visual tests. Meanwhile, uneven distributions including normal and lognormal distributions genearlly produce lower power.}\label{fig:different-x-dist-poly-power}
\end{figure}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

Motivated by the advice of regression analysis experts, that residual
plots as opposed to conventional tests are an indispensible methods for
assessing model fit, a human subjects experiment was conducted using
visual inference. The experiment tested two primary departures from good
residuals: non-linearity and heteroskedasticity.

The experiment found that conventional residual-based statistical tests
are more sensitive to weak residual departures from model assumptions
than visual tests as would be evaluated by humans. That is, conventional
tests conclude there are problems with the model fit almost twice as
often as humans would. They often reject when departures in the form of
non-linearity and heteroskedasticity are not visible to a human.

One might say that this is correct behaviour, but it can be argued that
the conventional tests are rejecting when it is not necessary. Many of
these rejections happen even when downstream analysis and results would
not be significantly affected by the small departures from a good fit.
The results from human evaluations provide a more practical solution,
which reinforces the statements from regression experts that residual
plots are an indispensible method for model diagnostics.

Now it is important to note that residual plots need to be delivered as
a lineup, where it is embedded in a field of null plots. A residual plot
may contains many visual features, but some are caused by the
characteristics of the predictors and the randomness of the error, not
by the violation of the model assumptions. These irrelevant visual
features have a chance to be filtered out by subjects with a comparison
to null plots, results in a set of more accurate visual findings. This
enables a careful calibration for reading structure in residual plots.

However, human evaluation of residuals is expensive. It is
time-consuming, laborious and unfriendly to vision-impaired people. This
is another reason why it often appears to be ignored. With the
availability of sophisticated computer vision algorithms today, the goal
of this work is to form the basis of providing automated residual plot
reading. The findings suggest the strong demand of graphical inspection
in regression diagnostics, so developing an automatic visual inference
system to evaluate lineups of residual plots is valuable. We plan to
build a completed open-source system in an \texttt{R} package and
provide a web interface such as a website for public to interact with.
Details about this system will be discussed in our next paper.

The experiment also revealed some interesting details about how residual
plots are read. For the most part, the visual test performed very
similarly to the appropriate conventional test only with the power curve
shifted in the less sensitive direction. Unlike the conventional tests,
where one needs to specifically test for non-linearity or
heteroskedasticity the visual test operated effectively across the range
of departures from good residuals.

As expected, if the fitted value distribution is not uniform, there is a
loss of power in the visual test. Structure is hardest to detect if
fitted values are lognormal. Also, complex structure are generally
harder to detect, but there are outliers. Under the designed scenarios
in this paper, we find the visual test to be a more robust test against
the change of fitted value distributions. A surprising finding was that
the direction of heteroskedasticity appears to affect the ability to
visually detect it, with wedge to the right being less detectable.

\hypertarget{acknowledgements}{%
\section*{Acknowledgements}\label{acknowledgements}}
\addcontentsline{toc}{section}{Acknowledgements}

These \texttt{R} packages are used for the work: \texttt{cli}
\citep{cli}, \texttt{curl} \citep{curl}, \texttt{dplyr} \citep{dplyr},
\texttt{ggplot2} \citep{ggplot2}, \texttt{jsonlite} \citep{jsonlite},
\texttt{lmtest} \citep{lmtest}, \texttt{mpoly} \citep{mpoly},
\texttt{progress} \citep{progress}, \texttt{tibble} \citep{tibble},
\texttt{ggmosaic} \citep{ggmosaic}, \texttt{purrr} \citep{purrr},
\texttt{tidyr} \citep{tidyr}, \texttt{readr} \citep{readr},
\texttt{stringr} \citep{stringr}, \texttt{here} \citep{here},
\texttt{kableExtra} \citep{kableextra}, \texttt{patchwork}
\citep{patchwork}, \texttt{rcartocolor} \citep{rcartocolor}. The study
website is powered by \texttt{PythonAnywhere} \citep{pythonanywhere} and
the \texttt{Flask} web framework \citep{flask}. The \texttt{jsPsych}
framework \citep{jspsych} is used to create behavioral experiments that
run in our study website.

The article was created with R packages \texttt{rticles}
\citep{rticles}, \texttt{knitr} \citep{knitr} and \texttt{rmarkdown}
\citep{rmarkdown}. The project's Github repository (link here) contains
all materials required to reproduce this article.

\hypertarget{supplementary-material}{%
\section*{Supplementary material}\label{supplementary-material}}
\addcontentsline{toc}{section}{Supplementary material}

The supplementary material is available at (link here). It includes more
details about the experimental setup, the derivation of the effect size,
the effect of data collection period, and the estimate of \(\alpha\).

\bibliographystyle{tfcad}
\bibliography{paper.bib}

\end{document}

### 2023-04-03 paper.Rmd
copy in the appendix

---
title: |
  A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol
type: ARTICLE TEMPLATE
author:
  - name: Weihao Li
    affil: a
    email: weihao.li@monash.edu
  - name: Dianne Cook
    affil: a
    email: dicook@monash.edu
  - name: Emi Tanaka
    affil: a
    email: emi.tanaka@monash.edu
  - name: Susan VanderPlas
    affil: b
    email: susan.vanderplas@unl.edu
affiliation:
  - num: a
    address: |
      Department of Econometrics and Business Statistics, Monash University, Clayton, VIC, Australia
  - num: b
    address: |
      Department of Statistics, University of Nebraska, Lincoln, Nebraska, USA
bibliography: paper.bib
abstract: |
  Abstract to fill.
keywords: |
  statistical graphics; data visualization; visual inference; hypothesis testing; reression analysis; cognitive perception; simulation; practical significance; effect size
header-includes: |
  \usepackage{lscape}
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
output: rticles::tf_article
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  echo = FALSE,
  fig.width = 8,
  fig.height = 6,
  out.width = "100%",
  fig.align = "center")
```

```{r}
# OOP supports needed by `visage`
# remotes::install_github("TengMCing/bandicoot")
# 
# Visual inference models and p-value calculation
# remotes::install_github("TengMCing/visage")

library(tidyverse)
library(visage)

# To control the simulation in this file
set.seed(10086)
```

```{r get-lineup-data}
# Create the dir folder
if (!dir.exists(here::here("data/"))) dir.create(here::here("data/"))

# The lineup data used to draw residual plot needed to be downloaded
# from the github repo. Cache it.
if (!file.exists(here::here("data/vi_lineup.rds"))) {
  vi_lineup <- get_vi_lineup()
  saveRDS(vi_lineup, here::here("data/vi_lineup.rds"))
} else {
  vi_lineup <- readRDS(here::here("data/vi_lineup.rds"))
}
```

```{r poly-conventional-simulation}
# Ensure the support of the predictor is [-1, 1]
stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

# Run simulations to get the behaviours of conventional tests
if (!file.exists(here::here("data/poly_conventional_simulation.rds"))) {
  poly_conv_sim <- list()

  # 100 bootstrap samples
  for (i in 1:100)
  {
    poly_conv_sim[[i]] <- 
      # Every sample contains 2000 lineups
      map(1:2000, function(i) {

        # Sample a set of parameters
        shape <- sample(1:4, 1)
        e_sigma <- sample(c(0.5, 1, 2, 4), 1)
        x_dist <- sample(c("uniform", 
                           "normal", 
                           "lognormal", 
                           "even_discrete"), 1)
        x <- switch(x_dist,
                    uniform = rand_uniform(-1, 1),
                    normal = {
                      raw_x <- rand_normal(sigma = 0.3)
                      closed_form(~stand_dist(raw_x))
                      },
                    lognormal = {
                      raw_x <- rand_lognormal(sigma = 0.6)
                      closed_form(~stand_dist(raw_x/3 - 1))
                      },
                    even_discrete = rand_uniform_d(k = 5, even = TRUE))

        # Build the model
        mod <- poly_model(shape, x = x, sigma = e_sigma)

        # Sample the number of observations
        n <- sample(c(50, 100, 300), 1)

        # Generate data from the model 
        tmp_dat <- mod$gen(n)

        # Return a data frame containing p-values of
        # F, RESET, BP and SW tests
        tibble(shape = shape,
               e_sigma = e_sigma,
               x_dist = x_dist,
               n = n,
               F_p_value = mod$test(tmp_dat)$p_value,
               RESET3_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:3, 
                                         power_type = "fitted")$p_value,
               RESET4_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:4, 
                                         power_type = "fitted")$p_value,
               RESET5_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:5, 
                                         power_type = "fitted")$p_value,
               RESET6_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:6, 
                                         power_type = "fitted")$p_value,
               RESET7_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:7, 
                                         power_type = "fitted")$p_value,
               RESET8_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:8, 
                                         power_type = "fitted")$p_value,
               RESET9_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:9, 
                                         power_type = "fitted")$p_value,
               RESET10_p_value = mod$test(tmp_dat, 
                                          test = "RESET", 
                                          power = 2:10, 
                                          power_type = "fitted")$p_value,
               BP_p_value = HETER_MODEL$test(tmp_dat)$p_value,
               SW_p_value = shapiro.test(tmp_dat$.resid)$p.value,
               boot_id = i)
        }) %>%
  reduce(bind_rows)
  }

  poly_conv_sim <- poly_conv_sim %>%
    reduce(bind_rows)

  saveRDS(poly_conv_sim, here::here("data/poly_conventional_simulation.rds"))
} else {
  poly_conv_sim <- readRDS(here::here("data/poly_conventional_simulation.rds"))
}
```

```{r heter-conventional-simulation}
# Run simulations to get the behaviours of conventional tests
if (!file.exists(here::here("data/heter_conventional_simulation.rds"))) {
  heter_conv_sim <- list()

  # 100 bootstrap samples
  for (i in 1:100)
  {
    heter_conv_sim[[i]] <-
      # Every sample contains 2000 lineups
      map(1:2000, function(x) {

        # Sample a set of parameters
        a <- sample(c(-1, 0, 1), 1)
        b <- sample(c(0.25, 1, 4, 16, 64), 1)
        x_dist <- sample(c("uniform", 
                           "normal", 
                           "lognormal", 
                           "even_discrete"), 1)
        x <- switch(x_dist,
                    uniform = rand_uniform(-1, 1),
                    normal = {
                      raw_x <- rand_normal(sigma = 0.3)
                      closed_form(~stand_dist(raw_x))
                      },
                    lognormal = {
                      raw_x <- rand_lognormal(sigma = 0.6)
                      closed_form(~stand_dist(raw_x/3 - 1))
                      },
                    even_discrete = rand_uniform_d(-1, 1, k = 5, even = TRUE))

        # Build the model
        mod <- heter_model(a = a, b = b, x = x)

        # Sample the number of observations
        n <- sample(c(50, 100, 300), 1)

        # Generate data from the model 
        tmp_dat <- mod$gen(n)

        # Return a data frame containing p-values of
        # F, RESET, BP and SW tests
        tibble(a = a,
               b = b,
               x_dist = x_dist,
               n = n,
               F_p_value = POLY_MODEL$test(

                 # Create a pseudo z to be able to use F-test
                 tmp_dat %>%
                   mutate(z = poly_model()$
                            gen(n, computed = select(tmp_dat, x)) %>%
                            pull(z))
                 )$p_value,
               RESET3_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:3, 
                                                power_type = "fitted")$p_value,
               RESET4_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:4, 
                                                power_type = "fitted")$p_value,
               RESET5_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:5, 
                                                power_type = "fitted")$p_value,
               RESET6_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:6, 
                                                power_type = "fitted")$p_value,
               RESET7_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:7, 
                                                power_type = "fitted")$p_value,
               RESET8_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:8, 
                                                power_type = "fitted")$p_value,
               RESET9_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:9, 
                                                power_type = "fitted")$p_value,
               RESET10_p_value = POLY_MODEL$test(tmp_dat, 
                                                 test = "RESET", 
                                                 power = 2:10, 
                                                 power_type = "fitted")$p_value,
               BP_p_value = mod$test(tmp_dat)$p_value,
               SW_p_value = shapiro.test(tmp_dat$.resid)$p.value,
               boot_id = i)
        }) %>%
  reduce(bind_rows)
  }

  heter_conv_sim <- heter_conv_sim %>%
    reduce(bind_rows)

  saveRDS(heter_conv_sim, here::here("data/heter_conventional_simulation.rds"))
} else {
  heter_conv_sim <- readRDS(here::here("data/heter_conventional_simulation.rds"))
}
```

```{r}
# Borrow effect size from the survey
poly_conv_sim <- poly_conv_sim %>%
  left_join(select(filter(vi_survey, type == "polynomial"), shape, e_sigma, n, x_dist, effect_size))

heter_conv_sim <- heter_conv_sim %>%
  left_join(select(filter(vi_survey, type == "heteroskedasticity"), a, b, n, x_dist, effect_size))
```

# Introduction

> *"Since all models are wrong the scientist must be alert to what is importantly wrong."* [@box1976science]

Diagnostics are the key to determining whether there is anything importantly wrong with a model. In linear regression analysis, residuals from the model fit are commonly used. Residuals summarise what is not captured by the model, and thus provide the capacity to identify what might be wrong.

We can assess residuals in multiple ways. Residuals may be plotted as a histogram or quantile-quantile plot to examine the distribution. Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, we consider it to be well behaved. However, if the distribution is skewed, bimodal, multimodal, or contains outliers, there is a cause for concern. One could also inspect the distribution by conducting a goodness of fit test, such as the Shapiro-Wilk Normality test [@shapiro1965analysis].

More typically, residuals will be plotted, as a scatter plot against the predicted values and each of the explanatory variables to scrutinize their relationships. If there are any visually discoverable patterns, the model is potentially misspecified. In general, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity. However, correctly judging a residual plot where no pattern exists can be a painstakingly difficult task for humans. It is especially common, particularly among students, to misinterpret patterns that are random noise and random deviation from a model [@loy2021bringing]. It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.

Abundance of literature describe appropriate diagnostic methods for linear regression, e.g. @draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals. All these writings consider plotting residuals to be a standard technique that should be examined routinely in all regression modelling problems. In addition, @draper1998applied and @belsley_regression_1980 believe that residual plots are usually revealing when the assumptions are violated. @cook_applied_1999 thinks formal tests and graphical procedures are complementary and both have a place in residual analysis, but they focus on graphical methods rather than on formal testing, as they are easier to use. @montgomery1982introduction even suggests that residual plots are more informative in most practical situations than the corresponding formal tests, and statistical tests on regression model residuals are not widely used based on their experience.

A common guidance by experts is that optimal method for diagnosing model fits is by plotting the data. The persistence of this advice to check the plots is curious, and investigating why this might be common advice is the subject of this paper. The paper is structured as follows. The next section describes the background on the types of departures that one expects to detect, and outlines a formal statistical process for reading residual plots, called visual inference. Section \ref{experimental-design} details the experimental design to compare the decisions made by formal hypothesis testing, and how humans would read diagnostic plots. The results are reported in Section \ref{results}. We conclude with a discussion of future work, in particular, how the responsibility for residual plot reading might be relieved.

# Background

## Departures from good residual plots

```{r residual-plot-common-departures, fig.width = 10, fig.height = 2.5, fig.cap = "Example residual vs fitted value plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0. The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted value plot."}

set.seed(10086)

# Base model
mod <- poly_model(include_z = FALSE, sigma = 0.25)

# Base data
dat <- mod$gen(300)

bind_rows(
  dat %>%
    mutate(type = "(A) Good residuals"),
  mod$set_prm("include_z", TRUE)$
    gen(300, computed = select(dat, x, e)) %>%
    mutate(type = "(B) Non-linearity"),
  heter_model(b = 64)$
    gen(300, computed = select(dat, x, e)) %>%
    mutate(type = "(C) Heteroskedasticity"),
  heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
    gen(300, computed = select(dat, x)) %>%
    mutate(type = "(D) Non-normality")
) %>%
  mod$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~type, ncol = 4, scales = "free") +
  xlab("Fitted values") +
  ylab("Residuals")
```

Graphical summaries in which residuals are plotted against fitted values or other functions of the predictors that are approximately orthogonal to the residuals are referred to as standard residual plots in @cook1982residuals. Figure \ref{fig:residual-plot-common-departures}A shows an ideal residual plot where the residuals are evenly distributed at both sides of the horizontal zero line, with no noticeable patterns.

There are various types of departures from an ideal residual plot. Non-linearity, heteroskedasticity and non-normality are perhaps the three mostly checked departures.

Non-linearity is a type of model misspecification caused by failing to include higher order terms of the predictors in the regression equation. Any non-linear functional form of residuals on fitted values in the residual plot could be indicative of non-linearity. An example residual plot containing visual pattern of non-linearity is shown in Figure \ref{fig:residual-plot-common-departures}B. One can clearly observe the "S-shape" from the residual plot as the cubic term is not captured by the misspecified model.

Heteroskedasticity refers to the presence of nonconstant error variance in a regression model. It is mostly due to the strict but false assumptions on the variance-covariance matrix of the error term. The usual pattern of heteroskedasticity on a residual plot is the inconsistent spread of the residuals across the horizontal axis. Visually, it sometimes results in the so-called "butterfly" shape as shown in Figure \ref{fig:residual-plot-common-departures}C, or the "left-triangle" and "right-triangle" shape where the smallest variance occurs at one side of the horizontal axis.

Compared to non-linearity and heteroskedasticity, non-normality is usually harder to detect from a residual plot since a scatter plot do not readily reveal the marginal distribution. A favourable graphical summary for this task is the quantile-quantile plot. As we mainly discuss residual plots, non-normality will not be the focus of this paper. For a consistent comparison, the residual plot of this departure is still presented in Figure \ref{fig:residual-plot-common-departures}D. When the number residuals below and above the horizontal axis are uneven across the local regions along the $x$-axis, we expect that the normality assumption is violated. For example, given a skewed error distribution, there will be fewer data points and more outliers on one side of the horizontal axis as shown in Figure \ref{fig:residual-plot-common-departures}D.

## Conventionally testing for departures

Other than checking diagnostic plots, analysts may perform formal hypothesis testing to detect model defects. A variety of tests can be applied; each testing for a specific violation of the null hypothesis. For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution. To test specific forms of non-linearity, one may apply the F-test as a model structural test to examine the significance of specific polynomial and non-linear forms of the predictors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969]. The Shapiro-Wilk test [@shapiro1965analysis] is the most widely used test of non-normality included by many of the statistical software programs. The Jarque--Bera test [@jarque1980efficient] is also used to directly check whether the sample skewness and kurtosis match a normal distribution.

We apply the RESET test, Breusch-Pagan test and Shapiro-Wilk test to the residual plots shown in Figure \ref{fig:residual-plot-common-departures}; the results of this is shown in Table \ref{tab:example-residual-plot-table}. The Breusch-Pagan test and the Shapiro-Wilk test both reject $H_0$ for residual plots that display non-linearity and heteroscedasticity, even though these are not the original intention of the test. As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are also sensitive to other types of departures. It is likely $H_0$ is correctly rejected but for the wrong reason, a phenomenon known as the "Type III error". Additionally, outliers will often incorrectly trigger the rejection of $H_0$ despite when majority of the residuals are well-behaved [@cook_applied_1999]. Furthermore, with a sufficiently large sample size, residual-based tests may reject $H_0$ due to a slight departure that is of little practical significance. These can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers and slight departures.

```{r}
set.seed(10086)

# Base model
mod <- poly_model(include_z = FALSE, sigma = 0.25)

# Base data
dat0 <- mod$gen(300)

# Replicate data in Figure 1
dat1 <- mod$set_prm("include_z", TRUE)$gen(300, computed = select(dat0, x, e))
dat2 <- heter_model(b = 64)$gen(300, computed = select(dat0, x, e))
dat3 <- heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
               gen(300, computed = select(dat0, x))

dat_list <- list(dat0, dat1, dat2, dat3)

table_dat <- data.frame(plot = c("A", "B", "C", "D"), 
                        model = c("None", 
                                  "Non-linearity", 
                                  "Heteroskedasticity", 
                                  "Non-normality"),
                        r = map_dbl(dat_list, 
                                    ~POLY_MODEL$test(.x, 
                                                     test = "RESET", 
                                                     power = 2:4)$p_value),
                        b = map_dbl(dat_list, 
                                    ~HETER_MODEL$test(.x)$p_value),

                        # SW test has not been included by visage
                        s = map_dbl(dat_list, 
                                    ~shapiro.test(.x$.resid)$p.value)) %>%

  # 3 decimal points
  mutate(across(r:s, ~ format(round(.x, digits = 3), nsmall = 3))) %>%

  # Italic if reject
  mutate(across(r:s, ~ kableExtra::cell_spec(.x, italic = as.numeric(.x) <= 0.05)))

  table_dat %>%
  knitr::kable(col.names = c("Plot", 
                             "Departures", 
                             "RESET", 
                             "Breusch-Pagan", 
                             "Shapiro–Wilk"),
               align = "llrrr",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Statistical significance testing for departures from good residuals for plots in Figure \\ref{fig:residual-plot-common-departures}. Shown are the $p$-values calculated for the RESET, the Breusch-Pagan and the Shapiro–Wilk tests. The good residual plot (A) is judged a good residual plot, as expected, by all tests. The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.', 
               label = "example-residual-plot-table" ) 
```

## Visual test procedure based on lineups

One may argue that reading diagnostic plots is to some extent subjective and indecisive compared to those rigorous statistical procedures as it relies on graphical perception - human ability to interpret and decode the information embedded in graph [@cleveland_graphical_1984]. Further, the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data. For instance, people 
over-interpret the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurrence [@roy_chowdhury_using_2015].

Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots. Later, @buja_statistical_2009 proposed the lineup protocol as a visual test inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people. The protocol consists of $m$ randomly placed plots, where one plot is the data plot, and the remaining $m - 1$ null plots have the identical graphical procedure except the data has been replaced with data consistent with $H_0$, that the model is correctly specified. Then, an observer who have not seen the data plot will be asked to point out the most different plot from the lineup. Under $H_0$, it is expected that the data plot would have no distinguishable difference from the null plots, and the probability that the observer correctly picks the data plot is $1/m$. If one rejects $H_0$ as the observer correctly picks the data plot, then the Type I error of this test is $1/m$.

```{r first-example-lineup, fig.height = 7, fig.width = 7, fig.cap = "Visual testing is conducted using a lineup, as in the example here. The residual plot computed from the observed data (plot $2^2 + 2$, exhibiting non-linearity) is embedded among 19 null plots, where the residuals are simulated from a standard error model. Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot. A small $p$-value would result from a substantial number selecting plot $2^2 + 2$."}

vi_lineup$poly_300$data %>%
  VI_MODEL$plot_lineup(theme = theme_light(), remove_axis = TRUE, remove_grid_line = TRUE)
```

Figure \ref{fig:first-example-lineup} is an example of a lineup protocol. If the data plot at position $2^2 + 2$ is identifiable, then it is evidence for the rejection of $H_0$. In fact, the actual residual plot is obtained from a misspecified regression model with missing non-linear terms.

Data used in the $m - 1$ null plots needs to be simulated. In regression diagnostics, sampling data consistent with $H_0$ is equivalent to sampling data from the assumed model. As @buja_statistical_2009 suggested, $H_0$ is usually a composite hypothesis controlled by nuisance parameters. Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to a so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling. The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009]. Essentially, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the predictors, then rescaling it by the ratio of residual sum of square in two regressions.

The effectiveness of lineup protocol for regression analysis is validated by @majumder_validation_2013 under relatively simple settings with up to two predictors. Their results suggest that visual tests are capable of testing the significance of a single predictor with a similar power as a t-test, though they express that in general it is unnecessary to use visual inference if there exists a conventional test, and they do not expect the visual test to perform equally well as the conventional test. In their third experiment, where there is not a conventional test, visual test outperforms the conventional test for a large margin. This is encouraging, as it promotes the use of visual inference in situations where there are no existing statistical testing procedures. Visual inference have also been integrated into diagnostic of hierarchical linear models by @loy2013diagnostic, @loy2014hlmdiag and @loy2015you. They use lineup protocols to judge the assumption of linearity, normality and constant error variance for both the level-1 and level-2 residuals.

# Calculation of statistical significance and test power

## What is being tested?

In diagnosing a model fit from residuals, we are generally interested in $H_0:$ *The regression model is correctly specified* against the broad alternative $H_a:$ *The regression model is misspecified*.

However, it is practically impossible to test this specific $H_0$ with conventional tests, which are constructed to measure specific departures. For example, the RESET test is formulated as $H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0$ against $H_a: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0$, from $y = \tau_0 + \sum_{i=1}^{p}\tau_px_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)$. 
Similarly, the Breusch-Pagan test is designed to specifically test $H_0:$ *error variances are all equal* ($\zeta_i=0 \text{ for } i=1,..,p$) versus the alternative $H_a:$ *that the error variances are a multiplicative function of one or more variables* ($\text{at least one } \zeta_i\neq 0$) from $e^2 = \zeta_0 + \sum_{i=1}^{p}\zeta_i x_i + u, ~ u\sim N(0,\sigma_u^2)$.

One of the potential benefits of the visual test, based on the lineup protocol, is that it should be able to detect a range of departures from good residuals.

## Statistical significance

In hypothesis testing, a $p$-value is defined as the probability of observing test results as least as extreme as the observed result given $H_0$ is true. Conventional hypothesis tests usually have an existing method to derive or compute $p$-value based on the null distribution. What we need to discuss in the following is the method to estimate $p$-value for a visual test.

Within the context of visual inference, by involving $k$ independent observers, the visual $p$-value can be interpreted as the probability of having as many or more subjects detect the data plot than the observed result.

Let $X_j = \{0,1\}$ be a Bernoulli random variable denoting whether subject $j$ correctly detecting the data plot, and $X = \sum_{j=1}^{K}X_j$ be the number of observers correctly picking the data plot. Then, by imposing a relatively strong assumption on the visual test that all $K$ evaluations are fully independent, under $H_0$, $X \sim \mathrm{Binom}_{K,1/m}$. Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as $P(X \geq x) = 1 - F(x) + f(x)$, where $F(.)$ is the binomial cumulative distribution function, $f(.)$ is the binomial probability mass function and $x$ is the realization of number of observers correctly picking the data plot [@majumder_validation_2013].

As pointed out by @vanderplas2021statistical, this basic binomial model does not take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup. And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup. @vanderplas2021statistical summarises three common scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects. Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experiment design. For Scenario 3, @vanderplas2021statistical models the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$. The number of times plot $i$ being selected in $K$ evaluations is denoted as $c_i$. In case subject $j$ makes multiple selections, $1/s_j$ will be added to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$. This ensures $\sum_{i}c_i=K$. Since we are only interested in the selections of the data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual $p$-value is given as

```{=tex}
\begin{equation} \label{eq:pvalue-beta-binomial}
P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+
\end{equation}
```

\noindent where $B(.)$ is the beta function defined as

```{=tex}
\begin{equation} \label{eq:betafunction}
B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0.
\end{equation}
```

Note that Equation \ref{eq:pvalue-beta-binomial} given in @vanderplas2021statistical only works with non-negative integer $c_i$. We extend the equation to non-negative real number $c_i$ by applying a linear approximation

```{=tex}
\begin{equation} \label{eq:pvalue-beta-binomial-approx}
P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+,
\end{equation}
```

where $P(C \geq \lceil c_i \rceil)$ is calculated using Equation \ref{eq:pvalue-beta-binomial} and $P(C = \lfloor c_i \rfloor)$ is calculated by

\begin{equation} \label{eq:pmf-beta-binomial}
P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+.
\end{equation}

Besides, the parameter $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} and \ref{eq:pmf-beta-binomial} is usually unknown and hence needs to be estimated from the survey data. For low values of $\alpha$, only a few plots are attractive to the observers and tend to be selected. For higher values of $\alpha$, the distribution of the probability of each plot being selected is more even. @vanderplas2021statistical defines that a plot is $c$-interesting if $c$ or more participants select the plot as the most different. Given the definition, The expected number of plots selected at least $c$ times, $E[Z_c]$, is calculated as

```{=tex}
\begin{equation} \label{eq:c-interesting-expectation}
E[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}
```

With Equation \ref{eq:c-interesting-expectation}, $\alpha$ can be estimated using maximum likelihood estimation. But for precise estimate of $\alpha$, additional human responses to Rorschach lineups, which is a type of lineup that consists of plots constructed from the same null data generating mechanism, are required.

## Power of the tests

The power of a model misspecification test is the probability that $H_0$ is rejected given the regression model is misspecified in a specific way. It is an important indicator
when one is concerned about whether model assumptions have been violated. Although in practice, one might be more interested in knowing how much the residuals deviate from the model assumptions, and whether this deviation is of practical significance.

The power of a conventional hypothesis test is affected by both the true parameter $\boldsymbol{\theta}$ and the sample size $n$. These two can be quantified in terms of effect size $E$ to measure the strength of the residual departures from the model assumptions. Details about the effect size is provided in Section \ref{effect-size} after the introduction of the simulation model used in our human subject experiment. The theoretical power of a test is sometimes not a trival solution, but it can be estimated if the data generating process is known. We use a predefined model to generate a large set of simulated data under different effect sizes, and record if the conventional test rejects $H_0$. The probability of the conventional test rejects $H_0$ is then fitted by a logistic regression formulated as

```{=tex}
\begin{equation} \label{eq:logistic-regression-1-1}
Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),
\end{equation}
```

\noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$. The effect size $E$ is the only predictor and the intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$, which is the desired significance level.

The power of a visual test on the other hand, may additionally depend on the ability of the particular subject, as the skill of the individual may affect the number of observers who identify the data plot from the lineup [@majumder_validation_2013]. To address this issue, @majumder_validation_2013 models the probability of a subject $j$ correctly picking the data plot from a lineup $l$ using a mixed-effect logistic regression, with subjects treated as random effect. Then, the estimated power of a visual test evaluated by a single subject is the predicted value obtained from the mix-effect model. However, this mix-effect model does not work with scenario where subjects are asked to select one or more most different plots. In this scenario, having the probability of a subject $j$ correctly picking the data plot from a lineup $l$ is insufficient to determine the power of a visual test because it does not provide information about the number of selections made by the subject for the calculation of the $p$-value (See Equation \ref{eq:pvalue-beta-binomial-approx}). Therefore, we directly estimate the probability of a lineup being rejected by assuming that individual skill has negligible effect on the variation of the power. This assumption is not necessary true, but it helps simplifying the model structure, thereby obviate a costly large-scale experiment to estimate complex covariance matrices. The same model given in Equation \ref{eq:logistic-regression-1-1} is applied to model the power of a visual test.

To study various factors contributing to the power of both tests, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors. These include the distribution of the fitted values, type of the simulation model and the shape of the residual departures.

# Experimental design

An experiment is conducted in three data collection periods to investigate the difference between conventional hypothesis testing and visual inference in the application of linear regression diagnostics. Two types of departures, non-linearity and heteroskedasticity, are collected during data collection periods I and II. The data collection period III was designed primarily to measure human responses to null lineups so that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated. Additional lineups for both non-linearity and heteroskedasticity, using uniform fitted value distribution, were included so that the participants were evaluating some lineups with signal also. It would be too frustrating for participants to only be assigned lineups with all null plots. Overall, we collected `r vi_survey  %>% filter(!attention_check) %>% nrow()` evaluations on `r vi_survey  %>% filter(!attention_check) %>% pull(unique_lineup_id) %>% unique() %>% length()` unique lineups performed by `r vi_survey  %>% filter(!attention_check) %>% count(exp, set) %>% nrow()` subjects throughout three data collection periods.

## Simulating departures from good residuals

### Non-linearity

Data collection period I is designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials on random vector $\boldsymbol{x}$ formulated as

\begin{align} \label{eq:nonlinearity-model}
\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
\end{align}

\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as

```{=tex}
\begin{equation} \label{eq:scaling-function}
g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x}))2k - k, \quad \text{for} \quad k > 0. 
\end{equation}
```

According to @abramowitz1964handbook, Hermite polynomials were initially defined by @de1820theorie, but named after Hermite [@hermite1864nouveau] because of the unrecognisable form of Laplace's work. When simulating $\boldsymbol{z}_{raw}$, function `hermite` from the R package `mpoly` [@mpoly] is used to generate Hermite polynomials.

The null regression model used to fit the realizations generated by the above model is formulated as

```{=tex}
\begin{equation} \label{eq:null-model}
\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
\end{equation}
```
\noindent where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.

Since $z = O(x^j)$, for $j > 1$, $z$ is a higher order term leaves out by the null regression, which will lead to model misspecification.

Visual patterns of non-linearity are simulated using four different orders of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$).  (A summary of the factors is given in Table \ref{tab:model-factor-table}.) The values of $j$ is chosen so that distinct shapes of non-linearity are included in the residual plot. These include "U", "S", "M" and "triple-U" shape as shown in Figure \ref{fig:different-shape-of-herimite}. A greater value of $j$ will result in a curve with more turning points. It is expected that the "U" shape will be the easiest one to detect because complex shape tends to be concealed by cluster of data points.

Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups used in non-linearity detection. This lineup is produced by the non-linearity model with $j = 6$. The data plot location is $2^3 - 4$. All five subjects correctly identify the data plot from this lineup.

```{r results='asis'}
data.frame(j = c("2", "3", "6", "18", ""),
           sigma = c("0.25", "1.00", "2.00", "4.00", ""),
           a = c("-1", "0", "1", "", ""),
           b = c("0.25", "1.00", "4.00", "16.00", "64.00"),
           n = c("50", "100", "300", "", ""), 
           x_dist = c("Uniform", "Normal", "Skewed", "Discrete uniform", "")) %>%
  knitr::kable(col.names = NULL,
               align = "rr|rr|rc",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Levels of the factors used in data collection periods I, II, III.', 
               label = "model-factor-table" ) %>%
  kableExtra::kable_styling(latex_options = "scale_down") %>%
  kableExtra::add_header_above(c("Poly Order (sym1)", #"Order of Hermite polynomial\n(sym1)", 
                                 "SD (sym3)", #"Standard deviation of polynomial model\n(sym3)",
                                 "Shape (sym4)", # "Shape of heteroskedasticity\n(sym4)", 
                                 "Ratio (sym5)", # Variance factor of heteroskedasticity model\n(sym5)",
                                 "Size (sym6)", 
                                 "Distribution of fitted values")) %>% #"Sample size\n(sym6)")) %>%
  kableExtra::add_header_above(c("Non-linearity" = 2, 
                               "Heteroskedasticity" = 2, "Common" = 2)) %>%
  as.character() %>%
  gsub("sym1", "$j$", .) %>%
  gsub("sym3", "$\\\\sigma$", .) %>%
  gsub("sym4", "$a$", .) %>%
  gsub("sym5", "$b$", .) %>%
  gsub("sym6", "$n$", .) %>%
  cat()
```

```{r different-shape-of-herimite, fig.height = 2, fig.cap = "Polynomial forms generated for the residual plots used to assess detecting non-linearity. The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$."}

# Facet label
shape_labels <- c('He[2]:"U"', 'He[3]:"S"', 'He[6]:"M"', 'He[18]:"triple-U"')

# Data for shape 1
dat_shape_1 <- poly_model(shape = 1, 
                          x = {
                            raw_x <- rand_uniform(-1, 1);
                            closed_form(~stand_dist(raw_x))
                            }, 
                          sigma = 0.05)$gen(300) %>%
  mutate(shape = shape_labels[1])

# Generate data for shape 2, 3 and 4. Reuse x and e.
map_df(2:4, function(shape) {
  poly_model(shape = shape, 
                  x = {
                    raw_x <- rand_uniform(-1, 1); 
                    closed_form(~stand_dist(raw_x))
                    }, 
                  sigma = 0.05)$
  gen(300, computed = select(dat_shape_1, x, e)) %>%
  mutate(shape = shape_labels[shape])
}) %>%

  # Combined with data for shape 1
  bind_rows(dat_shape_1) %>%
  mutate(shape = factor(shape, levels = shape_labels)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~shape, scales = "free", labeller = label_parsed, ncol = 4) +
  xlab("Fitted values") +
  ylab("Residuals")
```

```{r different-sigma, fig.height = 2, fig.cap = 'Examining the effect of $\\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform fitted value distribution and the "U" shape. As $\\sigma$ increases the signal strength decreases, to the point that the "U" is almost unrecognisable when $\\sigma=4$.'}

# Generate data for sigma 0.5
dat_sigma_05 <- poly_model(shape = 1, 
                           x = {
                             raw_x <- rand_uniform(-1, 1);
                             closed_form(~stand_dist(raw_x))
                             }, 
                           sigma = 0.5)$gen(300) %>%
  mutate(sigma = "sigma: 0.5")

# Generate data for other sigma
map(c(1, 2, 4), function(sigma) {
  poly_model(shape = 1, 
                  x = {
                    raw_x <- rand_uniform(-1, 1); 
                    closed_form(~stand_dist(raw_x))
                    }, 
                  sigma = sigma)$
  gen(300, computed = select(dat_sigma_05, x)) %>%
  mutate(sigma = paste0("sigma: ", sigma))
}) %>%

  # Combined with data for sigma 0.5
  bind_rows(dat_sigma_05) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) + 
  facet_wrap(~sigma, ncol = 4, scales = "free", labeller = label_parsed) +
  xlab("Fitted values") +
  ylab("Residuals")
```

```{r example-poly-lineup, fig.height = 7, fig.width = 7, fig.cap = "One of the lineups containing non-linearity patterns used in data collection period I. Can you spot the most different plot? The data plot is positioned at $2^3 + 1$."}
VI_MODEL$plot_lineup(vi_lineup$poly_24$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

### Heteroskedasticity

Data collection period II is designed to study the ability of human subjects to detect the appearance of a heteroskedasticity pattern under a simple linear regression model setting:

```{=tex}
\begin{align} \label{eq:heter-model}
\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}), 
\end{align}
```

\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vectors of size $n$ and $g(.)$ is the scaling function defined in Equation \ref{eq:scaling-function}.

The null regression model used to fit the realizations generated by the above model is formulated exactly the same as Equation \ref{eq:null-model}.

For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the predictor $\boldsymbol{x}$, which will lead to the presence of heteroskedasticity. Visual patterns of heteroskedasticity are simulated using three different shapes ($a$ = -1, 0, 1). (A summary of the factors can be found in Table \ref{tab:model-factor-table}.)

Since $supp(X) = [-1, 1]$, choosing $a$ to be $-1$, $0$ and $1$ can generate "left-triangle", "butterfly" and "right-triangle" shape as displayed in Figure \ref{fig:different-shape-of-heter}. The term $(2 - |a|)$ maintains the magnitude of residuals across different values of $a$.

```{r different-shape-of-heter, fig.height = 2.67, fig.cap = 'Heteroskedasticity forms used in the experiment. Three different shapes ($a = -1, 0, 1$) are used in the experiment to create left-triangle, "butterfly" and "right-triangle" shapes, respectively.'}

set.seed(10086)

a_labels <- c("left-triangle", "butterfly", "right-triangle")

# Generate data for a = -1
dat_a_n1 <- heter_model(a = -1, 
                        x = {
                          raw_x <- rand_uniform(-1, 1);
                          closed_form(~stand_dist(raw_x))
                          },
                        b = 128)$gen(300) %>%
  mutate(a = a_labels[1])

# Generate data for other a
map(c(0, 1), function(a) {
  heter_model(a = a,
              x = {
                raw_x <- rand_uniform(-1, 1); 
                closed_form(~stand_dist(raw_x))
                }, 
              b = 128)$
    gen(300, computed = select(dat_a_n1, x, e)) %>%
    mutate(a = a_labels[a + 2])
}) %>%

  # Combined with data for a = -1
  bind_rows(dat_a_n1) %>%
  mutate(a = factor(a, levels = a_labels)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~a, scales = "free", ncol = 3) +
  xlab("Fitted values") +
  ylab("Residuals")
```

```{r different-b, fig.height = 2, fig.cap = "Five different values of $b$ are used in heteroskedasticity simulation to control the strength of the signal. Larger values of $b$ yield a bigger difference in variation, and stus stronger heteroskedasticity signal."}

# Generate data for b = 0.25
dat_b025 <- heter_model(a = 0, 
                        x = {
                          raw_x <- rand_uniform(-1, 1);
                          closed_form(~stand_dist(raw_x))
                          }, 
                        b = 0.25)$gen(300) %>%
  mutate(b = "b: 0.25")

# Generate data for other b
map(c(1, 4, 16, 64), function(b) {
  heter_model(a = 0,
              x = {
                raw_x <- rand_uniform(-1, 1); 
                closed_form(~stand_dist(raw_x))
                }, 
              b = b)$
  gen(300, computed = select(dat_b025, x, e)) %>%
  mutate(b = paste0("b: ", b))
}) %>%

  # Combined with data for b = 0.25
  bind_rows(dat_b025) %>%
  mutate(b = factor(b, levels = paste0("b: ", c(0.25, 1, 4, 16, 64)))) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~b, scales = "free", ncol = 5) +
  xlab("Fitted values") +
  ylab("Residuals")
```

```{r example-heter-lineup, fig.height = 7, fig.width = 7, fig.cap = "One of the lineups containing heteroskedasticity pattern used in data collection period II. Can you spot the most different plot? The data plot is positioned at $3^3 - 3^2$"}

VI_MODEL$plot_lineup(vi_lineup$heter_471$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

An example lineup of this model used in data collection period II is shown in Figure \ref{fig:example-heter-lineup} with $a = -1$. The data plot location is $2^4 + 2$. Nine out of 11 subjects correctly identify the data plot from this lineup.

### Factors common to both data collection periods

Fitted values are a function of the independent variables, and the distribution of the observed values affects the distribution of the fitted values. In the best case scenario the fitted values will have a uniform distribution, which means that there is even coverage of possible observed values across all of the predictors. This is not always present in the collected data. Sometimes the fitted values are discrete because one or more predictors were measured discretely. The distribution may be relatively Gaussian, reflecting a linear combinatio of many predictors, adhering to the Central Limit Theorem. It is also common to see a skewed distribution of fitted values, if one or more of the predictors has a skewed distribution. This latter problem is usually corrected before modelling using a variable transformation. Our simulation assess this by using four different distributions to represent fitted values: (1) uniform, (2) normal, (3) skewed and (4) discrete uniform. This is constructed by defining the raw predictor $X_{raw}$ in four corresponding distributions: 
(1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$. We would expect that the best reading of residual plots occurs when the fitted values are uniformly distributed.

Three different sample sizes are used, $n=50, 100, 300$ across the experiments. We would expect considerable variation in the signal strength in the simulated data plots with smaller $n$. A sample size of 300 is typically enough for structure to be visible in a scatter plot reliably.

```{r different-dist, fig.height = 2, fig.cap = "Variations in fitted values, that might affect perception of residual plots. Four different distributions are used."}

# Data for uniform distribution
dat_dist_1 <- poly_model(shape = 1, 
                      x = {
                        raw_x <- rand_uniform(-1, 1);
                        closed_form(~stand_dist(raw_x))
                        }, 
                      sigma = 0.5)$gen(300) %>%
  mutate(x_dist = "Uniform")

# Generate data for other distributions
dat_dist_2 <- poly_model(shape = 1,
                         x = {
                           raw_x <- rand_normal(sigma = 0.3); 
                           closed_form(~stand_dist(raw_x))
                           }, 
                         sigma = 0.5)$
  gen(300, computed = select(dat_dist_1, e)) %>%
  mutate(x_dist = "Normal")

dat_dist_3 <- poly_model(shape = 1,
                         x = {
                           raw_x <- rand_lognormal(sigma = 0.6); 
                           closed_form(~stand_dist(raw_x/3 - 1))
                           }, 
                         sigma = 0.5)$
  gen(300, computed = select(dat_dist_1, e)) %>%
  mutate(x_dist = "Skewed")

dat_dist_4 <- poly_model(shape = 1,
                         x = {
                           raw_x <- rand_uniform_d(k = 5, even = TRUE); 
                           closed_form(~stand_dist(raw_x))
                           }, 
                         sigma = 0.5)$
  gen(300, computed = select(dat_dist_1, e)) %>%
  mutate(x_dist = "Discrete uniform")

# Generate and plot data for discrete uniform distribution
bind_rows(dat_dist_1, dat_dist_2, dat_dist_3, dat_dist_4) %>%
  mutate(x_dist = factor(x_dist, 
                         levels = c("Uniform", 
                                    "Normal", 
                                    "Skewed", 
                                    "Discrete uniform"))) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  xlab("Fitted values") +
  ylab("Residuals") +
  facet_wrap(~x_dist, ncol = 4, scales = "free")
```

```{r different-n, fig.height = 2.67, fig.cap = 'Examining the effect of signal strength for the three different values of $n$ used in the experiment, for non-linear structure with fixed $\\sigma = 1.5$, uniform fitted value distribution, and "S" shape. For these factor levels, only when $n = 300$ is the "S" shape clearly visible.'}

set.seed(666)

# Generate data for 300 observations
dat_n_300 <- poly_model(shape = 2, 
                        x = {
                          raw_x <- rand_uniform(-1, 1)
                          closed_form(~stand_dist(raw_x))
                        },
                        sigma = 1)$gen(300) %>%
  mutate(n = "N: 300")

# Generate data for other settings
map(c(50, 100), function(n) {
  poly_model(shape = 2, 
                   x = {
                     raw_x <- rand_uniform(-1, 1);
                     closed_form(~stand_dist(raw_x))
                     }, 
             sigma = 1)$
    gen(n, computed = select(dat_n_300[1:n, ], x, e)) %>%
    mutate(n = paste0("N: ", n))
}) %>%

  # Combined with data for 300 observations
  bind_rows(dat_n_300) %>%
  mutate(n = factor(n, levels = c("N: 50", "N: 100", "N: 300"))) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~n, scales = "free", ncol = 3) +
  xlab("Fitted values") +
  ylab("Residuals")
```

## Experimental setup

### Controlling the strength of the signal

As summarised in Table \ref{tab:model-factor-table}, three additional parameters $n$, $\sigma$ and $b$ are used to control the strength of the signal so that different difficulty levels of lineups are generated, and therefore, the estimated power curve will be smooth and continuous. Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ are used in data collection periods I and II respectively. Figure \ref{fig:different-sigma} and \ref{fig:different-b} demonstrate the impact of these two parameters. A large value of $\sigma$ will increase the variation of the error of the non-linearity model and decrease the visibility of the visual pattern. The parameter $b$ controls the standard deviation of the error across the support of the predictor. Given $x \neq a$, a larger value of $b$ will lead to a larger ratio of the variance at $x$ to the variance at $x - a = 0$, making the visual pattern more obvious.

Three different sample sizes are used (n = 50, 100, 300) in all three data collection periods. It can be observed from Figure \ref{fig:different-n} that with fewer data points drawn in a residual plot, the visual pattern is more difficult to be detected.

### Effect size

Effect size in statistics measures the strength of the signal relative to the noise. It is surprisingly difficult to quantify in general, even for simulated data as used in this experiment.

For the non-linearity model, the key items defining effect size are sample size ($n$) and variance of the error term ($\sigma^2$), and so effect size would be roughly calculated as $\sqrt{n}/{\sigma}$. As sample size increases the effect size would increase, but as variance increases the effect size decreases. However, it is not clear how the additional parameter for the model polynomial order, $k$, should be incorporated. Intuitively, the large $k$ means more complex pattern, which likely means effect size would decrease. For the purposes of our calculations we have chosen to use an approach based on Kullback-Leibler divergence [@kullback1951information], coupled with simulation. This formulation defines effect size to be:

$$E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right)$$
\noindent where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,
$\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a$ is the residual operator, $\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ is the hat matrix, $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients, and $\sigma^2\boldsymbol{I}$ is the assumed covariance matrix of the error term when $H_0$ is true.

In the heteroskedasticity model, the key elements for measuring effect size are sample size, $n$, and the ratio of the biggest variance to smallest variance, $b$. Larger values of both would produce higher effect size. However, it is not clear how to incorporate the additional shape parameter, $a$. Thus the same approach is used here, where the formula can be written as:

$$E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)$$
\noindent where $\boldsymbol{V}$ is the actual covariance matrix of the error term.

Derivations for these equations are provided in the Appendix.

To compute the effect size for each lineup we simulate a sufficient large number of samples from the same model, in each sample, the number of observations $n$ is fixed. We then compute the effect size for each sample and take the average as the final value. This ensures lineups constructed with the same experimental factors will share the same effect size.

### Subject allocation

As shown in Table \ref{tab:model-factor-table}, there are a total of $4 \times 4 \times 3 \times 4 = 192$ and $3 \times 5 \times 3 \times 4 = 180$ number of combinations of parameter values for non-linearity model and heteroskedasticity model respectively. Three replications are made for each of the combination results in $192 \times 3 = 576$ and $180 \times 3 = 540$ lineups. In addition, each lineup is designed to be evaluated by five different subjects. After attempting some pilot studies internally, we decide to present a block of 20 lineups to every subject. And to ensure the quality of the survey data, two lineups with obvious visual patterns are included as attention checks. Thus, $576 \times 5 / (20-2) = 160$ and $540 \times 5 / (20-2) = 150$ subjects are recruited to satisfy the design of the data collection period I and II respectively.

As mentioned in Section \ref{power-of-the-tests}, $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using null lineups. Three replications are made for $3 \times 4 = 12$ combinations of common factors $n$ and fitted value distribution, results in $12 \times 3 = 36$ lineups included in data collection period III. In these lineups, the data of the data plot is generated from a model with zero effect size, while the data of the 19 null plots are generated using the same simulation method discussed in Section \ref{visual-test-procedure-based-on-lineups}. This generation procedure differs from the canonical Rorschach lineup procedure, which requires that all 20 plots are generated directly from the null model. However, these lineups serve the same fundamental purpose: to assess the number of visually interesting plots generated under $H_0$.

To account for the fact that our simulation method for these lineups is not the Rorschach procedure, we use the method suggested in @vanderplas2021statistical for typical lineups containing a data plot to estimate $\alpha$. 
We have included a sensitivity analysis in the Appendix to examine the impact of the variance of the $\alpha$ estimate on our findings.

All lineups consist of only null plots are planned to be evaluated by 20 subjects. 
However, presenting only these lineups to subjects are considered to be bad practices as subjects will lose interest quickly. 
Therefore, we plan to collect 6 more evaluations on the 279 lineups with uniform fitted value distribution, result in $(36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133$ subjects recruited for data collection period III.

### Collecting results

Subjects for all three data collection periods are recruited from an crowdsourcing platform called Prolific [@palan2018prolific]. Prescreening procedure is applied during the recruitment, subjects are required to be fluent in English, with $98\%$ minimum approval rate and 10 minimum submissions in other studies.

During the experiment, every subject is presented with a block of 20 lineups. A lineup consists of a randomly placed data plot and 19 null plots, which are all residual plots drawn with raw residuals on the y-axis and fitted values on the x-axis. An additional horizontal red line is added at $y = 0$ as a helping line.

The data of the data plot is simulated from one of two models described in Section \ref{simulating-departures-from-good-residuals}, while the data of the remaining 19 null plots are generated by the residual rotation technique discussed in Section \ref{visual-test-procedure-based-on-lineups}.

In every lineup evaluation, the subject is asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots are from others. If there is no noticeable difference between plots in a lineup, subjects are permitted to select zero plots without providing the reason. No subject are shown the same lineup twice. Information about preferred pronoun, age group, education, and previous experience in visual experiments are also collected. A subject's submission is only accepted if the data plot is identified for at least one attention check. Data of rejected submissions are discarded automatically to maintain the overall data quality.

# Results

## Overview

There are `r 160 * 18`, `r 150 * 18` and `r 133 * 18 - nrow(filter(vi_survey, null_lineup, !attention_check))` lineups evaluation made by 160, 150 and 133 subjects recruited for data collection periods I, II and III respectively. In the total of `r nrow(filter(vi_survey, !attention_check))` lineup evaluations, `r nrow(filter(vi_survey, type == "polynomial", !attention_check))` use lineups produced by the non-linearity model, and `r nrow(filter(vi_survey, type == "heteroskedasticity", !attention_check))` use lineups produced by the heteroskedasticity model. Besides, there are `r nrow(filter(vi_survey, attention_check))` attention checks and `r nrow(filter(vi_survey, null_lineup, !attention_check))` evaluations on null lineups needed for the estimate of $\alpha$ not included in the analysis. The collated dataset is provided in `vi_survey` of the `visage` `R` package.

In the following analysis, lineups with uniform fitted values will be the focus. Visual patterns are more likely to be revealed under a uniform distribution. Additionally, we have collected extra evaluations on these lineups, which will result in a more reliable analysis. Analysis of lineups with other fitted value distributions can be found in Section \ref{effect-of-fitted-value-distributions}.

## Power comparison of different tests

```{r cache = TRUE}
# Define the minimum and maximum effect size
min_poly_es <- vi_survey %>% 
  filter(!null_lineup, 
         !attention_check,
         type == "polynomial", 
         x_dist == "uniform") %>%
  pull(effect_size) %>%
  min()
max_poly_es <- vi_survey %>% 
  filter(!null_lineup, 
         !attention_check,
         type == "polynomial", 
         x_dist == "uniform") %>%
  pull(effect_size) %>%
  max()

# The glm model for RESET test
reset_poly_glmmod <- poly_conv_sim %>%
  filter(x_dist == "uniform") %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(reject = as.numeric(RESET4_p_value <= 0.05)) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  select(x_dist, effect_size, log_effect_size, offset0, reject) %>%
  glm(reject ~ effect_size - 1, family = binomial(), data = ., offset = offset0)

# The prediction for all conventional tests on poly model
pred_conv_poly <- poly_conv_sim %>%
  filter(x_dist == "uniform") %>%
  select(-RESET3_p_value, -(RESET5_p_value:RESET10_p_value), -F_p_value) %>%
  rename(RESET_p_value = RESET4_p_value) %>%
  pivot_longer(RESET_p_value:SW_p_value) %>%
  mutate(name = gsub("_p_value", " test", name)) %>%
  mutate(reject = value <= 0.05) %>%
  select(effect_size, name, reject) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  nest(dat = c(effect_size, offset0, reject)) %>%
  mutate(mod = map(dat, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map(mod, function(mod) {
    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
                         offset0 = log(0.05/0.95))
    result$power <- predict(mod, type = "response", newdata = result)
    result
  })) %>%
  select(-dat, -mod) %>%
  unnest(power) %>%
  mutate(log_effect_size = log(effect_size))

# Data needed by the visual test
visual_poly_dat <- vi_survey %>%
  filter(!null_lineup, 
         !attention_check, 
         type == "polynomial", 
         x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  select(effect_size, offset0, reject)

# The glm model for visual test
visual_poly_glmmod <- glm(reject ~ effect_size - 1, 
                          family = binomial(), 
                          data = visual_poly_dat, 
                          offset = offset0)

# The prediction of the visual test
pred_visual_poly <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
                               offset0 = log(0.05/0.95)) %>%
  mutate(power = predict(visual_poly_glmmod, 
                         type = "response",
                         newdata = data.frame(effect_size = effect_size,
                                              offset0 = offset0))) %>%
  mutate(log_effect_size = log(effect_size))

# The prediction of the boot visual test
pred_visual_poly_boot <- map_df(1:500, function(boot_id) {
  slice_sample(visual_poly_dat, n = nrow(visual_poly_dat), replace = TRUE) %>%
    nest(dat = c(effect_size, offset0, reject)) %>%
    mutate(mod = map(dat, 
                     ~glm(reject ~ effect_size - 1, 
                          family = binomial(), 
                          data = .x,
                          offset = offset0))) %>%
    mutate(power = map(mod, function(mod) {
      result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
                           offset0 = log(0.05/0.95))
      result$power <- predict(mod, type = "response", newdata = result)
      result
    })) %>%
    select(-dat, -mod) %>%
    unnest(power) %>%
    mutate(log_effect_size = log(effect_size)) %>%
    mutate(boot_id = boot_id)
  })
```

Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with uniform fitted values, against the natural logarithm of the effect $log_e(E)$, with a 5% significance level. At the bottom of the figure \ref{fig:polypower}, there are a sequence of example residual plots with increasing levels of $log_e(E)$. Readers can evaluate them from left to right and determine at which level the departure from a good residual plot becomes detectable.

As discussed in Section \ref{conventionally-testing-for-departures}, many conventionally tests are available for detecting residual departures. Implementation-wise, the built-in R package `stats` provides some commonly used residual-based tests, such as Shapiro-Wilk test. A more comprehensive collection of regression diagnostics tests can be found in the R package `lmtest` [@lmtest]. In terms of heteroskedasticity diagnostics, the R package `skedastic` [@skedastic] collects and implements 25 existing conventional tests published since 1961.

We pick RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`. Among them, RESET test is the only exact and appropriate test in this scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are approximate and inappropriate tests. Their estimated power is shown in Figure \ref{fig:polypower}. To set up the RESET test, we include different powers of fitted values as proxies. According to @ramsey_tests_1969, there are no general rules for the power of the fitted values needed by the RESET test, but it finds power up to four is usually sufficient. Thus, we follow this guideline to conduct the RESET test. For the Breusch-Pagan test, the choice of predictors in the auxiliary regression is left to the user [@breusch_simple_1979]. But as @waldman1983note suggested, it is a good choice for the set of auxiliary predictors in the Breusch-Pagan test be the same as the White test. Thus, we include both $x$ and $x^2$ in the auxiliary regression.

Figure \ref{fig:heterpower} is similar to Figure \ref{fig:polypower}, but shows corresponding information on lineups produced by the heteroskedasticity model. In this scenario, the visual test is compared to an approximate test - Breusch-Pagan test, and two other inappropriate tests - RESET test and Shapiro-Wilk test.

For non-linearity patterns, the power curve of RESET test climbs aggressively from
`r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(0), offset0 = log(0.05/0.95)), type = "response")) * 100)`% to `r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(2.5), offset0 = log(0.05/0.95)), type = "response")) * 100)`% as $log_e(E)$ increases from 0 to 2.5, while power of other tests respond inactively to the change of effect, showing that RESET test is much more sensitive to weak non-linear structure. Meanwhile, no noticeable residual departures can be spotted from the example residual plots.

In terms of heteroskedasticity patterns, the power of Breusch-Pagan test is constantly greater than the power of visual test. At $log_e(E) \approx 2.5$, where the power curve of the visual test remains at a low level, the Breusch-Pagan test has around 50% chance of rejecting $H_0$. Similarly, the visual feature is nearly unobservable from the example residual plots at this level of effect size.

The power of visual test arises steadily as $log_e(E)$ increases from 3 to 5 for both non-linearity patterns and heteroskedasticity patterns, suggesting that the effect starts to make significant impact on the degree of the presence of the designed visual features. This can also be observed from the example residual plots that when $log_e(E) = 3.5$, a weak "S-shape" and a weak "right-triangle" shape are presented in Figure \ref{fig:polypower} and Figure \ref{fig:heterpower} respectively. The visual pattern becomes clearer as $log_e(E)$ increases. At $log_e(E) = 5$, the power of visual tests for both patterns reaches almost 100%.

Power curves of inappropriate tests show improvement as the effect increases but at a lower rate than the visual test in both scenarios. This coincides the point made by @cook1982residuals that residual-based tests for a specific type of model defect may be sensitive to other types of model defects. The power curve of RESET test remains at around 5% in Figure \ref{fig:heterpower} since there are no non-linear terms leave out in the heteroskedasticity model and $H_0$ of the test is always satisfied.

Overall, the power comparison suggests that conventional tests differs significantly from visual tests in two regression diagnostics scenarios designed by us. Visual test have much higher tolerance of the residual departures than the conventional test. Since fail to reject $H_0$ in a visual test literally means that there are no obvious visual discoveries found in the residual plot, analysts and the general public as the consumers of the output may not be convinced of the existence of significant residual departures in spite of the rejection of $H_0$ given by the conventional test. Even if the rejection is accepted, the model violation may be considered as impactless due to the fact that they are not clearly visible. Besides, the sensitivity of the conventional test to weak residual departures could also distract and discourage analysts from finding simple but good linear approximation to the data. The rejection of $H_0$ because of human acceptable and negligible residual departures is not practically meaningful and useful to analysts and decision makers. In contrast, if the strict correctness of the model assumption is of particular interest, conducting a conventional test is still the recommended choice based on the findings.

```{r}
# Analyse uniform polynomial data
p <- ggplot() +

  geom_point(data = visual_poly_dat,
             aes(log(effect_size), reject),
             alpha = 0.15) +

  # Draw conventional test power curve
  geom_line(data = pred_conv_poly,
            aes(log_effect_size, power, col = name),
            size = 1) +

  # Draw boot visual test curves
  geom_line(data = pred_visual_poly_boot,
            aes(log_effect_size, power, col = "Visual test", group = boot_id),
            size = 1,
            alpha = 0.01) +

  # Draw visual test power curve
  geom_line(data = pred_visual_poly,
            aes(log_effect_size, power, col = "Visual test"), 
            size = 1) +

  # Theme
  theme_light(base_size = 5) +
  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  labs(col = "", size = "# lineups")
```

```{r cache = TRUE}
# target_sigma is obtained by evaluating a vector of sigma and picking those 
# with expected effect 0 ~ 5

target_sigma <- c(3.65, 2.84, 2.21, 1.72, 1.34, 1.04, 0.81, 0.63, 0.49, 0.38, 0.3)

# Calculate the effect size 
es <- map_dbl(target_sigma, 
              ~log(poly_model(shape = 2, 
                              x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, 
                              sigma = .x)$
                     average_effect_size(n = 100, type = "kl")
                   )
              )
```

```{r polypower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for non-linear patterns (uniform fitted values only). Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups. Surrounding lines of the visual test show the estimated power of 500 bootstrap samples. Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by vertical lines in main plot. Where would you draw the line of too much non-linearity in the residuals? For the RESET test this is around log effect size 2, but for the visual test it is around 3.5.", dev = 'png', dpi = 450}

# 106
# 112
# 113
# 115
# 132
# 133
# 144!
# 147
set.seed(132)

# Base data
ex_dat <- poly_model(shape = 2, 
                     x = rand_uniform(-1, 1), 
                     sigma = 1)$gen(n = 100)
plist <- list()
j <- 0

# Obtain data at different effect. Reuse x and recompute e.
for (i in target_sigma) {
  j <- j + 1
plist[[j]] <- poly_model(shape = 2, 
                         x = rand_uniform())$
                  gen(n = 100, 
                      computed = ex_dat %>%
                        mutate(e = sqrt(i) * e) %>%
                        select(x, e)
                      ) %>%
  VI_MODEL$plot(theme = theme_light(base_size =  5), 
                remove_axis = TRUE, 
                remove_grid_line = TRUE, size = 0.25, stroke = 0.25) +
    ggtitle(round(es[j], 1)) +
    theme(aspect.ratio = 0.8)
}

# Use area() to design layout
map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
  reduce(c) %>%
  c(patchwork::area(3, 1, 20, 11), .) %>%
  patchwork::wrap_plots(append(plist, list(p), after = 0), 
                        design = .)
```

```{r cache = TRUE}
# Define the minimum and maximum effect size
min_heter_es <- vi_survey %>% 
  filter(!null_lineup, 
         !attention_check,
         type == "heteroskedasticity", 
         x_dist == "uniform") %>%
  pull(effect_size) %>%
  min()
max_heter_es <- vi_survey %>% 
  filter(!null_lineup, 
         !attention_check,
         type == "heteroskedasticity", 
         x_dist == "uniform") %>%
  pull(effect_size) %>%
  max()

# The prediction for all conventional tests on heter model
pred_conv_heter <- heter_conv_sim %>%
  filter(x_dist == "uniform") %>%
  select(-RESET3_p_value, -(RESET5_p_value:RESET10_p_value), -F_p_value) %>%
  rename(RESET_p_value = RESET4_p_value) %>%
  pivot_longer(RESET_p_value:SW_p_value) %>%
  mutate(name = gsub("_p_value", " test", name)) %>%
  mutate(reject = value <= 0.05) %>%
  select(effect_size, name, reject) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  nest(dat = c(effect_size, offset0, reject)) %>%
  mutate(mod = map(dat, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map(mod, function(mod) {
    result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
                         offset0 = log(0.05/0.95))
    result$power <- predict(mod, type = "response", newdata = result)
    result
  })) %>%
  select(-dat, -mod) %>%
  unnest(power) %>%
  mutate(log_effect_size = log(effect_size))

# Data needed by the visual test
visual_heter_dat <- vi_survey %>%
  filter(!null_lineup, 
         !attention_check, 
         type == "heteroskedasticity", 
         x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  select(effect_size, offset0, reject)

# The glm model for visual test
visual_heter_glmmod <- glm(reject ~ effect_size - 1, 
                           family = binomial(), 
                           data = visual_heter_dat, 
                           offset = offset0)

# The prediction of the visual test
pred_visual_heter <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
                                offset0 = log(0.05/0.95)) %>%
  mutate(power = predict(visual_heter_glmmod, 
                         type = "response",
                         newdata = data.frame(effect_size = effect_size,
                                              offset0 = offset0))) %>%
  mutate(log_effect_size = log(effect_size))

# The prediction of the boot visual test
pred_visual_heter_boot <- map_df(1:500, function(boot_id) {
  slice_sample(visual_heter_dat, n = nrow(visual_heter_dat), replace = TRUE) %>%
    nest(dat = c(effect_size, offset0, reject)) %>%
    mutate(mod = map(dat, 
                     ~glm(reject ~ effect_size - 1, 
                          family = binomial(), 
                          data = .x,
                          offset = offset0))) %>%
    mutate(power = map(mod, function(mod) {
      result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
                           offset0 = log(0.05/0.95))
      result$power <- predict(mod, type = "response", newdata = result)
      result
    })) %>%
    select(-dat, -mod) %>%
    unnest(power) %>%
    mutate(log_effect_size = log(effect_size)) %>%
    mutate(boot_id = boot_id)
  })
```

```{r}
# Analyse uniform heteroskedasticity data
p <- ggplot() +

  geom_point(data = visual_heter_dat,
             aes(log(effect_size), reject),
             alpha = 0.15) +

  # Draw conventional test power curve
  geom_line(data = pred_conv_heter,
            aes(log_effect_size, power, col = name),
            size = 1) +

  # Draw boot visual test curves
  geom_line(data = pred_visual_heter_boot,
            aes(log_effect_size, power, col = "Visual test", group = boot_id),
            size = 1,
            alpha = 0.01) +

  # Draw visual test power curve
  geom_line(data = pred_visual_heter,
            aes(log_effect_size, power, col = "Visual test"), 
            size = 1) +

  # Theme
  theme_light(base_size = 5) +
  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
  # scale_size_manual(values = seq(1, 9)) +
  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  labs(col = "", size = "# lineups")
```

```{r cache = TRUE}
# target_b is obtained by evaluating a vector of b and picking those 
# with expected effect 0 ~ 5

target_b <- c(0.17, 0.23, 0.32, 0.46, 0.67, 1.04, 1.7, 3, 6.3, 16.2, 61)

# Calculate the effect size 
es <- map_dbl(target_b, 
              ~log(heter_model(a = 1, 
                               b = .x, 
                               x = {raw_x <- rand_normal(sigma = 0.3);closed_form(~stand_dist(raw_x))})$
                     average_effect_size(n = 100, type = "kl")
                   )
              )
```

```{r heterpower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only). Main plot shows the power curves, with dots indicating human evaluations of lineups. Surrounding lines of the visual test show the estimated power of 500 bootstrap samples. Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by vertical lines in main plot. Where would you draw the line of too much heteroskedasticity in the residuals? For the BP test this is around log effect size 2.5, but for the visual test it is around 3.5.", dev = 'png', dpi = 450}

# 10091
# 101
# 105
# 106
# 112
# 119
# 133
set.seed(135)

# Base data
ex_dat <- heter_model(a = 1, 
                      x = rand_uniform(-1, 1), 
                      b = 1)$gen(n = 100)
plist <- list()
j <- 0

# Obtain data at different b
for (i in target_b) {
  j <- j + 1
  plist[[j]] <- heter_model(a = 1, 
                            b = i, 
                            x = rand_uniform(-1, 1))$
                  gen(n = 100, computed = select(ex_dat, x, e)) %>%
  VI_MODEL$plot(theme = theme_light(base_size =  5),
                remove_axis = TRUE,
                remove_grid_line = TRUE, size = 0.25, stroke = 0.25) +
    ggtitle(round(es[j], 1)) +
    theme(aspect.ratio = 0.8)
}

# Use area() to design layout
map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
  reduce(c) %>%
  c(patchwork::area(3, 1, 20, 11), .) %>%
patchwork::wrap_plots(append(plist, list(p), after = 0), 
                      design = .)
```

## Comparison of test decisions based on $p$-values

```{r}
# Dataset for p-value comparison
p_value_cmp_dat <- vi_survey %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  filter(x_dist == "uniform") %>%
  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
  mutate(conv_reject = ifelse(conventional_p_value <= 0.05, "Reject", "Not"),
         reject = ifelse(p_value <= 0.05, "Reject", "Not")) %>%
  mutate(conv_reject = factor(conv_reject, levels = c("Reject", "Not")),
         reject = factor(reject, levels = c("Reject", "Not"))) %>%
  select(type, conv_reject, reject)
```

The power comparison illustrates that appropriate conventional tests will reject $H_0$ more aggressively than visual tests. In this section, we explore how often they agree with each other by investigating the rejections for the two model designs based on $p$-values for each lineup.

Figure \ref{fig:p-value-comparison} provides a mosaic plot showing the rejection rate of visual tests and conventional tests for both non-linearity patterns and heteroskedasticity patterns.

For lineups containing non-linearity patterns, conventional tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% and visual tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time. Of the lineups rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by the visual test, that is, approximately half as many as the conventional test. There are no lineups that are rejected by the visual test but not by the conventional test.

In terms of lineups containing heteroskedasticity patterns, `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by conventional tests, while `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by visual tests. When the conventional test rejects a lineup, there is a great chance (`r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%) the visual test will also reject it.

```{r}
nr <- p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Not") %>% count(reject)
```

Surprisingly, the visual test rejects `r nr$n[1]` of the `r nr$n[1]+nr$n[2]` (`r round(nr$n[1]/(nr$n[1]+nr$n[2])*100, 0)`%) of lineups where the conventional test does not reject. Figure \ref{fig:heter-example} shows this lineup. The data plot in position seventeen displays a relatively strong heteroskedasticity pattern, and has a strong effect size ($log_e(E)=4.02$). This is reflected by the visual test $p\text{-value} = 0.026$. But the Breusch-Pagan test $p\text{-value} = 0.056$, is slightly above the significance cutoff of $0.05$. This lineup was evaluated by 11 subjects, it has experimental factors $a = 0$ ("butterfly" shape), $b = 64$ (large variance ratio), $n = 50$ (small sample size), and a uniform distribution for the fitted values. It must be the small sample size that may have resulted in the lack of detection.

```{r p-value-comparison, fig.width = 8, fig.height = 4, fig.cap = "Rejection rate ($p$-value $\\leq0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot. The visual test rejects less frequently than the conventional test. We would generally expect the visual test to only reject when the conventional test does. Surprisingly, one lineup containing a heteroskedasticity pattern does not follow this rule. "}

library(ggmosaic)

# Mosaic plot
p_value_cmp_dat %>%
  ggplot() +
  geom_mosaic(aes(x = ggmosaic::product(reject, conv_reject), 
                  fill = reject)) +
  facet_grid(~type) +
  ylab("Visual tests") +
  xlab("Conventional tests") +
  labs(fill = "Conventional tests") +
  scale_fill_brewer("", palette = "Dark2") +
  theme_bw() +
  theme(legend.position = "none") +
  coord_fixed()
```

```{r heter-example, fig.height = 7, fig.width = 7, fig.cap = 'The single heteroskedasticity lineup that is rejected by the visual test but not by the BP test. The data plot at panel 17 contains a "butterfly" shape. It has effect size $ = 3.76$, somewhat surprising that it is not detected by the BP test.'}
# plot heter-331
VI_MODEL$plot_lineup(vi_lineup$heter_331$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

## Effect of shape of non-linearity

A primary factor contributes to the non-linearity model is the shape of the non-linearity pattern. According to Figure \ref{fig:poly-power-uniform-j}, conventional tests have higher power in testing shapes constructed from lower orders of Hermite polynomials. But the chance of detecting the "triple-U" shape drops significantly compared to other shapes. To understand why this is, one needs to return to the way the RESET test is applied. It requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four [@ramsey_tests_1969]. However, the "triple-U" shape constructed from the Hermite polynomials use power up to 18. If the RESET test had been applied using a higher power no less than six, the power curve of "triple-U" shape will be only slight lower than the power curve of "M" shape. The recommendation of the polynomial power for the RESET test should be revised, perhaps. This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data generating process in order to set the parameters for the test.

For visual tests, based on the orders of the Hermite polynomials, we expect the "U" shape will be the easiest one to be detected by subjects followed by the "S", "M" and "triple-U" shape. From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves are mostly aligned with our expectation, except for the "M" shape, which is as easy to be detected as the "S" shape. This implies, unlike the conventional test, the visibility of the shape do not strictly follow the degree of the polynomials.

```{r poly-power-uniform-j, fig.height = 4, fig.cap = 'Power of conventional tests and visual tests on lineups containing four different non-linearity shapes. Power curves with higher order of non-linearity are drawn with deeper colours. The default RESET tests under-perform significantly in detecting "triple-U" shape. To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test. But this means the order needs to be known prior to testing.'}

dat_poly_j <- poly_conv_sim %>%
  filter(x_dist == "uniform") %>%
  mutate(reject = as.numeric(RESET4_p_value <= 0.05)) %>%
  mutate(shape = c("U", "S", "M", "triple-U")[shape]) %>%
  mutate(conv_or_not = "Conventional") %>%
  bind_rows(vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform", type == "polynomial") %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(shape = c("U", "S", "M", "triple-U")[shape]) %>%
  mutate(conv_or_not = "Visual")) %>%
  mutate(shape = factor(shape, levels = c("U", "S", "M", "triple-U"))) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  select(shape, effect_size, offset0, reject, conv_or_not)

dat_poly_j %>%
  nest(dat = c(effect_size, offset0, reject)) %>%
  mutate(mod = map(dat, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map(mod, function(mod) {
    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
                         offset0 = log(0.05/0.95))
    result$power <- predict(mod, type = "response", newdata = result)
    result
  })) %>%
  select(-dat, -mod) %>%
  unnest(power) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_line(aes(log_effect_size, power, col = shape), size = 1) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  scale_color_manual(values = rcartocolor::carto_pal(4, "TealGrn")) +
  theme_light() +
  theme(legend.position = "bottom") +
  facet_wrap(~conv_or_not) +
  labs(col = "Non-linearity shape")
```

## Effect of shape of heteroskedasticity

We have also investigated the impact of different heteroskedasticity shapes on power of conventional tests and visual tests. In theory, the "left-triangle" and the "right-triangle" shapes are functionally identical from the point of view of a Breusch-Pagan test. As shown in Figure \ref{fig:heter-power-uniform-a}, this is indeed the case where little difference between the power curves can be perceived. Similarly, visual tests should have the same power in detecting these two shapes if they are equally likely to be identified. However, it can be observed from Figure \ref{fig:heter-power-uniform-a} that the power curve of the "left-triangle" shape is higher than the one of the "right-triangle" shape, indicating a potential favour of orientation by human, which is worth to be explored in future studies. Besides, the chance of detecting the "butterfly" shape are higher than other two shapes in both conventional and visual testing.

```{r heter-power-uniform-a, fig.height = 4, fig.cap = 'Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$. The visual test has an asymmetric power in testing the "left-triangle" shape and "right-triangle" shape, while these two shapes are equivalent in conventional testing.'}

dat_heter_a <- heter_conv_sim %>%
  filter(x_dist == "uniform") %>%
  mutate(reject = as.numeric(BP_p_value <= 0.05)) %>%
  mutate(a = c("left-triangle", "butterfly", "right-triangle")[a + 2]) %>%
  mutate(conv_or_not = "Conventional") %>%
  bind_rows(vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform", type == "heteroskedasticity") %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(a = c("left-triangle", "butterfly", "right-triangle")[a + 2]) %>%
  mutate(conv_or_not = "Visual")) %>%
  mutate(a = factor(a, levels = c("left-triangle", "butterfly", "right-triangle"))) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  select(a, effect_size, offset0, reject, conv_or_not)

# Manually fit the model
dat_heter_a %>%
  nest(dat = c(effect_size, offset0, reject)) %>%
  mutate(mod = map(dat, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map(mod, function(mod) {
    result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
                         offset0 = log(0.05/0.95))
    result$power <- predict(mod, type = "response", newdata = result)
    result
  })) %>%
  select(-dat, -mod) %>%
  unnest(power) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_line(aes(log_effect_size, power, col = a), size = 1) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  scale_color_manual(values = c(rcartocolor::carto_pal(3, "Earth")[1], "grey", rcartocolor::carto_pal(3, "Earth")[3])) +
  theme_light() +
  theme(legend.position = "bottom") +
  facet_wrap(~conv_or_not) +
  labs(col = "Heteroskedasticity shape")
```

## Effect of fitted value distributions

The prediction in a regression model is $E(Y|X)$, that is, it is conditional on observed values of the predictors. The distribution of $X$, or consequently $\hat{Y}$, may however affect the ability to read any patterns in the residuals. The four distributions of fitted values, used in this experiment, were designed to examine this.

Figure \ref{fig:different-x-dist-poly-power} illustrates the impact of fitted value distribution on the power of conventional tests and visual tests. For conventional tests, only the power curves of appropriate tests which are RESET tests and Breusch-Pagan tests for non-linearity pattern and heteroskedasticity pattern respectively are shown. For visual tests, note that we have collected more evaluations on lineup with uniform fitted value distribution, and the number of evaluations on a lineup will affect the power of the visual test. To have a fair comparison, for lineups with uniform fitted value distribution, we randomly sample five evaluations from the total eleven evaluations to estimate the power curves. The sampling process is repeated 500 times to produce an indication of the variation of the power.

In terms of visual tests, we do not observe significant difference between the power curve of the uniform distribution and the power curve of the discrete uniform distribution. Normal and lognormal distributions produce lower power than uniform and discrete uniform distributions in regards of residual departure patterns, but the gap is larger when the residual plot contains heteroskedasticity pattern. This suggests that residual departures displayed with evenly distributed fitted values are visually more attractive and effective. Besides, the skewness of the fitted value distribution plays a role in the degree of presence of visual features as lognormal distribution produce the lowest power for both non-linearity and heteroskedasticity patterns. A symmetric fitted value distribution can better reveal the underlying shape.

Although it is unusual to consider fitted value distribution or predictor distribution in power analysis, it is not so surprised that the power of conventional tests vary because of fitted value distribution. For instance, the power of the RESET test which is effectively a F test depends on the realizations of the additional predictors included in the auxiliary regression equation [@jamshidian2007study; @olvera2019relationship; @zhang2018practical]. And those additional predictors are linear transformation of random vectors following certain predictor distributions. What is truly unexpected is the huge variability in power of conventional tests compared to visual tests. The discrete uniform distribution constantly produce the highest power while the normal distribution and the lognormal distribution produce the lowest power for non-linearity and heteroskedasticity pattern respectively. The difference in power corresponding to different distributions peak at around 90% when the effect size is at a moderate level. This huge difference is undesirable as is not uncommon for analysts to omit the factor of predictor distribution in power calculation and sample size calculation.

```{r get-5-eval, cache = TRUE}
dat_uniform_5 <- map_df(1:500, function(boot_id) {
  vi_survey %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  slice_sample(n = 5) %>%
  ungroup() %>%
  calc_p_value_multi(lineup_id = unique_lineup_id,
                     detect = detect,
                     n_sel = num_selection,
                     alpha = alpha) %>%
  mutate(boot_id = boot_id)
}) %>%
  left_join(select(vi_survey, unique_lineup_id, effect_size, x_dist, type)) %>%
  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(conv_or_not = "Visual") %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  select(-unique_lineup_id, -p_value)

dat_uniform_5_pred <- dat_uniform_5 %>%
  nest(dat = c(effect_size, offset0, reject)) %>%
  mutate(mod = map(dat, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map2(mod, type, function(mod, type) {
    this_min_es <- ifelse(type == "heteroskedasticity", min_heter_es, min_poly_es)
    this_max_es <- ifelse(type == "heteroskedasticity", max_heter_es, max_poly_es)
    result <- data.frame(effect_size = seq(this_min_es[1], this_max_es[1], 0.1),
                         offset0 = log(0.05/0.95))
    result$power <- predict(mod, type = "response", newdata = result)
    result
  })) %>%
  select(-dat, -mod) %>%
  unnest(power) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(type = factor(stringr::str_to_title(type), 
                       levels = c("Non-Linearity", "Heteroskedasticity")))
```

```{r different-x-dist-poly-power, fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for conventional tests and visual tests. Power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and Breusch-Pagan tests respectively. Power curves of visual tests are estimated using five evaluations on each lineup. For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations. The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines. The fitted value distribution has greater impact on the power of conventional tests than visual tests. Meanwhile, uneven distributions including normal and lognormal distributions genearlly produce lower power.", dev = 'png', dpi = 450}

diff_x_dist_power_pred <- list(
  # Visual inference data
  vi_survey %>%
    filter(!null_lineup) %>%
    filter(!attention_check) %>%
    group_by(unique_lineup_id) %>%
    summarise(across(everything(), first)) %>%
    filter(x_dist != "uniform") %>%
    mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
    mutate(reject = as.numeric(p_value <= 0.05)) %>%
    mutate(offset0 = log(0.05/0.95)) %>%
    select(type, x_dist, effect_size, offset0, reject) %>%
    mutate(conv_or_not = "Visual"),

  # Conventional simulation (poly)
  poly_conv_sim %>%
    mutate(reject = RESET4_p_value < 0.05) %>%
    mutate(offset0 = log(0.05/0.95)) %>%
    mutate(type = "non-linearity") %>%
    select(type, x_dist, effect_size, offset0, reject) %>%
    mutate(conv_or_not = "Conventional"),

  # Conventional simulation (heter)
  heter_conv_sim %>%
    mutate(reject = BP_p_value < 0.05) %>%
    mutate(offset0 = log(0.05/0.95)) %>%
    mutate(type = "heteroskedasticity") %>%
    select(type, x_dist, effect_size, offset0, reject) %>%
    mutate(conv_or_not = "Conventional")
) %>%

  # Bind rows
  map_df(~.x) %>%

  nest(dat = c(effect_size, offset0, reject)) %>%
  mutate(mod = map(dat, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map2(mod, type, function(mod, type) {
    this_min_es <- ifelse(type == "heteroskedasticity", min_heter_es, min_poly_es)
    this_max_es <- ifelse(type == "heteroskedasticity", max_heter_es, max_poly_es)
    result <- data.frame(effect_size = seq(this_min_es[1], this_max_es[1], 0.1),
                         offset0 = log(0.05/0.95))
    result$power <- predict(mod, type = "response", newdata = result)
    result
  })) %>%
  select(-dat, -mod) %>%
  unnest(power) %>%
  mutate(log_effect_size = log(effect_size))

diff_x_dist_power_pred %>%
  mutate(x_dist = ifelse(x_dist == "even_discrete", 
                         "Discrete uniform", 
                         stringr::str_to_title(x_dist))) %>%
  mutate(type = factor(stringr::str_to_title(type), 
                       levels = c("Non-Linearity", 
                                  "Heteroskedasticity"))) %>%
  mutate(x_dist = factor(x_dist, 
                         levels = c("Uniform", 
                                    "Normal", 
                                    "Lognormal", 
                                    "Discrete uniform"))) %>%

  ggplot() +

  # Draw all smooth lines
  geom_line(aes(log_effect_size, power, col = x_dist),
            size = 1) +

  # Draw boot uniform (5 eval) lines
  geom_line(data = dat_uniform_5_pred,
            aes(log_effect_size, power, col = "Uniform", group = boot_id),
            size = 1,
            alpha = 0.01) +

  # Theme
  theme_light() +
  scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(4,1,3,2)]) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  labs(col = "", linetype = "Test") +
  facet_grid(conv_or_not~type, scales = "free_x") +
  theme(legend.position = "bottom")
```

# Conclusion

Motivated by the advice of regression analysis experts, that residual plots as opposed to conventional tests are an indispensible methods for assessing model fit, a human subjects experiment was conducted using visual inference. The experiment tested two primary departures from good residuals: non-linearity and heteroskedasticity.

The experiment found that conventional residual-based statistical tests are more sensitive to weak residual departures from model assumptions than visual tests as would be evaluated by humans. That is, conventional tests conclude there are problems with the model fit almost twice as often as humans would. They often reject when departures in the form of non-linearity and heteroskedasticity are not visible to a human.

One might say that this is correct behaviour, but it can be argued that the conventional tests are rejecting when it is not necessary. Many of these rejections happen even when downstream analysis and results would not be significantly affected by the small departures from a good fit. The results from human evaluations provide a more practical solution, which reinforces the statements from regression experts that residual plots are an indispensible method for model diagnostics.

Now it is important to note that residual plots need to be delivered as a lineup, where it is embedded in a field of null plots. A residual plot may contains many visual features, but some are caused by the characteristics of the predictors and the randomness of the error, not by the violation of the model assumptions. These irrelevant visual features have a chance to be filtered out by subjects with a comparison to null plots, results in a set of more accurate visual findings. This enables a careful calibration for reading structure in residual plots.

However, human evaluation of residuals is expensive. It is time-consuming, laborious and unfriendly to vision-impaired people. This is another reason why it often appears to be ignored. With the availability of sophisticated computer vision algorithms today, the goal of this work is to form the basis of providing automated residual plot reading. The findings suggest the strong demand of graphical inspection in regression diagnostics, so developing an automatic visual inference system to evaluate lineups of residual plots is valuable. We plan to build a completed open-source system in an `R` package and provide a web interface such as a website for public to interact with. Details about this system will be discussed in our next paper.

The experiment also revealed some interesting details about how residual plots are read. For the most part, the visual test performed very similarly to the appropriate conventional test only with the power curve shifted in the less sensitive direction. Unlike the conventional tests, where one needs to specifically test for non-linearity or heteroskedasticity the visual test operated effectively across the range of departures from good residuals.

As expected, if the fitted value distribution is not uniform, there is a loss of power in the visual test. Structure is hardest to detect if fitted values are lognormal. Also, complex structure are generally harder to detect, but there are outliers. Under the designed scenarios in this paper, we find the visual test to be a more robust test against the change of fitted value distributions. A surprising finding was that the direction of heteroskedasticity appears to affect the ability to visually detect it, with wedge to the right being less detectable.

# Acknowledgements {-}

These `R` packages are used for the work: 
`cli` [@cli], `curl` [@curl], `dplyr` [@dplyr], `ggplot2` [@ggplot2], `jsonlite` [@jsonlite], `lmtest` [@lmtest], `mpoly` [@mpoly], `progress` [@progress], `tibble` [@tibble], `ggmosaic` [@ggmosaic], `purrr` [@purrr], `tidyr` [@tidyr], `readr` [@readr], `stringr` [@stringr], `here` [@here], `kableExtra` [@kableextra], `patchwork` [@patchwork], `rcartocolor` [@rcartocolor]. The study website is powered by `PythonAnywhere` [@pythonanywhere] and the `Flask` web framework [@flask]. The `jsPsych` framework [@jspsych] is used to create behavioral experiments that run in our study website.

The article was created with R packages `rticles` [@rticles], `knitr` [@knitr] and `rmarkdown` [@rmarkdown]. The project's Github repository (link here) contains all materials required to reproduce this article.

# Supplementary material {-}

The supplementary material is available at (link here). It includes more details about the experimental setup, the derivation of the effect size, the effect of data collection period, and the estimate of  $\alpha$.

\appendix

# Experiment setup

## Mapping of subjects to experimental factors

Mapping of subjects to experimental factors is an important part of experiment design. Essentially, we want to maximum the difference in factors exposed to a subject. For this purpose, we design an algorithm to conduct subject allocation. Let $L$ be a set of available lineups and $S$ be a set of available subjects. According to the experimental design, the availability of a lineup is associated with the number of subjects it can assign to. For lineups with uniform fitted value distribution, this value is 11. And other lineups can be allocated to at most five different subjects. The availability of a subject is associated with the number of lineups that being allocated to this subject. A subject can view at most 18 different lineups.

The algorithm starts from picking a random subject $s \in S$ with the minimum number of allocated lineups. It then tries to find a lineup $l \in L$ that can maximise the distance metric $D$ and allocate it to subject $s$. Set $L$ and $S$ will be updated and the picking process will be repeated until there is no available lineups or subjects.

Let $F_1,...,F_q$ be $q$ experimental factors, and $f_1, ...,f_q$ be the corresponding factor values. We say $f_i$ exists in $L_{s}$ if any lineup in $L_{s}$ has this factor value. Similarly, $f_if_j$ exists in $L_{s}$ if any lineup in $L_{s}$ has this pair of factor values. And $f_if_jf_k$ exists in $L_{s}$ if any lineup in $L_{s}$ has this trio of factor values. The distance metric $D$ is defined between a lineup $l$ and a set of lineups $L_{s}$ allocated to a subject $s$ if $L_{s}$ is non-empty:

\footnotesize

\begin{equation*}
D =
C - \sum_{1\leq i \leq q}I(f_i\text{ exists in }L_{s}) - \sum_{\substack{1\leq i \leq q-1 \\ i < j \leq q}}I(f_if_j\text{ exists in }L_{s}) - \sum_{\substack{1\leq i \leq q - 2 \\ i < j \leq q - 1 \\ j < k \leq q}}I(f_if_jf_k\text{ exists in }L_{s})
\end{equation*}

\normalsize
where $C$ is a sufficiently large constant such that $D > 0$. If $L_{s}$ is empty, we define $D = 0$.

The distance measures how different a lineup is from the set of lineups allocated to the subject in terms of factor values. Thus, the algorithm will try to allocate the most different lineup to a subject at each step.

## Data collection process

The survey data is collected via a self-hosted website designed by us. The complete architecture is provided in Figure \ref{fig:tech}. The website is built with the `Flask` [@flask] web framework and hosted on `PythonAnywhere` [@pythonanywhere]. It is configured to handle HTTP requests such that subjects can correctly receive webpages and submit responses. Embedded in the resources sent to subjects, the `jsPsych` front-end framework [@jspsych] instructs subjects' browsers to render an environment for running behavioral experiments. During the experiment, this framework will automatically collect common behavioral data such as response time and clicks on buttons. Subjects' responses are first validated by a scheduled `Python` script run on the server, then push to a Github repository. Lineup images shown to users are saved in multiple Github repositories and hosted in corresponding Github pages. The URLs to these images are resolved by `Flask` and bundled in HTML files.

Once the participant is recruited from Prolific [@palan2018prolific], it will be redirected to the entry page of our study website. An image of the entry page is provided in Figure \ref{fig:entry-page}. Then, the participant needs to submit the online consent form and fill in the demographic information as shown in \ref{fig:consent-form} and \ref{fig:metadata} respectively. Before evaluating lineups, participant also need to read the training page as provide in Figure \ref{fig:training-page} to understand the process. An example of the lineup page is given in Figure \ref{fig:lineup-page}. A half of the page is taken by the lineup image to attract participant's attention. The button to skip the selections for the current lineup is intentionally put in the corner of the bounding box with smaller font size, such that participants will not misuse this functionality.

```{r tech, out.width = "100%", fig.cap = "Diagram of online experiment setup. The server-side of the study website uses Flask as backend hosted on PythonAnywhere. And the client-side uses jsPsych to run experiment."}
knitr::include_graphics("figures/experiment_tech.pdf")
```

```{r entry-page, out.width = "100%", fig.cap = "The entry page of the study website."}
knitr::include_graphics("figures/message.png")
```

```{r consent-form, out.width = "100%", fig.cap = "The consent form provided in the study website."}
knitr::include_graphics("figures/consent_form.png")
```

```{r metadata, out.width = "100%", fig.cap = "The form to provide demographic information."}
knitr::include_graphics("figures/metadata.png")
```

```{r training-page, out.width = "100%", fig.cap = "The training page of the study website."}
knitr::include_graphics("figures/training.png")
```

```{r lineup-page, out.width = "100%", fig.cap = "The lineup page of the study website."}
knitr::include_graphics("figures/lineup1.png")
```

# Demographics

Along with the responses to lineups, we have collected a series of demographic information including age, pronoun, education background and previous experience in studies involved data visualization. Table \ref{tab:pronoun}, \ref{tab:age-group}, \ref{tab:education} and \ref{tab:experience} provide summary of the demographic data.

It can be observed from the tables that most participants have Diploma or Bachelor degrees, followed by High school or below and the survey data is gender balanced. Majority of participants are between 18 to 39 years old and there are slightly more participants who do not have previous experience than those who have.

```{r}
vi_survey %>%
  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
  group_by(exp, set) %>%
  summarise(across(everything(), first)) %>%
  count(pronoun) %>%
  mutate(exp = paste("Period", c("I", "II", "III")[exp - 2])) %>%
  pivot_wider(values_from = n, names_from = exp) %>%
  mutate(Total = `Period I` + `Period II` + `Period III`) %>%
  rename(Pronoun = pronoun) %>%
  add_row(Pronoun = "",
          `Period I` = sum(.$`Period I`),
          `Period II` = sum(.$`Period II`),
          `Period III` = sum(.$`Period III`),
          Total = sum(.$Total)) %>%
  mutate(`Period I %` = round(`Period I`/443*100, 1)) %>%
  mutate(`Period II %` = round(`Period II`/443*100, 1)) %>%
  mutate(`Period III %` = round(`Period III`/443*100, 1)) %>%
  mutate(`Total %` = round(Total/443*100, 1)) %>%
  select(Pronoun, 
         `Period I`, `Period I %`, 
         `Period II`, `Period II %`,
         `Period III`, `Period III %`,
         Total, `Total %`) %>%
  kableExtra::kable(format = "latex", 
                    col.names = c("Pronoun", 
                                  "Period I", "%", 
                                  "Period II", "%",
                                  "Period III", "%",
                                  "Total", "%"), 
                    booktabs = TRUE, label = "pronoun", caption = "Summary of pronoun distribution of subjects recuritted in this study.")
```

```{r}
vi_survey %>%
  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
  group_by(exp, set) %>%
  summarise(across(everything(), first)) %>%
  count(age_group) %>%
  mutate(exp = paste("Period", c("I", "II", "III")[exp - 2])) %>%
  pivot_wider(values_from = n, names_from = exp) %>%
  mutate(Total = `Period I` + `Period II` + `Period III`) %>%
  rename(`Age group` = age_group) %>%
  kableExtra::kable(format = "latex", booktabs = TRUE, label = "age-group", caption = "Summary of age distribution of subjects recuritted in this study.")
```

```{r}
vi_survey %>%
  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
  group_by(exp, set) %>%
  summarise(across(everything(), first)) %>%
  count(education) %>%
  mutate(exp = paste("Period", c("I", "II", "III")[exp - 2])) %>%
  pivot_wider(values_from = n, names_from = exp, values_fill = 0) %>%
  mutate(Total = `Period I` + `Period II` + `Period III`) %>%
  slice(c(2, 1, 3, 4, 5)) %>%
  rename(`Education` = education) %>%
  add_row(`Education` = "",
          `Period I` = sum(.$`Period I`),
          `Period II` = sum(.$`Period II`),
          `Period III` = sum(.$`Period III`),
          Total = sum(.$Total)) %>%
  mutate(`Period I` = paste0(`Period I`, " (", round(`Period I`/443*100, 1), "%)")) %>%
  mutate(`Period II` = paste0(`Period II`, " (", round(`Period II`/443*100, 1), "%)")) %>%
  mutate(`Period III` = paste0(`Period III`, " (", round(`Period III`/443*100, 1), "%)")) %>%
  mutate(Total = paste0(Total, " (", round(Total/443*100, 1), "%)")) %>%
  kableExtra::kable(format = "latex", booktabs = TRUE, label = "education", caption = "Summary of education distribution of subjects recuritted in this study.")
```

```{r}
vi_survey %>%
  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
  group_by(exp, set) %>%
  summarise(across(everything(), first)) %>%
  count(previous_experience) %>%
  mutate(exp = paste("Period", c("I", "II", "III")[exp - 2])) %>%
  pivot_wider(values_from = n, names_from = exp) %>%
  mutate(Total = `Period I` + `Period II` + `Period III`) %>%
  rename(`Previous experience` = previous_experience) %>%
  kableExtra::kable(format = "latex", booktabs = TRUE, label = "experience", caption = "Summary of previous experience distribution of subjects recuritted in this study.")
```

# Effect size derivation

Effect size can be defined as the difference of a parameter for a particular model or distribution,  or a statistic derived from a sample. Importantly, it needs to reflect the treatment we try to measure. Centred on a conventional statistical test, we usually can deduce the effect size from the test statistic by substituting the null parameter value. When considering the diagnostics of residual departures, there exist many possibilities of test statistics for a variety of model assumptions. Meanwhile, diagnostic plots such as the residual plot have no general agreement on measuring how strong a model violation pattern is. To build a bridge between various residual-based tests, and the visual test, we focus on the shared information embedded in the testing procedures, which is the distribution of residuals. When comes to comparison of distribution, Kullback-Leibler divergence [@kullback1951information] is a classical way to represent the information loss or entropy increase caused by the approximation to the true distribution, which in our case, the inefficiency due to the use of false model assumptions.

Following the terminology introduced by @kullback1951information, $P$ represents the measured probability distribution, and $Q$ represents the assumed probability distribution. The Kullback-Leibler divergence is defined as $\int_{-\infty}^{\infty}log(p(x)/q(x))p(x)dx$, where $p(.)$ and $q(.)$ denote probability densities of $P$ and $Q$.

Let $\boldsymbol{X}_a = (\boldsymbol{1}, \boldsymbol{X})$ denotes the $p$ regressors with $n$ observations, $\boldsymbol{R}_a = \boldsymbol{I} -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ denotes the residual operator, and let $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0},\sigma^2\boldsymbol{I})$ denotes the error. Using the Frisch–Waugh–Lovell theorem, residuals $\boldsymbol{e} = \boldsymbol{R}_a\boldsymbol{\varepsilon}$. Because $rank(\boldsymbol{R}_a) = n - p < n$, $e$ follows a degenerate multivariate normal distribution and does not have a density. Since the Kullback-Leibler divergence requires a proper density function, we need to simplify the covariance matrix of $\boldsymbol{e}$ by setting all the off-diagonal elements to 0. Then, the residuals will assumed to follow $N(\boldsymbol{0}, diag(\boldsymbol{R}_a\sigma^2))$ under the null hypothesis that the model is correctly specified. If the model is however misspecified due to omitted variables $\boldsymbol{Z}$, or a non-constant variance $\boldsymbol{V}$, the distribution of residuals can be derived as $N(\boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}_a\sigma^2))$ and $N(\boldsymbol{0}, diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a'))$ respectively.

By assuming both $P$ and $Q$ are multivariate normal density functions, the Kullback-Leibler divergence can be rewritten as 
$$KL = \frac{1}{2}\left(log\frac{|\Sigma_p|}{|\Sigma_q|} - n + tr(\Sigma_p^{-1}\Sigma_q) + (\mu_p - \mu_q)'\Sigma_p^{-1}(\mu_p - \mu_q)\right).$$

Then, we can combine the two residual departures into one formula

\small

\begin{equation}
\label{eq:effect-size}
KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}\boldsymbol{\mu}_z\right).
\end{equation}

\normalsize

When there are omitted variables but constant error variance, the formula can be reduced to

$$KL = \frac{1}{2}\left(\boldsymbol{\mu}_z^{T}(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right).$$

And when the model equation is correctly specified but the error variance is non-constant, the formula can be reduced to

$$KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2))\right).$$
Since we assume $\sigma = 1$ for the heteroskedasticity model, the final form of the formula is

$$KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right).$$

# Effect of data collection period

```{r poly-boxplot-lineup, fig.cap = 'A lineup of "letter-value" boxplots of weighted propotion of detect for lineups over different data collection periods for non-linearity model. Can you find the most different boxplot? The data plot is positioned in panel $2^3 - 1$.'}
library(lvplot)
sample_from_null <- function(dat, true_position) {
  dat %>%
    mutate(k = true_position) %>%
    bind_rows(map_df((1:20)[-true_position], function(k) {
    dat %>%
      mutate(exp = sample(exp),
             k = k)
  }))
}

vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform") %>%
  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
  mutate(exp = exp - 2) %>%
  group_by(exp, unique_lineup_id) %>%
  mutate(prop_detect = mean(weighted_detect)) %>%
  group_by(exp, unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  ungroup() %>%
  filter(type == "non-linearity") %>%
  select(type, exp, prop_detect) %>%
  mutate(exp = as.character(exp)) %>%
  sample_from_null(7) %>%
  ggplot() +
  geom_lv(aes(x = exp, y = prop_detect, fill = after_stat(LV)), varwidth = TRUE, k = 4) +
  scale_alpha_discrete(range = c(0.7, 0)) +
  facet_wrap(~k, ncol = 5) +
  theme_light() +
  scale_x_discrete(labels = c("I", "III")) +
  labs(fill = "Quantiles") +
  scale_fill_manual(labels = c("0.5", "[0.25, 0.75]", "[0.125, 0.875]", "[0.0625, 0.09375]"),
                    values = c("white", rev(RColorBrewer::brewer.pal(5, "Blues")))) +
  ylab("Weighted propotion of detect") +
  xlab("Data collection period")
```

```{r heter-boxplot-lineup, fig.cap = 'A lineup of "letter-value" boxplots of weighted propotion of detect for lineups over different data collection periods for heteroskedasticity model. Can you find the most different boxplot? The data plot is positioned in panel $2^4 - 2$.'}
vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform") %>%
  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
  mutate(exp = exp - 2) %>%
  group_by(exp, unique_lineup_id) %>%
  mutate(prop_detect = mean(weighted_detect)) %>%
  group_by(exp, unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  ungroup() %>%
  filter(type != "non-linearity") %>%
  select(type, exp, prop_detect) %>%
  mutate(exp = as.character(exp)) %>%
  sample_from_null(14) %>%
  ggplot() +
  geom_lv(aes(x = exp, y = prop_detect, fill = after_stat(LV)), varwidth = TRUE, k = 4) +
  facet_wrap(~k, ncol = 5) +
  theme_light() +
  scale_x_discrete(labels = c("II", "III")) +
  labs(fill = "Quantiles") +
  scale_fill_manual(labels = c("0.5", "[0.25, 0.75]", "[0.125, 0.875]", "[0.0625, 0.09375]"),
                    values = c("white", rev(RColorBrewer::brewer.pal(5, "Blues")))) +
  ylab("Weighted propotion of detect") +
  xlab("Data collection period")
```

We have the same type of model collected over different data collection periods, that may lead to unexpected batch effect.
Figure \ref{fig:poly-boxplot-lineup} and \ref{fig:heter-boxplot-lineup} provide two lineups to examine whether there is an actual difference across data collection periods for non-linearity model and heteroskedasticity model respectively. To emphasize the tail behaviour and display fewer outliers, we use the "letter-value" boxplot [@hofmann2017value] which is an extension of the number of "letter value" statistics to check the weighed proportion of detect over different data collection period. The weighted proportion of detect is calculated by taking the average of $c_i$ of a lineup over a data collection period. Within our research team, we can not identify the data plot from the null plots for these two lineups, result in $p$-values much greater than $5$%. Thus, there is no clear evidence of batch effect.

# Sensitivity analysis for $\alpha$

The parameter $\alpha$ used for the $p$-value calculation needs to be estimated from responses to null lineups. With a greater value of $\hat{\alpha}$, the $p$-value will be smaller, resulting in more lineups being rejected. However, The way we generate Rorschach lineup is not strictly the same as what suggested in @vanderplas2021statistical and @buja_statistical_2009. Therefore, we conduct a sensitivity analysis in this section to examine the impact of the variation of the estimator $\alpha$ on our primary findings.

The analysis is conducted by setting up several scenarios, where the $\alpha$ is under or overestimated by 12.5%, 25% and 50%. Using the adjusted $\hat{\alpha}$, we recalculate the $p$-value for every lineup and show the results in Figure \ref{fig:sensitivity}. It can be observed that there are some changes to $p$-values, especially when the $\hat{\alpha}$ is multiplied by 50%. However, Table \ref{tab:sensitivity} shows that adjusting $\hat{\alpha}$ will not result in a huge difference in rejection decisions. There are only a small percentage of cases where the rejection decision change. It is very unlikely the downstream findings will be affected because of the estimate of $\alpha$.

```{r}
upper_p <- map_df(c(0.125, 0.25, 0.5), function(modified) {
  calc_p_value_multi(vi_survey %>%
                       filter(!attention_check) %>%
                       filter(!null_lineup) %>%
                       filter(x_dist == "uniform") %>%
                       mutate(alpha_upper = alpha * (1 + modified)),
                     lineup_id = unique_lineup_id,
                     detect = detect,
                     n_sel = num_selection,
                     alpha = alpha_upper) %>%
  mutate(modified = 1 + modified) %>%
  mutate(adjusted_p_value = p_value)
})

lower_p <- map_df(c(0.125, 0.25, 0.5), function(modified) {
  calc_p_value_multi(vi_survey %>%
                       filter(!attention_check) %>%
                       filter(!null_lineup) %>%
                       filter(x_dist == "uniform") %>%
                       mutate(alpha_lower = alpha * (1 - modified)),
                     lineup_id = unique_lineup_id,
                     detect = detect,
                     n_sel = num_selection,
                     alpha = alpha_lower) %>%
  mutate(modified = 1 - modified) %>%
  mutate(adjusted_p_value = p_value)
})

upper_lower_p <- upper_p %>%
  bind_rows(lower_p) %>%
  select(-p_value)
```

```{r sensitivity, fig.cap = "Change of $p$-values with $\\hat{\\alpha}$ multiplied by $0.5$, $0.75$, $0.875$, $1.125$, $1.25$ and $1.5$. The vertical dashed line is to indicate $p\\text{-value} = 0.05$. The x-axis is drawn on logarithmic scale. For multiplier smaller than $1$, the adjusted $p$-value will incrase then decrase as $p$-value increases. The trend is the opposite for multiplier greater than $1$, but the difference will eventually reach $0$."}
upper_lower_p %>%
  left_join(vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(p_value = first(p_value))) %>%
  mutate(diff_p_value = adjusted_p_value - p_value) %>%
  ggplot() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0.05, col = "red", linetype = 2) +
  geom_point(aes(p_value, diff_p_value), alpha = 0.2) +
  scale_x_log10() +
  facet_wrap(~modified) +
  ylab("Difference between adjusted p-value and p-value") +
  xlab("p-value") +
  theme_light()
```

```{r}
upper_lower_p %>%
  left_join(vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(p_value = first(p_value))) %>%
  group_by(modified) %>%
  count(adjusted_reject = adjusted_p_value <= 0.05, reject = p_value <= 0.05) %>%
  complete(adjusted_reject, reject, fill = list(n = 0)) %>%
  ungroup() %>%
  filter(adjusted_reject != reject) %>%
  mutate(case = ifelse(!adjusted_reject, 'Proportion of lineups transforms to "not reject"', 'Proportion of lineups transforms to "reject"')) %>%
  rename(multiplier = modified) %>%
  mutate(n = paste0(round(n/279 * 100, 2), "\\%")) %>%
  select(-adjusted_reject, -reject) %>%
  pivot_wider(names_from = case, values_from = n) %>%
  kableExtra::kable(format = "latex", booktabs = TRUE, label = "sensitivity", 
                    escape = FALSE,
                    caption = "Change of rejection decision because of the modification of $\\hat{\\alpha}$.")
```

# Effect of number of evaluations on the power of a visual test

When comparing power of visual tests across different fitted value distributions, we have discussed the number of evaluations on a lineup will affect the power of the visual test. Using the lineups with uniform fitted value distribution, we show in Figure \ref{fig:n_eval} the change of power of visual tests due to different number of evaluations. It can be learned that as the number of evaluations increases, the power will increase but the margin will decrease. Considering we have eleven evaluations on lineups with uniform fitted value distribution, and five evaluations on other lineups, it is necessary to use the same number of evaluations for each lineup in comparison.

```{r}
min_poly_es <- vi_survey %>% 
  filter(!null_lineup, 
         !attention_check,
         type == "polynomial", 
         x_dist == "uniform") %>%
  pull(effect_size) %>%
  min()
max_poly_es <- vi_survey %>% 
  filter(!null_lineup, 
         !attention_check,
         type == "polynomial", 
         x_dist == "uniform") %>%
  pull(effect_size) %>%
  max()
```

```{r n_eval, fig.cap = "Change of power of visual tests for different number of evalutions on lineups with uniform fitted value distribution. The power will incrase as the number of evaluations increases, but the margin will decrase."}
eval_n <- function(n_eval, boot) {
  map_df(1:boot, function(boot_id) {
  vi_survey %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  filter(x_dist == "uniform") %>%
  filter(type == "polynomial") %>%
  group_by(unique_lineup_id) %>%
  slice_sample(n = n_eval) %>%
  ungroup() %>%
  calc_p_value_multi(lineup_id = unique_lineup_id,
                     detect = detect,
                     n_sel = num_selection,
                     alpha = alpha) %>%
  mutate(boot_id = boot_id)
}) %>%
  left_join(select(vi_survey, unique_lineup_id, effect_size)) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  select(-unique_lineup_id, -p_value) %>%
  mutate(n_eval = n_eval)
}

map_df(seq(1, 11, 2), ~eval_n(.x, 1)) %>%
  select(-boot_id) %>%
  nest(c(effect_size, reject, offset0)) %>%
  mutate(mod = map(data, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map(mod, function(mod) {
    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
                         offset0 = log(0.05/0.95))
    result$power <- predict(mod, type = "response", newdata = result)
    result
  })) %>%
  select(-data, -mod) %>%
  unnest(power) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_line(aes(log_effect_size, power, col = factor(n_eval)),
            size = 1) +
  labs(col = "Number of evalutaions") +
  xlab(expression(log[e] (Effect_size))) +
  theme_light() +
  scale_color_brewer(palette = "Reds")
```

# Power of a RESET test under different auxiliary regression formulas

It is found in the result that the power of a RESET test will be affected by the highest order of fitted values included in the auxiliary formula. And we suspect that the current recommendation of the highest order - four, is insufficient to test complex non-linear structures such as the "Triple-U" shape designed in this paper. Figure \ref{fig:reset} illustrates the change of power of RESET test while testing the "U" shape and the "Triple-U" shape with different highest orders. Clearly, when testing a simple shape like the "U" shape, the highest order has very little impact on the power. But for testing the "Triple-U" shape, there will be a loss of power if the recommended order is used. To avoid the loss of power, the highest order needs to be set to at least six.

```{r cache = TRUE}
reset_diff_dat <- poly_conv_sim %>%
  filter(x_dist == "uniform") %>%
  filter(shape == 4) %>%
  select(-F_p_value, -BP_p_value, -SW_p_value) %>%
  pivot_longer(RESET3_p_value:RESET10_p_value) %>%
  mutate(name = gsub("RESET(.*)_p_value", "\\1", name)) %>%
  mutate(reject = value <= 0.05) %>%
  select(effect_size, name, reject) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  nest(dat = c(effect_size, offset0, reject)) %>%
  mutate(mod = map(dat, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map(mod, function(mod) {
    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
                         offset0 = log(0.05/0.95))
    result$power <- predict(mod, type = "response", newdata = result)
    result
  })) %>%
  select(-dat, -mod) %>%
  unnest(power) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(shape = 4)
```

```{r cache = TRUE}
reset_diff_dat2 <- poly_conv_sim %>%
  filter(x_dist == "uniform") %>%
  filter(shape == 1) %>%
  select(-F_p_value, -BP_p_value, -SW_p_value) %>%
  pivot_longer(RESET3_p_value:RESET10_p_value) %>%
  mutate(name = gsub("RESET(.*)_p_value", "\\1", name)) %>%
  mutate(reject = value <= 0.05) %>%
  select(effect_size, name, reject) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  nest(dat = c(effect_size, offset0, reject)) %>%
  mutate(mod = map(dat, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map(mod, function(mod) {
    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
                         offset0 = log(0.05/0.95))
    result$power <- predict(mod, type = "response", newdata = result)
    result
  })) %>%
  select(-dat, -mod) %>%
  unnest(power) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(shape = 1)
```

```{r reset, fig.cap = 'Change of power of RESET tests for different orders of fitted values included in the auxiliary formula. The left panel is the power of testing the "U" shape and the right panel is the power of testing the "Triple-U" shape. The power will not be greatly affected by the highest order in the case of testing the "U" shape. In the case of testing the "Triple-U" shape, the highest order needs to be set to at least six to avoid the loss of power.', fig.height = 4}
reset_diff_dat %>%
  bind_rows(reset_diff_dat2) %>%
  mutate(shape = factor(ifelse(shape == 1, "U", "Triple-U"), levels = c("U", "Triple-U"))) %>%
  ggplot() +
  geom_line(aes(log_effect_size, power, col = as.factor(as.integer(name))),
            size = 1) +
  theme_light() +
  labs(col = "Highest order") +
  xlab(expression(log[e] (Effect_size))) +
  scale_color_brewer(palette = "Reds") +
  facet_wrap(~shape)
```

\clearpage

### 2023-04-03 paper.tex
copy in the appendix

@@
 
+\appendix
+
+\hypertarget{experiment-setup}{%
+\section{Experiment setup}\label{experiment-setup}}
+
+\hypertarget{mapping-of-subjects-to-experimental-factors}{%
+\subsection{Mapping of subjects to experimental
+factors}\label{mapping-of-subjects-to-experimental-factors}}
+
+Mapping of subjects to experimental factors is an important part of
+experiment design. Essentially, we want to maximum the difference in
+factors exposed to a subject. For this purpose, we design an algorithm
+to conduct subject allocation. Let \(L\) be a set of available lineups
+and \(S\) be a set of available subjects. According to the experimental
+design, the availability of a lineup is associated with the number of
+subjects it can assign to. For lineups with uniform fitted value
+distribution, this value is 11. And other lineups can be allocated to at
+most five different subjects. The availability of a subject is
+associated with the number of lineups that being allocated to this
+subject. A subject can view at most 18 different lineups.
+
+The algorithm starts from picking a random subject \(s \in S\) with the
+minimum number of allocated lineups. It then tries to find a lineup
+\(l \in L\) that can maximise the distance metric \(D\) and allocate it
+to subject \(s\). Set \(L\) and \(S\) will be updated and the picking
+process will be repeated until there is no available lineups or
+subjects.
+
+Let \(F_1,...,F_q\) be \(q\) experimental factors, and \(f_1, ...,f_q\)
+be the corresponding factor values. We say \(f_i\) exists in \(L_{s}\)
+if any lineup in \(L_{s}\) has this factor value. Similarly, \(f_if_j\)
+exists in \(L_{s}\) if any lineup in \(L_{s}\) has this pair of factor
+values. And \(f_if_jf_k\) exists in \(L_{s}\) if any lineup in \(L_{s}\)
+has this trio of factor values. The distance metric \(D\) is defined
+between a lineup \(l\) and a set of lineups \(L_{s}\) allocated to a
+subject \(s\) if \(L_{s}\) is non-empty:
+
+\footnotesize
+
+\begin{equation*}
+D =
+C - \sum_{1\leq i \leq q}I(f_i\text{ exists in }L_{s}) - \sum_{\substack{1\leq i \leq q-1 \\ i < j \leq q}}I(f_if_j\text{ exists in }L_{s}) - \sum_{\substack{1\leq i \leq q - 2 \\ i < j \leq q - 1 \\ j < k \leq q}}I(f_if_jf_k\text{ exists in }L_{s})
+\end{equation*}
+
+\normalsize
+
+where \(C\) is a sufficiently large constant such that \(D > 0\). If
+\(L_{s}\) is empty, we define \(D = 0\).
+
+The distance measures how different a lineup is from the set of lineups
+allocated to the subject in terms of factor values. Thus, the algorithm
+will try to allocate the most different lineup to a subject at each
+step.
+
+\hypertarget{data-collection-process}{%
+\subsection{Data collection process}\label{data-collection-process}}
+
+The survey data is collected via a self-hosted website designed by us.
+The complete architecture is provided in Figure \ref{fig:tech}. The
+website is built with the \texttt{Flask} \citep{flask} web framework and
+hosted on \texttt{PythonAnywhere} \citep{pythonanywhere}. It is
+configured to handle HTTP requests such that subjects can correctly
+receive webpages and submit responses. Embedded in the resources sent to
+subjects, the \texttt{jsPsych} front-end framework \citep{jspsych}
+instructs subjects' browsers to render an environment for running
+behavioral experiments. During the experiment, this framework will
+automatically collect common behavioral data such as response time and
+clicks on buttons. Subjects' responses are first validated by a
+scheduled \texttt{Python} script run on the server, then push to a
+Github repository. Lineup images shown to users are saved in multiple
+Github repositories and hosted in corresponding Github pages. The URLs
+to these images are resolved by \texttt{Flask} and bundled in HTML
+files.
+
+Once the participant is recruited from Prolific
+\citep{palan2018prolific}, it will be redirected to the entry page of
+our study website. An image of the entry page is provided in Figure
+\ref{fig:entry-page}. Then, the participant needs to submit the online
+consent form and fill in the demographic information as shown in
+\ref{fig:consent-form} and \ref{fig:metadata} respectively. Before
+evaluating lineups, participant also need to read the training page as
+provide in Figure \ref{fig:training-page} to understand the process. An
+example of the lineup page is given in Figure \ref{fig:lineup-page}. A
+half of the page is taken by the lineup image to attract participant's
+attention. The button to skip the selections for the current lineup is
+intentionally put in the corner of the bounding box with smaller font
+size, such that participants will not misuse this functionality.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures/experiment_tech}
+
+}
+
+\caption{Diagram of online experiment setup. The server-side of the study website uses Flask as backend hosted on PythonAnywhere. And the client-side uses jsPsych to run experiment.}\label{fig:tech}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures/message}
+
+}
+
+\caption{The entry page of the study website.}\label{fig:entry-page}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures/consent_form}
+
+}
+
+\caption{The consent form provided in the study website.}\label{fig:consent-form}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures/metadata}
+
+}
+
+\caption{The form to provide demographic information.}\label{fig:metadata}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures/training}
+
+}
+
+\caption{The training page of the study website.}\label{fig:training-page}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures/lineup1}
+
+}
+
+\caption{The lineup page of the study website.}\label{fig:lineup-page}
+\end{figure}
+
+\hypertarget{demographics}{%
+\section{Demographics}\label{demographics}}
+
+Along with the responses to lineups, we have collected a series of
+demographic information including age, pronoun, education background and
+previous experience in studies involved data visualization. Table
+\ref{tab:pronoun}, \ref{tab:age-group}, \ref{tab:education} and
+\ref{tab:experience} provide summary of the demographic data.
+
+It can be observed from the tables that most participants have Diploma
+or Bachelor degrees, followed by High school or below and the survey
+data is gender balanced. Majority of participants are between 18 to 39
+years old and there are slightly more participants who do not have
+previous experience than those who have.
+
+\begin{table}
+
+\caption{\label{tab:pronoun}Summary of pronoun distribution of subjects recuritted in this study.}
+\centering
+\begin{tabular}[t]{lrrrrrrrr}
+\toprule
+Pronoun & Period I & \% & Period II & \% & Period III & \% & Total & \%\\
+\midrule
+He & 77 & 17.4 & 79 & 17.8 & 61 & 13.8 & 217 & 49.0\\
+She & 78 & 17.6 & 77 & 17.4 & 61 & 13.8 & 216 & 48.8\\
+Other & 5 & 1.1 & 4 & 0.9 & 1 & 0.2 & 10 & 2.3\\
+ & 160 & 36.1 & 160 & 36.1 & 123 & 27.8 & 443 & 100.0\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:age-group}Summary of age distribution of subjects recuritted in this study.}
+\centering
+\begin{tabular}[t]{lrrrr}
+\toprule
+Age group & Period I & Period II & Period III & Total\\
+\midrule
+18-24 & 83 & 86 & 51 & 220\\
+25-39 & 69 & 63 & 63 & 195\\
+40-54 & 6 & 8 & 6 & 20\\
+55-64 & 2 & 3 & 3 & 8\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:education}Summary of education distribution of subjects recuritted in this study.}
+\centering
+\begin{tabular}[t]{lllll}
+\toprule
+Education & Period I & Period II & Period III & Total\\
+\midrule
+High School or below & 41 (9.3\%) & 53 (12\%) & 33 (7.4\%) & 127 (28.7\%)\\
+Diploma and Bachelor Degree & 92 (20.8\%) & 79 (17.8\%) & 66 (14.9\%) & 237 (53.5\%)\\
+Honours Degree & 6 (1.4\%) & 15 (3.4\%) & 6 (1.4\%) & 27 (6.1\%)\\
+Masters Degree & 21 (4.7\%) & 13 (2.9\%) & 16 (3.6\%) & 50 (11.3\%)\\
+Doctoral Degree & 0 (0\%) & 0 (0\%) & 2 (0.5\%) & 2 (0.5\%)\\
+\addlinespace
+ & 160 (36.1\%) & 160 (36.1\%) & 123 (27.8\%) & 443 (100\%)\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:experience}Summary of previous experience distribution of subjects recuritted in this study.}
+\centering
+\begin{tabular}[t]{lrrrr}
+\toprule
+Previous experience & Period I & Period II & Period III & Total\\
+\midrule
+No & 96 & 88 & 67 & 251\\
+Yes & 64 & 72 & 56 & 192\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\hypertarget{effect-size-derivation}{%
+\section{Effect size derivation}\label{effect-size-derivation}}
+
+Effect size can be defined as the difference of a parameter for a
+particular model or distribution, or a statistic derived from a sample.
+Importantly, it needs to reflect the treatment we try to measure.
+Centred on a conventional statistical test, we usually can deduce the
+effect size from the test statistic by substituting the null parameter
+value. When considering the diagnostics of residual departures, there
+exist many possibilities of test statistics for a variety of model
+assumptions. Meanwhile, diagnostic plots such as the residual plot have
+no general agreement on measuring how strong a model violation pattern
+is. To build a bridge between various residual-based tests, and the
+visual test, we focus on the shared information embedded in the testing
+procedures, which is the distribution of residuals. When comes to
+comparison of distribution, Kullback-Leibler divergence
+\citep{kullback1951information} is a classical way to represent the
+information loss or entropy increase caused by the approximation to the
+true distribution, which in our case, the inefficiency due to the use of
+false model assumptions.
+
+Following the terminology introduced by \citet{kullback1951information},
+\(P\) represents the measured probability distribution, and \(Q\)
+represents the assumed probability distribution. The Kullback-Leibler
+divergence is defined as
+\(\int_{-\infty}^{\infty}log(p(x)/q(x))p(x)dx\), where \(p(.)\) and
+\(q(.)\) denote probability densities of \(P\) and \(Q\).
+
+Let \(\boldsymbol{X}_a = (\boldsymbol{1}, \boldsymbol{X})\) denotes the
+\(p\) regressors with \(n\) observations,
+\(\boldsymbol{R}_a = \boldsymbol{I} -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\)
+denotes the residual operator, and let
+\(\boldsymbol{\varepsilon} \sim N(\boldsymbol{0},\sigma^2\boldsymbol{I})\)
+denotes the error. Using the Frisch--Waugh--Lovell theorem, residuals
+\(\boldsymbol{e} = \boldsymbol{R}_a\boldsymbol{\varepsilon}\). Because
+\(rank(\boldsymbol{R}_a) = n - p < n\), \(e\) follows a degenerate
+multivariate normal distribution and does not have a density. Since the
+Kullback-Leibler divergence requires a proper density function, we need
+to simplify the covariance matrix of \(\boldsymbol{e}\) by setting all
+the off-diagonal elements to 0. Then, the residuals will assumed to
+follow \(N(\boldsymbol{0}, diag(\boldsymbol{R}_a\sigma^2))\) under the
+null hypothesis that the model is correctly specified. If the model is
+however misspecified due to omitted variables \(\boldsymbol{Z}\), or a
+non-constant variance \(\boldsymbol{V}\), the distribution of residuals
+can be derived as
+\(N(\boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}_a\sigma^2))\)
+and
+\(N(\boldsymbol{0}, diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a'))\)
+respectively.
+
+By assuming both \(P\) and \(Q\) are multivariate normal density
+functions, the Kullback-Leibler divergence can be rewritten as
+\[KL = \frac{1}{2}\left(log\frac{|\Sigma_p|}{|\Sigma_q|} - n + tr(\Sigma_p^{-1}\Sigma_q) + (\mu_p - \mu_q)'\Sigma_p^{-1}(\mu_p - \mu_q)\right).\]
+
+Then, we can combine the two residual departures into one formula
+
+\small
+
+\begin{equation}
+\label{eq:effect-size}
+KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}\boldsymbol{\mu}_z\right).
+\end{equation}
+
+\normalsize
+
+When there are omitted variables but constant error variance, the
+formula can be reduced to
+
+\[KL = \frac{1}{2}\left(\boldsymbol{\mu}_z^{T}(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right).\]
+
+And when the model equation is correctly specified but the error
+variance is non-constant, the formula can be reduced to
+
+\[KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2))\right).\]
+Since we assume \(\sigma = 1\) for the heteroskedasticity model, the
+final form of the formula is
+
+\[KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right).\]
+
+\hypertarget{effect-of-data-collection-period}{%
+\section{Effect of data collection
+period}\label{effect-of-data-collection-period}}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/poly-boxplot-lineup-1}
+
+}
+
+\caption{A lineup of "letter-value" boxplots of weighted propotion of detect for lineups over different data collection periods for non-linearity model.
+Can you find the most different boxplot?
+The data plot is positioned in panel $2^3 - 1$.}\label{fig:poly-boxplot-lineup}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/heter-boxplot-lineup-1}
+
+}
+
+\caption{A lineup of "letter-value" boxplots of weighted propotion of detect for lineups over different data collection periods for heteroskedasticity model.
+Can you find the most different boxplot?
+The data plot is positioned in panel $2^4 - 2$.}\label{fig:heter-boxplot-lineup}
+\end{figure}
+
+We have the same type of model collected over different data collection
+periods, that may lead to unexpected batch effect. Figure
+\ref{fig:poly-boxplot-lineup} and \ref{fig:heter-boxplot-lineup} provide
+two lineups to examine whether there is an actual difference across data
+collection periods for non-linearity model and heteroskedasticity model
+respectively. To emphasize the tail behaviour and display fewer
+outliers, we use the ``letter-value'' boxplot \citep{hofmann2017value}
+which is an extension of the number of ``letter value'' statistics to
+check the weighed proportion of detect over different data collection
+period. The weighted proportion of detect is calculated by taking the
+average of \(c_i\) of a lineup over a data collection period. Within our
+research team, we can not identify the data plot from the null plots for
+these two lineups, result in \(p\)-values much greater than \(5\)\%.
+Thus, there is no clear evidence of batch effect.
+
+\hypertarget{sensitivity-analysis-for-alpha}{%
+\section{\texorpdfstring{Sensitivity analysis for
+\(\alpha\)}{Sensitivity analysis for \textbackslash alpha}}\label{sensitivity-analysis-for-alpha}}
+
+The parameter \(\alpha\) used for the \(p\)-value calculation needs to
+be estimated from responses to null lineups. With a greater value of
+\(\hat{\alpha}\), the \(p\)-value will be smaller, resulting in more
+lineups being rejected. However, The way we generate Rorschach lineup is
+not strictly the same as what suggested in
+\citet{vanderplas2021statistical} and \citet{buja_statistical_2009}.
+Therefore, we conduct a sensitivity analysis in this section to examine
+the impact of the variation of the estimator \(\alpha\) on our primary
+findings.
+
+The analysis is conducted by setting up several scenarios, where the
+\(\alpha\) is under or overestimated by 12.5\%, 25\% and 50\%. Using the
+adjusted \(\hat{\alpha}\), we recalculate the \(p\)-value for every
+lineup and show the results in Figure \ref{fig:sensitivity}. It can be
+observed that there are some changes to \(p\)-values, especially when
+the \(\hat{\alpha}\) is multiplied by 50\%. However, Table
+\ref{tab:sensitivity} shows that adjusting \(\hat{\alpha}\) will not
+result in a huge difference in rejection decisions. There are only a
+small percentage of cases where the rejection decision change. It is
+very unlikely the downstream findings will be affected because of the
+estimate of \(\alpha\).
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/sensitivity-1}
+
+}
+
+\caption{Change of $p$-values with $\hat{\alpha}$ multiplied by $0.5$, $0.75$, $0.875$, $1.125$, $1.25$ and $1.5$.
+The vertical dashed line is to indicate $p\text{-value} = 0.05$.
+The x-axis is drawn on logarithmic scale.
+For multiplier smaller than $1$, the adjusted $p$-value will incrase then decrase as $p$-value increases.
+The trend is the opposite for multiplier greater than $1$, but the difference will eventually reach $0$.}\label{fig:sensitivity}
+\end{figure}
+
+\begin{table}
+
+\caption{\label{tab:sensitivity}Change of rejection decision because of the modification of $\hat{\alpha}$.}
+\centering
+\begin{tabular}[t]{rll}
+\toprule
+multiplier & Proportion of lineups transforms to "not reject" & Proportion of lineups transforms to "reject"\\
+\midrule
+0.500 & 2.51\% & 0\%\\
+0.750 & 1.43\% & 0\%\\
+0.875 & 1.08\% & 0\%\\
+1.125 & 0\% & 1.08\%\\
+1.250 & 0\% & 1.43\%\\
+\addlinespace
+1.500 & 0\% & 1.79\%\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\hypertarget{effect-of-number-of-evaluations-on-the-power-of-a-visual-test}{%
+\section{Effect of number of evaluations on the power of a visual
+test}\label{effect-of-number-of-evaluations-on-the-power-of-a-visual-test}}
+
+When comparing power of visual tests across different fitted value
+distributions, we have discussed the number of evaluations on a lineup
+will affect the power of the visual test. Using the lineups with uniform
+fitted value distribution, we show in Figure \ref{fig:n_eval} the change
+of power of visual tests due to different number of evaluations. It can
+be learned that as the number of evaluations increases, the power will
+increase but the margin will decrease. Considering we have eleven
+evaluations on lineups with uniform fitted value distribution, and five
+evaluations on other lineups, it is necessary to use the same number of
+evaluations for each lineup in comparison.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/n_eval-1}
+
+}
+
+\caption{Change of power of visual tests for different number of evalutions on lineups with uniform fitted value distribution.
+The power will incrase as the number of evaluations increases, but the margin will decrase.}\label{fig:n_eval}
+\end{figure}
+
+\hypertarget{power-of-a-reset-test-under-different-auxiliary-regression-formulas}{%
+\section{Power of a RESET test under different auxiliary regression
+formulas}\label{power-of-a-reset-test-under-different-auxiliary-regression-formulas}}
+
+It is found in the result that the power of a RESET test will be
+affected by the highest order of fitted values included in the auxiliary
+formula. And we suspect that the current recommendation of the highest
+order - four, is insufficient to test complex non-linear structures such
+as the ``Triple-U'' shape designed in this paper. Figure \ref{fig:reset}
+illustrates the change of power of RESET test while testing the ``U''
+shape and the ``Triple-U'' shape with different highest orders. Clearly,
+when testing a simple shape like the ``U'' shape, the highest order has
+very little impact on the power. But for testing the ``Triple-U'' shape,
+there will be a loss of power if the recommended order is used. To avoid
+the loss of power, the highest order needs to be set to at least six.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/reset-1}
+
+}
+
+\caption{Change of power of RESET tests for different orders of fitted values included in the auxiliary formula.
+The left panel is the power of testing the "U" shape and the right panel is the power of testing the "Triple-U" shape.
+The power will not be greatly affected by the highest order in the case of testing the "U" shape.
+In the case of testing the "Triple-U" shape, the highest order needs to be set to at least six to avoid the loss of power.}\label{fig:reset}
+\end{figure}
+
+\clearpage
+
 \bibliographystyle{tfcad}

### 2023-04-03 paper.Rmd
add bounding box for legend

@@
   ylab("Weighted propotion of detect") +
-  xlab("Data collection period")
+  xlab("Data collection period") +
+  theme(legend.key = element_rect(colour = "black"))
 ```

### 2023-04-03 paper.Rmd
Draw p-value notation correctly

@@
   facet_wrap(~modified) +
-  ylab("Difference between adjusted p-value and p-value") +
-  xlab("p-value") +
+  ylab(expression("Adjusted"~italic(p)*"-value minus"~italic(p)*"-value")) +
+  xlab(expression(italic(p)*"-value")) +
   theme_light()
+
 ```

### 2023-04-03 appendix.Rmd
rearranged main paper and appendix into separate files again

--- 2023-04-03 paper.Rmd
@@
 title: |
-  A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol
-type: ARTICLE TEMPLATE
+  Appendix: A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol
 author:
@@
 bibliography: paper.bib
-abstract: |
-  Abstract to fill.
-keywords: |
-  statistical graphics; data visualization; visual inference; hypothesis testing; reression analysis; cognitive perception; simulation; practical significance; effect size
 header-includes: |
@@
 
-# Introduction
-
-> *"Since all models are wrong the scientist must be alert to what is importantly wrong."* [@box1976science]
-
-Diagnostics are the key to determining whether there is anything importantly wrong with a model.
-In linear regression analysis, residuals from the model fit are commonly used.
-Residuals summarise what is not captured by the model, and thus provide the capacity to identify what might be wrong.
-
-We can assess residuals in multiple ways.
-Residuals may be plotted as a histogram or quantile-quantile plot to examine the distribution.
-Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, we consider it to be well behaved.
-However, if the distribution is skewed, bimodal, multimodal, or contains outliers, there is a cause for concern.
-One could also inspect the distribution by conducting a goodness of fit test, such as the Shapiro-Wilk Normality test [@shapiro1965analysis].
-
-More typically, residuals will be plotted, as a scatter plot against the predicted values and each of the explanatory variables to scrutinize their relationships.
-If there are any visually discoverable patterns, the model is potentially misspecified.
-In general, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity.
-However, correctly judging a residual plot where no pattern exists can be a painstakingly difficult task for humans.
-It is especially common, particularly among students, to misinterpret patterns that are random noise and random deviation from a model [@loy2021bringing].
-It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.
-
-Abundance of literature describe appropriate diagnostic methods for linear regression, e.g.
-@draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals.
-All these writings consider plotting residuals to be a standard technique that should be examined routinely in all regression modelling problems.
-In addition, @draper1998applied and @belsley_regression_1980 believe that residual plots are usually revealing when the assumptions are violated.
-@cook_applied_1999 thinks formal tests and graphical procedures are complementary and both have a place in residual analysis, but they focus on graphical methods rather than on formal testing, as they are easier to use.
-@montgomery1982introduction even suggests that residual plots are more informative in most practical situations than the corresponding formal tests, and statistical tests on regression model residuals are not widely used based on their experience.
-
-A common guidance by experts is that optimal method for diagnosing model fits is by plotting the data.
-The persistence of this advice to check the plots is curious, and investigating why this might be common advice is the subject of this paper.
-The paper is structured as follows.
-The next section describes the background on the types of departures that one expects to detect, and outlines a formal statistical process for reading residual plots, called visual inference.
-Section \ref{experimental-design} details the experimental design to compare the decisions made by formal hypothesis testing, and how humans would read diagnostic plots.
-The results are reported in Section \ref{results}.
-We conclude with a discussion of future work, in particular, how the responsibility for residual plot reading might be relieved.
-
-# Background
-
-## Departures from good residual plots
-
-```{r residual-plot-common-departures, fig.width = 10, fig.height = 2.5, fig.cap = "Example residual vs fitted value plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
-The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted value plot."}
-
-set.seed(10086)
-
-# Base model
-mod <- poly_model(include_z = FALSE, sigma = 0.25)
-
-# Base data
-dat <- mod$gen(300)
-
-bind_rows(
-  dat %>%
-    mutate(type = "(A) Good residuals"),
-  mod$set_prm("include_z", TRUE)$
-    gen(300, computed = select(dat, x, e)) %>%
-    mutate(type = "(B) Non-linearity"),
-  heter_model(b = 64)$
-    gen(300, computed = select(dat, x, e)) %>%
-    mutate(type = "(C) Heteroskedasticity"),
-  heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
-    gen(300, computed = select(dat, x)) %>%
-    mutate(type = "(D) Non-normality")
-) %>%
-  mod$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
-  facet_wrap(~type, ncol = 4, scales = "free") +
-  xlab("Fitted values") +
-  ylab("Residuals")
-```
-
-Graphical summaries in which residuals are plotted against fitted values or other functions of the predictors that are approximately orthogonal to the residuals are referred to as standard residual plots in @cook1982residuals.
-Figure \ref{fig:residual-plot-common-departures}A shows an ideal residual plot where the residuals are evenly distributed at both sides of the horizontal zero line, with no noticeable patterns.
-
-There are various types of departures from an ideal residual plot. Non-linearity, heteroskedasticity and non-normality are perhaps the three mostly checked departures.
-
-Non-linearity is a type of model misspecification caused by failing to include higher order terms of the predictors in the regression equation.
-Any non-linear functional form of residuals on fitted values in the residual plot could be indicative of non-linearity.
-An example residual plot containing visual pattern of non-linearity is shown in Figure \ref{fig:residual-plot-common-departures}B.
-One can clearly observe the "S-shape" from the residual plot as the cubic term is not captured by the misspecified model.
-
-Heteroskedasticity refers to the presence of nonconstant error variance in a regression model.
-It is mostly due to the strict but false assumptions on the variance-covariance matrix of the error term.
-The usual pattern of heteroskedasticity on a residual plot is the inconsistent spread of the residuals across the horizontal axis.
-Visually, it sometimes results in the so-called "butterfly" shape as shown in Figure \ref{fig:residual-plot-common-departures}C, or the "left-triangle" and "right-triangle" shape where the smallest variance occurs at one side of the horizontal axis.
-
-Compared to non-linearity and heteroskedasticity, non-normality is usually harder to detect from a residual plot since a scatter plot do not readily reveal the marginal distribution.
-A favourable graphical summary for this task is the quantile-quantile plot.
-As we mainly discuss residual plots, non-normality will not be the focus of this paper.
-For a consistent comparison, the residual plot of this departure is still presented in Figure \ref{fig:residual-plot-common-departures}D.
-When the number residuals below and above the horizontal axis are uneven across the local regions along the $x$-axis, we expect that the normality assumption is violated.
-For example, given a skewed error distribution, there will be fewer data points and more outliers on one side of the horizontal axis as shown in Figure \ref{fig:residual-plot-common-departures}D.
-
-## Conventionally testing for departures
-
-Other than checking diagnostic plots, analysts may perform formal hypothesis testing to detect model defects.
-A variety of tests can be applied; each testing for a specific violation of the null hypothesis.
-For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution.
-To test specific forms of non-linearity, one may apply the F-test as a model structural test to examine the significance of specific polynomial and non-linear forms of the predictors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].
-The Shapiro-Wilk test [@shapiro1965analysis] is the most widely used test of non-normality included by many of the statistical software programs.
-The Jarque--Bera test [@jarque1980efficient] is also used to directly check whether the sample skewness and kurtosis match a normal distribution.
-
-We apply the RESET test, Breusch-Pagan test and Shapiro-Wilk test to the residual plots shown in Figure \ref{fig:residual-plot-common-departures}; the results of this is shown in Table \ref{tab:example-residual-plot-table}.
-The Breusch-Pagan test and the Shapiro-Wilk test both reject $H_0$ for residual plots that display non-linearity and heteroscedasticity, even though these are not the original intention of the test.
-As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are also sensitive to other types of departures.
-It is likely $H_0$ is correctly rejected but for the wrong reason, a phenomenon known as the "Type III error".
-Additionally, outliers will often incorrectly trigger the rejection of $H_0$ despite when majority of the residuals are well-behaved [@cook_applied_1999].
-Furthermore, with a sufficiently large sample size, residual-based tests may reject $H_0$ due to a slight departure that is of little practical significance.
-These can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers and slight departures.
+\appendix
+
+# Additional details of testing procedures
+## Effect size derivation
+
+Effect size can be defined as the difference of a parameter for a particular model or distribution,  or a statistic derived from a sample.
+Importantly, it needs to reflect the treatment we try to measure.
+Centred on a conventional statistical test, we usually can deduce the effect size from the test statistic by substituting the null parameter value.
+When considering the diagnostics of residual departures, there exist many possibilities of test statistics for a variety of model assumptions.
+Meanwhile, diagnostic plots such as the residual plot have no general agreement on measuring how strong a model violation pattern is.
+To build a bridge between various residual-based tests, and the visual test, we focus on the shared information embedded in the testing procedures, which is the distribution of residuals.
+When comes to comparison of distribution, Kullback-Leibler divergence [@kullback1951information] is a classical way to represent the information loss or entropy increase caused by the approximation to the true distribution, which in our case, the inefficiency due to the use of false model assumptions.
+
+Following the terminology introduced by @kullback1951information, $P$ represents the measured probability distribution, and $Q$ represents the assumed probability distribution.
+The Kullback-Leibler divergence is defined as $\int_{-\infty}^{\infty}log(p(x)/q(x))p(x)dx$, where $p(.)$ and $q(.)$ denote probability densities of $P$ and $Q$.
+
+Let $\boldsymbol{X}_a = (\boldsymbol{1}, \boldsymbol{X})$ denotes the $p$ regressors with $n$ observations, $\boldsymbol{R}_a = \boldsymbol{I} -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ denotes the residual operator, and let $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0},\sigma^2\boldsymbol{I})$ denotes the error.
+Using the Frisch–Waugh–Lovell theorem, residuals $\boldsymbol{e} = \boldsymbol{R}_a\boldsymbol{\varepsilon}$.
+Because $rank(\boldsymbol{R}_a) = n - p < n$, $e$ follows a degenerate multivariate normal distribution and does not have a density.
+Since the Kullback-Leibler divergence requires a proper density function, we need to simplify the covariance matrix of $\boldsymbol{e}$ by setting all the off-diagonal elements to 0.
+Then, the residuals will assumed to follow $N(\boldsymbol{0}, diag(\boldsymbol{R}_a\sigma^2))$ under the null hypothesis that the model is correctly specified.
+If the model is however misspecified due to omitted variables $\boldsymbol{Z}$, or a non-constant variance $\boldsymbol{V}$, the distribution of residuals can be derived as $N(\boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}_a\sigma^2))$ and $N(\boldsymbol{0}, diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a'))$ respectively.
+
+By assuming both $P$ and $Q$ are multivariate normal density functions, the Kullback-Leibler divergence can be rewritten as 
+$$KL = \frac{1}{2}\left(log\frac{|\Sigma_p|}{|\Sigma_q|} - n + tr(\Sigma_p^{-1}\Sigma_q) + (\mu_p - \mu_q)'\Sigma_p^{-1}(\mu_p - \mu_q)\right).$$
+
+Then, we can combine the two residual departures into one formula
+
+\small
+
+\begin{equation}
+\label{eq:effect-size}
+KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}\boldsymbol{\mu}_z\right).
+\end{equation}
+
+\normalsize
+
+When there are omitted variables but constant error variance, the formula can be reduced to
+
+$$KL = \frac{1}{2}\left(\boldsymbol{\mu}_z^{T}(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right).$$
+
+And when the model equation is correctly specified but the error variance is non-constant, the formula can be reduced to
+
+$$KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2))\right).$$
+Since we assume $\sigma = 1$ for the heteroskedasticity model, the final form of the formula is
+
+$$KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right).$$
+
+## Sensitivity analysis for $\alpha$
+
+The parameter $\alpha$ used for the $p$-value calculation needs to be estimated from responses to null lineups.
+With a greater value of $\hat{\alpha}$, the $p$-value will be smaller, resulting in more lineups being rejected.
+However, The way we generate Rorschach lineup is not strictly the same as what suggested in @vanderplas2021statistical and @buja_statistical_2009.
+Therefore, we conduct a sensitivity analysis in this section to examine the impact of the variation of the estimator $\alpha$ on our primary findings.
+
+The analysis is conducted by setting up several scenarios, where the $\alpha$ is under or overestimated by 12.5%, 25% and 50%.
+Using the adjusted $\hat{\alpha}$, we recalculate the $p$-value for every lineup and show the results in Figure \ref{fig:sensitivity}.
+It can be observed that there are some changes to $p$-values, especially when the $\hat{\alpha}$ is multiplied by 50%.
+However, Table \ref{tab:sensitivity} shows that adjusting $\hat{\alpha}$ will not result in a huge difference in rejection decisions.
+There are only a small percentage of cases where the rejection decision change.
+It is very unlikely the downstream findings will be affected because of the estimate of $\alpha$.
 
 ```{r}
-set.seed(10086)
-
-# Base model
-mod <- poly_model(include_z = FALSE, sigma = 0.25)
-
-# Base data
-dat0 <- mod$gen(300)
-
-# Replicate data in Figure 1
-dat1 <- mod$set_prm("include_z", TRUE)$gen(300, computed = select(dat0, x, e))
-dat2 <- heter_model(b = 64)$gen(300, computed = select(dat0, x, e))
-dat3 <- heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
-               gen(300, computed = select(dat0, x))
-
-dat_list <- list(dat0, dat1, dat2, dat3)
-
-table_dat <- data.frame(plot = c("A", "B", "C", "D"), 
-                        model = c("None", 
-                                  "Non-linearity", 
-                                  "Heteroskedasticity", 
-                                  "Non-normality"),
-                        r = map_dbl(dat_list, 
-                                    ~POLY_MODEL$test(.x, 
-                                                     test = "RESET", 
-                                                     power = 2:4)$p_value),
-                        b = map_dbl(dat_list, 
-                                    ~HETER_MODEL$test(.x)$p_value),
-
-                        # SW test has not been included by visage
-                        s = map_dbl(dat_list, 
-                                    ~shapiro.test(.x$.resid)$p.value)) %>%
-
-  # 3 decimal points
-  mutate(across(r:s, ~ format(round(.x, digits = 3), nsmall = 3))) %>%
-
-  # Italic if reject
-  mutate(across(r:s, ~ kableExtra::cell_spec(.x, italic = as.numeric(.x) <= 0.05)))
-
-  table_dat %>%
-  knitr::kable(col.names = c("Plot", 
-                             "Departures", 
-                             "RESET", 
-                             "Breusch-Pagan", 
-                             "Shapiro–Wilk"),
-               align = "llrrr",
+upper_p <- map_df(c(0.125, 0.25, 0.5), function(modified) {
+  calc_p_value_multi(vi_survey %>%
+                       filter(!attention_check) %>%
+                       filter(!null_lineup) %>%
+                       filter(x_dist == "uniform") %>%
+                       mutate(alpha_upper = alpha * (1 + modified)),
+                     lineup_id = unique_lineup_id,
+                     detect = detect,
+                     n_sel = num_selection,
+                     alpha = alpha_upper) %>%
+  mutate(modified = 1 + modified) %>%
+  mutate(adjusted_p_value = p_value)
+})
+
+lower_p <- map_df(c(0.125, 0.25, 0.5), function(modified) {
+  calc_p_value_multi(vi_survey %>%
+                       filter(!attention_check) %>%
+                       filter(!null_lineup) %>%
+                       filter(x_dist == "uniform") %>%
+                       mutate(alpha_lower = alpha * (1 - modified)),
+                     lineup_id = unique_lineup_id,
+                     detect = detect,
+                     n_sel = num_selection,
+                     alpha = alpha_lower) %>%
+  mutate(modified = 1 - modified) %>%
+  mutate(adjusted_p_value = p_value)
+})
+
+upper_lower_p <- upper_p %>%
+  bind_rows(lower_p) %>%
+  select(-p_value)
+```
+
+```{r sensitivity, fig.cap = "Change of $p$-values with $\\hat{\\alpha}$ multiplied by $0.5$, $0.75$, $0.875$, $1.125$, $1.25$ and $1.5$.
+The vertical dashed line is to indicate $p\\text{-value} = 0.05$.
+The x-axis is drawn on logarithmic scale.
+For multiplier smaller than $1$, the adjusted $p$-value will incrase then decrase as $p$-value increases.
+The trend is the opposite for multiplier greater than $1$, but the difference will eventually reach $0$."}
+upper_lower_p %>%
+  left_join(vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(x_dist == "uniform") %>%
+  group_by(unique_lineup_id) %>%
+  summarise(p_value = first(p_value))) %>%
+  mutate(diff_p_value = adjusted_p_value - p_value) %>%
+  ggplot() +
+  geom_hline(yintercept = 0) +
+  geom_vline(xintercept = 0.05, col = "red", linetype = 2) +
+  geom_point(aes(p_value, diff_p_value), alpha = 0.2) +
+  scale_x_log10() +
+  facet_wrap(~modified) +
+  ylab(expression("Adjusted"~italic(p)*"-value minus"~italic(p)*"-value")) +
+  xlab(expression(italic(p)*"-value")) +
+  theme_light()
+
+```
+
+```{r}
+upper_lower_p %>%
+  left_join(vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(x_dist == "uniform") %>%
+  group_by(unique_lineup_id) %>%
+  summarise(p_value = first(p_value))) %>%
+  group_by(modified) %>%
+  count(adjusted_reject = adjusted_p_value <= 0.05, reject = p_value <= 0.05) %>%
+  complete(adjusted_reject, reject, fill = list(n = 0)) %>%
+  ungroup() %>%
+  filter(adjusted_reject != reject) %>%
+  mutate(case = ifelse(!adjusted_reject, "reject to not reject", "not reject to reject")) %>%
+  rename(fraction = modified) %>%
+  mutate(n = paste0(round(n/279 * 100, 2))) %>%
+  select(-adjusted_reject, -reject) %>%
+  pivot_wider(names_from = case, values_from = n) %>%
+  kableExtra::kable(format = "latex", booktabs = TRUE, label = "sensitivity", 
                     escape = FALSE,
-               format = "latex",
-               booktabs = TRUE, 
-               caption  = 'Statistical significance testing for departures from good residuals for plots in Figure \\ref{fig:residual-plot-common-departures}.
-Shown are the $p$-values calculated for the RESET, the Breusch-Pagan and the Shapiro–Wilk tests.
-The good residual plot (A) is judged a good residual plot, as expected, by all tests.
-The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.',
-               label = "example-residual-plot-table" ) 
-```
-
-## Visual test procedure based on lineups
-
-One may argue that reading diagnostic plots is to some extent subjective and indecisive compared to those rigorous statistical procedures as it relies on graphical perception - human ability to interpret and decode the information embedded in graph [@cleveland_graphical_1984].
-Further, the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
-For instance, people
-over-interpret the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurrence [@roy_chowdhury_using_2015].
-
-Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots.
-Later, @buja_statistical_2009 proposed the lineup protocol as a visual test inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people.
-The protocol consists of $m$ randomly placed plots, where one plot is the data plot, and the remaining $m - 1$ null plots have the identical graphical procedure except the data has been replaced with data consistent with $H_0$, that the model is correctly specified.
-Then, an observer who have not seen the data plot will be asked to point out the most different plot from the lineup.
-Under $H_0$, it is expected that the data plot would have no distinguishable difference from the null plots, and the probability that the observer correctly picks the data plot is $1/m$.
-If one rejects $H_0$ as the observer correctly picks the data plot, then the Type I error of this test is $1/m$.
-
-```{r first-example-lineup, fig.height = 7, fig.width = 7, fig.cap = "Visual testing is conducted using a lineup, as in the example here.
-The residual plot computed from the observed data (plot $2^2 + 2$, exhibiting non-linearity) is embedded among 19 null plots, where the residuals are simulated from a standard error model.
-Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot.
-A small $p$-value would result from a substantial number selecting plot $2^2 + 2$."}
-
-vi_lineup$poly_300$data %>%
-  VI_MODEL$plot_lineup(theme = theme_light(), remove_axis = TRUE, remove_grid_line = TRUE)
-```
-
-Figure \ref{fig:first-example-lineup} is an example of a lineup protocol.
-If the data plot at position $2^2 + 2$ is identifiable, then it is evidence for the rejection of $H_0$.
-In fact, the actual residual plot is obtained from a misspecified regression model with missing non-linear terms.
-
-Data used in the $m - 1$ null plots needs to be simulated.
-In regression diagnostics, sampling data consistent with $H_0$ is equivalent to sampling data from the assumed model.
-As @buja_statistical_2009 suggested, $H_0$ is usually a composite hypothesis controlled by nuisance parameters.
-Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to a so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling.
-The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009].
-Essentially, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the predictors, then rescaling it by the ratio of residual sum of square in two regressions.
-
-The effectiveness of lineup protocol for regression analysis is validated by @majumder_validation_2013 under relatively simple settings with up to two predictors.
-Their results suggest that visual tests are capable of testing the significance of a single predictor with a similar power as a t-test, though they express that in general it is unnecessary to use visual inference if there exists a conventional test, and they do not expect the visual test to perform equally well as the conventional test.
-In their third experiment, where there is not a conventional test, visual test outperforms the conventional test for a large margin.
-This is encouraging, as it promotes the use of visual inference in situations where there are no existing statistical testing procedures.
-Visual inference have also been integrated into diagnostic of hierarchical linear models by @loy2013diagnostic, @loy2014hlmdiag and @loy2015you.
-They use lineup protocols to judge the assumption of linearity, normality and constant error variance for both the level-1 and level-2 residuals.
-
-# Calculation of statistical significance and test power
-
-## What is being tested?
-
-In diagnosing a model fit from residuals, we are generally interested in $H_0:$ *The regression model is correctly specified* against the broad alternative $H_a:$ *The regression model is misspecified*.
-
-However, it is practically impossible to test this specific $H_0$ with conventional tests, which are constructed to measure specific departures.
-For example, the RESET test is formulated as $H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0$ against $H_a: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0$, from $y = \tau_0 + \sum_{i=1}^{p}\tau_px_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)$.
-Similarly, the Breusch-Pagan test is designed to specifically test $H_0:$ *error variances are all equal* ($\zeta_i=0 \text{ for } i=1,..,p$) versus the alternative $H_a:$ *that the error variances are a multiplicative function of one or more variables* ($\text{at least one } \zeta_i\neq 0$) from $e^2 = \zeta_0 + \sum_{i=1}^{p}\zeta_i x_i + u, ~ u\sim N(0,\sigma_u^2)$.
-
-One of the potential benefits of the visual test, based on the lineup protocol, is that it should be able to detect a range of departures from good residuals.
-
-## Statistical significance
-
-In hypothesis testing, a $p$-value is defined as the probability of observing test results as least as extreme as the observed result given $H_0$ is true.
-Conventional hypothesis tests usually have an existing method to derive or compute $p$-value based on the null distribution.
-What we need to discuss in the following is the method to estimate $p$-value for a visual test.
-
-Within the context of visual inference, by involving $k$ independent observers, the visual $p$-value can be interpreted as the probability of having as many or more subjects detect the data plot than the observed result.
-
-Let $X_j = \{0,1\}$ be a Bernoulli random variable denoting whether subject $j$ correctly detecting the data plot, and $X = \sum_{j=1}^{K}X_j$ be the number of observers correctly picking the data plot.
-Then, by imposing a relatively strong assumption on the visual test that all $K$ evaluations are fully independent, under $H_0$, $X \sim \mathrm{Binom}_{K,1/m}$.
-Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as $P(X \geq x) = 1 - F(x) + f(x)$, where $F(.)$ is the binomial cumulative distribution function, $f(.)$ is the binomial probability mass function and $x$ is the realization of number of observers correctly picking the data plot [@majumder_validation_2013].
-
-As pointed out by @vanderplas2021statistical, this basic binomial model does not take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup.
-And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup.
-@vanderplas2021statistical summarises three common scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects.
-Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experiment design.
-For Scenario 3, @vanderplas2021statistical models the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$.
-The number of times plot $i$ being selected in $K$ evaluations is denoted as $c_i$.
-In case subject $j$ makes multiple selections, $1/s_j$ will be added to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$.
-This ensures $\sum_{i}c_i=K$.
-Since we are only interested in the selections of the data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual $p$-value is given as
-
-```{=tex}
-\begin{equation} \label{eq:pvalue-beta-binomial}
-P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+
-\end{equation}
-```
-
-\noindent where $B(.)$ is the beta function defined as
-
-```{=tex}
-\begin{equation} \label{eq:betafunction}
-B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0.
-\end{equation}
-```
-
-Note that Equation \ref{eq:pvalue-beta-binomial} given in @vanderplas2021statistical only works with non-negative integer $c_i$.
-We extend the equation to non-negative real number $c_i$ by applying a linear approximation
-
-```{=tex}
-\begin{equation} \label{eq:pvalue-beta-binomial-approx}
-P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+,
-\end{equation}
-```
-
-where $P(C \geq \lceil c_i \rceil)$ is calculated using Equation \ref{eq:pvalue-beta-binomial} and $P(C = \lfloor c_i \rfloor)$ is calculated by
-
-\begin{equation} \label{eq:pmf-beta-binomial}
-P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+.
-\end{equation}
-
-Besides, the parameter $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} and \ref{eq:pmf-beta-binomial} is usually unknown and hence needs to be estimated from the survey data.
-For low values of $\alpha$, only a few plots are attractive to the observers and tend to be selected.
-For higher values of $\alpha$, the distribution of the probability of each plot being selected is more even.
-@vanderplas2021statistical defines that a plot is $c$-interesting if $c$ or more participants select the plot as the most different.
-Given the definition, The expected number of plots selected at least $c$ times, $E[Z_c]$, is calculated as
-
-```{=tex}
-\begin{equation} \label{eq:c-interesting-expectation}
-E[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}
-```
-
-With Equation \ref{eq:c-interesting-expectation}, $\alpha$ can be estimated using maximum likelihood estimation.
-But for precise estimate of $\alpha$, additional human responses to Rorschach lineups, which is a type of lineup that consists of plots constructed from the same null data generating mechanism, are required.
-
-## Power of the tests
-
-The power of a model misspecification test is the probability that $H_0$ is rejected given the regression model is misspecified in a specific way. It is an important indicator
-when one is concerned about whether model assumptions have been violated.
-Although in practice, one might be more interested in knowing how much the residuals deviate from the model assumptions, and whether this deviation is of practical significance.
-
-The power of a conventional hypothesis test is affected by both the true parameter $\boldsymbol{\theta}$ and the sample size $n$.
-These two can be quantified in terms of effect size $E$ to measure the strength of the residual departures from the model assumptions.
-Details about the effect size is provided in Section \ref{effect-size} after the introduction of the simulation model used in our human subject experiment.
-The theoretical power of a test is sometimes not a trival solution, but it can be estimated if the data generating process is known.
-We use a predefined model to generate a large set of simulated data under different effect sizes, and record if the conventional test rejects $H_0$.
-The probability of the conventional test rejects $H_0$ is then fitted by a logistic regression formulated as
-
-```{=tex}
-\begin{equation} \label{eq:logistic-regression-1-1}
-Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),
-\end{equation}
-```
-
-\noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
-The effect size $E$ is the only predictor and the intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$, which is the desired significance level.
-
-The power of a visual test on the other hand, may additionally depend on the ability of the particular subject, as the skill of the individual may affect the number of observers who identify the data plot from the lineup [@majumder_validation_2013].
-To address this issue, @majumder_validation_2013 models the probability of a subject $j$ correctly picking the data plot from a lineup $l$ using a mixed-effect logistic regression, with subjects treated as random effect.
-Then, the estimated power of a visual test evaluated by a single subject is the predicted value obtained from the mix-effect model.
-However, this mix-effect model does not work with scenario where subjects are asked to select one or more most different plots.
-In this scenario, having the probability of a subject $j$ correctly picking the data plot from a lineup $l$ is insufficient to determine the power of a visual test because it does not provide information about the number of selections made by the subject for the calculation of the $p$-value (See Equation \ref{eq:pvalue-beta-binomial-approx}).
-Therefore, we directly estimate the probability of a lineup being rejected by assuming that individual skill has negligible effect on the variation of the power.
-This assumption is not necessary true, but it helps simplifying the model structure, thereby obviate a costly large-scale experiment to estimate complex covariance matrices.
-The same model given in Equation \ref{eq:logistic-regression-1-1} is applied to model the power of a visual test.
-
-To study various factors contributing to the power of both tests, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
-These include the distribution of the fitted values, type of the simulation model and the shape of the residual departures.
-
-# Experimental design
-
-An experiment is conducted in three data collection periods to investigate the difference between conventional hypothesis testing and visual inference in the application of linear regression diagnostics.
-Two types of departures, non-linearity and heteroskedasticity, are collected during data collection periods I and II.
-The data collection period III was designed primarily to measure human responses to null lineups so that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated.
-Additional lineups for both non-linearity and heteroskedasticity, using uniform fitted value distribution, were included so that the participants were evaluating some lineups with signal also.
-It would be too frustrating for participants to only be assigned lineups with all null plots.
-Overall, we collected `r vi_survey  %>% filter(!attention_check) %>% nrow()` evaluations on `r vi_survey  %>% filter(!attention_check) %>% pull(unique_lineup_id) %>% unique() %>% length()` unique lineups performed by `r vi_survey  %>% filter(!attention_check) %>% count(exp, set) %>% nrow()` subjects throughout three data collection periods.
-
-## Simulating departures from good residuals
-
-### Non-linearity
-
-Data collection period I is designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials on random vector $\boldsymbol{x}$ formulated as
-
-\begin{align} \label{eq:nonlinearity-model}
-\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
-\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
-\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
-\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
-\end{align}
-
-\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
-
-```{=tex}
-\begin{equation} \label{eq:scaling-function}
-g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x}))2k - k, \quad \text{for} \quad k > 0. 
-\end{equation}
-```
-
-According to @abramowitz1964handbook, Hermite polynomials were initially defined by @de1820theorie, but named after Hermite [@hermite1864nouveau] because of the unrecognisable form of Laplace's work.
-When simulating $\boldsymbol{z}_{raw}$, function `hermite` from the R package `mpoly` [@mpoly] is used to generate Hermite polynomials.
-
-The null regression model used to fit the realizations generated by the above model is formulated as
-
-```{=tex}
-\begin{equation} \label{eq:null-model}
-\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
-\end{equation}
-```
-\noindent where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
-
-Since $z = O(x^j)$, for $j > 1$, $z$ is a higher order term leaves out by the null regression, which will lead to model misspecification.
-
-Visual patterns of non-linearity are simulated using four different orders of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$).
-(A summary of the factors is given in Table \ref{tab:model-factor-table}.) The values of $j$ is chosen so that distinct shapes of non-linearity are included in the residual plot.
-These include "U", "S", "M" and "triple-U" shape as shown in Figure \ref{fig:different-shape-of-herimite}.
-A greater value of $j$ will result in a curve with more turning points.
-It is expected that the "U" shape will be the easiest one to detect because complex shape tends to be concealed by cluster of data points.
-
-Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups used in non-linearity detection.
-This lineup is produced by the non-linearity model with $j = 6$.
-The data plot location is $2^3 - 4$.
-All five subjects correctly identify the data plot from this lineup.
-
-```{r results='asis'}
-data.frame(j = c("2", "3", "6", "18", ""),
-           sigma = c("0.25", "1.00", "2.00", "4.00", ""),
-           a = c("-1", "0", "1", "", ""),
-           b = c("0.25", "1.00", "4.00", "16.00", "64.00"),
-           n = c("50", "100", "300", "", ""), 
-           x_dist = c("Uniform", "Normal", "Skewed", "Discrete uniform", "")) %>%
-  knitr::kable(col.names = NULL,
-               align = "rr|rr|rc",
-               escape = FALSE,
-               format = "latex",
-               booktabs = TRUE, 
-               caption  = 'Levels of the factors used in data collection periods I, II, III.', 
-               label = "model-factor-table" ) %>%
-  kableExtra::kable_styling(latex_options = "scale_down") %>%
-  kableExtra::add_header_above(c("Poly Order (sym1)", #"Order of Hermite polynomial\n(sym1)", 
-                                 "SD (sym3)", #"Standard deviation of polynomial model\n(sym3)",
-                                 "Shape (sym4)", # "Shape of heteroskedasticity\n(sym4)", 
-                                 "Ratio (sym5)", # Variance factor of heteroskedasticity model\n(sym5)",
-                                 "Size (sym6)", 
-                                 "Distribution of fitted values")) %>% #"Sample size\n(sym6)")) %>%
-  kableExtra::add_header_above(c("Non-linearity" = 2, 
-                               "Heteroskedasticity" = 2, "Common" = 2)) %>%
-  as.character() %>%
-  gsub("sym1", "$j$", .) %>%
-  gsub("sym3", "$\\\\sigma$", .) %>%
-  gsub("sym4", "$a$", .) %>%
-  gsub("sym5", "$b$", .) %>%
-  gsub("sym6", "$n$", .) %>%
-  cat()
-```
-
-```{r different-shape-of-herimite, fig.height = 2, fig.cap = "Polynomial forms generated for the residual plots used to assess detecting non-linearity.
-The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$."}
-
-# Facet label
-shape_labels <- c('He[2]:"U"', 'He[3]:"S"', 'He[6]:"M"', 'He[18]:"triple-U"')
-
-# Data for shape 1
-dat_shape_1 <- poly_model(shape = 1, 
-                          x = {
-                            raw_x <- rand_uniform(-1, 1);
-                            closed_form(~stand_dist(raw_x))
-                            }, 
-                          sigma = 0.05)$gen(300) %>%
-  mutate(shape = shape_labels[1])
-
-# Generate data for shape 2, 3 and 4. Reuse x and e.
-map_df(2:4, function(shape) {
-  poly_model(shape = shape, 
-                  x = {
-                    raw_x <- rand_uniform(-1, 1); 
-                    closed_form(~stand_dist(raw_x))
-                    }, 
-                  sigma = 0.05)$
-  gen(300, computed = select(dat_shape_1, x, e)) %>%
-  mutate(shape = shape_labels[shape])
-}) %>%
-
-  # Combined with data for shape 1
-  bind_rows(dat_shape_1) %>%
-  mutate(shape = factor(shape, levels = shape_labels)) %>%
-  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
-  facet_wrap(~shape, scales = "free", labeller = label_parsed, ncol = 4) +
-  xlab("Fitted values") +
-  ylab("Residuals")
-```
-
-```{r different-sigma, fig.height = 2, fig.cap = 'Examining the effect of $\\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform fitted value distribution and the "U" shape.
-As $\\sigma$ increases the signal strength decreases, to the point that the "U" is almost unrecognisable when $\\sigma=4$.'}
-
-# Generate data for sigma 0.5
-dat_sigma_05 <- poly_model(shape = 1, 
-                           x = {
-                             raw_x <- rand_uniform(-1, 1);
-                             closed_form(~stand_dist(raw_x))
-                             }, 
-                           sigma = 0.5)$gen(300) %>%
-  mutate(sigma = "sigma: 0.5")
-
-# Generate data for other sigma
-map(c(1, 2, 4), function(sigma) {
-  poly_model(shape = 1, 
-                  x = {
-                    raw_x <- rand_uniform(-1, 1); 
-                    closed_form(~stand_dist(raw_x))
-                    }, 
-                  sigma = sigma)$
-  gen(300, computed = select(dat_sigma_05, x)) %>%
-  mutate(sigma = paste0("sigma: ", sigma))
-}) %>%
-
-  # Combined with data for sigma 0.5
-  bind_rows(dat_sigma_05) %>%
-  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) + 
-  facet_wrap(~sigma, ncol = 4, scales = "free", labeller = label_parsed) +
-  xlab("Fitted values") +
-  ylab("Residuals")
-```
-
-```{r example-poly-lineup, fig.height = 7, fig.width = 7, fig.cap = "One of the lineups containing non-linearity patterns used in data collection period I.
-Can you spot the most different plot?
-The data plot is positioned at $2^3 + 1$."}
-VI_MODEL$plot_lineup(vi_lineup$poly_24$data, 
-                     remove_grid_line = TRUE, 
-                     theme = theme_light(),
-                     remove_axis = TRUE)
-```
-
-### Heteroskedasticity
-
-Data collection period II is designed to study the ability of human subjects to detect the appearance of a heteroskedasticity pattern under a simple linear regression model setting:
-
-```{=tex}
-\begin{align} \label{eq:heter-model}
-\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
-\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
-\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}), 
-\end{align}
-```
-
-\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vectors of size $n$ and $g(.)$ is the scaling function defined in Equation \ref{eq:scaling-function}.
-
-The null regression model used to fit the realizations generated by the above model is formulated exactly the same as Equation \ref{eq:null-model}.
-
-For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the predictor $\boldsymbol{x}$, which will lead to the presence of heteroskedasticity.
-Visual patterns of heteroskedasticity are simulated using three different shapes ($a$ = -1, 0, 1).
-(A summary of the factors can be found in Table \ref{tab:model-factor-table}.)
-
-Since $supp(X) = [-1, 1]$, choosing $a$ to be $-1$, $0$ and $1$ can generate "left-triangle", "butterfly" and "right-triangle" shape as displayed in Figure \ref{fig:different-shape-of-heter}.
-The term $(2 - |a|)$ maintains the magnitude of residuals across different values of $a$.
-
-```{r different-shape-of-heter, fig.height = 2.67, fig.cap = 'Heteroskedasticity forms used in the experiment.
-Three different shapes ($a = -1, 0, 1$) are used in the experiment to create left-triangle, "butterfly" and "right-triangle" shapes, respectively.'}
-
-set.seed(10086)
-
-a_labels <- c("left-triangle", "butterfly", "right-triangle")
-
-# Generate data for a = -1
-dat_a_n1 <- heter_model(a = -1, 
-                        x = {
-                          raw_x <- rand_uniform(-1, 1);
-                          closed_form(~stand_dist(raw_x))
-                          },
-                        b = 128)$gen(300) %>%
-  mutate(a = a_labels[1])
-
-# Generate data for other a
-map(c(0, 1), function(a) {
-  heter_model(a = a,
-              x = {
-                raw_x <- rand_uniform(-1, 1); 
-                closed_form(~stand_dist(raw_x))
-                }, 
-              b = 128)$
-    gen(300, computed = select(dat_a_n1, x, e)) %>%
-    mutate(a = a_labels[a + 2])
-}) %>%
-
-  # Combined with data for a = -1
-  bind_rows(dat_a_n1) %>%
-  mutate(a = factor(a, levels = a_labels)) %>%
-  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
-  facet_wrap(~a, scales = "free", ncol = 3) +
-  xlab("Fitted values") +
-  ylab("Residuals")
-```
-
-```{r different-b, fig.height = 2, fig.cap = "Five different values of $b$ are used in heteroskedasticity simulation to control the strength of the signal.
-Larger values of $b$ yield a bigger difference in variation, and stus stronger heteroskedasticity signal."}
-
-# Generate data for b = 0.25
-dat_b025 <- heter_model(a = 0, 
-                        x = {
-                          raw_x <- rand_uniform(-1, 1);
-                          closed_form(~stand_dist(raw_x))
-                          }, 
-                        b = 0.25)$gen(300) %>%
-  mutate(b = "b: 0.25")
-
-# Generate data for other b
-map(c(1, 4, 16, 64), function(b) {
-  heter_model(a = 0,
-              x = {
-                raw_x <- rand_uniform(-1, 1); 
-                closed_form(~stand_dist(raw_x))
-                }, 
-              b = b)$
-  gen(300, computed = select(dat_b025, x, e)) %>%
-  mutate(b = paste0("b: ", b))
-}) %>%
-
-  # Combined with data for b = 0.25
-  bind_rows(dat_b025) %>%
-  mutate(b = factor(b, levels = paste0("b: ", c(0.25, 1, 4, 16, 64)))) %>%
-  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
-  facet_wrap(~b, scales = "free", ncol = 5) +
-  xlab("Fitted values") +
-  ylab("Residuals")
-```
-
-```{r example-heter-lineup, fig.height = 7, fig.width = 7, fig.cap = "One of the lineups containing heteroskedasticity pattern used in data collection period II.
-Can you spot the most different plot?
-The data plot is positioned at $3^3 - 3^2$"}
-
-VI_MODEL$plot_lineup(vi_lineup$heter_471$data, 
-                     remove_grid_line = TRUE, 
-                     theme = theme_light(),
-                     remove_axis = TRUE)
-```
-
-An example lineup of this model used in data collection period II is shown in Figure \ref{fig:example-heter-lineup} with $a = -1$.
-The data plot location is $2^4 + 2$.
-Nine out of 11 subjects correctly identify the data plot from this lineup.
-
-### Factors common to both data collection periods
-
-Fitted values are a function of the independent variables, and the distribution of the observed values affects the distribution of the fitted values.
-In the best case scenario the fitted values will have a uniform distribution, which means that there is even coverage of possible observed values across all of the predictors.
-This is not always present in the collected data.
-Sometimes the fitted values are discrete because one or more predictors were measured discretely.
-The distribution may be relatively Gaussian, reflecting a linear combinatio of many predictors, adhering to the Central Limit Theorem.
-It is also common to see a skewed distribution of fitted values, if one or more of the predictors has a skewed distribution.
-This latter problem is usually corrected before modelling using a variable transformation.
-Our simulation assess this by using four different distributions to represent fitted values: (1) uniform, (2) normal, (3) skewed and (4) discrete uniform.
-This is constructed by defining the raw predictor $X_{raw}$ in four corresponding distributions:
-(1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$. We would expect that the best reading of residual plots occurs when the fitted values are uniformly distributed.
-
-Three different sample sizes are used, $n=50, 100, 300$ across the experiments.
-We would expect considerable variation in the signal strength in the simulated data plots with smaller $n$.
-A sample size of 300 is typically enough for structure to be visible in a scatter plot reliably.
-
-```{r different-dist, fig.height = 2, fig.cap = "Variations in fitted values, that might affect perception of residual plots. Four different distributions are used."}
-
-# Data for uniform distribution
-dat_dist_1 <- poly_model(shape = 1, 
-                      x = {
-                        raw_x <- rand_uniform(-1, 1);
-                        closed_form(~stand_dist(raw_x))
-                        }, 
-                      sigma = 0.5)$gen(300) %>%
-  mutate(x_dist = "Uniform")
-
-# Generate data for other distributions
-dat_dist_2 <- poly_model(shape = 1,
-                         x = {
-                           raw_x <- rand_normal(sigma = 0.3); 
-                           closed_form(~stand_dist(raw_x))
-                           }, 
-                         sigma = 0.5)$
-  gen(300, computed = select(dat_dist_1, e)) %>%
-  mutate(x_dist = "Normal")
-
-dat_dist_3 <- poly_model(shape = 1,
-                         x = {
-                           raw_x <- rand_lognormal(sigma = 0.6); 
-                           closed_form(~stand_dist(raw_x/3 - 1))
-                           }, 
-                         sigma = 0.5)$
-  gen(300, computed = select(dat_dist_1, e)) %>%
-  mutate(x_dist = "Skewed")
-
-dat_dist_4 <- poly_model(shape = 1,
-                         x = {
-                           raw_x <- rand_uniform_d(k = 5, even = TRUE); 
-                           closed_form(~stand_dist(raw_x))
-                           }, 
-                         sigma = 0.5)$
-  gen(300, computed = select(dat_dist_1, e)) %>%
-  mutate(x_dist = "Discrete uniform")
-
-# Generate and plot data for discrete uniform distribution
-bind_rows(dat_dist_1, dat_dist_2, dat_dist_3, dat_dist_4) %>%
-  mutate(x_dist = factor(x_dist, 
-                         levels = c("Uniform", 
-                                    "Normal", 
-                                    "Skewed", 
-                                    "Discrete uniform"))) %>%
-  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  facet_wrap(~x_dist, ncol = 4, scales = "free")
-```
-
-```{r different-n, fig.height = 2.67, fig.cap = 'Examining the effect of signal strength for the three different values of $n$ used in the experiment, for non-linear structure with fixed $\\sigma = 1.5$, uniform fitted value distribution, and "S" shape.
-For these factor levels, only when $n = 300$ is the "S" shape clearly visible.'}
-
-set.seed(666)
-
-# Generate data for 300 observations
-dat_n_300 <- poly_model(shape = 2, 
-                        x = {
-                          raw_x <- rand_uniform(-1, 1)
-                          closed_form(~stand_dist(raw_x))
-                        },
-                        sigma = 1)$gen(300) %>%
-  mutate(n = "N: 300")
-
-# Generate data for other settings
-map(c(50, 100), function(n) {
-  poly_model(shape = 2, 
-                   x = {
-                     raw_x <- rand_uniform(-1, 1);
-                     closed_form(~stand_dist(raw_x))
-                     }, 
-             sigma = 1)$
-    gen(n, computed = select(dat_n_300[1:n, ], x, e)) %>%
-    mutate(n = paste0("N: ", n))
-}) %>%
-
-  # Combined with data for 300 observations
-  bind_rows(dat_n_300) %>%
-  mutate(n = factor(n, levels = c("N: 50", "N: 100", "N: 300"))) %>%
-  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
-  facet_wrap(~n, scales = "free", ncol = 3) +
-  xlab("Fitted values") +
-  ylab("Residuals")
-```
-
-## Experimental setup
-
-### Controlling the strength of the signal
-
-As summarised in Table \ref{tab:model-factor-table}, three additional parameters $n$, $\sigma$ and $b$ are used to control the strength of the signal so that different difficulty levels of lineups are generated, and therefore, the estimated power curve will be smooth and continuous.
-Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ are used in data collection periods I and II respectively.
-Figure \ref{fig:different-sigma} and \ref{fig:different-b} demonstrate the impact of these two parameters.
-A large value of $\sigma$ will increase the variation of the error of the non-linearity model and decrease the visibility of the visual pattern.
-The parameter $b$ controls the standard deviation of the error across the support of the predictor.
-Given $x \neq a$, a larger value of $b$ will lead to a larger ratio of the variance at $x$ to the variance at $x - a = 0$, making the visual pattern more obvious.
-
-Three different sample sizes are used (n = 50, 100, 300) in all three data collection periods.
-It can be observed from Figure \ref{fig:different-n} that with fewer data points drawn in a residual plot, the visual pattern is more difficult to be detected.
-
-### Effect size
-
-Effect size in statistics measures the strength of the signal relative to the noise. It is surprisingly difficult to quantify in general, even for simulated data as used in this experiment.
-
-For the non-linearity model, the key items defining effect size are sample size ($n$) and variance of the error term ($\sigma^2$), and so effect size would be roughly calculated as $\sqrt{n}/{\sigma}$.
-As sample size increases the effect size would increase, but as variance increases the effect size decreases.
-However, it is not clear how the additional parameter for the model polynomial order, $k$, should be incorporated.
-Intuitively, the large $k$ means more complex pattern, which likely means effect size would decrease.
-For the purposes of our calculations we have chosen to use an approach based on Kullback-Leibler divergence [@kullback1951information], coupled with simulation.
-This formulation defines effect size to be:
-
-$$E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right)$$
-\noindent where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,
-$\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a$ is the residual operator, $\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ is the hat matrix, $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients, and $\sigma^2\boldsymbol{I}$ is the assumed covariance matrix of the error term when $H_0$ is true.
-
-In the heteroskedasticity model, the key elements for measuring effect size are sample size, $n$, and the ratio of the biggest variance to smallest variance, $b$.
-Larger values of both would produce higher effect size.
-However, it is not clear how to incorporate the additional shape parameter, $a$.
-Thus the same approach is used here, where the formula can be written as:
-
-$$E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)$$
-\noindent where $\boldsymbol{V}$ is the actual covariance matrix of the error term.
-
-Derivations for these equations are provided in the Appendix.
-
-To compute the effect size for each lineup we simulate a sufficient large number of samples from the same model, in each sample, the number of observations $n$ is fixed.
-We then compute the effect size for each sample and take the average as the final value.
-This ensures lineups constructed with the same experimental factors will share the same effect size.
-
-### Subject allocation
-
-As shown in Table \ref{tab:model-factor-table}, there are a total of $4 \times 4 \times 3 \times 4 = 192$ and $3 \times 5 \times 3 \times 4 = 180$ number of combinations of parameter values for non-linearity model and heteroskedasticity model respectively.
-Three replications are made for each of the combination results in $192 \times 3 = 576$ and $180 \times 3 = 540$ lineups.
-In addition, each lineup is designed to be evaluated by five different subjects.
-After attempting some pilot studies internally, we decide to present a block of 20 lineups to every subject.
-And to ensure the quality of the survey data, two lineups with obvious visual patterns are included as attention checks.
-Thus, $576 \times 5 / (20-2) = 160$ and $540 \times 5 / (20-2) = 150$ subjects are recruited to satisfy the design of the data collection period I and II respectively.
-
-As mentioned in Section \ref{power-of-the-tests}, $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using null lineups.
-Three replications are made for $3 \times 4 = 12$ combinations of common factors $n$ and fitted value distribution, results in $12 \times 3 = 36$ lineups included in data collection period III.
-In these lineups, the data of the data plot is generated from a model with zero effect size, while the data of the 19 null plots are generated using the same simulation method discussed in Section \ref{visual-test-procedure-based-on-lineups}.
-This generation procedure differs from the canonical Rorschach lineup procedure, which requires that all 20 plots are generated directly from the null model.
-However, these lineups serve the same fundamental purpose: to assess the number of visually interesting plots generated under $H_0$.
-
-To account for the fact that our simulation method for these lineups is not the Rorschach procedure, we use the method suggested in @vanderplas2021statistical for typical lineups containing a data plot to estimate $\alpha$.
-We have included a sensitivity analysis in the Appendix to examine the impact of the variance of the $\alpha$ estimate on our findings.
-
-All lineups consist of only null plots are planned to be evaluated by 20 subjects. 
-However, presenting only these lineups to subjects are considered to be bad practices as subjects will lose interest quickly. 
-Therefore, we plan to collect 6 more evaluations on the 279 lineups with uniform fitted value distribution, result in $(36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133$ subjects recruited for data collection period III.
-
-### Collecting results
-
-Subjects for all three data collection periods are recruited from an crowdsourcing platform called Prolific [@palan2018prolific].
-Prescreening procedure is applied during the recruitment, subjects are required to be fluent in English, with $98\%$ minimum approval rate and 10 minimum submissions in other studies.
-
-During the experiment, every subject is presented with a block of 20 lineups.
-A lineup consists of a randomly placed data plot and 19 null plots, which are all residual plots drawn with raw residuals on the y-axis and fitted values on the x-axis.
-An additional horizontal red line is added at $y = 0$ as a helping line.
-
-The data of the data plot is simulated from one of two models described in Section \ref{simulating-departures-from-good-residuals}, while the data of the remaining 19 null plots are generated by the residual rotation technique discussed in Section \ref{visual-test-procedure-based-on-lineups}.
-
-In every lineup evaluation, the subject is asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots are from others.
-If there is no noticeable difference between plots in a lineup, subjects are permitted to select zero plots without providing the reason.
-No subject are shown the same lineup twice.
-Information about preferred pronoun, age group, education, and previous experience in visual experiments are also collected.
-A subject's submission is only accepted if the data plot is identified for at least one attention check.
-Data of rejected submissions are discarded automatically to maintain the overall data quality.
-
-# Results
-
-## Overview
-
-There are `r 160 * 18`, `r 150 * 18` and `r 133 * 18 - nrow(filter(vi_survey, null_lineup, !attention_check))` lineups evaluation made by 160, 150 and 133 subjects recruited for data collection periods I, II and III respectively.
-In the total of `r nrow(filter(vi_survey, !attention_check))` lineup evaluations, `r nrow(filter(vi_survey, type == "polynomial", !attention_check))` use lineups produced by the non-linearity model, and `r nrow(filter(vi_survey, type == "heteroskedasticity", !attention_check))` use lineups produced by the heteroskedasticity model.
-Besides, there are `r nrow(filter(vi_survey, attention_check))` attention checks and `r nrow(filter(vi_survey, null_lineup, !attention_check))` evaluations on null lineups needed for the estimate of $\alpha$ not included in the analysis.
-The collated dataset is provided in `vi_survey` of the `visage` `R` package.
-
-In the following analysis, lineups with uniform fitted values will be the focus.
-Visual patterns are more likely to be revealed under a uniform distribution.
-Additionally, we have collected extra evaluations on these lineups, which will result in a more reliable analysis.
-Analysis of lineups with other fitted value distributions can be found in Section \ref{effect-of-fitted-value-distributions}.
-
-## Power comparison of different tests
-
-```{r cache = TRUE}
-# Define the minimum and maximum effect size
+                    caption = "Examining how decisions might change if $\\hat{\\alpha}$ was different.
+Percentage of lineups that where the $p$-value would switch to above or below 0.05, when $\\hat{\\alpha}$ is varied by a fractional amount.")
+```
+
+## Effect of number of evaluations on the power of a visual test
+
+When comparing power of visual tests across different fitted value distributions, we have discussed the number of evaluations on a lineup will affect the power of the visual test.
+Using the lineups with uniform fitted value distribution, we show in Figure \ref{fig:n_eval} the change of power of visual tests due to different number of evaluations.
+It can be learned that as the number of evaluations increases, the power will increase but the margin will decrease.
+Considering we have eleven evaluations on lineups with uniform fitted value distribution, and five evaluations on other lineups, it is necessary to use the same number of evaluations for each lineup in comparison.
+
+```{r}
 min_poly_es <- vi_survey %>% 
@@
   max()
-
-# The glm model for RESET test
-reset_poly_glmmod <- poly_conv_sim %>%
+```
+
+```{r n_eval, fig.cap = "Change of power of visual tests for different number of evalutions on lineups with uniform fitted value distribution.
+The power will incrase as the number of evaluations increases, but the margin will decrase.", fig.width=6, fig.height=4, out.width="70%"}
+eval_n <- function(n_eval, boot) {
+  map_df(1:boot, function(boot_id) {
+  vi_survey %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
   filter(x_dist == "uniform") %>%
+  filter(type == "polynomial") %>%
+  group_by(unique_lineup_id) %>%
+  slice_sample(n = n_eval) %>%
+  ungroup() %>%
+  calc_p_value_multi(lineup_id = unique_lineup_id,
+                     detect = detect,
+                     n_sel = num_selection,
+                     alpha = alpha) %>%
+  mutate(boot_id = boot_id)
+}) %>%
+  left_join(select(vi_survey, unique_lineup_id, effect_size)) %>%
+  mutate(reject = as.numeric(p_value <= 0.05)) %>%
+  mutate(offset0 = log(0.05/0.95)) %>%
+  select(-unique_lineup_id, -p_value) %>%
+  mutate(n_eval = n_eval)
+}
+
+map_df(seq(1, 11, 2), ~eval_n(.x, 1)) %>%
+  select(-boot_id) %>%
+  nest(c(effect_size, reject, offset0)) %>%
+  mutate(mod = map(data, 
+                   ~glm(reject ~ effect_size - 1, 
+                        family = binomial(), 
+                        data = .x,
+                        offset = offset0))) %>%
+  mutate(power = map(mod, function(mod) {
+    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
+                         offset0 = log(0.05/0.95))
+    result$power <- predict(mod, type = "response", newdata = result)
+    result
+  })) %>%
+  select(-data, -mod) %>%
+  unnest(power) %>%
   mutate(log_effect_size = log(effect_size)) %>%
-  mutate(reject = as.numeric(RESET4_p_value <= 0.05)) %>%
-  mutate(offset0 = log(0.05/0.95)) %>%
-  select(x_dist, effect_size, log_effect_size, offset0, reject) %>%
-  glm(reject ~ effect_size - 1, family = binomial(), data = ., offset = offset0)
-
-# The prediction for all conventional tests on poly model
-pred_conv_poly <- poly_conv_sim %>%
+  ggplot() +
+  geom_line(aes(log_effect_size, power, col = factor(n_eval)),
+            size = 1) +
+  labs(col = "N. evals") +
+  xlab(expression(log[e] (Effect_size))) +
+  theme_light() +
+  scale_color_brewer(palette = "Reds")
+```
+
+## Power of a RESET test under different auxiliary regression formulas
+
+It is found in the result that the power of a RESET test will be affected by the highest order of fitted values included in the auxiliary formula.
+And we suspect that the current recommendation of the highest order - four, is insufficient to test complex non-linear structures such as the "Triple-U" shape designed in this paper.
+Figure \ref{fig:reset} illustrates the change of power of RESET test while testing the "U" shape and the "Triple-U" shape with different highest orders.
+Clearly, when testing a simple shape like the "U" shape, the highest order has very little impact on the power.
+But for testing the "Triple-U" shape, there will be a loss of power if the recommended order is used.
+To avoid the loss of power, the highest order needs to be set to at least six.
+
+```{r cache = TRUE}
+reset_diff_dat <- poly_conv_sim %>%
   filter(x_dist == "uniform") %>%
-  select(-RESET3_p_value, -(RESET5_p_value:RESET10_p_value), -F_p_value) %>%
-  rename(RESET_p_value = RESET4_p_value) %>%
-  pivot_longer(RESET_p_value:SW_p_value) %>%
-  mutate(name = gsub("_p_value", " test", name)) %>%
+  filter(shape == 4) %>%
+  select(-F_p_value, -BP_p_value, -SW_p_value) %>%
+  pivot_longer(RESET3_p_value:RESET10_p_value) %>%
+  mutate(name = gsub("RESET(.*)_p_value", "\\1", name)) %>%
   mutate(reject = value <= 0.05) %>%
@@
   unnest(power) %>%
-  mutate(log_effect_size = log(effect_size))
-
-# Data needed by the visual test
-visual_poly_dat <- vi_survey %>%
-  filter(!null_lineup, 
-         !attention_check, 
-         type == "polynomial", 
-         x_dist == "uniform") %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  mutate(reject = as.numeric(p_value <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  mutate(shape = 4)
+```
+
+```{r cache = TRUE}
+reset_diff_dat2 <- poly_conv_sim %>%
+  filter(x_dist == "uniform") %>%
+  filter(shape == 1) %>%
+  select(-F_p_value, -BP_p_value, -SW_p_value) %>%
+  pivot_longer(RESET3_p_value:RESET10_p_value) %>%
+  mutate(name = gsub("RESET(.*)_p_value", "\\1", name)) %>%
+  mutate(reject = value <= 0.05) %>%
+  select(effect_size, name, reject) %>%
   mutate(offset0 = log(0.05/0.95)) %>%
-  select(effect_size, offset0, reject)
-
-# The glm model for visual test
-visual_poly_glmmod <- glm(reject ~ effect_size - 1, 
-                          family = binomial(), 
-                          data = visual_poly_dat, 
-                          offset = offset0)
-
-# The prediction of the visual test
-pred_visual_poly <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
-                               offset0 = log(0.05/0.95)) %>%
-  mutate(power = predict(visual_poly_glmmod, 
-                         type = "response",
-                         newdata = data.frame(effect_size = effect_size,
-                                              offset0 = offset0))) %>%
-  mutate(log_effect_size = log(effect_size))
-
-# The prediction of the boot visual test
-pred_visual_poly_boot <- map_df(1:500, function(boot_id) {
-  slice_sample(visual_poly_dat, n = nrow(visual_poly_dat), replace = TRUE) %>%
   nest(dat = c(effect_size, offset0, reject)) %>%
@@
   mutate(log_effect_size = log(effect_size)) %>%
-    mutate(boot_id = boot_id)
-  })
-```
-
-Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with uniform fitted values, against the natural logarithm of the effect $log_e(E)$, with a 5% significance level.
-At the bottom of the figure \ref{fig:polypower}, there are a sequence of example residual plots with increasing levels of $log_e(E)$.
-Readers can evaluate them from left to right and determine at which level the departure from a good residual plot becomes detectable.
-
-As discussed in Section \ref{conventionally-testing-for-departures}, many conventionally tests are available for detecting residual departures.
-Implementation-wise, the built-in R package `stats` provides some commonly used residual-based tests, such as Shapiro-Wilk test.
-A more comprehensive collection of regression diagnostics tests can be found in the R package `lmtest` [@lmtest].
-In terms of heteroskedasticity diagnostics, the R package `skedastic` [@skedastic] collects and implements 25 existing conventional tests published since 1961.
-
-We pick RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`.
-Among them, RESET test is the only exact and appropriate test in this scenario.
-Both the Breusch-Pagan test and the Shapiro-Wilk test are approximate and inappropriate tests.
-Their estimated power is shown in Figure \ref{fig:polypower}.
-To set up the RESET test, we include different powers of fitted values as proxies.
-According to @ramsey_tests_1969, there are no general rules for the power of the fitted values needed by the RESET test, but it finds power up to four is usually sufficient.
-Thus, we follow this guideline to conduct the RESET test.
-For the Breusch-Pagan test, the choice of predictors in the auxiliary regression is left to the user [@breusch_simple_1979].
-But as @waldman1983note suggested, it is a good choice for the set of auxiliary predictors in the Breusch-Pagan test be the same as the White test.
-Thus, we include both $x$ and $x^2$ in the auxiliary regression.
-
-Figure \ref{fig:heterpower} is similar to Figure \ref{fig:polypower}, but shows corresponding information on lineups produced by the heteroskedasticity model.
-In this scenario, the visual test is compared to an approximate test - Breusch-Pagan test, and two other inappropriate tests - RESET test and Shapiro-Wilk test.
-
-For non-linearity patterns, the power curve of RESET test climbs aggressively from
-`r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(0), offset0 = log(0.05/0.95)), type = "response")) * 100)`% to `r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(2.5), offset0 = log(0.05/0.95)), type = "response")) * 100)`% as $log_e(E)$ increases from 0 to 2.5, while power of other tests respond inactively to the change of effect, showing that RESET test is much more sensitive to weak non-linear structure.
-Meanwhile, no noticeable residual departures can be spotted from the example residual plots.
-
-In terms of heteroskedasticity patterns, the power of Breusch-Pagan test is constantly greater than the power of visual test.
-At $log_e(E) \approx 2.5$, where the power curve of the visual test remains at a low level, the Breusch-Pagan test has around 50% chance of rejecting $H_0$.
-Similarly, the visual feature is nearly unobservable from the example residual plots at this level of effect size.
-
-The power of visual test arises steadily as $log_e(E)$ increases from 3 to 5 for both non-linearity patterns and heteroskedasticity patterns, suggesting that the effect starts to make significant impact on the degree of the presence of the designed visual features.
-This can also be observed from the example residual plots that when $log_e(E) = 3.5$, a weak "S-shape" and a weak "right-triangle" shape are presented in Figure \ref{fig:polypower} and Figure \ref{fig:heterpower} respectively.
-The visual pattern becomes clearer as $log_e(E)$ increases.
-At $log_e(E) = 5$, the power of visual tests for both patterns reaches almost 100%.
-
-Power curves of inappropriate tests show improvement as the effect increases but at a lower rate than the visual test in both scenarios.
-This coincides the point made by @cook1982residuals that residual-based tests for a specific type of model defect may be sensitive to other types of model defects.
-The power curve of RESET test remains at around 5% in Figure \ref{fig:heterpower} since there are no non-linear terms leave out in the heteroskedasticity model and $H_0$ of the test is always satisfied.
-
-Overall, the power comparison suggests that conventional tests differs significantly from visual tests in two regression diagnostics scenarios designed by us.
-Visual test have much higher tolerance of the residual departures than the conventional test.
-Since fail to reject $H_0$ in a visual test literally means that there are no obvious visual discoveries found in the residual plot, analysts and the general public as the consumers of the output may not be convinced of the existence of significant residual departures in spite of the rejection of $H_0$ given by the conventional test.
-Even if the rejection is accepted, the model violation may be considered as impactless due to the fact that they are not clearly visible.
-Besides, the sensitivity of the conventional test to weak residual departures could also distract and discourage analysts from finding simple but good linear approximation to the data.
-The rejection of $H_0$ because of human acceptable and negligible residual departures is not practically meaningful and useful to analysts and decision makers.
-In contrast, if the strict correctness of the model assumption is of particular interest, conducting a conventional test is still the recommended choice based on the findings.
-
-```{r}
-# Analyse uniform polynomial data
-p <- ggplot() +
-
-  geom_point(data = visual_poly_dat,
-             aes(log(effect_size), reject),
-             alpha = 0.15) +
-
-  # Draw conventional test power curve
-  geom_line(data = pred_conv_poly,
-            aes(log_effect_size, power, col = name),
+  mutate(shape = 1)
+```
+
+```{r reset, fig.cap = 'Change of power of RESET tests for different orders of fitted values included in the auxiliary formula.
+The left panel is the power of testing the "U" shape and the right panel is the power of testing the "Triple-U" shape.
+The power will not be greatly affected by the highest order in the case of testing the "U" shape.
+In the case of testing the "Triple-U" shape, the highest order needs to be set to at least six to avoid the loss of power.', fig.height = 4}
+reset_diff_dat %>%
+  bind_rows(reset_diff_dat2) %>%
+  mutate(shape = factor(ifelse(shape == 1, "U", "Triple-U"), levels = c("U", "Triple-U"))) %>%
+  ggplot() +
+  geom_line(aes(log_effect_size, power, col = as.factor(as.integer(name))),
             size = 1) +
-
-  # Draw boot visual test curves
-  geom_line(data = pred_visual_poly_boot,
-            aes(log_effect_size, power, col = "Visual test", group = boot_id),
-            size = 1,
-            alpha = 0.01) +
-
-  # Draw visual test power curve
-  geom_line(data = pred_visual_poly,
-            aes(log_effect_size, power, col = "Visual test"), 
-            size = 1) +
-
-  # Theme
-  theme_light(base_size = 5) +
-  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
-  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
-  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
+  theme_light() +
+  labs(col = "Highest order") +
   xlab(expression(log[e] (Effect_size))) +
-  ylab("Power") +
-  labs(col = "", size = "# lineups")
-```
-
-```{r cache = TRUE}
-# target_sigma is obtained by evaluating a vector of sigma and picking those 
-# with expected effect 0 ~ 5
-
-target_sigma <- c(3.65, 2.84, 2.21, 1.72, 1.34, 1.04, 0.81, 0.63, 0.49, 0.38, 0.3)
-
-# Calculate the effect size 
-es <- map_dbl(target_sigma, 
-              ~log(poly_model(shape = 2, 
-                              x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, 
-                              sigma = .x)$
-                     average_effect_size(n = 100, type = "kl")
-                   )
-              )
-```
-
-```{r polypower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for non-linear patterns (uniform fitted values only).
-Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups.
-Surrounding lines of the visual test show the estimated power of 500 bootstrap samples.
-Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by vertical lines in main plot.
-Where would you draw the line of too much non-linearity in the residuals?
-For the RESET test this is around log effect size 2, but for the visual test it is around 3.5.", dev = 'png', dpi = 450}
-
-# 106
-# 112
-# 113
-# 115
-# 132
-# 133
-# 144!
-# 147
-set.seed(132)
-
-# Base data
-ex_dat <- poly_model(shape = 2, 
-                     x = rand_uniform(-1, 1), 
-                     sigma = 1)$gen(n = 100)
-plist <- list()
-j <- 0
-
-# Obtain data at different effect. Reuse x and recompute e.
-for (i in target_sigma) {
-  j <- j + 1
-plist[[j]] <- poly_model(shape = 2, 
-                         x = rand_uniform())$
-                  gen(n = 100, 
-                      computed = ex_dat %>%
-                        mutate(e = sqrt(i) * e) %>%
-                        select(x, e)
-                      ) %>%
-  VI_MODEL$plot(theme = theme_light(base_size =  5), 
-                remove_axis = TRUE, 
-                remove_grid_line = TRUE, size = 0.25, stroke = 0.25) +
-    ggtitle(round(es[j], 1)) +
-    theme(aspect.ratio = 0.8)
-}
-
-# Use area() to design layout
-map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
-  reduce(c) %>%
-  c(patchwork::area(3, 1, 20, 11), .) %>%
-  patchwork::wrap_plots(append(plist, list(p), after = 0), 
-                        design = .)
-```
-
-```{r cache = TRUE}
-# Define the minimum and maximum effect size
-min_heter_es <- vi_survey %>% 
-  filter(!null_lineup, 
-         !attention_check,
-         type == "heteroskedasticity", 
-         x_dist == "uniform") %>%
-  pull(effect_size) %>%
-  min()
-max_heter_es <- vi_survey %>% 
-  filter(!null_lineup, 
-         !attention_check,
-         type == "heteroskedasticity", 
-         x_dist == "uniform") %>%
-  pull(effect_size) %>%
-  max()
-
-# The prediction for all conventional tests on heter model
-pred_conv_heter <- heter_conv_sim %>%
-  filter(x_dist == "uniform") %>%
-  select(-RESET3_p_value, -(RESET5_p_value:RESET10_p_value), -F_p_value) %>%
-  rename(RESET_p_value = RESET4_p_value) %>%
-  pivot_longer(RESET_p_value:SW_p_value) %>%
-  mutate(name = gsub("_p_value", " test", name)) %>%
-  mutate(reject = value <= 0.05) %>%
-  select(effect_size, name, reject) %>%
-  mutate(offset0 = log(0.05/0.95)) %>%
-  nest(dat = c(effect_size, offset0, reject)) %>%
-  mutate(mod = map(dat, 
-                   ~glm(reject ~ effect_size - 1, 
-                        family = binomial(), 
-                        data = .x,
-                        offset = offset0))) %>%
-  mutate(power = map(mod, function(mod) {
-    result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
-                         offset0 = log(0.05/0.95))
-    result$power <- predict(mod, type = "response", newdata = result)
-    result
-  })) %>%
-  select(-dat, -mod) %>%
-  unnest(power) %>%
-  mutate(log_effect_size = log(effect_size))
-
-# Data needed by the visual test
-visual_heter_dat <- vi_survey %>%
-  filter(!null_lineup, 
-         !attention_check, 
-         type == "heteroskedasticity", 
-         x_dist == "uniform") %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  mutate(reject = as.numeric(p_value <= 0.05)) %>%
-  mutate(offset0 = log(0.05/0.95)) %>%
-  select(effect_size, offset0, reject)
-
-# The glm model for visual test
-visual_heter_glmmod <- glm(reject ~ effect_size - 1, 
-                           family = binomial(), 
-                           data = visual_heter_dat, 
-                           offset = offset0)
-
-# The prediction of the visual test
-pred_visual_heter <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
-                                offset0 = log(0.05/0.95)) %>%
-  mutate(power = predict(visual_heter_glmmod, 
-                         type = "response",
-                         newdata = data.frame(effect_size = effect_size,
-                                              offset0 = offset0))) %>%
-  mutate(log_effect_size = log(effect_size))
-
-# The prediction of the boot visual test
-pred_visual_heter_boot <- map_df(1:500, function(boot_id) {
-  slice_sample(visual_heter_dat, n = nrow(visual_heter_dat), replace = TRUE) %>%
-    nest(dat = c(effect_size, offset0, reject)) %>%
-    mutate(mod = map(dat, 
-                     ~glm(reject ~ effect_size - 1, 
-                          family = binomial(), 
-                          data = .x,
-                          offset = offset0))) %>%
-    mutate(power = map(mod, function(mod) {
-      result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
-                           offset0 = log(0.05/0.95))
-      result$power <- predict(mod, type = "response", newdata = result)
-      result
-    })) %>%
-    select(-dat, -mod) %>%
-    unnest(power) %>%
-    mutate(log_effect_size = log(effect_size)) %>%
-    mutate(boot_id = boot_id)
-  })
-```
-
-```{r}
-# Analyse uniform heteroskedasticity data
-p <- ggplot() +
-
-  geom_point(data = visual_heter_dat,
-             aes(log(effect_size), reject),
-             alpha = 0.15) +
-
-  # Draw conventional test power curve
-  geom_line(data = pred_conv_heter,
-            aes(log_effect_size, power, col = name),
-            size = 1) +
-
-  # Draw boot visual test curves
-  geom_line(data = pred_visual_heter_boot,
-            aes(log_effect_size, power, col = "Visual test", group = boot_id),
-            size = 1,
-            alpha = 0.01) +
-
-  # Draw visual test power curve
-  geom_line(data = pred_visual_heter,
-            aes(log_effect_size, power, col = "Visual test"), 
-            size = 1) +
-
-  # Theme
-  theme_light(base_size = 5) +
-  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
-  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
-  # scale_size_manual(values = seq(1, 9)) +
-  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
-  xlab(expression(log[e] (Effect_size))) +
-  ylab("Power") +
-  labs(col = "", size = "# lineups")
-```
-
-```{r cache = TRUE}
-# target_b is obtained by evaluating a vector of b and picking those 
-# with expected effect 0 ~ 5
-
-target_b <- c(0.17, 0.23, 0.32, 0.46, 0.67, 1.04, 1.7, 3, 6.3, 16.2, 61)
-
-# Calculate the effect size 
-es <- map_dbl(target_b, 
-              ~log(heter_model(a = 1, 
-                               b = .x, 
-                               x = {raw_x <- rand_normal(sigma = 0.3);closed_form(~stand_dist(raw_x))})$
-                     average_effect_size(n = 100, type = "kl")
-                   )
-              )
-```
-
-```{r heterpower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only).
-Main plot shows the power curves, with dots indicating human evaluations of lineups.
-Surrounding lines of the visual test show the estimated power of 500 bootstrap samples.
-Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by vertical lines in main plot.
-Where would you draw the line of too much heteroskedasticity in the residuals?
-For the BP test this is around log effect size 2.5, but for the visual test it is around 3.5.", dev = 'png', dpi = 450}
-
-# 10091
-# 101
-# 105
-# 106
-# 112
-# 119
-# 133
-set.seed(135)
-
-# Base data
-ex_dat <- heter_model(a = 1, 
-                      x = rand_uniform(-1, 1), 
-                      b = 1)$gen(n = 100)
-plist <- list()
-j <- 0
-
-# Obtain data at different b
-for (i in target_b) {
-  j <- j + 1
-  plist[[j]] <- heter_model(a = 1, 
-                            b = i, 
-                            x = rand_uniform(-1, 1))$
-                  gen(n = 100, computed = select(ex_dat, x, e)) %>%
-  VI_MODEL$plot(theme = theme_light(base_size =  5),
-                remove_axis = TRUE,
-                remove_grid_line = TRUE, size = 0.25, stroke = 0.25) +
-    ggtitle(round(es[j], 1)) +
-    theme(aspect.ratio = 0.8)
-}
-
-# Use area() to design layout
-map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
-  reduce(c) %>%
-  c(patchwork::area(3, 1, 20, 11), .) %>%
-patchwork::wrap_plots(append(plist, list(p), after = 0), 
-                      design = .)
-```
-
-## Comparison of test decisions based on $p$-values
-
-```{r}
-# Dataset for p-value comparison
-p_value_cmp_dat <- vi_survey %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  filter(x_dist == "uniform") %>%
-  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
-  mutate(conv_reject = ifelse(conventional_p_value <= 0.05, "Reject", "Not"),
-         reject = ifelse(p_value <= 0.05, "Reject", "Not")) %>%
-  mutate(conv_reject = factor(conv_reject, levels = c("Reject", "Not")),
-         reject = factor(reject, levels = c("Reject", "Not"))) %>%
-  select(type, conv_reject, reject)
-```
-
-The power comparison illustrates that appropriate conventional tests will reject $H_0$ more aggressively than visual tests.
-In this section, we explore how often they agree with each other by investigating the rejections for the two model designs based on $p$-values for each lineup.
-
-Figure \ref{fig:p-value-comparison} provides a mosaic plot showing the rejection rate of visual tests and conventional tests for both non-linearity patterns and heteroskedasticity patterns.
-
-For lineups containing non-linearity patterns, conventional tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% and visual tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time.
-Of the lineups rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by the visual test, that is, approximately half as many as the conventional test.
-There are no lineups that are rejected by the visual test but not by the conventional test.
-
-In terms of lineups containing heteroskedasticity patterns, `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by conventional tests, while `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by visual tests.
-When the conventional test rejects a lineup, there is a great chance (`r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%) the visual test will also reject it.
-
-```{r}
-nr <- p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Not") %>% count(reject)
-```
-
-Surprisingly, the visual test rejects `r nr$n[1]` of the `r nr$n[1]+nr$n[2]` (`r round(nr$n[1]/(nr$n[1]+nr$n[2])*100, 0)`%) of lineups where the conventional test does not reject.
-Figure \ref{fig:heter-example} shows this lineup.
-The data plot in position seventeen displays a relatively strong heteroskedasticity pattern, and has a strong effect size ($log_e(E)=4.02$).
-This is reflected by the visual test $p\text{-value} = 0.026$.
-But the Breusch-Pagan test $p\text{-value} = 0.056$, is slightly above the significance cutoff of $0.05$.
-This lineup was evaluated by 11 subjects, it has experimental factors $a = 0$ ("butterfly" shape), $b = 64$ (large variance ratio), $n = 50$ (small sample size), and a uniform distribution for the fitted values.
-It must be the small sample size that may have resulted in the lack of detection.
-
-```{r p-value-comparison, fig.width = 8, fig.height = 4, fig.cap = "Rejection rate ($p$-value $\\leq0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
-The visual test rejects less frequently than the conventional test.
-We would generally expect the visual test to only reject when the conventional test does.
-Surprisingly, one lineup containing a heteroskedasticity pattern does not follow this rule.
-"}
-
-library(ggmosaic)
-
-# Mosaic plot
-p_value_cmp_dat %>%
-  ggplot() +
-  geom_mosaic(aes(x = ggmosaic::product(reject, conv_reject), 
-                  fill = reject)) +
-  facet_grid(~type) +
-  ylab("Visual tests") +
-  xlab("Conventional tests") +
-  labs(fill = "Conventional tests") +
-  scale_fill_brewer("", palette = "Dark2") +
-  theme_bw() +
-  theme(legend.position = "none") +
-  coord_fixed()
-```
-
-```{r heter-example, fig.height = 7, fig.width = 7, fig.cap = 'The single heteroskedasticity lineup that is rejected by the visual test but not by the BP test.
-The data plot at panel 17 contains a "butterfly" shape.
-It has effect size $ = 3.76$, somewhat surprising that it is not detected by the BP test.'}
-# plot heter-331
-VI_MODEL$plot_lineup(vi_lineup$heter_331$data, 
-                     remove_grid_line = TRUE, 
-                     theme = theme_light(),
-                     remove_axis = TRUE)
-```
-
-## Effect of shape of non-linearity
-
-A primary factor contributes to the non-linearity model is the shape of the non-linearity pattern.
-According to Figure \ref{fig:poly-power-uniform-j}, conventional tests have higher power in testing shapes constructed from lower orders of Hermite polynomials.
-But the chance of detecting the "triple-U" shape drops significantly compared to other shapes.
-To understand why this is, one needs to return to the way the RESET test is applied.
-It requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four [@ramsey_tests_1969].
-However, the "triple-U" shape constructed from the Hermite polynomials use power up to 18.
-If the RESET test had been applied using a higher power no less than six, the power curve of "triple-U" shape will be only slight lower than the power curve of "M" shape.
-The recommendation of the polynomial power for the RESET test should be revised, perhaps.
-This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data generating process in order to set the parameters for the test.
-
-For visual tests, based on the orders of the Hermite polynomials, we expect the "U" shape will be the easiest one to be detected by subjects followed by the "S", "M" and "triple-U" shape.
-From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves are mostly aligned with our expectation, except for the "M" shape, which is as easy to be detected as the "S" shape.
-This implies, unlike the conventional test, the visibility of the shape do not strictly follow the degree of the polynomials.
-
-```{r poly-power-uniform-j, fig.height = 4, fig.cap = 'Power of conventional tests and visual tests on lineups containing four different non-linearity shapes.
-Power curves with higher order of non-linearity are drawn with deeper colours.
-The default RESET tests under-perform significantly in detecting "triple-U" shape.
-To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test.
-But this means the order needs to be known prior to testing.'}
-
-dat_poly_j <- poly_conv_sim %>%
-  filter(x_dist == "uniform") %>%
-  mutate(reject = as.numeric(RESET4_p_value <= 0.05)) %>%
-  mutate(shape = c("U", "S", "M", "triple-U")[shape]) %>%
-  mutate(conv_or_not = "Conventional") %>%
-  bind_rows(vi_survey %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(x_dist == "uniform", type == "polynomial") %>%
-  mutate(reject = as.numeric(p_value <= 0.05)) %>%
-  mutate(shape = c("U", "S", "M", "triple-U")[shape]) %>%
-  mutate(conv_or_not = "Visual")) %>%
-  mutate(shape = factor(shape, levels = c("U", "S", "M", "triple-U"))) %>%
-  mutate(offset0 = log(0.05/0.95)) %>%
-  select(shape, effect_size, offset0, reject, conv_or_not)
-
-dat_poly_j %>%
-  nest(dat = c(effect_size, offset0, reject)) %>%
-  mutate(mod = map(dat, 
-                   ~glm(reject ~ effect_size - 1, 
-                        family = binomial(), 
-                        data = .x,
-                        offset = offset0))) %>%
-  mutate(power = map(mod, function(mod) {
-    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
-                         offset0 = log(0.05/0.95))
-    result$power <- predict(mod, type = "response", newdata = result)
-    result
-  })) %>%
-  select(-dat, -mod) %>%
-  unnest(power) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  ggplot() +
-  geom_line(aes(log_effect_size, power, col = shape), size = 1) +
-  xlab(expression(log[e] (Effect_size))) +
-  ylab("Power") +
-  scale_color_manual(values = rcartocolor::carto_pal(4, "TealGrn")) +
-  theme_light() +
-  theme(legend.position = "bottom") +
-  facet_wrap(~conv_or_not) +
-  labs(col = "Non-linearity shape")
-```
-
-## Effect of shape of heteroskedasticity
-
-We have also investigated the impact of different heteroskedasticity shapes on power of conventional tests and visual tests.
-In theory, the "left-triangle" and the "right-triangle" shapes are functionally identical from the point of view of a Breusch-Pagan test.
-As shown in Figure \ref{fig:heter-power-uniform-a}, this is indeed the case where little difference between the power curves can be perceived.
-Similarly, visual tests should have the same power in detecting these two shapes if they are equally likely to be identified.
-However, it can be observed from Figure \ref{fig:heter-power-uniform-a} that the power curve of the "left-triangle" shape is higher than the one of the "right-triangle" shape, indicating a potential favour of orientation by human, which is worth to be explored in future studies.
-Besides, the chance of detecting the "butterfly" shape are higher than other two shapes in both conventional and visual testing.
-
-```{r heter-power-uniform-a, fig.height = 4, fig.cap = 'Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$.
-The visual test has an asymmetric power in testing the "left-triangle" shape and "right-triangle" shape, while these two shapes are equivalent in conventional testing.'}
-
-dat_heter_a <- heter_conv_sim %>%
-  filter(x_dist == "uniform") %>%
-  mutate(reject = as.numeric(BP_p_value <= 0.05)) %>%
-  mutate(a = c("left-triangle", "butterfly", "right-triangle")[a + 2]) %>%
-  mutate(conv_or_not = "Conventional") %>%
-  bind_rows(vi_survey %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(x_dist == "uniform", type == "heteroskedasticity") %>%
-  mutate(reject = as.numeric(p_value <= 0.05)) %>%
-  mutate(a = c("left-triangle", "butterfly", "right-triangle")[a + 2]) %>%
-  mutate(conv_or_not = "Visual")) %>%
-  mutate(a = factor(a, levels = c("left-triangle", "butterfly", "right-triangle"))) %>%
-  mutate(offset0 = log(0.05/0.95)) %>%
-  select(a, effect_size, offset0, reject, conv_or_not)
-
-# Manually fit the model
-dat_heter_a %>%
-  nest(dat = c(effect_size, offset0, reject)) %>%
-  mutate(mod = map(dat, 
-                   ~glm(reject ~ effect_size - 1, 
-                        family = binomial(), 
-                        data = .x,
-                        offset = offset0))) %>%
-  mutate(power = map(mod, function(mod) {
-    result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
-                         offset0 = log(0.05/0.95))
-    result$power <- predict(mod, type = "response", newdata = result)
-    result
-  })) %>%
-  select(-dat, -mod) %>%
-  unnest(power) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  ggplot() +
-  geom_line(aes(log_effect_size, power, col = a), size = 1) +
-  xlab(expression(log[e] (Effect_size))) +
-  ylab("Power") +
-  scale_color_manual(values = c(rcartocolor::carto_pal(3, "Earth")[1], "grey", rcartocolor::carto_pal(3, "Earth")[3])) +
-  theme_light() +
-  theme(legend.position = "bottom") +
-  facet_wrap(~conv_or_not) +
-  labs(col = "Heteroskedasticity shape")
-```
-
-## Effect of fitted value distributions
-
-The prediction in a regression model is $E(Y|X)$, that is, it is conditional on observed values of the predictors.
-The distribution of $X$, or consequently $\hat{Y}$, may however affect the ability to read any patterns in the residuals.
-The four distributions of fitted values, used in this experiment, were designed to examine this.
-
-Figure \ref{fig:different-x-dist-poly-power} illustrates the impact of fitted value distribution on the power of conventional tests and visual tests.
-For conventional tests, only the power curves of appropriate tests which are RESET tests and Breusch-Pagan tests for non-linearity pattern and heteroskedasticity pattern respectively are shown.
-For visual tests, note that we have collected more evaluations on lineup with uniform fitted value distribution, and the number of evaluations on a lineup will affect the power of the visual test.
-To have a fair comparison, for lineups with uniform fitted value distribution, we randomly sample five evaluations from the total eleven evaluations to estimate the power curves.
-The sampling process is repeated 500 times to produce an indication of the variation of the power.
-
-In terms of visual tests, we do not observe significant difference between the power curve of the uniform distribution and the power curve of the discrete uniform distribution.
-Normal and lognormal distributions produce lower power than uniform and discrete uniform distributions in regards of residual departure patterns, but the gap is larger when the residual plot contains heteroskedasticity pattern.
-This suggests that residual departures displayed with evenly distributed fitted values are visually more attractive and effective.
-Besides, the skewness of the fitted value distribution plays a role in the degree of presence of visual features as lognormal distribution produce the lowest power for both non-linearity and heteroskedasticity patterns.
-A symmetric fitted value distribution can better reveal the underlying shape.
-
-Although it is unusual to consider fitted value distribution or predictor distribution in power analysis, it is not so surprised that the power of conventional tests vary because of fitted value distribution.
-For instance, the power of the RESET test which is effectively a F test depends on the realizations of the additional predictors included in the auxiliary regression equation [@jamshidian2007study; @olvera2019relationship; @zhang2018practical].
-And those additional predictors are linear transformation of random vectors following certain predictor distributions.
-What is truly unexpected is the huge variability in power of conventional tests compared to visual tests.
-The discrete uniform distribution constantly produce the highest power while the normal distribution and the lognormal distribution produce the lowest power for non-linearity and heteroskedasticity pattern respectively.
-The difference in power corresponding to different distributions peak at around 90% when the effect size is at a moderate level.
-This huge difference is undesirable as is not uncommon for analysts to omit the factor of predictor distribution in power calculation and sample size calculation.
-
-```{r get-5-eval, cache = TRUE}
-dat_uniform_5 <- map_df(1:500, function(boot_id) {
-  vi_survey %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  filter(x_dist == "uniform") %>%
-  group_by(unique_lineup_id) %>%
-  slice_sample(n = 5) %>%
-  ungroup() %>%
-  calc_p_value_multi(lineup_id = unique_lineup_id,
-                     detect = detect,
-                     n_sel = num_selection,
-                     alpha = alpha) %>%
-  mutate(boot_id = boot_id)
-}) %>%
-  left_join(select(vi_survey, unique_lineup_id, effect_size, x_dist, type)) %>%
-  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  mutate(reject = as.numeric(p_value <= 0.05)) %>%
-  mutate(conv_or_not = "Visual") %>%
-  mutate(offset0 = log(0.05/0.95)) %>%
-  select(-unique_lineup_id, -p_value)
-
-dat_uniform_5_pred <- dat_uniform_5 %>%
-  nest(dat = c(effect_size, offset0, reject)) %>%
-  mutate(mod = map(dat, 
-                   ~glm(reject ~ effect_size - 1, 
-                        family = binomial(), 
-                        data = .x,
-                        offset = offset0))) %>%
-  mutate(power = map2(mod, type, function(mod, type) {
-    this_min_es <- ifelse(type == "heteroskedasticity", min_heter_es, min_poly_es)
-    this_max_es <- ifelse(type == "heteroskedasticity", max_heter_es, max_poly_es)
-    result <- data.frame(effect_size = seq(this_min_es[1], this_max_es[1], 0.1),
-                         offset0 = log(0.05/0.95))
-    result$power <- predict(mod, type = "response", newdata = result)
-    result
-  })) %>%
-  select(-dat, -mod) %>%
-  unnest(power) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  mutate(type = factor(stringr::str_to_title(type), 
-                       levels = c("Non-Linearity", "Heteroskedasticity")))
-```
-
-```{r different-x-dist-poly-power, fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for conventional tests and visual tests.
-Power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and Breusch-Pagan tests respectively.
-Power curves of visual tests are estimated using five evaluations on each lineup.
-For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations.
-The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines.
-The fitted value distribution has greater impact on the power of conventional tests than visual tests.
-Meanwhile, uneven distributions including normal and lognormal distributions genearlly produce lower power.", dev = 'png', dpi = 450}
-
-diff_x_dist_power_pred <- list(
-  # Visual inference data
-  vi_survey %>%
-    filter(!null_lineup) %>%
-    filter(!attention_check) %>%
-    group_by(unique_lineup_id) %>%
-    summarise(across(everything(), first)) %>%
-    filter(x_dist != "uniform") %>%
-    mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-    mutate(reject = as.numeric(p_value <= 0.05)) %>%
-    mutate(offset0 = log(0.05/0.95)) %>%
-    select(type, x_dist, effect_size, offset0, reject) %>%
-    mutate(conv_or_not = "Visual"),
-
-  # Conventional simulation (poly)
-  poly_conv_sim %>%
-    mutate(reject = RESET4_p_value < 0.05) %>%
-    mutate(offset0 = log(0.05/0.95)) %>%
-    mutate(type = "non-linearity") %>%
-    select(type, x_dist, effect_size, offset0, reject) %>%
-    mutate(conv_or_not = "Conventional"),
-
-  # Conventional simulation (heter)
-  heter_conv_sim %>%
-    mutate(reject = BP_p_value < 0.05) %>%
-    mutate(offset0 = log(0.05/0.95)) %>%
-    mutate(type = "heteroskedasticity") %>%
-    select(type, x_dist, effect_size, offset0, reject) %>%
-    mutate(conv_or_not = "Conventional")
-) %>%
-
-  # Bind rows
-  map_df(~.x) %>%
-
-  nest(dat = c(effect_size, offset0, reject)) %>%
-  mutate(mod = map(dat, 
-                   ~glm(reject ~ effect_size - 1, 
-                        family = binomial(), 
-                        data = .x,
-                        offset = offset0))) %>%
-  mutate(power = map2(mod, type, function(mod, type) {
-    this_min_es <- ifelse(type == "heteroskedasticity", min_heter_es, min_poly_es)
-    this_max_es <- ifelse(type == "heteroskedasticity", max_heter_es, max_poly_es)
-    result <- data.frame(effect_size = seq(this_min_es[1], this_max_es[1], 0.1),
-                         offset0 = log(0.05/0.95))
-    result$power <- predict(mod, type = "response", newdata = result)
-    result
-  })) %>%
-  select(-dat, -mod) %>%
-  unnest(power) %>%
-  mutate(log_effect_size = log(effect_size))
-
-diff_x_dist_power_pred %>%
-  mutate(x_dist = ifelse(x_dist == "even_discrete", 
-                         "Discrete uniform", 
-                         stringr::str_to_title(x_dist))) %>%
-  mutate(type = factor(stringr::str_to_title(type), 
-                       levels = c("Non-Linearity", 
-                                  "Heteroskedasticity"))) %>%
-  mutate(x_dist = factor(x_dist, 
-                         levels = c("Uniform", 
-                                    "Normal", 
-                                    "Lognormal", 
-                                    "Discrete uniform"))) %>%
-
-  ggplot() +
-
-  # Draw all smooth lines
-  geom_line(aes(log_effect_size, power, col = x_dist),
-            size = 1) +
-
-  # Draw boot uniform (5 eval) lines
-  geom_line(data = dat_uniform_5_pred,
-            aes(log_effect_size, power, col = "Uniform", group = boot_id),
-            size = 1,
-            alpha = 0.01) +
-
-  # Theme
-  theme_light() +
-  scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(4,1,3,2)]) +
-  xlab(expression(log[e] (Effect_size))) +
-  ylab("Power") +
-  labs(col = "", linetype = "Test") +
-  facet_grid(conv_or_not~type, scales = "free_x") +
-  theme(legend.position = "bottom")
-```
-
-# Conclusion
-
-Motivated by the advice of regression analysis experts, that residual plots as opposed to conventional tests are an indispensible methods for assessing model fit, a human subjects experiment was conducted using visual inference.
-The experiment tested two primary departures from good residuals: non-linearity and heteroskedasticity.
-
-The experiment found that conventional residual-based statistical tests are more sensitive to weak residual departures from model assumptions than visual tests as would be evaluated by humans.
-That is, conventional tests conclude there are problems with the model fit almost twice as often as humans would.
-They often reject when departures in the form of non-linearity and heteroskedasticity are not visible to a human.
-
-One might say that this is correct behaviour, but it can be argued that the conventional tests are rejecting when it is not necessary.
-Many of these rejections happen even when downstream analysis and results would not be significantly affected by the small departures from a good fit.
-The results from human evaluations provide a more practical solution, which reinforces the statements from regression experts that residual plots are an indispensible method for model diagnostics.
-
-Now it is important to note that residual plots need to be delivered as a lineup, where it is embedded in a field of null plots.
-A residual plot may contains many visual features, but some are caused by the characteristics of the predictors and the randomness of the error, not by the violation of the model assumptions.
-These irrelevant visual features have a chance to be filtered out by subjects with a comparison to null plots, results in a set of more accurate visual findings.
-This enables a careful calibration for reading structure in residual plots.
-
-However, human evaluation of residuals is expensive.
-It is time-consuming, laborious and unfriendly to vision-impaired people.
-This is another reason why it often appears to be ignored.
-With the availability of sophisticated computer vision algorithms today, the goal of this work is to form the basis of providing automated residual plot reading.
-The findings suggest the strong demand of graphical inspection in regression diagnostics, so developing an automatic visual inference system to evaluate lineups of residual plots is valuable.
-We plan to build a completed open-source system in an `R` package and provide a web interface such as a website for public to interact with.
-Details about this system will be discussed in our next paper.
-
-The experiment also revealed some interesting details about how residual plots are read.
-For the most part, the visual test performed very similarly to the appropriate conventional test only with the power curve shifted in the less sensitive direction.
-Unlike the conventional tests, where one needs to specifically test for non-linearity or heteroskedasticity the visual test operated effectively across the range of departures from good residuals.
-
-As expected, if the fitted value distribution is not uniform, there is a loss of power in the visual test.
-Structure is hardest to detect if fitted values are lognormal.
-Also, complex structure are generally harder to detect, but there are outliers.
-Under the designed scenarios in this paper, we find the visual test to be a more robust test against the change of fitted value distributions.
-A surprising finding was that the direction of heteroskedasticity appears to affect the ability to visually detect it, with wedge to the right being less detectable.
-
-# Acknowledgements {-}
-
-These `R` packages are used for the work: 
-`cli` [@cli], `curl` [@curl], `dplyr` [@dplyr], `ggplot2` [@ggplot2], `jsonlite` [@jsonlite], `lmtest` [@lmtest], `mpoly` [@mpoly], `progress` [@progress], `tibble` [@tibble], `ggmosaic` [@ggmosaic], `purrr` [@purrr], `tidyr` [@tidyr], `readr` [@readr], `stringr` [@stringr], `here` [@here], `kableExtra` [@kableextra], `patchwork` [@patchwork], `rcartocolor` [@rcartocolor].
-The study website is powered by `PythonAnywhere` [@pythonanywhere] and the `Flask` web framework [@flask].
-The `jsPsych` framework [@jspsych] is used to create behavioral experiments that run in our study website.
-
-The article was created with R packages `rticles` [@rticles], `knitr` [@knitr] and `rmarkdown` [@rmarkdown].
-The project's Github repository (link here) contains all materials required to reproduce this article.
-
-# Supplementary material {-}
-
-The supplementary material is available at (link here).
-It includes more details about the experimental setup, the derivation of the effect size, the effect of data collection period, and the estimate of  $\alpha$.
-
-\appendix
-
+  scale_color_brewer(palette = "Reds") +
+  facet_wrap(~shape)
+```
+
+\newpage
 # Experiment setup
@@
 
-# Demographics
+\newpage
+
+# Analysis of results relative to data collection process
+
+## Demographics
 
@@
 
-# Effect size derivation
-
-Effect size can be defined as the difference of a parameter for a particular model or distribution,  or a statistic derived from a sample.
-Importantly, it needs to reflect the treatment we try to measure.
-Centred on a conventional statistical test, we usually can deduce the effect size from the test statistic by substituting the null parameter value.
-When considering the diagnostics of residual departures, there exist many possibilities of test statistics for a variety of model assumptions.
-Meanwhile, diagnostic plots such as the residual plot have no general agreement on measuring how strong a model violation pattern is.
-To build a bridge between various residual-based tests, and the visual test, we focus on the shared information embedded in the testing procedures, which is the distribution of residuals.
-When comes to comparison of distribution, Kullback-Leibler divergence [@kullback1951information] is a classical way to represent the information loss or entropy increase caused by the approximation to the true distribution, which in our case, the inefficiency due to the use of false model assumptions.
-
-Following the terminology introduced by @kullback1951information, $P$ represents the measured probability distribution, and $Q$ represents the assumed probability distribution.
-The Kullback-Leibler divergence is defined as $\int_{-\infty}^{\infty}log(p(x)/q(x))p(x)dx$, where $p(.)$ and $q(.)$ denote probability densities of $P$ and $Q$.
-
-Let $\boldsymbol{X}_a = (\boldsymbol{1}, \boldsymbol{X})$ denotes the $p$ regressors with $n$ observations, $\boldsymbol{R}_a = \boldsymbol{I} -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ denotes the residual operator, and let $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0},\sigma^2\boldsymbol{I})$ denotes the error.
-Using the Frisch–Waugh–Lovell theorem, residuals $\boldsymbol{e} = \boldsymbol{R}_a\boldsymbol{\varepsilon}$.
-Because $rank(\boldsymbol{R}_a) = n - p < n$, $e$ follows a degenerate multivariate normal distribution and does not have a density.
-Since the Kullback-Leibler divergence requires a proper density function, we need to simplify the covariance matrix of $\boldsymbol{e}$ by setting all the off-diagonal elements to 0.
-Then, the residuals will assumed to follow $N(\boldsymbol{0}, diag(\boldsymbol{R}_a\sigma^2))$ under the null hypothesis that the model is correctly specified.
-If the model is however misspecified due to omitted variables $\boldsymbol{Z}$, or a non-constant variance $\boldsymbol{V}$, the distribution of residuals can be derived as $N(\boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}_a\sigma^2))$ and $N(\boldsymbol{0}, diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a'))$ respectively.
-
-By assuming both $P$ and $Q$ are multivariate normal density functions, the Kullback-Leibler divergence can be rewritten as 
-$$KL = \frac{1}{2}\left(log\frac{|\Sigma_p|}{|\Sigma_q|} - n + tr(\Sigma_p^{-1}\Sigma_q) + (\mu_p - \mu_q)'\Sigma_p^{-1}(\mu_p - \mu_q)\right).$$
-
-Then, we can combine the two residual departures into one formula
-
-\small
-
-\begin{equation}
-\label{eq:effect-size}
-KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}\boldsymbol{\mu}_z\right).
-\end{equation}
-
-\normalsize
-
-When there are omitted variables but constant error variance, the formula can be reduced to
-
-$$KL = \frac{1}{2}\left(\boldsymbol{\mu}_z^{T}(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right).$$
-
-And when the model equation is correctly specified but the error variance is non-constant, the formula can be reduced to
-
-$$KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2))\right).$$
-Since we assume $\sigma = 1$ for the heteroskedasticity model, the final form of the formula is
-
-$$KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right).$$
-
-# Effect of data collection period
+## Data collection periods
 
@@
 Thus, there is no clear evidence of batch effect.
-
-# Sensitivity analysis for $\alpha$
-
-The parameter $\alpha$ used for the $p$-value calculation needs to be estimated from responses to null lineups.
-With a greater value of $\hat{\alpha}$, the $p$-value will be smaller, resulting in more lineups being rejected.
-However, The way we generate Rorschach lineup is not strictly the same as what suggested in @vanderplas2021statistical and @buja_statistical_2009.
-Therefore, we conduct a sensitivity analysis in this section to examine the impact of the variation of the estimator $\alpha$ on our primary findings.
-
-The analysis is conducted by setting up several scenarios, where the $\alpha$ is under or overestimated by 12.5%, 25% and 50%.
-Using the adjusted $\hat{\alpha}$, we recalculate the $p$-value for every lineup and show the results in Figure \ref{fig:sensitivity}.
-It can be observed that there are some changes to $p$-values, especially when the $\hat{\alpha}$ is multiplied by 50%.
-However, Table \ref{tab:sensitivity} shows that adjusting $\hat{\alpha}$ will not result in a huge difference in rejection decisions.
-There are only a small percentage of cases where the rejection decision change.
-It is very unlikely the downstream findings will be affected because of the estimate of $\alpha$.
-
-```{r}
-upper_p <- map_df(c(0.125, 0.25, 0.5), function(modified) {
-  calc_p_value_multi(vi_survey %>%
-                       filter(!attention_check) %>%
-                       filter(!null_lineup) %>%
-                       filter(x_dist == "uniform") %>%
-                       mutate(alpha_upper = alpha * (1 + modified)),
-                     lineup_id = unique_lineup_id,
-                     detect = detect,
-                     n_sel = num_selection,
-                     alpha = alpha_upper) %>%
-  mutate(modified = 1 + modified) %>%
-  mutate(adjusted_p_value = p_value)
-})
-
-lower_p <- map_df(c(0.125, 0.25, 0.5), function(modified) {
-  calc_p_value_multi(vi_survey %>%
-                       filter(!attention_check) %>%
-                       filter(!null_lineup) %>%
-                       filter(x_dist == "uniform") %>%
-                       mutate(alpha_lower = alpha * (1 - modified)),
-                     lineup_id = unique_lineup_id,
-                     detect = detect,
-                     n_sel = num_selection,
-                     alpha = alpha_lower) %>%
-  mutate(modified = 1 - modified) %>%
-  mutate(adjusted_p_value = p_value)
-})
-
-upper_lower_p <- upper_p %>%
-  bind_rows(lower_p) %>%
-  select(-p_value)
-```
-
-```{r sensitivity, fig.cap = "Change of $p$-values with $\\hat{\\alpha}$ multiplied by $0.5$, $0.75$, $0.875$, $1.125$, $1.25$ and $1.5$.
-The vertical dashed line is to indicate $p\\text{-value} = 0.05$.
-The x-axis is drawn on logarithmic scale.
-For multiplier smaller than $1$, the adjusted $p$-value will incrase then decrase as $p$-value increases.
-The trend is the opposite for multiplier greater than $1$, but the difference will eventually reach $0$."}
-upper_lower_p %>%
-  left_join(vi_survey %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(x_dist == "uniform") %>%
-  group_by(unique_lineup_id) %>%
-  summarise(p_value = first(p_value))) %>%
-  mutate(diff_p_value = adjusted_p_value - p_value) %>%
-  ggplot() +
-  geom_hline(yintercept = 0) +
-  geom_vline(xintercept = 0.05, col = "red", linetype = 2) +
-  geom_point(aes(p_value, diff_p_value), alpha = 0.2) +
-  scale_x_log10() +
-  facet_wrap(~modified) +
-  ylab(expression("Adjusted"~italic(p)*"-value minus"~italic(p)*"-value")) +
-  xlab(expression(italic(p)*"-value")) +
-  theme_light()
-
-```
-
-```{r}
-upper_lower_p %>%
-  left_join(vi_survey %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(x_dist == "uniform") %>%
-  group_by(unique_lineup_id) %>%
-  summarise(p_value = first(p_value))) %>%
-  group_by(modified) %>%
-  count(adjusted_reject = adjusted_p_value <= 0.05, reject = p_value <= 0.05) %>%
-  complete(adjusted_reject, reject, fill = list(n = 0)) %>%
-  ungroup() %>%
-  filter(adjusted_reject != reject) %>%
-  mutate(case = ifelse(!adjusted_reject, 'Proportion of lineups transforms to "not reject"', 'Proportion of lineups transforms to "reject"')) %>%
-  rename(multiplier = modified) %>%
-  mutate(n = paste0(round(n/279 * 100, 2), "\\%")) %>%
-  select(-adjusted_reject, -reject) %>%
-  pivot_wider(names_from = case, values_from = n) %>%
-  kableExtra::kable(format = "latex", booktabs = TRUE, label = "sensitivity", 
-                    escape = FALSE,
-                    caption = "Change of rejection decision because of the modification of $\\hat{\\alpha}$.")
-```
-
-# Effect of number of evaluations on the power of a visual test
-
-When comparing power of visual tests across different fitted value distributions, we have discussed the number of evaluations on a lineup will affect the power of the visual test.
-Using the lineups with uniform fitted value distribution, we show in Figure \ref{fig:n_eval} the change of power of visual tests due to different number of evaluations.
-It can be learned that as the number of evaluations increases, the power will increase but the margin will decrease.
-Considering we have eleven evaluations on lineups with uniform fitted value distribution, and five evaluations on other lineups, it is necessary to use the same number of evaluations for each lineup in comparison.
-
-```{r}
-min_poly_es <- vi_survey %>% 
-  filter(!null_lineup, 
-         !attention_check,
-         type == "polynomial", 
-         x_dist == "uniform") %>%
-  pull(effect_size) %>%
-  min()
-max_poly_es <- vi_survey %>% 
-  filter(!null_lineup, 
-         !attention_check,
-         type == "polynomial", 
-         x_dist == "uniform") %>%
-  pull(effect_size) %>%
-  max()
-```
-
-```{r n_eval, fig.cap = "Change of power of visual tests for different number of evalutions on lineups with uniform fitted value distribution.
-The power will incrase as the number of evaluations increases, but the margin will decrase."}
-eval_n <- function(n_eval, boot) {
-  map_df(1:boot, function(boot_id) {
-  vi_survey %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  filter(x_dist == "uniform") %>%
-  filter(type == "polynomial") %>%
-  group_by(unique_lineup_id) %>%
-  slice_sample(n = n_eval) %>%
-  ungroup() %>%
-  calc_p_value_multi(lineup_id = unique_lineup_id,
-                     detect = detect,
-                     n_sel = num_selection,
-                     alpha = alpha) %>%
-  mutate(boot_id = boot_id)
-}) %>%
-  left_join(select(vi_survey, unique_lineup_id, effect_size)) %>%
-  mutate(reject = as.numeric(p_value <= 0.05)) %>%
-  mutate(offset0 = log(0.05/0.95)) %>%
-  select(-unique_lineup_id, -p_value) %>%
-  mutate(n_eval = n_eval)
-}
-
-map_df(seq(1, 11, 2), ~eval_n(.x, 1)) %>%
-  select(-boot_id) %>%
-  nest(c(effect_size, reject, offset0)) %>%
-  mutate(mod = map(data, 
-                   ~glm(reject ~ effect_size - 1, 
-                        family = binomial(), 
-                        data = .x,
-                        offset = offset0))) %>%
-  mutate(power = map(mod, function(mod) {
-    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
-                         offset0 = log(0.05/0.95))
-    result$power <- predict(mod, type = "response", newdata = result)
-    result
-  })) %>%
-  select(-data, -mod) %>%
-  unnest(power) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  ggplot() +
-  geom_line(aes(log_effect_size, power, col = factor(n_eval)),
-            size = 1) +
-  labs(col = "Number of evalutaions") +
-  xlab(expression(log[e] (Effect_size))) +
-  theme_light() +
-  scale_color_brewer(palette = "Reds")
-```
-
-# Power of a RESET test under different auxiliary regression formulas
-
-It is found in the result that the power of a RESET test will be affected by the highest order of fitted values included in the auxiliary formula.
-And we suspect that the current recommendation of the highest order - four, is insufficient to test complex non-linear structures such as the "Triple-U" shape designed in this paper.
-Figure \ref{fig:reset} illustrates the change of power of RESET test while testing the "U" shape and the "Triple-U" shape with different highest orders.
-Clearly, when testing a simple shape like the "U" shape, the highest order has very little impact on the power.
-But for testing the "Triple-U" shape, there will be a loss of power if the recommended order is used.
-To avoid the loss of power, the highest order needs to be set to at least six.
-
-```{r cache = TRUE}
-reset_diff_dat <- poly_conv_sim %>%
-  filter(x_dist == "uniform") %>%
-  filter(shape == 4) %>%
-  select(-F_p_value, -BP_p_value, -SW_p_value) %>%
-  pivot_longer(RESET3_p_value:RESET10_p_value) %>%
-  mutate(name = gsub("RESET(.*)_p_value", "\\1", name)) %>%
-  mutate(reject = value <= 0.05) %>%
-  select(effect_size, name, reject) %>%
-  mutate(offset0 = log(0.05/0.95)) %>%
-  nest(dat = c(effect_size, offset0, reject)) %>%
-  mutate(mod = map(dat, 
-                   ~glm(reject ~ effect_size - 1, 
-                        family = binomial(), 
-                        data = .x,
-                        offset = offset0))) %>%
-  mutate(power = map(mod, function(mod) {
-    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
-                         offset0 = log(0.05/0.95))
-    result$power <- predict(mod, type = "response", newdata = result)
-    result
-  })) %>%
-  select(-dat, -mod) %>%
-  unnest(power) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  mutate(shape = 4)
-```
-
-```{r cache = TRUE}
-reset_diff_dat2 <- poly_conv_sim %>%
-  filter(x_dist == "uniform") %>%
-  filter(shape == 1) %>%
-  select(-F_p_value, -BP_p_value, -SW_p_value) %>%
-  pivot_longer(RESET3_p_value:RESET10_p_value) %>%
-  mutate(name = gsub("RESET(.*)_p_value", "\\1", name)) %>%
-  mutate(reject = value <= 0.05) %>%
-  select(effect_size, name, reject) %>%
-  mutate(offset0 = log(0.05/0.95)) %>%
-  nest(dat = c(effect_size, offset0, reject)) %>%
-  mutate(mod = map(dat, 
-                   ~glm(reject ~ effect_size - 1, 
-                        family = binomial(), 
-                        data = .x,
-                        offset = offset0))) %>%
-  mutate(power = map(mod, function(mod) {
-    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
-                         offset0 = log(0.05/0.95))
-    result$power <- predict(mod, type = "response", newdata = result)
-    result
-  })) %>%
-  select(-dat, -mod) %>%
-  unnest(power) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  mutate(shape = 1)
-```
-
-```{r reset, fig.cap = 'Change of power of RESET tests for different orders of fitted values included in the auxiliary formula.
-The left panel is the power of testing the "U" shape and the right panel is the power of testing the "Triple-U" shape.
-The power will not be greatly affected by the highest order in the case of testing the "U" shape.
-In the case of testing the "Triple-U" shape, the highest order needs to be set to at least six to avoid the loss of power.', fig.height = 4}
-reset_diff_dat %>%
-  bind_rows(reset_diff_dat2) %>%
-  mutate(shape = factor(ifelse(shape == 1, "U", "Triple-U"), levels = c("U", "Triple-U"))) %>%
-  ggplot() +
-  geom_line(aes(log_effect_size, power, col = as.factor(as.integer(name))),
-            size = 1) +
-  theme_light() +
-  labs(col = "Highest order") +
-  xlab(expression(log[e] (Effect_size))) +
-  scale_color_brewer(palette = "Reds") +
-  facet_wrap(~shape)
-```
-
-\clearpage

### 2023-04-03 appendix.tex
rearranged main paper and appendix into separate files again

--- 2023-04-03 paper.tex
@@
 
-\articletype{ARTICLE TEMPLATE}
-
-\title{A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics
-with the Lineup Protocol}
+\articletype{}
+
+\title{Appendix: A Plot is Worth a Thousand Tests: Assessing Residual
+Diagnostics with the Lineup Protocol}
 
@@
 
-\begin{abstract}
-Abstract to fill.
-\end{abstract}
-
-\begin{keywords}
-statistical graphics; data visualization; visual inference; hypothesis
-testing; reression analysis; cognitive perception; simulation; practical
-significance; effect size
-\end{keywords}
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-\begin{quote}
-\emph{``Since all models are wrong the scientist must be alert to what
-is importantly wrong.''} \citep{box1976science}
-\end{quote}
-
-Diagnostics are the key to determining whether there is anything
-importantly wrong with a model. In linear regression analysis, residuals
-from the model fit are commonly used. Residuals summarise what is not
-captured by the model, and thus provide the capacity to identify what
-might be wrong.
-
-We can assess residuals in multiple ways. Residuals may be plotted as a
-histogram or quantile-quantile plot to examine the distribution. Using
-the classical normal linear regression model as an example, if the
-distribution is symmetric and unimodal, we consider it to be well
-behaved. However, if the distribution is skewed, bimodal, multimodal, or
-contains outliers, there is a cause for concern. One could also inspect
-the distribution by conducting a goodness of fit test, such as the
-Shapiro-Wilk Normality test \citep{shapiro1965analysis}.
-
-More typically, residuals will be plotted, as a scatter plot against the
-predicted values and each of the explanatory variables to scrutinize
-their relationships. If there are any visually discoverable patterns,
-the model is potentially misspecified. In general, one looks for
-noticeable departures from the model like non-linear dependency or
-heteroskedasticity. However, correctly judging a residual plot where no
-pattern exists can be a painstakingly difficult task for humans. It is
-especially common, particularly among students, to misinterpret patterns
-that are random noise and random deviation from a model
-\citep{loy2021bringing}. It is also possible to conduct hypothesis tests
-for non-linear dependence \citep{ramsey_tests_1969}, and use a
-Breusch-Pagan test \citep{breusch_simple_1979} for heteroskedasticity.
-
-Abundance of literature describe appropriate diagnostic methods for
-linear regression, e.g. \citet{draper1998applied},
-\citet{montgomery1982introduction}, \citet{belsley_regression_1980},
-\citet{cook_applied_1999} and \citet{cook1982residuals}. All these
-writings consider plotting residuals to be a standard technique that
-should be examined routinely in all regression modelling problems. In
-addition, \citet{draper1998applied} and \citet{belsley_regression_1980}
-believe that residual plots are usually revealing when the assumptions
-are violated. \citet{cook_applied_1999} thinks formal tests and
-graphical procedures are complementary and both have a place in residual
-analysis, but they focus on graphical methods rather than on formal
-testing, as they are easier to use. \citet{montgomery1982introduction}
-even suggests that residual plots are more informative in most practical
-situations than the corresponding formal tests, and statistical tests on
-regression model residuals are not widely used based on their
-experience.
-
-A common guidance by experts is that optimal method for diagnosing model
-fits is by plotting the data. The persistence of this advice to check
-the plots is curious, and investigating why this might be common advice
-is the subject of this paper. The paper is structured as follows. The
-next section describes the background on the types of departures that
-one expects to detect, and outlines a formal statistical process for
-reading residual plots, called visual inference. Section
-\ref{experimental-design} details the experimental design to compare the
-decisions made by formal hypothesis testing, and how humans would read
-diagnostic plots. The results are reported in Section \ref{results}. We
-conclude with a discussion of future work, in particular, how the
-responsibility for residual plot reading might be relieved.
-
-\hypertarget{background}{%
-\section{Background}\label{background}}
-
-\hypertarget{departures-from-good-residual-plots}{%
-\subsection{Departures from good residual
-plots}\label{departures-from-good-residual-plots}}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/residual-plot-common-departures-1}
-
-}
-
-\caption{Example residual vs fitted value plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
-The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted value plot.}\label{fig:residual-plot-common-departures}
-\end{figure}
-
-Graphical summaries in which residuals are plotted against fitted values
-or other functions of the predictors that are approximately orthogonal
-to the residuals are referred to as standard residual plots in
-\citet{cook1982residuals}. Figure
-\ref{fig:residual-plot-common-departures}A shows an ideal residual plot
-where the residuals are evenly distributed at both sides of the
-horizontal zero line, with no noticeable patterns.
-
-There are various types of departures from an ideal residual plot.
-Non-linearity, heteroskedasticity and non-normality are perhaps the
-three mostly checked departures.
-
-Non-linearity is a type of model misspecification caused by failing to
-include higher order terms of the predictors in the regression equation.
-Any non-linear functional form of residuals on fitted values in the
-residual plot could be indicative of non-linearity. An example residual
-plot containing visual pattern of non-linearity is shown in Figure
-\ref{fig:residual-plot-common-departures}B. One can clearly observe the
-``S-shape'' from the residual plot as the cubic term is not captured by
-the misspecified model.
-
-Heteroskedasticity refers to the presence of nonconstant error variance
-in a regression model. It is mostly due to the strict but false
-assumptions on the variance-covariance matrix of the error term. The
-usual pattern of heteroskedasticity on a residual plot is the
-inconsistent spread of the residuals across the horizontal axis.
-Visually, it sometimes results in the so-called ``butterfly'' shape as
-shown in Figure \ref{fig:residual-plot-common-departures}C, or the
-``left-triangle'' and ``right-triangle'' shape where the smallest
-variance occurs at one side of the horizontal axis.
-
-Compared to non-linearity and heteroskedasticity, non-normality is
-usually harder to detect from a residual plot since a scatter plot do
-not readily reveal the marginal distribution. A favourable graphical
-summary for this task is the quantile-quantile plot. As we mainly
-discuss residual plots, non-normality will not be the focus of this
-paper. For a consistent comparison, the residual plot of this departure
-is still presented in Figure \ref{fig:residual-plot-common-departures}D.
-When the number residuals below and above the horizontal axis are uneven
-across the local regions along the \(x\)-axis, we expect that the
-normality assumption is violated. For example, given a skewed error
-distribution, there will be fewer data points and more outliers on one
-side of the horizontal axis as shown in Figure
-\ref{fig:residual-plot-common-departures}D.
-
-\hypertarget{conventionally-testing-for-departures}{%
-\subsection{Conventionally testing for
-departures}\label{conventionally-testing-for-departures}}
-
-Other than checking diagnostic plots, analysts may perform formal
-hypothesis testing to detect model defects. A variety of tests can be
-applied; each testing for a specific violation of the null hypothesis.
-For example, the presence of heteroskedasticity can usually be tested by
-applying the White test \citep{white_heteroskedasticity-consistent_1980}
-or the Breusch-Pagan test \citep{breusch_simple_1979}, which are both
-derived from the Lagrange multiplier test \citep{silvey1959lagrangian}
-principle that relies on the asymptotic properties of the null
-distribution. To test specific forms of non-linearity, one may apply the
-F-test as a model structural test to examine the significance of
-specific polynomial and non-linear forms of the predictors, or the
-significance of proxy variables as in the Ramsey Regression Equation
-Specification Error Test (RESET) \citep{ramsey_tests_1969}. The
-Shapiro-Wilk test \citep{shapiro1965analysis} is the most widely used
-test of non-normality included by many of the statistical software
-programs. The Jarque--Bera test \citep{jarque1980efficient} is also used
-to directly check whether the sample skewness and kurtosis match a
-normal distribution.
-
-We apply the RESET test, Breusch-Pagan test and Shapiro-Wilk test to the
-residual plots shown in Figure
-\ref{fig:residual-plot-common-departures}; the results of this is shown
-in Table \ref{tab:example-residual-plot-table}. The Breusch-Pagan test
-and the Shapiro-Wilk test both reject \(H_0\) for residual plots that
-display non-linearity and heteroscedasticity, even though these are not
-the original intention of the test. As discussed in
-\citet{cook1982residuals}, most residual-based tests for a particular
-type of departure from model assumptions are also sensitive to other
-types of departures. It is likely \(H_0\) is correctly rejected but for
-the wrong reason, a phenomenon known as the ``Type III error''.
-Additionally, outliers will often incorrectly trigger the rejection of
-\(H_0\) despite when majority of the residuals are well-behaved
-\citep{cook_applied_1999}. Furthermore, with a sufficiently large sample
-size, residual-based tests may reject \(H_0\) due to a slight departure
-that is of little practical significance. These can be largely avoided
-in diagnostic plots as experienced analysts can evaluate the
-acceptability of assumptions flexibly, even in the presence of outliers
-and slight departures.
-
-\begin{table}
-
-\caption{\label{tab:example-residual-plot-table}Statistical significance testing for departures from good residuals for plots in Figure \ref{fig:residual-plot-common-departures}.
-Shown are the $p$-values calculated for the RESET, the Breusch-Pagan and the Shapiro–Wilk tests.
-The good residual plot (A) is judged a good residual plot, as expected, by all tests.
-The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.}
-\centering
-\begin{tabular}[t]{llrrr}
-\toprule
-Plot & Departures & RESET & Breusch-Pagan & Shapiro–Wilk\\
-\midrule
-A & None & 0.779 & 0.133 & 0.728\\
-B & Non-linearity & \em{0.000} & \em{0.000} & \em{0.039}\\
-C & Heteroskedasticity & 0.658 & \em{0.000} & \em{0.000}\\
-D & Non-normality & 0.863 & 0.736 & \em{0.000}\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\hypertarget{visual-test-procedure-based-on-lineups}{%
-\subsection{Visual test procedure based on
-lineups}\label{visual-test-procedure-based-on-lineups}}
-
-One may argue that reading diagnostic plots is to some extent subjective
-and indecisive compared to those rigorous statistical procedures as it
-relies on graphical perception - human ability to interpret and decode
-the information embedded in graph \citep{cleveland_graphical_1984}.
-Further, the degree of the presence of the visual features typically can
-not be measured quantitatively and objectively, which may lead to over
-or under-interpretations of the data. For instance, people
-over-interpret the separation between gene groups in a two-dimensional
-projection from a linear discriminant analysis when in fact there are no
-differences in the expression levels between the gene groups and
-separation is not an uncommon occurrence
-\citep{roy_chowdhury_using_2015}.
-
-Visual inference was first introduced in a 1999 Joint Statistical
-Meetings (JSM) talk with the title ``Inference for Data Visualization''
-by \citet{buja_inference_1999} as an idea to address the issue of valid
-inference for visual discoveries of data plots. Later,
-\citet{buja_statistical_2009} proposed the lineup protocol as a visual
-test inspired by the ``police lineup'' or ``identity parade'' which is
-the act of asking the eyewitness to identify criminal suspect from a
-group of irrelevant people. The protocol consists of \(m\) randomly
-placed plots, where one plot is the data plot, and the remaining
-\(m - 1\) null plots have the identical graphical procedure except the
-data has been replaced with data consistent with \(H_0\), that the model
-is correctly specified. Then, an observer who have not seen the data
-plot will be asked to point out the most different plot from the lineup.
-Under \(H_0\), it is expected that the data plot would have no
-distinguishable difference from the null plots, and the probability that
-the observer correctly picks the data plot is \(1/m\). If one rejects
-\(H_0\) as the observer correctly picks the data plot, then the Type I
-error of this test is \(1/m\).
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/first-example-lineup-1}
-
-}
-
-\caption{Visual testing is conducted using a lineup, as in the example here.
-The residual plot computed from the observed data (plot $2^2 + 2$, exhibiting non-linearity) is embedded among 19 null plots, where the residuals are simulated from a standard error model.
-Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot.
-A small $p$-value would result from a substantial number selecting plot $2^2 + 2$.}\label{fig:first-example-lineup}
-\end{figure}
-
-Figure \ref{fig:first-example-lineup} is an example of a lineup
-protocol. If the data plot at position \(2^2 + 2\) is identifiable, then
-it is evidence for the rejection of \(H_0\). In fact, the actual
-residual plot is obtained from a misspecified regression model with
-missing non-linear terms.
-
-Data used in the \(m - 1\) null plots needs to be simulated. In
-regression diagnostics, sampling data consistent with \(H_0\) is
-equivalent to sampling data from the assumed model. As
-\citet{buja_statistical_2009} suggested, \(H_0\) is usually a composite
-hypothesis controlled by nuisance parameters. Since regression models
-can have various forms, there is no general solution to this problem,
-but it sometimes can be reduced to a so called ``reference
-distribution'' by applying one of the three methods: (i) sampling from a
-conditional distribution given a minimal sufficient statistic under
-\(H_0\), (ii) parametric bootstrap sampling with nuisance parameters
-estimated under \(H_0\), and (iii) Bayesian posterior predictive
-sampling. The conditional distribution given a minimal sufficient
-statistic is the best justified reference distribution among the three
-\citep{buja_statistical_2009}. Essentially, null residuals can be
-simulated by regressing \(N\) i.i.d standard normal random draws on the
-predictors, then rescaling it by the ratio of residual sum of square in
-two regressions.
-
-The effectiveness of lineup protocol for regression analysis is
-validated by \citet{majumder_validation_2013} under relatively simple
-settings with up to two predictors. Their results suggest that visual
-tests are capable of testing the significance of a single predictor with
-a similar power as a t-test, though they express that in general it is
-unnecessary to use visual inference if there exists a conventional test,
-and they do not expect the visual test to perform equally well as the
-conventional test. In their third experiment, where there is not a
-conventional test, visual test outperforms the conventional test for a
-large margin. This is encouraging, as it promotes the use of visual
-inference in situations where there are no existing statistical testing
-procedures. Visual inference have also been integrated into diagnostic
-of hierarchical linear models by \citet{loy2013diagnostic},
-\citet{loy2014hlmdiag} and \citet{loy2015you}. They use lineup protocols
-to judge the assumption of linearity, normality and constant error
-variance for both the level-1 and level-2 residuals.
-
-\hypertarget{calculation-of-statistical-significance-and-test-power}{%
-\section{Calculation of statistical significance and test
-power}\label{calculation-of-statistical-significance-and-test-power}}
-
-\hypertarget{what-is-being-tested}{%
-\subsection{What is being tested?}\label{what-is-being-tested}}
-
-In diagnosing a model fit from residuals, we are generally interested in
-\(H_0:\) \emph{The regression model is correctly specified} against the
-broad alternative \(H_a:\) \emph{The regression model is misspecified}.
-
-However, it is practically impossible to test this specific \(H_0\) with
-conventional tests, which are constructed to measure specific
-departures. For example, the RESET test is formulated as
-\(H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0\) against
-\(H_a: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0\),
-from
-\(y = \tau_0 + \sum_{i=1}^{p}\tau_px_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)\).
-Similarly, the Breusch-Pagan test is designed to specifically test
-\(H_0:\) \emph{error variances are all equal}
-(\(\zeta_i=0 \text{ for } i=1,..,p\)) versus the alternative \(H_a:\)
-\emph{that the error variances are a multiplicative function of one or
-more variables} (\(\text{at least one } \zeta_i\neq 0\)) from
-\(e^2 = \zeta_0 + \sum_{i=1}^{p}\zeta_i x_i + u, ~ u\sim N(0,\sigma_u^2)\).
-
-One of the potential benefits of the visual test, based on the lineup
-protocol, is that it should be able to detect a range of departures from
-good residuals.
-
-\hypertarget{statistical-significance}{%
-\subsection{Statistical significance}\label{statistical-significance}}
-
-In hypothesis testing, a \(p\)-value is defined as the probability of
-observing test results as least as extreme as the observed result given
-\(H_0\) is true. Conventional hypothesis tests usually have an existing
-method to derive or compute \(p\)-value based on the null distribution.
-What we need to discuss in the following is the method to estimate
-\(p\)-value for a visual test.
-
-Within the context of visual inference, by involving \(k\) independent
-observers, the visual \(p\)-value can be interpreted as the probability
-of having as many or more subjects detect the data plot than the
-observed result.
-
-Let \(X_j = \{0,1\}\) be a Bernoulli random variable denoting whether
-subject \(j\) correctly detecting the data plot, and
-\(X = \sum_{j=1}^{K}X_j\) be the number of observers correctly picking
-the data plot. Then, by imposing a relatively strong assumption on the
-visual test that all \(K\) evaluations are fully independent, under
-\(H_0\), \(X \sim \mathrm{Binom}_{K,1/m}\). Therefore, the \(p\)-value
-of a lineup of size \(m\) evaluated by \(K\) observer is given as
-\(P(X \geq x) = 1 - F(x) + f(x)\), where \(F(.)\) is the binomial
-cumulative distribution function, \(f(.)\) is the binomial probability
-mass function and \(x\) is the realization of number of observers
-correctly picking the data plot \citep{majumder_validation_2013}.
-
-As pointed out by \citet{vanderplas2021statistical}, this basic binomial
-model does not take into account the possible dependencies in the visual
-test due to repeated evaluations of the same lineup. And it is
-inapplicable to visual test where subjects are asked to select one or
-more ``most different'' plots from the lineup.
-\citet{vanderplas2021statistical} summarises three common scenarios in
-visual inference: (1) \(K\) different lineups are shown to \(K\)
-subjects, (2) \(K\) lineups with different null plots but the same data
-plot are shown to \(K\) subjects, and (3) the same lineup is shown to
-\(K\) subjects. Out of these three scenarios, Scenario 3 is the most
-common in previous studies as it puts the least constraints on the
-experiment design. For Scenario 3, \citet{vanderplas2021statistical}
-models the probability of a plot \(i\) being selected from a lineup as
-\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
-\(i=1,...,m\) and \(\alpha > 0\). The number of times plot \(i\) being
-selected in \(K\) evaluations is denoted as \(c_i\). In case subject
-\(j\) makes multiple selections, \(1/s_j\) will be added to \(c_i\)
-instead of one, where \(s_j\) is the number of plots subject \(j\)
-selected for \(j=1,...K\). This ensures \(\sum_{i}c_i=K\). Since we are
-only interested in the selections of the data plot \(i\), the marginal
-model can be simplified to a beta-binomial model and thus the visual
-\(p\)-value is given as
-
-\begin{equation} \label{eq:pvalue-beta-binomial}
-P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+
-\end{equation}
-
-\noindent where \(B(.)\) is the beta function defined as
-
-\begin{equation} \label{eq:betafunction}
-B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0.
-\end{equation}
-
-Note that Equation \ref{eq:pvalue-beta-binomial} given in
-\citet{vanderplas2021statistical} only works with non-negative integer
-\(c_i\). We extend the equation to non-negative real number \(c_i\) by
-applying a linear approximation
-
-\begin{equation} \label{eq:pvalue-beta-binomial-approx}
-P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+,
-\end{equation}
-
-where \(P(C \geq \lceil c_i \rceil)\) is calculated using Equation
-\ref{eq:pvalue-beta-binomial} and \(P(C = \lfloor c_i \rfloor)\) is
-calculated by
-
-\begin{equation} \label{eq:pmf-beta-binomial}
-P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+.
-\end{equation}
-
-Besides, the parameter \(\alpha\) used in Equation
-\ref{eq:pvalue-beta-binomial} and \ref{eq:pmf-beta-binomial} is usually
-unknown and hence needs to be estimated from the survey data. For low
-values of \(\alpha\), only a few plots are attractive to the observers
-and tend to be selected. For higher values of \(\alpha\), the
-distribution of the probability of each plot being selected is more
-even. \citet{vanderplas2021statistical} defines that a plot is
-\(c\)-interesting if \(c\) or more participants select the plot as the
-most different. Given the definition, The expected number of plots
-selected at least \(c\) times, \(E[Z_c]\), is calculated as
-
-\begin{equation} \label{eq:c-interesting-expectation}
-E[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}
-
-With Equation \ref{eq:c-interesting-expectation}, \(\alpha\) can be
-estimated using maximum likelihood estimation. But for precise estimate
-of \(\alpha\), additional human responses to Rorschach lineups, which is
-a type of lineup that consists of plots constructed from the same null
-data generating mechanism, are required.
-
-\hypertarget{power-of-the-tests}{%
-\subsection{Power of the tests}\label{power-of-the-tests}}
-
-The power of a model misspecification test is the probability that
-\(H_0\) is rejected given the regression model is misspecified in a
-specific way. It is an important indicator when one is concerned about
-whether model assumptions have been violated. Although in practice, one
-might be more interested in knowing how much the residuals deviate from
-the model assumptions, and whether this deviation is of practical
-significance.
-
-The power of a conventional hypothesis test is affected by both the true
-parameter \(\boldsymbol{\theta}\) and the sample size \(n\). These two
-can be quantified in terms of effect size \(E\) to measure the strength
-of the residual departures from the model assumptions. Details about the
-effect size is provided in Section \ref{effect-size} after the
-introduction of the simulation model used in our human subject
-experiment. The theoretical power of a test is sometimes not a trival
-solution, but it can be estimated if the data generating process is
-known. We use a predefined model to generate a large set of simulated
-data under different effect sizes, and record if the conventional test
-rejects \(H_0\). The probability of the conventional test rejects
-\(H_0\) is then fitted by a logistic regression formulated as
-
-\begin{equation} \label{eq:logistic-regression-1-1}
-Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),
-\end{equation}
-
-\noindent where \(\Lambda(.)\) is the standard logistic function given
-as \(\Lambda(z) = exp(z)/(1+exp(z))\). The effect size \(E\) is the only
-predictor and the intercept is fixed to \(log(0.05/0.95)\) so that
-\(\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05\), which is the desired
-significance level.
-
-The power of a visual test on the other hand, may additionally depend on
-the ability of the particular subject, as the skill of the individual
-may affect the number of observers who identify the data plot from the
-lineup \citep{majumder_validation_2013}. To address this issue,
-\citet{majumder_validation_2013} models the probability of a subject
-\(j\) correctly picking the data plot from a lineup \(l\) using a
-mixed-effect logistic regression, with subjects treated as random
-effect. Then, the estimated power of a visual test evaluated by a single
-subject is the predicted value obtained from the mix-effect model.
-However, this mix-effect model does not work with scenario where
-subjects are asked to select one or more most different plots. In this
-scenario, having the probability of a subject \(j\) correctly picking
-the data plot from a lineup \(l\) is insufficient to determine the power
-of a visual test because it does not provide information about the
-number of selections made by the subject for the calculation of the
-\(p\)-value (See Equation \ref{eq:pvalue-beta-binomial-approx}).
-Therefore, we directly estimate the probability of a lineup being
-rejected by assuming that individual skill has negligible effect on the
-variation of the power. This assumption is not necessary true, but it
-helps simplifying the model structure, thereby obviate a costly
-large-scale experiment to estimate complex covariance matrices. The same
-model given in Equation \ref{eq:logistic-regression-1-1} is applied to
-model the power of a visual test.
-
-To study various factors contributing to the power of both tests, the
-same logistic regression model is fit on different subsets of the
-collated data grouped by levels of factors. These include the
-distribution of the fitted values, type of the simulation model and the
-shape of the residual departures.
-
-\hypertarget{experimental-design}{%
-\section{Experimental design}\label{experimental-design}}
-
-An experiment is conducted in three data collection periods to
-investigate the difference between conventional hypothesis testing and
-visual inference in the application of linear regression diagnostics.
-Two types of departures, non-linearity and heteroskedasticity, are
-collected during data collection periods I and II. The data collection
-period III was designed primarily to measure human responses to null
-lineups so that the parameter \(\alpha\) in Equation
-\ref{eq:pvalue-beta-binomial} can be estimated. Additional lineups for
-both non-linearity and heteroskedasticity, using uniform fitted value
-distribution, were included so that the participants were evaluating
-some lineups with signal also. It would be too frustrating for
-participants to only be assigned lineups with all null plots. Overall,
-we collected 7974 evaluations on 1152 unique lineups performed by 443
-subjects throughout three data collection periods.
-
-\hypertarget{simulating-departures-from-good-residuals}{%
-\subsection{Simulating departures from good
-residuals}\label{simulating-departures-from-good-residuals}}
-
-\hypertarget{non-linearity}{%
-\subsubsection{Non-linearity}\label{non-linearity}}
-
-Data collection period I is designed to study the ability of human
-subjects to detect the effect of a non-linear term \(\boldsymbol{z}\)
-constructed using Hermite polynomials on random vector
-\(\boldsymbol{x}\) formulated as
-
-\begin{align} \label{eq:nonlinearity-model}
-\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
-\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
-\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
-\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
-\end{align}
-
-\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
-\(\boldsymbol{\varepsilon}\), \(\boldsymbol{x}_{raw}\),
-\(\boldsymbol{z}_{raw}\) are vectors of size \(n\), \(He_{j}(.)\) is the
-\(j\)th-order probabilist's Hermite polynomials,
-\(\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\), and
-\(g(\boldsymbol{x}, k)\) is a scaling function to enforce the support of
-the random vector to be \([-k, k]^n\) defined as
-
-\begin{equation} \label{eq:scaling-function}
-g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x}))2k - k, \quad \text{for} \quad k > 0. 
-\end{equation}
-
-According to \citet{abramowitz1964handbook}, Hermite polynomials were
-initially defined by \citet{de1820theorie}, but named after Hermite
-\citep{hermite1864nouveau} because of the unrecognisable form of
-Laplace's work. When simulating \(\boldsymbol{z}_{raw}\), function
-\texttt{hermite} from the R package \texttt{mpoly} \citep{mpoly} is used
-to generate Hermite polynomials.
-
-The null regression model used to fit the realizations generated by the
-above model is formulated as
-
-\begin{equation} \label{eq:null-model}
-\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
-\end{equation}
-
-\noindent where
-\(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\).
-
-Since \(z = O(x^j)\), for \(j > 1\), \(z\) is a higher order term leaves
-out by the null regression, which will lead to model misspecification.
-
-Visual patterns of non-linearity are simulated using four different
-orders of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)). (A
-summary of the factors is given in Table \ref{tab:model-factor-table}.)
-The values of \(j\) is chosen so that distinct shapes of non-linearity
-are included in the residual plot. These include ``U'', ``S'', ``M'' and
-``triple-U'' shape as shown in Figure
-\ref{fig:different-shape-of-herimite}. A greater value of \(j\) will
-result in a curve with more turning points. It is expected that the
-``U'' shape will be the easiest one to detect because complex shape
-tends to be concealed by cluster of data points.
-
-Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups
-used in non-linearity detection. This lineup is produced by the
-non-linearity model with \(j = 6\). The data plot location is
-\(2^3 - 4\). All five subjects correctly identify the data plot from
-this lineup.
-
-\begin{table}
-
-\caption{\label{tab:model-factor-table}Levels of the factors used in data collection periods I, II, III.}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{rr|rr|rcrr|rr|rcrr|rr|rcrr|rr|rcrr|rr|rcrr|rr|rc}
-\toprule
-\multicolumn{2}{c}{Non-linearity} & \multicolumn{2}{c}{Heteroskedasticity} & \multicolumn{2}{c}{Common} \\
-\cmidrule(l{3pt}r{3pt}){1-2} \cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6}
-\multicolumn{1}{c}{Poly Order ($j$)} & \multicolumn{1}{c}{SD ($\sigma$)} & \multicolumn{1}{c}{Shape ($a$)} & \multicolumn{1}{c}{Ratio ($b$)} & \multicolumn{1}{c}{Size ($n$)} & \multicolumn{1}{c}{Distribution of fitted values} \\
-\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
-2 & 0.25 & -1 & 0.25 & 50 & Uniform\\
-3 & 1.00 & 0 & 1.00 & 100 & Normal\\
-6 & 2.00 & 1 & 4.00 & 300 & Skewed\\
-18 & 4.00 &  & 16.00 &  & Discrete uniform\\
- &  &  & 64.00 &  & \\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-shape-of-herimite-1}
-
-}
-
-\caption{Polynomial forms generated for the residual plots used to assess detecting non-linearity.
-The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$.}\label{fig:different-shape-of-herimite}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-sigma-1}
-
-}
-
-\caption{Examining the effect of $\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform fitted value distribution and the "U" shape.
-As $\sigma$ increases the signal strength decreases, to the point that the "U" is almost unrecognisable when $\sigma=4$.}\label{fig:different-sigma}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/example-poly-lineup-1}
-
-}
-
-\caption{One of the lineups containing non-linearity patterns used in data collection period I.
-Can you spot the most different plot?
-The data plot is positioned at $2^3 + 1$.}\label{fig:example-poly-lineup}
-\end{figure}
-
-\hypertarget{heteroskedasticity}{%
-\subsubsection{Heteroskedasticity}\label{heteroskedasticity}}
-
-Data collection period II is designed to study the ability of human
-subjects to detect the appearance of a heteroskedasticity pattern under
-a simple linear regression model setting:
-
-\begin{align} \label{eq:heter-model}
-\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
-\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
-\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}), 
-\end{align}
-
-\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
-\(\boldsymbol{\varepsilon}\) are vectors of size \(n\) and \(g(.)\) is
-the scaling function defined in Equation \ref{eq:scaling-function}.
-
-The null regression model used to fit the realizations generated by the
-above model is formulated exactly the same as Equation
-\ref{eq:null-model}.
-
-For \(b \neq 0\), the variance-covariance matrix of the error term
-\(\boldsymbol{\varepsilon}\) is correlated with the predictor
-\(\boldsymbol{x}\), which will lead to the presence of
-heteroskedasticity. Visual patterns of heteroskedasticity are simulated
-using three different shapes (\(a\) = -1, 0, 1). (A summary of the
-factors can be found in Table \ref{tab:model-factor-table}.)
-
-Since \(supp(X) = [-1, 1]\), choosing \(a\) to be \(-1\), \(0\) and
-\(1\) can generate ``left-triangle'', ``butterfly'' and
-``right-triangle'' shape as displayed in Figure
-\ref{fig:different-shape-of-heter}. The term \((2 - |a|)\) maintains the
-magnitude of residuals across different values of \(a\).
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-shape-of-heter-1}
-
-}
-
-\caption{Heteroskedasticity forms used in the experiment.
-Three different shapes ($a = -1, 0, 1$) are used in the experiment to create left-triangle, "butterfly" and "right-triangle" shapes, respectively.}\label{fig:different-shape-of-heter}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-b-1}
-
-}
-
-\caption{Five different values of $b$ are used in heteroskedasticity simulation to control the strength of the signal.
-Larger values of $b$ yield a bigger difference in variation, and stus stronger heteroskedasticity signal.}\label{fig:different-b}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/example-heter-lineup-1}
-
-}
-
-\caption{One of the lineups containing heteroskedasticity pattern used in data collection period II.
-Can you spot the most different plot?
-The data plot is positioned at $3^3 - 3^2$}\label{fig:example-heter-lineup}
-\end{figure}
-
-An example lineup of this model used in data collection period II is
-shown in Figure \ref{fig:example-heter-lineup} with \(a = -1\). The data
-plot location is \(2^4 + 2\). Nine out of 11 subjects correctly identify
-the data plot from this lineup.
-
-\hypertarget{factors-common-to-both-data-collection-periods}{%
-\subsubsection{Factors common to both data collection
-periods}\label{factors-common-to-both-data-collection-periods}}
-
-Fitted values are a function of the independent variables, and the
-distribution of the observed values affects the distribution of the
-fitted values. In the best case scenario the fitted values will have a
-uniform distribution, which means that there is even coverage of
-possible observed values across all of the predictors. This is not
-always present in the collected data. Sometimes the fitted values are
-discrete because one or more predictors were measured discretely. The
-distribution may be relatively Gaussian, reflecting a linear combinatio
-of many predictors, adhering to the Central Limit Theorem. It is also
-common to see a skewed distribution of fitted values, if one or more of
-the predictors has a skewed distribution. This latter problem is usually
-corrected before modelling using a variable transformation. Our
-simulation assess this by using four different distributions to
-represent fitted values: (1) uniform, (2) normal, (3) skewed and (4)
-discrete uniform. This is constructed by defining the raw predictor
-\(X_{raw}\) in four corresponding distributions: (1) \(U(-1, 1)\), (2)
-\(N(0, 0.3^2)\), (3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). We
-would expect that the best reading of residual plots occurs when the
-fitted values are uniformly distributed.
-
-Three different sample sizes are used, \(n=50, 100, 300\) across the
-experiments. We would expect considerable variation in the signal
-strength in the simulated data plots with smaller \(n\). A sample size
-of 300 is typically enough for structure to be visible in a scatter plot
-reliably.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-dist-1}
-
-}
-
-\caption{Variations in fitted values, that might affect perception of residual plots. Four different distributions are used.}\label{fig:different-dist}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-n-1}
-
-}
-
-\caption{Examining the effect of signal strength for the three different values of $n$ used in the experiment, for non-linear structure with fixed $\sigma = 1.5$, uniform fitted value distribution, and "S" shape.
-For these factor levels, only when $n = 300$ is the "S" shape clearly visible.}\label{fig:different-n}
-\end{figure}
-
-\hypertarget{experimental-setup}{%
-\subsection{Experimental setup}\label{experimental-setup}}
-
-\hypertarget{controlling-the-strength-of-the-signal}{%
-\subsubsection{Controlling the strength of the
-signal}\label{controlling-the-strength-of-the-signal}}
-
-As summarised in Table \ref{tab:model-factor-table}, three additional
-parameters \(n\), \(\sigma\) and \(b\) are used to control the strength
-of the signal so that different difficulty levels of lineups are
-generated, and therefore, the estimated power curve will be smooth and
-continuous. Parameter \(\sigma \in \{0.5, 1, 2, 4\}\) and
-\(b \in \{0.25, 1, 4, 16, 64\}\) are used in data collection periods I
-and II respectively. Figure \ref{fig:different-sigma} and
-\ref{fig:different-b} demonstrate the impact of these two parameters. A
-large value of \(\sigma\) will increase the variation of the error of
-the non-linearity model and decrease the visibility of the visual
-pattern. The parameter \(b\) controls the standard deviation of the
-error across the support of the predictor. Given \(x \neq a\), a larger
-value of \(b\) will lead to a larger ratio of the variance at \(x\) to
-the variance at \(x - a = 0\), making the visual pattern more obvious.
-
-Three different sample sizes are used (n = 50, 100, 300) in all three
-data collection periods. It can be observed from Figure
-\ref{fig:different-n} that with fewer data points drawn in a residual
-plot, the visual pattern is more difficult to be detected.
-
-\hypertarget{effect-size}{%
-\subsubsection{Effect size}\label{effect-size}}
-
-Effect size in statistics measures the strength of the signal relative
-to the noise. It is surprisingly difficult to quantify in general, even
-for simulated data as used in this experiment.
-
-For the non-linearity model, the key items defining effect size are
-sample size (\(n\)) and variance of the error term (\(\sigma^2\)), and
-so effect size would be roughly calculated as \(\sqrt{n}/{\sigma}\). As
-sample size increases the effect size would increase, but as variance
-increases the effect size decreases. However, it is not clear how the
-additional parameter for the model polynomial order, \(k\), should be
-incorporated. Intuitively, the large \(k\) means more complex pattern,
-which likely means effect size would decrease. For the purposes of our
-calculations we have chosen to use an approach based on Kullback-Leibler
-divergence \citep{kullback1951information}, coupled with simulation.
-This formulation defines effect size to be:
-
-\[E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right)\]
-\noindent where \(diag(.)\) is the diagonal matrix constructed from the
-diagonal elements of a matrix,
-\(\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a\) is the
-residual operator,
-\(\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'\)
-is the hat matrix,
-\(\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z\)
-is the expected values of residuals with \(\boldsymbol{Z}\) be any
-higher order terms of \(\boldsymbol{X}\) leave out by the regression
-equation and \(\boldsymbol{\beta}_z\) be the corresponding coefficients,
-and \(\sigma^2\boldsymbol{I}\) is the assumed covariance matrix of the
-error term when \(H_0\) is true.
-
-In the heteroskedasticity model, the key elements for measuring effect
-size are sample size, \(n\), and the ratio of the biggest variance to
-smallest variance, \(b\). Larger values of both would produce higher
-effect size. However, it is not clear how to incorporate the additional
-shape parameter, \(a\). Thus the same approach is used here, where the
-formula can be written as:
-
-\[E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)\]
-\noindent where \(\boldsymbol{V}\) is the actual covariance matrix of
-the error term.
-
-Derivations for these equations are provided in the Appendix.
-
-To compute the effect size for each lineup we simulate a sufficient
-large number of samples from the same model, in each sample, the number
-of observations \(n\) is fixed. We then compute the effect size for each
-sample and take the average as the final value. This ensures lineups
-constructed with the same experimental factors will share the same
-effect size.
-
-\hypertarget{subject-allocation}{%
-\subsubsection{Subject allocation}\label{subject-allocation}}
-
-As shown in Table \ref{tab:model-factor-table}, there are a total of
-\(4 \times 4 \times 3 \times 4 = 192\) and
-\(3 \times 5 \times 3 \times 4 = 180\) number of combinations of
-parameter values for non-linearity model and heteroskedasticity model
-respectively. Three replications are made for each of the combination
-results in \(192 \times 3 = 576\) and \(180 \times 3 = 540\) lineups. In
-addition, each lineup is designed to be evaluated by five different
-subjects. After attempting some pilot studies internally, we decide to
-present a block of 20 lineups to every subject. And to ensure the
-quality of the survey data, two lineups with obvious visual patterns are
-included as attention checks. Thus, \(576 \times 5 / (20-2) = 160\) and
-\(540 \times 5 / (20-2) = 150\) subjects are recruited to satisfy the
-design of the data collection period I and II respectively.
-
-As mentioned in Section \ref{power-of-the-tests}, \(\alpha\) used in
-Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using null
-lineups. Three replications are made for \(3 \times 4 = 12\)
-combinations of common factors \(n\) and fitted value distribution,
-results in \(12 \times 3 = 36\) lineups included in data collection
-period III. In these lineups, the data of the data plot is generated
-from a model with zero effect size, while the data of the 19 null plots
-are generated using the same simulation method discussed in Section
-\ref{visual-test-procedure-based-on-lineups}. This generation procedure
-differs from the canonical Rorschach lineup procedure, which requires
-that all 20 plots are generated directly from the null model. However,
-these lineups serve the same fundamental purpose: to assess the number
-of visually interesting plots generated under \(H_0\).
-
-To account for the fact that our simulation method for these lineups is
-not the Rorschach procedure, we use the method suggested in
-\citet{vanderplas2021statistical} for typical lineups containing a data
-plot to estimate \(\alpha\). We have included a sensitivity analysis in
-the Appendix to examine the impact of the variance of the \(\alpha\)
-estimate on our findings.
-
-All lineups consist of only null plots are planned to be evaluated by 20
-subjects. However, presenting only these lineups to subjects are
-considered to be bad practices as subjects will lose interest quickly.
-Therefore, we plan to collect 6 more evaluations on the 279 lineups with
-uniform fitted value distribution, result in
-\((36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133\) subjects
-recruited for data collection period III.
-
-\hypertarget{collecting-results}{%
-\subsubsection{Collecting results}\label{collecting-results}}
-
-Subjects for all three data collection periods are recruited from an
-crowdsourcing platform called Prolific \citep{palan2018prolific}.
-Prescreening procedure is applied during the recruitment, subjects are
-required to be fluent in English, with \(98\%\) minimum approval rate
-and 10 minimum submissions in other studies.
-
-During the experiment, every subject is presented with a block of 20
-lineups. A lineup consists of a randomly placed data plot and 19 null
-plots, which are all residual plots drawn with raw residuals on the
-y-axis and fitted values on the x-axis. An additional horizontal red
-line is added at \(y = 0\) as a helping line.
-
-The data of the data plot is simulated from one of two models described
-in Section \ref{simulating-departures-from-good-residuals}, while the
-data of the remaining 19 null plots are generated by the residual
-rotation technique discussed in Section
-\ref{visual-test-procedure-based-on-lineups}.
-
-In every lineup evaluation, the subject is asked to select one or more
-plots that are most different from others, provide a reason for their
-selections, and evaluate how different they think the selected plots are
-from others. If there is no noticeable difference between plots in a
-lineup, subjects are permitted to select zero plots without providing
-the reason. No subject are shown the same lineup twice. Information
-about preferred pronoun, age group, education, and previous experience
-in visual experiments are also collected. A subject's submission is only
-accepted if the data plot is identified for at least one attention
-check. Data of rejected submissions are discarded automatically to
-maintain the overall data quality.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{overview}{%
-\subsection{Overview}\label{overview}}
-
-There are 2880, 2700 and 1674 lineups evaluation made by 160, 150 and
-133 subjects recruited for data collection periods I, II and III
-respectively. In the total of 7974 lineup evaluations, 3744 use lineups
-produced by the non-linearity model, and 4230 use lineups produced by
-the heteroskedasticity model. Besides, there are 886 attention checks
-and 720 evaluations on null lineups needed for the estimate of
-\(\alpha\) not included in the analysis. The collated dataset is
-provided in \texttt{vi\_survey} of the \texttt{visage} \texttt{R}
-package.
-
-In the following analysis, lineups with uniform fitted values will be
-the focus. Visual patterns are more likely to be revealed under a
-uniform distribution. Additionally, we have collected extra evaluations
-on these lineups, which will result in a more reliable analysis.
-Analysis of lineups with other fitted value distributions can be found
-in Section \ref{effect-of-fitted-value-distributions}.
-
-\hypertarget{power-comparison-of-different-tests}{%
-\subsection{Power comparison of different
-tests}\label{power-comparison-of-different-tests}}
-
-Figure \ref{fig:polypower} shows the estimated power of visual test on
-lineups produced by the non-linearity model with uniform fitted values,
-against the natural logarithm of the effect \(log_e(E)\), with a 5\%
-significance level. At the bottom of the figure \ref{fig:polypower},
-there are a sequence of example residual plots with increasing levels of
-\(log_e(E)\). Readers can evaluate them from left to right and determine
-at which level the departure from a good residual plot becomes
-detectable.
-
-As discussed in Section \ref{conventionally-testing-for-departures},
-many conventionally tests are available for detecting residual
-departures. Implementation-wise, the built-in R package \texttt{stats}
-provides some commonly used residual-based tests, such as Shapiro-Wilk
-test. A more comprehensive collection of regression diagnostics tests
-can be found in the R package \texttt{lmtest} \citep{lmtest}. In terms
-of heteroskedasticity diagnostics, the R package \texttt{skedastic}
-\citep{skedastic} collects and implements 25 existing conventional tests
-published since 1961.
-
-We pick RESET test (\texttt{resettest}) and Breusch-Pagan test
-(\texttt{bptest}) from the R package \texttt{lmtest}, and Shapiro-Wilk
-test (\texttt{shapiro.test}) from the built-in R package \texttt{stats}.
-Among them, RESET test is the only exact and appropriate test in this
-scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are
-approximate and inappropriate tests. Their estimated power is shown in
-Figure \ref{fig:polypower}. To set up the RESET test, we include
-different powers of fitted values as proxies. According to
-\citet{ramsey_tests_1969}, there are no general rules for the power of
-the fitted values needed by the RESET test, but it finds power up to
-four is usually sufficient. Thus, we follow this guideline to conduct
-the RESET test. For the Breusch-Pagan test, the choice of predictors in
-the auxiliary regression is left to the user
-\citep{breusch_simple_1979}. But as \citet{waldman1983note} suggested,
-it is a good choice for the set of auxiliary predictors in the
-Breusch-Pagan test be the same as the White test. Thus, we include both
-\(x\) and \(x^2\) in the auxiliary regression.
-
-Figure \ref{fig:heterpower} is similar to Figure \ref{fig:polypower},
-but shows corresponding information on lineups produced by the
-heteroskedasticity model. In this scenario, the visual test is compared
-to an approximate test - Breusch-Pagan test, and two other inappropriate
-tests - RESET test and Shapiro-Wilk test.
-
-For non-linearity patterns, the power curve of RESET test climbs
-aggressively from 7\% to 84\% as \(log_e(E)\) increases from 0 to 2.5,
-while power of other tests respond inactively to the change of effect,
-showing that RESET test is much more sensitive to weak non-linear
-structure. Meanwhile, no noticeable residual departures can be spotted
-from the example residual plots.
-
-In terms of heteroskedasticity patterns, the power of Breusch-Pagan test
-is constantly greater than the power of visual test. At
-\(log_e(E) \approx 2.5\), where the power curve of the visual test
-remains at a low level, the Breusch-Pagan test has around 50\% chance of
-rejecting \(H_0\). Similarly, the visual feature is nearly unobservable
-from the example residual plots at this level of effect size.
-
-The power of visual test arises steadily as \(log_e(E)\) increases from
-3 to 5 for both non-linearity patterns and heteroskedasticity patterns,
-suggesting that the effect starts to make significant impact on the
-degree of the presence of the designed visual features. This can also be
-observed from the example residual plots that when \(log_e(E) = 3.5\), a
-weak ``S-shape'' and a weak ``right-triangle'' shape are presented in
-Figure \ref{fig:polypower} and Figure \ref{fig:heterpower} respectively.
-The visual pattern becomes clearer as \(log_e(E)\) increases. At
-\(log_e(E) = 5\), the power of visual tests for both patterns reaches
-almost 100\%.
-
-Power curves of inappropriate tests show improvement as the effect
-increases but at a lower rate than the visual test in both scenarios.
-This coincides the point made by \citet{cook1982residuals} that
-residual-based tests for a specific type of model defect may be
-sensitive to other types of model defects. The power curve of RESET test
-remains at around 5\% in Figure \ref{fig:heterpower} since there are no
-non-linear terms leave out in the heteroskedasticity model and \(H_0\)
-of the test is always satisfied.
-
-Overall, the power comparison suggests that conventional tests differs
-significantly from visual tests in two regression diagnostics scenarios
-designed by us. Visual test have much higher tolerance of the residual
-departures than the conventional test. Since fail to reject \(H_0\) in a
-visual test literally means that there are no obvious visual discoveries
-found in the residual plot, analysts and the general public as the
-consumers of the output may not be convinced of the existence of
-significant residual departures in spite of the rejection of \(H_0\)
-given by the conventional test. Even if the rejection is accepted, the
-model violation may be considered as impactless due to the fact that
-they are not clearly visible. Besides, the sensitivity of the
-conventional test to weak residual departures could also distract and
-discourage analysts from finding simple but good linear approximation to
-the data. The rejection of \(H_0\) because of human acceptable and
-negligible residual departures is not practically meaningful and useful
-to analysts and decision makers. In contrast, if the strict correctness
-of the model assumption is of particular interest, conducting a
-conventional test is still the recommended choice based on the findings.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/polypower-1}
-
-}
-
-\caption{Comparison of power between different tests for non-linear patterns (uniform fitted values only).
-Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups.
-Surrounding lines of the visual test show the estimated power of 500 bootstrap samples.
-Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by vertical lines in main plot.
-Where would you draw the line of too much non-linearity in the residuals?
-For the RESET test this is around log effect size 2, but for the visual test it is around 3.5.}\label{fig:polypower}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/heterpower-1}
-
-}
-
-\caption{Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only).
-Main plot shows the power curves, with dots indicating human evaluations of lineups.
-Surrounding lines of the visual test show the estimated power of 500 bootstrap samples.
-Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by vertical lines in main plot.
-Where would you draw the line of too much heteroskedasticity in the residuals?
-For the BP test this is around log effect size 2.5, but for the visual test it is around 3.5.}\label{fig:heterpower}
-\end{figure}
-
-\hypertarget{comparison-of-test-decisions-based-on-p-values}{%
-\subsection{\texorpdfstring{Comparison of test decisions based on
-\(p\)-values}{Comparison of test decisions based on p-values}}\label{comparison-of-test-decisions-based-on-p-values}}
-
-The power comparison illustrates that appropriate conventional tests
-will reject \(H_0\) more aggressively than visual tests. In this
-section, we explore how often they agree with each other by
-investigating the rejections for the two model designs based on
-\(p\)-values for each lineup.
-
-Figure \ref{fig:p-value-comparison} provides a mosaic plot showing the
-rejection rate of visual tests and conventional tests for both
-non-linearity patterns and heteroskedasticity patterns.
-
-For lineups containing non-linearity patterns, conventional tests reject
-69\% and visual tests reject 32\% of the time. Of the lineups rejected
-by the conventional test, 46\% are rejected by the visual test, that is,
-approximately half as many as the conventional test. There are no
-lineups that are rejected by the visual test but not by the conventional
-test.
-
-In terms of lineups containing heteroskedasticity patterns, 76\% are
-rejected by conventional tests, while 56\% are rejected by visual tests.
-When the conventional test rejects a lineup, there is a great chance
-(73\%) the visual test will also reject it.
-
-Surprisingly, the visual test rejects 1 of the 33 (3\%) of lineups where
-the conventional test does not reject. Figure \ref{fig:heter-example}
-shows this lineup. The data plot in position seventeen displays a
-relatively strong heteroskedasticity pattern, and has a strong effect
-size (\(log_e(E)=4.02\)). This is reflected by the visual test
-\(p\text{-value} = 0.026\). But the Breusch-Pagan test
-\(p\text{-value} = 0.056\), is slightly above the significance cutoff of
-\(0.05\). This lineup was evaluated by 11 subjects, it has experimental
-factors \(a = 0\) (``butterfly'' shape), \(b = 64\) (large variance
-ratio), \(n = 50\) (small sample size), and a uniform distribution for
-the fitted values. It must be the small sample size that may have
-resulted in the lack of detection.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/p-value-comparison-1}
-
-}
-
-\caption{Rejection rate ($p$-value $\leq0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
-The visual test rejects less frequently than the conventional test.
-We would generally expect the visual test to only reject when the conventional test does.
-Surprisingly, one lineup containing a heteroskedasticity pattern does not follow this rule.
-}\label{fig:p-value-comparison}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/heter-example-1}
-
-}
-
-\caption{The single heteroskedasticity lineup that is rejected by the visual test but not by the BP test.
-The data plot at panel 17 contains a "butterfly" shape.
-It has effect size $ = 3.76$, somewhat surprising that it is not detected by the BP test.}\label{fig:heter-example}
-\end{figure}
-
-\hypertarget{effect-of-shape-of-non-linearity}{%
-\subsection{Effect of shape of
-non-linearity}\label{effect-of-shape-of-non-linearity}}
-
-A primary factor contributes to the non-linearity model is the shape of
-the non-linearity pattern. According to Figure
-\ref{fig:poly-power-uniform-j}, conventional tests have higher power in
-testing shapes constructed from lower orders of Hermite polynomials. But
-the chance of detecting the ``triple-U'' shape drops significantly
-compared to other shapes. To understand why this is, one needs to return
-to the way the RESET test is applied. It requires a parameter indicating
-degree of fitted values to test for, and the recommendation is to
-generically use four \citep{ramsey_tests_1969}. However, the
-``triple-U'' shape constructed from the Hermite polynomials use power up
-to 18. If the RESET test had been applied using a higher power no less
-than six, the power curve of ``triple-U'' shape will be only slight
-lower than the power curve of ``M'' shape. The recommendation of the
-polynomial power for the RESET test should be revised, perhaps. This
-illustrates the sensitivity of conventional testing to the parameters,
-and it also points to a limitation that one needs to know the data
-generating process in order to set the parameters for the test.
-
-For visual tests, based on the orders of the Hermite polynomials, we
-expect the ``U'' shape will be the easiest one to be detected by
-subjects followed by the ``S'', ``M'' and ``triple-U'' shape. From
-Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power
-curves are mostly aligned with our expectation, except for the ``M''
-shape, which is as easy to be detected as the ``S'' shape. This implies,
-unlike the conventional test, the visibility of the shape do not
-strictly follow the degree of the polynomials.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/poly-power-uniform-j-1}
-
-}
-
-\caption{Power of conventional tests and visual tests on lineups containing four different non-linearity shapes.
-Power curves with higher order of non-linearity are drawn with deeper colours.
-The default RESET tests under-perform significantly in detecting "triple-U" shape.
-To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test.
-But this means the order needs to be known prior to testing.}\label{fig:poly-power-uniform-j}
-\end{figure}
-
-\hypertarget{effect-of-shape-of-heteroskedasticity}{%
-\subsection{Effect of shape of
-heteroskedasticity}\label{effect-of-shape-of-heteroskedasticity}}
-
-We have also investigated the impact of different heteroskedasticity
-shapes on power of conventional tests and visual tests. In theory, the
-``left-triangle'' and the ``right-triangle'' shapes are functionally
-identical from the point of view of a Breusch-Pagan test. As shown in
-Figure \ref{fig:heter-power-uniform-a}, this is indeed the case where
-little difference between the power curves can be perceived. Similarly,
-visual tests should have the same power in detecting these two shapes if
-they are equally likely to be identified. However, it can be observed
-from Figure \ref{fig:heter-power-uniform-a} that the power curve of the
-``left-triangle'' shape is higher than the one of the ``right-triangle''
-shape, indicating a potential favour of orientation by human, which is
-worth to be explored in future studies. Besides, the chance of detecting
-the ``butterfly'' shape are higher than other two shapes in both
-conventional and visual testing.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/heter-power-uniform-a-1}
-
-}
-
-\caption{Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$.
-The visual test has an asymmetric power in testing the "left-triangle" shape and "right-triangle" shape, while these two shapes are equivalent in conventional testing.}\label{fig:heter-power-uniform-a}
-\end{figure}
-
-\hypertarget{effect-of-fitted-value-distributions}{%
-\subsection{Effect of fitted value
-distributions}\label{effect-of-fitted-value-distributions}}
-
-The prediction in a regression model is \(E(Y|X)\), that is, it is
-conditional on observed values of the predictors. The distribution of
-\(X\), or consequently \(\hat{Y}\), may however affect the ability to
-read any patterns in the residuals. The four distributions of fitted
-values, used in this experiment, were designed to examine this.
-
-Figure \ref{fig:different-x-dist-poly-power} illustrates the impact of
-fitted value distribution on the power of conventional tests and visual
-tests. For conventional tests, only the power curves of appropriate
-tests which are RESET tests and Breusch-Pagan tests for non-linearity
-pattern and heteroskedasticity pattern respectively are shown. For
-visual tests, note that we have collected more evaluations on lineup
-with uniform fitted value distribution, and the number of evaluations on
-a lineup will affect the power of the visual test. To have a fair
-comparison, for lineups with uniform fitted value distribution, we
-randomly sample five evaluations from the total eleven evaluations to
-estimate the power curves. The sampling process is repeated 500 times to
-produce an indication of the variation of the power.
-
-In terms of visual tests, we do not observe significant difference
-between the power curve of the uniform distribution and the power curve
-of the discrete uniform distribution. Normal and lognormal distributions
-produce lower power than uniform and discrete uniform distributions in
-regards of residual departure patterns, but the gap is larger when the
-residual plot contains heteroskedasticity pattern. This suggests that
-residual departures displayed with evenly distributed fitted values are
-visually more attractive and effective. Besides, the skewness of the
-fitted value distribution plays a role in the degree of presence of
-visual features as lognormal distribution produce the lowest power for
-both non-linearity and heteroskedasticity patterns. A symmetric fitted
-value distribution can better reveal the underlying shape.
-
-Although it is unusual to consider fitted value distribution or
-predictor distribution in power analysis, it is not so surprised that
-the power of conventional tests vary because of fitted value
-distribution. For instance, the power of the RESET test which is
-effectively a F test depends on the realizations of the additional
-predictors included in the auxiliary regression equation
-\citep{jamshidian2007study, olvera2019relationship, zhang2018practical}.
-And those additional predictors are linear transformation of random
-vectors following certain predictor distributions. What is truly
-unexpected is the huge variability in power of conventional tests
-compared to visual tests. The discrete uniform distribution constantly
-produce the highest power while the normal distribution and the
-lognormal distribution produce the lowest power for non-linearity and
-heteroskedasticity pattern respectively. The difference in power
-corresponding to different distributions peak at around 90\% when the
-effect size is at a moderate level. This huge difference is undesirable
-as is not uncommon for analysts to omit the factor of predictor
-distribution in power calculation and sample size calculation.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-x-dist-poly-power-1}
-
-}
-
-\caption{Comparison of power on lineups with different fitted value distributions for conventional tests and visual tests.
-Power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and Breusch-Pagan tests respectively.
-Power curves of visual tests are estimated using five evaluations on each lineup.
-For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations.
-The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines.
-The fitted value distribution has greater impact on the power of conventional tests than visual tests.
-Meanwhile, uneven distributions including normal and lognormal distributions genearlly produce lower power.}\label{fig:different-x-dist-poly-power}
-\end{figure}
-
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
-
-Motivated by the advice of regression analysis experts, that residual
-plots as opposed to conventional tests are an indispensible methods for
-assessing model fit, a human subjects experiment was conducted using
-visual inference. The experiment tested two primary departures from good
-residuals: non-linearity and heteroskedasticity.
-
-The experiment found that conventional residual-based statistical tests
-are more sensitive to weak residual departures from model assumptions
-than visual tests as would be evaluated by humans. That is, conventional
-tests conclude there are problems with the model fit almost twice as
-often as humans would. They often reject when departures in the form of
-non-linearity and heteroskedasticity are not visible to a human.
-
-One might say that this is correct behaviour, but it can be argued that
-the conventional tests are rejecting when it is not necessary. Many of
-these rejections happen even when downstream analysis and results would
-not be significantly affected by the small departures from a good fit.
-The results from human evaluations provide a more practical solution,
-which reinforces the statements from regression experts that residual
-plots are an indispensible method for model diagnostics.
-
-Now it is important to note that residual plots need to be delivered as
-a lineup, where it is embedded in a field of null plots. A residual plot
-may contains many visual features, but some are caused by the
-characteristics of the predictors and the randomness of the error, not
-by the violation of the model assumptions. These irrelevant visual
-features have a chance to be filtered out by subjects with a comparison
-to null plots, results in a set of more accurate visual findings. This
-enables a careful calibration for reading structure in residual plots.
-
-However, human evaluation of residuals is expensive. It is
-time-consuming, laborious and unfriendly to vision-impaired people. This
-is another reason why it often appears to be ignored. With the
-availability of sophisticated computer vision algorithms today, the goal
-of this work is to form the basis of providing automated residual plot
-reading. The findings suggest the strong demand of graphical inspection
-in regression diagnostics, so developing an automatic visual inference
-system to evaluate lineups of residual plots is valuable. We plan to
-build a completed open-source system in an \texttt{R} package and
-provide a web interface such as a website for public to interact with.
-Details about this system will be discussed in our next paper.
-
-The experiment also revealed some interesting details about how residual
-plots are read. For the most part, the visual test performed very
-similarly to the appropriate conventional test only with the power curve
-shifted in the less sensitive direction. Unlike the conventional tests,
-where one needs to specifically test for non-linearity or
-heteroskedasticity the visual test operated effectively across the range
-of departures from good residuals.
-
-As expected, if the fitted value distribution is not uniform, there is a
-loss of power in the visual test. Structure is hardest to detect if
-fitted values are lognormal. Also, complex structure are generally
-harder to detect, but there are outliers. Under the designed scenarios
-in this paper, we find the visual test to be a more robust test against
-the change of fitted value distributions. A surprising finding was that
-the direction of heteroskedasticity appears to affect the ability to
-visually detect it, with wedge to the right being less detectable.
-
-\hypertarget{acknowledgements}{%
-\section*{Acknowledgements}\label{acknowledgements}}
-\addcontentsline{toc}{section}{Acknowledgements}
-
-These \texttt{R} packages are used for the work: \texttt{cli}
-\citep{cli}, \texttt{curl} \citep{curl}, \texttt{dplyr} \citep{dplyr},
-\texttt{ggplot2} \citep{ggplot2}, \texttt{jsonlite} \citep{jsonlite},
-\texttt{lmtest} \citep{lmtest}, \texttt{mpoly} \citep{mpoly},
-\texttt{progress} \citep{progress}, \texttt{tibble} \citep{tibble},
-\texttt{ggmosaic} \citep{ggmosaic}, \texttt{purrr} \citep{purrr},
-\texttt{tidyr} \citep{tidyr}, \texttt{readr} \citep{readr},
-\texttt{stringr} \citep{stringr}, \texttt{here} \citep{here},
-\texttt{kableExtra} \citep{kableextra}, \texttt{patchwork}
-\citep{patchwork}, \texttt{rcartocolor} \citep{rcartocolor}. The study
-website is powered by \texttt{PythonAnywhere} \citep{pythonanywhere} and
-the \texttt{Flask} web framework \citep{flask}. The \texttt{jsPsych}
-framework \citep{jspsych} is used to create behavioral experiments that
-run in our study website.
-
-The article was created with R packages \texttt{rticles}
-\citep{rticles}, \texttt{knitr} \citep{knitr} and \texttt{rmarkdown}
-\citep{rmarkdown}. The project's Github repository (link here) contains
-all materials required to reproduce this article.
-
-\hypertarget{supplementary-material}{%
-\section*{Supplementary material}\label{supplementary-material}}
-\addcontentsline{toc}{section}{Supplementary material}
-
-The supplementary material is available at (link here). It includes more
-details about the experimental setup, the derivation of the effect size,
-the effect of data collection period, and the estimate of \(\alpha\).
-
 \appendix
 
-\hypertarget{experiment-setup}{%
-\section{Experiment setup}\label{experiment-setup}}
-
-\hypertarget{mapping-of-subjects-to-experimental-factors}{%
-\subsection{Mapping of subjects to experimental
-factors}\label{mapping-of-subjects-to-experimental-factors}}
-
-Mapping of subjects to experimental factors is an important part of
-experiment design. Essentially, we want to maximum the difference in
-factors exposed to a subject. For this purpose, we design an algorithm
-to conduct subject allocation. Let \(L\) be a set of available lineups
-and \(S\) be a set of available subjects. According to the experimental
-design, the availability of a lineup is associated with the number of
-subjects it can assign to. For lineups with uniform fitted value
-distribution, this value is 11. And other lineups can be allocated to at
-most five different subjects. The availability of a subject is
-associated with the number of lineups that being allocated to this
-subject. A subject can view at most 18 different lineups.
-
-The algorithm starts from picking a random subject \(s \in S\) with the
-minimum number of allocated lineups. It then tries to find a lineup
-\(l \in L\) that can maximise the distance metric \(D\) and allocate it
-to subject \(s\). Set \(L\) and \(S\) will be updated and the picking
-process will be repeated until there is no available lineups or
-subjects.
-
-Let \(F_1,...,F_q\) be \(q\) experimental factors, and \(f_1, ...,f_q\)
-be the corresponding factor values. We say \(f_i\) exists in \(L_{s}\)
-if any lineup in \(L_{s}\) has this factor value. Similarly, \(f_if_j\)
-exists in \(L_{s}\) if any lineup in \(L_{s}\) has this pair of factor
-values. And \(f_if_jf_k\) exists in \(L_{s}\) if any lineup in \(L_{s}\)
-has this trio of factor values. The distance metric \(D\) is defined
-between a lineup \(l\) and a set of lineups \(L_{s}\) allocated to a
-subject \(s\) if \(L_{s}\) is non-empty:
-
-\footnotesize
-
-\begin{equation*}
-D =
-C - \sum_{1\leq i \leq q}I(f_i\text{ exists in }L_{s}) - \sum_{\substack{1\leq i \leq q-1 \\ i < j \leq q}}I(f_if_j\text{ exists in }L_{s}) - \sum_{\substack{1\leq i \leq q - 2 \\ i < j \leq q - 1 \\ j < k \leq q}}I(f_if_jf_k\text{ exists in }L_{s})
-\end{equation*}
-
-\normalsize
-
-where \(C\) is a sufficiently large constant such that \(D > 0\). If
-\(L_{s}\) is empty, we define \(D = 0\).
-
-The distance measures how different a lineup is from the set of lineups
-allocated to the subject in terms of factor values. Thus, the algorithm
-will try to allocate the most different lineup to a subject at each
-step.
-
-\hypertarget{data-collection-process}{%
-\subsection{Data collection process}\label{data-collection-process}}
-
-The survey data is collected via a self-hosted website designed by us.
-The complete architecture is provided in Figure \ref{fig:tech}. The
-website is built with the \texttt{Flask} \citep{flask} web framework and
-hosted on \texttt{PythonAnywhere} \citep{pythonanywhere}. It is
-configured to handle HTTP requests such that subjects can correctly
-receive webpages and submit responses. Embedded in the resources sent to
-subjects, the \texttt{jsPsych} front-end framework \citep{jspsych}
-instructs subjects' browsers to render an environment for running
-behavioral experiments. During the experiment, this framework will
-automatically collect common behavioral data such as response time and
-clicks on buttons. Subjects' responses are first validated by a
-scheduled \texttt{Python} script run on the server, then push to a
-Github repository. Lineup images shown to users are saved in multiple
-Github repositories and hosted in corresponding Github pages. The URLs
-to these images are resolved by \texttt{Flask} and bundled in HTML
-files.
-
-Once the participant is recruited from Prolific
-\citep{palan2018prolific}, it will be redirected to the entry page of
-our study website. An image of the entry page is provided in Figure
-\ref{fig:entry-page}. Then, the participant needs to submit the online
-consent form and fill in the demographic information as shown in
-\ref{fig:consent-form} and \ref{fig:metadata} respectively. Before
-evaluating lineups, participant also need to read the training page as
-provide in Figure \ref{fig:training-page} to understand the process. An
-example of the lineup page is given in Figure \ref{fig:lineup-page}. A
-half of the page is taken by the lineup image to attract participant's
-attention. The button to skip the selections for the current lineup is
-intentionally put in the corner of the bounding box with smaller font
-size, such that participants will not misuse this functionality.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figures/experiment_tech}
-
-}
-
-\caption{Diagram of online experiment setup. The server-side of the study website uses Flask as backend hosted on PythonAnywhere. And the client-side uses jsPsych to run experiment.}\label{fig:tech}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figures/message}
-
-}
-
-\caption{The entry page of the study website.}\label{fig:entry-page}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figures/consent_form}
-
-}
-
-\caption{The consent form provided in the study website.}\label{fig:consent-form}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figures/metadata}
-
-}
-
-\caption{The form to provide demographic information.}\label{fig:metadata}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figures/training}
-
-}
-
-\caption{The training page of the study website.}\label{fig:training-page}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figures/lineup1}
-
-}
-
-\caption{The lineup page of the study website.}\label{fig:lineup-page}
-\end{figure}
-
-\hypertarget{demographics}{%
-\section{Demographics}\label{demographics}}
-
-Along with the responses to lineups, we have collected a series of
-demographic information including age, pronoun, education background and
-previous experience in studies involved data visualization. Table
-\ref{tab:pronoun}, \ref{tab:age-group}, \ref{tab:education} and
-\ref{tab:experience} provide summary of the demographic data.
-
-It can be observed from the tables that most participants have Diploma
-or Bachelor degrees, followed by High school or below and the survey
-data is gender balanced. Majority of participants are between 18 to 39
-years old and there are slightly more participants who do not have
-previous experience than those who have.
-
-\begin{table}
-
-\caption{\label{tab:pronoun}Summary of pronoun distribution of subjects recuritted in this study.}
-\centering
-\begin{tabular}[t]{lrrrrrrrr}
-\toprule
-Pronoun & Period I & \% & Period II & \% & Period III & \% & Total & \%\\
-\midrule
-He & 77 & 17.4 & 79 & 17.8 & 61 & 13.8 & 217 & 49.0\\
-She & 78 & 17.6 & 77 & 17.4 & 61 & 13.8 & 216 & 48.8\\
-Other & 5 & 1.1 & 4 & 0.9 & 1 & 0.2 & 10 & 2.3\\
- & 160 & 36.1 & 160 & 36.1 & 123 & 27.8 & 443 & 100.0\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{table}
-
-\caption{\label{tab:age-group}Summary of age distribution of subjects recuritted in this study.}
-\centering
-\begin{tabular}[t]{lrrrr}
-\toprule
-Age group & Period I & Period II & Period III & Total\\
-\midrule
-18-24 & 83 & 86 & 51 & 220\\
-25-39 & 69 & 63 & 63 & 195\\
-40-54 & 6 & 8 & 6 & 20\\
-55-64 & 2 & 3 & 3 & 8\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{table}
-
-\caption{\label{tab:education}Summary of education distribution of subjects recuritted in this study.}
-\centering
-\begin{tabular}[t]{lllll}
-\toprule
-Education & Period I & Period II & Period III & Total\\
-\midrule
-High School or below & 41 (9.3\%) & 53 (12\%) & 33 (7.4\%) & 127 (28.7\%)\\
-Diploma and Bachelor Degree & 92 (20.8\%) & 79 (17.8\%) & 66 (14.9\%) & 237 (53.5\%)\\
-Honours Degree & 6 (1.4\%) & 15 (3.4\%) & 6 (1.4\%) & 27 (6.1\%)\\
-Masters Degree & 21 (4.7\%) & 13 (2.9\%) & 16 (3.6\%) & 50 (11.3\%)\\
-Doctoral Degree & 0 (0\%) & 0 (0\%) & 2 (0.5\%) & 2 (0.5\%)\\
-\addlinespace
- & 160 (36.1\%) & 160 (36.1\%) & 123 (27.8\%) & 443 (100\%)\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{table}
-
-\caption{\label{tab:experience}Summary of previous experience distribution of subjects recuritted in this study.}
-\centering
-\begin{tabular}[t]{lrrrr}
-\toprule
-Previous experience & Period I & Period II & Period III & Total\\
-\midrule
-No & 96 & 88 & 67 & 251\\
-Yes & 64 & 72 & 56 & 192\\
-\bottomrule
-\end{tabular}
-\end{table}
+\hypertarget{additional-details-of-testing-procedures}{%
+\section{Additional details of testing
+procedures}\label{additional-details-of-testing-procedures}}
 
 \hypertarget{effect-size-derivation}{%
-\section{Effect size derivation}\label{effect-size-derivation}}
+\subsection{Effect size derivation}\label{effect-size-derivation}}
 
@@
 
-\hypertarget{effect-of-data-collection-period}{%
-\section{Effect of data collection
-period}\label{effect-of-data-collection-period}}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/poly-boxplot-lineup-1}
+\hypertarget{sensitivity-analysis-for-alpha}{%
+\subsection{\texorpdfstring{Sensitivity analysis for
+\(\alpha\)}{Sensitivity analysis for \textbackslash alpha}}\label{sensitivity-analysis-for-alpha}}
+
+The parameter \(\alpha\) used for the \(p\)-value calculation needs to
+be estimated from responses to null lineups. With a greater value of
+\(\hat{\alpha}\), the \(p\)-value will be smaller, resulting in more
+lineups being rejected. However, The way we generate Rorschach lineup is
+not strictly the same as what suggested in
+\citet{vanderplas2021statistical} and \citet{buja_statistical_2009}.
+Therefore, we conduct a sensitivity analysis in this section to examine
+the impact of the variation of the estimator \(\alpha\) on our primary
+findings.
+
+The analysis is conducted by setting up several scenarios, where the
+\(\alpha\) is under or overestimated by 12.5\%, 25\% and 50\%. Using the
+adjusted \(\hat{\alpha}\), we recalculate the \(p\)-value for every
+lineup and show the results in Figure \ref{fig:sensitivity}. It can be
+observed that there are some changes to \(p\)-values, especially when
+the \(\hat{\alpha}\) is multiplied by 50\%. However, Table
+\ref{tab:sensitivity} shows that adjusting \(\hat{\alpha}\) will not
+result in a huge difference in rejection decisions. There are only a
+small percentage of cases where the rejection decision change. It is
+very unlikely the downstream findings will be affected because of the
+estimate of \(\alpha\).
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{appendix_files/figure-latex/sensitivity-1}
+
+}
+
+\caption{Change of $p$-values with $\hat{\alpha}$ multiplied by $0.5$, $0.75$, $0.875$, $1.125$, $1.25$ and $1.5$.
+The vertical dashed line is to indicate $p\text{-value} = 0.05$.
+The x-axis is drawn on logarithmic scale.
+For multiplier smaller than $1$, the adjusted $p$-value will incrase then decrase as $p$-value increases.
+The trend is the opposite for multiplier greater than $1$, but the difference will eventually reach $0$.}\label{fig:sensitivity}
+\end{figure}
+
+\begin{table}
+
+\caption{\label{tab:sensitivity}Examining how decisions might change if $\hat{\alpha}$ was different.
+Percentage of lineups that where the $p$-value would switch to above or below 0.05, when $\hat{\alpha}$ is varied by a fractional amount.}
+\centering
+\begin{tabular}[t]{rll}
+\toprule
+fraction & reject to not reject & not reject to reject\\
+\midrule
+0.500 & 2.51 & 0\\
+0.750 & 1.43 & 0\\
+0.875 & 1.08 & 0\\
+1.125 & 0 & 1.08\\
+1.250 & 0 & 1.43\\
+\addlinespace
+1.500 & 0 & 1.79\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\hypertarget{effect-of-number-of-evaluations-on-the-power-of-a-visual-test}{%
+\subsection{Effect of number of evaluations on the power of a visual
+test}\label{effect-of-number-of-evaluations-on-the-power-of-a-visual-test}}
+
+When comparing power of visual tests across different fitted value
+distributions, we have discussed the number of evaluations on a lineup
+will affect the power of the visual test. Using the lineups with uniform
+fitted value distribution, we show in Figure \ref{fig:n_eval} the change
+of power of visual tests due to different number of evaluations. It can
+be learned that as the number of evaluations increases, the power will
+increase but the margin will decrease. Considering we have eleven
+evaluations on lineups with uniform fitted value distribution, and five
+evaluations on other lineups, it is necessary to use the same number of
+evaluations for each lineup in comparison.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.7\linewidth]{appendix_files/figure-latex/n_eval-1}
+
+}
+
+\caption{Change of power of visual tests for different number of evalutions on lineups with uniform fitted value distribution.
+The power will incrase as the number of evaluations increases, but the margin will decrase.}\label{fig:n_eval}
+\end{figure}
+
+\hypertarget{power-of-a-reset-test-under-different-auxiliary-regression-formulas}{%
+\subsection{Power of a RESET test under different auxiliary regression
+formulas}\label{power-of-a-reset-test-under-different-auxiliary-regression-formulas}}
+
+It is found in the result that the power of a RESET test will be
+affected by the highest order of fitted values included in the auxiliary
+formula. And we suspect that the current recommendation of the highest
+order - four, is insufficient to test complex non-linear structures such
+as the ``Triple-U'' shape designed in this paper. Figure \ref{fig:reset}
+illustrates the change of power of RESET test while testing the ``U''
+shape and the ``Triple-U'' shape with different highest orders. Clearly,
+when testing a simple shape like the ``U'' shape, the highest order has
+very little impact on the power. But for testing the ``Triple-U'' shape,
+there will be a loss of power if the recommended order is used. To avoid
+the loss of power, the highest order needs to be set to at least six.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{appendix_files/figure-latex/reset-1}
+
+}
+
+\caption{Change of power of RESET tests for different orders of fitted values included in the auxiliary formula.
+The left panel is the power of testing the "U" shape and the right panel is the power of testing the "Triple-U" shape.
+The power will not be greatly affected by the highest order in the case of testing the "U" shape.
+In the case of testing the "Triple-U" shape, the highest order needs to be set to at least six to avoid the loss of power.}\label{fig:reset}
+\end{figure}
+
+\newpage
+
+\hypertarget{experiment-setup}{%
+\section{Experiment setup}\label{experiment-setup}}
+
+\hypertarget{mapping-of-subjects-to-experimental-factors}{%
+\subsection{Mapping of subjects to experimental
+factors}\label{mapping-of-subjects-to-experimental-factors}}
+
+Mapping of subjects to experimental factors is an important part of
+experiment design. Essentially, we want to maximum the difference in
+factors exposed to a subject. For this purpose, we design an algorithm
+to conduct subject allocation. Let \(L\) be a set of available lineups
+and \(S\) be a set of available subjects. According to the experimental
+design, the availability of a lineup is associated with the number of
+subjects it can assign to. For lineups with uniform fitted value
+distribution, this value is 11. And other lineups can be allocated to at
+most five different subjects. The availability of a subject is
+associated with the number of lineups that being allocated to this
+subject. A subject can view at most 18 different lineups.
+
+The algorithm starts from picking a random subject \(s \in S\) with the
+minimum number of allocated lineups. It then tries to find a lineup
+\(l \in L\) that can maximise the distance metric \(D\) and allocate it
+to subject \(s\). Set \(L\) and \(S\) will be updated and the picking
+process will be repeated until there is no available lineups or
+subjects.
+
+Let \(F_1,...,F_q\) be \(q\) experimental factors, and \(f_1, ...,f_q\)
+be the corresponding factor values. We say \(f_i\) exists in \(L_{s}\)
+if any lineup in \(L_{s}\) has this factor value. Similarly, \(f_if_j\)
+exists in \(L_{s}\) if any lineup in \(L_{s}\) has this pair of factor
+values. And \(f_if_jf_k\) exists in \(L_{s}\) if any lineup in \(L_{s}\)
+has this trio of factor values. The distance metric \(D\) is defined
+between a lineup \(l\) and a set of lineups \(L_{s}\) allocated to a
+subject \(s\) if \(L_{s}\) is non-empty:
+
+\footnotesize
+
+\begin{equation*}
+D =
+C - \sum_{1\leq i \leq q}I(f_i\text{ exists in }L_{s}) - \sum_{\substack{1\leq i \leq q-1 \\ i < j \leq q}}I(f_if_j\text{ exists in }L_{s}) - \sum_{\substack{1\leq i \leq q - 2 \\ i < j \leq q - 1 \\ j < k \leq q}}I(f_if_jf_k\text{ exists in }L_{s})
+\end{equation*}
+
+\normalsize
+
+where \(C\) is a sufficiently large constant such that \(D > 0\). If
+\(L_{s}\) is empty, we define \(D = 0\).
+
+The distance measures how different a lineup is from the set of lineups
+allocated to the subject in terms of factor values. Thus, the algorithm
+will try to allocate the most different lineup to a subject at each
+step.
+
+\hypertarget{data-collection-process}{%
+\subsection{Data collection process}\label{data-collection-process}}
+
+The survey data is collected via a self-hosted website designed by us.
+The complete architecture is provided in Figure \ref{fig:tech}. The
+website is built with the \texttt{Flask} \citep{flask} web framework and
+hosted on \texttt{PythonAnywhere} \citep{pythonanywhere}. It is
+configured to handle HTTP requests such that subjects can correctly
+receive webpages and submit responses. Embedded in the resources sent to
+subjects, the \texttt{jsPsych} front-end framework \citep{jspsych}
+instructs subjects' browsers to render an environment for running
+behavioral experiments. During the experiment, this framework will
+automatically collect common behavioral data such as response time and
+clicks on buttons. Subjects' responses are first validated by a
+scheduled \texttt{Python} script run on the server, then push to a
+Github repository. Lineup images shown to users are saved in multiple
+Github repositories and hosted in corresponding Github pages. The URLs
+to these images are resolved by \texttt{Flask} and bundled in HTML
+files.
+
+Once the participant is recruited from Prolific
+\citep{palan2018prolific}, it will be redirected to the entry page of
+our study website. An image of the entry page is provided in Figure
+\ref{fig:entry-page}. Then, the participant needs to submit the online
+consent form and fill in the demographic information as shown in
+\ref{fig:consent-form} and \ref{fig:metadata} respectively. Before
+evaluating lineups, participant also need to read the training page as
+provide in Figure \ref{fig:training-page} to understand the process. An
+example of the lineup page is given in Figure \ref{fig:lineup-page}. A
+half of the page is taken by the lineup image to attract participant's
+attention. The button to skip the selections for the current lineup is
+intentionally put in the corner of the bounding box with smaller font
+size, such that participants will not misuse this functionality.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures/experiment_tech}
+
+}
+
+\caption{Diagram of online experiment setup. The server-side of the study website uses Flask as backend hosted on PythonAnywhere. And the client-side uses jsPsych to run experiment.}\label{fig:tech}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures/message}
+
+}
+
+\caption{The entry page of the study website.}\label{fig:entry-page}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures/consent_form}
+
+}
+
+\caption{The consent form provided in the study website.}\label{fig:consent-form}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures/metadata}
+
+}
+
+\caption{The form to provide demographic information.}\label{fig:metadata}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures/training}
+
+}
+
+\caption{The training page of the study website.}\label{fig:training-page}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures/lineup1}
+
+}
+
+\caption{The lineup page of the study website.}\label{fig:lineup-page}
+\end{figure}
+
+\newpage
+
+\hypertarget{analysis-of-results-relative-to-data-collection-process}{%
+\section{Analysis of results relative to data collection
+process}\label{analysis-of-results-relative-to-data-collection-process}}
+
+\hypertarget{demographics}{%
+\subsection{Demographics}\label{demographics}}
+
+Along with the responses to lineups, we have collected a series of
+demographic information including age, pronoun, education background and
+previous experience in studies involved data visualization. Table
+\ref{tab:pronoun}, \ref{tab:age-group}, \ref{tab:education} and
+\ref{tab:experience} provide summary of the demographic data.
+
+It can be observed from the tables that most participants have Diploma
+or Bachelor degrees, followed by High school or below and the survey
+data is gender balanced. Majority of participants are between 18 to 39
+years old and there are slightly more participants who do not have
+previous experience than those who have.
+
+\begin{table}
+
+\caption{\label{tab:pronoun}Summary of pronoun distribution of subjects recuritted in this study.}
+\centering
+\begin{tabular}[t]{lrrrrrrrr}
+\toprule
+Pronoun & Period I & \% & Period II & \% & Period III & \% & Total & \%\\
+\midrule
+He & 77 & 17.4 & 79 & 17.8 & 61 & 13.8 & 217 & 49.0\\
+She & 78 & 17.6 & 77 & 17.4 & 61 & 13.8 & 216 & 48.8\\
+Other & 5 & 1.1 & 4 & 0.9 & 1 & 0.2 & 10 & 2.3\\
+ & 160 & 36.1 & 160 & 36.1 & 123 & 27.8 & 443 & 100.0\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:age-group}Summary of age distribution of subjects recuritted in this study.}
+\centering
+\begin{tabular}[t]{lrrrr}
+\toprule
+Age group & Period I & Period II & Period III & Total\\
+\midrule
+18-24 & 83 & 86 & 51 & 220\\
+25-39 & 69 & 63 & 63 & 195\\
+40-54 & 6 & 8 & 6 & 20\\
+55-64 & 2 & 3 & 3 & 8\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:education}Summary of education distribution of subjects recuritted in this study.}
+\centering
+\begin{tabular}[t]{lllll}
+\toprule
+Education & Period I & Period II & Period III & Total\\
+\midrule
+High School or below & 41 (9.3\%) & 53 (12\%) & 33 (7.4\%) & 127 (28.7\%)\\
+Diploma and Bachelor Degree & 92 (20.8\%) & 79 (17.8\%) & 66 (14.9\%) & 237 (53.5\%)\\
+Honours Degree & 6 (1.4\%) & 15 (3.4\%) & 6 (1.4\%) & 27 (6.1\%)\\
+Masters Degree & 21 (4.7\%) & 13 (2.9\%) & 16 (3.6\%) & 50 (11.3\%)\\
+Doctoral Degree & 0 (0\%) & 0 (0\%) & 2 (0.5\%) & 2 (0.5\%)\\
+\addlinespace
+ & 160 (36.1\%) & 160 (36.1\%) & 123 (27.8\%) & 443 (100\%)\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:experience}Summary of previous experience distribution of subjects recuritted in this study.}
+\centering
+\begin{tabular}[t]{lrrrr}
+\toprule
+Previous experience & Period I & Period II & Period III & Total\\
+\midrule
+No & 96 & 88 & 67 & 251\\
+Yes & 64 & 72 & 56 & 192\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\hypertarget{data-collection-periods}{%
+\subsection{Data collection periods}\label{data-collection-periods}}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{appendix_files/figure-latex/poly-boxplot-lineup-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/heter-boxplot-lineup-1}
+{\centering \includegraphics[width=1\linewidth]{appendix_files/figure-latex/heter-boxplot-lineup-1}
 
@@
 
-\hypertarget{sensitivity-analysis-for-alpha}{%
-\section{\texorpdfstring{Sensitivity analysis for
-\(\alpha\)}{Sensitivity analysis for \textbackslash alpha}}\label{sensitivity-analysis-for-alpha}}
-
-The parameter \(\alpha\) used for the \(p\)-value calculation needs to
-be estimated from responses to null lineups. With a greater value of
-\(\hat{\alpha}\), the \(p\)-value will be smaller, resulting in more
-lineups being rejected. However, The way we generate Rorschach lineup is
-not strictly the same as what suggested in
-\citet{vanderplas2021statistical} and \citet{buja_statistical_2009}.
-Therefore, we conduct a sensitivity analysis in this section to examine
-the impact of the variation of the estimator \(\alpha\) on our primary
-findings.
-
-The analysis is conducted by setting up several scenarios, where the
-\(\alpha\) is under or overestimated by 12.5\%, 25\% and 50\%. Using the
-adjusted \(\hat{\alpha}\), we recalculate the \(p\)-value for every
-lineup and show the results in Figure \ref{fig:sensitivity}. It can be
-observed that there are some changes to \(p\)-values, especially when
-the \(\hat{\alpha}\) is multiplied by 50\%. However, Table
-\ref{tab:sensitivity} shows that adjusting \(\hat{\alpha}\) will not
-result in a huge difference in rejection decisions. There are only a
-small percentage of cases where the rejection decision change. It is
-very unlikely the downstream findings will be affected because of the
-estimate of \(\alpha\).
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/sensitivity-1}
-
-}
-
-\caption{Change of $p$-values with $\hat{\alpha}$ multiplied by $0.5$, $0.75$, $0.875$, $1.125$, $1.25$ and $1.5$.
-The vertical dashed line is to indicate $p\text{-value} = 0.05$.
-The x-axis is drawn on logarithmic scale.
-For multiplier smaller than $1$, the adjusted $p$-value will incrase then decrase as $p$-value increases.
-The trend is the opposite for multiplier greater than $1$, but the difference will eventually reach $0$.}\label{fig:sensitivity}
-\end{figure}
-
-\begin{table}
-
-\caption{\label{tab:sensitivity}Change of rejection decision because of the modification of $\hat{\alpha}$.}
-\centering
-\begin{tabular}[t]{rll}
-\toprule
-multiplier & Proportion of lineups transforms to "not reject" & Proportion of lineups transforms to "reject"\\
-\midrule
-0.500 & 2.51\% & 0\%\\
-0.750 & 1.43\% & 0\%\\
-0.875 & 1.08\% & 0\%\\
-1.125 & 0\% & 1.08\%\\
-1.250 & 0\% & 1.43\%\\
-\addlinespace
-1.500 & 0\% & 1.79\%\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\hypertarget{effect-of-number-of-evaluations-on-the-power-of-a-visual-test}{%
-\section{Effect of number of evaluations on the power of a visual
-test}\label{effect-of-number-of-evaluations-on-the-power-of-a-visual-test}}
-
-When comparing power of visual tests across different fitted value
-distributions, we have discussed the number of evaluations on a lineup
-will affect the power of the visual test. Using the lineups with uniform
-fitted value distribution, we show in Figure \ref{fig:n_eval} the change
-of power of visual tests due to different number of evaluations. It can
-be learned that as the number of evaluations increases, the power will
-increase but the margin will decrease. Considering we have eleven
-evaluations on lineups with uniform fitted value distribution, and five
-evaluations on other lineups, it is necessary to use the same number of
-evaluations for each lineup in comparison.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/n_eval-1}
-
-}
-
-\caption{Change of power of visual tests for different number of evalutions on lineups with uniform fitted value distribution.
-The power will incrase as the number of evaluations increases, but the margin will decrase.}\label{fig:n_eval}
-\end{figure}
-
-\hypertarget{power-of-a-reset-test-under-different-auxiliary-regression-formulas}{%
-\section{Power of a RESET test under different auxiliary regression
-formulas}\label{power-of-a-reset-test-under-different-auxiliary-regression-formulas}}
-
-It is found in the result that the power of a RESET test will be
-affected by the highest order of fitted values included in the auxiliary
-formula. And we suspect that the current recommendation of the highest
-order - four, is insufficient to test complex non-linear structures such
-as the ``Triple-U'' shape designed in this paper. Figure \ref{fig:reset}
-illustrates the change of power of RESET test while testing the ``U''
-shape and the ``Triple-U'' shape with different highest orders. Clearly,
-when testing a simple shape like the ``U'' shape, the highest order has
-very little impact on the power. But for testing the ``Triple-U'' shape,
-there will be a loss of power if the recommended order is used. To avoid
-the loss of power, the highest order needs to be set to at least six.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/reset-1}
-
-}
-
-\caption{Change of power of RESET tests for different orders of fitted values included in the auxiliary formula.
-The left panel is the power of testing the "U" shape and the right panel is the power of testing the "Triple-U" shape.
-The power will not be greatly affected by the highest order in the case of testing the "U" shape.
-In the case of testing the "Triple-U" shape, the highest order needs to be set to at least six to avoid the loss of power.}\label{fig:reset}
-\end{figure}
-
-\clearpage
-
 \bibliographystyle{tfcad}

### 2023-04-03 paper.Rmd
rearranged main paper and appendix into separate files again

@@
 $$E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)$$
-\noindent where $\boldsymbol{V}$ is the actual covariance matrix of the error term.
-
-Derivations for these equations are provided in the Appendix.
+\noindent where $\boldsymbol{V}$ is the actual covariance matrix of the error term. (Derivations for these equations are provided in the Appendix.)
 
@@
 To account for the fact that our simulation method for these lineups is not the Rorschach procedure, we use the method suggested in @vanderplas2021statistical for typical lineups containing a data plot to estimate $\alpha$.
-We have included a sensitivity analysis in the Appendix to examine the impact of the variance of the $\alpha$ estimate on our findings.
+(We have included a sensitivity analysis in the Appendix to examine the impact of the variance of the $\alpha$ estimate on our findings.)
 
@@
 It includes more details about the experimental setup, the derivation of the effect size, the effect of data collection period, and the estimate of  $\alpha$.
-
-\appendix
 

### 2023-04-03 paper.tex
rearranged main paper and appendix into separate files again

@@
 \noindent where \(\boldsymbol{V}\) is the actual covariance matrix of
-the error term.
-
-Derivations for these equations are provided in the Appendix.
+the error term. (Derivations for these equations are provided in the
+Appendix.)
 
@@
 \citet{vanderplas2021statistical} for typical lineups containing a data
-plot to estimate \(\alpha\). We have included a sensitivity analysis in
+plot to estimate \(\alpha\). (We have included a sensitivity analysis in
 the Appendix to examine the impact of the variance of the \(\alpha\)
-estimate on our findings.
+estimate on our findings.)
 
@@
 the effect of data collection period, and the estimate of \(\alpha\).
-
-\appendix
 

### 2023-04-03 paper.Rmd
Auto stash before merge of "master" and "origin/master"

@@
 # Visual inference models and p-value calculation
-# remotes::install_github("TengMCing/visage")
+# remotes::install_url("https://github.com/TengMCing/visage/raw/master/built/visage_0.1.0.tar.gz")
 
@@
   geom_point(aes(p_value, diff_p_value), alpha = 0.2) +
-  scale_x_log10() +
   facet_wrap(~modified) +
@@
   xlab(expression(italic(p)*"-value")) +
-  theme_light()
-
+  theme_light() +
+  scale_x_continuous(breaks = c(0.0001, 0.001, 0.01, 0.1, 1),
+                     minor_breaks = NULL,
+                     trans = "log10",
+                     labels = c("0.01%", "0.1%", "1%", "10%", "100%"))
 ```
@@
 ```{r n_eval, fig.cap = "Change of power of visual tests for different number of evalutions on lineups with uniform fitted value distribution.
-The power will incrase as the number of evaluations increases, but the margin will decrase."}
+The power will incrase as the number of evaluations increases, but the margin will decrase.", fig.height = 4}
 eval_n <- function(n_eval, boot) {

### 2023-04-03 paper.Rmd
delete the appendix from the main article

@@
 It includes more details about the experimental setup, the derivation of the effect size, the effect of data collection period, and the estimate of  $\alpha$.
-
-# Experiment setup
-
-## Mapping of subjects to experimental factors
-
-Mapping of subjects to experimental factors is an important part of experiment design.
-Essentially, we want to maximum the difference in factors exposed to a subject.
-For this purpose, we design an algorithm to conduct subject allocation.
-Let $L$ be a set of available lineups and $S$ be a set of available subjects.
-According to the experimental design, the availability of a lineup is associated with the number of subjects it can assign to.
-For lineups with uniform fitted value distribution, this value is 11.
-And other lineups can be allocated to at most five different subjects.
-The availability of a subject is associated with the number of lineups that being allocated to this subject.
-A subject can view at most 18 different lineups.
-
-The algorithm starts from picking a random subject $s \in S$ with the minimum number of allocated lineups.
-It then tries to find a lineup $l \in L$ that can maximise the distance metric $D$ and allocate it to subject $s$.
-Set $L$ and $S$ will be updated and the picking process will be repeated until there is no available lineups or subjects.
-
-Let $F_1,...,F_q$ be $q$ experimental factors, and $f_1, ...,f_q$ be the corresponding factor values.
-We say $f_i$ exists in $L_{s}$ if any lineup in $L_{s}$ has this factor value.
-Similarly, $f_if_j$ exists in $L_{s}$ if any lineup in $L_{s}$ has this pair of factor values.
-And $f_if_jf_k$ exists in $L_{s}$ if any lineup in $L_{s}$ has this trio of factor values.
-The distance metric $D$ is defined between a lineup $l$ and a set of lineups $L_{s}$ allocated to a subject $s$ if $L_{s}$ is non-empty:
-
-\footnotesize
-
-\begin{equation*}
-D =
-C - \sum_{1\leq i \leq q}I(f_i\text{ exists in }L_{s}) - \sum_{\substack{1\leq i \leq q-1 \\ i < j \leq q}}I(f_if_j\text{ exists in }L_{s}) - \sum_{\substack{1\leq i \leq q - 2 \\ i < j \leq q - 1 \\ j < k \leq q}}I(f_if_jf_k\text{ exists in }L_{s})
-\end{equation*}
-
-\normalsize
-where $C$ is a sufficiently large constant such that $D > 0$. If $L_{s}$ is empty, we define $D = 0$.
-
-The distance measures how different a lineup is from the set of lineups allocated to the subject in terms of factor values.
-Thus, the algorithm will try to allocate the most different lineup to a subject at each step.
-
-## Data collection process
-
-The survey data is collected via a self-hosted website designed by us.
-The complete architecture is provided in Figure \ref{fig:tech}.
-The website is built with the `Flask` [@flask] web framework and hosted on `PythonAnywhere` [@pythonanywhere].
-It is configured to handle HTTP requests such that subjects can correctly receive webpages and submit responses.
-Embedded in the resources sent to subjects, the `jsPsych` front-end framework [@jspsych] instructs subjects' browsers to render an environment for running behavioral experiments.
-During the experiment, this framework will automatically collect common behavioral data such as response time and clicks on buttons.
-Subjects' responses are first validated by a scheduled `Python` script run on the server, then push to a Github repository.
-Lineup images shown to users are saved in multiple Github repositories and hosted in corresponding Github pages.
-The URLs to these images are resolved by `Flask` and bundled in HTML files.
-
-Once the participant is recruited from Prolific [@palan2018prolific], it will be redirected to the entry page of our study website.
-An image of the entry page is provided in Figure \ref{fig:entry-page}.
-Then, the participant needs to submit the online consent form and fill in the demographic information as shown in \ref{fig:consent-form} and \ref{fig:metadata} respectively.
-Before evaluating lineups, participant also need to read the training page as provide in Figure \ref{fig:training-page} to understand the process.
-An example of the lineup page is given in Figure \ref{fig:lineup-page}.
-A half of the page is taken by the lineup image to attract participant's attention.
-The button to skip the selections for the current lineup is intentionally put in the corner of the bounding box with smaller font size, such that participants will not misuse this functionality.
-
-```{r tech, out.width = "100%", fig.cap = "Diagram of online experiment setup.
-The server-side of the study website uses Flask as backend hosted on PythonAnywhere.
-And the client-side uses jsPsych to run experiment."}
-knitr::include_graphics("figures/experiment_tech.pdf")
-```
-
-```{r entry-page, out.width = "100%", fig.cap = "The entry page of the study website."}
-knitr::include_graphics("figures/message.png")
-```
-
-```{r consent-form, out.width = "100%", fig.cap = "The consent form provided in the study website."}
-knitr::include_graphics("figures/consent_form.png")
-```
-
-```{r metadata, out.width = "100%", fig.cap = "The form to provide demographic information."}
-knitr::include_graphics("figures/metadata.png")
-```
-
-```{r training-page, out.width = "100%", fig.cap = "The training page of the study website."}
-knitr::include_graphics("figures/training.png")
-```
-
-```{r lineup-page, out.width = "100%", fig.cap = "The lineup page of the study website."}
-knitr::include_graphics("figures/lineup1.png")
-```
-
-# Demographics
-
-Along with the responses to lineups, we have collected a series of demographic information including age, pronoun, education background and previous experience in studies involved data visualization.
-Table \ref{tab:pronoun}, \ref{tab:age-group}, \ref{tab:education} and \ref{tab:experience} provide summary of the demographic data.
-
-It can be observed from the tables that most participants have Diploma or Bachelor degrees, followed by High school or below and the survey data is gender balanced.
-Majority of participants are between 18 to 39 years old and there are slightly more participants who do not have previous experience than those who have.
-
-```{r}
-vi_survey %>%
-  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
-  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
-  group_by(exp, set) %>%
-  summarise(across(everything(), first)) %>%
-  count(pronoun) %>%
-  mutate(exp = paste("Period", c("I", "II", "III")[exp - 2])) %>%
-  pivot_wider(values_from = n, names_from = exp) %>%
-  mutate(Total = `Period I` + `Period II` + `Period III`) %>%
-  rename(Pronoun = pronoun) %>%
-  add_row(Pronoun = "",
-          `Period I` = sum(.$`Period I`),
-          `Period II` = sum(.$`Period II`),
-          `Period III` = sum(.$`Period III`),
-          Total = sum(.$Total)) %>%
-  mutate(`Period I %` = round(`Period I`/443*100, 1)) %>%
-  mutate(`Period II %` = round(`Period II`/443*100, 1)) %>%
-  mutate(`Period III %` = round(`Period III`/443*100, 1)) %>%
-  mutate(`Total %` = round(Total/443*100, 1)) %>%
-  select(Pronoun, 
-         `Period I`, `Period I %`, 
-         `Period II`, `Period II %`,
-         `Period III`, `Period III %`,
-         Total, `Total %`) %>%
-  kableExtra::kable(format = "latex", 
-                    col.names = c("Pronoun", 
-                                  "Period I", "%", 
-                                  "Period II", "%",
-                                  "Period III", "%",
-                                  "Total", "%"), 
-                    booktabs = TRUE, label = "pronoun", caption = "Summary of pronoun distribution of subjects recuritted in this study.")
-```
-
-```{r}
-vi_survey %>%
-  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
-  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
-  group_by(exp, set) %>%
-  summarise(across(everything(), first)) %>%
-  count(age_group) %>%
-  mutate(exp = paste("Period", c("I", "II", "III")[exp - 2])) %>%
-  pivot_wider(values_from = n, names_from = exp) %>%
-  mutate(Total = `Period I` + `Period II` + `Period III`) %>%
-  rename(`Age group` = age_group) %>%
-  kableExtra::kable(format = "latex", booktabs = TRUE, label = "age-group", caption = "Summary of age distribution of subjects recuritted in this study.")
-```
-
-```{r}
-vi_survey %>%
-  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
-  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
-  group_by(exp, set) %>%
-  summarise(across(everything(), first)) %>%
-  count(education) %>%
-  mutate(exp = paste("Period", c("I", "II", "III")[exp - 2])) %>%
-  pivot_wider(values_from = n, names_from = exp, values_fill = 0) %>%
-  mutate(Total = `Period I` + `Period II` + `Period III`) %>%
-  slice(c(2, 1, 3, 4, 5)) %>%
-  rename(`Education` = education) %>%
-  add_row(`Education` = "",
-          `Period I` = sum(.$`Period I`),
-          `Period II` = sum(.$`Period II`),
-          `Period III` = sum(.$`Period III`),
-          Total = sum(.$Total)) %>%
-  mutate(`Period I` = paste0(`Period I`, " (", round(`Period I`/443*100, 1), "%)")) %>%
-  mutate(`Period II` = paste0(`Period II`, " (", round(`Period II`/443*100, 1), "%)")) %>%
-  mutate(`Period III` = paste0(`Period III`, " (", round(`Period III`/443*100, 1), "%)")) %>%
-  mutate(Total = paste0(Total, " (", round(Total/443*100, 1), "%)")) %>%
-  kableExtra::kable(format = "latex", booktabs = TRUE, label = "education", caption = "Summary of education distribution of subjects recuritted in this study.")
-```
-
-```{r}
-vi_survey %>%
-  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
-  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
-  group_by(exp, set) %>%
-  summarise(across(everything(), first)) %>%
-  count(previous_experience) %>%
-  mutate(exp = paste("Period", c("I", "II", "III")[exp - 2])) %>%
-  pivot_wider(values_from = n, names_from = exp) %>%
-  mutate(Total = `Period I` + `Period II` + `Period III`) %>%
-  rename(`Previous experience` = previous_experience) %>%
-  kableExtra::kable(format = "latex", booktabs = TRUE, label = "experience", caption = "Summary of previous experience distribution of subjects recuritted in this study.")
-```
-
-# Effect size derivation
-
-Effect size can be defined as the difference of a parameter for a particular model or distribution,  or a statistic derived from a sample.
-Importantly, it needs to reflect the treatment we try to measure.
-Centred on a conventional statistical test, we usually can deduce the effect size from the test statistic by substituting the null parameter value.
-When considering the diagnostics of residual departures, there exist many possibilities of test statistics for a variety of model assumptions.
-Meanwhile, diagnostic plots such as the residual plot have no general agreement on measuring how strong a model violation pattern is.
-To build a bridge between various residual-based tests, and the visual test, we focus on the shared information embedded in the testing procedures, which is the distribution of residuals.
-When comes to comparison of distribution, Kullback-Leibler divergence [@kullback1951information] is a classical way to represent the information loss or entropy increase caused by the approximation to the true distribution, which in our case, the inefficiency due to the use of false model assumptions.
-
-Following the terminology introduced by @kullback1951information, $P$ represents the measured probability distribution, and $Q$ represents the assumed probability distribution.
-The Kullback-Leibler divergence is defined as $\int_{-\infty}^{\infty}log(p(x)/q(x))p(x)dx$, where $p(.)$ and $q(.)$ denote probability densities of $P$ and $Q$.
-
-Let $\boldsymbol{X}_a = (\boldsymbol{1}, \boldsymbol{X})$ denotes the $p$ regressors with $n$ observations, $\boldsymbol{R}_a = \boldsymbol{I} -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ denotes the residual operator, and let $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0},\sigma^2\boldsymbol{I})$ denotes the error.
-Using the Frisch–Waugh–Lovell theorem, residuals $\boldsymbol{e} = \boldsymbol{R}_a\boldsymbol{\varepsilon}$.
-Because $rank(\boldsymbol{R}_a) = n - p < n$, $e$ follows a degenerate multivariate normal distribution and does not have a density.
-Since the Kullback-Leibler divergence requires a proper density function, we need to simplify the covariance matrix of $\boldsymbol{e}$ by setting all the off-diagonal elements to 0.
-Then, the residuals will assumed to follow $N(\boldsymbol{0}, diag(\boldsymbol{R}_a\sigma^2))$ under the null hypothesis that the model is correctly specified.
-If the model is however misspecified due to omitted variables $\boldsymbol{Z}$, or a non-constant variance $\boldsymbol{V}$, the distribution of residuals can be derived as $N(\boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}_a\sigma^2))$ and $N(\boldsymbol{0}, diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a'))$ respectively.
-
-By assuming both $P$ and $Q$ are multivariate normal density functions, the Kullback-Leibler divergence can be rewritten as 
-$$KL = \frac{1}{2}\left(log\frac{|\Sigma_p|}{|\Sigma_q|} - n + tr(\Sigma_p^{-1}\Sigma_q) + (\mu_p - \mu_q)'\Sigma_p^{-1}(\mu_p - \mu_q)\right).$$
-
-Then, we can combine the two residual departures into one formula
-
-\small
-
-\begin{equation}
-\label{eq:effect-size}
-KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}\boldsymbol{\mu}_z\right).
-\end{equation}
-
-\normalsize
-
-When there are omitted variables but constant error variance, the formula can be reduced to
-
-$$KL = \frac{1}{2}\left(\boldsymbol{\mu}_z^{T}(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right).$$
-
-And when the model equation is correctly specified but the error variance is non-constant, the formula can be reduced to
-
-$$KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2))\right).$$
-Since we assume $\sigma = 1$ for the heteroskedasticity model, the final form of the formula is
-
-$$KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right).$$
-
-# Effect of data collection period
-
-```{r poly-boxplot-lineup, fig.cap = 'A lineup of "letter-value" boxplots of weighted propotion of detect for lineups over different data collection periods for non-linearity model.
-Can you find the most different boxplot?
-The data plot is positioned in panel $2^3 - 1$.'}
-library(lvplot)
-sample_from_null <- function(dat, true_position) {
-  dat %>%
-    mutate(k = true_position) %>%
-    bind_rows(map_df((1:20)[-true_position], function(k) {
-    dat %>%
-      mutate(exp = sample(exp),
-             k = k)
-  }))
-}
-
-vi_survey %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(x_dist == "uniform") %>%
-  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
-  mutate(exp = exp - 2) %>%
-  group_by(exp, unique_lineup_id) %>%
-  mutate(prop_detect = mean(weighted_detect)) %>%
-  group_by(exp, unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  ungroup() %>%
-  filter(type == "non-linearity") %>%
-  select(type, exp, prop_detect) %>%
-  mutate(exp = as.character(exp)) %>%
-  sample_from_null(7) %>%
-  ggplot() +
-  geom_lv(aes(x = exp, y = prop_detect, fill = after_stat(LV)), varwidth = TRUE, k = 4) +
-  scale_alpha_discrete(range = c(0.7, 0)) +
-  facet_wrap(~k, ncol = 5) +
-  theme_light() +
-  scale_x_discrete(labels = c("I", "III")) +
-  labs(fill = "Quantiles") +
-  scale_fill_manual(labels = c("0.5", "[0.25, 0.75]", "[0.125, 0.875]", "[0.0625, 0.09375]"),
-                    values = c("white", rev(RColorBrewer::brewer.pal(5, "Blues")))) +
-  ylab("Weighted propotion of detect") +
-  xlab("Data collection period") +
-  theme(legend.key = element_rect(colour = "black"))
-```
-
-```{r heter-boxplot-lineup, fig.cap = 'A lineup of "letter-value" boxplots of weighted propotion of detect for lineups over different data collection periods for heteroskedasticity model.
-Can you find the most different boxplot?
-The data plot is positioned in panel $2^4 - 2$.'}
-vi_survey %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(x_dist == "uniform") %>%
-  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
-  mutate(exp = exp - 2) %>%
-  group_by(exp, unique_lineup_id) %>%
-  mutate(prop_detect = mean(weighted_detect)) %>%
-  group_by(exp, unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  ungroup() %>%
-  filter(type != "non-linearity") %>%
-  select(type, exp, prop_detect) %>%
-  mutate(exp = as.character(exp)) %>%
-  sample_from_null(14) %>%
-  ggplot() +
-  geom_lv(aes(x = exp, y = prop_detect, fill = after_stat(LV)), varwidth = TRUE, k = 4) +
-  facet_wrap(~k, ncol = 5) +
-  theme_light() +
-  scale_x_discrete(labels = c("II", "III")) +
-  labs(fill = "Quantiles") +
-  scale_fill_manual(labels = c("0.5", "[0.25, 0.75]", "[0.125, 0.875]", "[0.0625, 0.09375]"),
-                    values = c("white", rev(RColorBrewer::brewer.pal(5, "Blues")))) +
-  ylab("Weighted propotion of detect") +
-  xlab("Data collection period") +
-  theme(legend.key = element_rect(colour = "black"))
-```
-
-We have the same type of model collected over different data collection periods, that may lead to unexpected batch effect.
-Figure \ref{fig:poly-boxplot-lineup} and \ref{fig:heter-boxplot-lineup} provide two lineups to examine whether there is an actual difference across data collection periods for non-linearity model and heteroskedasticity model respectively.
-To emphasize the tail behaviour and display fewer outliers, we use the "letter-value" boxplot [@hofmann2017value] which is an extension of the number of "letter value" statistics to check the weighed proportion of detect over different data collection period.
-The weighted proportion of detect is calculated by taking the average of $c_i$ of a lineup over a data collection period.
-Within our research team, we can not identify the data plot from the null plots for these two lineups, result in $p$-values much greater than $5$%.
-Thus, there is no clear evidence of batch effect.
-
-# Sensitivity analysis for $\alpha$
-
-The parameter $\alpha$ used for the $p$-value calculation needs to be estimated from responses to null lineups.
-With a greater value of $\hat{\alpha}$, the $p$-value will be smaller, resulting in more lineups being rejected.
-However, The way we generate Rorschach lineup is not strictly the same as what suggested in @vanderplas2021statistical and @buja_statistical_2009.
-Therefore, we conduct a sensitivity analysis in this section to examine the impact of the variation of the estimator $\alpha$ on our primary findings.
-
-The analysis is conducted by setting up several scenarios, where the $\alpha$ is under or overestimated by 12.5%, 25% and 50%.
-Using the adjusted $\hat{\alpha}$, we recalculate the $p$-value for every lineup and show the results in Figure \ref{fig:sensitivity}.
-It can be observed that there are some changes to $p$-values, especially when the $\hat{\alpha}$ is multiplied by 50%.
-However, Table \ref{tab:sensitivity} shows that adjusting $\hat{\alpha}$ will not result in a huge difference in rejection decisions.
-There are only a small percentage of cases where the rejection decision change.
-It is very unlikely the downstream findings will be affected because of the estimate of $\alpha$.
-
-```{r}
-upper_p <- map_df(c(0.125, 0.25, 0.5), function(modified) {
-  calc_p_value_multi(vi_survey %>%
-                       filter(!attention_check) %>%
-                       filter(!null_lineup) %>%
-                       filter(x_dist == "uniform") %>%
-                       mutate(alpha_upper = alpha * (1 + modified)),
-                     lineup_id = unique_lineup_id,
-                     detect = detect,
-                     n_sel = num_selection,
-                     alpha = alpha_upper) %>%
-  mutate(modified = 1 + modified) %>%
-  mutate(adjusted_p_value = p_value)
-})
-
-lower_p <- map_df(c(0.125, 0.25, 0.5), function(modified) {
-  calc_p_value_multi(vi_survey %>%
-                       filter(!attention_check) %>%
-                       filter(!null_lineup) %>%
-                       filter(x_dist == "uniform") %>%
-                       mutate(alpha_lower = alpha * (1 - modified)),
-                     lineup_id = unique_lineup_id,
-                     detect = detect,
-                     n_sel = num_selection,
-                     alpha = alpha_lower) %>%
-  mutate(modified = 1 - modified) %>%
-  mutate(adjusted_p_value = p_value)
-})
-
-upper_lower_p <- upper_p %>%
-  bind_rows(lower_p) %>%
-  select(-p_value)
-```
-
-```{r sensitivity, fig.cap = "Change of $p$-values with $\\hat{\\alpha}$ multiplied by $0.5$, $0.75$, $0.875$, $1.125$, $1.25$ and $1.5$.
-The vertical dashed line is to indicate $p\\text{-value} = 0.05$.
-The x-axis is drawn on logarithmic scale.
-For multiplier smaller than $1$, the adjusted $p$-value will incrase then decrase as $p$-value increases.
-The trend is the opposite for multiplier greater than $1$, but the difference will eventually reach $0$."}
-upper_lower_p %>%
-  left_join(vi_survey %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(x_dist == "uniform") %>%
-  group_by(unique_lineup_id) %>%
-  summarise(p_value = first(p_value))) %>%
-  mutate(diff_p_value = adjusted_p_value - p_value) %>%
-  ggplot() +
-  geom_hline(yintercept = 0) +
-  geom_vline(xintercept = 0.05, col = "red", linetype = 2) +
-  geom_point(aes(p_value, diff_p_value), alpha = 0.2) +
-  facet_wrap(~modified) +
-  ylab(expression("Adjusted"~italic(p)*"-value minus"~italic(p)*"-value")) +
-  xlab(expression(italic(p)*"-value")) +
-  theme_light() +
-  scale_x_continuous(breaks = c(0.0001, 0.001, 0.01, 0.1, 1),
-                     minor_breaks = NULL,
-                     trans = "log10",
-                     labels = c("0.01%", "0.1%", "1%", "10%", "100%"))
-```
-
-```{r}
-upper_lower_p %>%
-  left_join(vi_survey %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(x_dist == "uniform") %>%
-  group_by(unique_lineup_id) %>%
-  summarise(p_value = first(p_value))) %>%
-  group_by(modified) %>%
-  count(adjusted_reject = adjusted_p_value <= 0.05, reject = p_value <= 0.05) %>%
-  complete(adjusted_reject, reject, fill = list(n = 0)) %>%
-  ungroup() %>%
-  filter(adjusted_reject != reject) %>%
-  mutate(case = ifelse(!adjusted_reject, 'Proportion of lineups transforms to "not reject"', 'Proportion of lineups transforms to "reject"')) %>%
-  rename(multiplier = modified) %>%
-  mutate(n = paste0(round(n/279 * 100, 2), "\\%")) %>%
-  select(-adjusted_reject, -reject) %>%
-  pivot_wider(names_from = case, values_from = n) %>%
-  kableExtra::kable(format = "latex", booktabs = TRUE, label = "sensitivity", 
-                    escape = FALSE,
-                    caption = "Change of rejection decision because of the modification of $\\hat{\\alpha}$.")
-```
-
-# Effect of number of evaluations on the power of a visual test
-
-When comparing power of visual tests across different fitted value distributions, we have discussed the number of evaluations on a lineup will affect the power of the visual test.
-Using the lineups with uniform fitted value distribution, we show in Figure \ref{fig:n_eval} the change of power of visual tests due to different number of evaluations.
-It can be learned that as the number of evaluations increases, the power will increase but the margin will decrease.
-Considering we have eleven evaluations on lineups with uniform fitted value distribution, and five evaluations on other lineups, it is necessary to use the same number of evaluations for each lineup in comparison.
-
-```{r}
-min_poly_es <- vi_survey %>% 
-  filter(!null_lineup, 
-         !attention_check,
-         type == "polynomial", 
-         x_dist == "uniform") %>%
-  pull(effect_size) %>%
-  min()
-max_poly_es <- vi_survey %>% 
-  filter(!null_lineup, 
-         !attention_check,
-         type == "polynomial", 
-         x_dist == "uniform") %>%
-  pull(effect_size) %>%
-  max()
-```
-
-```{r n_eval, fig.cap = "Change of power of visual tests for different number of evalutions on lineups with uniform fitted value distribution.
-The power will incrase as the number of evaluations increases, but the margin will decrase.", fig.height = 4}
-eval_n <- function(n_eval, boot) {
-  map_df(1:boot, function(boot_id) {
-  vi_survey %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  filter(x_dist == "uniform") %>%
-  filter(type == "polynomial") %>%
-  group_by(unique_lineup_id) %>%
-  slice_sample(n = n_eval) %>%
-  ungroup() %>%
-  calc_p_value_multi(lineup_id = unique_lineup_id,
-                     detect = detect,
-                     n_sel = num_selection,
-                     alpha = alpha) %>%
-  mutate(boot_id = boot_id)
-}) %>%
-  left_join(select(vi_survey, unique_lineup_id, effect_size)) %>%
-  mutate(reject = as.numeric(p_value <= 0.05)) %>%
-  mutate(offset0 = log(0.05/0.95)) %>%
-  select(-unique_lineup_id, -p_value) %>%
-  mutate(n_eval = n_eval)
-}
-
-map_df(seq(1, 11, 2), ~eval_n(.x, 1)) %>%
-  select(-boot_id) %>%
-  nest(c(effect_size, reject, offset0)) %>%
-  mutate(mod = map(data, 
-                   ~glm(reject ~ effect_size - 1, 
-                        family = binomial(), 
-                        data = .x,
-                        offset = offset0))) %>%
-  mutate(power = map(mod, function(mod) {
-    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
-                         offset0 = log(0.05/0.95))
-    result$power <- predict(mod, type = "response", newdata = result)
-    result
-  })) %>%
-  select(-data, -mod) %>%
-  unnest(power) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  ggplot() +
-  geom_line(aes(log_effect_size, power, col = factor(n_eval)),
-            size = 1) +
-  labs(col = "Number of evalutaions") +
-  xlab(expression(log[e] (Effect_size))) +
-  theme_light() +
-  scale_color_brewer(palette = "Reds")
-```
-
-# Power of a RESET test under different auxiliary regression formulas
-
-It is found in the result that the power of a RESET test will be affected by the highest order of fitted values included in the auxiliary formula.
-And we suspect that the current recommendation of the highest order - four, is insufficient to test complex non-linear structures such as the "Triple-U" shape designed in this paper.
-Figure \ref{fig:reset} illustrates the change of power of RESET test while testing the "U" shape and the "Triple-U" shape with different highest orders.
-Clearly, when testing a simple shape like the "U" shape, the highest order has very little impact on the power.
-But for testing the "Triple-U" shape, there will be a loss of power if the recommended order is used.
-To avoid the loss of power, the highest order needs to be set to at least six.
-
-```{r cache = TRUE}
-reset_diff_dat <- poly_conv_sim %>%
-  filter(x_dist == "uniform") %>%
-  filter(shape == 4) %>%
-  select(-F_p_value, -BP_p_value, -SW_p_value) %>%
-  pivot_longer(RESET3_p_value:RESET10_p_value) %>%
-  mutate(name = gsub("RESET(.*)_p_value", "\\1", name)) %>%
-  mutate(reject = value <= 0.05) %>%
-  select(effect_size, name, reject) %>%
-  mutate(offset0 = log(0.05/0.95)) %>%
-  nest(dat = c(effect_size, offset0, reject)) %>%
-  mutate(mod = map(dat, 
-                   ~glm(reject ~ effect_size - 1, 
-                        family = binomial(), 
-                        data = .x,
-                        offset = offset0))) %>%
-  mutate(power = map(mod, function(mod) {
-    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
-                         offset0 = log(0.05/0.95))
-    result$power <- predict(mod, type = "response", newdata = result)
-    result
-  })) %>%
-  select(-dat, -mod) %>%
-  unnest(power) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  mutate(shape = 4)
-```
-
-```{r cache = TRUE}
-reset_diff_dat2 <- poly_conv_sim %>%
-  filter(x_dist == "uniform") %>%
-  filter(shape == 1) %>%
-  select(-F_p_value, -BP_p_value, -SW_p_value) %>%
-  pivot_longer(RESET3_p_value:RESET10_p_value) %>%
-  mutate(name = gsub("RESET(.*)_p_value", "\\1", name)) %>%
-  mutate(reject = value <= 0.05) %>%
-  select(effect_size, name, reject) %>%
-  mutate(offset0 = log(0.05/0.95)) %>%
-  nest(dat = c(effect_size, offset0, reject)) %>%
-  mutate(mod = map(dat, 
-                   ~glm(reject ~ effect_size - 1, 
-                        family = binomial(), 
-                        data = .x,
-                        offset = offset0))) %>%
-  mutate(power = map(mod, function(mod) {
-    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
-                         offset0 = log(0.05/0.95))
-    result$power <- predict(mod, type = "response", newdata = result)
-    result
-  })) %>%
-  select(-dat, -mod) %>%
-  unnest(power) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  mutate(shape = 1)
-```
-
-```{r reset, fig.cap = 'Change of power of RESET tests for different orders of fitted values included in the auxiliary formula.
-The left panel is the power of testing the "U" shape and the right panel is the power of testing the "Triple-U" shape.
-The power will not be greatly affected by the highest order in the case of testing the "U" shape.
-In the case of testing the "Triple-U" shape, the highest order needs to be set to at least six to avoid the loss of power.', fig.height = 4}
-reset_diff_dat %>%
-  bind_rows(reset_diff_dat2) %>%
-  mutate(shape = factor(ifelse(shape == 1, "U", "Triple-U"), levels = c("U", "Triple-U"))) %>%
-  ggplot() +
-  geom_line(aes(log_effect_size, power, col = as.factor(as.integer(name))),
-            size = 1) +
-  theme_light() +
-  labs(col = "Highest order") +
-  xlab(expression(log[e] (Effect_size))) +
-  scale_color_brewer(palette = "Reds") +
-  facet_wrap(~shape)
-```
-
-\clearpage

### 2023-04-04 paper.Rmd
working on section 5.6

@@
 
-The prediction in a regression model is $E(Y|X)$, that is, it is conditional on observed values of the predictors.
-The distribution of $X$, or consequently $\hat{Y}$, may however affect the ability to read any patterns in the residuals.
-The four distributions of fitted values, used in this experiment, were designed to examine this.
-
-Figure \ref{fig:different-x-dist-poly-power} illustrates the impact of fitted value distribution on the power of conventional tests and visual tests.
-For conventional tests, only the power curves of appropriate tests which are RESET tests and Breusch-Pagan tests for non-linearity pattern and heteroskedasticity pattern respectively are shown.
-For visual tests, note that we have collected more evaluations on lineup with uniform fitted value distribution, and the number of evaluations on a lineup will affect the power of the visual test.
-To have a fair comparison, for lineups with uniform fitted value distribution, we randomly sample five evaluations from the total eleven evaluations to estimate the power curves.
-The sampling process is repeated 500 times to produce an indication of the variation of the power.
-
-In terms of visual tests, we do not observe significant difference between the power curve of the uniform distribution and the power curve of the discrete uniform distribution.
-Normal and lognormal distributions produce lower power than uniform and discrete uniform distributions in regards of residual departure patterns, but the gap is larger when the residual plot contains heteroskedasticity pattern.
-This suggests that residual departures displayed with evenly distributed fitted values are visually more attractive and effective.
-Besides, the skewness of the fitted value distribution plays a role in the degree of presence of visual features as lognormal distribution produce the lowest power for both non-linearity and heteroskedasticity patterns.
-A symmetric fitted value distribution can better reveal the underlying shape.
-
-Although it is unusual to consider fitted value distribution or predictor distribution in power analysis, it is not so surprised that the power of conventional tests vary because of fitted value distribution.
-For instance, the power of the RESET test which is effectively a F test depends on the realizations of the additional predictors included in the auxiliary regression equation [@jamshidian2007study; @olvera2019relationship; @zhang2018practical].
-And those additional predictors are linear transformation of random vectors following certain predictor distributions.
-What is truly unexpected is the huge variability in power of conventional tests compared to visual tests.
-The discrete uniform distribution constantly produce the highest power while the normal distribution and the lognormal distribution produce the lowest power for non-linearity and heteroskedasticity pattern respectively.
-The difference in power corresponding to different distributions peak at around 90% when the effect size is at a moderate level.
-This huge difference is undesirable as is not uncommon for analysts to omit the factor of predictor distribution in power calculation and sample size calculation.
+In regression predictions are conditional on the observed values of the predictors, that is, the conditional mean of the dependent variable Y given the value of the independent variable X, $E(Y|X)$.
+This is an often forgotten element of regression analysis but it is important.
+Where X is observed, particularly the distribution of the X values in the sample, or consequently $\hat{Y}$, may affect the ability to read any patterns in the residual plots.
+This experiment was constructed to assess this, based on four different distributions of fitted values: uniform, normal, discrete and lognormal (skewed).
+We would expect that if $\hat{Y}$ has a uniform distribution, this would make it easier to read the relationship with the residuals.
 
@@
 The fitted value distribution has greater impact on the power of conventional tests than visual tests.
-Meanwhile, uneven distributions including normal and lognormal distributions genearlly produce lower power.", dev = 'png', dpi = 450}
+Meanwhile, uneven distributions including normal and lognormal distributions genearlly produce lower power.", dev = 'png', dpi = 450, fig.pos="t"}
 
@@
 
+Figure \ref{fig:different-x-dist-poly-power} examines the impact of the fitted value distribution on the power of conventional and visual tests.
+For conventional tests, only the power curves of appropriate tests which are RESET tests and Breusch-Pagan tests for non-linearity pattern and heteroskedasticity pattern respectively are shown.
+For visual tests, note that more evaluations on lineup with uniform fitted value distribution were collected, so to have a fair comparison, we randomly sample five from the total eleven evaluations to estimate the power curves.
+Re-sampling produces the multiple curves for the uniform, and provides an indication of the variation of the power.
+
+In terms of visual tests, we do not observe significant difference between the power curve of the uniform distribution and the power curve of the discrete uniform distribution.
+Normal and lognormal distributions produce lower power than uniform and discrete uniform distributions in regards of residual departure patterns, but the gap is larger when the residual plot contains heteroskedasticity pattern.
+This suggests that residual departures displayed with evenly distributed fitted values are visually more attractive and effective.
+Besides, the skewness of the fitted value distribution plays a role in the degree of presence of visual features as lognormal distribution produce the lowest power for both non-linearity and heteroskedasticity patterns.
+A symmetric fitted value distribution can better reveal the underlying shape.
+
+Although it is unusual to consider fitted value distribution or predictor distribution in power analysis, it is not so surprised that the power of conventional tests vary because of fitted value distribution.
+For instance, the power of the RESET test which is effectively a F test depends on the realizations of the additional predictors included in the auxiliary regression equation [@jamshidian2007study; @olvera2019relationship; @zhang2018practical].
+And those additional predictors are linear transformation of random vectors following certain predictor distributions.
+What is truly unexpected is the huge variability in power of conventional tests compared to visual tests.
+The discrete uniform distribution constantly produce the highest power while the normal distribution and the lognormal distribution produce the lowest power for non-linearity and heteroskedasticity pattern respectively.
+The difference in power corresponding to different distributions peak at around 90% when the effect size is at a moderate level.
+This huge difference is undesirable as is not uncommon for analysts to omit the factor of predictor distribution in power calculation and sample size calculation.
+
 # Conclusion

### 2023-04-04 paper.tex
working on section 5.6

@@
 
-The prediction in a regression model is \(E(Y|X)\), that is, it is
-conditional on observed values of the predictors. The distribution of
-\(X\), or consequently \(\hat{Y}\), may however affect the ability to
-read any patterns in the residuals. The four distributions of fitted
-values, used in this experiment, were designed to examine this.
-
-Figure \ref{fig:different-x-dist-poly-power} illustrates the impact of
-fitted value distribution on the power of conventional tests and visual
-tests. For conventional tests, only the power curves of appropriate
-tests which are RESET tests and Breusch-Pagan tests for non-linearity
-pattern and heteroskedasticity pattern respectively are shown. For
-visual tests, note that we have collected more evaluations on lineup
-with uniform fitted value distribution, and the number of evaluations on
-a lineup will affect the power of the visual test. To have a fair
-comparison, for lineups with uniform fitted value distribution, we
-randomly sample five evaluations from the total eleven evaluations to
-estimate the power curves. The sampling process is repeated 500 times to
-produce an indication of the variation of the power.
+In regression predictions are conditional on the observed values of the
+predictors, that is, the conditional mean of the dependent variable Y
+given the value of the independent variable X, \(E(Y|X)\). This is an
+often forgotten element of regression analysis but it is important.
+Where X is observed, particularly the distribution of the X values in
+the sample, or consequently \(\hat{Y}\), may affect the ability to read
+any patterns in the residual plots. This experiment was constructed to
+assess this, based on four different distributions of fitted values:
+uniform, normal, discrete and lognormal (skewed). We would expect that
+if \(\hat{Y}\) has a uniform distribution, this would make it easier to
+read the relationship with the residuals.
+
+\begin{figure}[t]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-x-dist-poly-power-1}
+
+}
+
+\caption{Comparison of power on lineups with different fitted value distributions for conventional tests and visual tests.
+Power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and Breusch-Pagan tests respectively.
+Power curves of visual tests are estimated using five evaluations on each lineup.
+For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations.
+The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines.
+The fitted value distribution has greater impact on the power of conventional tests than visual tests.
+Meanwhile, uneven distributions including normal and lognormal distributions genearlly produce lower power.}\label{fig:different-x-dist-poly-power}
+\end{figure}
+
+Figure \ref{fig:different-x-dist-poly-power} examines the impact of the
+fitted value distribution on the power of conventional and visual tests.
+For conventional tests, only the power curves of appropriate tests which
+are RESET tests and Breusch-Pagan tests for non-linearity pattern and
+heteroskedasticity pattern respectively are shown. For visual tests,
+note that more evaluations on lineup with uniform fitted value
+distribution were collected, so to have a fair comparison, we randomly
+sample five from the total eleven evaluations to estimate the power
+curves. Re-sampling produces the multiple curves for the uniform, and
+provides an indication of the variation of the power.
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-x-dist-poly-power-1}
-
-}
-
-\caption{Comparison of power on lineups with different fitted value distributions for conventional tests and visual tests.
-Power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and Breusch-Pagan tests respectively.
-Power curves of visual tests are estimated using five evaluations on each lineup.
-For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations.
-The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines.
-The fitted value distribution has greater impact on the power of conventional tests than visual tests.
-Meanwhile, uneven distributions including normal and lognormal distributions genearlly produce lower power.}\label{fig:different-x-dist-poly-power}
-\end{figure}
-
 \hypertarget{conclusion}{%

### 2023-04-04 paper.Rmd
finished revising section 5.6

@@
 
-Figure \ref{fig:different-x-dist-poly-power} examines the impact of the fitted value distribution on the power of conventional and visual tests.
+Figure \ref{fig:different-x-dist-poly-power} examines the impact of the fitted value distribution on the power of conventional (top row) and visual (bottom row) tests for both the non-linearity (left column) and heteroskedasticity (right column) patterns.
 For conventional tests, only the power curves of appropriate tests which are RESET tests and Breusch-Pagan tests for non-linearity pattern and heteroskedasticity pattern respectively are shown.
-For visual tests, note that more evaluations on lineup with uniform fitted value distribution were collected, so to have a fair comparison, we randomly sample five from the total eleven evaluations to estimate the power curves.
+For visual tests, note that, more evaluations on lineup with uniform fitted value distribution were collected, so to have a fair comparison, we randomly sample five from the total eleven evaluations to estimate the power curves.
 Re-sampling produces the multiple curves for the uniform, and provides an indication of the variation of the power.
 
-In terms of visual tests, we do not observe significant difference between the power curve of the uniform distribution and the power curve of the discrete uniform distribution.
+Perhaps surprisingly, the visual tests have more consistent power across the different fitted value distributions.
+For the non-linear pattern, there is almost no power difference.
+For the heteroskedastic pattern, uniform and discrete have higher power than normal and lognormal.
+The likely reason is that these latter two have less observations in the tails where the heteroskedastic pattern needs to be detected.
+
+<!-- In terms of visual tests, we do not observe significant difference between the power curve of the uniform distribution and the power curve of the discrete uniform distribution.
 Normal and lognormal distributions produce lower power than uniform and discrete uniform distributions in regards of residual departure patterns, but the gap is larger when the residual plot contains heteroskedasticity pattern.
@@
 Besides, the skewness of the fitted value distribution plays a role in the degree of presence of visual features as lognormal distribution produce the lowest power for both non-linearity and heteroskedasticity patterns.
-A symmetric fitted value distribution can better reveal the underlying shape.
-
-Although it is unusual to consider fitted value distribution or predictor distribution in power analysis, it is not so surprised that the power of conventional tests vary because of fitted value distribution.
+A symmetric fitted value distribution can better reveal the underlying shape.-->
+
+The variation in power in the conventional tests is at first sight, shocking.
+However, it is discussed, albeit rarely, in the testing literature.
+See, for example, @jamshidian2007study, @olvera2019relationship and  @zhang2018practical which show derivations and use simulation to assess the effect of the observed distribution of the predictors on test power.
+The big differences in power curves that are seen in Figure \ref{fig:different-x-dist-poly-power} echoes the results reported in these articles.
+
+<!-- Although it is unusual to consider fitted value distribution or predictor distribution in power analysis, it is not so surprised that the power of conventional tests vary because of fitted value distribution.
 For instance, the power of the RESET test which is effectively a F test depends on the realizations of the additional predictors included in the auxiliary regression equation [@jamshidian2007study; @olvera2019relationship; @zhang2018practical].
@@
 This huge difference is undesirable as is not uncommon for analysts to omit the factor of predictor distribution in power calculation and sample size calculation.
+-->
 

### 2023-04-04 paper.tex
finished revising section 5.6

@@
 Figure \ref{fig:different-x-dist-poly-power} examines the impact of the
-fitted value distribution on the power of conventional and visual tests.
-For conventional tests, only the power curves of appropriate tests which
-are RESET tests and Breusch-Pagan tests for non-linearity pattern and
-heteroskedasticity pattern respectively are shown. For visual tests,
-note that more evaluations on lineup with uniform fitted value
-distribution were collected, so to have a fair comparison, we randomly
-sample five from the total eleven evaluations to estimate the power
-curves. Re-sampling produces the multiple curves for the uniform, and
-provides an indication of the variation of the power.
-
-In terms of visual tests, we do not observe significant difference
-between the power curve of the uniform distribution and the power curve
-of the discrete uniform distribution. Normal and lognormal distributions
-produce lower power than uniform and discrete uniform distributions in
-regards of residual departure patterns, but the gap is larger when the
-residual plot contains heteroskedasticity pattern. This suggests that
-residual departures displayed with evenly distributed fitted values are
-visually more attractive and effective. Besides, the skewness of the
-fitted value distribution plays a role in the degree of presence of
-visual features as lognormal distribution produce the lowest power for
-both non-linearity and heteroskedasticity patterns. A symmetric fitted
-value distribution can better reveal the underlying shape.
-
-Although it is unusual to consider fitted value distribution or
-predictor distribution in power analysis, it is not so surprised that
-the power of conventional tests vary because of fitted value
-distribution. For instance, the power of the RESET test which is
-effectively a F test depends on the realizations of the additional
-predictors included in the auxiliary regression equation
-\citep{jamshidian2007study, olvera2019relationship, zhang2018practical}.
-And those additional predictors are linear transformation of random
-vectors following certain predictor distributions. What is truly
-unexpected is the huge variability in power of conventional tests
-compared to visual tests. The discrete uniform distribution constantly
-produce the highest power while the normal distribution and the
-lognormal distribution produce the lowest power for non-linearity and
-heteroskedasticity pattern respectively. The difference in power
-corresponding to different distributions peak at around 90\% when the
-effect size is at a moderate level. This huge difference is undesirable
-as is not uncommon for analysts to omit the factor of predictor
-distribution in power calculation and sample size calculation.
+fitted value distribution on the power of conventional (top row) and
+visual (bottom row) tests for both the non-linearity (left column) and
+heteroskedasticity (right column) patterns. For conventional tests, only
+the power curves of appropriate tests which are RESET tests and
+Breusch-Pagan tests for non-linearity pattern and heteroskedasticity
+pattern respectively are shown. For visual tests, note that, more
+evaluations on lineup with uniform fitted value distribution were
+collected, so to have a fair comparison, we randomly sample five from
+the total eleven evaluations to estimate the power curves. Re-sampling
+produces the multiple curves for the uniform, and provides an indication
+of the variation of the power.
+
+Perhaps surprisingly, the visual tests have more consistent power across
+the different fitted value distributions. For the non-linear pattern,
+there is almost no power difference. For the heteroskedastic pattern,
+uniform and discrete have higher power than normal and lognormal. The
+likely reason is that these latter two have less observations in the
+tails where the heteroskedastic pattern needs to be detected.
+
+The variation in power in the conventional tests is at first sight,
+shocking. However, it is discussed, albeit rarely, in the testing
+literature. See, for example, \citet{jamshidian2007study},
+\citet{olvera2019relationship} and \citet{zhang2018practical} which show
+derivations and use simulation to assess the effect of the observed
+distribution of the predictors on test power. The big differences in
+power curves that are seen in Figure
+\ref{fig:different-x-dist-poly-power} echoes the results reported in
+these articles.
 

### 2023-04-04 appendix.Rmd
update the table precentage

@@
                                   "Total", "%"), 
-                    booktabs = TRUE, label = "pronoun", caption = "Summary of pronoun distribution of subjects recuritted in this study.")
+                    booktabs = TRUE, 
+                    label = "pronoun", 
+                    caption = "Summary of pronoun distribution of subjects recuritted in this study.") %>%
+  kableExtra::kable_styling(latex_options = "scale_down")
 ```
@@
   rename(`Age group` = age_group) %>%
-  kableExtra::kable(format = "latex", booktabs = TRUE, label = "age-group", caption = "Summary of age distribution of subjects recuritted in this study.")
+  add_row(`Age group` = "",
+          `Period I` = sum(.$`Period I`),
+          `Period II` = sum(.$`Period II`),
+          `Period III` = sum(.$`Period III`),
+          Total = sum(.$Total)) %>%
+  mutate(`Period I %` = round(`Period I`/443*100, 1)) %>%
+  mutate(`Period II %` = round(`Period II`/443*100, 1)) %>%
+  mutate(`Period III %` = round(`Period III`/443*100, 1)) %>%
+  mutate(`Total %` = round(Total/443*100, 1)) %>%
+  select(`Age group`, 
+         `Period I`, `Period I %`, 
+         `Period II`, `Period II %`,
+         `Period III`, `Period III %`,
+         Total, `Total %`) %>%
+  kableExtra::kable(format = "latex", 
+                    col.names = c("Age group", 
+                                  "Period I", "%", 
+                                  "Period II", "%",
+                                  "Period III", "%",
+                                  "Total", "%"),
+                    booktabs = TRUE, 
+                    label = "age-group", 
+                    caption = "Summary of age distribution of subjects recuritted in this study.") %>%
+  kableExtra::kable_styling(latex_options = "scale_down")
 ```
@@
   rename(`Education` = education) %>%
-  add_row(`Education` = "",
+  add_row(Education = "",
           `Period I` = sum(.$`Period I`),
@@
           Total = sum(.$Total)) %>%
-  mutate(`Period I` = paste0(`Period I`, " (", round(`Period I`/443*100, 1), "%)")) %>%
-  mutate(`Period II` = paste0(`Period II`, " (", round(`Period II`/443*100, 1), "%)")) %>%
-  mutate(`Period III` = paste0(`Period III`, " (", round(`Period III`/443*100, 1), "%)")) %>%
-  mutate(Total = paste0(Total, " (", round(Total/443*100, 1), "%)")) %>%
-  kableExtra::kable(format = "latex", booktabs = TRUE, label = "education", caption = "Summary of education distribution of subjects recuritted in this study.")
+  mutate(`Period I %` = round(`Period I`/443*100, 1)) %>%
+  mutate(`Period II %` = round(`Period II`/443*100, 1)) %>%
+  mutate(`Period III %` = round(`Period III`/443*100, 1)) %>%
+  mutate(`Total %` = round(Total/443*100, 1)) %>%
+  select(Education, 
+         `Period I`, `Period I %`, 
+         `Period II`, `Period II %`,
+         `Period III`, `Period III %`,
+         Total, `Total %`) %>%
+  kableExtra::kable(format = "latex", 
+                    col.names = c("Education", 
+                                  "Period I", "%", 
+                                  "Period II", "%",
+                                  "Period III", "%",
+                                  "Total", "%"),
+                    booktabs = TRUE, 
+                    label = "education", 
+                    caption = "Summary of education distribution of subjects recuritted in this study.") %>%
+  kableExtra::kable_styling(latex_options = "scale_down")
 ```
@@
   rename(`Previous experience` = previous_experience) %>%
-  kableExtra::kable(format = "latex", booktabs = TRUE, label = "experience", caption = "Summary of previous experience distribution of subjects recuritted in this study.")
+  add_row(`Previous experience` = "",
+          `Period I` = sum(.$`Period I`),
+          `Period II` = sum(.$`Period II`),
+          `Period III` = sum(.$`Period III`),
+          Total = sum(.$Total)) %>%
+  mutate(`Period I %` = round(`Period I`/443*100, 1)) %>%
+  mutate(`Period II %` = round(`Period II`/443*100, 1)) %>%
+  mutate(`Period III %` = round(`Period III`/443*100, 1)) %>%
+  mutate(`Total %` = round(Total/443*100, 1)) %>%
+  select(`Previous experience`, 
+         `Period I`, `Period I %`, 
+         `Period II`, `Period II %`,
+         `Period III`, `Period III %`,
+         Total, `Total %`) %>%
+  kableExtra::kable(format = "latex", 
+                    col.names = c("Previous experience", 
+                                  "Period I", "%", 
+                                  "Period II", "%",
+                                  "Period III", "%",
+                                  "Total", "%"),
+                    booktabs = TRUE, 
+                    label = "experience", 
+                    caption = "Summary of previous experience distribution of subjects recuritted in this study.") %>%
+  kableExtra::kable_styling(latex_options = "scale_down")
 ```
@@
   geom_lv(aes(x = exp, y = prop_detect, fill = after_stat(LV)), varwidth = TRUE, k = 4) +
-  scale_alpha_discrete(range = c(0.7, 0)) +
   facet_wrap(~k, ncol = 5) +

### 2023-04-04 appendix.tex
update the table precentage

@@
 \centering
+\resizebox{\linewidth}{!}{
 \begin{tabular}[t]{lrrrrrrrr}
@@
 \bottomrule
-\end{tabular}
+\end{tabular}}
 \end{table}
@@
 \centering
-\begin{tabular}[t]{lrrrr}
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lrrrrrrrr}
 \toprule
-Age group & Period I & Period II & Period III & Total\\
+Age group & Period I & \% & Period II & \% & Period III & \% & Total & \%\\
 \midrule
-18-24 & 83 & 86 & 51 & 220\\
-25-39 & 69 & 63 & 63 & 195\\
-40-54 & 6 & 8 & 6 & 20\\
-55-64 & 2 & 3 & 3 & 8\\
+18-24 & 83 & 18.7 & 86 & 19.4 & 51 & 11.5 & 220 & 49.7\\
+25-39 & 69 & 15.6 & 63 & 14.2 & 63 & 14.2 & 195 & 44.0\\
+40-54 & 6 & 1.4 & 8 & 1.8 & 6 & 1.4 & 20 & 4.5\\
+55-64 & 2 & 0.5 & 3 & 0.7 & 3 & 0.7 & 8 & 1.8\\
+ & 160 & 36.1 & 160 & 36.1 & 123 & 27.8 & 443 & 100.0\\
 \bottomrule
-\end{tabular}
+\end{tabular}}
 \end{table}
@@
 \centering
-\begin{tabular}[t]{lllll}
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lrrrrrrrr}
 \toprule
-Education & Period I & Period II & Period III & Total\\
+Education & Period I & \% & Period II & \% & Period III & \% & Total & \%\\
 \midrule
-High School or below & 41 (9.3\%) & 53 (12\%) & 33 (7.4\%) & 127 (28.7\%)\\
-Diploma and Bachelor Degree & 92 (20.8\%) & 79 (17.8\%) & 66 (14.9\%) & 237 (53.5\%)\\
-Honours Degree & 6 (1.4\%) & 15 (3.4\%) & 6 (1.4\%) & 27 (6.1\%)\\
-Masters Degree & 21 (4.7\%) & 13 (2.9\%) & 16 (3.6\%) & 50 (11.3\%)\\
-Doctoral Degree & 0 (0\%) & 0 (0\%) & 2 (0.5\%) & 2 (0.5\%)\\
+High School or below & 41 & 9.3 & 53 & 12.0 & 33 & 7.4 & 127 & 28.7\\
+Diploma and Bachelor Degree & 92 & 20.8 & 79 & 17.8 & 66 & 14.9 & 237 & 53.5\\
+Honours Degree & 6 & 1.4 & 15 & 3.4 & 6 & 1.4 & 27 & 6.1\\
+Masters Degree & 21 & 4.7 & 13 & 2.9 & 16 & 3.6 & 50 & 11.3\\
+Doctoral Degree & 0 & 0.0 & 0 & 0.0 & 2 & 0.5 & 2 & 0.5\\
 \addlinespace
- & 160 (36.1\%) & 160 (36.1\%) & 123 (27.8\%) & 443 (100\%)\\
+ & 160 & 36.1 & 160 & 36.1 & 123 & 27.8 & 443 & 100.0\\
 \bottomrule
-\end{tabular}
+\end{tabular}}
 \end{table}
@@
 \centering
-\begin{tabular}[t]{lrrrr}
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lrrrrrrrr}
 \toprule
-Previous experience & Period I & Period II & Period III & Total\\
+Previous experience & Period I & \% & Period II & \% & Period III & \% & Total & \%\\
 \midrule
-No & 96 & 88 & 67 & 251\\
-Yes & 64 & 72 & 56 & 192\\
+No & 96 & 21.7 & 88 & 19.9 & 67 & 15.1 & 251 & 56.7\\
+Yes & 64 & 14.4 & 72 & 16.3 & 56 & 12.6 & 192 & 43.3\\
+ & 160 & 36.1 & 160 & 36.1 & 123 & 27.8 & 443 & 100.0\\
 \bottomrule
-\end{tabular}
+\end{tabular}}
 \end{table}

### 2023-04-04 appendix.Rmd
Modify the scale for the p-value cmp plot.
 Also draw an area for adj. p-value < 0.05

@@
 ```{r sensitivity, fig.cap = "Change of $p$-values with $\\hat{\\alpha}$ multiplied by $0.5$, $0.75$, $0.875$, $1.125$, $1.25$ and $1.5$.
-The vertical dashed line is to indicate $p\\text{-value} = 0.05$.
+The vertical dashed line indicates $p\\text{-value} = 0.05$.
+The area coloured in red indicates $\\text{adjusted }p\\text{-value} < 0.05$.
 The x-axis is drawn on logarithmic scale.
@@
 The trend is the opposite for multiplier greater than $1$, but the difference will eventually reach $0$."}
+
+smallest_p_value <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(x_dist == "uniform") %>%
+  pull(p_value) %>%
+  min()
+
 upper_lower_p %>%
@@
   geom_hline(yintercept = 0) +
-  geom_vline(xintercept = 0.05, col = "red", linetype = 2) +
+  geom_vline(xintercept = 0.05, col = "black", linetype = 2, alpha = 0.7) +
+  geom_ribbon(data = expand.grid(x = 10^seq(-10, 0, 0.01),
+                                y = seq(-0.1, 0.1, 0.0001)) %>%
+              filter(x + y < 0.05) %>%
+              group_by(x) %>%
+              summarise(y = max(y)),
+            aes(x, ymax = y, ymin = -0.06),
+            fill = "red",
+            alpha = 0.2) +
+  annotate("text", 
+           x = 0.0005, 
+           y = -0.03, 
+           label = '"Adjusted"~italic(p)*"-value"~"< 5%"', 
+           parse = TRUE,
+           col = "red") +
   geom_point(aes(p_value, diff_p_value), alpha = 0.2) +
-  scale_x_log10() +
   facet_wrap(~modified) +
@@
   xlab(expression(italic(p)*"-value")) +
-  theme_light()
-
+  theme_light() +
+  scale_x_continuous(breaks = c(0.0001, 0.001, 0.01, 0.1, 1),
+                     minor_breaks = NULL,
+                     trans = "log10",
+                     labels = c("0.01%", "0.1%", "1%", "10%", "100%"),
+                     limits = c(smallest_p_value, 1)) +
+  scale_y_continuous(limits = c(-0.06, 0.06))
 ```

### 2023-04-04 appendix.Rmd
fix notation in ES formula

@@
 
-Let $\boldsymbol{X}_a = (\boldsymbol{1}, \boldsymbol{X})$ denotes the $p$ regressors with $n$ observations, $\boldsymbol{R}_a = \boldsymbol{I} -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ denotes the residual operator, and let $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0},\sigma^2\boldsymbol{I})$ denotes the error.
-Using the Frisch–Waugh–Lovell theorem, residuals $\boldsymbol{e} = \boldsymbol{R}_a\boldsymbol{\varepsilon}$.
-Because $rank(\boldsymbol{R}_a) = n - p < n$, $e$ follows a degenerate multivariate normal distribution and does not have a density.
+Let $\boldsymbol{X}$ denotes the $p + 1$ regressors with $n$ observations, $\boldsymbol{b} = (\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\boldsymbol{y}$ denotes the OLS solution, $\boldsymbol{R} = \boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ denotes the residual operator, and let $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0},\sigma^2\boldsymbol{I})$ denotes the error.
+The residual vector
+
+\begin{align*}
+\boldsymbol{e} &= \boldsymbol{y} - \boldsymbol{X}\boldsymbol{b} \\
+               &= \boldsymbol{y} - \boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\boldsymbol{y} \\
+               &= (\boldsymbol{I} -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}')\boldsymbol{y} \\
+               &= \boldsymbol{R}\boldsymbol{y} \\
+               &= \boldsymbol{R}(\boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon}) \\
+               &= \boldsymbol{R}\boldsymbol{\varepsilon}.
+\end{align*}
+
+Because $rank(\boldsymbol{R}) = n - p - 1 < n$, $\boldsymbol{e}$ follows a degenerate multivariate normal distribution and does not have a density.
 Since the Kullback-Leibler divergence requires a proper density function, we need to simplify the covariance matrix of $\boldsymbol{e}$ by setting all the off-diagonal elements to 0.
-Then, the residuals will assumed to follow $N(\boldsymbol{0}, diag(\boldsymbol{R}_a\sigma^2))$ under the null hypothesis that the model is correctly specified.
-If the model is however misspecified due to omitted variables $\boldsymbol{Z}$, or a non-constant variance $\boldsymbol{V}$, the distribution of residuals can be derived as $N(\boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}_a\sigma^2))$ and $N(\boldsymbol{0}, diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a'))$ respectively.
+Then, the residuals will assumed to follow $N(\boldsymbol{0}, diag(\boldsymbol{R}\sigma^2))$ under the null hypothesis that the model is correctly specified.
+If the model is however misspecified due to omitted variables $\boldsymbol{Z}$, or a non-constant variance $\boldsymbol{V}$, the distribution of residuals can be derived as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}\sigma^2))$ and $N(\boldsymbol{0}, diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'))$ respectively.
 
@@
 \label{eq:effect-size}
-KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}\boldsymbol{\mu}_z\right).
+KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|diag(\boldsymbol{R}\sigma^2)|} - n + tr(diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}diag(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\boldsymbol{\mu}_z\right).
 \end{equation}
@@
 
-$$KL = \frac{1}{2}\left(\boldsymbol{\mu}_z^{T}(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right).$$
+$$KL = \frac{1}{2}\left(\boldsymbol{\mu}_z^{T}(diag(\boldsymbol{R}\sigma^2))^{-1}\boldsymbol{\mu}_z\right).$$
 
@@
 
-$$KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2))\right).$$
+$$KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|diag(\boldsymbol{R}\sigma^2)|} - n + tr(diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}diag(\boldsymbol{R}\sigma^2))\right).$$
 Since we assume $\sigma = 1$ for the heteroskedasticity model, the final form of the formula is
 
-$$KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right).$$
+$$KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|diag(\boldsymbol{R})|} - n + tr(diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}diag(\boldsymbol{R}))\right).$$
 
@@
             alpha = 0.2) +
-  annotate("text", 
-           x = 0.0005, 
-           y = -0.03, 
-           label = '"Adjusted"~italic(p)*"-value"~"< 5%"', 
-           parse = TRUE,
-           col = "red") +
   geom_point(aes(p_value, diff_p_value), alpha = 0.2) +

### 2023-04-04 appendix.tex
Modify the scale for the p-value cmp plot.
 Also draw an area for adj. p-value < 0.05
fix notation in ES formula

@@
 
-Let \(\boldsymbol{X}_a = (\boldsymbol{1}, \boldsymbol{X})\) denotes the
-\(p\) regressors with \(n\) observations,
-\(\boldsymbol{R}_a = \boldsymbol{I} -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\)
+Let \(\boldsymbol{X}\) denotes the \(p + 1\) regressors with \(n\)
+observations,
+\(\boldsymbol{b} = (\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\boldsymbol{y}\)
+denotes the OLS solution,
+\(\boldsymbol{R} = \boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\)
 denotes the residual operator, and let
 \(\boldsymbol{\varepsilon} \sim N(\boldsymbol{0},\sigma^2\boldsymbol{I})\)
-denotes the error. Using the Frisch--Waugh--Lovell theorem, residuals
-\(\boldsymbol{e} = \boldsymbol{R}_a\boldsymbol{\varepsilon}\). Because
-\(rank(\boldsymbol{R}_a) = n - p < n\), \(e\) follows a degenerate
-multivariate normal distribution and does not have a density. Since the
-Kullback-Leibler divergence requires a proper density function, we need
-to simplify the covariance matrix of \(\boldsymbol{e}\) by setting all
-the off-diagonal elements to 0. Then, the residuals will assumed to
-follow \(N(\boldsymbol{0}, diag(\boldsymbol{R}_a\sigma^2))\) under the
-null hypothesis that the model is correctly specified. If the model is
+denotes the error. The residual vector
+
+\begin{align*}
+\boldsymbol{e} &= \boldsymbol{y} - \boldsymbol{X}\boldsymbol{b} \\
+               &= \boldsymbol{y} - \boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\boldsymbol{y} \\
+               &= (\boldsymbol{I} -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}')\boldsymbol{y} \\
+               &= \boldsymbol{R}\boldsymbol{y} \\
+               &= \boldsymbol{R}(\boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon}) \\
+               &= \boldsymbol{R}\boldsymbol{\varepsilon}.
+\end{align*}
+
+Because \(rank(\boldsymbol{R}) = n - p - 1 < n\), \(\boldsymbol{e}\)
+follows a degenerate multivariate normal distribution and does not have
+a density. Since the Kullback-Leibler divergence requires a proper
+density function, we need to simplify the covariance matrix of
+\(\boldsymbol{e}\) by setting all the off-diagonal elements to 0. Then,
+the residuals will assumed to follow
+\(N(\boldsymbol{0}, diag(\boldsymbol{R}\sigma^2))\) under the null
+hypothesis that the model is correctly specified. If the model is
 however misspecified due to omitted variables \(\boldsymbol{Z}\), or a
@@
 can be derived as
-\(N(\boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}_a\sigma^2))\)
+\(N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}\sigma^2))\)
 and
-\(N(\boldsymbol{0}, diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a'))\)
+\(N(\boldsymbol{0}, diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'))\)
 respectively.
@@
 
-\[KL = \frac{1}{2}\left(\boldsymbol{\mu}_z^{T}(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right).\]
+\[KL = \frac{1}{2}\left(\boldsymbol{\mu}_z^{T}(diag(\boldsymbol{R}\sigma^2))^{-1}\boldsymbol{\mu}_z\right).\]
 
@@
 
-\[KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2))\right).\]
+\[KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|diag(\boldsymbol{R}\sigma^2)|} - n + tr(diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}diag(\boldsymbol{R}\sigma^2))\right).\]
 Since we assume \(\sigma = 1\) for the heteroskedasticity model, the
@@
 
-\[KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right).\]
+\[KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|diag(\boldsymbol{R})|} - n + tr(diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}diag(\boldsymbol{R}))\right).\]
 
@@
 \caption{Change of $p$-values with $\hat{\alpha}$ multiplied by $0.5$, $0.75$, $0.875$, $1.125$, $1.25$ and $1.5$.
-The vertical dashed line is to indicate $p\text{-value} = 0.05$.
+The vertical dashed line indicates $p\text{-value} = 0.05$.
+The area coloured in red indicates $\text{adjusted }p\text{-value} < 0.05$.
 The x-axis is drawn on logarithmic scale.

### 2023-04-04 paper.Rmd
finished revising section 5.6
fix notation in ES formula

@@
 
-$$E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right)$$
+$$E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}\sigma^2))^{-1}\boldsymbol{\mu}_z\right)$$
 \noindent where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,
-$\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a$ is the residual operator, $\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ is the hat matrix, $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients, and $\sigma^2\boldsymbol{I}$ is the assumed covariance matrix of the error term when $H_0$ is true.
+$\boldsymbol{R} = \boldsymbol{I}_n - \boldsymbol{H}$ is the residual operator, $\boldsymbol{H} = \boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the hat matrix, $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients, and $\sigma^2\boldsymbol{I}$ is the assumed covariance matrix of the error term when $H_0$ is true.
 
@@
 
-$$E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)$$
+$$E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|diag(\boldsymbol{R})|} - n + tr(diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}diag(\boldsymbol{R}))\right)$$
 \noindent where $\boldsymbol{V}$ is the actual covariance matrix of the error term. (Derivations for these equations are provided in the Appendix.)
@@
 See, for example, @jamshidian2007study, @olvera2019relationship and  @zhang2018practical which show derivations and use simulation to assess the effect of the observed distribution of the predictors on test power.
-The big differences in power curves that are seen in Figure \ref{fig:different-x-dist-poly-power} echoes the results reported in these articles.
+The big differences in power curves that are seen in Figure \ref{fig:different-x-dist-poly-power} is echoed in the results reported in these articles.
 

### 2023-04-04 paper.tex
finished revising section 5.6
fix notation in ES formula

@@
 
-\[E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right)\]
+\[E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}\sigma^2))^{-1}\boldsymbol{\mu}_z\right)\]
 \noindent where \(diag(.)\) is the diagonal matrix constructed from the
 diagonal elements of a matrix,
-\(\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a\) is the
-residual operator,
-\(\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'\)
+\(\boldsymbol{R} = \boldsymbol{I}_n - \boldsymbol{H}\) is the residual
+operator,
+\(\boldsymbol{H} = \boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\)
 is the hat matrix,
-\(\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z\)
+\(\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z\)
 is the expected values of residuals with \(\boldsymbol{Z}\) be any
@@
 
-\[E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)\]
+\[E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|diag(\boldsymbol{R})|} - n + tr(diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}diag(\boldsymbol{R}))\right)\]
 \noindent where \(\boldsymbol{V}\) is the actual covariance matrix of
@@
 power curves that are seen in Figure
-\ref{fig:different-x-dist-poly-power} echoes the results reported in
-these articles.
+\ref{fig:different-x-dist-poly-power} is echoed in the results reported
+in these articles.
 

### 2023-04-04 paper.Rmd
small typos in section 5.6

@@
 
-In regression predictions are conditional on the observed values of the predictors, that is, the conditional mean of the dependent variable Y given the value of the independent variable X, $E(Y|X)$.
+In regression analysis, predictions are conditional on the observed values of the predictors, that is, the conditional mean of the dependent variable $Y$ given the value of the independent variable $X$, $E(Y|X)$.
 This is an often forgotten element of regression analysis but it is important.
-Where X is observed, particularly the distribution of the X values in the sample, or consequently $\hat{Y}$, may affect the ability to read any patterns in the residual plots.
+Where $X$ is observed, particularly the distribution of the $X$ values in the sample, or consequently $\hat{Y}$, may affect the ability to read any patterns in the residual plots.
 This experiment was constructed to assess this, based on four different distributions of fitted values: uniform, normal, discrete and lognormal (skewed).
@@
 
-```{r different-x-dist-poly-power, fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for conventional tests and visual tests.
-Power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and Breusch-Pagan tests respectively.
+```{r different-x-dist-poly-power, fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for conventional and visual tests for non-linearity and heteroskedasticity patterns.
+The power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and Breusch-Pagan tests, respectively.
 Power curves of visual tests are estimated using five evaluations on each lineup.
-For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations.
-The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines.
-The fitted value distribution has greater impact on the power of conventional tests than visual tests.
-Meanwhile, uneven distributions including normal and lognormal distributions genearlly produce lower power.", dev = 'png', dpi = 450, fig.pos="t"}
+For lineups with a uniform fitted value distribution, the five evaluations are repeatedly sampled from the total eleven evaluations to give multiple power curves (grey).
+Surprisingly, the fitted value distribution has produces more variability in the power of conventional tests than visual tests.
+Uneven distributions, normal and lognormal distributions, tend to yield lower power.", dev = 'png', dpi = 450, fig.pos="t"}
 

### 2023-04-04 paper.tex
small typos in section 5.6

@@
 
-In regression predictions are conditional on the observed values of the
-predictors, that is, the conditional mean of the dependent variable Y
-given the value of the independent variable X, \(E(Y|X)\). This is an
-often forgotten element of regression analysis but it is important.
-Where X is observed, particularly the distribution of the X values in
-the sample, or consequently \(\hat{Y}\), may affect the ability to read
-any patterns in the residual plots. This experiment was constructed to
-assess this, based on four different distributions of fitted values:
-uniform, normal, discrete and lognormal (skewed). We would expect that
-if \(\hat{Y}\) has a uniform distribution, this would make it easier to
-read the relationship with the residuals.
+In regression analysis, predictions are conditional on the observed
+values of the predictors, that is, the conditional mean of the dependent
+variable \(Y\) given the value of the independent variable \(X\),
+\(E(Y|X)\). This is an often forgotten element of regression analysis
+but it is important. Where \(X\) is observed, particularly the
+distribution of the \(X\) values in the sample, or consequently
+\(\hat{Y}\), may affect the ability to read any patterns in the residual
+plots. This experiment was constructed to assess this, based on four
+different distributions of fitted values: uniform, normal, discrete and
+lognormal (skewed). We would expect that if \(\hat{Y}\) has a uniform
+distribution, this would make it easier to read the relationship with
+the residuals.
 

### 2023-04-04 paper.tex
caption for Fig 17

@@
 
-\caption{Comparison of power on lineups with different fitted value distributions for conventional tests and visual tests.
-Power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and Breusch-Pagan tests respectively.
+\caption{Comparison of power on lineups with different fitted value distributions for conventional and visual tests for non-linearity and heteroskedasticity patterns.
+The power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and Breusch-Pagan tests, respectively.
 Power curves of visual tests are estimated using five evaluations on each lineup.
-For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations.
-The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines.
-The fitted value distribution has greater impact on the power of conventional tests than visual tests.
-Meanwhile, uneven distributions including normal and lognormal distributions genearlly produce lower power.}\label{fig:different-x-dist-poly-power}
+For lineups with a uniform fitted value distribution, the five evaluations are repeatedly sampled from the total eleven evaluations to give multiple power curves (grey).
+Surprisingly, the fitted value distribution has produces more variability in the power of conventional tests than visual tests.
+Uneven distributions, normal and lognormal distributions, tend to yield lower power.}\label{fig:different-x-dist-poly-power}
 \end{figure}

### 2023-04-04 paper.Rmd
section 5.5 revised

@@
 
-We have also investigated the impact of different heteroskedasticity shapes on power of conventional tests and visual tests.
-In theory, the "left-triangle" and the "right-triangle" shapes are functionally identical from the point of view of a Breusch-Pagan test.
-As shown in Figure \ref{fig:heter-power-uniform-a}, this is indeed the case where little difference between the power curves can be perceived.
-Similarly, visual tests should have the same power in detecting these two shapes if they are equally likely to be identified.
-However, it can be observed from Figure \ref{fig:heter-power-uniform-a} that the power curve of the "left-triangle" shape is higher than the one of the "right-triangle" shape, indicating a potential favour of orientation by human, which is worth to be explored in future studies.
-Besides, the chance of detecting the "butterfly" shape are higher than other two shapes in both conventional and visual testing.
-
-```{r heter-power-uniform-a, fig.height = 4, fig.cap = 'Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$.
-The visual test has an asymmetric power in testing the "left-triangle" shape and "right-triangle" shape, while these two shapes are equivalent in conventional testing.'}
+Figure \ref{fig:heter-power-uniform-a} examines the impact of the shape of the heteroskedasticity on the power of of both tests.
+The butterfly shape has higher power on both types of tests.
+The "left-triangle" and the "right-triangle" shapes are functionally identical, and this is observed for the conventional test, where the power curves are identical.
+Interestingly there is a difference for the visual test, where the power curve of the "left-triangle" shape is slightly higher than that of the "right-triangle" shape.
+This indicates a bias in perceiving heteroskedasticity depending on the direction.
+This would be worth investigating further.
+
+```{r heter-power-uniform-a, fig.height = 4, fig.cap = 'The effect of heteroskedasticity shape (parameter $a$) on the power of conventional and visual tests.
+The butterfly has higher power in both tests.
+Curiously, the visual test has slightly higher power for the "left-triangle" than the "right-triangle" shape, when it would be expected that they should be identical, which is observed in conventional testing.'}
 

### 2023-04-04 paper.tex
section 5.5 revised

@@
 
-We have also investigated the impact of different heteroskedasticity
-shapes on power of conventional tests and visual tests. In theory, the
-``left-triangle'' and the ``right-triangle'' shapes are functionally
-identical from the point of view of a Breusch-Pagan test. As shown in
-Figure \ref{fig:heter-power-uniform-a}, this is indeed the case where
-little difference between the power curves can be perceived. Similarly,
-visual tests should have the same power in detecting these two shapes if
-they are equally likely to be identified. However, it can be observed
-from Figure \ref{fig:heter-power-uniform-a} that the power curve of the
-``left-triangle'' shape is higher than the one of the ``right-triangle''
-shape, indicating a potential favour of orientation by human, which is
-worth to be explored in future studies. Besides, the chance of detecting
-the ``butterfly'' shape are higher than other two shapes in both
-conventional and visual testing.
+Figure \ref{fig:heter-power-uniform-a} examines the impact of the shape
+of the heteroskedasticity on the power of of both tests. The butterfly
+shape has higher power on both types of tests. The ``left-triangle'' and
+the ``right-triangle'' shapes are functionally identical, and this is
+observed for the conventional test, where the power curves are
+identical. Interestingly there is a difference for the visual test,
+where the power curve of the ``left-triangle'' shape is slightly higher
+than that of the ``right-triangle'' shape. This indicates a bias in
+perceiving heteroskedasticity depending on the direction. This would be
+worth investigating further.
 
@@
 
-\caption{Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$.
-The visual test has an asymmetric power in testing the "left-triangle" shape and "right-triangle" shape, while these two shapes are equivalent in conventional testing.}\label{fig:heter-power-uniform-a}
+\caption{The effect of heteroskedasticity shape (parameter $a$) on the power of conventional and visual tests.
+The butterfly has higher power in both tests.
+Curiously, the visual test has slightly higher power for the "left-triangle" than the "right-triangle" shape, when it would be expected that they should be identical, which is observed in conventional testing.}\label{fig:heter-power-uniform-a}
 \end{figure}

### 2023-04-04 paper.Rmd
section 5.4 revised

@@
 
-## Effect of shape of non-linearity
-
-A primary factor contributes to the non-linearity model is the shape of the non-linearity pattern.
-According to Figure \ref{fig:poly-power-uniform-j}, conventional tests have higher power in testing shapes constructed from lower orders of Hermite polynomials.
-But the chance of detecting the "triple-U" shape drops significantly compared to other shapes.
+## Effect of amount of non-linearity
+
+The order of the polynomial is a primary factor contributing to the pattern produced by the non-linearity model.
+Figure \ref{fig:poly-power-uniform-j} explores the relationship between polynomial order and power of the tests.
+The conventional tests have higher power for lower orders of Hermite polynomials, and it drops substantially for the "triple-U" shape.
 To understand why this is, one needs to return to the way the RESET test is applied.
 It requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four [@ramsey_tests_1969].
-However, the "triple-U" shape constructed from the Hermite polynomials use power up to 18.
-If the RESET test had been applied using a higher power no less than six, the power curve of "triple-U" shape will be only slight lower than the power curve of "M" shape.
-The recommendation of the polynomial power for the RESET test should be revised, perhaps.
-This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data generating process in order to set the parameters for the test.
-
-For visual tests, based on the orders of the Hermite polynomials, we expect the "U" shape will be the easiest one to be detected by subjects followed by the "S", "M" and "triple-U" shape.
-From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves are mostly aligned with our expectation, except for the "M" shape, which is as easy to be detected as the "S" shape.
-This implies, unlike the conventional test, the visibility of the shape do not strictly follow the degree of the polynomials.
-
-```{r poly-power-uniform-j, fig.height = 4, fig.cap = 'Power of conventional tests and visual tests on lineups containing four different non-linearity shapes.
-Power curves with higher order of non-linearity are drawn with deeper colours.
-The default RESET tests under-perform significantly in detecting "triple-U" shape.
-To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test.
-But this means the order needs to be known prior to testing.'}
+However, the "triple-U" shape is constructed from the Hermite polynomials using power up to 18.
+If the RESET test had been applied using a higher power no less than six, the power curve of "triple-U" shape will be closer to other power curves.
+This illustrates the sensitivity of the conventional test to the parameter choice, and highlights a limitation that it helps to know the data generating process to set the parameters for the test, which is unrealistic.
+However, we examined this in more detail (see Appendix) and found that there is no harm for setting the parameter higher than four on the tests' operation for lower order polynomial shapes.
+Using a parameter value of six, instead of four, yields higher power regardless of generating process, and would be recommended.
+
+For visual tests, we expect the "U" shape to be detected more readily, followed by the "S", "M" and "triple-U" shape.
+From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves mostly align with these expectations, except for the "M" shape, which is as easy to be detected as the "S" shape.
+This suggests a benefit of  the visual test, the knowing the shape ahead of time is not needed for its application.
+
+```{r poly-power-uniform-j, fig.height = 4, fig.cap = 'The effect of the order of the polynomial on the power of conventional and visual tests.
+Deeper colour indicates higher order.
+The default RESET tests under-performs significantly in detecting the "triple-U" shape.
+To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test, but this higher than the recommended value.'}
 

### 2023-04-04 paper.tex
section 5.4 revised

@@
 
-\hypertarget{effect-of-shape-of-non-linearity}{%
-\subsection{Effect of shape of
-non-linearity}\label{effect-of-shape-of-non-linearity}}
-
-A primary factor contributes to the non-linearity model is the shape of
-the non-linearity pattern. According to Figure
-\ref{fig:poly-power-uniform-j}, conventional tests have higher power in
-testing shapes constructed from lower orders of Hermite polynomials. But
-the chance of detecting the ``triple-U'' shape drops significantly
-compared to other shapes. To understand why this is, one needs to return
-to the way the RESET test is applied. It requires a parameter indicating
-degree of fitted values to test for, and the recommendation is to
-generically use four \citep{ramsey_tests_1969}. However, the
-``triple-U'' shape constructed from the Hermite polynomials use power up
-to 18. If the RESET test had been applied using a higher power no less
-than six, the power curve of ``triple-U'' shape will be only slight
-lower than the power curve of ``M'' shape. The recommendation of the
-polynomial power for the RESET test should be revised, perhaps. This
-illustrates the sensitivity of conventional testing to the parameters,
-and it also points to a limitation that one needs to know the data
-generating process in order to set the parameters for the test.
-
-For visual tests, based on the orders of the Hermite polynomials, we
-expect the ``U'' shape will be the easiest one to be detected by
-subjects followed by the ``S'', ``M'' and ``triple-U'' shape. From
-Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power
-curves are mostly aligned with our expectation, except for the ``M''
-shape, which is as easy to be detected as the ``S'' shape. This implies,
-unlike the conventional test, the visibility of the shape do not
-strictly follow the degree of the polynomials.
+\hypertarget{effect-of-amount-of-non-linearity}{%
+\subsection{Effect of amount of
+non-linearity}\label{effect-of-amount-of-non-linearity}}
+
+The order of the polynomial is a primary factor contributing to the
+pattern produced by the non-linearity model. Figure
+\ref{fig:poly-power-uniform-j} explores the relationship between
+polynomial order and power of the tests. The conventional tests have
+higher power for lower orders of Hermite polynomials, and it drops
+substantially for the ``triple-U'' shape. To understand why this is, one
+needs to return to the way the RESET test is applied. It requires a
+parameter indicating degree of fitted values to test for, and the
+recommendation is to generically use four \citep{ramsey_tests_1969}.
+However, the ``triple-U'' shape is constructed from the Hermite
+polynomials using power up to 18. If the RESET test had been applied
+using a higher power no less than six, the power curve of ``triple-U''
+shape will be closer to other power curves. This illustrates the
+sensitivity of the conventional test to the parameter choice, and
+highlights a limitation that it helps to know the data generating
+process to set the parameters for the test, which is unrealistic.
+However, we examined this in more detail (see Appendix) and found that
+there is no harm for setting the parameter higher than four on the
+tests' operation for lower order polynomial shapes. Using a parameter
+value of six, instead of four, yields higher power regardless of
+generating process, and would be recommended.
+
+For visual tests, we expect the ``U'' shape to be detected more readily,
+followed by the ``S'', ``M'' and ``triple-U'' shape. From Figure
+\ref{fig:poly-power-uniform-j}, it can be observed that the power curves
+mostly align with these expectations, except for the ``M'' shape, which
+is as easy to be detected as the ``S'' shape. This suggests a benefit of
+the visual test, the knowing the shape ahead of time is not needed for
+its application.
 
@@
 
-\caption{Power of conventional tests and visual tests on lineups containing four different non-linearity shapes.
-Power curves with higher order of non-linearity are drawn with deeper colours.
-The default RESET tests under-perform significantly in detecting "triple-U" shape.
-To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test.
-But this means the order needs to be known prior to testing.}\label{fig:poly-power-uniform-j}
+\caption{The effect of the order of the polynomial on the power of conventional and visual tests.
+Deeper colour indicates higher order.
+The default RESET tests under-performs significantly in detecting the "triple-U" shape.
+To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test, but this higher than the recommended value.}\label{fig:poly-power-uniform-j}
 \end{figure}

### 2023-04-04 appendix.Rmd
update caption for sensitivity analysis

@@
 
-\begin{equation}
+\begin{equation*}
 \label{eq:effect-size}
 KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|diag(\boldsymbol{R}\sigma^2)|} - n + tr(diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}diag(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\boldsymbol{\mu}_z\right).
-\end{equation}
+\end{equation*}
 
@@
 The x-axis is drawn on logarithmic scale.
-For multiplier smaller than $1$, the adjusted $p$-value will incrase then decrase as $p$-value increases.
-The trend is the opposite for multiplier greater than $1$, but the difference will eventually reach $0$."}
+For multipliers smaller than 1, the adjusted $p\\text{-value}$ will initially increase and decline when the $p\\text{-value}$ increases.
+The trend is opposite with multipliers greater than $1$, but the difference eventually reaches $0$."}
 

### 2023-04-04 appendix.tex
update caption for sensitivity analysis

@@
 The x-axis is drawn on logarithmic scale.
-For multiplier smaller than $1$, the adjusted $p$-value will incrase then decrase as $p$-value increases.
-The trend is the opposite for multiplier greater than $1$, but the difference will eventually reach $0$.}\label{fig:sensitivity}
+For multipliers smaller than 1, the adjusted $p\text{-value}$ will initially increase and decline when the $p\text{-value}$ increases.
+The trend is opposite with multipliers greater than $1$, but the difference eventually reaches $0$.}\label{fig:sensitivity}
 \end{figure}

### 2023-04-04 appendix.Rmd
modify the alpha table

@@
 ```{r sensitivity, fig.cap = "Change of $p$-values with $\\hat{\\alpha}$ multiplied by $0.5$, $0.75$, $0.875$, $1.125$, $1.25$ and $1.5$.
+Only lineups with uniform fitted value distribution is used.
 The vertical dashed line indicates $p\\text{-value} = 0.05$.
@@
   mutate(case = ifelse(!adjusted_reject, "reject to not reject", "not reject to reject")) %>%
-  rename(fraction = modified) %>%
-  mutate(n = paste0(round(n/279 * 100, 2))) %>%
+  rename(multiplier = modified) %>%
   select(-adjusted_reject, -reject) %>%
   pivot_wider(names_from = case, values_from = n) %>%
-  kableExtra::kable(format = "latex", booktabs = TRUE, label = "sensitivity", 
-                    escape = FALSE,
+  mutate(reject_not_reject_per = round(`reject to not reject`/279*100, 2)) %>%
+  mutate(not_reject_reject_per = round(`not reject to reject`/279*100, 2)) %>%
+  select(multiplier,
+         `reject to not reject`,
+         reject_not_reject_per,
+         `not reject to reject`,
+         not_reject_reject_per) %>%
+  kableExtra::kable(format = "latex", 
+                    booktabs = TRUE,
+                    col.names = c("Multiplier",
+                                  "Reject to not reject",
+                                  "%",
+                                  "Not reject to reject",
+                                  "%"),
+                    label = "sensitivity", 
                     caption = "Examining how decisions might change if $\\hat{\\alpha}$ was different.
-Percentage of lineups that where the $p$-value would switch to above or below 0.05, when $\\hat{\\alpha}$ is varied by a fractional amount.")
+Percentage of lineups that where the $p$-value would switch to above or below 0.05, when $\\hat{\\alpha}$ is multiplied by a multiplier.") %>%
+  kableExtra::kable_styling(latex_options = "scale_down")
 ```
@@
             size = 1) +
-  labs(col = "N. evals") +
+  labs(col = "#Evaluations") +
+  ylab("Power") +
   xlab(expression(log[e] (Effect_size))) +

### 2023-04-04 appendix.tex
modify the alpha table

@@
 \caption{Change of $p$-values with $\hat{\alpha}$ multiplied by $0.5$, $0.75$, $0.875$, $1.125$, $1.25$ and $1.5$.
+Only lineups with uniform fitted value distribution is used.
 The vertical dashed line indicates $p\text{-value} = 0.05$.
@@
 \caption{\label{tab:sensitivity}Examining how decisions might change if $\hat{\alpha}$ was different.
-Percentage of lineups that where the $p$-value would switch to above or below 0.05, when $\hat{\alpha}$ is varied by a fractional amount.}
+Percentage of lineups that where the $p$-value would switch to above or below 0.05, when $\hat{\alpha}$ is multiplied by a multiplier.}
 \centering
-\begin{tabular}[t]{rll}
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{rrrrr}
 \toprule
-fraction & reject to not reject & not reject to reject\\
+Multiplier & Reject to not reject & \% & Not reject to reject & \%\\
 \midrule
-0.500 & 2.51 & 0\\
-0.750 & 1.43 & 0\\
-0.875 & 1.08 & 0\\
-1.125 & 0 & 1.08\\
-1.250 & 0 & 1.43\\
+0.500 & 7 & 2.51 & 0 & 0.00\\
+0.750 & 4 & 1.43 & 0 & 0.00\\
+0.875 & 3 & 1.08 & 0 & 0.00\\
+1.125 & 0 & 0.00 & 3 & 1.08\\
+1.250 & 0 & 0.00 & 4 & 1.43\\
 \addlinespace
-1.500 & 0 & 1.79\\
+1.500 & 0 & 0.00 & 5 & 1.79\\
 \bottomrule
-\end{tabular}
+\end{tabular}}
 \end{table}

### 2023-04-04 README.Rmd
update README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
metadata <- rmarkdown::yaml_front_matter("paper.Rmd")
```

# `r metadata$title`

This repo contains the materials to fully reproduce the paper titled "`r metadata$title`" by `r knitr::combine_words(purrr::map_chr(metadata$author, ~.x$name))`.

## Citation

Please cite the paper as:

Li et al. (2023) `r metadata$title`. URL:  https://github.com/TengMCing/lineup_residual_diagnostics

### 2023-04-04 paper.Rmd
small section heading change
change affiliation for ET

@@
   - name: Emi Tanaka
-    affil: a
-    email: emi.tanaka@monash.edu
+    affil: a, b, c
+    email: emi.tanaka@anu.edu.au
   - name: Susan VanderPlas
-    affil: b
+    affil: d
     email: susan.vanderplas@unl.edu
@@
   - num: b
+    address: |
+      Biological Data Science Institute, Australian National Institute, Acton, ACT, Australia
+  - num: c
+    address: |
+      Research School of Finance, Actuarial Studies and Statistics, Australian National Institute, Acton, ACT, Australia
+  - num: d
     address: |
@@
 
-## Power comparison of different tests
+## Power comparison of the tests
 

### 2023-04-04 paper.Rmd
fix university name

--- 2023-04-04 paper.Rmd
@@
   - num: b
+    address: |
+      Biological Data Science Institute, Australian National University, Acton, ACT, Australia
+  - num: c
+    address: |
+      Research School of Finance, Actuarial Studies and Statistics, Australian National University, Acton, ACT, Australia
+  - num: d
     address: |

### 2023-04-04 paper.Rmd
fixing narrative in Background section 3

@@
 
-One may argue that reading diagnostic plots is to some extent subjective and indecisive compared to those rigorous statistical procedures as it relies on graphical perception - human ability to interpret and decode the information embedded in graph [@cleveland_graphical_1984].
-Further, the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
-For instance, people
-over-interpret the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurrence [@roy_chowdhury_using_2015].
-
-Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots.
-Later, @buja_statistical_2009 proposed the lineup protocol as a visual test inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people.
-The protocol consists of $m$ randomly placed plots, where one plot is the data plot, and the remaining $m - 1$ null plots have the identical graphical procedure except the data has been replaced with data consistent with $H_0$, that the model is correctly specified.
+The examination of data plots to infer signal or patterns (or lack thereof) is fraught with variation in the human ability to interpret and decode the information embedded in graph [@cleveland_graphical_1984].
+Such examination thus feels there is a subjective nature to evaluate, say, diagnostic plots to infer appropriateness of statistical models, although arguably we would expect such evaluation should be done objectively and such human-to-human variation in "reading" data plots should be minimised.
+
+In practice, over-interpretation of a single plot is common.
+For instance, in @roy_chowdhury_using_2015, some over-interpreted the separation between gene groups in a two-dimensional projection of a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups.
+To mitigate this over-interpretation, @buja_statistical_2009 proposed a line-up protocol to assess plots in a manner analogous to the null hypothesis significance testing framework.
+More specifically, the protocol consists of $m$ randomly placed plots, where one plot is the data plot, and the remaining $m - 1$ null plots have the identical graphical procedure except the data has been replaced with data consistent with $H_0$, that the model is correctly specified.
 Then, an observer who have not seen the data plot will be asked to point out the most different plot from the lineup.

### 2023-04-05 paper.Rmd
clear up language for NHST + vis inf

@@
 For instance, in @roy_chowdhury_using_2015, some over-interpreted the separation between gene groups in a two-dimensional projection of a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups.
-To mitigate this over-interpretation, @buja_statistical_2009 proposed a line-up protocol to assess plots in a manner analogous to the null hypothesis significance testing framework.
-More specifically, the protocol consists of $m$ randomly placed plots, where one plot is the data plot, and the remaining $m - 1$ null plots have the identical graphical procedure except the data has been replaced with data consistent with $H_0$, that the model is correctly specified.
+To mitigate this over-interpretation, @buja_statistical_2009 proposed a line-up protocol to assess plots in a manner analogous to the null hypothesis significance testing (NHST) framework.
+More specifically, the protocol consists of $m$ randomly placed plots, where one plot is the data plot, and the remaining $m - 1$ plots, referred to as the _null plots_, have the identical graphical procedure as the data plot except the data is replaced with a data generation mechanism that is consistent with the null hypothesis, $H_0$.
 Then, an observer who have not seen the data plot will be asked to point out the most different plot from the lineup.
@@
 If one rejects $H_0$ as the observer correctly picks the data plot, then the Type I error of this test is $1/m$.
+This protocol requires apriori specification of $H_0$ (or at least a null data generating mechanism) much like the requirement of knowing the distribution of the test statistic in NHST.
 

### 2023-04-07 paper.Rmd
start on revising section 5

@@
 
-## Overview
-
-There are `r 160 * 18`, `r 150 * 18` and `r 133 * 18 - nrow(filter(vi_survey, null_lineup, !attention_check))` lineups evaluation made by 160, 150 and 133 subjects recruited for data collection periods I, II and III respectively.
-In the total of `r nrow(filter(vi_survey, !attention_check))` lineup evaluations, `r nrow(filter(vi_survey, type == "polynomial", !attention_check))` use lineups produced by the non-linearity model, and `r nrow(filter(vi_survey, type == "heteroskedasticity", !attention_check))` use lineups produced by the heteroskedasticity model.
-Besides, there are `r nrow(filter(vi_survey, attention_check))` attention checks and `r nrow(filter(vi_survey, null_lineup, !attention_check))` evaluations on null lineups needed for the estimate of $\alpha$ not included in the analysis.
-The collated dataset is provided in `vi_survey` of the `visage` `R` package.
-
-In the following analysis, lineups with uniform fitted values will be the focus.
-Visual patterns are more likely to be revealed under a uniform distribution.
-Additionally, we have collected extra evaluations on these lineups, which will result in a more reliable analysis.
-Analysis of lineups with other fitted value distributions can be found in Section \ref{effect-of-fitted-value-distributions}.
-
-## Power comparison of the tests
+Data collection resulted in a total of `r nrow(filter(vi_survey, !attention_check))` lineup evaluations, `r nrow(filter(vi_survey, type == "polynomial", !attention_check))` for the non-linearity model, and `r nrow(filter(vi_survey, type == "heteroskedasticity", !attention_check))` for the heteroskedasticity model.
+This was done in three collection periods, called I, II, III, with `r 160 * 18`, `r 150 * 18` and `r 133 * 18 - nrow(filter(vi_survey, null_lineup, !attention_check))` lineups evaluated by 160, 150 and 133 participants, respectively.
+There were `r nrow(filter(vi_survey, attention_check))` attention check lineups used to filter false participant attempts.
+To estimate $\alpha$ for calculating statistical significance there were `r nrow(filter(vi_survey, null_lineup, !attention_check))` evaluations of null lineups.
+Neither the attention checks nor null lineups were used for the subsequent analysis.
+The de-identified data, `vi_survey`,  is made available in the R package, `visage`.
+
+For clarity and consistency, the analysis in Sections \ref{power-analysis}-\ref{hetero-analysis}, uses only on results from lineups generated with uniform fitted values, a total of `r nrow(filter(vi_survey, x_dist == "uniform"))` lineups and XXX evaluation.
+More evaluations were made using the uniform fitted value distribution, and it avoids potentially confounding results of the analysis by combining values from different power curves.
+We also expect that visual patterns are more clearly visible when the fitted value distribution is uniform.
+Section \ref{effect-of-fitted-value-distributions} contains comparison of the effect of fitted value distribution on power.
+
+## Power comparison of the tests~\label{power-analysis}
 
@@
 
-## Comparison of test decisions based on $p$-values
+## Comparison of test decisions based on $p$-values~\label{p-value}
 
@@
 
-## Effect of amount of non-linearity
+## Effect of amount of non-linearity~\label{nonlin-analysis}
 
@@
 
-## Effect of shape of heteroskedasticity
+## Effect of shape of heteroskedasticity~\label{hetero-analysis}
 

### 2023-04-07 paper.tex
small section heading change
start on revising section 5

@@
 
-\author{\name{Weihao Li$^{a}$, Dianne Cook$^{a}$, Emi
-Tanaka$^{a}$, Susan VanderPlas$^{b}$}
+\author{\name{Weihao Li$^{a}$, Dianne Cook$^{a}$, Emi Tanaka$^{a, b,
+c}$, Susan VanderPlas$^{d}$}
 \affil{$^{a}$Department of Econometrics and Business Statistics, Monash
-University, Clayton, VIC, Australia; $^{b}$Department of Statistics,
-University of Nebraska, Lincoln, Nebraska, USA}
+University, Clayton, VIC, Australia; $^{b}$Biological Data Science
+Institute, Australian National University, Acton, ACT,
+Australia; $^{c}$Research School of Finance, Actuarial Studies and
+Statistics, Australian National University, Acton, ACT,
+Australia; $^{d}$Department of Statistics, University of Nebraska,
+Lincoln, Nebraska, USA}
 }
@@
 Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}, Emi
-Tanaka. Email: \href{mailto:emi.tanaka@monash.edu}{\nolinkurl{emi.tanaka@monash.edu}}, Susan
+Tanaka. Email: \href{mailto:emi.tanaka@anu.edu.au}{\nolinkurl{emi.tanaka@anu.edu.au}}, Susan
 VanderPlas. Email: \href{mailto:susan.vanderplas@unl.edu}{\nolinkurl{susan.vanderplas@unl.edu}}}
@@
 
-One may argue that reading diagnostic plots is to some extent subjective
-and indecisive compared to those rigorous statistical procedures as it
-relies on graphical perception - human ability to interpret and decode
-the information embedded in graph \citep{cleveland_graphical_1984}.
-Further, the degree of the presence of the visual features typically can
-not be measured quantitatively and objectively, which may lead to over
-or under-interpretations of the data. For instance, people
-over-interpret the separation between gene groups in a two-dimensional
-projection from a linear discriminant analysis when in fact there are no
-differences in the expression levels between the gene groups and
-separation is not an uncommon occurrence
-\citep{roy_chowdhury_using_2015}.
-
-Visual inference was first introduced in a 1999 Joint Statistical
-Meetings (JSM) talk with the title ``Inference for Data Visualization''
-by \citet{buja_inference_1999} as an idea to address the issue of valid
-inference for visual discoveries of data plots. Later,
-\citet{buja_statistical_2009} proposed the lineup protocol as a visual
-test inspired by the ``police lineup'' or ``identity parade'' which is
-the act of asking the eyewitness to identify criminal suspect from a
-group of irrelevant people. The protocol consists of \(m\) randomly
-placed plots, where one plot is the data plot, and the remaining
-\(m - 1\) null plots have the identical graphical procedure except the
-data has been replaced with data consistent with \(H_0\), that the model
-is correctly specified. Then, an observer who have not seen the data
-plot will be asked to point out the most different plot from the lineup.
-Under \(H_0\), it is expected that the data plot would have no
-distinguishable difference from the null plots, and the probability that
-the observer correctly picks the data plot is \(1/m\). If one rejects
-\(H_0\) as the observer correctly picks the data plot, then the Type I
-error of this test is \(1/m\).
+The examination of data plots to infer signal or patterns (or lack
+thereof) is fraught with variation in the human ability to interpret and
+decode the information embedded in graph
+\citep{cleveland_graphical_1984}. Such examination thus feels there is a
+subjective nature to evaluate, say, diagnostic plots to infer
+appropriateness of statistical models, although arguably we would expect
+such evaluation should be done objectively and such human-to-human
+variation in ``reading'' data plots should be minimised.
+
+In practice, over-interpretation of a single plot is common. For
+instance, in \citet{roy_chowdhury_using_2015}, some over-interpreted the
+separation between gene groups in a two-dimensional projection of a
+linear discriminant analysis when in fact there are no differences in
+the expression levels between the gene groups. To mitigate this
+over-interpretation, \citet{buja_statistical_2009} proposed a line-up
+protocol to assess plots in a manner analogous to the null hypothesis
+significance testing (NHST) framework. More specifically, the protocol
+consists of \(m\) randomly placed plots, where one plot is the data
+plot, and the remaining \(m - 1\) plots, referred to as the \emph{null
+plots}, have the identical graphical procedure as the data plot except
+the data is replaced with a data generation mechanism that is consistent
+with the null hypothesis, \(H_0\). Then, an observer who have not seen
+the data plot will be asked to point out the most different plot from
+the lineup. Under \(H_0\), it is expected that the data plot would have
+no distinguishable difference from the null plots, and the probability
+that the observer correctly picks the data plot is \(1/m\). If one
+rejects \(H_0\) as the observer correctly picks the data plot, then the
+Type I error of this test is \(1/m\). This protocol requires apriori
+specification of \(H_0\) (or at least a null data generating mechanism)
+much like the requirement of knowing the distribution of the test
+statistic in NHST.
 
@@
 
-\hypertarget{overview}{%
-\subsection{Overview}\label{overview}}
-
-There are 2880, 2700 and 1674 lineups evaluation made by 160, 150 and
-133 subjects recruited for data collection periods I, II and III
-respectively. In the total of 7974 lineup evaluations, 3744 use lineups
-produced by the non-linearity model, and 4230 use lineups produced by
-the heteroskedasticity model. Besides, there are 886 attention checks
-and 720 evaluations on null lineups needed for the estimate of
-\(\alpha\) not included in the analysis. The collated dataset is
-provided in \texttt{vi\_survey} of the \texttt{visage} \texttt{R}
-package.
-
-In the following analysis, lineups with uniform fitted values will be
-the focus. Visual patterns are more likely to be revealed under a
-uniform distribution. Additionally, we have collected extra evaluations
-on these lineups, which will result in a more reliable analysis.
-Analysis of lineups with other fitted value distributions can be found
-in Section \ref{effect-of-fitted-value-distributions}.
-
-\hypertarget{power-comparison-of-different-tests}{%
-\subsection{Power comparison of different
-tests}\label{power-comparison-of-different-tests}}
+Data collection resulted in a total of 7974 lineup evaluations, 3744 for
+the non-linearity model, and 4230 for the heteroskedasticity model. This
+was done in three collection periods, called I, II, III, with 2880, 2700
+and 1674 lineups evaluated by 160, 150 and 133 participants,
+respectively. There were 886 attention check lineups used to filter
+false participant attempts. To estimate \(\alpha\) for calculating
+statistical significance there were 720 evaluations of null lineups.
+Neither the attention checks nor null lineups were used for the
+subsequent analysis. The de-identified data, \texttt{vi\_survey}, is
+made available in the R package, \texttt{visage}.
+
+For clarity and consistency, the analysis in Sections
+\ref{power-analysis}-\ref{hetero-analysis}, uses only on results from
+lineups generated with uniform fitted values, a total of 4135 lineups
+and XXX evaluation. More evaluations were made using the uniform fitted
+value distribution, and it avoids potentially confounding results of the
+analysis by combining values from different power curves. We also expect
+that visual patterns are more clearly visible when the fitted value
+distribution is uniform. Section
+\ref{effect-of-fitted-value-distributions} contains comparison of the
+effect of fitted value distribution on power.
+
+\hypertarget{power-comparison-of-the-tests}{%
+\subsection{\texorpdfstring{Power comparison of the
+tests\textasciitilde{}\label{power-analysis}}{Power comparison of the tests\textasciitilde{}}}\label{power-comparison-of-the-tests}}
 
@@
 \subsection{\texorpdfstring{Comparison of test decisions based on
-\(p\)-values}{Comparison of test decisions based on p-values}}\label{comparison-of-test-decisions-based-on-p-values}}
+\(p\)-values\textasciitilde{}\label{p-value}}{Comparison of test decisions based on p-values\textasciitilde{}}}\label{comparison-of-test-decisions-based-on-p-values}}
 
@@
 \hypertarget{effect-of-amount-of-non-linearity}{%
-\subsection{Effect of amount of
-non-linearity}\label{effect-of-amount-of-non-linearity}}
+\subsection{\texorpdfstring{Effect of amount of
+non-linearity\textasciitilde{}\label{nonlin-analysis}}{Effect of amount of non-linearity\textasciitilde{}}}\label{effect-of-amount-of-non-linearity}}
 
@@
 \hypertarget{effect-of-shape-of-heteroskedasticity}{%
-\subsection{Effect of shape of
-heteroskedasticity}\label{effect-of-shape-of-heteroskedasticity}}
+\subsection{\texorpdfstring{Effect of shape of
+heteroskedasticity\textasciitilde{}\label{hetero-analysis}}{Effect of shape of heteroskedasticity\textasciitilde{}}}\label{effect-of-shape-of-heteroskedasticity}}
 

### 2023-04-07 paper.Rmd
small typo in top of Section 5

@@
 
-For clarity and consistency, the analysis in Sections \ref{power-analysis}-\ref{hetero-analysis}, uses only on results from lineups generated with uniform fitted values, a total of `r nrow(filter(vi_survey, x_dist == "uniform"))` lineups and XXX evaluation.
+For clarity and consistency, the analysis in Sections \ref{power-analysis}-\ref{hetero-analysis}, uses only on results from lineups generated with uniform fitted values, a total of `r nrow(filter(vi_survey, x_dist == "uniform"))` lineups and XXX evaluations.
 More evaluations were made using the uniform fitted value distribution, and it avoids potentially confounding results of the analysis by combining values from different power curves.
 We also expect that visual patterns are more clearly visible when the fitted value distribution is uniform.
-Section \ref{effect-of-fitted-value-distributions} contains comparison of the effect of fitted value distribution on power.
+The results in Section \ref{effect-of-fitted-value-distributions} examine the effect of fitted value distribution on power.
 

### 2023-04-07 paper.tex
small typo in top of Section 5

@@
 lineups generated with uniform fitted values, a total of 4135 lineups
-and XXX evaluation. More evaluations were made using the uniform fitted
+and XXX evaluations. More evaluations were made using the uniform fitted
 value distribution, and it avoids potentially confounding results of the
@@
 that visual patterns are more clearly visible when the fitted value
-distribution is uniform. Section
-\ref{effect-of-fitted-value-distributions} contains comparison of the
-effect of fitted value distribution on power.
+distribution is uniform. The results in Section
+\ref{effect-of-fitted-value-distributions} examine the effect of fitted
+value distribution on power.
 

### 2023-04-07 paper.Rmd
on section 5.1 might need some rearranging with some details of computations shifted back to methods section

@@
 
-where $P(C \geq \lceil c_i \rceil)$ is calculated using Equation \ref{eq:pvalue-beta-binomial} and $P(C = \lfloor c_i \rfloor)$ is calculated by
+\noindent where $P(C \geq \lceil c_i \rceil)$ is calculated using Equation \ref{eq:pvalue-beta-binomial} and $P(C = \lfloor c_i \rfloor)$ is calculated by
 
@@
 
-Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with uniform fitted values, against the natural logarithm of the effect $log_e(E)$, with a 5% significance level.
-At the bottom of the figure \ref{fig:polypower}, there are a sequence of example residual plots with increasing levels of $log_e(E)$.
-Readers can evaluate them from left to right and determine at which level the departure from a good residual plot becomes detectable.
-
-As discussed in Section \ref{conventionally-testing-for-departures}, many conventionally tests are available for detecting residual departures.
+Figure \ref{fig:polypower} displays the estimated power of a visual test, computed on evaluations of the lineups generated by the non-linearity model with uniform fitted values.
+Power is plotted against the natural logarithm of the effect size $log_e(E)$, computed at a 5% significance level.
+The small plots at the bottom of this figure shows a sequence of example residual plots of increasing effect size.
+These are for the readers to compare what a residual plot might look like at particular effect sizes.
+
+<!-- Where would you draw the line of too much non-linearity in the residuals? For the RESET test this is around log effect size 2, but for the visual test it is around 3.5.-->
+
+As discussed in Section \ref{conventionally-testing-for-departures}, many conventional tests are available for detecting residual departures.
 Implementation-wise, the built-in R package `stats` provides some commonly used residual-based tests, such as Shapiro-Wilk test.
@@
 ```{r polypower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for non-linear patterns (uniform fitted values only).
-Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups.
-Surrounding lines of the visual test show the estimated power of 500 bootstrap samples.
-Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by vertical lines in main plot.
-Where would you draw the line of too much non-linearity in the residuals?
-For the RESET test this is around log effect size 2, but for the visual test it is around 3.5.", dev = 'png', dpi = 450}
+The power curves are estimated using logistic regression, and the horizontal lines of dots represent human evaluations of lineups, with 0, 1 indicating non-detect and detect, respectively.
+lines The visual test has multiple power curves estimated from bootstrap samples.
+The row of scatterplots plots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.", dev = 'png', dpi = 450}
 

### 2023-04-07 paper.tex
on section 5.1 might need some rearranging with some details of computations shifted back to methods section

@@
 
-where \(P(C \geq \lceil c_i \rceil)\) is calculated using Equation
-\ref{eq:pvalue-beta-binomial} and \(P(C = \lfloor c_i \rfloor)\) is
-calculated by
+\noindent where \(P(C \geq \lceil c_i \rceil)\) is calculated using
+Equation \ref{eq:pvalue-beta-binomial} and
+\(P(C = \lfloor c_i \rfloor)\) is calculated by
 
@@
 
-Figure \ref{fig:polypower} shows the estimated power of visual test on
-lineups produced by the non-linearity model with uniform fitted values,
-against the natural logarithm of the effect \(log_e(E)\), with a 5\%
-significance level. At the bottom of the figure \ref{fig:polypower},
-there are a sequence of example residual plots with increasing levels of
-\(log_e(E)\). Readers can evaluate them from left to right and determine
-at which level the departure from a good residual plot becomes
-detectable.
+Figure \ref{fig:polypower} displays the estimated power of a visual
+test, computed on evaluations of the lineups generated by the
+non-linearity model with uniform fitted values. Power is plotted against
+the natural logarithm of the effect size \(log_e(E)\), computed at a 5\%
+significance level. The small plots at the bottom of this figure shows a
+sequence of example residual plots of increasing effect size. These are
+for the readers to compare what a residual plot might look like at
+particular effect sizes.
 
 As discussed in Section \ref{conventionally-testing-for-departures},
-many conventionally tests are available for detecting residual
-departures. Implementation-wise, the built-in R package \texttt{stats}
-provides some commonly used residual-based tests, such as Shapiro-Wilk
-test. A more comprehensive collection of regression diagnostics tests
-can be found in the R package \texttt{lmtest} \citep{lmtest}. In terms
-of heteroskedasticity diagnostics, the R package \texttt{skedastic}
+many conventional tests are available for detecting residual departures.
+Implementation-wise, the built-in R package \texttt{stats} provides some
+commonly used residual-based tests, such as Shapiro-Wilk test. A more
+comprehensive collection of regression diagnostics tests can be found in
+the R package \texttt{lmtest} \citep{lmtest}. In terms of
+heteroskedasticity diagnostics, the R package \texttt{skedastic}
 \citep{skedastic} collects and implements 25 existing conventional tests
@@
 \caption{Comparison of power between different tests for non-linear patterns (uniform fitted values only).
-Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups.
-Surrounding lines of the visual test show the estimated power of 500 bootstrap samples.
-Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by vertical lines in main plot.
-Where would you draw the line of too much non-linearity in the residuals?
-For the RESET test this is around log effect size 2, but for the visual test it is around 3.5.}\label{fig:polypower}
+The power curves are estimated using logistic regression, and the horizontal lines of dots represent human evaluations of lineups, with 0, 1 indicating non-detect and detect, respectively.
+lines The visual test has multiple power curves estimated from bootstrap samples.
+The row of scatterplots plots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}\label{fig:polypower}
 \end{figure}

### 2023-04-07 paper.Rmd
small changes in 3.1

@@
 
-In diagnosing a model fit from residuals, we are generally interested in $H_0:$ *The regression model is correctly specified* against the broad alternative $H_a:$ *The regression model is misspecified*.
+In diagnosing a model fit from residuals, we are generally interested in *the regression model is correctly specified* ($H_0$) against the broad alternative  *the regression model is misspecified* ($H_a$).
 
@@
 
-One of the potential benefits of the visual test, based on the lineup protocol, is that it should be able to detect a range of departures from good residuals.
+One of the potential benefits of the visual test, based on the lineup protocol, is that it works as an omnibus test,  able to detect a range of departures from good residuals.
 

### 2023-04-07 paper.tex
small changes in 3.1

@@
 In diagnosing a model fit from residuals, we are generally interested in
-\(H_0:\) \emph{The regression model is correctly specified} against the
-broad alternative \(H_a:\) \emph{The regression model is misspecified}.
+\emph{the regression model is correctly specified} (\(H_0\)) against the
+broad alternative \emph{the regression model is misspecified} (\(H_a\)).
 
@@
 One of the potential benefits of the visual test, based on the lineup
-protocol, is that it should be able to detect a range of departures from
-good residuals.
+protocol, is that it works as an omnibus test, able to detect a range of
+departures from good residuals.
 

### 2023-04-08 paper.Rmd
fill for result counts

@@
 
-For clarity and consistency, the analysis in Sections \ref{power-analysis}-\ref{hetero-analysis}, uses only on results from lineups generated with uniform fitted values, a total of `r nrow(filter(vi_survey, x_dist == "uniform"))` lineups and XXX evaluations.
+For clarity and consistency, the analysis in Sections \ref{power-analysis}-\ref{hetero-analysis}, uses only on results from lineups generated with uniform fitted values, a total of `r unique(pull(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"), unique_lineup_id))` lineups and `nrow(filter(vi_survey, !attention_check, !null_lineup))` evaluations.
 More evaluations were made using the uniform fitted value distribution, and it avoids potentially confounding results of the analysis by combining values from different power curves.

### 2023-04-08 paper.Rmd
fix polypower caption

@@
 ```{r polypower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for non-linear patterns (uniform fitted values only).
-The power curves are estimated using logistic regression, and the horizontal lines of dots represent human evaluations of lineups, with 0, 1 indicating non-detect and detect, respectively.
-lines The visual test has multiple power curves estimated from bootstrap samples.
+The power curves are estimated using logistic regression, and the horizontal lines of dots represent decisions made by visual tests, with 0, 1 indicating non-reject and reject, respectively.
+The visual test has multiple power curves estimated from bootstrap samples.
 The row of scatterplots plots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.", dev = 'png', dpi = 450}

### 2023-04-08 paper.Rmd
fix unique lineups count

@@
 
-For clarity and consistency, the analysis in Sections \ref{power-analysis}-\ref{hetero-analysis}, uses only on results from lineups generated with uniform fitted values, a total of `r unique(pull(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"), unique_lineup_id))` lineups and `nrow(filter(vi_survey, !attention_check, !null_lineup))` evaluations.
+For clarity and consistency, the analysis in Sections \ref{power-analysis}-\ref{hetero-analysis}, uses only on results from lineups generated with uniform fitted values, a total of `r length(unique(pull(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"), unique_lineup_id)))` lineups and `nrow(filter(vi_survey, !attention_check, !null_lineup))` evaluations.
 More evaluations were made using the uniform fitted value distribution, and it avoids potentially confounding results of the analysis by combining values from different power curves.

### 2023-04-08 paper.Rmd
Use discrete instead of discrete uniform across the paper

@@
            n = c("50", "100", "300", "", ""), 
-           x_dist = c("Uniform", "Normal", "Skewed", "Discrete uniform", "")) %>%
+           x_dist = c("Uniform", "Normal", "Skewed", "Discrete", "")) %>%
   knitr::kable(col.names = NULL,
@@
 This latter problem is usually corrected before modelling using a variable transformation.
-Our simulation assess this by using four different distributions to represent fitted values: (1) uniform, (2) normal, (3) skewed and (4) discrete uniform.
+Our simulation assess this by using four different distributions to represent fitted values: (1) uniform, (2) normal, (3) skewed and (4) discrete.
 This is constructed by defining the raw predictor $X_{raw}$ in four corresponding distributions:
@@
 
-Three different sample sizes are used, $n=50, 100, 300$ across the experiments.
+Three different sample sizes are used, $n = 50, 100, 300$ across the experiments.
 We would expect considerable variation in the signal strength in the simulated data plots with smaller $n$.
@@
   gen(300, computed = select(dat_dist_1, e)) %>%
-  mutate(x_dist = "Discrete uniform")
+  mutate(x_dist = "Discrete")
 
@@
                                     "Skewed", 
-                                    "Discrete uniform"))) %>%
+                                    "Discrete"))) %>%
   VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
@@
   mutate(x_dist = ifelse(x_dist == "even_discrete", 
-                         "Discrete uniform", 
+                         "Discrete", 
                          stringr::str_to_title(x_dist))) %>%
@@
                                     "Lognormal", 
-                                    "Discrete uniform"))) %>%
+                                    "Discrete"))) %>%
 

### 2023-04-08 paper.Rmd
section 5 intro finished

--- 2023-04-08 paper.Rmd
@@
 
-## Statistical significance
+## Statistical significance~\label{sig}
 
@@
 
-Data collection resulted in a total of `r nrow(filter(vi_survey, !attention_check))` lineup evaluations, `r nrow(filter(vi_survey, type == "polynomial", !attention_check))` for the non-linearity model, and `r nrow(filter(vi_survey, type == "heteroskedasticity", !attention_check))` for the heteroskedasticity model.
-This was done in three collection periods, called I, II, III, with `r 160 * 18`, `r 150 * 18` and `r 133 * 18 - nrow(filter(vi_survey, null_lineup, !attention_check))` lineups evaluated by 160, 150 and 133 participants, respectively.
-There were `r nrow(filter(vi_survey, attention_check))` attention check lineups used to filter false participant attempts.
-To estimate $\alpha$ for calculating statistical significance there were `r nrow(filter(vi_survey, null_lineup, !attention_check))` evaluations of null lineups.
+Data collection used a total of `r vi_survey %>% count(unique_lineup_id) %>% nrow()` lineups, and resulted in a total of `r nrow(filter(vi_survey, !attention_check))` evaluations from `r vi_survey %>% count(set, exp) %>% nrow()` participants.
+Roughly half corresponded to the two models, non-linearity and heteroskedasticiy, and the three collection periods had similar numbers in each.
+<!-- `r nrow(filter(vi_survey, type == "polynomial", !attention_check))` for the non-linearity model, and `r nrow(filter(vi_survey, type == "heteroskedasticity", !attention_check))` for the heteroskedasticity model.
+This was done in three collection periods (I, II, III), with `r 160 * 18`, `r 150 * 18` and `r 133 * 18 - nrow(filter(vi_survey, null_lineup, !attention_check))` lineups evaluated by 160, 150 and 133 participants, respectively.--> Each participant received two of the `r vi_survey %>% filter(attention_check) %>% count(unique_lineup_id) %>% nrow()` attention check lineups which were used to filter results of participants who were clearly not making an honest effort (only `r vi_survey %>% filter(attention_check, !detect) %>% count(exp, set) %>% nrow()` of `r vi_survey %>% count(set, exp) %>% nrow()`).
+To estimate $\alpha$ for calculating statistical significance (see Section \ref{sig}) there were `r nrow(filter(vi_survey, null_lineup, !attention_check))` evaluations of `r vi_survey %>% filter(null_lineup, !attention_check) %>% count(unique_lineup_id) %>% nrow()` null lineups.
 Neither the attention checks nor null lineups were used for the subsequent analysis.
@@
 
-For clarity and consistency, the analysis in Sections \ref{power-analysis}-\ref{hetero-analysis}, uses only on results from lineups generated with uniform fitted values, a total of `r unique(pull(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"), unique_lineup_id))` lineups and `nrow(filter(vi_survey, !attention_check, !null_lineup))` evaluations.
-More evaluations were made using the uniform fitted value distribution, and it avoids potentially confounding results of the analysis by combining values from different power curves.
-We also expect that visual patterns are more clearly visible when the fitted value distribution is uniform.
-The results in Section \ref{effect-of-fitted-value-distributions} examine the effect of fitted value distribution on power.
+The data was collected on lineups constructed from four different fitted value distributions: uniform, normal, skewed and discrete.
+More data was collected on the uniform (each evaluated by 11 participants) than the others (each evaluated by 5 participants).
+The analysis in Sections \ref{power-analysis}-\ref{hetero-analysis}, uses only results from lineups generated with uniform fitted values, a total <!--of `r length(unique(pull(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"), unique_lineup_id)))` lineups and--> `r nrow(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"))` lineup evaluations.
+This was decided in order to compare the conventional and visual test performance for an optimal scenario.
+Section \ref{effect-of-fitted-value-distributions} examines how the results may be affected if the fitted value distribution was different.
 

### 2023-04-08 paper.tex
section 5 intro finished

@@
 \hypertarget{statistical-significance}{%
-\subsection{Statistical significance}\label{statistical-significance}}
+\subsection{\texorpdfstring{Statistical
+significance\textasciitilde{}\label{sig}}{Statistical significance\textasciitilde{}}}\label{statistical-significance}}
 
@@
 
-Data collection resulted in a total of 7974 lineup evaluations, 3744 for
-the non-linearity model, and 4230 for the heteroskedasticity model. This
-was done in three collection periods, called I, II, III, with 2880, 2700
-and 1674 lineups evaluated by 160, 150 and 133 participants,
-respectively. There were 886 attention check lineups used to filter
-false participant attempts. To estimate \(\alpha\) for calculating
-statistical significance there were 720 evaluations of null lineups.
-Neither the attention checks nor null lineups were used for the
-subsequent analysis. The de-identified data, \texttt{vi\_survey}, is
-made available in the R package, \texttt{visage}.
-
-For clarity and consistency, the analysis in Sections
-\ref{power-analysis}-\ref{hetero-analysis}, uses only on results from
-lineups generated with uniform fitted values, a total of 4135 lineups
-and XXX evaluations. More evaluations were made using the uniform fitted
-value distribution, and it avoids potentially confounding results of the
-analysis by combining values from different power curves. We also expect
-that visual patterns are more clearly visible when the fitted value
-distribution is uniform. The results in Section
-\ref{effect-of-fitted-value-distributions} examine the effect of fitted
-value distribution on power.
+Data collection used a total of 1176 lineups, and resulted in a total of
+7974 evaluations from 443 participants. Roughly half corresponded to the
+two models, non-linearity and heteroskedasticiy, and the three
+collection periods had similar numbers in each. Each participant
+received two of the 24 attention check lineups which were used to filter
+results of participants who were clearly not making an honest effort
+(only 4 of 443). To estimate \(\alpha\) for calculating statistical
+significance (see Section \ref{sig}) there were 720 evaluations of 36
+null lineups. Neither the attention checks nor null lineups were used
+for the subsequent analysis. The de-identified data,
+\texttt{vi\_survey}, is made available in the R package,
+\texttt{visage}.
+
+The data was collected on lineups constructed from four different fitted
+value distributions: uniform, normal, skewed and discrete. More data was
+collected on the uniform (each evaluated by 11 participants) than the
+others (each evaluated by 5 participants). The analysis in Sections
+\ref{power-analysis}-\ref{hetero-analysis}, uses only results from
+lineups generated with uniform fitted values, a total 3069 lineup
+evaluations. This was decided in order to compare the conventional and
+visual test performance for an optimal scenario. Section
+\ref{effect-of-fitted-value-distributions} examines how the results may
+be affected if the fitted value distribution was different.
 
@@
 \caption{Comparison of power between different tests for non-linear patterns (uniform fitted values only).
-The power curves are estimated using logistic regression, and the horizontal lines of dots represent human evaluations of lineups, with 0, 1 indicating non-detect and detect, respectively.
-lines The visual test has multiple power curves estimated from bootstrap samples.
+The power curves are estimated using logistic regression, and the horizontal lines of dots represent decisions made by visual tests, with 0, 1 indicating non-reject and reject, respectively.
+The visual test has multiple power curves estimated from bootstrap samples.
 The row of scatterplots plots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}\label{fig:polypower}

### 2023-04-08 paper.Rmd

--- 2023-04-08 paper.Rmd
@@
 
-For clarity and consistency, the analysis in Sections \ref{power-analysis}-\ref{hetero-analysis}, uses only on results from lineups generated with uniform fitted values, a total of `r length(unique(pull(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"), unique_lineup_id)))` lineups and `nrow(filter(vi_survey, !attention_check, !null_lineup))` evaluations.
-More evaluations were made using the uniform fitted value distribution, and it avoids potentially confounding results of the analysis by combining values from different power curves.
-We also expect that visual patterns are more clearly visible when the fitted value distribution is uniform.
-The results in Section \ref{effect-of-fitted-value-distributions} examine the effect of fitted value distribution on power.
+The data was collected on lineups constructed from four different fitted value distributions: uniform, normal, skewed and discrete.
+More data was collected on the uniform (each evaluated by 11 participants) than the others (each evaluated by 5 participants).
+The analysis in Sections \ref{power-analysis}-\ref{hetero-analysis}, uses only results from lineups generated with uniform fitted values, a total <!--of `r length(unique(pull(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"), unique_lineup_id)))` lineups and--> `r nrow(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"))` lineup evaluations.
+This was decided in order to compare the conventional and visual test performance for an optimal scenario.
+Section \ref{effect-of-fitted-value-distributions} examines how the results may be affected if the fitted value distribution was different.
 

### 2023-04-08 paper.tex
section 5 intro finished

@@
 6 & 2.00 & 1 & 4.00 & 300 & Skewed\\
-18 & 4.00 &  & 16.00 &  & Discrete uniform\\
+18 & 4.00 &  & 16.00 &  & Discrete\\
  &  &  & 64.00 &  & \\
@@
 represent fitted values: (1) uniform, (2) normal, (3) skewed and (4)
-discrete uniform. This is constructed by defining the raw predictor
-\(X_{raw}\) in four corresponding distributions: (1) \(U(-1, 1)\), (2)
+discrete. This is constructed by defining the raw predictor \(X_{raw}\)
+in four corresponding distributions: (1) \(U(-1, 1)\), (2)
 \(N(0, 0.3^2)\), (3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). We
@@
 
-Three different sample sizes are used, \(n=50, 100, 300\) across the
+Three different sample sizes are used, \(n = 50, 100, 300\) across the
 experiments. We would expect considerable variation in the signal

### 2023-04-08 paper.Rmd
re-working section 2.2 and 5.1, 2.2 still needs more

@@
 
-Other than checking diagnostic plots, analysts may perform formal hypothesis testing to detect model defects.
-A variety of tests can be applied; each testing for a specific violation of the null hypothesis.
+Many different hypothesis tests are available to detect specific model defects.
 For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution.
@@
 
-We apply the RESET test, Breusch-Pagan test and Shapiro-Wilk test to the residual plots shown in Figure \ref{fig:residual-plot-common-departures}; the results of this is shown in Table \ref{tab:example-residual-plot-table}.
-The Breusch-Pagan test and the Shapiro-Wilk test both reject $H_0$ for residual plots that display non-linearity and heteroscedasticity, even though these are not the original intention of the test.
-As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are also sensitive to other types of departures.
-It is likely $H_0$ is correctly rejected but for the wrong reason, a phenomenon known as the "Type III error".
-Additionally, outliers will often incorrectly trigger the rejection of $H_0$ despite when majority of the residuals are well-behaved [@cook_applied_1999].
-Furthermore, with a sufficiently large sample size, residual-based tests may reject $H_0$ due to a slight departure that is of little practical significance.
-These can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers and slight departures.
+Table \ref{tab:example-residual-plot-table} shows the RESET, Breusch-Pagan and Shapiro-Wilk test results ($p$-values) for the residual plots shown in Figure \ref{fig:residual-plot-common-departures}.
+The RESET test  and Breusch-Pagan test were computed using the `resettest` and `bptest` functions from R package `lmtest`, respectively.
+The Shapiro-Wilk test  was computed using the `shapiro.test` from the core R package `stats`.
+(The the R package `skedastic` [@skedastic] contains a large collection of tests for heteroskedasticity.) The RESET test is exact, but the power parameter must be chosen.
+@ramsey_tests_1969 suggests using a power of four, which is what we used.
+The Breusch-Pagan and Shapiro-Wilk tests are approximate.
+
+We would expect the RESET test for non-linearity to reject residual plot B, the Breusch-Pagan test for heteroskedasticity to reject the residual plot C, and Shapiro-Wilk test for non-normality to reject residual plot D, which they all do and correctly fail to reject residual plot A.
+Interestingly, the Breusch-Pagan and Shapiro-Wilk tests also reject the residual plots exhibiting structure that they weren't designed for.
+@cook1982residuals discusses that most residual-based tests for particular types of departure from model assumptions are also sensitive to other types of departures, that is, correctly rejected but for the wrong reason, a phenomenon known as the "Type III error".
+Also, some types of departure can also have elements of other types of departure, for example, non-linearity could be viewed as heteroskedasticity.
+Additionally, other problesms such as outliers can trigger the rejection [@cook_applied_1999].
+With a sufficiently large sample size, residual-based tests may reject when there is only a slight departure.
+This is a situation that is addressed by the results in this paper because from a practical perspective we would only want to flag the model as problematic when ther is a XXX diagnostic  practical significance.
+<!--These can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers and slight departures.
+-->
+ XXX check heteroscedastic spelling
+
+<!--
+As discussed in Section \ref{conventionally-testing-for-departures}, many conventional tests are available for detecting residual departures.
+Implementation-wise, the built-in R package `stats` provides some commonly used residual-based tests, such as Shapiro-Wilk test.
+A more comprehensive collection of regression diagnostics tests can be found in the R package `lmtest` [@lmtest].
+In terms of heteroskedasticity diagnostics, the R package `skedastic` [@skedastic] collects and implements 25 existing conventional tests published since 1961.
+
+We pick RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`.
+Among them, RESET test is the only exact and appropriate test in this scenario.
+Both the Breusch-Pagan test and the Shapiro-Wilk test are approximate and inappropriate tests.
+Their estimated power is shown in Figure \ref{fig:polypower}.
+To set up the RESET test, we include different powers of fitted values as proxies.
+According to @ramsey_tests_1969, there are no general rules for the power of the fitted values needed by the RESET test, but it finds power up to four is usually sufficient.
+Thus, we follow this guideline to conduct the RESET test.
+For the Breusch-Pagan test, the choice of predictors in the auxiliary regression is left to the user [@breusch_simple_1979].
+But as @waldman1983note suggested, it is a good choice for the set of auxiliary predictors in the Breusch-Pagan test be the same as the White test.
+Thus, we include both $x$ and $x^2$ in the auxiliary regression.
+-->
 
@@
 
-As discussed in Section \ref{conventionally-testing-for-departures}, many conventional tests are available for detecting residual departures.
-Implementation-wise, the built-in R package `stats` provides some commonly used residual-based tests, such as Shapiro-Wilk test.
-A more comprehensive collection of regression diagnostics tests can be found in the R package `lmtest` [@lmtest].
-In terms of heteroskedasticity diagnostics, the R package `skedastic` [@skedastic] collects and implements 25 existing conventional tests published since 1961.
-
-We pick RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`.
-Among them, RESET test is the only exact and appropriate test in this scenario.
-Both the Breusch-Pagan test and the Shapiro-Wilk test are approximate and inappropriate tests.
-Their estimated power is shown in Figure \ref{fig:polypower}.
-To set up the RESET test, we include different powers of fitted values as proxies.
-According to @ramsey_tests_1969, there are no general rules for the power of the fitted values needed by the RESET test, but it finds power up to four is usually sufficient.
-Thus, we follow this guideline to conduct the RESET test.
-For the Breusch-Pagan test, the choice of predictors in the auxiliary regression is left to the user [@breusch_simple_1979].
-But as @waldman1983note suggested, it is a good choice for the set of auxiliary predictors in the Breusch-Pagan test be the same as the White test.
-Thus, we include both $x$ and $x^2$ in the auxiliary regression.
-
 Figure \ref{fig:heterpower} is similar to Figure \ref{fig:polypower}, but shows corresponding information on lineups produced by the heteroskedasticity model.

### 2023-04-08 paper.tex
re-working section 2.2 and 5.1, 2.2 still needs more

@@
 
-Other than checking diagnostic plots, analysts may perform formal
-hypothesis testing to detect model defects. A variety of tests can be
-applied; each testing for a specific violation of the null hypothesis.
-For example, the presence of heteroskedasticity can usually be tested by
-applying the White test \citep{white_heteroskedasticity-consistent_1980}
-or the Breusch-Pagan test \citep{breusch_simple_1979}, which are both
-derived from the Lagrange multiplier test \citep{silvey1959lagrangian}
-principle that relies on the asymptotic properties of the null
-distribution. To test specific forms of non-linearity, one may apply the
-F-test as a model structural test to examine the significance of
-specific polynomial and non-linear forms of the predictors, or the
-significance of proxy variables as in the Ramsey Regression Equation
-Specification Error Test (RESET) \citep{ramsey_tests_1969}. The
-Shapiro-Wilk test \citep{shapiro1965analysis} is the most widely used
-test of non-normality included by many of the statistical software
-programs. The Jarque--Bera test \citep{jarque1980efficient} is also used
-to directly check whether the sample skewness and kurtosis match a
-normal distribution.
-
-We apply the RESET test, Breusch-Pagan test and Shapiro-Wilk test to the
-residual plots shown in Figure
-\ref{fig:residual-plot-common-departures}; the results of this is shown
-in Table \ref{tab:example-residual-plot-table}. The Breusch-Pagan test
-and the Shapiro-Wilk test both reject \(H_0\) for residual plots that
-display non-linearity and heteroscedasticity, even though these are not
-the original intention of the test. As discussed in
+Many different hypothesis tests are available to detect specific model
+defects. For example, the presence of heteroskedasticity can usually be
+tested by applying the White test
+\citep{white_heteroskedasticity-consistent_1980} or the Breusch-Pagan
+test \citep{breusch_simple_1979}, which are both derived from the
+Lagrange multiplier test \citep{silvey1959lagrangian} principle that
+relies on the asymptotic properties of the null distribution. To test
+specific forms of non-linearity, one may apply the F-test as a model
+structural test to examine the significance of specific polynomial and
+non-linear forms of the predictors, or the significance of proxy
+variables as in the Ramsey Regression Equation Specification Error Test
+(RESET) \citep{ramsey_tests_1969}. The Shapiro-Wilk test
+\citep{shapiro1965analysis} is the most widely used test of
+non-normality included by many of the statistical software programs. The
+Jarque--Bera test \citep{jarque1980efficient} is also used to directly
+check whether the sample skewness and kurtosis match a normal
+distribution.
+
+Table \ref{tab:example-residual-plot-table} shows the RESET,
+Breusch-Pagan and Shapiro-Wilk test results for the residual plots shown
+in Figure \ref{fig:residual-plot-common-departures}. The RESET test and
+Breusch-Pagan test were computed using the \texttt{resettest} and
+\texttt{bptest} functions from R package \texttt{lmtest}, respectively.
+The Shapiro-Wilk test was computed using the \texttt{shapiro.test} from
+the core R package \texttt{stats}. (The the R package \texttt{skedastic}
+\citep{skedastic} contains a large collection of tests for
+heteroskedasticity.) The RESET test is exact, but the power parameter
+must be chosen. \citet{ramsey_tests_1969} suggests using a power of
+four, which is what we used. The Breusch-Pagan and Shapiro-Wilk tests
+are approximate.
+
+The Breusch-Pagan test and the Shapiro-Wilk test both reject \(H_0\) for
+residual plots that display non-linearity and heteroscedasticity, even
+though these are not the original intention of the test. As discussed in
 \citet{cook1982residuals}, most residual-based tests for a particular
@@
 
-As discussed in Section \ref{conventionally-testing-for-departures},
-many conventional tests are available for detecting residual departures.
-Implementation-wise, the built-in R package \texttt{stats} provides some
-commonly used residual-based tests, such as Shapiro-Wilk test. A more
-comprehensive collection of regression diagnostics tests can be found in
-the R package \texttt{lmtest} \citep{lmtest}. In terms of
-heteroskedasticity diagnostics, the R package \texttt{skedastic}
-\citep{skedastic} collects and implements 25 existing conventional tests
-published since 1961.
-
-We pick RESET test (\texttt{resettest}) and Breusch-Pagan test
-(\texttt{bptest}) from the R package \texttt{lmtest}, and Shapiro-Wilk
-test (\texttt{shapiro.test}) from the built-in R package \texttt{stats}.
-Among them, RESET test is the only exact and appropriate test in this
-scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are
-approximate and inappropriate tests. Their estimated power is shown in
-Figure \ref{fig:polypower}. To set up the RESET test, we include
-different powers of fitted values as proxies. According to
-\citet{ramsey_tests_1969}, there are no general rules for the power of
-the fitted values needed by the RESET test, but it finds power up to
-four is usually sufficient. Thus, we follow this guideline to conduct
-the RESET test. For the Breusch-Pagan test, the choice of predictors in
-the auxiliary regression is left to the user
-\citep{breusch_simple_1979}. But as \citet{waldman1983note} suggested,
-it is a good choice for the set of auxiliary predictors in the
-Breusch-Pagan test be the same as the White test. Thus, we include both
-\(x\) and \(x^2\) in the auxiliary regression.
-
 Figure \ref{fig:heterpower} is similar to Figure \ref{fig:polypower},

### 2023-04-08 paper.tex
re-working section 2.2 and 5.1, 2.2 still needs more

@@
 Table \ref{tab:example-residual-plot-table} shows the RESET,
-Breusch-Pagan and Shapiro-Wilk test results for the residual plots shown
-in Figure \ref{fig:residual-plot-common-departures}. The RESET test and
+Breusch-Pagan and Shapiro-Wilk test results (\(p\)-values) for the
+residual plots shown in Figure
+\ref{fig:residual-plot-common-departures}. The RESET test and
 Breusch-Pagan test were computed using the \texttt{resettest} and
@@
 
-The Breusch-Pagan test and the Shapiro-Wilk test both reject \(H_0\) for
-residual plots that display non-linearity and heteroscedasticity, even
-though these are not the original intention of the test. As discussed in
-\citet{cook1982residuals}, most residual-based tests for a particular
-type of departure from model assumptions are also sensitive to other
-types of departures. It is likely \(H_0\) is correctly rejected but for
-the wrong reason, a phenomenon known as the ``Type III error''.
-Additionally, outliers will often incorrectly trigger the rejection of
-\(H_0\) despite when majority of the residuals are well-behaved
-\citep{cook_applied_1999}. Furthermore, with a sufficiently large sample
-size, residual-based tests may reject \(H_0\) due to a slight departure
-that is of little practical significance. These can be largely avoided
-in diagnostic plots as experienced analysts can evaluate the
-acceptability of assumptions flexibly, even in the presence of outliers
-and slight departures.
+We would expect the RESET test for non-linearity to reject residual plot
+B, the Breusch-Pagan test for heteroskedasticity to reject the residual
+plot C, and Shapiro-Wilk test for non-normality to reject residual plot
+D, which they all do and correctly fail to reject residual plot A.
+Interestingly, the Breusch-Pagan and Shapiro-Wilk tests also reject the
+residual plots exhibiting structure that they weren't designed for.
+\citet{cook1982residuals} discusses that most residual-based tests for
+particular types of departure from model assumptions are also sensitive
+to other types of departures, that is, correctly rejected but for the
+wrong reason, a phenomenon known as the ``Type III error''. Also, some
+types of departure can also have elements of other types of departure,
+for example, non-linearity could be viewed as heteroskedasticity.
+Additionally, other problesms such as outliers can trigger the rejection
+\citep{cook_applied_1999}. With a sufficiently large sample size,
+residual-based tests may reject when there is only a slight departure.
+This is a situation that is addressed by the results in this paper
+because from a practical perspective we would only want to flag the
+model as problematic when ther is a XXX diagnostic practical
+significance. XXX check heteroscedastic spelling
 

### 2023-04-08 paper.Rmd
section 2.2 revised

@@
 
-Table \ref{tab:example-residual-plot-table} shows the RESET, Breusch-Pagan and Shapiro-Wilk test results ($p$-values) for the residual plots shown in Figure \ref{fig:residual-plot-common-departures}.
-The RESET test  and Breusch-Pagan test were computed using the `resettest` and `bptest` functions from R package `lmtest`, respectively.
+Table \ref{tab:example-residual-plot-table} displays the $p$-values from the RESET, Breusch-Pagan and Shapiro-Wilk tests applied to the residual plots in Figure \ref{fig:residual-plot-common-departures}.
+The RESET test and Breusch-Pagan test were computed using the `resettest` and `bptest` functions from the R package `lmtest`, respectively.
 The Shapiro-Wilk test was computed using the `shapiro.test` from the core R package `stats`.
-(The the R package `skedastic` [@skedastic] contains a large collection of tests for heteroskedasticity.) The RESET test is exact, but the power parameter must be chosen.
-@ramsey_tests_1969 suggests using a power of four, which is what we used.
+(The the R package `skedastic` [@skedastic] contains a large collection of tests for heteroskedasticity.) Although, the RESET test is exact, it requires the selection of a power parameter.
+According to @ramsey_tests_1969, a power of four is recommended, which we adopted in our analysis.
 The Breusch-Pagan and Shapiro-Wilk tests are approximate.
@@
 Additionally, other problesms such as outliers can trigger the rejection [@cook_applied_1999].
-With a sufficiently large sample size, residual-based tests may reject when there is only a slight departure.
-This is a situation that is addressed by the results in this paper because from a practical perspective we would only want to flag the model as problematic when ther is a XXX diagnostic  practical significance.
-<!--These can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers and slight departures.
+
+With large sample sizes, hypothesis tests are may reject a residual plot when there is only a slight departure (XXX is there a reference for this statement?).
+While such rejections may be statistically correct, their sensitivity may render the results impractical.
+Therefore, a key goal of residual plot diagnostics is to identify potential issues that could lead to incorrect conclusions or errors in subsequent analyses.
+However, minor defects in the model are unlikely to have a significant impact and may be best disregarded for practical purposes.
+In fact, the experiment discussed in this paper specifically addresses this.
+
+<!--that is addressed by the results in this paper because from a practical perspective we would only want to flag the model as problematic when ther is a XXX diagnostic  practical significance.
+These can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers and slight departures.
 -->
- XXX check heteroscedastic spelling
 

### 2023-04-08 paper.tex
section 2.2 revised

@@
 
-Table \ref{tab:example-residual-plot-table} shows the RESET,
-Breusch-Pagan and Shapiro-Wilk test results (\(p\)-values) for the
-residual plots shown in Figure
-\ref{fig:residual-plot-common-departures}. The RESET test and
-Breusch-Pagan test were computed using the \texttt{resettest} and
-\texttt{bptest} functions from R package \texttt{lmtest}, respectively.
-The Shapiro-Wilk test was computed using the \texttt{shapiro.test} from
-the core R package \texttt{stats}. (The the R package \texttt{skedastic}
-\citep{skedastic} contains a large collection of tests for
-heteroskedasticity.) The RESET test is exact, but the power parameter
-must be chosen. \citet{ramsey_tests_1969} suggests using a power of
-four, which is what we used. The Breusch-Pagan and Shapiro-Wilk tests
-are approximate.
+Table \ref{tab:example-residual-plot-table} displays the \(p\)-values
+from the RESET, Breusch-Pagan and Shapiro-Wilk tests applied to the
+residual plots in Figure \ref{fig:residual-plot-common-departures}. The
+RESET test and Breusch-Pagan test were computed using the
+\texttt{resettest} and \texttt{bptest} functions from the R package
+\texttt{lmtest}, respectively. The Shapiro-Wilk test was computed using
+the \texttt{shapiro.test} from the core R package \texttt{stats}. (The
+the R package \texttt{skedastic} \citep{skedastic} contains a large
+collection of tests for heteroskedasticity.) Although, the RESET test is
+exact, it requires the selection of a power parameter. According to
+\citet{ramsey_tests_1969}, a power of four is recommended, which we
+adopted in our analysis. The Breusch-Pagan and Shapiro-Wilk tests are
+approximate.
 
@@
 Additionally, other problesms such as outliers can trigger the rejection
-\citep{cook_applied_1999}. With a sufficiently large sample size,
-residual-based tests may reject when there is only a slight departure.
-This is a situation that is addressed by the results in this paper
-because from a practical perspective we would only want to flag the
-model as problematic when ther is a XXX diagnostic practical
-significance. XXX check heteroscedastic spelling
+\citep{cook_applied_1999}.
+
+With large sample sizes, hypothesis tests are may reject a residual plot
+when there is only a slight departure (XXX is there a reference for this
+statement?). While such rejections may be statistically correct, their
+sensitivity may render the results impractical. Therefore, a key goal of
+residual plot diagnostics is to identify potential issues that could
+lead to incorrect conclusions or errors in subsequent analyses. However,
+minor defects in the model are unlikely to have a significant impact and
+may be best disregarded for practical purposes. In fact, the experiment
+discussed in this paper specifically addresses this.
 

### 2023-04-08 paper.Rmd
working on section 5.1

@@
 
-Figure \ref{fig:polypower} displays the estimated power of a visual test, computed on evaluations of the lineups generated by the non-linearity model with uniform fitted values.
+Figures \ref{fig:polypower} and \ref{fig:heterpower} present the power curves of various tests plotted against the effect size in the residuals.
+The effect size was computed at a 5% significance level and plotted on a natural logarithmic scale.
+To facilitate visual calibration of effect size values with the corresponding diagnostic plots, a sequence of example residual plots with increasing effect sizes is provided at the bottom of these figures.
+These plots serve as a visual aid to help readers understand how different effect size values translate to changes in the diagnostic plots.
+
+displays the estimated power of a visual test, computed on evaluations of the lineups generated by the non-linearity model with uniform fitted values.
 Power is plotted against the natural logarithm of the effect size $log_e(E)$, computed at a 5% significance level.

### 2023-04-08 paper.tex
working on section 5.1

@@
 
-Figure \ref{fig:polypower} displays the estimated power of a visual
-test, computed on evaluations of the lineups generated by the
-non-linearity model with uniform fitted values. Power is plotted against
-the natural logarithm of the effect size \(log_e(E)\), computed at a 5\%
-significance level. The small plots at the bottom of this figure shows a
-sequence of example residual plots of increasing effect size. These are
-for the readers to compare what a residual plot might look like at
-particular effect sizes.
+Figures \ref{fig:polypower} and \ref{fig:heterpower} present the power
+curves of various tests plotted against the effect size in the
+residuals. The effect size was computed at a 5\% significance level and
+plotted on a natural logarithmic scale. To facilitate visual calibration
+of effect size values with the corresponding diagnostic plots, a
+sequence of example residual plots with increasing effect sizes is
+provided at the bottom of these figures. These plots serve as a visual
+aid to help readers understand how different effect size values
+translate to changes in the diagnostic plots.
+
+displays the estimated power of a visual test, computed on evaluations
+of the lineups generated by the non-linearity model with uniform fitted
+values. Power is plotted against the natural logarithm of the effect
+size \(log_e(E)\), computed at a 5\% significance level. The small plots
+at the bottom of this figure shows a sequence of example residual plots
+of increasing effect size. These are for the readers to compare what a
+residual plot might look like at particular effect sizes.
 

### 2023-04-09 paper.Rmd
finished with 5.1

@@
 
-Figures \ref{fig:polypower} and \ref{fig:heterpower} present the power curves of various tests plotted against the effect size in the residuals.
-The effect size was computed at a 5% significance level and plotted on a natural logarithmic scale.
-To facilitate visual calibration of effect size values with the corresponding diagnostic plots, a sequence of example residual plots with increasing effect sizes is provided at the bottom of these figures.
-These plots serve as a visual aid to help readers understand how different effect size values translate to changes in the diagnostic plots.
-
-displays the estimated power of a visual test, computed on evaluations of the lineups generated by the non-linearity model with uniform fitted values.
-Power is plotted against the natural logarithm of the effect size $log_e(E)$, computed at a 5% significance level.
-The small plots at the bottom of this figure shows a sequence of example residual plots of increasing effect size.
-These are for the readers to compare what a residual plot might look like at particular effect sizes.
-
-<!-- Where would you draw the line of too much non-linearity in the residuals? For the RESET test this is around log effect size 2, but for the visual test it is around 3.5.-->
-
-Figure \ref{fig:heterpower} is similar to Figure \ref{fig:polypower}, but shows corresponding information on lineups produced by the heteroskedasticity model.
-In this scenario, the visual test is compared to an approximate test - Breusch-Pagan test, and two other inappropriate tests - RESET test and Shapiro-Wilk test.
-
-For non-linearity patterns, the power curve of RESET test climbs aggressively from
-`r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(0), offset0 = log(0.05/0.95)), type = "response")) * 100)`% to `r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(2.5), offset0 = log(0.05/0.95)), type = "response")) * 100)`% as $log_e(E)$ increases from 0 to 2.5, while power of other tests respond inactively to the change of effect, showing that RESET test is much more sensitive to weak non-linear structure.
-Meanwhile, no noticeable residual departures can be spotted from the example residual plots.
-
-In terms of heteroskedasticity patterns, the power of Breusch-Pagan test is constantly greater than the power of visual test.
-At $log_e(E) \approx 2.5$, where the power curve of the visual test remains at a low level, the Breusch-Pagan test has around 50% chance of rejecting $H_0$.
-Similarly, the visual feature is nearly unobservable from the example residual plots at this level of effect size.
-
-The power of visual test arises steadily as $log_e(E)$ increases from 3 to 5 for both non-linearity patterns and heteroskedasticity patterns, suggesting that the effect starts to make significant impact on the degree of the presence of the designed visual features.
-This can also be observed from the example residual plots that when $log_e(E) = 3.5$, a weak "S-shape" and a weak "right-triangle" shape are presented in Figure \ref{fig:polypower} and Figure \ref{fig:heterpower} respectively.
-The visual pattern becomes clearer as $log_e(E)$ increases.
-At $log_e(E) = 5$, the power of visual tests for both patterns reaches almost 100%.
-
-Power curves of inappropriate tests show improvement as the effect increases but at a lower rate than the visual test in both scenarios.
-This coincides the point made by @cook1982residuals that residual-based tests for a specific type of model defect may be sensitive to other types of model defects.
-The power curve of RESET test remains at around 5% in Figure \ref{fig:heterpower} since there are no non-linear terms leave out in the heteroskedasticity model and $H_0$ of the test is always satisfied.
-
-Overall, the power comparison suggests that conventional tests differs significantly from visual tests in two regression diagnostics scenarios designed by us.
-Visual test have much higher tolerance of the residual departures than the conventional test.
-Since fail to reject $H_0$ in a visual test literally means that there are no obvious visual discoveries found in the residual plot, analysts and the general public as the consumers of the output may not be convinced of the existence of significant residual departures in spite of the rejection of $H_0$ given by the conventional test.
-Even if the rejection is accepted, the model violation may be considered as impactless due to the fact that they are not clearly visible.
-Besides, the sensitivity of the conventional test to weak residual departures could also distract and discourage analysts from finding simple but good linear approximation to the data.
-The rejection of $H_0$ because of human acceptable and negligible residual departures is not practically meaningful and useful to analysts and decision makers.
-In contrast, if the strict correctness of the model assumption is of particular interest, conducting a conventional test is still the recommended choice based on the findings.
-
 ```{r}
@@
 ```{r polypower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for non-linear patterns (uniform fitted values only).
-The power curves are estimated using logistic regression, and the horizontal lines of dots represent decisions made by visual tests, with 0, 1 indicating non-reject and reject, respectively.
+The power curves are estimated using logistic regression, and the horizontal lines of dots represent non-detect and detect results from human observers for each lineup.
 The visual test has multiple power curves estimated from bootstrap samples.
@@
 ```{r heterpower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only).
-Main plot shows the power curves, with dots indicating human evaluations of lineups.
-Surrounding lines of the visual test show the estimated power of 500 bootstrap samples.
-Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by vertical lines in main plot.
-Where would you draw the line of too much heteroskedasticity in the residuals?
-For the BP test this is around log effect size 2.5, but for the visual test it is around 3.5.", dev = 'png', dpi = 450}
+Main plot shows the power curves, with dots indicating non-detect and detect in human evaluations of lineups.
+The multiple lines for the visual test arise from estimating the power on many bootstrap samples.
+The row of scatterplots plots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.", dev = 'png', dpi = 450}
 
@@
 ```
+
+Figures \ref{fig:polypower} and \ref{fig:heterpower} present the power curves of various tests plotted against the effect size in the residuals, for non-linearity and heteroskedasticity, respectively.
+In each case the power of visual test is re-calculated for multiple bootstrap samples leading to the many (orange) curves.
+The effect size was computed at a 5% significance level and plotted on a natural logarithmic scale.
+To facilitate visual calibration of effect size values with the corresponding diagnostic plots, a sequence of example residual plots with increasing effect sizes is provided at the bottom of these figures.
+These plots serve as a visual aid to help readers understand how different effect size values translate to changes in the diagnostic plots.
+The horizontal lines of dots at 0 and 1 represent the detects or non-detects made by human observers for each lineup.
+
+Figures \ref{fig:polypower} compares the power for the different tests for non-linear structure in the residuals.
+The test with the uniformly higher power is the RESET test, one that specifically tests for non-linearity.
+The power curves for the visual test are effectively a shift right from that of the RESET test.
+This means that the RESET test will reject a lower effect size (less structure) than the visual test, but otherwise the performance will be similar.
+In other words, the RESET test is more sensitive than the visual test.
+This is not necessarily a good feature, for the purposes of diagnosing model defects.
+If one scans the residual plot examples at the bottom, we might argue that the non-linearity is not sufficiently problematic until an effect size of around 3 or 3.5.
+The RESET test would reject closer to an effect size of 2, but the visual test would reject closer 3.25.
+As expected the Breusch-Pagan and Shapiro-Wilk tests have much lower power - they are not designed to detect non-linearity.
+
+For the heteroskedasticity pattern, the power of Breusch-Pagan test, designed for detecting heteroskedasticity, is uniformly higher than the other tests.
+The visual test power curve is a right shift.
+This shows a similar story to the power curves for non-linearity pattern - the conventional test is more sensitive than the visual test.
+From the example residual plots at the bottom we might argue that the heteroskedasticity becomes noticeably visible around an effect size of 3 or 3.5.
+However the Breusch-Pagan test would reject at around effect size 2.5.
+Interestingly, the power curve for the Shapiro-Wilk test (for non-normality) is only sighly different to that of the visual test, suggesting that it performs reasonably for detecting heteroskedasticity, too.
+The power curve for the Breusch-Pagan test suggests it is not useful for detecting heteroskeadsticity, as expected.
+
+Overall, the results show that the conventional tests are more sensitive than the visual test.
+The conventional tests do have higher power for the patterns they are designed to detect, and generaly unable to detect other patterns.
+The visual test doesn't require specifiying the pattern ahead of time, relying purely on whether the observed residual plot is detectably different from "good" residual plots.
+They will perform equally well regardless of the type of model defect.
+This aligns with the advice of experts on residual analysis, who consider residual plot analysis to be an indispensable tool for diagnosing model problems.
+What we gain from using a visual test for this purpose is the removal of subjective argument about whether a pattern is visible or not.
+The lineup protocol provides the calibration of detecting patterns, that if the pattern in the data plot cannot be distinguished from patterns in good residual plots, then no discernible problem with the model exists.
 
@@
 
-# Conclusion
+# Conclusions
+
+<!-- Make a note about "arguing" that there is no problem apparent in residual plot - reading a single plot -  vs the visual test which removes the argument, and provides objectivity.-->
 

### 2023-04-09 paper.Rmd
Correct the number of people filtered by attention checks
These filtered observations are not recorded by our database so there is no way to tell how many of them rather than checking the rejected people via the Prolific system.

--- 2023-04-08 paper.Rmd
@@
 
-Data collection used a total of `r vi_survey %>% count(unique_lineup_id) %>% nrow()` lineups, and resulted in a total of `r nrow(filter(vi_survey, !attention_check))` evaluations from `r vi_survey %>% count(set, exp) %>% nrow()` participants.
+Data collection used a total of `r filter(vi_survey, !attention_check) %>% count(unique_lineup_id) %>% nrow()` lineups, and resulted in a total of `r nrow(filter(vi_survey, !attention_check))` evaluations from `r vi_survey %>% count(set, exp) %>% nrow()` participants.
 Roughly half corresponded to the two models, non-linearity and heteroskedasticiy, and the three collection periods had similar numbers in each.
-<!-- `r nrow(filter(vi_survey, type == "polynomial", !attention_check))` for the non-linearity model, and `r nrow(filter(vi_survey, type == "heteroskedasticity", !attention_check))` for the heteroskedasticity model.
-This was done in three collection periods (I, II, III), with `r 160 * 18`, `r 150 * 18` and `r 133 * 18 - nrow(filter(vi_survey, null_lineup, !attention_check))` lineups evaluated by 160, 150 and 133 participants, respectively.--> Each participant received two of the `r vi_survey %>% filter(attention_check) %>% count(unique_lineup_id) %>% nrow()` attention check lineups which were used to filter results of participants who were clearly not making an honest effort (only `r vi_survey %>% filter(attention_check, !detect) %>% count(exp, set) %>% nrow()` of `r vi_survey %>% count(set, exp) %>% nrow()`).
+Each participant received two of the `r vi_survey %>% filter(attention_check) %>% count(unique_lineup_id) %>% nrow()` attention check lineups which were used to filter results of participants who were clearly not making an honest effort (only 11 of `r vi_survey %>% count(set, exp) %>% nrow() + 11`).
 To estimate $\alpha$ for calculating statistical significance (see Section \ref{sig}) there were `r nrow(filter(vi_survey, null_lineup, !attention_check))` evaluations of `r vi_survey %>% filter(null_lineup, !attention_check) %>% count(unique_lineup_id) %>% nrow()` null lineups.
@@
 More data was collected on the uniform (each evaluated by 11 participants) than the others (each evaluated by 5 participants).
-The analysis in Sections \ref{power-analysis}-\ref{hetero-analysis}, uses only results from lineups generated with uniform fitted values, a total <!--of `r length(unique(pull(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"), unique_lineup_id)))` lineups and--> `r nrow(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"))` lineup evaluations.
+The analysis in Sections \ref{power-analysis}-\ref{hetero-analysis}, uses only results from lineups generated with uniform fitted values, a total `r nrow(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"))` lineup evaluations.
 This was decided in order to compare the conventional and visual test performance for an optimal scenario.

### 2023-04-09 paper.tex

@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/polypower-1}
+
+}
+
+\caption{Comparison of power between different tests for non-linear patterns (uniform fitted values only).
+The power curves are estimated using logistic regression, and the horizontal lines of dots represent non-detect and detect results from human observers for each lineup.
+The visual test has multiple power curves estimated from bootstrap samples.
+The row of scatterplots plots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}\label{fig:polypower}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/heterpower-1}
+
+}
+
+\caption{Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only).
+Main plot shows the power curves, with dots indicating non-detect and detect in human evaluations of lineups.
+The multiple lines for the visual test arise from estimating the power on many bootstrap samples.
+The row of scatterplots plots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}\label{fig:heterpower}
+\end{figure}
+
 Figures \ref{fig:polypower} and \ref{fig:heterpower} present the power
 curves of various tests plotted against the effect size in the
-residuals. The effect size was computed at a 5\% significance level and
-plotted on a natural logarithmic scale. To facilitate visual calibration
-of effect size values with the corresponding diagnostic plots, a
-sequence of example residual plots with increasing effect sizes is
-provided at the bottom of these figures. These plots serve as a visual
-aid to help readers understand how different effect size values
-translate to changes in the diagnostic plots.
-
-displays the estimated power of a visual test, computed on evaluations
-of the lineups generated by the non-linearity model with uniform fitted
-values. Power is plotted against the natural logarithm of the effect
-size \(log_e(E)\), computed at a 5\% significance level. The small plots
-at the bottom of this figure shows a sequence of example residual plots
-of increasing effect size. These are for the readers to compare what a
-residual plot might look like at particular effect sizes.
-
-Figure \ref{fig:heterpower} is similar to Figure \ref{fig:polypower},
-but shows corresponding information on lineups produced by the
-heteroskedasticity model. In this scenario, the visual test is compared
-to an approximate test - Breusch-Pagan test, and two other inappropriate
-tests - RESET test and Shapiro-Wilk test.
-
-For non-linearity patterns, the power curve of RESET test climbs
-aggressively from 7\% to 84\% as \(log_e(E)\) increases from 0 to 2.5,
-while power of other tests respond inactively to the change of effect,
-showing that RESET test is much more sensitive to weak non-linear
-structure. Meanwhile, no noticeable residual departures can be spotted
-from the example residual plots.
-
-In terms of heteroskedasticity patterns, the power of Breusch-Pagan test
-is constantly greater than the power of visual test. At
-\(log_e(E) \approx 2.5\), where the power curve of the visual test
-remains at a low level, the Breusch-Pagan test has around 50\% chance of
-rejecting \(H_0\). Similarly, the visual feature is nearly unobservable
-from the example residual plots at this level of effect size.
-
-The power of visual test arises steadily as \(log_e(E)\) increases from
-3 to 5 for both non-linearity patterns and heteroskedasticity patterns,
-suggesting that the effect starts to make significant impact on the
-degree of the presence of the designed visual features. This can also be
-observed from the example residual plots that when \(log_e(E) = 3.5\), a
-weak ``S-shape'' and a weak ``right-triangle'' shape are presented in
-Figure \ref{fig:polypower} and Figure \ref{fig:heterpower} respectively.
-The visual pattern becomes clearer as \(log_e(E)\) increases. At
-\(log_e(E) = 5\), the power of visual tests for both patterns reaches
-almost 100\%.
-
-Power curves of inappropriate tests show improvement as the effect
-increases but at a lower rate than the visual test in both scenarios.
-This coincides the point made by \citet{cook1982residuals} that
-residual-based tests for a specific type of model defect may be
-sensitive to other types of model defects. The power curve of RESET test
-remains at around 5\% in Figure \ref{fig:heterpower} since there are no
-non-linear terms leave out in the heteroskedasticity model and \(H_0\)
-of the test is always satisfied.
-
-Overall, the power comparison suggests that conventional tests differs
-significantly from visual tests in two regression diagnostics scenarios
-designed by us. Visual test have much higher tolerance of the residual
-departures than the conventional test. Since fail to reject \(H_0\) in a
-visual test literally means that there are no obvious visual discoveries
-found in the residual plot, analysts and the general public as the
-consumers of the output may not be convinced of the existence of
-significant residual departures in spite of the rejection of \(H_0\)
-given by the conventional test. Even if the rejection is accepted, the
-model violation may be considered as impactless due to the fact that
-they are not clearly visible. Besides, the sensitivity of the
-conventional test to weak residual departures could also distract and
-discourage analysts from finding simple but good linear approximation to
-the data. The rejection of \(H_0\) because of human acceptable and
-negligible residual departures is not practically meaningful and useful
-to analysts and decision makers. In contrast, if the strict correctness
-of the model assumption is of particular interest, conducting a
-conventional test is still the recommended choice based on the findings.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/polypower-1}
-
-}
-
-\caption{Comparison of power between different tests for non-linear patterns (uniform fitted values only).
-The power curves are estimated using logistic regression, and the horizontal lines of dots represent decisions made by visual tests, with 0, 1 indicating non-reject and reject, respectively.
-The visual test has multiple power curves estimated from bootstrap samples.
-The row of scatterplots plots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}\label{fig:polypower}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/heterpower-1}
-
-}
-
-\caption{Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only).
-Main plot shows the power curves, with dots indicating human evaluations of lineups.
-Surrounding lines of the visual test show the estimated power of 500 bootstrap samples.
-Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by vertical lines in main plot.
-Where would you draw the line of too much heteroskedasticity in the residuals?
-For the BP test this is around log effect size 2.5, but for the visual test it is around 3.5.}\label{fig:heterpower}
-\end{figure}
+residuals, for non-linearity and heteroskedasticity, respectively. In
+each case the power of visual test is re-calculated for multiple
+bootstrap samples leading to the many (orange) curves. The effect size
+was computed at a 5\% significance level and plotted on a natural
+logarithmic scale. To facilitate visual calibration of effect size
+values with the corresponding diagnostic plots, a sequence of example
+residual plots with increasing effect sizes is provided at the bottom of
+these figures. These plots serve as a visual aid to help readers
+understand how different effect size values translate to changes in the
+diagnostic plots. The horizontal lines of dots at 0 and 1 represent the
+detects or non-detects made by human observers for each lineup.
+
+Figures \ref{fig:polypower} compares the power for the different tests
+for non-linear structure in the residuals. The test with the uniformly
+higher power is the RESET test, one that specifically tests for
+non-linearity. The power curves for the visual test are effectively a
+shift right from that of the RESET test. This means that the RESET test
+will reject a lower effect size (less structure) than the visual test,
+but otherwise the performance will be similar. In other words, the RESET
+test is more sensitive than the visual test. This is not necessarily a
+good feature, for the purposes of diagnosing model defects. If one scans
+the residual plot examples at the bottom, we might argue that the
+non-linearity is not sufficiently problematic until an effect size of
+around 3 or 3.5. The RESET test would reject closer to an effect size of
+2, but the visual test would reject closer 3.25. As expected the
+Breusch-Pagan and Shapiro-Wilk tests have much lower power - they are
+not designed to detect non-linearity.
+
+For the heteroskedasticity pattern, the power of Breusch-Pagan test,
+designed for detecting heteroskedasticity, is uniformly higher than the
+other tests. The visual test power curve is a right shift. This shows a
+similar story to the power curves for non-linearity pattern - the
+conventional test is more sensitive than the visual test. From the
+example residual plots at the bottom we might argue that the
+heteroskedasticity becomes noticeably visible around an effect size of 3
+or 3.5. However the Breusch-Pagan test would reject at around effect
+size 2.5. Interestingly, the power curve for the Shapiro-Wilk test (for
+non-normality) is only sighly different to that of the visual test,
+suggesting that it performs reasonably for detecting heteroskedasticity,
+too. The power curve for the Breusch-Pagan test suggests it is not
+useful for detecting heteroskeadsticity, as expected.
+
+Overall, the results show that the conventional tests are more sensitive
+than the visual test. The conventional tests do have higher power for
+the patterns they are designed to detect, and generaly unable to detect
+other patterns. The visual test doesn't require specifiying the pattern
+ahead of time, relying purely on whether the observed residual plot is
+detectably different from ``good'' residual plots. They will perform
+equally well regardless of the type of model defect. This aligns with
+the advice of experts on residual analysis, who consider residual plot
+analysis to be an indispensable tool for diagnosing model problems. What
+we gain from using a visual test for this purpose is the removal of
+subjective argument about whether a pattern is visible or not. The
+lineup protocol provides the calibration of detecting patterns, that if
+the pattern in the data plot cannot be distinguished from patterns in
+good residual plots, then no discernible problem with the model exists.
 
@@
 
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
+\hypertarget{conclusions}{%
+\section{Conclusions}\label{conclusions}}
 

### 2023-04-09 paper.Rmd
finished 5.2

--- 2023-04-09 paper.Rmd
@@
 
-```{r first-example-lineup, fig.height = 7, fig.width = 7, fig.cap = "Visual testing is conducted using a lineup, as in the example here.
+```{r first-example-lineup, fig.pos="t!", fig.height = 7, fig.width = 7, fig.cap = "Visual testing is conducted using a lineup, as in the example here.
 The residual plot computed from the observed data (plot $2^2 + 2$, exhibiting non-linearity) is embedded among 19 null plots, where the residuals are simulated from a standard error model.
@@
 
-```{r example-poly-lineup, fig.height = 7, fig.width = 7, fig.cap = "One of the lineups containing non-linearity patterns used in data collection period I.
+```{r example-poly-lineup, fig.pos="t!", fig.height = 7, fig.width = 7, fig.cap = "One of the lineups containing non-linearity patterns used in data collection period I.
 Can you spot the most different plot?
@@
 
-```{r example-heter-lineup, fig.height = 7, fig.width = 7, fig.cap = "One of the lineups containing heteroskedasticity pattern used in data collection period II.
+```{r example-heter-lineup, fig.pos="t!", fig.height = 7, fig.width = 7, fig.cap = "One of the lineups containing heteroskedasticity pattern used in data collection period II.
 Can you spot the most different plot?
@@
 
-The power comparison illustrates that appropriate conventional tests will reject $H_0$ more aggressively than visual tests.
-In this section, we explore how often they agree with each other by investigating the rejections for the two model designs based on $p$-values for each lineup.
-
-Figure \ref{fig:p-value-comparison} provides a mosaic plot showing the rejection rate of visual tests and conventional tests for both non-linearity patterns and heteroskedasticity patterns.
+The power comparison demonstrated that the appropriate conventional tests will reject more aggressively than visual tests, but we don't know how the decisions for each lineup would agree or disagree.
+Here we compare the reject or fail to reject decisions of these tests, across all the lineups.
+Figure \ref{fig:p-value-comparison} shows the agreement of the conventional and visual tests using a mosaic plot for both non-linearity patterns and heteroskedasticity patterns.
+
+For both patterns the lineups resulting in a rejection by the visual test are *all* also rejected by the conventional test, except for one from the heteroskedasticity model.
+This reflects exactly the story from the previous section, that the conventional tests reject more aggressively than the visual test.
 
@@
 In terms of lineups containing heteroskedasticity patterns, `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by conventional tests, while `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by visual tests.
-When the conventional test rejects a lineup, there is a great chance (`r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%) the visual test will also reject it.
+When the conventional test rejects a lineup, there is a good chance (`r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%) the visual test will also reject it.
 
@@
 This lineup was evaluated by 11 subjects, it has experimental factors $a = 0$ ("butterfly" shape), $b = 64$ (large variance ratio), $n = 50$ (small sample size), and a uniform distribution for the fitted values.
-It must be the small sample size that may have resulted in the lack of detection.
+It may have been the small sample size and the presence of a few outliers that may have resulted in the lack of detection by the conventional test.
 
@@
 
-```{r heter-example, fig.height = 7, fig.width = 7, fig.cap = 'The single heteroskedasticity lineup that is rejected by the visual test but not by the BP test.
+```{r heter-example, fig.pos="t!", fig.height = 7, fig.width = 7, fig.cap = 'The single heteroskedasticity lineup that is rejected by the visual test but not by the BP test.
 The data plot at panel 17 contains a "butterfly" shape.
@@
 
-```{r different-x-dist-poly-power, fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for conventional and visual tests for non-linearity and heteroskedasticity patterns.
+```{r different-x-dist-poly-power, fig.pos="t!", fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for conventional and visual tests for non-linearity and heteroskedasticity patterns.
 The power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and Breusch-Pagan tests, respectively.
@@
 Surprisingly, the fitted value distribution has produces more variability in the power of conventional tests than visual tests.
-Uneven distributions, normal and lognormal distributions, tend to yield lower power.", dev = 'png', dpi = 450, fig.pos="t"}
+Uneven distributions, normal and lognormal distributions, tend to yield lower power.", dev = 'png', dpi = 450}
 

### 2023-04-09 paper.tex
finished 5.2

@@
 
-\begin{figure}
+\begin{figure}[t!]
 
@@
 
-The power comparison illustrates that appropriate conventional tests
-will reject \(H_0\) more aggressively than visual tests. In this
-section, we explore how often they agree with each other by
-investigating the rejections for the two model designs based on
-\(p\)-values for each lineup.
-
-Figure \ref{fig:p-value-comparison} provides a mosaic plot showing the
-rejection rate of visual tests and conventional tests for both
+The power comparison demonstrated that the appropriate conventional
+tests will reject more aggressively than visual tests, but we don't know
+how the decisions for each lineup would agree or disagree. Here we
+compare the reject or fail to reject decisions of these tests, across
+all the lineups. Figure \ref{fig:p-value-comparison} shows the agreement
+of the conventional and visual tests using a mosaic plot for both
 non-linearity patterns and heteroskedasticity patterns.
+
+For both patterns the lineups resulting in a rejection by the visual
+test are \emph{all} also rejected by the conventional test, except for
+one from the heteroskedasticity model. This reflects exactly the story
+from the previous section, that the conventional tests reject more
+aggressively than the visual test.
 
@@
 rejected by conventional tests, while 56\% are rejected by visual tests.
-When the conventional test rejects a lineup, there is a great chance
+When the conventional test rejects a lineup, there is a good chance
 (73\%) the visual test will also reject it.
@@
 ratio), \(n = 50\) (small sample size), and a uniform distribution for
-the fitted values. It must be the small sample size that may have
-resulted in the lack of detection.
+the fitted values. It may have been the small sample size and the
+presence of a few outliers that may have resulted in the lack of
+detection by the conventional test.
 
@@
 
-\begin{figure}[t]
+\begin{figure}[t!]
 

### 2023-04-09 paper.tex
finished 5.2

@@
 
-Data collection used a total of 1176 lineups, and resulted in a total of
+Data collection used a total of 1152 lineups, and resulted in a total of
 7974 evaluations from 443 participants. Roughly half corresponded to the
@@
 results of participants who were clearly not making an honest effort
-(only 4 of 443). To estimate \(\alpha\) for calculating statistical
+(only 11 of 454). To estimate \(\alpha\) for calculating statistical
 significance (see Section \ref{sig}) there were 720 evaluations of 36

### 2023-04-09 paper.Rmd
placement of figures 11, 12
fixing spelling in 5.1

@@
 
+Figures \ref{fig:polypower} and \ref{fig:heterpower} present the power curves of various tests plotted against the effect size in the residuals, for non-linearity and heteroskedasticity, respectively.
+In each case the power of visual test is re-calculated for multiple bootstrap samples leading to the many (orange) curves.
+The effect size was computed at a 5% significance level and plotted on a natural logarithmic scale.
+To facilitate visual calibration of effect size values with the corresponding diagnostic plots, a sequence of example residual plots with increasing effect sizes is provided at the bottom of these figures.
+These plots serve as a visual aid to help readers understand how different effect size values translate to changes in the diagnostic plots.
+The horizontal lines of dots at 0 and 1 represent the detects or non-detects made by human observers for each lineup.
+
 ```{r cache = TRUE}
@@
 
-Figures \ref{fig:polypower} and \ref{fig:heterpower} present the power curves of various tests plotted against the effect size in the residuals, for non-linearity and heteroskedasticity, respectively.
-In each case the power of visual test is re-calculated for multiple bootstrap samples leading to the many (orange) curves.
-The effect size was computed at a 5% significance level and plotted on a natural logarithmic scale.
-To facilitate visual calibration of effect size values with the corresponding diagnostic plots, a sequence of example residual plots with increasing effect sizes is provided at the bottom of these figures.
-These plots serve as a visual aid to help readers understand how different effect size values translate to changes in the diagnostic plots.
-The horizontal lines of dots at 0 and 1 represent the detects or non-detects made by human observers for each lineup.
-
 Figures \ref{fig:polypower} compares the power for the different tests for non-linear structure in the residuals.
@@
 Overall, the results show that the conventional tests are more sensitive than the visual test.
-The conventional tests do have higher power for the patterns they are designed to detect, and generaly unable to detect other patterns.
-The visual test doesn't require specifiying the pattern ahead of time, relying purely on whether the observed residual plot is detectably different from "good" residual plots.
+The conventional tests do have higher power for the patterns they are designed to detect, and are generally unable to detect other patterns.
+The visual test doesn't require specifying the pattern ahead of time, relying purely on whether the observed residual plot is detectably different from "good" residual plots.
 They will perform equally well regardless of the type of model defect.

### 2023-04-09 paper.tex
placement of figures 11, 12
fixing spelling in 5.1

@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/polypower-1}
-
-}
-
-\caption{Comparison of power between different tests for non-linear patterns (uniform fitted values only).
-The power curves are estimated using logistic regression, and the horizontal lines of dots represent non-detect and detect results from human observers for each lineup.
-The visual test has multiple power curves estimated from bootstrap samples.
-The row of scatterplots plots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}\label{fig:polypower}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/heterpower-1}
-
-}
-
-\caption{Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only).
-Main plot shows the power curves, with dots indicating non-detect and detect in human evaluations of lineups.
-The multiple lines for the visual test arise from estimating the power on many bootstrap samples.
-The row of scatterplots plots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}\label{fig:heterpower}
-\end{figure}
-
 Figures \ref{fig:polypower} and \ref{fig:heterpower} present the power
@@
 detects or non-detects made by human observers for each lineup.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/polypower-1}
+
+}
+
+\caption{Comparison of power between different tests for non-linear patterns (uniform fitted values only).
+The power curves are estimated using logistic regression, and the horizontal lines of dots represent non-detect and detect results from human observers for each lineup.
+The visual test has multiple power curves estimated from bootstrap samples.
+The row of scatterplots plots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}\label{fig:polypower}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/heterpower-1}
+
+}
+
+\caption{Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only).
+Main plot shows the power curves, with dots indicating non-detect and detect in human evaluations of lineups.
+The multiple lines for the visual test arise from estimating the power on many bootstrap samples.
+The row of scatterplots plots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}\label{fig:heterpower}
+\end{figure}
 
@@
 than the visual test. The conventional tests do have higher power for
-the patterns they are designed to detect, and generaly unable to detect
-other patterns. The visual test doesn't require specifiying the pattern
-ahead of time, relying purely on whether the observed residual plot is
-detectably different from ``good'' residual plots. They will perform
-equally well regardless of the type of model defect. This aligns with
-the advice of experts on residual analysis, who consider residual plot
-analysis to be an indispensable tool for diagnosing model problems. What
-we gain from using a visual test for this purpose is the removal of
+the patterns they are designed to detect, and are generally unable to
+detect other patterns. The visual test doesn't require specifying the
+pattern ahead of time, relying purely on whether the observed residual
+plot is detectably different from ``good'' residual plots. They will
+perform equally well regardless of the type of model defect. This aligns
+with the advice of experts on residual analysis, who consider residual
+plot analysis to be an indispensable tool for diagnosing model problems.
+What we gain from using a visual test for this purpose is the removal of
 subjective argument about whether a pattern is visible or not. The

### 2023-04-09 paper.Rmd
fixing spelling in 5.1

@@
 This aligns with the advice of experts on residual analysis, who consider residual plot analysis to be an indispensable tool for diagnosing model problems.
-What we gain from using a visual test for this purpose is the removal of subjective argument about whether a pattern is visible or not.
-The lineup protocol provides the calibration of detecting patterns, that if the pattern in the data plot cannot be distinguished from patterns in good residual plots, then no discernible problem with the model exists.
+What we gain from using a visual test for this purpose is the removal of any subjective arguments about whether a pattern is visible or not.
+The lineup protocol provides the calibration for detecting patterns: that if the pattern in the data plot cannot be distinguished from patterns in good residual plots, then no discernible problem with the model exists.
 

### 2023-04-09 paper.tex
fixing spelling in 5.1

@@
 What we gain from using a visual test for this purpose is the removal of
-subjective argument about whether a pattern is visible or not. The
-lineup protocol provides the calibration of detecting patterns, that if
+any subjective arguments about whether a pattern is visible or not. The
+lineup protocol provides the calibration for detecting patterns: that if
 the pattern in the data plot cannot be distinguished from patterns in

### 2023-04-09 paper.Rmd
abb BP and SW

@@
 However, if the distribution is skewed, bimodal, multimodal, or contains outliers, there is a cause for concern.
-One could also inspect the distribution by conducting a goodness of fit test, such as the Shapiro-Wilk Normality test [@shapiro1965analysis].
+One could also inspect the distribution by conducting a goodness of fit test, such as the Shapiro-Wilk (SW) Normality test [@shapiro1965analysis].
 
@@
 It is especially common, particularly among students, to misinterpret patterns that are random noise and random deviation from a model [@loy2021bringing].
-It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.
+It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan (BP) test [@breusch_simple_1979] for heteroskedasticity.
 
@@
 Many different hypothesis tests are available to detect specific model defects.
-For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution.
+For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the BP test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution.
 To test specific forms of non-linearity, one may apply the F-test as a model structural test to examine the significance of specific polynomial and non-linear forms of the predictors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].
-The Shapiro-Wilk test [@shapiro1965analysis] is the most widely used test of non-normality included by many of the statistical software programs.
+The SW test [@shapiro1965analysis] is the most widely used test of non-normality included by many of the statistical software programs.
 The Jarque--Bera test [@jarque1980efficient] is also used to directly check whether the sample skewness and kurtosis match a normal distribution.
 
-Table \ref{tab:example-residual-plot-table} displays the $p$-values from the RESET, Breusch-Pagan and Shapiro-Wilk tests applied to the residual plots in Figure \ref{fig:residual-plot-common-departures}.
-The RESET test and Breusch-Pagan test were computed using the `resettest` and `bptest` functions from the R package `lmtest`, respectively.
-The Shapiro-Wilk test was computed using the `shapiro.test` from the core R package `stats`.
+Table \ref{tab:example-residual-plot-table} displays the $p$-values from the RESET, BP and SW tests applied to the residual plots in Figure \ref{fig:residual-plot-common-departures}.
+The RESET test and BP test were computed using the `resettest` and `bptest` functions from the R package `lmtest`, respectively.
+The SW test was computed using the `shapiro.test` from the core R package `stats`.
 (The the R package `skedastic` [@skedastic] contains a large collection of tests for heteroskedasticity.) Although, the RESET test is exact, it requires the selection of a power parameter.
 According to @ramsey_tests_1969, a power of four is recommended, which we adopted in our analysis.
-The Breusch-Pagan and Shapiro-Wilk tests are approximate.
-
-We would expect the RESET test for non-linearity to reject residual plot B, the Breusch-Pagan test for heteroskedasticity to reject the residual plot C, and Shapiro-Wilk test for non-normality to reject residual plot D, which they all do and correctly fail to reject residual plot A.
-Interestingly, the Breusch-Pagan and Shapiro-Wilk tests also reject the residual plots exhibiting structure that they weren't designed for.
+The BP and SW tests are approximate.
+
+We would expect the RESET test for non-linearity to reject residual plot B, the BP test for heteroskedasticity to reject the residual plot C, and SW test for non-normality to reject residual plot D, which they all do and correctly fail to reject residual plot A.
+Interestingly, the BP and SW tests also reject the residual plots exhibiting structure that they weren't designed for.
 @cook1982residuals discusses that most residual-based tests for particular types of departure from model assumptions are also sensitive to other types of departures, that is, correctly rejected but for the wrong reason, a phenomenon known as the "Type III error".
@@
 As discussed in Section \ref{conventionally-testing-for-departures}, many conventional tests are available for detecting residual departures.
-Implementation-wise, the built-in R package `stats` provides some commonly used residual-based tests, such as Shapiro-Wilk test.
+Implementation-wise, the built-in R package `stats` provides some commonly used residual-based tests, such as SW test.
 A more comprehensive collection of regression diagnostics tests can be found in the R package `lmtest` [@lmtest].
@@
 
-We pick RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`.
+We pick RESET test (`resettest`) and BP test (`bptest`) from the R package `lmtest`, and SW test (`shapiro.test`) from the built-in R package `stats`.
 Among them, RESET test is the only exact and appropriate test in this scenario.
-Both the Breusch-Pagan test and the Shapiro-Wilk test are approximate and inappropriate tests.
+Both the BP test and the SW test are approximate and inappropriate tests.
 Their estimated power is shown in Figure \ref{fig:polypower}.
@@
 Thus, we follow this guideline to conduct the RESET test.
-For the Breusch-Pagan test, the choice of predictors in the auxiliary regression is left to the user [@breusch_simple_1979].
-But as @waldman1983note suggested, it is a good choice for the set of auxiliary predictors in the Breusch-Pagan test be the same as the White test.
+For the BP test, the choice of predictors in the auxiliary regression is left to the user [@breusch_simple_1979].
+But as @waldman1983note suggested, it is a good choice for the set of auxiliary predictors in the BP test be the same as the White test.
 Thus, we include both $x$ and $x^2$ in the auxiliary regression.
@@
                              "RESET", 
-                             "Breusch-Pagan", 
-                             "Shapiro–Wilk"),
+                             "BP", 
+                             "SW"),
                align = "llrrr",
@@
                caption  = 'Statistical significance testing for departures from good residuals for plots in Figure \\ref{fig:residual-plot-common-departures}.
-Shown are the $p$-values calculated for the RESET, the Breusch-Pagan and the Shapiro–Wilk tests.
+Shown are the $p$-values calculated for the RESET, the BP and the SW tests.
 The good residual plot (A) is judged a good residual plot, as expected, by all tests.
@@
 For example, the RESET test is formulated as $H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0$ against $H_a: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0$, from $y = \tau_0 + \sum_{i=1}^{p}\tau_px_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)$.
-Similarly, the Breusch-Pagan test is designed to specifically test $H_0:$ *error variances are all equal* ($\zeta_i=0 \text{ for } i=1,..,p$) versus the alternative $H_a:$ *that the error variances are a multiplicative function of one or more variables* ($\text{at least one } \zeta_i\neq 0$) from $e^2 = \zeta_0 + \sum_{i=1}^{p}\zeta_i x_i + u, ~ u\sim N(0,\sigma_u^2)$.
+Similarly, the BP test is designed to specifically test $H_0:$ *error variances are all equal* ($\zeta_i=0 \text{ for } i=1,..,p$) versus the alternative $H_a:$ *that the error variances are a multiplicative function of one or more variables* ($\text{at least one } \zeta_i\neq 0$) from $e^2 = \zeta_0 + \sum_{i=1}^{p}\zeta_i x_i + u, ~ u\sim N(0,\sigma_u^2)$.
 
@@
 The RESET test would reject closer to an effect size of 2, but the visual test would reject closer 3.25.
-As expected the Breusch-Pagan and Shapiro-Wilk tests have much lower power - they are not designed to detect non-linearity.
-
-For the heteroskedasticity pattern, the power of Breusch-Pagan test, designed for detecting heteroskedasticity, is uniformly higher than the other tests.
+As expected the BP and SW tests have much lower power - they are not designed to detect non-linearity.
+
+For the heteroskedasticity pattern, the power of BP test, designed for detecting heteroskedasticity, is uniformly higher than the other tests.
 The visual test power curve is a right shift.
@@
 From the example residual plots at the bottom we might argue that the heteroskedasticity becomes noticeably visible around an effect size of 3 or 3.5.
-However the Breusch-Pagan test would reject at around effect size 2.5.
-Interestingly, the power curve for the Shapiro-Wilk test (for non-normality) is only sighly different to that of the visual test, suggesting that it performs reasonably for detecting heteroskedasticity, too.
-The power curve for the Breusch-Pagan test suggests it is not useful for detecting heteroskeadsticity, as expected.
+However the BP test would reject at around effect size 2.5.
+Interestingly, the power curve for the SW test (for non-normality) is only sighly different to that of the visual test, suggesting that it performs reasonably for detecting heteroskedasticity, too.
+The power curve for the BP test suggests it is not useful for detecting heteroskeadsticity, as expected.
 
@@
 This is reflected by the visual test $p\text{-value} = 0.026$.
-But the Breusch-Pagan test $p\text{-value} = 0.056$, is slightly above the significance cutoff of $0.05$.
+But the BP test $p\text{-value} = 0.056$, is slightly above the significance cutoff of $0.05$.
 This lineup was evaluated by 11 subjects, it has experimental factors $a = 0$ ("butterfly" shape), $b = 64$ (large variance ratio), $n = 50$ (small sample size), and a uniform distribution for the fitted values.
@@
 ```{r different-x-dist-poly-power, fig.pos="t!", fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for conventional and visual tests for non-linearity and heteroskedasticity patterns.
-The power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and Breusch-Pagan tests, respectively.
+The power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and BP tests, respectively.
 Power curves of visual tests are estimated using five evaluations on each lineup.
@@
 Figure \ref{fig:different-x-dist-poly-power} examines the impact of the fitted value distribution on the power of conventional (top row) and visual (bottom row) tests for both the non-linearity (left column) and heteroskedasticity (right column) patterns.
-For conventional tests, only the power curves of appropriate tests which are RESET tests and Breusch-Pagan tests for non-linearity pattern and heteroskedasticity pattern respectively are shown.
+For conventional tests, only the power curves of appropriate tests which are RESET tests and BP tests for non-linearity pattern and heteroskedasticity pattern respectively are shown.
 For visual tests, note that, more evaluations on lineup with uniform fitted value distribution were collected, so to have a fair comparison, we randomly sample five from the total eleven evaluations to estimate the power curves.

### 2023-04-09 paper.Rmd
fixing spelling in 5.2

--- 2023-04-09 paper.Rmd
@@
 In terms of lineups containing heteroskedasticity patterns, `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by conventional tests, while `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by visual tests.
-When the conventional test rejects a lineup, there is a good chance (`r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%) the visual test will also reject it.
+The visual test rejects `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the lineups that the conventional visual test reject.
 

### 2023-04-09 paper.Rmd
fixing typos in 5.3

@@
 
-## Statistical significance~\label{sig}
+## Statistical significance\label{sig}
 
@@
 
-## Power comparison of the tests~\label{power-analysis}
+## Power comparison of the tests\label{power-analysis}
 
@@
 
-## Comparison of test decisions based on $p$-values~\label{p-value}
+## Comparison of test decisions based on $p$-values\label{p-value}
 
@@
 
-## Effect of amount of non-linearity~\label{nonlin-analysis}
+## Effect of amount of non-linearity\label{nonlin-analysis}
 
@@
 From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves mostly align with these expectations, except for the "M" shape, which is as easy to be detected as the "S" shape.
-This suggests a benefit of the visual test, the knowing the shape ahead of time is not needed for its application.
+This suggests a benefit of the visual test: knowing the shape ahead of time is *not* needed for its application.
 
@@
 
-## Effect of shape of heteroskedasticity~\label{hetero-analysis}
+## Effect of shape of heteroskedasticity\label{hetero-analysis}
 

### 2023-04-09 paper.tex
fixing spelling in 5.2
fixing typos in 5.3

@@
 the distribution by conducting a goodness of fit test, such as the
-Shapiro-Wilk Normality test \citep{shapiro1965analysis}.
+Shapiro-Wilk (SW) Normality test \citep{shapiro1965analysis}.
 
@@
 for non-linear dependence \citep{ramsey_tests_1969}, and use a
-Breusch-Pagan test \citep{breusch_simple_1979} for heteroskedasticity.
+Breusch-Pagan (BP) test \citep{breusch_simple_1979} for
+heteroskedasticity.
 
@@
 tested by applying the White test
-\citep{white_heteroskedasticity-consistent_1980} or the Breusch-Pagan
-test \citep{breusch_simple_1979}, which are both derived from the
-Lagrange multiplier test \citep{silvey1959lagrangian} principle that
-relies on the asymptotic properties of the null distribution. To test
-specific forms of non-linearity, one may apply the F-test as a model
-structural test to examine the significance of specific polynomial and
-non-linear forms of the predictors, or the significance of proxy
-variables as in the Ramsey Regression Equation Specification Error Test
-(RESET) \citep{ramsey_tests_1969}. The Shapiro-Wilk test
-\citep{shapiro1965analysis} is the most widely used test of
-non-normality included by many of the statistical software programs. The
-Jarque--Bera test \citep{jarque1980efficient} is also used to directly
-check whether the sample skewness and kurtosis match a normal
-distribution.
+\citep{white_heteroskedasticity-consistent_1980} or the BP test
+\citep{breusch_simple_1979}, which are both derived from the Lagrange
+multiplier test \citep{silvey1959lagrangian} principle that relies on
+the asymptotic properties of the null distribution. To test specific
+forms of non-linearity, one may apply the F-test as a model structural
+test to examine the significance of specific polynomial and non-linear
+forms of the predictors, or the significance of proxy variables as in
+the Ramsey Regression Equation Specification Error Test (RESET)
+\citep{ramsey_tests_1969}. The SW test \citep{shapiro1965analysis} is
+the most widely used test of non-normality included by many of the
+statistical software programs. The Jarque--Bera test
+\citep{jarque1980efficient} is also used to directly check whether the
+sample skewness and kurtosis match a normal distribution.
 
 Table \ref{tab:example-residual-plot-table} displays the \(p\)-values
-from the RESET, Breusch-Pagan and Shapiro-Wilk tests applied to the
-residual plots in Figure \ref{fig:residual-plot-common-departures}. The
-RESET test and Breusch-Pagan test were computed using the
-\texttt{resettest} and \texttt{bptest} functions from the R package
-\texttt{lmtest}, respectively. The Shapiro-Wilk test was computed using
-the \texttt{shapiro.test} from the core R package \texttt{stats}. (The
-the R package \texttt{skedastic} \citep{skedastic} contains a large
-collection of tests for heteroskedasticity.) Although, the RESET test is
-exact, it requires the selection of a power parameter. According to
-\citet{ramsey_tests_1969}, a power of four is recommended, which we
-adopted in our analysis. The Breusch-Pagan and Shapiro-Wilk tests are
-approximate.
+from the RESET, BP and SW tests applied to the residual plots in Figure
+\ref{fig:residual-plot-common-departures}. The RESET test and BP test
+were computed using the \texttt{resettest} and \texttt{bptest} functions
+from the R package \texttt{lmtest}, respectively. The SW test was
+computed using the \texttt{shapiro.test} from the core R package
+\texttt{stats}. (The the R package \texttt{skedastic} \citep{skedastic}
+contains a large collection of tests for heteroskedasticity.) Although,
+the RESET test is exact, it requires the selection of a power parameter.
+According to \citet{ramsey_tests_1969}, a power of four is recommended,
+which we adopted in our analysis. The BP and SW tests are approximate.
 
 We would expect the RESET test for non-linearity to reject residual plot
-B, the Breusch-Pagan test for heteroskedasticity to reject the residual
-plot C, and Shapiro-Wilk test for non-normality to reject residual plot
-D, which they all do and correctly fail to reject residual plot A.
-Interestingly, the Breusch-Pagan and Shapiro-Wilk tests also reject the
-residual plots exhibiting structure that they weren't designed for.
-\citet{cook1982residuals} discusses that most residual-based tests for
-particular types of departure from model assumptions are also sensitive
-to other types of departures, that is, correctly rejected but for the
-wrong reason, a phenomenon known as the ``Type III error''. Also, some
-types of departure can also have elements of other types of departure,
-for example, non-linearity could be viewed as heteroskedasticity.
-Additionally, other problesms such as outliers can trigger the rejection
-\citep{cook_applied_1999}.
+B, the BP test for heteroskedasticity to reject the residual plot C, and
+SW test for non-normality to reject residual plot D, which they all do
+and correctly fail to reject residual plot A. Interestingly, the BP and
+SW tests also reject the residual plots exhibiting structure that they
+weren't designed for. \citet{cook1982residuals} discusses that most
+residual-based tests for particular types of departure from model
+assumptions are also sensitive to other types of departures, that is,
+correctly rejected but for the wrong reason, a phenomenon known as the
+``Type III error''. Also, some types of departure can also have elements
+of other types of departure, for example, non-linearity could be viewed
+as heteroskedasticity. Additionally, other problesms such as outliers
+can trigger the rejection \citep{cook_applied_1999}.
 
@@
 \caption{\label{tab:example-residual-plot-table}Statistical significance testing for departures from good residuals for plots in Figure \ref{fig:residual-plot-common-departures}.
-Shown are the $p$-values calculated for the RESET, the Breusch-Pagan and the Shapiro–Wilk tests.
+Shown are the $p$-values calculated for the RESET, the BP and the SW tests.
 The good residual plot (A) is judged a good residual plot, as expected, by all tests.
@@
 \toprule
-Plot & Departures & RESET & Breusch-Pagan & Shapiro–Wilk\\
+Plot & Departures & RESET & BP & SW\\
 \midrule
@@
 \(y = \tau_0 + \sum_{i=1}^{p}\tau_px_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)\).
-Similarly, the Breusch-Pagan test is designed to specifically test
-\(H_0:\) \emph{error variances are all equal}
+Similarly, the BP test is designed to specifically test \(H_0:\)
+\emph{error variances are all equal}
 (\(\zeta_i=0 \text{ for } i=1,..,p\)) versus the alternative \(H_a:\)
@@
 around 3 or 3.5. The RESET test would reject closer to an effect size of
-2, but the visual test would reject closer 3.25. As expected the
-Breusch-Pagan and Shapiro-Wilk tests have much lower power - they are
-not designed to detect non-linearity.
-
-For the heteroskedasticity pattern, the power of Breusch-Pagan test,
-designed for detecting heteroskedasticity, is uniformly higher than the
-other tests. The visual test power curve is a right shift. This shows a
-similar story to the power curves for non-linearity pattern - the
-conventional test is more sensitive than the visual test. From the
-example residual plots at the bottom we might argue that the
-heteroskedasticity becomes noticeably visible around an effect size of 3
-or 3.5. However the Breusch-Pagan test would reject at around effect
-size 2.5. Interestingly, the power curve for the Shapiro-Wilk test (for
-non-normality) is only sighly different to that of the visual test,
-suggesting that it performs reasonably for detecting heteroskedasticity,
-too. The power curve for the Breusch-Pagan test suggests it is not
-useful for detecting heteroskeadsticity, as expected.
+2, but the visual test would reject closer 3.25. As expected the BP and
+SW tests have much lower power - they are not designed to detect
+non-linearity.
+
+For the heteroskedasticity pattern, the power of BP test, designed for
+detecting heteroskedasticity, is uniformly higher than the other tests.
+The visual test power curve is a right shift. This shows a similar story
+to the power curves for non-linearity pattern - the conventional test is
+more sensitive than the visual test. From the example residual plots at
+the bottom we might argue that the heteroskedasticity becomes noticeably
+visible around an effect size of 3 or 3.5. However the BP test would
+reject at around effect size 2.5. Interestingly, the power curve for the
+SW test (for non-normality) is only sighly different to that of the
+visual test, suggesting that it performs reasonably for detecting
+heteroskedasticity, too. The power curve for the BP test suggests it is
+not useful for detecting heteroskeadsticity, as expected.
 
@@
 rejected by conventional tests, while 56\% are rejected by visual tests.
-When the conventional test rejects a lineup, there is a good chance
-(73\%) the visual test will also reject it.
+The visual test rejects 73\% of the lineups that the conventional visual
+test reject.
 
@@
 size (\(log_e(E)=4.02\)). This is reflected by the visual test
-\(p\text{-value} = 0.026\). But the Breusch-Pagan test
-\(p\text{-value} = 0.056\), is slightly above the significance cutoff of
-\(0.05\). This lineup was evaluated by 11 subjects, it has experimental
-factors \(a = 0\) (``butterfly'' shape), \(b = 64\) (large variance
-ratio), \(n = 50\) (small sample size), and a uniform distribution for
-the fitted values. It may have been the small sample size and the
-presence of a few outliers that may have resulted in the lack of
-detection by the conventional test.
+\(p\text{-value} = 0.026\). But the BP test \(p\text{-value} = 0.056\),
+is slightly above the significance cutoff of \(0.05\). This lineup was
+evaluated by 11 subjects, it has experimental factors \(a = 0\)
+(``butterfly'' shape), \(b = 64\) (large variance ratio), \(n = 50\)
+(small sample size), and a uniform distribution for the fitted values.
+It may have been the small sample size and the presence of a few
+outliers that may have resulted in the lack of detection by the
+conventional test.
 
@@
 is as easy to be detected as the ``S'' shape. This suggests a benefit of
-the visual test, the knowing the shape ahead of time is not needed for
-its application.
+the visual test: knowing the shape ahead of time is \emph{not} needed
+for its application.
 
@@
 \caption{Comparison of power on lineups with different fitted value distributions for conventional and visual tests for non-linearity and heteroskedasticity patterns.
-The power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and Breusch-Pagan tests, respectively.
+The power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and BP tests, respectively.
 Power curves of visual tests are estimated using five evaluations on each lineup.
@@
 heteroskedasticity (right column) patterns. For conventional tests, only
-the power curves of appropriate tests which are RESET tests and
-Breusch-Pagan tests for non-linearity pattern and heteroskedasticity
-pattern respectively are shown. For visual tests, note that, more
-evaluations on lineup with uniform fitted value distribution were
-collected, so to have a fair comparison, we randomly sample five from
-the total eleven evaluations to estimate the power curves. Re-sampling
-produces the multiple curves for the uniform, and provides an indication
-of the variation of the power.
+the power curves of appropriate tests which are RESET tests and BP tests
+for non-linearity pattern and heteroskedasticity pattern respectively
+are shown. For visual tests, note that, more evaluations on lineup with
+uniform fitted value distribution were collected, so to have a fair
+comparison, we randomly sample five from the total eleven evaluations to
+estimate the power curves. Re-sampling produces the multiple curves for
+the uniform, and provides an indication of the variation of the power.
 

### 2023-04-09 paper.tex
removed accidental ~ in headings

@@
 \subsection{\texorpdfstring{Statistical
-significance\textasciitilde{}\label{sig}}{Statistical significance\textasciitilde{}}}\label{statistical-significance}}
+significance\label{sig}}{Statistical significance}}\label{statistical-significance}}
 
@@
 \subsection{\texorpdfstring{Power comparison of the
-tests\textasciitilde{}\label{power-analysis}}{Power comparison of the tests\textasciitilde{}}}\label{power-comparison-of-the-tests}}
+tests\label{power-analysis}}{Power comparison of the tests}}\label{power-comparison-of-the-tests}}
 
@@
 \subsection{\texorpdfstring{Comparison of test decisions based on
-\(p\)-values\textasciitilde{}\label{p-value}}{Comparison of test decisions based on p-values\textasciitilde{}}}\label{comparison-of-test-decisions-based-on-p-values}}
+\(p\)-values\label{p-value}}{Comparison of test decisions based on p-values}}\label{comparison-of-test-decisions-based-on-p-values}}
 
@@
 \subsection{\texorpdfstring{Effect of amount of
-non-linearity\textasciitilde{}\label{nonlin-analysis}}{Effect of amount of non-linearity\textasciitilde{}}}\label{effect-of-amount-of-non-linearity}}
+non-linearity\label{nonlin-analysis}}{Effect of amount of non-linearity}}\label{effect-of-amount-of-non-linearity}}
 
@@
 \subsection{\texorpdfstring{Effect of shape of
-heteroskedasticity\textasciitilde{}\label{hetero-analysis}}{Effect of shape of heteroskedasticity\textasciitilde{}}}\label{effect-of-shape-of-heteroskedasticity}}
+heteroskedasticity\label{hetero-analysis}}{Effect of shape of heteroskedasticity}}\label{effect-of-shape-of-heteroskedasticity}}
 

### 2023-04-09 paper.Rmd
slight word-smitting of intro

@@
 However, if the distribution is skewed, bimodal, multimodal, or contains outliers, there is a cause for concern.
-One could also inspect the distribution by conducting a goodness of fit test, such as the Shapiro-Wilk (SW) Normality test [@shapiro1965analysis].
-
-More typically, residuals will be plotted, as a scatter plot against the predicted values and each of the explanatory variables to scrutinize their relationships.
-If there are any visually discoverable patterns, the model is potentially misspecified.
+One could also inspect the distribution by conducting a goodness-of-fit test, such as the Shapiro-Wilk (SW) Normality test [@shapiro1965analysis].
+
+More typically, a scatterplot of residuals plotted against the fitted values and each of the explanatory variables, is made to scrutinize their relationships.
+If there are any visually discoverable patterns, the model is potentially inadequate or incorrectly specifed.
 In general, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity.
-However, correctly judging a residual plot where no pattern exists can be a painstakingly difficult task for humans.
+However, correctly judging a residual plot where no pattern exists can be a difficult task for humans.
 It is especially common, particularly among students, to misinterpret patterns that are random noise and random deviation from a model [@loy2021bringing].
-It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan (BP) test [@breusch_simple_1979] for heteroskedasticity.
-
-Abundance of literature describe appropriate diagnostic methods for linear regression, e.g.
+There are also numerous conventional hypothesis tests available to test for non-linear dependence [@ramsey_tests_1969], and heteroskedasticity  [@breusch_simple_1979].
+
+An abundance of literature describe appropriate diagnostic methods for linear regression, e.g.
 @draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals.
-All these writings consider plotting residuals to be a standard technique that should be examined routinely in all regression modelling problems.
+All these writings consider plotting residuals to be an essential tool for diagnosing regression modelling problems.
 In addition, @draper1998applied and @belsley_regression_1980 believe that residual plots are usually revealing when the assumptions are violated.
 @cook_applied_1999 thinks formal tests and graphical procedures are complementary and both have a place in residual analysis, but they focus on graphical methods rather than on formal testing, as they are easier to use.
-@montgomery1982introduction even suggests that residual plots are more informative in most practical situations than the corresponding formal tests, and statistical tests on regression model residuals are not widely used based on their experience.
+@montgomery1982introduction even suggests that residual plots are more informative in most practical situations than the corresponding conventional hypothesis tests.
 
@@
 To address this issue, @majumder_validation_2013 models the probability of a subject $j$ correctly picking the data plot from a lineup $l$ using a mixed-effect logistic regression, with subjects treated as random effect.
-Then, the estimated power of a visual test evaluated by a single subject is the predicted value obtained from the mix-effect model.
-However, this mix-effect model does not work with scenario where subjects are asked to select one or more most different plots.
+Then, the estimated power of a visual test evaluated by a single subject is the predicted value obtained from the mixed effects model.
+However, this mixed effects model does not work with scenario where subjects are asked to select one or more most different plots.
 In this scenario, having the probability of a subject $j$ correctly picking the data plot from a lineup $l$ is insufficient to determine the power of a visual test because it does not provide information about the number of selections made by the subject for the calculation of the $p$-value (See Equation \ref{eq:pvalue-beta-binomial-approx}).

### 2023-04-09 paper.tex
slight word-smitting of intro

@@
 contains outliers, there is a cause for concern. One could also inspect
-the distribution by conducting a goodness of fit test, such as the
+the distribution by conducting a goodness-of-fit test, such as the
 Shapiro-Wilk (SW) Normality test \citep{shapiro1965analysis}.
 
-More typically, residuals will be plotted, as a scatter plot against the
-predicted values and each of the explanatory variables to scrutinize
+More typically, a scatterplot of residuals plotted against the fitted
+values and each of the explanatory variables, is made to scrutinize
 their relationships. If there are any visually discoverable patterns,
-the model is potentially misspecified. In general, one looks for
-noticeable departures from the model like non-linear dependency or
-heteroskedasticity. However, correctly judging a residual plot where no
-pattern exists can be a painstakingly difficult task for humans. It is
+the model is potentially inadequate or incorrectly specifed. In general,
+one looks for noticeable departures from the model like non-linear
+dependency or heteroskedasticity. However, correctly judging a residual
+plot where no pattern exists can be a difficult task for humans. It is
 especially common, particularly among students, to misinterpret patterns
 that are random noise and random deviation from a model
-\citep{loy2021bringing}. It is also possible to conduct hypothesis tests
-for non-linear dependence \citep{ramsey_tests_1969}, and use a
-Breusch-Pagan (BP) test \citep{breusch_simple_1979} for
-heteroskedasticity.
-
-Abundance of literature describe appropriate diagnostic methods for
+\citep{loy2021bringing}. There are also numerous conventional hypothesis
+tests available to test for non-linear dependence
+\citep{ramsey_tests_1969}, and heteroskedasticity
+\citep{breusch_simple_1979}.
+
+An abundance of literature describe appropriate diagnostic methods for
 linear regression, e.g. \citet{draper1998applied},
@@
 \citet{cook_applied_1999} and \citet{cook1982residuals}. All these
-writings consider plotting residuals to be a standard technique that
-should be examined routinely in all regression modelling problems. In
-addition, \citet{draper1998applied} and \citet{belsley_regression_1980}
-believe that residual plots are usually revealing when the assumptions
-are violated. \citet{cook_applied_1999} thinks formal tests and
-graphical procedures are complementary and both have a place in residual
-analysis, but they focus on graphical methods rather than on formal
-testing, as they are easier to use. \citet{montgomery1982introduction}
-even suggests that residual plots are more informative in most practical
-situations than the corresponding formal tests, and statistical tests on
-regression model residuals are not widely used based on their
-experience.
+writings consider plotting residuals to be an essential tool for
+diagnosing regression modelling problems. In addition,
+\citet{draper1998applied} and \citet{belsley_regression_1980} believe
+that residual plots are usually revealing when the assumptions are
+violated. \citet{cook_applied_1999} thinks formal tests and graphical
+procedures are complementary and both have a place in residual analysis,
+but they focus on graphical methods rather than on formal testing, as
+they are easier to use. \citet{montgomery1982introduction} even suggests
+that residual plots are more informative in most practical situations
+than the corresponding conventional hypothesis tests.
 
@@
 effect. Then, the estimated power of a visual test evaluated by a single
-subject is the predicted value obtained from the mix-effect model.
-However, this mix-effect model does not work with scenario where
+subject is the predicted value obtained from the mixed effects model.
+However, this mixed effects model does not work with scenario where
 subjects are asked to select one or more most different plots. In this

### 2023-04-10 paper.Rmd
add link to the repo

--- 2023-04-09 paper.Rmd
@@
 In terms of lineups containing heteroskedasticity patterns, `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by conventional tests, while `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by visual tests.
-The visual test rejects `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the lineups that the conventional visual test reject.
+When the conventional test rejects a lineup, there is a good chance (`r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%) the visual test will also reject it.
 
@@
 The article was created with R packages `rticles` [@rticles], `knitr` [@knitr] and `rmarkdown` [@rmarkdown].
-The project's Github repository (link here) contains all materials required to reproduce this article.
+The project's Github repository (https://github.com/TengMCing/lineup_residual_diagnostics) contains all materials required to reproduce this article.
 
@@
 
-The supplementary material is available at (link here).
+The supplementary material is available at https://github.com/TengMCing/lineup_residual_diagnostics/blob/master/appendix.pdf.
 It includes more details about the experimental setup, the derivation of the effect size, the effect of data collection period, and the estimate of $\alpha$.

### 2023-04-10 paper.tex
add link to the repo

--- 2023-04-09 paper.tex
@@
 rejected by conventional tests, while 56\% are rejected by visual tests.
-The visual test rejects 73\% of the lineups that the conventional visual
-test reject.
+When the conventional test rejects a lineup, there is a good chance
+(73\%) the visual test will also reject it.
 
@@
 is as easy to be detected as the ``S'' shape. This suggests a benefit of
-the visual test: knowing the shape ahead of time is \emph{not} needed
-for its application.
+the visual test, the knowing the shape ahead of time is not needed for
+its application.
 
@@
 \citep{rticles}, \texttt{knitr} \citep{knitr} and \texttt{rmarkdown}
-\citep{rmarkdown}. The project's Github repository (link here) contains
-all materials required to reproduce this article.
+\citep{rmarkdown}. The project's Github repository
+(\url{https://github.com/TengMCing/lineup_residual_diagnostics})
+contains all materials required to reproduce this article.
 
@@
 
-The supplementary material is available at (link here). It includes more
-details about the experimental setup, the derivation of the effect size,
-the effect of data collection period, and the estimate of \(\alpha\).
+The supplementary material is available at
+\url{https://github.com/TengMCing/lineup_residual_diagnostics/blob/master/appendix.pdf}.
+It includes more details about the experimental setup, the derivation of
+the effect size, the effect of data collection period, and the estimate
+of \(\alpha\).
 

### 2023-04-10 appendix.Rmd
a table in appendix for number of lineups and evaluations

@@
 
+Throughout the study, we have collected `r vi_survey %>% filter(!attention_check, !null_lineup) %>% nrow()` evaluations on `r vi_survey %>% filter(!attention_check, !null_lineup) %>% count(unique_lineup_id) %>% nrow()` non-null lineups.
+Table \ref{tab:count-lineup} gives further details about the number of evaluations and lineups over departure types and data collection periods.
+
+```{r}
+matrix(c(c(576, 0, 144, 0, 540, 135, 1116),
+         c(2880, 0, 864, 0, 2700, 810, 7254)),
+       ncol = 7,
+       byrow = TRUE) %>%
+  data.frame() %>%
+  add_column(name = c("Lineups", "Evaluations"), .before = 1) %>%
+  kableExtra::kable(format = "latex", 
+                    booktabs = TRUE,
+                    label = "count-lineup",
+                    caption = "Count of lineups and evaluations over departure types and data collection periods.",
+                    col.names = c("Number", "I", "II", "III", "I", "II", "II", "Total")) %>%
+  kableExtra::add_header_above(c(" " = 1, "Non-linearity" = 3, "Heteroskedasticity" = 3))
+```
+
 Along with the responses to lineups, we have collected a series of demographic information including age, pronoun, education background and previous experience in studies involved data visualization.

### 2023-04-10 appendix.tex
a table in appendix for number of lineups and evaluations

@@
 
+Throughout the study, we have collected 7254 evaluations on 1116
+non-null lineups. Table \ref{tab:count-lineup} gives further details
+about the number of evaluations and lineups over departure types and
+data collection periods.
+
+\begin{table}
+
+\caption{\label{tab:count-lineup}Count of lineups and evaluations over departure types and data collection periods.}
+\centering
+\begin{tabular}[t]{lrrrrrrr}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Non-linearity} & \multicolumn{3}{c}{Heteroskedasticity} \\
+\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
+Number & I & II & III & I & II & II & Total\\
+\midrule
+Lineups & 576 & 0 & 144 & 0 & 540 & 135 & 1116\\
+Evaluations & 2880 & 0 & 864 & 0 & 2700 & 810 & 7254\\
+\bottomrule
+\end{tabular}
+\end{table}
+
 Along with the responses to lineups, we have collected a series of

### 2023-04-10 appendix.Rmd
add participant info

--- 2023-04-04 appendix.Rmd
@@
 
+Throughout the study, we have collected `r vi_survey %>% filter(!attention_check, !null_lineup) %>% nrow()` evaluations on `r vi_survey %>% filter(!attention_check, !null_lineup) %>% count(unique_lineup_id) %>% nrow()` non-null lineups.
+Table \ref{tab:count-lineup} gives further details about the number of evaluations, lineups and participants over departure types and data collection periods.
+
+```{r}
+matrix(c(c(576, 0, 144, 0, 540, 135, 1116),
+         c(2880, 0, 864, 0, 2700, 810, 7254),
+         c(160, 0, 123, 0, 160, 123, 443)),
+       ncol = 7,
+       byrow = TRUE) %>%
+  data.frame() %>%
+  add_column(name = c("Lineups", "Evaluations", "Participants"), .before = 1) %>%
+  kableExtra::kable(format = "latex", 
+                    booktabs = TRUE,
+                    label = "count-lineup",
+                    caption = "Count of lineups, evaluations and participants over departure types and data collection periods.",
+                    col.names = c("Number", "I", "II", "III", "I", "II", "II", "Total")) %>%
+  kableExtra::add_header_above(c(" " = 1, "Non-linearity" = 3, "Heteroskedasticity" = 3))
+```
+
 Along with the responses to lineups, we have collected a series of demographic information including age, pronoun, education background and previous experience in studies involved data visualization.

### 2023-04-10 appendix.tex
add participant info

--- 2023-04-04 appendix.tex
@@
 
+Throughout the study, we have collected 7254 evaluations on 1116
+non-null lineups. Table \ref{tab:count-lineup} gives further details
+about the number of evaluations, lineups and participants over departure
+types and data collection periods.
+
+\begin{table}
+
+\caption{\label{tab:count-lineup}Count of lineups, evaluations and participants over departure types and data collection periods.}
+\centering
+\begin{tabular}[t]{lrrrrrrr}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Non-linearity} & \multicolumn{3}{c}{Heteroskedasticity} \\
+\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
+Number & I & II & III & I & II & II & Total\\
+\midrule
+Lineups & 576 & 0 & 144 & 0 & 540 & 135 & 1116\\
+Evaluations & 2880 & 0 & 864 & 0 & 2700 & 810 & 7254\\
+Participants & 160 & 0 & 123 & 0 & 160 & 123 & 443\\
+\bottomrule
+\end{tabular}
+\end{table}
+
 Along with the responses to lineups, we have collected a series of

### 2023-04-10 appendix.Rmd
checking table C1

--- 2023-04-04 appendix.Rmd
@@
 
+Throughout the study, we have collected `r vi_survey %>% filter(!attention_check, !null_lineup) %>% nrow()` evaluations on `r vi_survey %>% filter(!attention_check, !null_lineup) %>% count(unique_lineup_id) %>% nrow()` non-null lineups.
+Table \ref{tab:count-lineup} gives further details about the number of evaluations, lineups and participants over pattern types and data collection periods.
+
+```{r}
+# Calculate these numbers
+matrix(c(c(576, 0, 144, 0, 540, 135, 1116),
+         c(2880, 0, 864, 0, 2700, 810, 7254),
+         c(160, 0, 123, 0, 160, 123, 443)),
+       ncol = 7,
+       byrow = TRUE) %>%
+  data.frame() %>%
+  add_column(name = c("Lineups", "Evaluations", "Participants"), .before = 1) %>%
+  kableExtra::kable(format = "latex", 
+                    booktabs = TRUE,
+                    label = "count-lineup",
+                    caption = "Count of lineups, evaluations and participants over departure types and data collection periods.",
+                    col.names = c("Number", "I", "II", "III", "I", "II", "II", "Total")) %>%
+  kableExtra::add_header_above(c(" " = 1, "Non-linearity" = 3, "Heteroskedasticity" = 3))
+```
+
 Along with the responses to lineups, we have collected a series of demographic information including age, pronoun, education background and previous experience in studies involved data visualization.

### 2023-04-10 appendix.tex
checking table C1

@@
 non-null lineups. Table \ref{tab:count-lineup} gives further details
-about the number of evaluations, lineups and participants over departure
+about the number of evaluations, lineups and participants over pattern
 types and data collection periods.

### 2023-04-11 paper.Rmd
correct the caption in 5.1

@@
 These plots serve as a visual aid to help readers understand how different effect size values translate to changes in the diagnostic plots.
-The horizontal lines of dots at 0 and 1 represent the detects or non-detects made by human observers for each lineup.
+The horizontal lines of dots at 0 and 1 represent the non-reject or reject made by visual tests for each lineup.
 
@@
 ```{r polypower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for non-linear patterns (uniform fitted values only).
-The power curves are estimated using logistic regression, and the horizontal lines of dots represent non-detect and detect results from human observers for each lineup.
+The power curves are estimated using logistic regression, and the horizontal lines of dots represent non-reject and reject results from visual tests for each lineup.
 The visual test has multiple power curves estimated from bootstrap samples.
@@
 ```{r heterpower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only).
-Main plot shows the power curves, with dots indicating non-detect and detect in human evaluations of lineups.
+Main plot shows the power curves, with dots indicating non-reject and reject in visual testing of lineups.
 The multiple lines for the visual test arise from estimating the power on many bootstrap samples.

### 2023-04-11 paper.Rmd
correct caption

@@
 The visual test has multiple power curves estimated from bootstrap samples.
-The row of scatterplots plots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.", dev = 'png', dpi = 450}
+The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.", dev = 'png', dpi = 450}
 
@@
 The multiple lines for the visual test arise from estimating the power on many bootstrap samples.
-The row of scatterplots plots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.", dev = 'png', dpi = 450}
+The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.", dev = 'png', dpi = 450}
 

### 2023-04-11 appendix.Rmd
replace with calculated numbers

@@
 # Calculate these numbers
-matrix(c(c(576, 0, 144, 0, 540, 135, 1116),
-         c(2880, 0, 864, 0, 2700, 810, 7254),
-         c(160, 0, 123, 0, 160, 123, 443)),
+a1 <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(type == "polynomial") %>%
+  filter(exp == 3) %>%
+  count(unique_lineup_id) %>%
+  nrow()
+
+a2 <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(type == "polynomial") %>%
+  filter(exp == 4) %>%
+  count(unique_lineup_id) %>%
+  nrow()
+
+a3 <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(type == "polynomial") %>%
+  filter(exp == 5) %>%
+  count(unique_lineup_id) %>%
+  nrow()
+
+a4 <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(type != "polynomial") %>%
+  filter(exp == 3) %>%
+  count(unique_lineup_id) %>%
+  nrow()
+
+a5 <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(type != "polynomial") %>%
+  filter(exp == 4) %>%
+  count(unique_lineup_id) %>%
+  nrow()
+
+a6 <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(type != "polynomial") %>%
+  filter(exp == 5) %>%
+  count(unique_lineup_id) %>%
+  nrow()
+
+a7 <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  count(unique_lineup_id) %>%
+  nrow()
+
+b1 <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(type == "polynomial") %>%
+  filter(exp == 3) %>%
+  nrow()
+
+b2 <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(type == "polynomial") %>%
+  filter(exp == 4) %>%
+  nrow()
+
+b3 <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(type == "polynomial") %>%
+  filter(exp == 5) %>%
+  nrow()
+
+b4 <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(type != "polynomial") %>%
+  filter(exp == 3) %>%
+  nrow()
+
+b5 <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(type != "polynomial") %>%
+  filter(exp == 4) %>%
+  nrow()
+
+b6 <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(type != "polynomial") %>%
+  filter(exp == 5) %>%
+  nrow()
+
+b7 <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  nrow()
+
+c1 <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(type == "polynomial") %>%
+  filter(exp == 3) %>%
+  count(set) %>%
+  nrow()
+
+c2 <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(type == "polynomial") %>%
+  filter(exp == 4) %>%
+  count(set) %>%
+  nrow()
+
+c3 <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(type == "polynomial") %>%
+  filter(exp == 5) %>%
+  count(set) %>%
+  nrow()
+
+c4 <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(type != "polynomial") %>%
+  filter(exp == 3) %>%
+  count(set) %>%
+  nrow()
+
+c5 <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(type != "polynomial") %>%
+  filter(exp == 4) %>%
+  count(set) %>%
+  nrow()
+
+c6 <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(type != "polynomial") %>%
+  filter(exp == 5) %>%
+  count(set) %>%
+  nrow()
+
+c7 <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  count(exp, set) %>%
+  nrow()
+
+matrix(c(c(a1, a2, a3, a4, a5, a6, a7),
+         c(b1, b2, b3, b4, b5, b6, b7),
+         c(c1, c2, c3, c4, c5, c6, c7)),
        ncol = 7,

### 2023-04-14 paper.Rmd
fig 13, 14 caption

@@
 ```{r p-value-comparison, fig.width = 8, fig.height = 4, fig.cap = "Rejection rate ($p$-value $\\leq0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
-The visual test rejects less frequently than the conventional test.
-We would generally expect the visual test to only reject when the conventional test does.
-Surprisingly, one lineup containing a heteroskedasticity pattern does not follow this rule.
-"}
+The visual test rejects less frequently than the conventional test, and (almost) only rejects when the conventional test does.
+Surprisingly, one lineup in the heteroskedasticity group is rejected by the visual test but NOT the conventional test."}
 
@@
 ```{r heter-example, fig.pos="t!", fig.height = 7, fig.width = 7, fig.cap = 'The single heteroskedasticity lineup that is rejected by the visual test but not by the BP test.
-The data plot at panel 17 contains a "butterfly" shape.
-It has effect size $ = 3.76$, somewhat surprising that it is not detected by the BP test.'}
+The data plot (position 17) contains a ``butterfly" shape.
+It has a log effect size of $ 3.76$, and visibly displays heteroskedasticity, making it somewhat surprising that it is not detected by the BP test.'}
 # plot heter-331

### 2023-04-14 paper.tex
fig 13, 14 caption

@@
 diagnostic plots. The horizontal lines of dots at 0 and 1 represent the
-detects or non-detects made by human observers for each lineup.
+non-reject or reject made by visual tests for each lineup.
 
@@
 \caption{Comparison of power between different tests for non-linear patterns (uniform fitted values only).
-The power curves are estimated using logistic regression, and the horizontal lines of dots represent non-detect and detect results from human observers for each lineup.
+The power curves are estimated using logistic regression, and the horizontal lines of dots represent non-reject and reject results from visual tests for each lineup.
 The visual test has multiple power curves estimated from bootstrap samples.
-The row of scatterplots plots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}\label{fig:polypower}
+The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}\label{fig:polypower}
 \end{figure}
@@
 \caption{Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only).
-Main plot shows the power curves, with dots indicating non-detect and detect in human evaluations of lineups.
+Main plot shows the power curves, with dots indicating non-reject and reject in visual testing of lineups.
 The multiple lines for the visual test arise from estimating the power on many bootstrap samples.
-The row of scatterplots plots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}\label{fig:heterpower}
+The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}\label{fig:heterpower}
 \end{figure}
@@
 \caption{Rejection rate ($p$-value $\leq0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
-The visual test rejects less frequently than the conventional test.
-We would generally expect the visual test to only reject when the conventional test does.
-Surprisingly, one lineup containing a heteroskedasticity pattern does not follow this rule.
-}\label{fig:p-value-comparison}
+The visual test rejects less frequently than the conventional test, and (almost) only rejects when the conventional test does.
+Surprisingly, one lineup in the heteroskedasticity group is rejected by the visual test but NOT the conventional test.}\label{fig:p-value-comparison}
 \end{figure}
@@
 \caption{The single heteroskedasticity lineup that is rejected by the visual test but not by the BP test.
-The data plot at panel 17 contains a "butterfly" shape.
-It has effect size $ = 3.76$, somewhat surprising that it is not detected by the BP test.}\label{fig:heter-example}
+The data plot (position 17) contains a ``butterfly" shape.
+It has a log effect size of $ 3.76$, and visibly displays heteroskedasticity, making it somewhat surprising that it is not detected by the BP test.}\label{fig:heter-example}
 \end{figure}

### 2023-04-14 paper.Rmd
tweaked the intro and placement of fig 1

@@
 Diagnostics are the key to determining whether there is anything importantly wrong with a model.
-In linear regression analysis, residuals from the model fit are commonly used.
+In linear regression analysis, studying the residuals from a model fit is a common diagnostic activity.
 Residuals summarise what is not captured by the model, and thus provide the capacity to identify what might be wrong.
@@
 We can assess residuals in multiple ways.
-Residuals may be plotted as a histogram or quantile-quantile plot to examine the distribution.
-Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, we consider it to be well behaved.
-However, if the distribution is skewed, bimodal, multimodal, or contains outliers, there is a cause for concern.
+To examine the their univariate distribution, residuals may be plotted as a histogram or normal probability plot.
+Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, we would consider it to be well-behaved.
+However, if the distribution is skewed, bimodal, multimodal, or contains outliers, there would be cause for concern.
 One could also inspect the distribution by conducting a goodness-of-fit test, such as the Shapiro-Wilk (SW) Normality test [@shapiro1965analysis].
 
-More typically, a scatterplot of residuals plotted against the fitted values and each of the explanatory variables, is made to scrutinize their relationships.
-If there are any visually discoverable patterns, the model is potentially inadequate or incorrectly specifed.
-In general, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity.
-However, correctly judging a residual plot where no pattern exists can be a difficult task for humans.
-It is especially common, particularly among students, to misinterpret patterns that are random noise and random deviation from a model [@loy2021bringing].
-There are also numerous conventional hypothesis tests available to test for non-linear dependence [@ramsey_tests_1969], and heteroskedasticity  [@breusch_simple_1979].
-
-An abundance of literature describe appropriate diagnostic methods for linear regression, e.g.
+In addition, scatterplots of residuals plotted against the fitted values and each of the explanatory variables, are made to scrutinize their relationships.
+If there are any visually discoverable patterns, the model is potentially inadequate or incorrectly specified.
+In general, one looks for noticeable departures from the model such as non-linear dependency or heteroskedasticity.
+A non-linear dependency would suggest that the model needs to have some additional non-linear terms.
+Heteroskedasticity suggests that the error is dependent on the predictors, and hence violates the independence assumption.
+However, correctly judging whether NO pattern exists in a residual plot is a difficult task for humans.
+We humans will almost always will see a pattern, so the question that really needs answering is whether any pattern perceived is consistent with randomness, purely sampling variability or noise.
+It is especially difficult to teach this to new analysts and students [@loy2021bringing].
+To answer this, there have been numerous conventional hypothesis tests made available to test for non-linear dependence [e.g.
+@ramsey_tests_1969], and heteroskedasticity  [e.g.
+@breusch_simple_1979].
+
+Linear regression is a well-established procedure, and there is considerable literature describing diagnostic procedures, e.g.
 @draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals.
-All these writings consider plotting residuals to be an essential tool for diagnosing regression modelling problems.
-In addition, @draper1998applied and @belsley_regression_1980 believe that residual plots are usually revealing when the assumptions are violated.
-@cook_applied_1999 thinks formal tests and graphical procedures are complementary and both have a place in residual analysis, but they focus on graphical methods rather than on formal testing, as they are easier to use.
+An interesting details, is that despite the abundance of conventional tests, ALL of these writings advise that plotting residuals is an essential tool for diagnosing regression model problems.
+@draper1998applied and @belsley_regression_1980 explain that residual plots are usually revealing when the assumptions are violated.
+@cook_applied_1999 thinks formal tests and graphical procedures are complementary and both have a place in residual analysis, but they focus on graphical methods rather than on formal testing.
 @montgomery1982introduction even suggests that residual plots are more informative in most practical situations than the corresponding conventional hypothesis tests.
 
-A common guidance by experts is that optimal method for diagnosing model fits is by plotting the data.
-The persistence of this advice to check the plots is curious, and investigating why this might be common advice is the subject of this paper.
+The common wisdom of experts is that the optimal method for diagnosing model fits is by plotting the data.
+The persistence of this advice to check the plots is *curious*, and investigating why this is, is the subject of this paper.
+
 The paper is structured as follows.
@@
 
-```{r residual-plot-common-departures, fig.width = 10, fig.height = 2.5, fig.cap = "Example residual vs fitted value plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
+Graphical summaries in which residuals are plotted against fitted values or other functions of the predictors that are approximately orthogonal to the residuals are referred to as standard residual plots in @cook1982residuals.
+Figure \ref{fig:residual-plot-common-departures}A shows an ideal residual plot where the residuals are evenly distributed at both sides of the horizontal zero line, with no noticeable patterns.
+
+There are various types of departures from an ideal residual plot. Non-linearity, heteroskedasticity and non-normality are perhaps the three mostly checked departures.
+
+```{r residual-plot-common-departures, fig.pos="t!", fig.width = 10, fig.height = 2.5, fig.cap = "Example residual vs fitted value plots (red line indicates 0): (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
 The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted value plot."}
@@
 ```
-
-Graphical summaries in which residuals are plotted against fitted values or other functions of the predictors that are approximately orthogonal to the residuals are referred to as standard residual plots in @cook1982residuals.
-Figure \ref{fig:residual-plot-common-departures}A shows an ideal residual plot where the residuals are evenly distributed at both sides of the horizontal zero line, with no noticeable patterns.
-
-There are various types of departures from an ideal residual plot. Non-linearity, heteroskedasticity and non-normality are perhaps the three mostly checked departures.
 

### 2023-04-14 paper.tex
tweaked the intro and placement of fig 1

@@
 Diagnostics are the key to determining whether there is anything
-importantly wrong with a model. In linear regression analysis, residuals
-from the model fit are commonly used. Residuals summarise what is not
-captured by the model, and thus provide the capacity to identify what
-might be wrong.
-
-We can assess residuals in multiple ways. Residuals may be plotted as a
-histogram or quantile-quantile plot to examine the distribution. Using
-the classical normal linear regression model as an example, if the
-distribution is symmetric and unimodal, we consider it to be well
-behaved. However, if the distribution is skewed, bimodal, multimodal, or
-contains outliers, there is a cause for concern. One could also inspect
-the distribution by conducting a goodness-of-fit test, such as the
-Shapiro-Wilk (SW) Normality test \citep{shapiro1965analysis}.
-
-More typically, a scatterplot of residuals plotted against the fitted
-values and each of the explanatory variables, is made to scrutinize
-their relationships. If there are any visually discoverable patterns,
-the model is potentially inadequate or incorrectly specifed. In general,
-one looks for noticeable departures from the model like non-linear
-dependency or heteroskedasticity. However, correctly judging a residual
-plot where no pattern exists can be a difficult task for humans. It is
-especially common, particularly among students, to misinterpret patterns
-that are random noise and random deviation from a model
-\citep{loy2021bringing}. There are also numerous conventional hypothesis
-tests available to test for non-linear dependence
-\citep{ramsey_tests_1969}, and heteroskedasticity
-\citep{breusch_simple_1979}.
-
-An abundance of literature describe appropriate diagnostic methods for
-linear regression, e.g. \citet{draper1998applied},
-\citet{montgomery1982introduction}, \citet{belsley_regression_1980},
-\citet{cook_applied_1999} and \citet{cook1982residuals}. All these
-writings consider plotting residuals to be an essential tool for
-diagnosing regression modelling problems. In addition,
-\citet{draper1998applied} and \citet{belsley_regression_1980} believe
-that residual plots are usually revealing when the assumptions are
-violated. \citet{cook_applied_1999} thinks formal tests and graphical
-procedures are complementary and both have a place in residual analysis,
-but they focus on graphical methods rather than on formal testing, as
-they are easier to use. \citet{montgomery1982introduction} even suggests
-that residual plots are more informative in most practical situations
-than the corresponding conventional hypothesis tests.
-
-A common guidance by experts is that optimal method for diagnosing model
-fits is by plotting the data. The persistence of this advice to check
-the plots is curious, and investigating why this might be common advice
-is the subject of this paper. The paper is structured as follows. The
-next section describes the background on the types of departures that
-one expects to detect, and outlines a formal statistical process for
-reading residual plots, called visual inference. Section
-\ref{experimental-design} details the experimental design to compare the
-decisions made by formal hypothesis testing, and how humans would read
-diagnostic plots. The results are reported in Section \ref{results}. We
-conclude with a discussion of future work, in particular, how the
-responsibility for residual plot reading might be relieved.
+importantly wrong with a model. In linear regression analysis, studying
+the residuals from a model fit is a common diagnostic activity.
+Residuals summarise what is not captured by the model, and thus provide
+the capacity to identify what might be wrong.
+
+We can assess residuals in multiple ways. To examine the their
+univariate distribution, residuals may be plotted as a histogram or
+normal probability plot. Using the classical normal linear regression
+model as an example, if the distribution is symmetric and unimodal, we
+would consider it to be well-behaved. However, if the distribution is
+skewed, bimodal, multimodal, or contains outliers, there would be cause
+for concern. One could also inspect the distribution by conducting a
+goodness-of-fit test, such as the Shapiro-Wilk (SW) Normality test
+\citep{shapiro1965analysis}.
+
+In addition, scatterplots of residuals plotted against the fitted values
+and each of the explanatory variables, are made to scrutinize their
+relationships. If there are any visually discoverable patterns, the
+model is potentially inadequate or incorrectly specified. In general,
+one looks for noticeable departures from the model such as non-linear
+dependency or heteroskedasticity. A non-linear dependency would suggest
+that the model needs to have some additional non-linear terms.
+Heteroskedasticity suggests that the error is dependent on the
+predictors, and hence violates the independence assumption. However,
+correctly judging whether NO pattern exists in a residual plot is a
+difficult task for humans. We humans will almost always will see a
+pattern, so the question that really needs answering is whether any
+pattern perceived is consistent with randomness, purely sampling
+variability or noise. It is especially difficult to teach this to new
+analysts and students \citep{loy2021bringing}. To answer this, there
+have been numerous conventional hypothesis tests made available to test
+for non-linear dependence \citep[e.g.][]{ramsey_tests_1969}, and
+heteroskedasticity \citep[e.g.][]{breusch_simple_1979}.
+
+Linear regression is a well-established procedure, and there is
+considerable literature describing diagnostic procedures, e.g.
+\citet{draper1998applied}, \citet{montgomery1982introduction},
+\citet{belsley_regression_1980}, \citet{cook_applied_1999} and
+\citet{cook1982residuals}. An interesting details, is that despite the
+abundance of conventional tests, ALL of these writings advise that
+plotting residuals is an essential tool for diagnosing regression model
+problems. \citet{draper1998applied} and \citet{belsley_regression_1980}
+explain that residual plots are usually revealing when the assumptions
+are violated. \citet{cook_applied_1999} thinks formal tests and
+graphical procedures are complementary and both have a place in residual
+analysis, but they focus on graphical methods rather than on formal
+testing. \citet{montgomery1982introduction} even suggests that residual
+plots are more informative in most practical situations than the
+corresponding conventional hypothesis tests.
+
+The common wisdom of experts is that the optimal method for diagnosing
+model fits is by plotting the data. The persistence of this advice to
+check the plots is \emph{curious}, and investigating why this is, is the
+subject of this paper.
+
+The paper is structured as follows. The next section describes the
+background on the types of departures that one expects to detect, and
+outlines a formal statistical process for reading residual plots, called
+visual inference. Section \ref{experimental-design} details the
+experimental design to compare the decisions made by formal hypothesis
+testing, and how humans would read diagnostic plots. The results are
+reported in Section \ref{results}. We conclude with a discussion of
+future work, in particular, how the responsibility for residual plot
+reading might be relieved.
 
@@
 plots}\label{departures-from-good-residual-plots}}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/residual-plot-common-departures-1}
-
-}
-
-\caption{Example residual vs fitted value plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
-The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted value plot.}\label{fig:residual-plot-common-departures}
-\end{figure}
 
@@
 three mostly checked departures.
+
+\begin{figure}[t!]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/residual-plot-common-departures-1}
+
+}
+
+\caption{Example residual vs fitted value plots (red line indicates 0): (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
+The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted value plot.}\label{fig:residual-plot-common-departures}
+\end{figure}
 

### 2023-04-14 paper.Rmd
section 2.1

@@
 The results are reported in Section \ref{results}.
-We conclude with a discussion of future work, in particular, how the responsibility for residual plot reading might be relieved.
+We conclude with a discussion of the presented work, and ideas for future directions.
 
@@
 
-Graphical summaries in which residuals are plotted against fitted values or other functions of the predictors that are approximately orthogonal to the residuals are referred to as standard residual plots in @cook1982residuals.
-Figure \ref{fig:residual-plot-common-departures}A shows an ideal residual plot where the residuals are evenly distributed at both sides of the horizontal zero line, with no noticeable patterns.
-
-There are various types of departures from an ideal residual plot. Non-linearity, heteroskedasticity and non-normality are perhaps the three mostly checked departures.
+Graphical summaries where residuals are plotted against fitted values, or other functions of the predictors (expected to be approximately orthogonal to the residuals) are referred to as standard residual plots in @cook1982residuals.
+Figure \ref{fig:residual-plot-common-departures}A shows an example of an ideal residual plot where points are symmetrically distributed around the horizontal zero line (red), with no discernible patterns.
+There can be various types of departures from this ideal pattern.
+Non-linearity, heteroskedasticity and non-normality are three commonly checked departures.
 
@@
 
-Non-linearity is a type of model misspecification caused by failing to include higher order terms of the predictors in the regression equation.
-Any non-linear functional form of residuals on fitted values in the residual plot could be indicative of non-linearity.
+Model misspecification occurs if functions of predictors that needed to accurately describe the relationship between response and predictors were omitted.
+Any non-linear pattern visible in the residual plot could be indicative of this problem.
 An example residual plot containing visual pattern of non-linearity is shown in Figure \ref{fig:residual-plot-common-departures}B.
-One can clearly observe the "S-shape" from the residual plot as the cubic term is not captured by the misspecified model.
-
-Heteroskedasticity refers to the presence of nonconstant error variance in a regression model.
-It is mostly due to the strict but false assumptions on the variance-covariance matrix of the error term.
-The usual pattern of heteroskedasticity on a residual plot is the inconsistent spread of the residuals across the horizontal axis.
-Visually, it sometimes results in the so-called "butterfly" shape as shown in Figure \ref{fig:residual-plot-common-departures}C, or the "left-triangle" and "right-triangle" shape where the smallest variance occurs at one side of the horizontal axis.
-
-Compared to non-linearity and heteroskedasticity, non-normality is usually harder to detect from a residual plot since a scatter plot do not readily reveal the marginal distribution.
-A favourable graphical summary for this task is the quantile-quantile plot.
-As we mainly discuss residual plots, non-normality will not be the focus of this paper.
-For a consistent comparison, the residual plot of this departure is still presented in Figure \ref{fig:residual-plot-common-departures}D.
-When the number residuals below and above the horizontal axis are uneven across the local regions along the $x$-axis, we expect that the normality assumption is violated.
-For example, given a skewed error distribution, there will be fewer data points and more outliers on one side of the horizontal axis as shown in Figure \ref{fig:residual-plot-common-departures}D.
+One can clearly observe the "S-shape" from the residual plot, which corresponds to cubic term that should have been included in the model.
+
+Heteroskedasticity refers to the presence of non-constant error variance in a regression model.
+It indicates that the distribution of residuals depends on the predictors, violating the independence assumption.
+This can be seen in a residual plot as an inconsistent spread of the residuals relative to the fitted values or predictors.
+An example is the "butterfly" shape  shown in Figure \ref{fig:residual-plot-common-departures}C, or a "left-triangle" and "right-triangle" shape where the smallest variance occurs at one side of the horizontal axis.
+
+Figure \ref{fig:residual-plot-common-departures}D) shows a scatterplot where the residuals have a skewed distribution, as seen by the uneven vertical spread.
+Unlike non-linearity and heteroskedasticity, non-normality is usually detected with a different type of residual plot, a histogram or normal probability plot.
+Because we focus on scatterplots, non-normality is not one of the departures examined in this paper.
 

### 2023-04-14 paper.tex
section 2.1

@@
 testing, and how humans would read diagnostic plots. The results are
-reported in Section \ref{results}. We conclude with a discussion of
-future work, in particular, how the responsibility for residual plot
-reading might be relieved.
+reported in Section \ref{results}. We conclude with a discussion of the
+presented work, and ideas for future directions.
 
@@
 
-Graphical summaries in which residuals are plotted against fitted values
-or other functions of the predictors that are approximately orthogonal
-to the residuals are referred to as standard residual plots in
-\citet{cook1982residuals}. Figure
-\ref{fig:residual-plot-common-departures}A shows an ideal residual plot
-where the residuals are evenly distributed at both sides of the
-horizontal zero line, with no noticeable patterns.
-
-There are various types of departures from an ideal residual plot.
-Non-linearity, heteroskedasticity and non-normality are perhaps the
-three mostly checked departures.
+Graphical summaries where residuals are plotted against fitted values,
+or other functions of the predictors (expected to be approximately
+orthogonal to the residuals) are referred to as standard residual plots
+in \citet{cook1982residuals}. Figure
+\ref{fig:residual-plot-common-departures}A shows an example of an ideal
+residual plot where points are symmetrically distributed around the
+horizontal zero line (red), with no discernible patterns. There can be
+various types of departures from this ideal pattern. Non-linearity,
+heteroskedasticity and non-normality are three commonly checked
+departures.
 
@@
 
-Non-linearity is a type of model misspecification caused by failing to
-include higher order terms of the predictors in the regression equation.
-Any non-linear functional form of residuals on fitted values in the
-residual plot could be indicative of non-linearity. An example residual
-plot containing visual pattern of non-linearity is shown in Figure
+Model misspecification occurs if functions of predictors that needed to
+accurately describe the relationship between response and predictors
+were omitted. Any non-linear pattern visible in the residual plot could
+be indicative of this problem. An example residual plot containing
+visual pattern of non-linearity is shown in Figure
 \ref{fig:residual-plot-common-departures}B. One can clearly observe the
-``S-shape'' from the residual plot as the cubic term is not captured by
-the misspecified model.
-
-Heteroskedasticity refers to the presence of nonconstant error variance
-in a regression model. It is mostly due to the strict but false
-assumptions on the variance-covariance matrix of the error term. The
-usual pattern of heteroskedasticity on a residual plot is the
-inconsistent spread of the residuals across the horizontal axis.
-Visually, it sometimes results in the so-called ``butterfly'' shape as
-shown in Figure \ref{fig:residual-plot-common-departures}C, or the
-``left-triangle'' and ``right-triangle'' shape where the smallest
-variance occurs at one side of the horizontal axis.
-
-Compared to non-linearity and heteroskedasticity, non-normality is
-usually harder to detect from a residual plot since a scatter plot do
-not readily reveal the marginal distribution. A favourable graphical
-summary for this task is the quantile-quantile plot. As we mainly
-discuss residual plots, non-normality will not be the focus of this
-paper. For a consistent comparison, the residual plot of this departure
-is still presented in Figure \ref{fig:residual-plot-common-departures}D.
-When the number residuals below and above the horizontal axis are uneven
-across the local regions along the \(x\)-axis, we expect that the
-normality assumption is violated. For example, given a skewed error
-distribution, there will be fewer data points and more outliers on one
-side of the horizontal axis as shown in Figure
-\ref{fig:residual-plot-common-departures}D.
+``S-shape'' from the residual plot, which corresponds to cubic term that
+should have been included in the model.
+
+Heteroskedasticity refers to the presence of non-constant error variance
+in a regression model. It indicates that the distribution of residuals
+depends on the predictors, violating the independence assumption. This
+can be seen in a residual plot as an inconsistent spread of the
+residuals relative to the fitted values or predictors. An example is the
+``butterfly'' shape shown in Figure
+\ref{fig:residual-plot-common-departures}C, or a ``left-triangle'' and
+``right-triangle'' shape where the smallest variance occurs at one side
+of the horizontal axis.
+
+Figure \ref{fig:residual-plot-common-departures}D) shows a scatterplot
+where the residuals have a skewed distribution, as seen by the uneven
+vertical spread. Unlike non-linearity and heteroskedasticity,
+non-normality is usually detected with a different type of residual
+plot, a histogram or normal probability plot. Because we focus on
+scatterplots, non-normality is not one of the departures examined in
+this paper.
 

### 2023-04-14 paper.Rmd
section 2.2

@@
 Unlike non-linearity and heteroskedasticity, non-normality is usually detected with a different type of residual plot, a histogram or normal probability plot.
-Because we focus on scatterplots, non-normality is not one of the departures examined in this paper.
+Because we focus on scatterplots, non-normality is not one of the departures examined in depth in this paper.
 
@@
 Interestingly, the BP and SW tests also reject the residual plots exhibiting structure that they weren't designed for.
-@cook1982residuals discusses that most residual-based tests for particular types of departure from model assumptions are also sensitive to other types of departures, that is, correctly rejected but for the wrong reason, a phenomenon known as the "Type III error".
-Also, some types of departure can also have elements of other types of departure, for example, non-linearity could be viewed as heteroskedasticity.
+@cook1982residuals discusses that most residual-based tests for particular types of departure from model assumptions are also sensitive to other types of departures.
+This could be considered a Type III error, where the null hypothesis of good residuals is correctly rejected but for the wrong reason.
+Also, some types of departure can have elements of other types of departure, for example, non-linearity could be viewed as heteroskedasticity.
 Additionally, other problesms such as outliers can trigger the rejection [@cook_applied_1999].
 
-With large sample sizes, hypothesis tests are may reject a residual plot when there is only a slight departure (XXX is there a reference for this statement?).
-While such rejections may be statistically correct, their sensitivity may render the results impractical.
+With large sample sizes, hypothesis tests may reject the null hypothesis when there is only a small effect.
+(This is common knowledge but a good discussion can be found in @kirk1996.) While such rejections may be statistically correct, their sensitivity may render the results impractical.
+This is especially pertinent for residual diagnostics.
 Therefore, a key goal of residual plot diagnostics is to identify potential issues that could lead to incorrect conclusions or errors in subsequent analyses.

### 2023-04-14 paper.tex
section 2.2

@@
 scatterplots, non-normality is not one of the departures examined in
-this paper.
+depth in this paper.
 
@@
 residual-based tests for particular types of departure from model
-assumptions are also sensitive to other types of departures, that is,
-correctly rejected but for the wrong reason, a phenomenon known as the
-``Type III error''. Also, some types of departure can also have elements
-of other types of departure, for example, non-linearity could be viewed
-as heteroskedasticity. Additionally, other problesms such as outliers
-can trigger the rejection \citep{cook_applied_1999}.
-
-With large sample sizes, hypothesis tests are may reject a residual plot
-when there is only a slight departure (XXX is there a reference for this
-statement?). While such rejections may be statistically correct, their
-sensitivity may render the results impractical. Therefore, a key goal of
-residual plot diagnostics is to identify potential issues that could
-lead to incorrect conclusions or errors in subsequent analyses. However,
-minor defects in the model are unlikely to have a significant impact and
-may be best disregarded for practical purposes. In fact, the experiment
-discussed in this paper specifically addresses this.
+assumptions are also sensitive to other types of departures. This could
+be considered a Type III error, where the null hypothesis of good
+residuals is correctly rejected but for the wrong reason. Also, some
+types of departure can have elements of other types of departure, for
+example, non-linearity could be viewed as heteroskedasticity.
+Additionally, other problesms such as outliers can trigger the rejection
+\citep{cook_applied_1999}.
+
+With large sample sizes, hypothesis tests may reject the null hypothesis
+when there is only a small effect. (This is common knowledge but a good
+discussion can be found in \citet{kirk1996}.) While such rejections may
+be statistically correct, their sensitivity may render the results
+impractical. This is especially pertinent for residual diagnostics.
+Therefore, a key goal of residual plot diagnostics is to identify
+potential issues that could lead to incorrect conclusions or errors in
+subsequent analyses. However, minor defects in the model are unlikely to
+have a significant impact and may be best disregarded for practical
+purposes. In fact, the experiment discussed in this paper specifically
+addresses this.
 

### 2023-04-14 paper.Rmd
section 2.3 but needs more work

@@
 
-The examination of data plots to infer signal or patterns (or lack thereof) is fraught with variation in the human ability to interpret and decode the information embedded in graph [@cleveland_graphical_1984].
-Such examination thus feels there is a subjective nature to evaluate, say, diagnostic plots to infer appropriateness of statistical models, although arguably we would expect such evaluation should be done objectively and such human-to-human variation in "reading" data plots should be minimised.
+The examination of data plots to infer signal or patterns (or lack thereof) is fraught with variation in the human ability to interpret and decode the information embedded in a graph [@cleveland_graphical_1984].
+To rely on human examination thus feels as though there is a subjective nature to using diagnostic plots to infer appropriateness of statistical models.
+Ideally we would expect visual evaluation should be done objectively and human-to-human variation in reading data plots small, but it cannot be assured.
 
 In practice, over-interpretation of a single plot is common.
-For instance, in @roy_chowdhury_using_2015, some over-interpreted the separation between gene groups in a two-dimensional projection of a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups.
-To mitigate this over-interpretation, @buja_statistical_2009 proposed a line-up protocol to assess plots in a manner analogous to the null hypothesis significance testing (NHST) framework.
+For instance, in @roy_chowdhury_using_2015, described an published example where the authors over-interpreted separation between gene groups from a two-dimensional projection of a linear discriminant analysis.
+@roy_chowdhury_using_2015 showed that there were no differences in the expression levels between the gene groups.
+The lineup protocol proposed in @buja_statistical_2009 provides a solution to over-interpretation.
+@majumder_validation_2013 showed that its use is analogous to the null hypothesis significance testing (NHST) framework.
 More specifically, the protocol consists of $m$ randomly placed plots, where one plot is the data plot, and the remaining $m - 1$ plots, referred to as the _null plots_, have the identical graphical procedure as the data plot except the data is replaced with a data generation mechanism that is consistent with the null hypothesis, $H_0$.

### 2023-04-14 paper.tex
section 2.3 but needs more work

@@
 thereof) is fraught with variation in the human ability to interpret and
-decode the information embedded in graph
-\citep{cleveland_graphical_1984}. Such examination thus feels there is a
-subjective nature to evaluate, say, diagnostic plots to infer
-appropriateness of statistical models, although arguably we would expect
-such evaluation should be done objectively and such human-to-human
-variation in ``reading'' data plots should be minimised.
+decode the information embedded in a graph
+\citep{cleveland_graphical_1984}. To rely on human examination thus
+feels as though there is a subjective nature to using diagnostic plots
+to infer appropriateness of statistical models. Ideally we would expect
+visual evaluation should be done objectively and human-to-human
+variation in reading data plots small, but it cannot be assured.
 
 In practice, over-interpretation of a single plot is common. For
-instance, in \citet{roy_chowdhury_using_2015}, some over-interpreted the
-separation between gene groups in a two-dimensional projection of a
-linear discriminant analysis when in fact there are no differences in
-the expression levels between the gene groups. To mitigate this
-over-interpretation, \citet{buja_statistical_2009} proposed a line-up
-protocol to assess plots in a manner analogous to the null hypothesis
-significance testing (NHST) framework. More specifically, the protocol
-consists of \(m\) randomly placed plots, where one plot is the data
-plot, and the remaining \(m - 1\) plots, referred to as the \emph{null
-plots}, have the identical graphical procedure as the data plot except
-the data is replaced with a data generation mechanism that is consistent
-with the null hypothesis, \(H_0\). Then, an observer who have not seen
-the data plot will be asked to point out the most different plot from
-the lineup. Under \(H_0\), it is expected that the data plot would have
-no distinguishable difference from the null plots, and the probability
-that the observer correctly picks the data plot is \(1/m\). If one
-rejects \(H_0\) as the observer correctly picks the data plot, then the
-Type I error of this test is \(1/m\). This protocol requires apriori
+instance, in \citet{roy_chowdhury_using_2015}, described an published
+example where the authors over-interpreted separation between gene
+groups from a two-dimensional projection of a linear discriminant
+analysis. \citet{roy_chowdhury_using_2015} showed that there were no
+differences in the expression levels between the gene groups. The lineup
+protocol proposed in \citet{buja_statistical_2009} provides a solution
+to over-interpretation. \citet{majumder_validation_2013} showed that its
+use is analogous to the null hypothesis significance testing (NHST)
+framework. More specifically, the protocol consists of \(m\) randomly
+placed plots, where one plot is the data plot, and the remaining
+\(m - 1\) plots, referred to as the \emph{null plots}, have the
+identical graphical procedure as the data plot except the data is
+replaced with a data generation mechanism that is consistent with the
+null hypothesis, \(H_0\). Then, an observer who have not seen the data
+plot will be asked to point out the most different plot from the lineup.
+Under \(H_0\), it is expected that the data plot would have no
+distinguishable difference from the null plots, and the probability that
+the observer correctly picks the data plot is \(1/m\). If one rejects
+\(H_0\) as the observer correctly picks the data plot, then the Type I
+error of this test is \(1/m\). This protocol requires apriori
 specification of \(H_0\) (or at least a null data generating mechanism)

### 2023-04-14 paper.Rmd
quick sketch of abstract

@@
 abstract: |
-  Abstract to fill.
+  Regression experts consistently recommend plotting residuals for model diagnosis, despite the existence of numerous hypothesis test procedures.
+This paper provides evidence for why this is good advice, using data from a visual inference experiment.
+We show how conventional tests are too sensitive to be practically useful, and how the lineup protocol can be used to yield reliable and consistent reading of residual plots for better practical application.
 keywords: |
@@
 In practice, over-interpretation of a single plot is common.
-For instance, in @roy_chowdhury_using_2015, described an published example where the authors over-interpreted separation between gene groups from a two-dimensional projection of a linear discriminant analysis.
+For instance, @roy_chowdhury_using_2015, described a published example where the authors over-interpreted separation between gene groups from a two-dimensional projection of a linear discriminant analysis.
 @roy_chowdhury_using_2015 showed that there were no differences in the expression levels between the gene groups.

### 2023-04-14 paper.tex
quick sketch of abstract

@@
 \begin{abstract}
-Abstract to fill.
+Regression experts consistently recommend plotting residuals for model
+diagnosis, despite the existence of numerous hypothesis test procedures.
+This paper provides evidence for why this is good advice, using data
+from a visual inference experiment. We show how conventional tests are
+too sensitive to be practically useful, and how the lineup protocol can
+be used to yield reliable and consistent reading of residual plots for
+better practical application.
 \end{abstract}
@@
 In practice, over-interpretation of a single plot is common. For
-instance, in \citet{roy_chowdhury_using_2015}, described an published
+instance, \citet{roy_chowdhury_using_2015}, described a published
 example where the authors over-interpreted separation between gene

### 2023-04-16 paper.Rmd
switched facetting in Fig 17 to match the previous two figures

@@
 This paper provides evidence for why this is good advice, using data from a visual inference experiment.
-We show how conventional tests are too sensitive to be practically useful, and how the lineup protocol can be used to yield reliable and consistent reading of residual plots for better practical application.
+We show how conventional tests are too sensitive to be practically useful, and how the lineup protocol can be used to yield reliable and consistent reading of residual plots for better model diagnosis.
 keywords: |
@@
 
-```{r different-x-dist-poly-power, fig.pos="t!", fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for conventional and visual tests for non-linearity and heteroskedasticity patterns.
+```{r different-x-dist-poly-power, fig.pos="t!", fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for conventional and visual tests (columns) for non-linearity and heteroskedasticity patterns (rows).
 The power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and BP tests, respectively.
@@
   labs(col = "", linetype = "Test") +
-  facet_grid(conv_or_not~type, scales = "free_x") +
+  facet_grid(type~conv_or_not, scales = "free_x") +
   theme(legend.position = "bottom")
@@
 
-Figure \ref{fig:different-x-dist-poly-power} examines the impact of the fitted value distribution on the power of conventional (top row) and visual (bottom row) tests for both the non-linearity (left column) and heteroskedasticity (right column) patterns.
+Figure \ref{fig:different-x-dist-poly-power} examines the impact of the fitted value distribution on the power of conventional (left) and visual (right) tests for both the non-linearity (top) and heteroskedasticity (bottom) patterns.
 For conventional tests, only the power curves of appropriate tests which are RESET tests and BP tests for non-linearity pattern and heteroskedasticity pattern respectively are shown.
@@
 
-Motivated by the advice of regression analysis experts, that residual plots as opposed to conventional tests are an indispensible methods for assessing model fit, a human subjects experiment was conducted using visual inference.
+Motivated by the advice of regression analysis experts, that residual plots as opposed to conventional tests are indispensible methods for assessing model fit, a human subjects experiment was conducted using visual inference.
 The experiment tested two primary departures from good residuals: non-linearity and heteroskedasticity.
 
-The experiment found that conventional residual-based statistical tests are more sensitive to weak residual departures from model assumptions than visual tests as would be evaluated by humans.
-That is, conventional tests conclude there are problems with the model fit almost twice as often as humans would.
-They often reject when departures in the form of non-linearity and heteroskedasticity are not visible to a human.
-
-One might say that this is correct behaviour, but it can be argued that the conventional tests are rejecting when it is not necessary.
+The experiment found that conventional residual-based statistical tests are more sensitive to weak departures from model assumptions than visual tests as would be evaluated by humans.
+That is, a conventional test conclude there are problems with the model fit almost twice as often as a human would.
+They often reject when departures in the form of non-linearity and heteroskedasticity are not visibly different from null residual plots.
+
+While it might be argued that the conventional tests are correctly detecting the small but real effects, it can also be considered that the conventional tests are rejecting when it is not necessary.
 Many of these rejections happen even when downstream analysis and results would not be significantly affected by the small departures from a good fit.

### 2023-04-16 paper.tex
switched facetting in Fig 17 to match the previous two figures

@@
 be used to yield reliable and consistent reading of residual plots for
-better practical application.
+better model diagnosis.
 \end{abstract}
@@
 
-\caption{Comparison of power on lineups with different fitted value distributions for conventional and visual tests for non-linearity and heteroskedasticity patterns.
+\caption{Comparison of power on lineups with different fitted value distributions for conventional and visual tests (columns) for non-linearity and heteroskedasticity patterns (rows).
 The power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and BP tests, respectively.
@@
 Figure \ref{fig:different-x-dist-poly-power} examines the impact of the
-fitted value distribution on the power of conventional (top row) and
-visual (bottom row) tests for both the non-linearity (left column) and
-heteroskedasticity (right column) patterns. For conventional tests, only
-the power curves of appropriate tests which are RESET tests and BP tests
-for non-linearity pattern and heteroskedasticity pattern respectively
-are shown. For visual tests, note that, more evaluations on lineup with
-uniform fitted value distribution were collected, so to have a fair
-comparison, we randomly sample five from the total eleven evaluations to
-estimate the power curves. Re-sampling produces the multiple curves for
-the uniform, and provides an indication of the variation of the power.
+fitted value distribution on the power of conventional (left) and visual
+(right) tests for both the non-linearity (top) and heteroskedasticity
+(bottom) patterns. For conventional tests, only the power curves of
+appropriate tests which are RESET tests and BP tests for non-linearity
+pattern and heteroskedasticity pattern respectively are shown. For
+visual tests, note that, more evaluations on lineup with uniform fitted
+value distribution were collected, so to have a fair comparison, we
+randomly sample five from the total eleven evaluations to estimate the
+power curves. Re-sampling produces the multiple curves for the uniform,
+and provides an indication of the variation of the power.
 
@@
 Motivated by the advice of regression analysis experts, that residual
-plots as opposed to conventional tests are an indispensible methods for
+plots as opposed to conventional tests are indispensible methods for
 assessing model fit, a human subjects experiment was conducted using
@@
 The experiment found that conventional residual-based statistical tests
-are more sensitive to weak residual departures from model assumptions
-than visual tests as would be evaluated by humans. That is, conventional
-tests conclude there are problems with the model fit almost twice as
-often as humans would. They often reject when departures in the form of
-non-linearity and heteroskedasticity are not visible to a human.
-
-One might say that this is correct behaviour, but it can be argued that
-the conventional tests are rejecting when it is not necessary. Many of
-these rejections happen even when downstream analysis and results would
-not be significantly affected by the small departures from a good fit.
-The results from human evaluations provide a more practical solution,
-which reinforces the statements from regression experts that residual
-plots are an indispensible method for model diagnostics.
+are more sensitive to weak departures from model assumptions than visual
+tests as would be evaluated by humans. That is, a conventional test
+conclude there are problems with the model fit almost twice as often as
+a human would. They often reject when departures in the form of
+non-linearity and heteroskedasticity are not visibly different from null
+residual plots.
+
+While it might be argued that the conventional tests are correctly
+detecting the small but real effects, it can also be considered that the
+conventional tests are rejecting when it is not necessary. Many of these
+rejections happen even when downstream analysis and results would not be
+significantly affected by the small departures from a good fit. The
+results from human evaluations provide a more practical solution, which
+reinforces the statements from regression experts that residual plots
+are an indispensible method for model diagnostics.
 

### 2023-04-16 appendix.Rmd
added section to appendix exploring the rejection rate by sig level

@@
                     caption = "Examining how decisions might change if $\\hat{\\alpha}$ was different.
-Percentage of lineups that where the $p$-value would switch to above or below 0.05, when $\\hat{\\alpha}$ is multiplied by a multiplier.") %>%
-  kableExtra::kable_styling(latex_options = "scale_down")
+Percentage of lineups that where the $p$-value would switch to above or below 0.05, when $\\hat{\\alpha}$ is multiplied by a multiplier.",
+                    linesep = c("", "", "\\addlinespace"))
 ```
@@
 ```{r n_eval, fig.cap = "Change of power of visual tests for different number of evalutions on lineups with uniform fitted value distribution.
-The power will incrase as the number of evaluations increases, but the margin will decrase.", fig.width=6, fig.height=4, out.width="70%"}
+The power will increase as the number of evaluations increases, but the margin will decrease.", fig.width=6, fig.height=4, out.width="70%"}
 eval_n <- function(n_eval, boot) {
@@
 
-\newpage
+## Conventional test rejection rate for varying significance levels
+
+In the main paper, Sections 5.1 and 5.2 compared the power, and the decisions made by the conventional tests and the visual test.
+The power curves for the visual test is effectively a right-shift from the conventional test.
+The effect is that the visual test rejects less often than the conventional test, at the same significance level.
+We also saw that the visual test rejected a subset of those that the conventional tests rejected.
+This means that they agreed quite well - only residual plots rejected by the contentional tests were rejected by the visual test.
+There was little disagreement, where residual plots not rejected by the conventional test were rejected by the visual test.
+The question arises whether the decisions made conventional test could be made similar to that of the visual test by reducing the significance level.
+Reducing the significance level from 0.05, to 0.01, 0.001, ...
+will have the effect of rejecting fewer of the residual plots.
+
+It would be interesting if a different conventional test significance level results in both the visual tests and conventional tests reject only the same residual plots, and fails to reject the same residual plots.
+This would be a state where both systems agree perfectly.
+Figure \ref{fig:rej-rates} examines this.
+Plot A shows the percentage of residual plots rejected by the visual test, given the conventional test rejected (solid lines) or failed to reject (dashed lines).
+The vertical grey line marks significance level 0.05.
+When the signicance level gets smaller, the it is possible te see that the visual tests reject (nearly) 100% of the time that the conventional test rejects.
+However, there is not agreement, because the visual tests also increasingly reject residual plots where the conventional test failed to reject.
+Plot B is comparable to an ROC curve, where the percentage visual test rejection conditional on conventional test decision is plotted: Reject conditional on reject is plotted against reject conditional on fail to reject, for different significance levels.
+The non-linearity pattern results are close to being ideal, that the percentage of reject relative to fail to reject increases very slowly as the reject relative to reject converges to 100.
+The heteroskedasticity pattern is more problematic, and shows that the cost of rejecting less with the conventional test is disagreement with the visual test.
+
+```{r rej-rates, fig.pos="t!", fig.width=8, fig.height=5, out.width="100%", fig.cap="Changing the significance level of the conventional test will change the rejection rate.
+A: Percentage of conventional tests also rejected by the visual test: rejected (solid), not rejected (dashed).
+As the significance level is reduced, the percentage rejected by the visual test that has been rejected by the conventional test approaches 100.
+The percentage of tests not rejected by the conventional test also increases, as expected, but the percentage of these that are rejected by the visual test increases too.
+B: ROC curve shows that forcing the conventional test to not reject creates a discrepancy with the visual test.
+Many of the residual plots not rejected by the conventional test are rejected by the visual test.
+It isn't possible to vary the significance level of the conventional test to match the decisions made by the visual test."}
+# compute rejection rate for different significance levels, holding visual test constant
+library(ggpubr)
+nl_rates <- NULL
+h_rates <- NULL
+for (i in c(0.1, 0.05, 0.01, 0.001, 0.0001, 0.00001, 0.000001, 0.0000001, 0.00000001, 0.000000001,
+            0.0000000001, 0.00000000001)) {
+p_value_cmp_dat <- vi_survey %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  filter(x_dist == "uniform") %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
+  mutate(conv_reject = ifelse(conventional_p_value <= i, "Reject", "Not"),
+         reject = ifelse(p_value <= 0.05, "Reject", "Not")) %>%
+  mutate(conv_reject = factor(conv_reject, levels = c("Reject", "Not")),
+         reject = factor(reject, levels = c("Reject", "Not"))) %>%
+  select(type, conv_reject, reject) 
+nl <- p_value_cmp_dat %>% 
+  filter(type == "non-linearity") %>% 
+  count(conv_reject, reject, .drop=FALSE) %>%
+  summarise(rej_pct=round(n[1]/(n[1]+n[2])*100, 0),
+            nrej_pct=round(n[3]/(n[3]+n[4])*100, 0)) %>%
+  bind_cols(alpha=i)
+  nl_rates <- bind_rows(nl_rates, nl)
+  # Heteroskedasticity 
+  h <- p_value_cmp_dat %>% 
+  filter(type == "heteroskedasticity") %>% 
+  count(conv_reject, reject, .drop=FALSE) %>%
+  summarise(rej_pct=round(n[1]/(n[1]+n[2])*100, 0),
+            nrej_pct=round(n[3]/(n[3]+n[4])*100, 0)) %>%
+  bind_cols(alpha=i)
+  h_rates <- bind_rows(h_rates, h)
+}
+nl_rates <- nl_rates %>% mutate(type = "non-linearity")
+h_rates <- h_rates %>% mutate(type = "heteroskedasticity")
+rates <- bind_rows(nl_rates, h_rates) %>%
+  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity")))
+rej_by_sig <- ggplot(rates) +
+  geom_vline(xintercept=0.05, colour="grey50") +
+  geom_line(aes(x=alpha, y=rej_pct, colour=type), alpha=0.8) +
+  geom_line(aes(x=alpha, y=nrej_pct, colour=type), linetype=2) +
+  scale_x_log10(breaks=c(0.1, 0.001, 0.00001, 0.0000001, 0.000000001, 0.00000000001)) + 
+  ylab("Visual test rejects (%)") + xlab("Significance level") +
+  scale_colour_brewer("", palette = "Dark2") +
+  ggtitle("A")
+roc <- ggplot(rates) +
+  geom_path(aes(x=nrej_pct, y=rej_pct, colour=type)) +
+  ylab("Reject | Reject (%)") + xlab("Reject | Fail to reject (%)") +
+  ylim(0, 100) +
+  scale_colour_brewer("", palette = "Dark2") +
+  coord_equal() +
+  ggtitle("B")
+ggarrange(rej_by_sig, roc, ncol=2, common.legend=TRUE, legend = "bottom")
+```
+
 # Experiment setup

### 2023-04-16 appendix.tex
added section to appendix exploring the rejection rate by sig level

@@
 \centering
-\resizebox{\linewidth}{!}{
 \begin{tabular}[t]{rrrrr}
@@
 0.875 & 3 & 1.08 & 0 & 0.00\\
+\addlinespace
 1.125 & 0 & 0.00 & 3 & 1.08\\
 1.250 & 0 & 0.00 & 4 & 1.43\\
-\addlinespace
 1.500 & 0 & 0.00 & 5 & 1.79\\
 \bottomrule
-\end{tabular}}
+\end{tabular}
 \end{table}
@@
 \caption{Change of power of visual tests for different number of evalutions on lineups with uniform fitted value distribution.
-The power will incrase as the number of evaluations increases, but the margin will decrase.}\label{fig:n_eval}
+The power will increase as the number of evaluations increases, but the margin will decrease.}\label{fig:n_eval}
 \end{figure}
@@
 
-\newpage
+\hypertarget{conventional-test-rejection-rate-for-varying-significance-levels}{%
+\subsection{Conventional test rejection rate for varying significance
+levels}\label{conventional-test-rejection-rate-for-varying-significance-levels}}
+
+In the main paper, Sections 5.1 and 5.2 compared the power, and the
+decisions made by the conventional tests and the visual test. The power
+curves for the visual test is effectively a right-shift from the
+conventional test. The effect is that the visual test rejects less often
+than the conventional test, at the same significance level. We also saw
+that the visual test rejected a subset of those that the conventional
+tests rejected. This means that they agreed quite well - only residual
+plots rejected by the contentional tests were rejected by the visual
+test. There was little disagreement, where residual plots not rejected
+by the conventional test were rejected by the visual test. The question
+arises whether the decisions made conventional test could be made
+similar to that of the visual test by reducing the significance level.
+Reducing the significance level from 0.05, to 0.01, 0.001, \ldots{} will
+have the effect of rejecting fewer of the residual plots.
+
+It would be interesting if a different conventional test significance
+level results in both the visual tests and conventional tests reject
+only the same residual plots, and fails to reject the same residual
+plots. This would be a state where both systems agree perfectly. Figure
+\ref{fig:rej-rates} examines this. Plot A shows the percentage of
+residual plots rejected by the visual test, given the conventional test
+rejected (solid lines) or failed to reject (dashed lines). The vertical
+grey line marks significance level 0.05. When the signicance level gets
+smaller, the it is possible te see that the visual tests reject (nearly)
+100\% of the time that the conventional test rejects. However, there is
+not agreement, because the visual tests also increasingly reject
+residual plots where the conventional test failed to reject. Plot B is
+comparable to an ROC curve, where the percentage visual test rejection
+conditional on conventional test decision is plotted: Reject conditional
+on reject is plotted against reject conditional on fail to reject, for
+different significance levels. The non-linearity pattern results are
+close to being ideal, that the percentage of reject relative to fail to
+reject increases very slowly as the reject relative to reject converges
+to 100. The heteroskedasticity pattern is more problematic, and shows
+that the cost of rejecting less with the conventional test is
+disagreement with the visual test.
+
+\begin{figure}[t!]
+
+{\centering \includegraphics[width=1\linewidth]{appendix_files/figure-latex/rej-rates-1}
+
+}
+
+\caption{Changing the significance level of the conventional test will change the rejection rate.
+A: Percentage of conventional tests also rejected by the visual test: rejected (solid), not rejected (dashed).
+As the significance level is reduced, the percentage rejected by the visual test that has been rejected by the conventional test approaches 100.
+The percentage of tests not rejected by the conventional test also increases, as expected, but the percentage of these that are rejected by the visual test increases too.
+B: ROC curve shows that forcing the conventional test to not reject creates a discrepancy with the visual test.
+Many of the residual plots not rejected by the conventional test are rejected by the visual test.
+It isn't possible to vary the significance level of the conventional test to match the decisions made by the visual test.}\label{fig:rej-rates}
+\end{figure}
 

### 2023-04-16 paper.Rmd
added section to appendix exploring the rejection rate by sig level

@@
 Figures \ref{fig:polypower} and \ref{fig:heterpower} present the power curves of various tests plotted against the effect size in the residuals, for non-linearity and heteroskedasticity, respectively.
-In each case the power of visual test is re-calculated for multiple bootstrap samples leading to the many (orange) curves.
+In each case the power of visual test is calculated for multiple bootstrap samples leading to the many (orange) curves.
 The effect size was computed at a 5% significance level and plotted on a natural logarithmic scale.
@@
 The test with the uniformly higher power is the RESET test, one that specifically tests for non-linearity.
-The power curves for the visual test are effectively a shift right from that of the RESET test.
-This means that the RESET test will reject a lower effect size (less structure) than the visual test, but otherwise the performance will be similar.
+(Note that the BP and SW tests have much lower power, which is expected because they are not designed to detect non-linearity.) The bootstrapped power curves for the visual test are effectively a shift right from that of the RESET test.
+This means that the RESET test will reject at a lower effect size (less structure) than the visual test, but otherwise the performance will be similar.
 In other words, the RESET test is more sensitive than the visual test.
-This is not necessarily a good feature, for the purposes of diagnosing model defects.
+This is not necessarily a good feature for the purposes of diagnosing model defects.
 If one scans the residual plot examples at the bottom, we might argue that the non-linearity is not sufficiently problematic until an effect size of around 3 or 3.5.
-The RESET test would reject closer to an effect size of 2, but the visual test would reject closer 3.25.
-As expected the BP and SW tests have much lower power - they are not designed to detect non-linearity.
+The RESET test would reject closer to an effect size of 2, but the visual test would reject closer 3.25, for a significance level of 0.05.
+It is also interesting to note that setting the significance level much lower, .001, .0001, .00001, ...
+for the RESET test, so that it rejects less does shift the power curve towards that of the visual test.
+However, it never reaches the matches the performance of the visual test.
 

### 2023-04-16 paper.Rmd
conclusion tweaking and paragraph on changing sig levels

@@
 
-Figures \ref{fig:polypower} compares the power for the different tests for non-linear structure in the residuals.
+Figure \ref{fig:polypower} compares the power for the different tests for non-linear structure in the residuals.
 The test with the uniformly higher power is the RESET test, one that specifically tests for non-linearity.
@@
 The RESET test would reject closer to an effect size of 2, but the visual test would reject closer 3.25, for a significance level of 0.05.
-It is also interesting to note that setting the significance level much lower, .001, .0001, .00001, ...
-for the RESET test, so that it rejects less does shift the power curve towards that of the visual test.
-However, it never reaches the matches the performance of the visual test.
 
@@
 However the BP test would reject at around effect size 2.5.
-Interestingly, the power curve for the SW test (for non-normality) is only sighly different to that of the visual test, suggesting that it performs reasonably for detecting heteroskedasticity, too.
+Interestingly, the power curve for the SW test (for non-normality) is only slightly different to that of the visual test, suggesting that it performs reasonably for detecting heteroskedasticity, too.
 The power curve for the BP test suggests it is not useful for detecting heteroskeadsticity, as expected.
@@
 ```
+
+Because the power curve of the visual tests are a shift to the right of the conventional test (Figures \ref{fig:polypower} and \ref{fig:heter-example}) it would be curious to know whether adjusting the significance level (to .001, .0001, .00001, ...) of the conventional test would generate similar decisions to that of the visual test.
+Interestingly, it doesn't.
+Despite resulting less rejections, neither the RESET or BP tests come to complete agreement with the visual test (See Appendix).
 
@@
 The experiment found that conventional residual-based statistical tests are more sensitive to weak departures from model assumptions than visual tests as would be evaluated by humans.
-That is, a conventional test conclude there are problems with the model fit almost twice as often as a human would.
+That is, a conventional test concludes there are problems with the model fit almost twice as often as a human would.
 They often reject when departures in the form of non-linearity and heteroskedasticity are not visibly different from null residual plots.
@@
 Now it is important to note that residual plots need to be delivered as a lineup, where it is embedded in a field of null plots.
-A residual plot may contains many visual features, but some are caused by the characteristics of the predictors and the randomness of the error, not by the violation of the model assumptions.
-These irrelevant visual features have a chance to be filtered out by subjects with a comparison to null plots, results in a set of more accurate visual findings.
-This enables a careful calibration for reading structure in residual plots.
-
-However, human evaluation of residuals is expensive.
-It is time-consuming, laborious and unfriendly to vision-impaired people.
-This is another reason why it often appears to be ignored.
-With the availability of sophisticated computer vision algorithms today, the goal of this work is to form the basis of providing automated residual plot reading.
-The findings suggest the strong demand of graphical inspection in regression diagnostics, so developing an automatic visual inference system to evaluate lineups of residual plots is valuable.
+A residual plot may contain many visual features, but some are caused by the characteristics of the predictors and the randomness of the error, not by the violation of the model assumptions.
+These irrelevant visual features have a chance to be filtered out by subjects with a comparison to null plots, resulting in more accurate reading.
+The lineup enables a careful calibration for reading structure in residual plots.
+
+Human evaluation of residuals is expensive, though.
+It is time-consuming and laborious.
+This is possibly why residual plot analysis is often not done, in practice.
+However, with the emergence of effective computer vision this work could form the basis to provide automated residual plot reading.
+<!--The findings suggest the strong demand of graphical inspection in regression diagnostics, so developing an automatic visual inference system to evaluate lineups of residual plots is valuable.
 We plan to build a completed open-source system in an `R` package and provide a web interface such as a website for public to interact with.
-Details about this system will be discussed in our next paper.
-
-The experiment also revealed some interesting details about how residual plots are read.
-For the most part, the visual test performed very similarly to the appropriate conventional test only with the power curve shifted in the less sensitive direction.
+Details about this system will be discussed in our next paper.-->
+
+The experiment also revealed some interesting results.
+For the most part, the visual test performed very similarly to the appropriate conventional test except that it has a power curve shifted in the less sensitive direction.
 Unlike the conventional tests, where one needs to specifically test for non-linearity or heteroskedasticity the visual test operated effectively across the range of departures from good residuals.
-
-As expected, if the fitted value distribution is not uniform, there is a loss of power in the visual test.
-Structure is hardest to detect if fitted values are lognormal.
-Also, complex structure are generally harder to detect, but there are outliers.
-Under the designed scenarios in this paper, we find the visual test to be a more robust test against the change of fitted value distributions.
-A surprising finding was that the direction of heteroskedasticity appears to affect the ability to visually detect it, with wedge to the right being less detectable.
+If the fitted value distribution is not uniform, there is a small loss of power in the visual test.
+But quite a surprise to us was the big difference in power of the conventional test, which appear to be quite sensitive to how the predictors are sampled.
+Another surprising finding was that the direction of heteroskedasticity appears to affect the ability to visually detect it: both triangles being more difficult to detect than the butterfly, and a small difference in detection between left- and right-triangle.
 

### 2023-04-18 appendix.Rmd
presubmission change

--- 2023-04-11 appendix.Rmd
@@
 It is configured to handle HTTP requests such that subjects can correctly receive webpages and submit responses.
-Embedded in the resources sent to subjects, the `jsPsych` front-end framework [@jspsych] instructs subjects' browsers to render an environment for running behavioral experiments.
+Embedded in the resources sent to subjects, the `jsPsych` front-end framework [@jspsych] instructs subjects' browsers to render an environment for running behavioural experiments.
 During the experiment, this framework will automatically collect common behavioral data such as response time and clicks on buttons.

### 2023-04-18 paper.Rmd
presubmission change

--- 2023-04-16 paper.Rmd
@@
 abstract: |
-  Regression experts consistently recommend plotting residuals for model diagnosis, despite the existence of numerous hypothesis test procedures.
-This paper provides evidence for why this is good advice, using data from a visual inference experiment.
-We show how conventional tests are too sensitive to be practically useful, and how the lineup protocol can be used to yield reliable and consistent reading of residual plots for better model diagnosis.
+  Abstract to fill.
 keywords: |
@@
 Diagnostics are the key to determining whether there is anything importantly wrong with a model.
-In linear regression analysis, studying the residuals from a model fit is a common diagnostic activity.
+In linear regression analysis, residuals from the model fit are commonly used.
 Residuals summarise what is not captured by the model, and thus provide the capacity to identify what might be wrong.
@@
 We can assess residuals in multiple ways.
-To examine the their univariate distribution, residuals may be plotted as a histogram or normal probability plot.
-Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, we would consider it to be well-behaved.
-However, if the distribution is skewed, bimodal, multimodal, or contains outliers, there would be cause for concern.
+Residuals may be plotted as a histogram or quantile-quantile plot to examine the distribution.
+Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, we consider it to be well behaved.
+However, if the distribution is skewed, bimodal, multimodal, or contains outliers, there is a cause for concern.
 One could also inspect the distribution by conducting a goodness-of-fit test, such as the Shapiro-Wilk (SW) Normality test [@shapiro1965analysis].
 
-In addition, scatterplots of residuals plotted against the fitted values and each of the explanatory variables, are made to scrutinize their relationships.
+More typically, a scatter plot of residuals plotted against the fitted values and each of the explanatory variables, is made to scrutinize their relationships.
 If there are any visually discoverable patterns, the model is potentially inadequate or incorrectly specified.
-In general, one looks for noticeable departures from the model such as non-linear dependency or heteroskedasticity.
-A non-linear dependency would suggest that the model needs to have some additional non-linear terms.
-Heteroskedasticity suggests that the error is dependent on the predictors, and hence violates the independence assumption.
-However, correctly judging whether NO pattern exists in a residual plot is a difficult task for humans.
-We humans will almost always will see a pattern, so the question that really needs answering is whether any pattern perceived is consistent with randomness, purely sampling variability or noise.
-It is especially difficult to teach this to new analysts and students [@loy2021bringing].
-To answer this, there have been numerous conventional hypothesis tests made available to test for non-linear dependence [e.g.
-@ramsey_tests_1969], and heteroskedasticity  [e.g.
-@breusch_simple_1979].
-
-Linear regression is a well-established procedure, and there is considerable literature describing diagnostic procedures, e.g.
+In general, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity.
+However, correctly judging a residual plot where no pattern exists can be a difficult task for humans.
+It is especially common, particularly among students, to misinterpret patterns that are random noise and random deviation from a model [@loy2021bringing].
+There are also numerous conventional hypothesis tests available to test for non-linear dependence [@ramsey_tests_1969], and heteroskedasticity  [@breusch_simple_1979].
+
+An abundance of literature describe appropriate diagnostic methods for linear regression, e.g.
 @draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals.
-An interesting details, is that despite the abundance of conventional tests, ALL of these writings advise that plotting residuals is an essential tool for diagnosing regression model problems.
-@draper1998applied and @belsley_regression_1980 explain that residual plots are usually revealing when the assumptions are violated.
-@cook_applied_1999 thinks formal tests and graphical procedures are complementary and both have a place in residual analysis, but they focus on graphical methods rather than on formal testing.
+All these writings consider plotting residuals to be an essential tool for diagnosing regression modelling problems.
+In addition, @draper1998applied and @belsley_regression_1980 believe that residual plots are usually revealing when the assumptions are violated.
+@cook1982residuals thinks formal tests and graphical procedures are complementary and both have a place in residual analysis, but they focus on graphical methods rather than on formal testing, as they are easier to use.
 @montgomery1982introduction even suggests that residual plots are more informative in most practical situations than the corresponding conventional hypothesis tests.
 
-The common wisdom of experts is that the optimal method for diagnosing model fits is by plotting the data.
-The persistence of this advice to check the plots is *curious*, and investigating why this is, is the subject of this paper.
-
+A common guidance by experts is that optimal method for diagnosing model fits is by plotting the data.
+The persistence of this advice to check the plots is curious, and investigating why this might be common advice is the subject of this paper.
 The paper is structured as follows.
@@
 The results are reported in Section \ref{results}.
-We conclude with a discussion of the presented work, and ideas for future directions.
+We conclude with a discussion of future work, in particular, how the responsibility for residual plot reading might be relieved.
 
@@
 
-Graphical summaries where residuals are plotted against fitted values, or other functions of the predictors (expected to be approximately orthogonal to the residuals) are referred to as standard residual plots in @cook1982residuals.
-Figure \ref{fig:residual-plot-common-departures}A shows an example of an ideal residual plot where points are symmetrically distributed around the horizontal zero line (red), with no discernible patterns.
-There can be various types of departures from this ideal pattern.
-Non-linearity, heteroskedasticity and non-normality are three commonly checked departures.
-
-```{r residual-plot-common-departures, fig.pos="t!", fig.width = 10, fig.height = 2.5, fig.cap = "Example residual vs fitted value plots (red line indicates 0): (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
+```{r residual-plot-common-departures, fig.width = 10, fig.height = 2.5, fig.cap = "Example residual vs fitted value plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
 The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted value plot."}
@@
 
-Model misspecification occurs if functions of predictors that needed to accurately describe the relationship between response and predictors were omitted.
-Any non-linear pattern visible in the residual plot could be indicative of this problem.
+Graphical summaries in which residuals are plotted against fitted values or other functions of the predictors that are approximately orthogonal to the residuals are referred to as standard residual plots in @cook1982residuals.
+Figure \ref{fig:residual-plot-common-departures}A shows an ideal residual plot where the residuals are evenly distributed at both sides of the horizontal zero line, with no noticeable patterns.
+
+There are various types of departures from an ideal residual plot. Non-linearity, heteroskedasticity and non-normality are perhaps the three mostly checked departures.
+
+Non-linearity is a type of model misspecification caused by failing to include higher order terms of the predictors in the regression equation.
+Any non-linear functional form of residuals on fitted values in the residual plot could be indicative of non-linearity.
 An example residual plot containing visual pattern of non-linearity is shown in Figure \ref{fig:residual-plot-common-departures}B.
-One can clearly observe the "S-shape" from the residual plot, which corresponds to cubic term that should have been included in the model.
+One can clearly observe the "S-shape" from the residual plot as the cubic term is not captured by the misspecified model.
 
 Heteroskedasticity refers to the presence of non-constant error variance in a regression model.
-It indicates that the distribution of residuals depends on the predictors, violating the independence assumption.
-This can be seen in a residual plot as an inconsistent spread of the residuals relative to the fitted values or predictors.
-An example is the "butterfly" shape  shown in Figure \ref{fig:residual-plot-common-departures}C, or a "left-triangle" and "right-triangle" shape where the smallest variance occurs at one side of the horizontal axis.
-
-Figure \ref{fig:residual-plot-common-departures}D) shows a scatterplot where the residuals have a skewed distribution, as seen by the uneven vertical spread.
-Unlike non-linearity and heteroskedasticity, non-normality is usually detected with a different type of residual plot, a histogram or normal probability plot.
-Because we focus on scatterplots, non-normality is not one of the departures examined in depth in this paper.
+It is mostly due to the strict but false assumptions on the variance-covariance matrix of the error term.
+The usual pattern of heteroskedasticity on a residual plot is the inconsistent spread of the residuals across the horizontal axis.
+Visually, it sometimes results in the so-called "butterfly" shape as shown in Figure \ref{fig:residual-plot-common-departures}C, or the "left-triangle" and "right-triangle" shape where the smallest variance occurs at one side of the horizontal axis.
+
+Compared to non-linearity and heteroskedasticity, non-normality is usually harder to detect from a residual plot since a scatter plot do not readily reveal the marginal distribution.
+A favourable graphical summary for this task is the quantile-quantile plot.
+As we mainly discuss residual plots, non-normality will not be the focus of this paper.
+For a consistent comparison, the residual plot of this departure is still presented in Figure \ref{fig:residual-plot-common-departures}D.
+When the number residuals below and above the horizontal axis are uneven across the local regions along the $x$-axis, we expect that the normality assumption is violated.
+For example, given a skewed error distribution, there will be fewer data points and more outliers on one side of the horizontal axis as shown in Figure \ref{fig:residual-plot-common-departures}D.
 
@@
 Interestingly, the BP and SW tests also reject the residual plots exhibiting structure that they weren't designed for.
-@cook1982residuals discusses that most residual-based tests for particular types of departure from model assumptions are also sensitive to other types of departures.
-This could be considered a Type III error, where the null hypothesis of good residuals is correctly rejected but for the wrong reason.
-Also, some types of departure can have elements of other types of departure, for example, non-linearity could be viewed as heteroskedasticity.
-Additionally, other problesms such as outliers can trigger the rejection [@cook_applied_1999].
-
-With large sample sizes, hypothesis tests may reject the null hypothesis when there is only a small effect.
-(This is common knowledge but a good discussion can be found in @kirk1996.) While such rejections may be statistically correct, their sensitivity may render the results impractical.
-This is especially pertinent for residual diagnostics.
+@cook1982residuals discusses that most residual-based tests for particular types of departure from model assumptions are also sensitive to other types of departures, that is, correctly rejected but for the wrong reason, a phenomenon known as the "Type III error".
+Also, some types of departure can also have elements of other types of departure, for example, non-linearity could be viewed as heteroskedasticity.
+Additionally, other problems such as outliers can trigger the rejection [@cook_applied_1999].
+
+With large sample sizes, hypothesis tests may reject a residual plot when there is only a slight departure [@belsley_regression_1980].
+While such rejections may be statistically correct, their sensitivity may render the results impractical.
 Therefore, a key goal of residual plot diagnostics is to identify potential issues that could lead to incorrect conclusions or errors in subsequent analyses.
@@
 
-The examination of data plots to infer signal or patterns (or lack thereof) is fraught with variation in the human ability to interpret and decode the information embedded in a graph [@cleveland_graphical_1984].
-To rely on human examination thus feels as though there is a subjective nature to using diagnostic plots to infer appropriateness of statistical models.
-Ideally we would expect visual evaluation should be done objectively and human-to-human variation in reading data plots small, but it cannot be assured.
+The examination of data plots to infer signal or patterns (or lack thereof) is fraught with variation in the human ability to interpret and decode the information embedded in graph [@cleveland_graphical_1984].
+Such examination thus feels there is a subjective nature to evaluate, say, diagnostic plots to infer appropriateness of statistical models, although arguably we would expect such evaluation should be done objectively and such human-to-human variation in "reading" data plots should be minimised.
 
 In practice, over-interpretation of a single plot is common.
-For instance, @roy_chowdhury_using_2015, described a published example where the authors over-interpreted separation between gene groups from a two-dimensional projection of a linear discriminant analysis.
-@roy_chowdhury_using_2015 showed that there were no differences in the expression levels between the gene groups.
-The lineup protocol proposed in @buja_statistical_2009 provides a solution to over-interpretation.
-@majumder_validation_2013 showed that its use is analogous to the null hypothesis significance testing (NHST) framework.
+For instance, in @roy_chowdhury_using_2015, some over-interpreted the separation between gene groups in a two-dimensional projection of a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups.
+To mitigate this over-interpretation, @buja_statistical_2009 proposed a line-up protocol to assess plots in a manner analogous to the null hypothesis significance testing (NHST) framework.
 More specifically, the protocol consists of $m$ randomly placed plots, where one plot is the data plot, and the remaining $m - 1$ plots, referred to as the _null plots_, have the identical graphical procedure as the data plot except the data is replaced with a data generation mechanism that is consistent with the null hypothesis, $H_0$.
@@
 The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009].
-Essentially, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the predictors, then rescaling it by the ratio of residual sum of square in two regressions.
+Essentially, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the predictors, then re-scaling it by the ratio of residual sum of square in two regressions.
 
@@
 Details about the effect size is provided in Section \ref{effect-size} after the introduction of the simulation model used in our human subject experiment.
-The theoretical power of a test is sometimes not a trival solution, but it can be estimated if the data generating process is known.
+The theoretical power of a test is sometimes not a trivial solution, but it can be estimated if the data generating process is known.
 We use a predefined model to generate a large set of simulated data under different effect sizes, and record if the conventional test rejects $H_0$.
@@
 Sometimes the fitted values are discrete because one or more predictors were measured discretely.
-The distribution may be relatively Gaussian, reflecting a linear combinatio of many predictors, adhering to the Central Limit Theorem.
+The distribution may be relatively Gaussian, reflecting a linear combination of many predictors, adhering to the Central Limit Theorem.
 It is also common to see a skewed distribution of fitted values, if one or more of the predictors has a skewed distribution.
@@
 
-Subjects for all three data collection periods are recruited from an crowdsourcing platform called Prolific [@palan2018prolific].
-Prescreening procedure is applied during the recruitment, subjects are required to be fluent in English, with $98\%$ minimum approval rate and 10 minimum submissions in other studies.
+Subjects for all three data collection periods are recruited from an crowd-sourcing platform called Prolific [@palan2018prolific].
+Pre-screening procedure is applied during the recruitment, subjects are required to be fluent in English, with $98\%$ minimum approval rate and 10 minimum submissions in other studies.
 
@@
 However the BP test would reject at around effect size 2.5.
-Interestingly, the power curve for the SW test (for non-normality) is only sighly different to that of the visual test, suggesting that it performs reasonably for detecting heteroskedasticity, too.
-The power curve for the BP test suggests it is not useful for detecting heteroskeadsticity, as expected.
+Interestingly, the power curve for the SW test (for non-normality) is only slightly different to that of the visual test, suggesting that it performs reasonably for detecting heteroskedasticity, too.
+The power curve for the BP test suggests it is not useful for detecting heteroskedasticity, as expected.
 
@@
 ```{r p-value-comparison, fig.width = 8, fig.height = 4, fig.cap = "Rejection rate ($p$-value $\\leq0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
-The visual test rejects less frequently than the conventional test, and (almost) only rejects when the conventional test does.
-Surprisingly, one lineup in the heteroskedasticity group is rejected by the visual test but NOT the conventional test."}
+The visual test rejects less frequently than the conventional test.
+We would generally expect the visual test to only reject when the conventional test does.
+Surprisingly, one lineup containing a heteroskedasticity pattern does not follow this rule.
+"}
 
@@
 ```{r heter-example, fig.pos="t!", fig.height = 7, fig.width = 7, fig.cap = 'The single heteroskedasticity lineup that is rejected by the visual test but not by the BP test.
-The data plot (position 17) contains a ``butterfly" shape.
-It has a log effect size of $ 3.76$, and visibly displays heteroskedasticity, making it somewhat surprising that it is not detected by the BP test.'}
+The data plot at panel 17 contains a "butterfly" shape.
+It has effect size $ = 3.76$, somewhat surprising that it is not detected by the BP test.'}
 # plot heter-331
@@
 
-```{r different-x-dist-poly-power, fig.pos="t!", fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for conventional and visual tests (columns) for non-linearity and heteroskedasticity patterns (rows).
+```{r different-x-dist-poly-power, fig.pos="t!", fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for conventional and visual tests for non-linearity and heteroskedasticity patterns.
 The power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and BP tests, respectively.
@@
   labs(col = "", linetype = "Test") +
-  facet_grid(type~conv_or_not, scales = "free_x") +
+  facet_grid(conv_or_not~type, scales = "free_x") +
   theme(legend.position = "bottom")
@@
 
-Figure \ref{fig:different-x-dist-poly-power} examines the impact of the fitted value distribution on the power of conventional (left) and visual (right) tests for both the non-linearity (top) and heteroskedasticity (bottom) patterns.
+Figure \ref{fig:different-x-dist-poly-power} examines the impact of the fitted value distribution on the power of conventional (top row) and visual (bottom row) tests for both the non-linearity (left column) and heteroskedasticity (right column) patterns.
 For conventional tests, only the power curves of appropriate tests which are RESET tests and BP tests for non-linearity pattern and heteroskedasticity pattern respectively are shown.
@@
 
-Motivated by the advice of regression analysis experts, that residual plots as opposed to conventional tests are indispensible methods for assessing model fit, a human subjects experiment was conducted using visual inference.
+Motivated by the advice of regression analysis experts, that residual plots as opposed to conventional tests are an indispensable methods for assessing model fit, a human subjects experiment was conducted using visual inference.
 The experiment tested two primary departures from good residuals: non-linearity and heteroskedasticity.
 
-The experiment found that conventional residual-based statistical tests are more sensitive to weak departures from model assumptions than visual tests as would be evaluated by humans.
-That is, a conventional test conclude there are problems with the model fit almost twice as often as a human would.
-They often reject when departures in the form of non-linearity and heteroskedasticity are not visibly different from null residual plots.
-
-While it might be argued that the conventional tests are correctly detecting the small but real effects, it can also be considered that the conventional tests are rejecting when it is not necessary.
+The experiment found that conventional residual-based statistical tests are more sensitive to weak residual departures from model assumptions than visual tests as would be evaluated by humans.
+That is, conventional tests conclude there are problems with the model fit almost twice as often as humans would.
+They often reject when departures in the form of non-linearity and heteroskedasticity are not visible to a human.
+
+One might say that this is correct behaviour, but it can be argued that the conventional tests are rejecting when it is not necessary.
 Many of these rejections happen even when downstream analysis and results would not be significantly affected by the small departures from a good fit.
-The results from human evaluations provide a more practical solution, which reinforces the statements from regression experts that residual plots are an indispensible method for model diagnostics.
+The results from human evaluations provide a more practical solution, which reinforces the statements from regression experts that residual plots are an indispensable method for model diagnostics.
 

### 2023-04-18 paper.tex

@@
 residuals, for non-linearity and heteroskedasticity, respectively. In
-each case the power of visual test is re-calculated for multiple
-bootstrap samples leading to the many (orange) curves. The effect size
-was computed at a 5\% significance level and plotted on a natural
+each case the power of visual test is calculated for multiple bootstrap
+samples leading to the many (orange) curves. The effect size was
+computed at a 5\% significance level and plotted on a natural
 logarithmic scale. To facilitate visual calibration of effect size
@@
 
-Figures \ref{fig:polypower} compares the power for the different tests
+Figure \ref{fig:polypower} compares the power for the different tests
 for non-linear structure in the residuals. The test with the uniformly
 higher power is the RESET test, one that specifically tests for
-non-linearity. The power curves for the visual test are effectively a
-shift right from that of the RESET test. This means that the RESET test
-will reject a lower effect size (less structure) than the visual test,
-but otherwise the performance will be similar. In other words, the RESET
-test is more sensitive than the visual test. This is not necessarily a
-good feature, for the purposes of diagnosing model defects. If one scans
-the residual plot examples at the bottom, we might argue that the
-non-linearity is not sufficiently problematic until an effect size of
-around 3 or 3.5. The RESET test would reject closer to an effect size of
-2, but the visual test would reject closer 3.25. As expected the BP and
-SW tests have much lower power - they are not designed to detect
-non-linearity.
+non-linearity. (Note that the BP and SW tests have much lower power,
+which is expected because they are not designed to detect
+non-linearity.) The bootstrapped power curves for the visual test are
+effectively a shift right from that of the RESET test. This means that
+the RESET test will reject at a lower effect size (less structure) than
+the visual test, but otherwise the performance will be similar. In other
+words, the RESET test is more sensitive than the visual test. This is
+not necessarily a good feature for the purposes of diagnosing model
+defects. If one scans the residual plot examples at the bottom, we might
+argue that the non-linearity is not sufficiently problematic until an
+effect size of around 3 or 3.5. The RESET test would reject closer to an
+effect size of 2, but the visual test would reject closer 3.25, for a
+significance level of 0.05.
 
@@
 reject at around effect size 2.5. Interestingly, the power curve for the
-SW test (for non-normality) is only sighly different to that of the
+SW test (for non-normality) is only slightly different to that of the
 visual test, suggesting that it performs reasonably for detecting
@@
 \end{figure}
+
+Because the power curve of the visual tests are a shift to the right of
+the conventional test (Figures \ref{fig:polypower} and
+\ref{fig:heter-example}) it would be curious to know whether adjusting
+the significance level (to .001, .0001, .00001, \ldots) of the
+conventional test would generate similar decisions to that of the visual
+test. Interestingly, it doesn't. Despite resulting less rejections,
+neither the RESET or BP tests come to complete agreement with the visual
+test (See Appendix).
 
@@
 tests as would be evaluated by humans. That is, a conventional test
-conclude there are problems with the model fit almost twice as often as
+concludes there are problems with the model fit almost twice as often as
 a human would. They often reject when departures in the form of
@@
 a lineup, where it is embedded in a field of null plots. A residual plot
-may contains many visual features, but some are caused by the
+may contain many visual features, but some are caused by the
 characteristics of the predictors and the randomness of the error, not
@@
 features have a chance to be filtered out by subjects with a comparison
-to null plots, results in a set of more accurate visual findings. This
-enables a careful calibration for reading structure in residual plots.
-
-However, human evaluation of residuals is expensive. It is
-time-consuming, laborious and unfriendly to vision-impaired people. This
-is another reason why it often appears to be ignored. With the
-availability of sophisticated computer vision algorithms today, the goal
-of this work is to form the basis of providing automated residual plot
-reading. The findings suggest the strong demand of graphical inspection
-in regression diagnostics, so developing an automatic visual inference
-system to evaluate lineups of residual plots is valuable. We plan to
-build a completed open-source system in an \texttt{R} package and
-provide a web interface such as a website for public to interact with.
-Details about this system will be discussed in our next paper.
-
-The experiment also revealed some interesting details about how residual
-plots are read. For the most part, the visual test performed very
-similarly to the appropriate conventional test only with the power curve
-shifted in the less sensitive direction. Unlike the conventional tests,
-where one needs to specifically test for non-linearity or
-heteroskedasticity the visual test operated effectively across the range
-of departures from good residuals.
-
-As expected, if the fitted value distribution is not uniform, there is a
-loss of power in the visual test. Structure is hardest to detect if
-fitted values are lognormal. Also, complex structure are generally
-harder to detect, but there are outliers. Under the designed scenarios
-in this paper, we find the visual test to be a more robust test against
-the change of fitted value distributions. A surprising finding was that
-the direction of heteroskedasticity appears to affect the ability to
-visually detect it, with wedge to the right being less detectable.
+to null plots, resulting in more accurate reading. The lineup enables a
+careful calibration for reading structure in residual plots.
+
+Human evaluation of residuals is expensive, though. It is time-consuming
+and laborious. This is possibly why residual plot analysis is often not
+done, in practice. However, with the emergence of effective computer
+vision this work could form the basis to provide automated residual plot
+reading.
+
+The experiment also revealed some interesting results. For the most
+part, the visual test performed very similarly to the appropriate
+conventional test except that it has a power curve shifted in the less
+sensitive direction. Unlike the conventional tests, where one needs to
+specifically test for non-linearity or heteroskedasticity the visual
+test operated effectively across the range of departures from good
+residuals. If the fitted value distribution is not uniform, there is a
+small loss of power in the visual test. But quite a surprise to us was
+the big difference in power of the conventional test, which appear to be
+quite sensitive to how the predictors are sampled. Another surprising
+finding was that the direction of heteroskedasticity appears to affect
+the ability to visually detect it: both triangles being more difficult
+to detect than the butterfly, and a small difference in detection
+between left- and right-triangle.
 

### 2023-04-22 paper.Rmd
done with susans suggestions for introduction

--- 2023-04-16 paper.Rmd
@@
 
-Diagnostics are the key to determining whether there is anything importantly wrong with a model.
+Diagnosing a model is an important part of analysing data.
 In linear regression analysis, studying the residuals from a model fit is a common diagnostic activity.
@@
 However, if the distribution is skewed, bimodal, multimodal, or contains outliers, there would be cause for concern.
-One could also inspect the distribution by conducting a goodness-of-fit test, such as the Shapiro-Wilk (SW) Normality test [@shapiro1965analysis].
-
-In addition, scatterplots of residuals plotted against the fitted values and each of the explanatory variables, are made to scrutinize their relationships.
+We can also inspect the distribution by conducting a goodness-of-fit test, such as the Shapiro-Wilk (SW) normality test [@shapiro1965analysis].
+
+In addition, scatterplots of residuals plotted against the fitted values and each of the explanatory variables, can be used to scrutinize their relationships.
 If there are any visually discoverable patterns, the model is potentially inadequate or incorrectly specified.
-In general, one looks for noticeable departures from the model such as non-linear dependency or heteroskedasticity.
-A non-linear dependency would suggest that the model needs to have some additional non-linear terms.
+In general, one looks for noticeable departures from the model such as non-linear pattern or heteroskedasticity.
+A non-linear pattern would suggest that the model needs to have some additional non-linear terms.
 Heteroskedasticity suggests that the error is dependent on the predictors, and hence violates the independence assumption.
 However, correctly judging whether NO pattern exists in a residual plot is a difficult task for humans.
-We humans will almost always will see a pattern, so the question that really needs answering is whether any pattern perceived is consistent with randomness, purely sampling variability or noise.
+Humans will almost always will see a pattern (see @kahneman), so the question that really needs answering is whether any pattern perceived is consistent with randomness, purely sampling variability, or noise.
 It is especially difficult to teach this to new analysts and students [@loy2021bringing].
-To answer this, there have been numerous conventional hypothesis tests made available to test for non-linear dependence [e.g.
+To address this, statistical tests are available to check for non-linear patterns [e.g.
 @ramsey_tests_1969], and heteroskedasticity  [e.g.
@@
 @draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals.
-An interesting details, is that despite the abundance of conventional tests, ALL of these writings advise that plotting residuals is an essential tool for diagnosing regression model problems.
-@draper1998applied and @belsley_regression_1980 explain that residual plots are usually revealing when the assumptions are violated.
-@cook_applied_1999 thinks formal tests and graphical procedures are complementary and both have a place in residual analysis, but they focus on graphical methods rather than on formal testing.
-@montgomery1982introduction even suggests that residual plots are more informative in most practical situations than the corresponding conventional hypothesis tests.
-
-The common wisdom of experts is that the optimal method for diagnosing model fits is by plotting the data.
-The persistence of this advice to check the plots is *curious*, and investigating why this is, is the subject of this paper.
+Interestingly, despite the abundance of conventional tests, ALL of these writings advise that plotting residuals is an essential tool for diagnosing regression model problems:
+
+> *Some most useful checks allowed by data plots should be done on a routine basis for every regression.* [@draper1998applied]
+
+> *Such formal and informal procedures are complementary, and both have a place in residual analysis.* [@cook1982residuals]
+
+> *In our experience, statistical tests on regression model residuals are not widely used.
+In most practical situations the residual plots are more informative than the corresponding tests.
+However, since residual plots do require skill and experience to interpret, the statistical tests may occasionally prove useful.* [@montgomery1982introduction]
+
+\noindent The common wisdom of experts is that plotting the residuals is indispensable for diagnosing model fits.
+The ubiquity of this advice is *curious*: investigating why, is the subject of this paper.
 
@@
 
-Graphical summaries where residuals are plotted against fitted values, or other functions of the predictors (expected to be approximately orthogonal to the residuals) are referred to as standard residual plots in @cook1982residuals.
+Graphical summaries where residuals are plotted against fitted values, or other functions of the predictors (expected to be approximately orthogonal to the residuals) are considered to be the most important residual plots by @cook_applied_1999.
 Figure \ref{fig:residual-plot-common-departures}A shows an example of an ideal residual plot where points are symmetrically distributed around the horizontal zero line (red), with no discernible patterns.

### 2023-04-22 paper.tex
done with susans suggestions for introduction

@@
 
-Diagnostics are the key to determining whether there is anything
-importantly wrong with a model. In linear regression analysis, studying
-the residuals from a model fit is a common diagnostic activity.
-Residuals summarise what is not captured by the model, and thus provide
-the capacity to identify what might be wrong.
+Diagnosing a model is an important part of analysing data. In linear
+regression analysis, studying the residuals from a model fit is a common
+diagnostic activity. Residuals summarise what is not captured by the
+model, and thus provide the capacity to identify what might be wrong.
 
@@
 skewed, bimodal, multimodal, or contains outliers, there would be cause
-for concern. One could also inspect the distribution by conducting a
-goodness-of-fit test, such as the Shapiro-Wilk (SW) Normality test
+for concern. We can also inspect the distribution by conducting a
+goodness-of-fit test, such as the Shapiro-Wilk (SW) normality test
 \citep{shapiro1965analysis}.
@@
 In addition, scatterplots of residuals plotted against the fitted values
-and each of the explanatory variables, are made to scrutinize their
+and each of the explanatory variables, can be used to scrutinize their
 relationships. If there are any visually discoverable patterns, the
@@
 one looks for noticeable departures from the model such as non-linear
-dependency or heteroskedasticity. A non-linear dependency would suggest
-that the model needs to have some additional non-linear terms.
+pattern or heteroskedasticity. A non-linear pattern would suggest that
+the model needs to have some additional non-linear terms.
 Heteroskedasticity suggests that the error is dependent on the
@@
 correctly judging whether NO pattern exists in a residual plot is a
-difficult task for humans. We humans will almost always will see a
-pattern, so the question that really needs answering is whether any
-pattern perceived is consistent with randomness, purely sampling
-variability or noise. It is especially difficult to teach this to new
-analysts and students \citep{loy2021bringing}. To answer this, there
-have been numerous conventional hypothesis tests made available to test
-for non-linear dependence \citep[e.g.][]{ramsey_tests_1969}, and
-heteroskedasticity \citep[e.g.][]{breusch_simple_1979}.
+difficult task for humans. Humans will almost always will see a pattern
+(see \citet{kahneman}), so the question that really needs answering is
+whether any pattern perceived is consistent with randomness, purely
+sampling variability, or noise. It is especially difficult to teach this
+to new analysts and students \citep{loy2021bringing}. To address this,
+statistical tests are available to check for non-linear patterns
+\citep[e.g.][]{ramsey_tests_1969}, and heteroskedasticity
+\citep[e.g.][]{breusch_simple_1979}.
 
@@
 \citet{belsley_regression_1980}, \citet{cook_applied_1999} and
-\citet{cook1982residuals}. An interesting details, is that despite the
-abundance of conventional tests, ALL of these writings advise that
-plotting residuals is an essential tool for diagnosing regression model
-problems. \citet{draper1998applied} and \citet{belsley_regression_1980}
-explain that residual plots are usually revealing when the assumptions
-are violated. \citet{cook_applied_1999} thinks formal tests and
-graphical procedures are complementary and both have a place in residual
-analysis, but they focus on graphical methods rather than on formal
-testing. \citet{montgomery1982introduction} even suggests that residual
-plots are more informative in most practical situations than the
-corresponding conventional hypothesis tests.
-
-The common wisdom of experts is that the optimal method for diagnosing
-model fits is by plotting the data. The persistence of this advice to
-check the plots is \emph{curious}, and investigating why this is, is the
-subject of this paper.
+\citet{cook1982residuals}. Interestingly, despite the abundance of
+conventional tests, ALL of these writings advise that plotting residuals
+is an essential tool for diagnosing regression model problems:
+
+\begin{quote}
+\emph{Some most useful checks allowed by data plots should be done on a
+routine basis for every regression.} \citep{draper1998applied}
+\end{quote}
+
+\begin{quote}
+\emph{Such formal and informal procedures are complementary, and both
+have a place in residual analysis.} \citep{cook1982residuals}
+\end{quote}
+
+\begin{quote}
+\emph{In our experience, statistical tests on regression model residuals
+are not widely used. In most practical situations the residual plots are
+more informative than the corresponding tests. However, since residual
+plots do require skill and experience to interpret, the statistical
+tests may occasionally prove useful.} \citep{montgomery1982introduction}
+\end{quote}
+
+\noindent The common wisdom of experts is that plotting the residuals is
+indispensable for diagnosing model fits. The ubiquity of this advice is
+\emph{curious}: investigating why, is the subject of this paper.
 
@@
 or other functions of the predictors (expected to be approximately
-orthogonal to the residuals) are referred to as standard residual plots
-in \citet{cook1982residuals}. Figure
+orthogonal to the residuals) are considered to be the most important
+residual plots by \citet{cook_applied_1999}. Figure
 \ref{fig:residual-plot-common-departures}A shows an example of an ideal

### 2023-04-22 paper.Rmd
up to 2.3

@@
 However, if the distribution is skewed, bimodal, multimodal, or contains outliers, there would be cause for concern.
-We can also inspect the distribution by conducting a goodness-of-fit test, such as the Shapiro-Wilk (SW) normality test [@shapiro1965analysis].
+We can also inspect the distribution by conducting a goodness-of-fit test, such as the Shapiro-Wilk normality test [@shapiro1965analysis].
 
@@
 There can be various types of departures from this ideal pattern.
-Non-linearity, heteroskedasticity and non-normality are three commonly checked departures.
+Non-linearity, heteroskedasticity and non-normality, shown in Figure \ref{fig:residual-plot-common-departures}B,C,D are three commonly checked departures.
 
@@
 
-Model misspecification occurs if functions of predictors that needed to accurately describe the relationship between response and predictors were omitted.
+Model misspecification occurs if functions of predictors that needed to accurately describe the relationship between response and predictors are omitted.
 Any non-linear pattern visible in the residual plot could be indicative of this problem.
@@
 Figure \ref{fig:residual-plot-common-departures}D) shows a scatterplot where the residuals have a skewed distribution, as seen by the uneven vertical spread.
-Unlike non-linearity and heteroskedasticity, non-normality is usually detected with a different type of residual plot, a histogram or normal probability plot.
+Unlike non-linearity and heteroskedasticity, non-normality is usually detected with a different type of residual plot: a histogram or normal probability plot.
 Because we focus on scatterplots, non-normality is not one of the departures examined in depth in this paper.
+(@loy_variations_2016 and @loy2013diagnostic discuss related work on non-normality checking.)
 
@@
 Many different hypothesis tests are available to detect specific model defects.
-For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the BP test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution.
+For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan (BP) test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution.
 To test specific forms of non-linearity, one may apply the F-test as a model structural test to examine the significance of specific polynomial and non-linear forms of the predictors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].
-The SW test [@shapiro1965analysis] is the most widely used test of non-normality included by many of the statistical software programs.
-The Jarque--Bera test [@jarque1980efficient] is also used to directly check whether the sample skewness and kurtosis match a normal distribution.
+The Shapiro-Wilk (SW) normality test [@shapiro1965analysis] is the most widely used test of non-normality included by many of the statistical software programs.
+The Jarque-Bera test [@jarque1980efficient] is also used to directly check whether the sample skewness and kurtosis match a normal distribution.
 
@@
 The SW test was computed using the `shapiro.test` from the core R package `stats`.
-(The the R package `skedastic` [@skedastic] contains a large collection of tests for heteroskedasticity.) Although, the RESET test is exact, it requires the selection of a power parameter.
-According to @ramsey_tests_1969, a power of four is recommended, which we adopted in our analysis.
+(Although we didn't use it, it's useful to know that the R package `skedastic` [@skedastic] also contains a large collection of functions to test for heteroskedasticity.) Although, the RESET test is exact, it requires the selection of a power parameter.
+@ramsey_tests_1969 recommends a power of four, which we adopted in our analysis.
 The BP and SW tests are approximate.
 
-We would expect the RESET test for non-linearity to reject residual plot B, the BP test for heteroskedasticity to reject the residual plot C, and SW test for non-normality to reject residual plot D, which they all do and correctly fail to reject residual plot A.
+We would expect the RESET test for non-linearity to reject residual plot B, the BP test for heteroskedasticity to reject the residual plot C, and the SW test for non-normality to reject residual plot D, which they all do and all tests also correctly fail to reject residual plot A.
 Interestingly, the BP and SW tests also reject the residual plots exhibiting structure that they weren't designed for.
-@cook1982residuals discusses that most residual-based tests for particular types of departure from model assumptions are also sensitive to other types of departures.
-This could be considered a Type III error, where the null hypothesis of good residuals is correctly rejected but for the wrong reason.
+@cook1982residuals explain that most residual-based tests for particular types of departure from model assumptions are also sensitive to other types of departures.
+This could be considered a Type III error (XXXREF), where the null hypothesis of good residuals is correctly rejected but for the wrong reason.
 Also, some types of departure can have elements of other types of departure, for example, non-linearity could be viewed as heteroskedasticity.
-Additionally, other problesms such as outliers can trigger the rejection [@cook_applied_1999].
+Additionally, other data problems such as outliers can trigger rejection (or not) of the null hypothesis [@cook_applied_1999].
 
 With large sample sizes, hypothesis tests may reject the null hypothesis when there is only a small effect.
-(This is common knowledge but a good discussion can be found in @kirk1996.) While such rejections may be statistically correct, their sensitivity may render the results impractical.
-This is especially pertinent for residual diagnostics.
-Therefore, a key goal of residual plot diagnostics is to identify potential issues that could lead to incorrect conclusions or errors in subsequent analyses.
-However, minor defects in the model are unlikely to have a significant impact and may be best disregarded for practical purposes.
-In fact, the experiment discussed in this paper specifically addresses this.
+(A good discussion can be found in @kirk1996.) While such rejections may be statistically correct, their sensitivity may render the results impractical.
+A key goal of residual plot diagnostics is to identify potential issues that could lead to incorrect conclusions or errors in subsequent analyses, but minor defects in the model are unlikely to have a significant impact and may be best disregarded for practical purposes.
+The experiment discussed in this paper specifically addresses this tension between statistical and practical statistics.
 

### 2023-04-22 paper.tex
up to 2.3

@@
 for concern. We can also inspect the distribution by conducting a
-goodness-of-fit test, such as the Shapiro-Wilk (SW) normality test
+goodness-of-fit test, such as the Shapiro-Wilk normality test
 \citep{shapiro1965analysis}.
@@
 various types of departures from this ideal pattern. Non-linearity,
-heteroskedasticity and non-normality are three commonly checked
-departures.
+heteroskedasticity and non-normality, shown in Figure
+\ref{fig:residual-plot-common-departures}B,C,D are three commonly
+checked departures.
 
@@
 Model misspecification occurs if functions of predictors that needed to
-accurately describe the relationship between response and predictors
-were omitted. Any non-linear pattern visible in the residual plot could
-be indicative of this problem. An example residual plot containing
-visual pattern of non-linearity is shown in Figure
+accurately describe the relationship between response and predictors are
+omitted. Any non-linear pattern visible in the residual plot could be
+indicative of this problem. An example residual plot containing visual
+pattern of non-linearity is shown in Figure
 \ref{fig:residual-plot-common-departures}B. One can clearly observe the
@@
 non-normality is usually detected with a different type of residual
-plot, a histogram or normal probability plot. Because we focus on
+plot: a histogram or normal probability plot. Because we focus on
 scatterplots, non-normality is not one of the departures examined in
-depth in this paper.
+depth in this paper. (\citet{loy_variations_2016} and
+\citet{loy2013diagnostic} discuss related work on non-normality
+checking.)
 
@@
 tested by applying the White test
-\citep{white_heteroskedasticity-consistent_1980} or the BP test
-\citep{breusch_simple_1979}, which are both derived from the Lagrange
-multiplier test \citep{silvey1959lagrangian} principle that relies on
-the asymptotic properties of the null distribution. To test specific
-forms of non-linearity, one may apply the F-test as a model structural
-test to examine the significance of specific polynomial and non-linear
-forms of the predictors, or the significance of proxy variables as in
-the Ramsey Regression Equation Specification Error Test (RESET)
-\citep{ramsey_tests_1969}. The SW test \citep{shapiro1965analysis} is
-the most widely used test of non-normality included by many of the
-statistical software programs. The Jarque--Bera test
-\citep{jarque1980efficient} is also used to directly check whether the
-sample skewness and kurtosis match a normal distribution.
+\citep{white_heteroskedasticity-consistent_1980} or the Breusch-Pagan
+(BP) test \citep{breusch_simple_1979}, which are both derived from the
+Lagrange multiplier test \citep{silvey1959lagrangian} principle that
+relies on the asymptotic properties of the null distribution. To test
+specific forms of non-linearity, one may apply the F-test as a model
+structural test to examine the significance of specific polynomial and
+non-linear forms of the predictors, or the significance of proxy
+variables as in the Ramsey Regression Equation Specification Error Test
+(RESET) \citep{ramsey_tests_1969}. The Shapiro-Wilk (SW) normality test
+\citep{shapiro1965analysis} is the most widely used test of
+non-normality included by many of the statistical software programs. The
+Jarque-Bera test \citep{jarque1980efficient} is also used to directly
+check whether the sample skewness and kurtosis match a normal
+distribution.
 
@@
 computed using the \texttt{shapiro.test} from the core R package
-\texttt{stats}. (The the R package \texttt{skedastic} \citep{skedastic}
-contains a large collection of tests for heteroskedasticity.) Although,
-the RESET test is exact, it requires the selection of a power parameter.
-According to \citet{ramsey_tests_1969}, a power of four is recommended,
-which we adopted in our analysis. The BP and SW tests are approximate.
+\texttt{stats}. (Although we didn't use it, it's useful to know that the
+R package \texttt{skedastic} \citep{skedastic} also contains a large
+collection of functions to test for heteroskedasticity.) Although, the
+RESET test is exact, it requires the selection of a power parameter.
+\citet{ramsey_tests_1969} recommends a power of four, which we adopted
+in our analysis. The BP and SW tests are approximate.
 
@@
 B, the BP test for heteroskedasticity to reject the residual plot C, and
-SW test for non-normality to reject residual plot D, which they all do
-and correctly fail to reject residual plot A. Interestingly, the BP and
-SW tests also reject the residual plots exhibiting structure that they
-weren't designed for. \citet{cook1982residuals} discusses that most
-residual-based tests for particular types of departure from model
-assumptions are also sensitive to other types of departures. This could
-be considered a Type III error, where the null hypothesis of good
-residuals is correctly rejected but for the wrong reason. Also, some
-types of departure can have elements of other types of departure, for
-example, non-linearity could be viewed as heteroskedasticity.
-Additionally, other problesms such as outliers can trigger the rejection
+the SW test for non-normality to reject residual plot D, which they all
+do and all tests also correctly fail to reject residual plot A.
+Interestingly, the BP and SW tests also reject the residual plots
+exhibiting structure that they weren't designed for.
+\citet{cook1982residuals} explain that most residual-based tests for
+particular types of departure from model assumptions are also sensitive
+to other types of departures. This could be considered a Type III error
+(XXXREF), where the null hypothesis of good residuals is correctly
+rejected but for the wrong reason. Also, some types of departure can
+have elements of other types of departure, for example, non-linearity
+could be viewed as heteroskedasticity. Additionally, other data problems
+such as outliers can trigger rejection (or not) of the null hypothesis
 \citep{cook_applied_1999}.
@@
 With large sample sizes, hypothesis tests may reject the null hypothesis
-when there is only a small effect. (This is common knowledge but a good
-discussion can be found in \citet{kirk1996}.) While such rejections may
-be statistically correct, their sensitivity may render the results
-impractical. This is especially pertinent for residual diagnostics.
-Therefore, a key goal of residual plot diagnostics is to identify
-potential issues that could lead to incorrect conclusions or errors in
-subsequent analyses. However, minor defects in the model are unlikely to
-have a significant impact and may be best disregarded for practical
-purposes. In fact, the experiment discussed in this paper specifically
-addresses this.
+when there is only a small effect. (A good discussion can be found in
+\citet{kirk1996}.) While such rejections may be statistically correct,
+their sensitivity may render the results impractical. A key goal of
+residual plot diagnostics is to identify potential issues that could
+lead to incorrect conclusions or errors in subsequent analyses, but
+minor defects in the model are unlikely to have a significant impact and
+may be best disregarded for practical purposes. The experiment discussed
+in this paper specifically addresses this tension between statistical
+and practical statistics.
 

### 2023-04-22 paper.Rmd
up to section 3

@@
 
-Figure \ref{fig:residual-plot-common-departures}D) shows a scatterplot where the residuals have a skewed distribution, as seen by the uneven vertical spread.
+Figure \ref{fig:residual-plot-common-departures}D shows a scatterplot where the residuals have a skewed distribution, as seen by the uneven vertical spread.
 Unlike non-linearity and heteroskedasticity, non-normality is usually detected with a different type of residual plot: a histogram or normal probability plot.
 Because we focus on scatterplots, non-normality is not one of the departures examined in depth in this paper.
-(@loy_variations_2016 and @loy2013diagnostic discuss related work on non-normality checking.)
+(loy_variations_2016 discuss related work on non-normality checking.)
 
@@
 The examination of data plots to infer signal or patterns (or lack thereof) is fraught with variation in the human ability to interpret and decode the information embedded in a graph [@cleveland_graphical_1984].
-To rely on human examination thus feels as though there is a subjective nature to using diagnostic plots to infer appropriateness of statistical models.
-Ideally we would expect visual evaluation should be done objectively and human-to-human variation in reading data plots small, but it cannot be assured.
+Human examination of diagnostic plots can feel subjective.
+Ideally we would expect visual evaluation to be done objectively with small amounts of human-to-human variation in reading data plots, but we cannot assume this is the case.
 
 In practice, over-interpretation of a single plot is common.
-For instance, @roy_chowdhury_using_2015, described a published example where the authors over-interpreted separation between gene groups from a two-dimensional projection of a linear discriminant analysis.
+For instance, @roy_chowdhury_using_2015 described a published example where authors over-interpreted separation between gene groups from a two-dimensional projection of a linear discriminant analysis.
 @roy_chowdhury_using_2015 showed that there were no differences in the expression levels between the gene groups.
-The lineup protocol proposed in @buja_statistical_2009 provides a solution to over-interpretation.
-@majumder_validation_2013 showed that its use is analogous to the null hypothesis significance testing (NHST) framework.
-More specifically, the protocol consists of $m$ randomly placed plots, where one plot is the data plot, and the remaining $m - 1$ plots, referred to as the _null plots_, have the identical graphical procedure as the data plot except the data is replaced with a data generation mechanism that is consistent with the null hypothesis, $H_0$.
-Then, an observer who have not seen the data plot will be asked to point out the most different plot from the lineup.
+One solution to over-interpretation is to examine the plot in the context of natural sampling variability assumed by the model, called the lineup protocol, as proposed in @buja_statistical_2009.
+@majumder_validation_2013 showed that the lineup protocol is analogous to the null hypothesis significance testing framework.
+The protocol consists of $m$ randomly placed plots, where one plot is the data plot, and the remaining $m - 1$ plots, referred to as the _null plots_, are constructed using the same graphical procedure as the data plot but the data is replaced with null data that is generated in a manner consistent with the null hypothesis, $H_0$.
+Then, an observer who has not seen the data plot is asked to point out the most different plot from the lineup.
 Under $H_0$, it is expected that the data plot would have no distinguishable difference from the null plots, and the probability that the observer correctly picks the data plot is $1/m$.
 If one rejects $H_0$ as the observer correctly picks the data plot, then the Type I error of this test is $1/m$.
-This protocol requires apriori specification of $H_0$ (or at least a null data generating mechanism) much like the requirement of knowing the distribution of the test statistic in NHST.
+This protocol requires a priori specification of $H_0$ (or at least a null data generating mechanism), much like the requirement of knowing the sampling distribution of the test statistic in null hypothesis significance testing framework.
 
@@
 
-Data used in the $m - 1$ null plots needs to be simulated.
+Data used in the $m - 1$ null plots is typically simulated.
 In regression diagnostics, sampling data consistent with $H_0$ is equivalent to sampling data from the assumed model.
@@
 
-The effectiveness of lineup protocol for regression analysis is validated by @majumder_validation_2013 under relatively simple settings with up to two predictors.
-Their results suggest that visual tests are capable of testing the significance of a single predictor with a similar power as a t-test, though they express that in general it is unnecessary to use visual inference if there exists a conventional test, and they do not expect the visual test to perform equally well as the conventional test.
-In their third experiment, where there is not a conventional test, visual test outperforms the conventional test for a large margin.
-This is encouraging, as it promotes the use of visual inference in situations where there are no existing statistical testing procedures.
-Visual inference have also been integrated into diagnostic of hierarchical linear models by @loy2013diagnostic, @loy2014hlmdiag and @loy2015you.
-They use lineup protocols to judge the assumption of linearity, normality and constant error variance for both the level-1 and level-2 residuals.
+The effectiveness of lineup protocol for regression analysis has been validated by @majumder_validation_2013 under relatively simple settings with up to two predictors.
+Their results suggest that visual tests are capable of testing the significance of a single predictor with a similar power to a t-test, though they express that in general it is unnecessary to use visual inference if there exists a conventional test, and they do not expect the visual test to perform equally well as the conventional test.
+In their third experiment, where there is not a conventional test, visual test outperforms the conventional test by a large margin.
+This is encouraging, as it supports the use of visual inference in situations where there are no existing statistical testing procedures.
+Visual inference has also been integrated into diagnostics for hierarchical linear models where the lineup protocol is used to judge the assumptions of linearity, normality and constant error variance for both the level-1 and level-2 residuals [@loy2013diagnostic, @loy2014hlmdiag and @loy2015you].
 

### 2023-04-22 paper.tex
up to section 3

@@
 
-Figure \ref{fig:residual-plot-common-departures}D) shows a scatterplot
+Figure \ref{fig:residual-plot-common-departures}D shows a scatterplot
 where the residuals have a skewed distribution, as seen by the uneven
@@
 scatterplots, non-normality is not one of the departures examined in
-depth in this paper. (\citet{loy_variations_2016} and
-\citet{loy2013diagnostic} discuss related work on non-normality
-checking.)
+depth in this paper. (loy\_variations\_2016 discuss related work on
+non-normality checking.)
 
@@
 decode the information embedded in a graph
-\citep{cleveland_graphical_1984}. To rely on human examination thus
-feels as though there is a subjective nature to using diagnostic plots
-to infer appropriateness of statistical models. Ideally we would expect
-visual evaluation should be done objectively and human-to-human
-variation in reading data plots small, but it cannot be assured.
+\citep{cleveland_graphical_1984}. Human examination of diagnostic plots
+can feel subjective. Ideally we would expect visual evaluation to be
+done objectively with small amounts of human-to-human variation in
+reading data plots, but we cannot assume this is the case.
 
 In practice, over-interpretation of a single plot is common. For
-instance, \citet{roy_chowdhury_using_2015}, described a published
-example where the authors over-interpreted separation between gene
-groups from a two-dimensional projection of a linear discriminant
-analysis. \citet{roy_chowdhury_using_2015} showed that there were no
-differences in the expression levels between the gene groups. The lineup
-protocol proposed in \citet{buja_statistical_2009} provides a solution
-to over-interpretation. \citet{majumder_validation_2013} showed that its
-use is analogous to the null hypothesis significance testing (NHST)
-framework. More specifically, the protocol consists of \(m\) randomly
-placed plots, where one plot is the data plot, and the remaining
-\(m - 1\) plots, referred to as the \emph{null plots}, have the
-identical graphical procedure as the data plot except the data is
-replaced with a data generation mechanism that is consistent with the
-null hypothesis, \(H_0\). Then, an observer who have not seen the data
-plot will be asked to point out the most different plot from the lineup.
-Under \(H_0\), it is expected that the data plot would have no
-distinguishable difference from the null plots, and the probability that
-the observer correctly picks the data plot is \(1/m\). If one rejects
-\(H_0\) as the observer correctly picks the data plot, then the Type I
-error of this test is \(1/m\). This protocol requires apriori
-specification of \(H_0\) (or at least a null data generating mechanism)
-much like the requirement of knowing the distribution of the test
-statistic in NHST.
+instance, \citet{roy_chowdhury_using_2015} described a published example
+where authors over-interpreted separation between gene groups from a
+two-dimensional projection of a linear discriminant analysis.
+\citet{roy_chowdhury_using_2015} showed that there were no differences
+in the expression levels between the gene groups. One solution to
+over-interpretation is to examine the plot in the context of natural
+sampling variability assumed by the model, called the lineup protocol,
+as proposed in \citet{buja_statistical_2009}.
+\citet{majumder_validation_2013} showed that the lineup protocol is
+analogous to the null hypothesis significance testing framework. The
+protocol consists of \(m\) randomly placed plots, where one plot is the
+data plot, and the remaining \(m - 1\) plots, referred to as the
+\emph{null plots}, are constructed using the same graphical procedure as
+the data plot but the data is replaced with null data that is generated
+in a manner consistent with the null hypothesis, \(H_0\). Then, an
+observer who has not seen the data plot is asked to point out the most
+different plot from the lineup. Under \(H_0\), it is expected that the
+data plot would have no distinguishable difference from the null plots,
+and the probability that the observer correctly picks the data plot is
+\(1/m\). If one rejects \(H_0\) as the observer correctly picks the data
+plot, then the Type I error of this test is \(1/m\). This protocol
+requires a priori specification of \(H_0\) (or at least a null data
+generating mechanism), much like the requirement of knowing the sampling
+distribution of the test statistic in null hypothesis significance
+testing framework.
 
@@
 
-Data used in the \(m - 1\) null plots needs to be simulated. In
+Data used in the \(m - 1\) null plots is typically simulated. In
 regression diagnostics, sampling data consistent with \(H_0\) is
@@
 
-The effectiveness of lineup protocol for regression analysis is
+The effectiveness of lineup protocol for regression analysis has been
 validated by \citet{majumder_validation_2013} under relatively simple
@@
 tests are capable of testing the significance of a single predictor with
-a similar power as a t-test, though they express that in general it is
+a similar power to a t-test, though they express that in general it is
 unnecessary to use visual inference if there exists a conventional test,
@@
 conventional test. In their third experiment, where there is not a
-conventional test, visual test outperforms the conventional test for a
-large margin. This is encouraging, as it promotes the use of visual
+conventional test, visual test outperforms the conventional test by a
+large margin. This is encouraging, as it supports the use of visual
 inference in situations where there are no existing statistical testing
-procedures. Visual inference have also been integrated into diagnostic
-of hierarchical linear models by \citet{loy2013diagnostic},
-\citet{loy2014hlmdiag} and \citet{loy2015you}. They use lineup protocols
-to judge the assumption of linearity, normality and constant error
-variance for both the level-1 and level-2 residuals.
+procedures. Visual inference has also been integrated into diagnostics
+for hierarchical linear models where the lineup protocol is used to
+judge the assumptions of linearity, normality and constant error
+variance for both the level-1 and level-2 residuals
+\citep[\citet{loy2014hlmdiag} and
+\citet{loy2015you}]{loy2013diagnostic}.
 

### 2023-04-22 paper.Rmd
up to section 3.3

@@
 
-In diagnosing a model fit from residuals, we are generally interested in *the regression model is correctly specified* ($H_0$) against the broad alternative  *the regression model is misspecified* ($H_a$).
-
-However, it is practically impossible to test this specific $H_0$ with conventional tests, which are constructed to measure specific departures.
-For example, the RESET test is formulated as $H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0$ against $H_a: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0$, from $y = \tau_0 + \sum_{i=1}^{p}\tau_px_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)$.
+In diagnosing a model fit using the residuals, we are generally interested in testing whether "*the regression model is correctly specified*" ($H_0$) against the broad alternative  "*the regression model is misspecified*" ($H_a$).
+
+However, it is practically impossible to test this broad $H_0$ with conventional tests, because they need specific structure causing the departure to be quantifiable in order to be computable.
+For example, the RESET test for detecting non-linear departures is formulated by fitting $y = \tau_0 + \sum_{i=1}^{p}\tau_px_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)$ in order to test $H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0$ against $H_a: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0$.
 Similarly, the BP test is designed to specifically test $H_0:$ *error variances are all equal* ($\zeta_i=0 \text{ for } i=1,..,p$) versus the alternative $H_a:$ *that the error variances are a multiplicative function of one or more variables* ($\text{at least one } \zeta_i\neq 0$) from $e^2 = \zeta_0 + \sum_{i=1}^{p}\zeta_i x_i + u, ~ u\sim N(0,\sigma_u^2)$.
 
-One of the potential benefits of the visual test, based on the lineup protocol, is that it works as an omnibus test,  able to detect a range of departures from good residuals.
+While a battery of conventional tests for different types of departures could be applied, this is intrinsic to the lineup protocol.
+The lineup protocol operates as an omnibus test,  able to detect a range of departures from good residuals in a single application.
 
@@
 
-In hypothesis testing, a $p$-value is defined as the probability of observing test results as least as extreme as the observed result given $H_0$ is true.
-Conventional hypothesis tests usually have an existing method to derive or compute $p$-value based on the null distribution.
-What we need to discuss in the following is the method to estimate $p$-value for a visual test.
-
-Within the context of visual inference, by involving $k$ independent observers, the visual $p$-value can be interpreted as the probability of having as many or more subjects detect the data plot than the observed result.
-
-Let $X_j = \{0,1\}$ be a Bernoulli random variable denoting whether subject $j$ correctly detecting the data plot, and $X = \sum_{j=1}^{K}X_j$ be the number of observers correctly picking the data plot.
-Then, by imposing a relatively strong assumption on the visual test that all $K$ evaluations are fully independent, under $H_0$, $X \sim \mathrm{Binom}_{K,1/m}$.
-Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as $P(X \geq x) = 1 - F(x) + f(x)$, where $F(.)$ is the binomial cumulative distribution function, $f(.)$ is the binomial probability mass function and $x$ is the realization of number of observers correctly picking the data plot [@majumder_validation_2013].
-
-As pointed out by @vanderplas2021statistical, this basic binomial model does not take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup.
-And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup.
-@vanderplas2021statistical summarises three common scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects.
-Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experiment design.
-For Scenario 3, @vanderplas2021statistical models the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$.
+In hypothesis testing, a $p$-value is defined as the probability of observing test results at least as extreme as the observed result assuming $H_0$ is true.
+Conventional hypothesis tests usually have an existing method to derive or compute the $p$-value based on the null distribution.
+Here we establish the method to estimate a $p$-value for a visual test.
+
+Within the context of visual inference, with $K$ independent observers, the visual $p$-value can be seen as the probability of having as many or more subjects detect the data plot than the observed result.
+
+The approach used in @majumder_validation_2013 is as follows.
+Define $X_j = \{0,1\}$ to be a Bernoulli random variable measuring whether subject $j$ detected the data plot, and $X = \sum_{j=1}^{K}X_j$ be the total number of observers who detected the data plot.
+Then, by imposing a relatively strong assumption that all $K$ evaluations are fully independent, under $H_0$ $X \sim \mathrm{Binom}_{K,1/m}$.
+Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is estimated with $P(X \geq x) = 1 - F(x) + f(x)$, where $F(.)$ is the binomial cumulative distribution function, $f(.)$ is the binomial probability mass function and $x$ is the realization of number of observers choosing the data plot.
+
+As pointed out by @vanderplas2021statistical, this basic binomial model is deficient.
+It does not take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup, or account for when subjects are offered the option to select one or more "most different" plots, or none, from a lineup.
+They suggest three common lineup scenarios: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects.
+Scenario 3 is the most feasible to apply, but has the most dependencies to accommodate for the $p$-value calculation.
+For Scenario 3, VanderPlas et al propose modeling the probability of plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$.
 The number of times plot $i$ being selected in $K$ evaluations is denoted as $c_i$.
@@
 
-Note that Equation \ref{eq:pvalue-beta-binomial} given in @vanderplas2021statistical only works with non-negative integer $c_i$.
-We extend the equation to non-negative real number $c_i$ by applying a linear approximation
+\noindent We extend the equation to non-negative real number $c_i$ by applying a linear approximation
 
@@
 
-Besides, the parameter $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} and \ref{eq:pmf-beta-binomial} is usually unknown and hence needs to be estimated from the survey data.
-For low values of $\alpha$, only a few plots are attractive to the observers and tend to be selected.
-For higher values of $\alpha$, the distribution of the probability of each plot being selected is more even.
-@vanderplas2021statistical defines that a plot is $c$-interesting if $c$ or more participants select the plot as the most different.
-Given the definition, The expected number of plots selected at least $c$ times, $E[Z_c]$, is calculated as
+The parameter $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} and \ref{eq:pmf-beta-binomial} is usually unknown and will need to be estimated from the survey data.
+An interpretation of $\alpha$ is that when it is low only a few plots are attractive to the observers and tend to be selected, and when high, most plots are equally likely to be chosen.
+VanderPlas et al define $c$-interesting plot to be if $c$ or more participants select the plot as the most different.
+The expected number of plots selected at least $c$ times, $E[Z_c]$, is then calculated as
 
@@
 With Equation \ref{eq:c-interesting-expectation}, $\alpha$ can be estimated using maximum likelihood estimation.
-But for precise estimate of $\alpha$, additional human responses to Rorschach lineups, which is a type of lineup that consists of plots constructed from the same null data generating mechanism, are required.
+Precise estimation of $\alpha$, is aided by evaluation of Rorschach lineups, where all plots are null plots.
+In a Rorschach, in theory all plots should be equally likely, but in practice some (irrelevant) visual elements may be more eye-catching than others.
+This is what $alpha$ captures, the capacity for extraneous features to distract the observer for a particular type of plot display.
 

### 2023-04-22 paper.tex
up to section 3.3

@@
 
-In diagnosing a model fit from residuals, we are generally interested in
-\emph{the regression model is correctly specified} (\(H_0\)) against the
-broad alternative \emph{the regression model is misspecified} (\(H_a\)).
-
-However, it is practically impossible to test this specific \(H_0\) with
-conventional tests, which are constructed to measure specific
-departures. For example, the RESET test is formulated as
-\(H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0\) against
-\(H_a: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0\),
-from
-\(y = \tau_0 + \sum_{i=1}^{p}\tau_px_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)\).
+In diagnosing a model fit using the residuals, we are generally
+interested in testing whether ``\emph{the regression model is correctly
+specified}'' (\(H_0\)) against the broad alternative ``\emph{the
+regression model is misspecified}'' (\(H_a\)).
+
+However, it is practically impossible to test this broad \(H_0\) with
+conventional tests, because they need specific structure causing the
+departure to be quantifiable in order to be computable. For example, the
+RESET test for detecting non-linear departures is formulated by fitting
+\(y = \tau_0 + \sum_{i=1}^{p}\tau_px_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)\)
+in order to test \(H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0\) against
+\(H_a: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0\).
 Similarly, the BP test is designed to specifically test \(H_0:\)
@@
 
-One of the potential benefits of the visual test, based on the lineup
-protocol, is that it works as an omnibus test, able to detect a range of
-departures from good residuals.
+While a battery of conventional tests for different types of departures
+could be applied, this is intrinsic to the lineup protocol. The lineup
+protocol operates as an omnibus test, able to detect a range of
+departures from good residuals in a single application.
 
@@
 In hypothesis testing, a \(p\)-value is defined as the probability of
-observing test results as least as extreme as the observed result given
-\(H_0\) is true. Conventional hypothesis tests usually have an existing
-method to derive or compute \(p\)-value based on the null distribution.
-What we need to discuss in the following is the method to estimate
-\(p\)-value for a visual test.
-
-Within the context of visual inference, by involving \(k\) independent
-observers, the visual \(p\)-value can be interpreted as the probability
-of having as many or more subjects detect the data plot than the
-observed result.
-
-Let \(X_j = \{0,1\}\) be a Bernoulli random variable denoting whether
-subject \(j\) correctly detecting the data plot, and
-\(X = \sum_{j=1}^{K}X_j\) be the number of observers correctly picking
-the data plot. Then, by imposing a relatively strong assumption on the
-visual test that all \(K\) evaluations are fully independent, under
-\(H_0\), \(X \sim \mathrm{Binom}_{K,1/m}\). Therefore, the \(p\)-value
-of a lineup of size \(m\) evaluated by \(K\) observer is given as
+observing test results at least as extreme as the observed result
+assuming \(H_0\) is true. Conventional hypothesis tests usually have an
+existing method to derive or compute the \(p\)-value based on the null
+distribution. Here we establish the method to estimate a \(p\)-value for
+a visual test.
+
+Within the context of visual inference, with \(K\) independent
+observers, the visual \(p\)-value can be seen as the probability of
+having as many or more subjects detect the data plot than the observed
+result.
+
+The approach used in \citet{majumder_validation_2013} is as follows.
+Define \(X_j = \{0,1\}\) to be a Bernoulli random variable measuring
+whether subject \(j\) detected the data plot, and
+\(X = \sum_{j=1}^{K}X_j\) be the total number of observers who detected
+the data plot. Then, by imposing a relatively strong assumption that all
+\(K\) evaluations are fully independent, under \(H_0\)
+\(X \sim \mathrm{Binom}_{K,1/m}\). Therefore, the \(p\)-value of a
+lineup of size \(m\) evaluated by \(K\) observer is estimated with
 \(P(X \geq x) = 1 - F(x) + f(x)\), where \(F(.)\) is the binomial
@@
 mass function and \(x\) is the realization of number of observers
-correctly picking the data plot \citep{majumder_validation_2013}.
+choosing the data plot.
 
 As pointed out by \citet{vanderplas2021statistical}, this basic binomial
-model does not take into account the possible dependencies in the visual
-test due to repeated evaluations of the same lineup. And it is
-inapplicable to visual test where subjects are asked to select one or
-more ``most different'' plots from the lineup.
-\citet{vanderplas2021statistical} summarises three common scenarios in
-visual inference: (1) \(K\) different lineups are shown to \(K\)
-subjects, (2) \(K\) lineups with different null plots but the same data
-plot are shown to \(K\) subjects, and (3) the same lineup is shown to
-\(K\) subjects. Out of these three scenarios, Scenario 3 is the most
-common in previous studies as it puts the least constraints on the
-experiment design. For Scenario 3, \citet{vanderplas2021statistical}
-models the probability of a plot \(i\) being selected from a lineup as
-\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
-\(i=1,...,m\) and \(\alpha > 0\). The number of times plot \(i\) being
-selected in \(K\) evaluations is denoted as \(c_i\). In case subject
-\(j\) makes multiple selections, \(1/s_j\) will be added to \(c_i\)
-instead of one, where \(s_j\) is the number of plots subject \(j\)
-selected for \(j=1,...K\). This ensures \(\sum_{i}c_i=K\). Since we are
-only interested in the selections of the data plot \(i\), the marginal
-model can be simplified to a beta-binomial model and thus the visual
-\(p\)-value is given as
+model is deficient. It does not take into account the possible
+dependencies in the visual test due to repeated evaluations of the same
+lineup, or account for when subjects are offered the option to select
+one or more ``most different'' plots, or none, from a lineup. They
+suggest three common lineup scenarios: (1) \(K\) different lineups are
+shown to \(K\) subjects, (2) \(K\) lineups with different null plots but
+the same data plot are shown to \(K\) subjects, and (3) the same lineup
+is shown to \(K\) subjects. Scenario 3 is the most feasible to apply,
+but has the most dependencies to accommodate for the \(p\)-value
+calculation. For Scenario 3, VanderPlas et al propose modeling the
+probability of plot \(i\) being selected from a lineup as \(\theta_i\),
+where \(\theta_i \sim Dirichlet(\alpha)\) for \(i=1,...,m\) and
+\(\alpha > 0\). The number of times plot \(i\) being selected in \(K\)
+evaluations is denoted as \(c_i\). In case subject \(j\) makes multiple
+selections, \(1/s_j\) will be added to \(c_i\) instead of one, where
+\(s_j\) is the number of plots subject \(j\) selected for \(j=1,...K\).
+This ensures \(\sum_{i}c_i=K\). Since we are only interested in the
+selections of the data plot \(i\), the marginal model can be simplified
+to a beta-binomial model and thus the visual \(p\)-value is given as
 
@@
 
-Note that Equation \ref{eq:pvalue-beta-binomial} given in
-\citet{vanderplas2021statistical} only works with non-negative integer
-\(c_i\). We extend the equation to non-negative real number \(c_i\) by
+\noindent We extend the equation to non-negative real number \(c_i\) by
 applying a linear approximation
@@
 
-Besides, the parameter \(\alpha\) used in Equation
-\ref{eq:pvalue-beta-binomial} and \ref{eq:pmf-beta-binomial} is usually
-unknown and hence needs to be estimated from the survey data. For low
-values of \(\alpha\), only a few plots are attractive to the observers
-and tend to be selected. For higher values of \(\alpha\), the
-distribution of the probability of each plot being selected is more
-even. \citet{vanderplas2021statistical} defines that a plot is
-\(c\)-interesting if \(c\) or more participants select the plot as the
-most different. Given the definition, The expected number of plots
-selected at least \(c\) times, \(E[Z_c]\), is calculated as
+The parameter \(\alpha\) used in Equation \ref{eq:pvalue-beta-binomial}
+and \ref{eq:pmf-beta-binomial} is usually unknown and will need to be
+estimated from the survey data. An interpretation of \(\alpha\) is that
+when it is low only a few plots are attractive to the observers and tend
+to be selected, and when high, most plots are equally likely to be
+chosen. VanderPlas et al define \(c\)-interesting plot to be if \(c\) or
+more participants select the plot as the most different. The expected
+number of plots selected at least \(c\) times, \(E[Z_c]\), is then
+calculated as
 
@@
 With Equation \ref{eq:c-interesting-expectation}, \(\alpha\) can be
-estimated using maximum likelihood estimation. But for precise estimate
-of \(\alpha\), additional human responses to Rorschach lineups, which is
-a type of lineup that consists of plots constructed from the same null
-data generating mechanism, are required.
+estimated using maximum likelihood estimation. Precise estimation of
+\(\alpha\), is aided by evaluation of Rorschach lineups, where all plots
+are null plots. In a Rorschach, in theory all plots should be equally
+likely, but in practice some (irrelevant) visual elements may be more
+eye-catching than others. This is what \(alpha\) captures, the capacity
+for extraneous features to distract the observer for a particular type
+of plot display.
 

### 2023-04-22 paper.Rmd
up to section 4

@@
 when one is concerned about whether model assumptions have been violated.
-Although in practice, one might be more interested in knowing how much the residuals deviate from the model assumptions, and whether this deviation is of practical significance.
+In practice, one might be more interested in knowing how much the residuals deviate from the model assumptions, and whether this deviation is of practical significance.
 
@@
 These two can be quantified in terms of effect size $E$ to measure the strength of the residual departures from the model assumptions.
-Details about the effect size is provided in Section \ref{effect-size} after the introduction of the simulation model used in our human subject experiment.
+Details about the effect size are provided in Section \ref{effect-size} after the introduction of the simulation model used in our experiment.
 The theoretical power of a test is sometimes not a trival solution, but it can be estimated if the data generating process is known.
@@
 \noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
-The effect size $E$ is the only predictor and the intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$, which is the desired significance level.
-
-The power of a visual test on the other hand, may additionally depend on the ability of the particular subject, as the skill of the individual may affect the number of observers who identify the data plot from the lineup [@majumder_validation_2013].
-To address this issue, @majumder_validation_2013 models the probability of a subject $j$ correctly picking the data plot from a lineup $l$ using a mixed-effect logistic regression, with subjects treated as random effect.
+The effect size $E$ is the only predictor and the intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$, the desired significance level.
+
+The power of a visual test on the other hand, may additionally depend on the ability of the particular subject, as the skill of each individual may affect the number of observers who identify the data plot from the lineup [@majumder_validation_2013].
+To address this issue, @majumder_validation_2013 models the probability of subject $j$ correctly picking the data plot from lineup $l$ using a mixed-effect logistic regression, with subjects treated as random effect.
 Then, the estimated power of a visual test evaluated by a single subject is the predicted value obtained from the mixed effects model.
@@
 Therefore, we directly estimate the probability of a lineup being rejected by assuming that individual skill has negligible effect on the variation of the power.
-This assumption is not necessary true, but it helps simplifying the model structure, thereby obviate a costly large-scale experiment to estimate complex covariance matrices.
+This assumption is not necessary true, but it helps to simplify the model structure, thereby obviating a costly large-scale experiment to estimate complex covariance matrices.
 The same model given in Equation \ref{eq:logistic-regression-1-1} is applied to model the power of a visual test.

### 2023-04-22 paper.tex
up to section 4

@@
 specific way. It is an important indicator when one is concerned about
-whether model assumptions have been violated. Although in practice, one
-might be more interested in knowing how much the residuals deviate from
-the model assumptions, and whether this deviation is of practical
-significance.
+whether model assumptions have been violated. In practice, one might be
+more interested in knowing how much the residuals deviate from the model
+assumptions, and whether this deviation is of practical significance.
 
@@
 of the residual departures from the model assumptions. Details about the
-effect size is provided in Section \ref{effect-size} after the
-introduction of the simulation model used in our human subject
-experiment. The theoretical power of a test is sometimes not a trival
-solution, but it can be estimated if the data generating process is
-known. We use a predefined model to generate a large set of simulated
-data under different effect sizes, and record if the conventional test
-rejects \(H_0\). The probability of the conventional test rejects
-\(H_0\) is then fitted by a logistic regression formulated as
+effect size are provided in Section \ref{effect-size} after the
+introduction of the simulation model used in our experiment. The
+theoretical power of a test is sometimes not a trival solution, but it
+can be estimated if the data generating process is known. We use a
+predefined model to generate a large set of simulated data under
+different effect sizes, and record if the conventional test rejects
+\(H_0\). The probability of the conventional test rejects \(H_0\) is
+then fitted by a logistic regression formulated as
 
@@
 predictor and the intercept is fixed to \(log(0.05/0.95)\) so that
-\(\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05\), which is the desired
+\(\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05\), the desired
 significance level.
@@
 The power of a visual test on the other hand, may additionally depend on
-the ability of the particular subject, as the skill of the individual
+the ability of the particular subject, as the skill of each individual
 may affect the number of observers who identify the data plot from the
 lineup \citep{majumder_validation_2013}. To address this issue,
-\citet{majumder_validation_2013} models the probability of a subject
-\(j\) correctly picking the data plot from a lineup \(l\) using a
-mixed-effect logistic regression, with subjects treated as random
-effect. Then, the estimated power of a visual test evaluated by a single
-subject is the predicted value obtained from the mixed effects model.
-However, this mixed effects model does not work with scenario where
-subjects are asked to select one or more most different plots. In this
-scenario, having the probability of a subject \(j\) correctly picking
-the data plot from a lineup \(l\) is insufficient to determine the power
-of a visual test because it does not provide information about the
-number of selections made by the subject for the calculation of the
-\(p\)-value (See Equation \ref{eq:pvalue-beta-binomial-approx}).
-Therefore, we directly estimate the probability of a lineup being
-rejected by assuming that individual skill has negligible effect on the
-variation of the power. This assumption is not necessary true, but it
-helps simplifying the model structure, thereby obviate a costly
-large-scale experiment to estimate complex covariance matrices. The same
-model given in Equation \ref{eq:logistic-regression-1-1} is applied to
-model the power of a visual test.
+\citet{majumder_validation_2013} models the probability of subject \(j\)
+correctly picking the data plot from lineup \(l\) using a mixed-effect
+logistic regression, with subjects treated as random effect. Then, the
+estimated power of a visual test evaluated by a single subject is the
+predicted value obtained from the mixed effects model. However, this
+mixed effects model does not work with scenario where subjects are asked
+to select one or more most different plots. In this scenario, having the
+probability of a subject \(j\) correctly picking the data plot from a
+lineup \(l\) is insufficient to determine the power of a visual test
+because it does not provide information about the number of selections
+made by the subject for the calculation of the \(p\)-value (See Equation
+\ref{eq:pvalue-beta-binomial-approx}). Therefore, we directly estimate
+the probability of a lineup being rejected by assuming that individual
+skill has negligible effect on the variation of the power. This
+assumption is not necessary true, but it helps to simplify the model
+structure, thereby obviating a costly large-scale experiment to estimate
+complex covariance matrices. The same model given in Equation
+\ref{eq:logistic-regression-1-1} is applied to model the power of a
+visual test.
 

### 2023-04-23 paper.Rmd
done with section 4.1.1

@@
 
-An experiment is conducted in three data collection periods to investigate the difference between conventional hypothesis testing and visual inference in the application of linear regression diagnostics.
-Two types of departures, non-linearity and heteroskedasticity, are collected during data collection periods I and II.
+Our experiment was conducted over three data collection periods to investigate the difference between conventional hypothesis testing and visual inference in the application of linear regression diagnostics.
+Two types of departures, non-linearity and heteroskedasticity, were collected during data collection periods I and II.
 The data collection period III was designed primarily to measure human responses to null lineups so that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated.
-Additional lineups for both non-linearity and heteroskedasticity, using uniform fitted value distribution, were included so that the participants were evaluating some lineups with signal also.
-It would be too frustrating for participants to only be assigned lineups with all null plots.
+Additional lineups for both non-linearity and heteroskedasticity, using uniform fitted value distributions, were included for additional data, and to avoid participant frustration of too many difficult tasks.
 Overall, we collected `r vi_survey  %>% filter(!attention_check) %>% nrow()` evaluations on `r vi_survey  %>% filter(!attention_check) %>% pull(unique_lineup_id) %>% unique() %>% length()` unique lineups performed by `r vi_survey  %>% filter(!attention_check) %>% count(exp, set) %>% nrow()` subjects throughout three data collection periods.
 
+A summary of the factors used in the experiment can be found in Table \ref{tab:model-factor-table}.
+There were four levels of the non-linear structure, and three levels of heteroskedastic structure.
+The signal strength was controlled by error variance ($\sigma$) for the non-linear pattern, and by a ratio ($b$) parameter for the heteroskedasticity.
+Additionally, three levels of sample size ($n$) and four different fitted value distributions were incorporated.
+
 ## Simulating departures from good residuals
@@
 ### Non-linearity
-
-Data collection period I is designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials on random vector $\boldsymbol{x}$ formulated as
-
-\begin{align} \label{eq:nonlinearity-model}
-\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
-\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
-\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
-\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
-\end{align}
-
-\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
-
-```{=tex}
-\begin{equation} \label{eq:scaling-function}
-g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x}))2k - k, \quad \text{for} \quad k > 0. 
-\end{equation}
-```
-
-According to @abramowitz1964handbook, Hermite polynomials were initially defined by @de1820theorie, but named after Hermite [@hermite1864nouveau] because of the unrecognisable form of Laplace's work.
-When simulating $\boldsymbol{z}_{raw}$, function `hermite` from the R package `mpoly` [@mpoly] is used to generate Hermite polynomials.
-
-The null regression model used to fit the realizations generated by the above model is formulated as
-
-```{=tex}
-\begin{equation} \label{eq:null-model}
-\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
-\end{equation}
-```
-\noindent where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
-
-Since $z = O(x^j)$, for $j > 1$, $z$ is a higher order term leaves out by the null regression, which will lead to model misspecification.
-
-Visual patterns of non-linearity are simulated using four different orders of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$).
-(A summary of the factors is given in Table \ref{tab:model-factor-table}.) The values of $j$ is chosen so that distinct shapes of non-linearity are included in the residual plot.
-These include "U", "S", "M" and "triple-U" shape as shown in Figure \ref{fig:different-shape-of-herimite}.
-A greater value of $j$ will result in a curve with more turning points.
-It is expected that the "U" shape will be the easiest one to detect because complex shape tends to be concealed by cluster of data points.
-
-Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups used in non-linearity detection.
-This lineup is produced by the non-linearity model with $j = 6$.
-The data plot location is $2^3 - 4$.
-All five subjects correctly identify the data plot from this lineup.
 
@@
 
-```{r different-shape-of-herimite, fig.height = 2, fig.cap = "Polynomial forms generated for the residual plots used to assess detecting non-linearity.
+Data collection period I was designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials on random vector $\boldsymbol{x}$ formulated as
+
+\begin{align*} \label{eq:nonlinearity-model}
+\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
+\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
+\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
+\end{align*}
+
+\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
+
+```{=tex}
+\begin{equation} \label{eq:scaling-function}
+g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x}))2k - k, \quad \text{for} \quad k > 0. 
+\end{equation}
+```
+
+According to @abramowitz1964handbook, Hermite polynomials were initially defined by @de1820theorie, but named after @hermite1864nouveau because of the unrecognisable form of Laplace's work.
+The function `hermite` from the R package `mpoly` [@mpoly] is used to  simulate $\boldsymbol{z}_{raw}$ to generate Hermite polynomials.
+
+The null regression model used to fit the realizations generated by the above model is formulated as
+
+```{=tex}
+\begin{equation} \label{eq:null-model}
+\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
+\end{equation}
+```
+\noindent where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
+Since $z = O(x^j)$, for $j > 1$, $z$ is a higher order term left out of the null regression, which will lead to model misspecification.
+
+```{r different-shape-of-herimite, fig.pos="!h", fig.height = 2, fig.cap = "Polynomial forms generated for the residual plots used to assess detecting non-linearity.
 The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$."}
@@
 
-```{r different-sigma, fig.height = 2, fig.cap = 'Examining the effect of $\\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform fitted value distribution and the "U" shape.
+```{r different-sigma, fig.pos="!h", fig.height = 2, fig.cap = 'Examining the effect of $\\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform fitted value distribution and the "U" shape.
 As $\\sigma$ increases the signal strength decreases, to the point that the "U" is almost unrecognisable when $\\sigma=4$.'}
@@
 
+Visual patterns of non-linearity are simulated using four different orders of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$).
+The values of $j$ were chosen so that distinct shapes of non-linearity were included in the residual plot.
+These include "U", "S", "M" and "triple-U" shape as shown in Figure \ref{fig:different-shape-of-herimite}.
+A greater value of $j$ will result in a curve with more turning points.
+It is expected that the "U" shape will be the easiest detect, as the shape gets more complex it will be harder to perceive in a scatterplot, particularly when there is noise.
+Figure \ref{fig:different-sigma} shows the "U" shape for different amounts of noise ($\sigma$).
+
+Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups used in non-linearity detection.
+This lineup is produced by the non-linearity model with $j = 6$.
+The data plot location is $2^3 - 4$.
+All five subjects correctly identify the data plot from this lineup.
+
 ```{r example-poly-lineup, fig.pos="t!", fig.height = 7, fig.width = 7, fig.cap = "One of the lineups containing non-linearity patterns used in data collection period I.
@@
 Visual patterns of heteroskedasticity are simulated using three different shapes ($a$ = -1, 0, 1).
-(A summary of the factors can be found in Table \ref{tab:model-factor-table}.)
 

### 2023-04-23 paper.tex
done with section 4.1.1

@@
 
-An experiment is conducted in three data collection periods to
+Our experiment was conducted over three data collection periods to
 investigate the difference between conventional hypothesis testing and
 visual inference in the application of linear regression diagnostics.
-Two types of departures, non-linearity and heteroskedasticity, are
+Two types of departures, non-linearity and heteroskedasticity, were
 collected during data collection periods I and II. The data collection
@@
 both non-linearity and heteroskedasticity, using uniform fitted value
-distribution, were included so that the participants were evaluating
-some lineups with signal also. It would be too frustrating for
-participants to only be assigned lineups with all null plots. Overall,
-we collected 7974 evaluations on 1152 unique lineups performed by 443
+distributions, were included for additional data, and to avoid
+participant frustration of too many difficult tasks. Overall, we
+collected 7974 evaluations on 1152 unique lineups performed by 443
 subjects throughout three data collection periods.
+
+A summary of the factors used in the experiment can be found in Table
+\ref{tab:model-factor-table}. There were four levels of the non-linear
+structure, and three levels of heteroskedastic structure. The signal
+strength was controlled by error variance (\(\sigma\)) for the
+non-linear pattern, and by a ratio (\(b\)) parameter for the
+heteroskedasticity. Additionally, three levels of sample size (\(n\))
+and four different fitted value distributions were incorporated.
 
@@
 \subsubsection{Non-linearity}\label{non-linearity}}
-
-Data collection period I is designed to study the ability of human
-subjects to detect the effect of a non-linear term \(\boldsymbol{z}\)
-constructed using Hermite polynomials on random vector
-\(\boldsymbol{x}\) formulated as
-
-\begin{align} \label{eq:nonlinearity-model}
-\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
-\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
-\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
-\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
-\end{align}
-
-\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
-\(\boldsymbol{\varepsilon}\), \(\boldsymbol{x}_{raw}\),
-\(\boldsymbol{z}_{raw}\) are vectors of size \(n\), \(He_{j}(.)\) is the
-\(j\)th-order probabilist's Hermite polynomials,
-\(\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\), and
-\(g(\boldsymbol{x}, k)\) is a scaling function to enforce the support of
-the random vector to be \([-k, k]^n\) defined as
-
-\begin{equation} \label{eq:scaling-function}
-g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x}))2k - k, \quad \text{for} \quad k > 0. 
-\end{equation}
-
-According to \citet{abramowitz1964handbook}, Hermite polynomials were
-initially defined by \citet{de1820theorie}, but named after Hermite
-\citep{hermite1864nouveau} because of the unrecognisable form of
-Laplace's work. When simulating \(\boldsymbol{z}_{raw}\), function
-\texttt{hermite} from the R package \texttt{mpoly} \citep{mpoly} is used
-to generate Hermite polynomials.
-
-The null regression model used to fit the realizations generated by the
-above model is formulated as
-
-\begin{equation} \label{eq:null-model}
-\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
-\end{equation}
-
-\noindent where
-\(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\).
-
-Since \(z = O(x^j)\), for \(j > 1\), \(z\) is a higher order term leaves
-out by the null regression, which will lead to model misspecification.
-
-Visual patterns of non-linearity are simulated using four different
-orders of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)). (A
-summary of the factors is given in Table \ref{tab:model-factor-table}.)
-The values of \(j\) is chosen so that distinct shapes of non-linearity
-are included in the residual plot. These include ``U'', ``S'', ``M'' and
-``triple-U'' shape as shown in Figure
-\ref{fig:different-shape-of-herimite}. A greater value of \(j\) will
-result in a curve with more turning points. It is expected that the
-``U'' shape will be the easiest one to detect because complex shape
-tends to be concealed by cluster of data points.
-
-Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups
-used in non-linearity detection. This lineup is produced by the
-non-linearity model with \(j = 6\). The data plot location is
-\(2^3 - 4\). All five subjects correctly identify the data plot from
-this lineup.
 
@@
 
-\begin{figure}
+Data collection period I was designed to study the ability of human
+subjects to detect the effect of a non-linear term \(\boldsymbol{z}\)
+constructed using Hermite polynomials on random vector
+\(\boldsymbol{x}\) formulated as
+
+\begin{align*} \label{eq:nonlinearity-model}
+\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
+\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
+\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
+\end{align*}
+
+\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
+\(\boldsymbol{\varepsilon}\), \(\boldsymbol{x}_{raw}\),
+\(\boldsymbol{z}_{raw}\) are vectors of size \(n\), \(He_{j}(.)\) is the
+\(j\)th-order probabilist's Hermite polynomials,
+\(\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\), and
+\(g(\boldsymbol{x}, k)\) is a scaling function to enforce the support of
+the random vector to be \([-k, k]^n\) defined as
+
+\begin{equation} \label{eq:scaling-function}
+g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x}))2k - k, \quad \text{for} \quad k > 0. 
+\end{equation}
+
+According to \citet{abramowitz1964handbook}, Hermite polynomials were
+initially defined by \citet{de1820theorie}, but named after
+\citet{hermite1864nouveau} because of the unrecognisable form of
+Laplace's work. The function \texttt{hermite} from the R package
+\texttt{mpoly} \citep{mpoly} is used to simulate
+\(\boldsymbol{z}_{raw}\) to generate Hermite polynomials.
+
+The null regression model used to fit the realizations generated by the
+above model is formulated as
+
+\begin{equation} \label{eq:null-model}
+\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
+\end{equation}
+
+\noindent where
+\(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\).
+Since \(z = O(x^j)\), for \(j > 1\), \(z\) is a higher order term left
+out of the null regression, which will lead to model misspecification.
+
+\begin{figure}[!h]
 
@@
 
-\begin{figure}
+\begin{figure}[!h]
 
@@
 \end{figure}
+
+Visual patterns of non-linearity are simulated using four different
+orders of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)). The
+values of \(j\) were chosen so that distinct shapes of non-linearity
+were included in the residual plot. These include ``U'', ``S'', ``M''
+and ``triple-U'' shape as shown in Figure
+\ref{fig:different-shape-of-herimite}. A greater value of \(j\) will
+result in a curve with more turning points. It is expected that the
+``U'' shape will be the easiest detect, as the shape gets more complex
+it will be harder to perceive in a scatterplot, particularly when there
+is noise. Figure \ref{fig:different-sigma} shows the ``U'' shape for
+different amounts of noise (\(\sigma\)).
+
+Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups
+used in non-linearity detection. This lineup is produced by the
+non-linearity model with \(j = 6\). The data plot location is
+\(2^3 - 4\). All five subjects correctly identify the data plot from
+this lineup.
 
@@
 heteroskedasticity. Visual patterns of heteroskedasticity are simulated
-using three different shapes (\(a\) = -1, 0, 1). (A summary of the
-factors can be found in Table \ref{tab:model-factor-table}.)
+using three different shapes (\(a\) = -1, 0, 1).
 

### 2023-04-23 paper.Rmd
done with section 4.1.3

@@
 
-Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups used in non-linearity detection.
-This lineup is produced by the non-linearity model with $j = 6$.
-The data plot location is $2^3 - 4$.
-All five subjects correctly identify the data plot from this lineup.
+Figure \ref{fig:example-poly-lineup} is one of the non-linearity lineups used in experiment.
+This data plot is produced by the non-linearity model with $j = 6$, and is in location $2^3+1$.
+All five subjects who viewed the lineup chose this plot as most different, thus all detected the data plot.
 
@@
 
-Data collection period II is designed to study the ability of human subjects to detect the appearance of a heteroskedasticity pattern under a simple linear regression model setting:
-
-```{=tex}
-\begin{align} \label{eq:heter-model}
+Data collection period II was designed to study the ability of human subjects to detect the heteroskedasticity under a simple linear regression model setting:
+
+\begin{align*} \label{eq:heter-model}
 \boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
@@
 \boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}), 
-\end{align}
-```
+\end{align*}
 
 \noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vectors of size $n$ and $g(.)$ is the scaling function defined in Equation \ref{eq:scaling-function}.
-
 The null regression model used to fit the realizations generated by the above model is formulated exactly the same as Equation \ref{eq:null-model}.
@@
 
-Since $supp(X) = [-1, 1]$, choosing $a$ to be $-1$, $0$ and $1$ can generate "left-triangle", "butterfly" and "right-triangle" shape as displayed in Figure \ref{fig:different-shape-of-heter}.
+Since $supp(X) = [-1, 1]$, choosing $a$ to be $-1$, $0$ and $1$ can generate "left-triangle", "butterfly" and "right-triangle" shapes as displayed in Figure \ref{fig:different-shape-of-heter}.
 The term $(2 - |a|)$ maintains the magnitude of residuals across different values of $a$.
-
-```{r different-shape-of-heter, fig.height = 2.67, fig.cap = 'Heteroskedasticity forms used in the experiment.
+Figure \ref{fig:different-b} shows the butterfly shape as the ratio parameter ($b$) is changed.
+
+```{r different-shape-of-heter, fig.pos="!h", fig.height = 2.67, fig.cap = 'Heteroskedasticity forms used in the experiment.
 Three different shapes ($a = -1, 0, 1$) are used in the experiment to create left-triangle, "butterfly" and "right-triangle" shapes, respectively.'}
@@
 
-```{r different-b, fig.height = 2, fig.cap = "Five different values of $b$ are used in heteroskedasticity simulation to control the strength of the signal.
+```{r different-b, fig.pos="!h", fig.height = 2, fig.cap = "Five different values of $b$ are used in heteroskedasticity simulation to control the strength of the signal.
 Larger values of $b$ yield a bigger difference in variation, and stus stronger heteroskedasticity signal."}
@@
 An example lineup of this model used in data collection period II is shown in Figure \ref{fig:example-heter-lineup} with $a = -1$.
-The data plot location is $2^4 + 2$.
-Nine out of 11 subjects correctly identify the data plot from this lineup.
+The data plot location is in position $2^4 + 2$.
+Nine out of 11 subjects chose this plot from the lineup.
 
@@
 Sometimes the fitted values are discrete because one or more predictors were measured discretely.
-The distribution may be relatively Gaussian, reflecting a linear combinatio of many predictors, adhering to the Central Limit Theorem.
-It is also common to see a skewed distribution of fitted values, if one or more of the predictors has a skewed distribution.
-This latter problem is usually corrected before modelling using a variable transformation.
-Our simulation assess this by using four different distributions to represent fitted values: (1) uniform, (2) normal, (3) skewed and (4) discrete.
-This is constructed by defining the raw predictor $X_{raw}$ in four corresponding distributions:
-(1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$. We would expect that the best reading of residual plots occurs when the fitted values are uniformly distributed.
-
-Three different sample sizes are used, $n = 50, 100, 300$ across the experiments.
-We would expect considerable variation in the signal strength in the simulated data plots with smaller $n$.
+The distribution may be relatively Gaussian, reflecting a linear combination of many predictors, adhering to the Central Limit Theorem.
+It is also common to see a skewed distribution of fitted values if one or more of the predictors has a skewed distribution.
+This latter problem is usually corrected before modelling, using a variable transformation.
+Our simulation assess this by using four different distributions to represent fitted values, constructed by different sampling of the raw predictor $X_{raw}$:
+
+- uniform, $U(-1, 1)$, 
+- normal, $N(0, 0.3^2)$, 
+- skewed, $lognormal(0, 0.6^2)/3$, and 
+- discrete, $U\{1, 5\}$.
+
+\noindent We would expect that structure in residual plots would be easier to perceive when the fitted values are uniformly distributed.
+
+Three different sample sizes were used in our experiment: $n = 50, 100, 300$.
+We expect considerable variation in the signal strength in the simulated data plots with smaller $n$.
 A sample size of 300 is typically enough for structure to be visible in a scatter plot reliably.
+Beyond 300, structure can be obscured by over-plotting, and a different type of display that included transparency, binning or density mapping, might be needed.
 

### 2023-04-23 paper.tex
done with section 4.1.3

@@
 
-Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups
-used in non-linearity detection. This lineup is produced by the
-non-linearity model with \(j = 6\). The data plot location is
-\(2^3 - 4\). All five subjects correctly identify the data plot from
-this lineup.
+Figure \ref{fig:example-poly-lineup} is one of the non-linearity lineups
+used in experiment. This data plot is produced by the non-linearity
+model with \(j = 6\), and is in location \(2^3+1\). All five subjects
+who viewed the lineup chose this plot as most different, thus all
+detected the data plot.
 
@@
 
-Data collection period II is designed to study the ability of human
-subjects to detect the appearance of a heteroskedasticity pattern under
-a simple linear regression model setting:
-
-\begin{align} \label{eq:heter-model}
+Data collection period II was designed to study the ability of human
+subjects to detect the heteroskedasticity under a simple linear
+regression model setting:
+
+\begin{align*} \label{eq:heter-model}
 \boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
@@
 \boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}), 
-\end{align}
+\end{align*}
 
@@
 \(\boldsymbol{\varepsilon}\) are vectors of size \(n\) and \(g(.)\) is
-the scaling function defined in Equation \ref{eq:scaling-function}.
-
-The null regression model used to fit the realizations generated by the
+the scaling function defined in Equation \ref{eq:scaling-function}. The
+null regression model used to fit the realizations generated by the
 above model is formulated exactly the same as Equation
@@
 \(1\) can generate ``left-triangle'', ``butterfly'' and
-``right-triangle'' shape as displayed in Figure
+``right-triangle'' shapes as displayed in Figure
 \ref{fig:different-shape-of-heter}. The term \((2 - |a|)\) maintains the
-magnitude of residuals across different values of \(a\).
-
-\begin{figure}
+magnitude of residuals across different values of \(a\). Figure
+\ref{fig:different-b} shows the butterfly shape as the ratio parameter
+(\(b\)) is changed.
+
+\begin{figure}[!h]
 
@@
 shown in Figure \ref{fig:example-heter-lineup} with \(a = -1\). The data
-plot location is \(2^4 + 2\). Nine out of 11 subjects correctly identify
-the data plot from this lineup.
+plot location is in position \(2^4 + 2\). Nine out of 11 subjects chose
+this plot from the lineup.
 
@@
 discrete because one or more predictors were measured discretely. The
-distribution may be relatively Gaussian, reflecting a linear combinatio
+distribution may be relatively Gaussian, reflecting a linear combination
 of many predictors, adhering to the Central Limit Theorem. It is also
-common to see a skewed distribution of fitted values, if one or more of
+common to see a skewed distribution of fitted values if one or more of
 the predictors has a skewed distribution. This latter problem is usually
-corrected before modelling using a variable transformation. Our
+corrected before modelling, using a variable transformation. Our
 simulation assess this by using four different distributions to
-represent fitted values: (1) uniform, (2) normal, (3) skewed and (4)
-discrete. This is constructed by defining the raw predictor \(X_{raw}\)
-in four corresponding distributions: (1) \(U(-1, 1)\), (2)
-\(N(0, 0.3^2)\), (3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). We
-would expect that the best reading of residual plots occurs when the
-fitted values are uniformly distributed.
-
-Three different sample sizes are used, \(n = 50, 100, 300\) across the
-experiments. We would expect considerable variation in the signal
+represent fitted values, constructed by different sampling of the raw
+predictor \(X_{raw}\):
+
+\begin{itemize}
+\tightlist
+\item
+  uniform, \(U(-1, 1)\),
+\item
+  normal, \(N(0, 0.3^2)\),
+\item
+  skewed, \(lognormal(0, 0.6^2)/3\), and
+\item
+  discrete, \(U\{1, 5\}\).
+\end{itemize}
+
+\noindent We would expect that structure in residual plots would be
+easier to perceive when the fitted values are uniformly distributed.
+
+Three different sample sizes were used in our experiment:
+\(n = 50, 100, 300\). We expect considerable variation in the signal
 strength in the simulated data plots with smaller \(n\). A sample size
 of 300 is typically enough for structure to be visible in a scatter plot
-reliably.
+reliably. Beyond 300, structure can be obscured by over-plotting, and a
+different type of display that included transparency, binning or density
+mapping, might be needed.
 

### 2023-04-23 paper.Rmd
revised 4.1.3 more, in light of big changes needed in 4.2

@@
 
-\noindent We would expect that structure in residual plots would be easier to perceive when the fitted values are uniformly distributed.
+\noindent Figure \ref{fig:different-dist} shows the non-linear pattern, a "U" shape, with the different fitted value distributions.
+We would expect that structure in residual plots would be easier to perceive when the fitted values are uniformly distributed.
 
 Three different sample sizes were used in our experiment: $n = 50, 100, 300$.
-We expect considerable variation in the signal strength in the simulated data plots with smaller $n$.
+Figure \ref{fig:different-n} shows the non-linear "S" shape for different sample sizes.
+We expect signal strength to decline in the simulated data plots with smaller $n$.
 A sample size of 300 is typically enough for structure to be visible in a scatter plot reliably.

### 2023-04-23 paper.tex
revised 4.1.3 more, in light of big changes needed in 4.2

@@
 
-\noindent We would expect that structure in residual plots would be
-easier to perceive when the fitted values are uniformly distributed.
+\noindent Figure \ref{fig:different-dist} shows the non-linear pattern,
+a ``U'' shape, with the different fitted value distributions. We would
+expect that structure in residual plots would be easier to perceive when
+the fitted values are uniformly distributed.
 
 Three different sample sizes were used in our experiment:
-\(n = 50, 100, 300\). We expect considerable variation in the signal
-strength in the simulated data plots with smaller \(n\). A sample size
-of 300 is typically enough for structure to be visible in a scatter plot
+\(n = 50, 100, 300\). Figure \ref{fig:different-n} shows the non-linear
+``S'' shape for different sample sizes. We expect signal strength to
+decline in the simulated data plots with smaller \(n\). A sample size of
+300 is typically enough for structure to be visible in a scatter plot
 reliably. Beyond 300, structure can be obscured by over-plotting, and a

### 2023-04-23 appendix.Rmd
moved signal strength into expt design section and swapped participant to subject

--- 2023-04-16 appendix.Rmd
@@
 
-## Mapping of subjects to experimental factors
-
-Mapping of subjects to experimental factors is an important part of experiment design.
-Essentially, we want to maximum the difference in factors exposed to a subject.
-For this purpose, we design an algorithm to conduct subject allocation.
-Let $L$ be a set of available lineups and $S$ be a set of available subjects.
-According to the experimental design, the availability of a lineup is associated with the number of subjects it can assign to.
+## Mapping of participants to experimental factors
+
+Mapping of participants to experimental factors is an important part of experiment design.
+Essentially, we want to maximum the difference in factors exposed to a participant.
+For this purpose, we design an algorithm to conduct participant allocation.
+Let $L$ be a set of available lineups and $S$ be a set of available participants.
+According to the experimental design, the availability of a lineup is associated with the number of participants it can assign to.
 For lineups with uniform fitted value distribution, this value is 11.
-And other lineups can be allocated to at most five different subjects.
-The availability of a subject is associated with the number of lineups that being allocated to this subject.
-A subject can view at most 18 different lineups.
-
-The algorithm starts from picking a random subject $s \in S$ with the minimum number of allocated lineups.
-It then tries to find a lineup $l \in L$ that can maximise the distance metric $D$ and allocate it to subject $s$.
-Set $L$ and $S$ will be updated and the picking process will be repeated until there is no available lineups or subjects.
+And other lineups can be allocated to at most five different participants.
+The availability of a participant is associated with the number of lineups that being allocated to this participant.
+A participant can view at most 18 different lineups.
+
+The algorithm starts from picking a random participant $s \in S$ with the minimum number of allocated lineups.
+It then tries to find a lineup $l \in L$ that can maximise the distance metric $D$ and allocate it to participant $s$.
+Set $L$ and $S$ will be updated and the picking process will be repeated until there is no available lineups or participants.
 
@@
 And $f_if_jf_k$ exists in $L_{s}$ if any lineup in $L_{s}$ has this trio of factor values.
-The distance metric $D$ is defined between a lineup $l$ and a set of lineups $L_{s}$ allocated to a subject $s$ if $L_{s}$ is non-empty:
+The distance metric $D$ is defined between a lineup $l$ and a set of lineups $L_{s}$ allocated to a participant $s$ if $L_{s}$ is non-empty:
 
@@
 
-The distance measures how different a lineup is from the set of lineups allocated to the subject in terms of factor values.
-Thus, the algorithm will try to allocate the most different lineup to a subject at each step.
+The distance measures how different a lineup is from the set of lineups allocated to the participant in terms of factor values.
+Thus, the algorithm will try to allocate the most different lineup to a participant at each step.
 
@@
 The website is built with the `Flask` [@flask] web framework and hosted on `PythonAnywhere` [@pythonanywhere].
-It is configured to handle HTTP requests such that subjects can correctly receive webpages and submit responses.
-Embedded in the resources sent to subjects, the `jsPsych` front-end framework [@jspsych] instructs subjects' browsers to render an environment for running behavioral experiments.
+It is configured to handle HTTP requests such that participants can correctly receive webpages and submit responses.
+Embedded in the resources sent to participants, the `jsPsych` front-end framework [@jspsych] instructs participants' browsers to render an environment for running behavioral experiments.
 During the experiment, this framework will automatically collect common behavioral data such as response time and clicks on buttons.
-Subjects' responses are first validated by a scheduled `Python` script run on the server, then push to a Github repository.
+participants' responses are first validated by a scheduled `Python` script run on the server, then push to a Github repository.
 Lineup images shown to users are saved in multiple Github repositories and hosted in corresponding Github pages.
@@
                     label = "pronoun", 
-                    caption = "Summary of pronoun distribution of subjects recuritted in this study.") %>%
+                    caption = "Summary of pronoun distribution of participants recruited in this study.") %>%
   kableExtra::kable_styling(latex_options = "scale_down")
@@
                     label = "age-group", 
-                    caption = "Summary of age distribution of subjects recuritted in this study.") %>%
+                    caption = "Summary of age distribution of participants recruited in this study.") %>%
   kableExtra::kable_styling(latex_options = "scale_down")
@@
                     label = "education", 
-                    caption = "Summary of education distribution of subjects recuritted in this study.") %>%
+                    caption = "Summary of education distribution of participants recruited in this study.") %>%
   kableExtra::kable_styling(latex_options = "scale_down")
@@
                     label = "experience", 
-                    caption = "Summary of previous experience distribution of subjects recuritted in this study.") %>%
+                    caption = "Summary of previous experience distribution of participants recruited in this study.") %>%
   kableExtra::kable_styling(latex_options = "scale_down")

### 2023-04-23 appendix.tex
moved signal strength into expt design section and swapped participant to subject

@@
 
-\hypertarget{mapping-of-subjects-to-experimental-factors}{%
-\subsection{Mapping of subjects to experimental
-factors}\label{mapping-of-subjects-to-experimental-factors}}
+\hypertarget{mapping-of-participants-to-experimental-factors}{%
+\subsection{Mapping of participants to experimental
+factors}\label{mapping-of-participants-to-experimental-factors}}
 
-Mapping of subjects to experimental factors is an important part of
+Mapping of participants to experimental factors is an important part of
 experiment design. Essentially, we want to maximum the difference in
-factors exposed to a subject. For this purpose, we design an algorithm
-to conduct subject allocation. Let \(L\) be a set of available lineups
-and \(S\) be a set of available subjects. According to the experimental
-design, the availability of a lineup is associated with the number of
-subjects it can assign to. For lineups with uniform fitted value
-distribution, this value is 11. And other lineups can be allocated to at
-most five different subjects. The availability of a subject is
-associated with the number of lineups that being allocated to this
-subject. A subject can view at most 18 different lineups.
+factors exposed to a participant. For this purpose, we design an
+algorithm to conduct participant allocation. Let \(L\) be a set of
+available lineups and \(S\) be a set of available participants.
+According to the experimental design, the availability of a lineup is
+associated with the number of participants it can assign to. For lineups
+with uniform fitted value distribution, this value is 11. And other
+lineups can be allocated to at most five different participants. The
+availability of a participant is associated with the number of lineups
+that being allocated to this participant. A participant can view at most
+18 different lineups.
 
-The algorithm starts from picking a random subject \(s \in S\) with the
-minimum number of allocated lineups. It then tries to find a lineup
+The algorithm starts from picking a random participant \(s \in S\) with
+the minimum number of allocated lineups. It then tries to find a lineup
 \(l \in L\) that can maximise the distance metric \(D\) and allocate it
-to subject \(s\). Set \(L\) and \(S\) will be updated and the picking
-process will be repeated until there is no available lineups or
-subjects.
+to participant \(s\). Set \(L\) and \(S\) will be updated and the
+picking process will be repeated until there is no available lineups or
+participants.
 
@@
 between a lineup \(l\) and a set of lineups \(L_{s}\) allocated to a
-subject \(s\) if \(L_{s}\) is non-empty:
+participant \(s\) if \(L_{s}\) is non-empty:
 
@@
 The distance measures how different a lineup is from the set of lineups
-allocated to the subject in terms of factor values. Thus, the algorithm
-will try to allocate the most different lineup to a subject at each
-step.
+allocated to the participant in terms of factor values. Thus, the
+algorithm will try to allocate the most different lineup to a
+participant at each step.
 
@@
 hosted on \texttt{PythonAnywhere} \citep{pythonanywhere}. It is
-configured to handle HTTP requests such that subjects can correctly
+configured to handle HTTP requests such that participants can correctly
 receive webpages and submit responses. Embedded in the resources sent to
-subjects, the \texttt{jsPsych} front-end framework \citep{jspsych}
-instructs subjects' browsers to render an environment for running
+participants, the \texttt{jsPsych} front-end framework \citep{jspsych}
+instructs participants' browsers to render an environment for running
 behavioral experiments. During the experiment, this framework will
 automatically collect common behavioral data such as response time and
-clicks on buttons. Subjects' responses are first validated by a
+clicks on buttons. participants' responses are first validated by a
 scheduled \texttt{Python} script run on the server, then push to a
@@
 
-\caption{\label{tab:pronoun}Summary of pronoun distribution of subjects recuritted in this study.}
+\caption{\label{tab:pronoun}Summary of pronoun distribution of participants recruited in this study.}
 \centering
@@
 
-\caption{\label{tab:age-group}Summary of age distribution of subjects recuritted in this study.}
+\caption{\label{tab:age-group}Summary of age distribution of participants recruited in this study.}
 \centering
@@
 
-\caption{\label{tab:education}Summary of education distribution of subjects recuritted in this study.}
+\caption{\label{tab:education}Summary of education distribution of participants recruited in this study.}
 \centering
@@
 
-\caption{\label{tab:experience}Summary of previous experience distribution of subjects recuritted in this study.}
+\caption{\label{tab:experience}Summary of previous experience distribution of participants recruited in this study.}
 \centering

### 2023-04-23 paper.Rmd
moved signal strength into expt design section and swapped participant to subject

@@
 
-Within the context of visual inference, with $K$ independent observers, the visual $p$-value can be seen as the probability of having as many or more subjects detect the data plot than the observed result.
+Within the context of visual inference, with $K$ independent observers, the visual $p$-value can be seen as the probability of having as many or more participants detect the data plot than the observed result.
 
 The approach used in @majumder_validation_2013 is as follows.
-Define $X_j = \{0,1\}$ to be a Bernoulli random variable measuring whether subject $j$ detected the data plot, and $X = \sum_{j=1}^{K}X_j$ be the total number of observers who detected the data plot.
+Define $X_j = \{0,1\}$ to be a Bernoulli random variable measuring whether participant $j$ detected the data plot, and $X = \sum_{j=1}^{K}X_j$ be the total number of observers who detected the data plot.
 Then, by imposing a relatively strong assumption that all $K$ evaluations are fully independent, under $H_0$ $X \sim \mathrm{Binom}_{K,1/m}$.
@@
 As pointed out by @vanderplas2021statistical, this basic binomial model is deficient.
-It does not take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup, or account for when subjects are offered the option to select one or more "most different" plots, or none, from a lineup.
-They suggest three common lineup scenarios: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects.
+It does not take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup, or account for when participants are offered the option to select one or more "most different" plots, or none, from a lineup.
+They suggest three common lineup scenarios: (1) $K$ different lineups are shown to $K$ participants, (2) $K$ lineups with different null plots but the same data plot are shown to $K$ participants, and (3) the same lineup is shown to $K$ participants.
 Scenario 3 is the most feasible to apply, but has the most dependencies to accommodate for the $p$-value calculation.
@@
 The number of times plot $i$ being selected in $K$ evaluations is denoted as $c_i$.
-In case subject $j$ makes multiple selections, $1/s_j$ will be added to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$.
+In case participant $j$ makes multiple selections, $1/s_j$ will be added to $c_i$ instead of one, where $s_j$ is the number of plots participant $j$ selected for $j=1,...K$.
 This ensures $\sum_{i}c_i=K$.
@@
 
-The power of a visual test on the other hand, may additionally depend on the ability of the particular subject, as the skill of each individual may affect the number of observers who identify the data plot from the lineup [@majumder_validation_2013].
-To address this issue, @majumder_validation_2013 models the probability of subject $j$ correctly picking the data plot from lineup $l$ using a mixed-effect logistic regression, with subjects treated as random effect.
-Then, the estimated power of a visual test evaluated by a single subject is the predicted value obtained from the mixed effects model.
-However, this mixed effects model does not work with scenario where subjects are asked to select one or more most different plots.
-In this scenario, having the probability of a subject $j$ correctly picking the data plot from a lineup $l$ is insufficient to determine the power of a visual test because it does not provide information about the number of selections made by the subject for the calculation of the $p$-value (See Equation \ref{eq:pvalue-beta-binomial-approx}).
+The power of a visual test on the other hand, may additionally depend on the ability of the particular participant, as the skill of each individual may affect the number of observers who identify the data plot from the lineup [@majumder_validation_2013].
+To address this issue, @majumder_validation_2013 models the probability of participant $j$ correctly picking the data plot from lineup $l$ using a mixed-effect logistic regression, with participants treated as random effect.
+Then, the estimated power of a visual test evaluated by a single participant is the predicted value obtained from the mixed effects model.
+However, this mixed effects model does not work with scenario where participants are asked to select one or more most different plots.
+In this scenario, having the probability of a participant $j$ correctly picking the data plot from a lineup $l$ is insufficient to determine the power of a visual test because it does not provide information about the number of selections made by the participant for the calculation of the $p$-value (See Equation \ref{eq:pvalue-beta-binomial-approx}).
 Therefore, we directly estimate the probability of a lineup being rejected by assuming that individual skill has negligible effect on the variation of the power.
@@
 Additional lineups for both non-linearity and heteroskedasticity, using uniform fitted value distributions, were included for additional data, and to avoid participant frustration of too many difficult tasks.
-Overall, we collected `r vi_survey  %>% filter(!attention_check) %>% nrow()` evaluations on `r vi_survey  %>% filter(!attention_check) %>% pull(unique_lineup_id) %>% unique() %>% length()` unique lineups performed by `r vi_survey  %>% filter(!attention_check) %>% count(exp, set) %>% nrow()` subjects throughout three data collection periods.
+Overall, we collected `r vi_survey  %>% filter(!attention_check) %>% nrow()` evaluations on `r vi_survey  %>% filter(!attention_check) %>% pull(unique_lineup_id) %>% unique() %>% length()` unique lineups performed by `r vi_survey  %>% filter(!attention_check) %>% count(exp, set) %>% nrow()` participants throughout three data collection periods.
 
@@
 
-Data collection period I was designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials on random vector $\boldsymbol{x}$ formulated as
+Data collection period I was designed to study the ability of participants to detect the effect of a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials on random vector $\boldsymbol{x}$ formulated as
 
@@
 This data plot is produced by the non-linearity model with $j = 6$, and is in location $2^3+1$.
-All five subjects who viewed the lineup chose this plot as most different, thus all detected the data plot.
+All five participants who viewed the lineup chose this plot as most different, thus all detected the data plot.
 
@@
 
-Data collection period II was designed to study the ability of human subjects to detect the heteroskedasticity under a simple linear regression model setting:
+Data collection period II was designed to study the ability of participants to detect the heteroskedasticity under a simple linear regression model setting:
 
@@
 Can you spot the most different plot?
-The data plot is positioned at $3^3 - 3^2$"}
+The data plot is positioned at $3^3 - 3^2$."}
 
@@
 The data plot location is in position $2^4 + 2$.
-Nine out of 11 subjects chose this plot from the lineup.
+Nine out of 11 participants chose this plot from the lineup.
 
@@
 
+### Controlling the signal strength of the signal
+
+The three parameters $n$, $\sigma$ and $b$ are important for controlling the strength of the signal to generate lineups with a variety difficulty levels.
+This will ensure that  estimated power curves will be smooth and continuous, and that participants are allocated a set of lineups with a range of difficulty.
+
+Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ are used in data collection periods I and II respectively.
+A large value of $\sigma$ will increase the variation of the error of the non-linearity model and decrease the visibility of the visual pattern.
+The parameter $b$ controls the variation in the standard deviation of the error across the support of the predictor.
+A larger value of $b$ generates a larger ratio between it's smallest and highest values, making the visual pattern more obvious.
+The sample size $n$ sharpens (or blurs) a pattern when it is larger (or smaller).
+Figures \ref{fig:different-sigma}, \ref{fig:different-b} and \ref{fig:different-n} demonstrate the impact of these parameters on signal strength.
+
 ## Experimental setup
 
-### Controlling the strength of the signal
-
-As summarised in Table \ref{tab:model-factor-table}, three additional parameters $n$, $\sigma$ and $b$ are used to control the strength of the signal so that different difficulty levels of lineups are generated, and therefore, the estimated power curve will be smooth and continuous.
-Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ are used in data collection periods I and II respectively.
-Figure \ref{fig:different-sigma} and \ref{fig:different-b} demonstrate the impact of these two parameters.
-A large value of $\sigma$ will increase the variation of the error of the non-linearity model and decrease the visibility of the visual pattern.
-The parameter $b$ controls the standard deviation of the error across the support of the predictor.
-Given $x \neq a$, a larger value of $b$ will lead to a larger ratio of the variance at $x$ to the variance at $x - a = 0$, making the visual pattern more obvious.
-
-Three different sample sizes are used (n = 50, 100, 300) in all three data collection periods.
-It can be observed from Figure \ref{fig:different-n} that with fewer data points drawn in a residual plot, the visual pattern is more difficult to be detected.
+The lineups are allocated to participants in a manner that covers the experimental design and is relatively uniform.
+To manage this, we use effect size to measure the signal strength, which helps in assigning a set of lineups with a range of difficulties to each participant.
 
@@
 
-### Subject allocation
+### Lineup allocation to participants
 
@@
 Three replications are made for each of the combination results in $192 \times 3 = 576$ and $180 \times 3 = 540$ lineups.
-In addition, each lineup is designed to be evaluated by five different subjects.
-After attempting some pilot studies internally, we decide to present a block of 20 lineups to every subject.
+In addition, each lineup is designed to be evaluated by five different participants.
+After attempting some pilot studies internally, we decide to present a block of 20 lineups to every participant.
 And to ensure the quality of the survey data, two lineups with obvious visual patterns are included as attention checks.
-Thus, $576 \times 5 / (20-2) = 160$ and $540 \times 5 / (20-2) = 150$ subjects are recruited to satisfy the design of the data collection period I and II respectively.
+Thus, $576 \times 5 / (20-2) = 160$ and $540 \times 5 / (20-2) = 150$ participants are recruited to satisfy the design of the data collection period I and II respectively.
 
@@
 
-All lineups consist of only null plots are planned to be evaluated by 20 subjects. 
-However, presenting only these lineups to subjects are considered to be bad practices as subjects will lose interest quickly. 
-Therefore, we plan to collect 6 more evaluations on the 279 lineups with uniform fitted value distribution, result in $(36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133$ subjects recruited for data collection period III.
+All lineups consist of only null plots are planned to be evaluated by 20 participants. 
+However, presenting only these lineups to participants are considered to be bad practices as participants will lose interest quickly. 
+Therefore, we plan to collect 6 more evaluations on the 279 lineups with uniform fitted value distribution, result in $(36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133$ participants recruited for data collection period III.
 
@@
 
-Subjects for all three data collection periods are recruited from an crowdsourcing platform called Prolific [@palan2018prolific].
-Prescreening procedure is applied during the recruitment, subjects are required to be fluent in English, with $98\%$ minimum approval rate and 10 minimum submissions in other studies.
-
-During the experiment, every subject is presented with a block of 20 lineups.
+Participants for all three data collection periods are recruited from an crowdsourcing platform called Prolific [@palan2018prolific].
+Prescreening procedure is applied during the recruitment, participants are required to be fluent in English, with $98\%$ minimum approval rate and 10 minimum submissions in other studies.
+
+During the experiment, every participant is presented with a block of 20 lineups.
 A lineup consists of a randomly placed data plot and 19 null plots, which are all residual plots drawn with raw residuals on the y-axis and fitted values on the x-axis.
@@
 
-In every lineup evaluation, the subject is asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots are from others.
-If there is no noticeable difference between plots in a lineup, subjects are permitted to select zero plots without providing the reason.
-No subject are shown the same lineup twice.
+In every lineup evaluation, the participant is asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots are from others.
+If there is no noticeable difference between plots in a lineup, participants are permitted to select zero plots without providing the reason.
+No participant are shown the same lineup twice.
 Information about preferred pronoun, age group, education, and previous experience in visual experiments are also collected.
-A subject's submission is only accepted if the data plot is identified for at least one attention check.
+A participant's submission is only accepted if the data plot is identified for at least one attention check.
 Data of rejected submissions are discarded automatically to maintain the overall data quality.
@@
 But the BP test $p\text{-value} = 0.056$, is slightly above the significance cutoff of $0.05$.
-This lineup was evaluated by 11 subjects, it has experimental factors $a = 0$ ("butterfly" shape), $b = 64$ (large variance ratio), $n = 50$ (small sample size), and a uniform distribution for the fitted values.
+This lineup was evaluated by 11 participants, it has experimental factors $a = 0$ ("butterfly" shape), $b = 64$ (large variance ratio), $n = 50$ (small sample size), and a uniform distribution for the fitted values.
 It may have been the small sample size and the presence of a few outliers that may have resulted in the lack of detection by the conventional test.
@@
 A residual plot may contain many visual features, but some are caused by the characteristics of the predictors and the randomness of the error, not by the violation of the model assumptions.
-These irrelevant visual features have a chance to be filtered out by subjects with a comparison to null plots, resulting in more accurate reading.
+These irrelevant visual features have a chance to be filtered out by participants with a comparison to null plots, resulting in more accurate reading.
 The lineup enables a careful calibration for reading structure in residual plots.

### 2023-04-23 paper.tex
moved signal strength into expt design section and swapped participant to subject

@@
 observers, the visual \(p\)-value can be seen as the probability of
-having as many or more subjects detect the data plot than the observed
-result.
+having as many or more participants detect the data plot than the
+observed result.
 
@@
 Define \(X_j = \{0,1\}\) to be a Bernoulli random variable measuring
-whether subject \(j\) detected the data plot, and
+whether participant \(j\) detected the data plot, and
 \(X = \sum_{j=1}^{K}X_j\) be the total number of observers who detected
@@
 dependencies in the visual test due to repeated evaluations of the same
-lineup, or account for when subjects are offered the option to select
-one or more ``most different'' plots, or none, from a lineup. They
-suggest three common lineup scenarios: (1) \(K\) different lineups are
-shown to \(K\) subjects, (2) \(K\) lineups with different null plots but
-the same data plot are shown to \(K\) subjects, and (3) the same lineup
-is shown to \(K\) subjects. Scenario 3 is the most feasible to apply,
-but has the most dependencies to accommodate for the \(p\)-value
-calculation. For Scenario 3, VanderPlas et al propose modeling the
-probability of plot \(i\) being selected from a lineup as \(\theta_i\),
-where \(\theta_i \sim Dirichlet(\alpha)\) for \(i=1,...,m\) and
-\(\alpha > 0\). The number of times plot \(i\) being selected in \(K\)
-evaluations is denoted as \(c_i\). In case subject \(j\) makes multiple
-selections, \(1/s_j\) will be added to \(c_i\) instead of one, where
-\(s_j\) is the number of plots subject \(j\) selected for \(j=1,...K\).
-This ensures \(\sum_{i}c_i=K\). Since we are only interested in the
-selections of the data plot \(i\), the marginal model can be simplified
-to a beta-binomial model and thus the visual \(p\)-value is given as
+lineup, or account for when participants are offered the option to
+select one or more ``most different'' plots, or none, from a lineup.
+They suggest three common lineup scenarios: (1) \(K\) different lineups
+are shown to \(K\) participants, (2) \(K\) lineups with different null
+plots but the same data plot are shown to \(K\) participants, and (3)
+the same lineup is shown to \(K\) participants. Scenario 3 is the most
+feasible to apply, but has the most dependencies to accommodate for the
+\(p\)-value calculation. For Scenario 3, VanderPlas et al propose
+modeling the probability of plot \(i\) being selected from a lineup as
+\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
+\(i=1,...,m\) and \(\alpha > 0\). The number of times plot \(i\) being
+selected in \(K\) evaluations is denoted as \(c_i\). In case participant
+\(j\) makes multiple selections, \(1/s_j\) will be added to \(c_i\)
+instead of one, where \(s_j\) is the number of plots participant \(j\)
+selected for \(j=1,...K\). This ensures \(\sum_{i}c_i=K\). Since we are
+only interested in the selections of the data plot \(i\), the marginal
+model can be simplified to a beta-binomial model and thus the visual
+\(p\)-value is given as
 
@@
 The power of a visual test on the other hand, may additionally depend on
-the ability of the particular subject, as the skill of each individual
-may affect the number of observers who identify the data plot from the
-lineup \citep{majumder_validation_2013}. To address this issue,
-\citet{majumder_validation_2013} models the probability of subject \(j\)
-correctly picking the data plot from lineup \(l\) using a mixed-effect
-logistic regression, with subjects treated as random effect. Then, the
-estimated power of a visual test evaluated by a single subject is the
-predicted value obtained from the mixed effects model. However, this
-mixed effects model does not work with scenario where subjects are asked
-to select one or more most different plots. In this scenario, having the
-probability of a subject \(j\) correctly picking the data plot from a
-lineup \(l\) is insufficient to determine the power of a visual test
-because it does not provide information about the number of selections
-made by the subject for the calculation of the \(p\)-value (See Equation
+the ability of the particular participant, as the skill of each
+individual may affect the number of observers who identify the data plot
+from the lineup \citep{majumder_validation_2013}. To address this issue,
+\citet{majumder_validation_2013} models the probability of participant
+\(j\) correctly picking the data plot from lineup \(l\) using a
+mixed-effect logistic regression, with participants treated as random
+effect. Then, the estimated power of a visual test evaluated by a single
+participant is the predicted value obtained from the mixed effects
+model. However, this mixed effects model does not work with scenario
+where participants are asked to select one or more most different plots.
+In this scenario, having the probability of a participant \(j\)
+correctly picking the data plot from a lineup \(l\) is insufficient to
+determine the power of a visual test because it does not provide
+information about the number of selections made by the participant for
+the calculation of the \(p\)-value (See Equation
 \ref{eq:pvalue-beta-binomial-approx}). Therefore, we directly estimate
@@
 collected 7974 evaluations on 1152 unique lineups performed by 443
-subjects throughout three data collection periods.
+participants throughout three data collection periods.
 
@@
 
-Data collection period I was designed to study the ability of human
-subjects to detect the effect of a non-linear term \(\boldsymbol{z}\)
-constructed using Hermite polynomials on random vector
-\(\boldsymbol{x}\) formulated as
+Data collection period I was designed to study the ability of
+participants to detect the effect of a non-linear term
+\(\boldsymbol{z}\) constructed using Hermite polynomials on random
+vector \(\boldsymbol{x}\) formulated as
 
@@
 used in experiment. This data plot is produced by the non-linearity
-model with \(j = 6\), and is in location \(2^3+1\). All five subjects
-who viewed the lineup chose this plot as most different, thus all
-detected the data plot.
+model with \(j = 6\), and is in location \(2^3+1\). All five
+participants who viewed the lineup chose this plot as most different,
+thus all detected the data plot.
 
@@
 
-Data collection period II was designed to study the ability of human
-subjects to detect the heteroskedasticity under a simple linear
+Data collection period II was designed to study the ability of
+participants to detect the heteroskedasticity under a simple linear
 regression model setting:
@@
 Can you spot the most different plot?
-The data plot is positioned at $3^3 - 3^2$}\label{fig:example-heter-lineup}
+The data plot is positioned at $3^3 - 3^2$.}\label{fig:example-heter-lineup}
 \end{figure}
@@
 shown in Figure \ref{fig:example-heter-lineup} with \(a = -1\). The data
-plot location is in position \(2^4 + 2\). Nine out of 11 subjects chose
-this plot from the lineup.
+plot location is in position \(2^4 + 2\). Nine out of 11 participants
+chose this plot from the lineup.
 
@@
 
+\hypertarget{controlling-the-signal-strength-of-the-signal}{%
+\subsubsection{Controlling the signal strength of the
+signal}\label{controlling-the-signal-strength-of-the-signal}}
+
+The three parameters \(n\), \(\sigma\) and \(b\) are important for
+controlling the strength of the signal to generate lineups with a
+variety difficulty levels. This will ensure that estimated power curves
+will be smooth and continuous, and that participants are allocated a set
+of lineups with a range of difficulty.
+
+Parameter \(\sigma \in \{0.5, 1, 2, 4\}\) and
+\(b \in \{0.25, 1, 4, 16, 64\}\) are used in data collection periods I
+and II respectively. A large value of \(\sigma\) will increase the
+variation of the error of the non-linearity model and decrease the
+visibility of the visual pattern. The parameter \(b\) controls the
+variation in the standard deviation of the error across the support of
+the predictor. A larger value of \(b\) generates a larger ratio between
+it's smallest and highest values, making the visual pattern more
+obvious. The sample size \(n\) sharpens (or blurs) a pattern when it is
+larger (or smaller). Figures \ref{fig:different-sigma},
+\ref{fig:different-b} and \ref{fig:different-n} demonstrate the impact
+of these parameters on signal strength.
+
 \hypertarget{experimental-setup}{%
@@
 
-\hypertarget{controlling-the-strength-of-the-signal}{%
-\subsubsection{Controlling the strength of the
-signal}\label{controlling-the-strength-of-the-signal}}
-
-As summarised in Table \ref{tab:model-factor-table}, three additional
-parameters \(n\), \(\sigma\) and \(b\) are used to control the strength
-of the signal so that different difficulty levels of lineups are
-generated, and therefore, the estimated power curve will be smooth and
-continuous. Parameter \(\sigma \in \{0.5, 1, 2, 4\}\) and
-\(b \in \{0.25, 1, 4, 16, 64\}\) are used in data collection periods I
-and II respectively. Figure \ref{fig:different-sigma} and
-\ref{fig:different-b} demonstrate the impact of these two parameters. A
-large value of \(\sigma\) will increase the variation of the error of
-the non-linearity model and decrease the visibility of the visual
-pattern. The parameter \(b\) controls the standard deviation of the
-error across the support of the predictor. Given \(x \neq a\), a larger
-value of \(b\) will lead to a larger ratio of the variance at \(x\) to
-the variance at \(x - a = 0\), making the visual pattern more obvious.
-
-Three different sample sizes are used (n = 50, 100, 300) in all three
-data collection periods. It can be observed from Figure
-\ref{fig:different-n} that with fewer data points drawn in a residual
-plot, the visual pattern is more difficult to be detected.
+The lineups are allocated to participants in a manner that covers the
+experimental design and is relatively uniform. To manage this, we use
+effect size to measure the signal strength, which helps in assigning a
+set of lineups with a range of difficulties to each participant.
 
@@
 
-\hypertarget{subject-allocation}{%
-\subsubsection{Subject allocation}\label{subject-allocation}}
+\hypertarget{lineup-allocation-to-participants}{%
+\subsubsection{Lineup allocation to
+participants}\label{lineup-allocation-to-participants}}
 
@@
 addition, each lineup is designed to be evaluated by five different
-subjects. After attempting some pilot studies internally, we decide to
-present a block of 20 lineups to every subject. And to ensure the
+participants. After attempting some pilot studies internally, we decide
+to present a block of 20 lineups to every participant. And to ensure the
 quality of the survey data, two lineups with obvious visual patterns are
 included as attention checks. Thus, \(576 \times 5 / (20-2) = 160\) and
-\(540 \times 5 / (20-2) = 150\) subjects are recruited to satisfy the
-design of the data collection period I and II respectively.
+\(540 \times 5 / (20-2) = 150\) participants are recruited to satisfy
+the design of the data collection period I and II respectively.
 
@@
 All lineups consist of only null plots are planned to be evaluated by 20
-subjects. However, presenting only these lineups to subjects are
-considered to be bad practices as subjects will lose interest quickly.
-Therefore, we plan to collect 6 more evaluations on the 279 lineups with
-uniform fitted value distribution, result in
-\((36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133\) subjects
+participants. However, presenting only these lineups to participants are
+considered to be bad practices as participants will lose interest
+quickly. Therefore, we plan to collect 6 more evaluations on the 279
+lineups with uniform fitted value distribution, result in
+\((36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133\) participants
 recruited for data collection period III.
@@
 
-Subjects for all three data collection periods are recruited from an
+Participants for all three data collection periods are recruited from an
 crowdsourcing platform called Prolific \citep{palan2018prolific}.
-Prescreening procedure is applied during the recruitment, subjects are
-required to be fluent in English, with \(98\%\) minimum approval rate
-and 10 minimum submissions in other studies.
-
-During the experiment, every subject is presented with a block of 20
+Prescreening procedure is applied during the recruitment, participants
+are required to be fluent in English, with \(98\%\) minimum approval
+rate and 10 minimum submissions in other studies.
+
+During the experiment, every participant is presented with a block of 20
 lineups. A lineup consists of a randomly placed data plot and 19 null
@@
 
-In every lineup evaluation, the subject is asked to select one or more
-plots that are most different from others, provide a reason for their
-selections, and evaluate how different they think the selected plots are
-from others. If there is no noticeable difference between plots in a
-lineup, subjects are permitted to select zero plots without providing
-the reason. No subject are shown the same lineup twice. Information
-about preferred pronoun, age group, education, and previous experience
-in visual experiments are also collected. A subject's submission is only
-accepted if the data plot is identified for at least one attention
-check. Data of rejected submissions are discarded automatically to
-maintain the overall data quality.
+In every lineup evaluation, the participant is asked to select one or
+more plots that are most different from others, provide a reason for
+their selections, and evaluate how different they think the selected
+plots are from others. If there is no noticeable difference between
+plots in a lineup, participants are permitted to select zero plots
+without providing the reason. No participant are shown the same lineup
+twice. Information about preferred pronoun, age group, education, and
+previous experience in visual experiments are also collected. A
+participant's submission is only accepted if the data plot is identified
+for at least one attention check. Data of rejected submissions are
+discarded automatically to maintain the overall data quality.
 
@@
 is slightly above the significance cutoff of \(0.05\). This lineup was
-evaluated by 11 subjects, it has experimental factors \(a = 0\)
+evaluated by 11 participants, it has experimental factors \(a = 0\)
 (``butterfly'' shape), \(b = 64\) (large variance ratio), \(n = 50\)
@@
 by the violation of the model assumptions. These irrelevant visual
-features have a chance to be filtered out by subjects with a comparison
-to null plots, resulting in more accurate reading. The lineup enables a
-careful calibration for reading structure in residual plots.
+features have a chance to be filtered out by participants with a
+comparison to null plots, resulting in more accurate reading. The lineup
+enables a careful calibration for reading structure in residual plots.
 

### 2023-04-23 paper.Rmd
working hard on revising the experimental setup section

@@
 These two can be quantified in terms of effect size $E$ to measure the strength of the residual departures from the model assumptions.
-Details about the effect size are provided in Section \ref{effect-size} after the introduction of the simulation model used in our experiment.
+Details about the calculation of effect size are provided in Section \ref{effect-size} after the introduction of the simulation model used in our experiment.
 The theoretical power of a test is sometimes not a trival solution, but it can be estimated if the data generating process is known.
@@
 
-### Controlling the signal strength of the signal
+### Controlling the signal strength
 
@@
 
-Effect size in statistics measures the strength of the signal relative to the noise. It is surprisingly difficult to quantify in general, even for simulated data as used in this experiment.
+Effect size in statistics measures the strength of the signal relative to the noise. It is surprisingly difficult to quantify, even for simulated data as used in this experiment.
 
@@
 
-$$E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}\sigma^2))^{-1}\boldsymbol{\mu}_z\right)$$
+\begin{equation}
+E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}\sigma^2))^{-1}\boldsymbol{\mu}_z\right)
+\label{eq:nonlin-effect-size}
+\end{equation}
+
 \noindent where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,
-$\boldsymbol{R} = \boldsymbol{I}_n - \boldsymbol{H}$ is the residual operator, $\boldsymbol{H} = \boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the hat matrix, $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients, and $\sigma^2\boldsymbol{I}$ is the assumed covariance matrix of the error term when $H_0$ is true.
+$\boldsymbol{R} = \boldsymbol{I}_n - \boldsymbol{H}$ is the residual operator, $\boldsymbol{H} = \boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the hat matrix, $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals where $\boldsymbol{Z}$ contain any higher order terms of $\boldsymbol{X}$ left out of the regression equation, $\boldsymbol{\beta}_z$ contains the corresponding coefficients, and $\sigma^2\boldsymbol{I}$ is the assumed covariance matrix of the error term when $H_0$ is true.
 
@@
 
-$$E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|diag(\boldsymbol{R})|} - n + tr(diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}diag(\boldsymbol{R}))\right)$$
-\noindent where $\boldsymbol{V}$ is the actual covariance matrix of the error term. (Derivations for these equations are provided in the Appendix.)
+\begin{equation}
+E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|diag(\boldsymbol{R})|} - n + tr(diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}diag(\boldsymbol{R}))\right)
+\label{eq:hetero-effect-size}
+\end{equation}
+
+\noindent where $\boldsymbol{V}$ is the actual covariance matrix of the error term. Derivations for these equations are provided in the Appendix.
 
@@
 
-As shown in Table \ref{tab:model-factor-table}, there are a total of $4 \times 4 \times 3 \times 4 = 192$ and $3 \times 5 \times 3 \times 4 = 180$ number of combinations of parameter values for non-linearity model and heteroskedasticity model respectively.
-Three replications are made for each of the combination results in $192 \times 3 = 576$ and $180 \times 3 = 540$ lineups.
-In addition, each lineup is designed to be evaluated by five different participants.
-After attempting some pilot studies internally, we decide to present a block of 20 lineups to every participant.
-And to ensure the quality of the survey data, two lineups with obvious visual patterns are included as attention checks.
-Thus, $576 \times 5 / (20-2) = 160$ and $540 \times 5 / (20-2) = 150$ participants are recruited to satisfy the design of the data collection period I and II respectively.
+From Table \ref{tab:model-factor-table} it can be computed that there are a total of $4 \times 4 \times 3 \times 4 = 192$ and $3 \times 5 \times 3 \times 4 = 180$ combinations of parameter values for the non-linearity model and heteroskedasticity models respectively.
+Three replications for each combination results in $192 \times 3 = 576$ and $180 \times 3 = 540$ lineups, respectively.
+
+Each lineup needs to be evaluated by at least five participants.
+From previous work, and additional pilot studies for this experiment, we decided to record evaluations from 20 lineups for each participant.
+Two of the 20 lineups with clear visual patterns were used as attention checks, to ensure quality data.
+Thus, $576 \times 5 / (20-2) = 160$ and $540 \times 5 / (20-2) = 150$ participants were needed to cover the experimental design for the data collection periods I and II, respectively.
+The factor levels and range of difficulty was assigned relatively equally among participants.
+The Appendix contains graphical summaries of these assignments for each subject.
 

### 2023-04-23 paper.tex
working hard on revising the experimental setup section

@@
 of the residual departures from the model assumptions. Details about the
-effect size are provided in Section \ref{effect-size} after the
-introduction of the simulation model used in our experiment. The
-theoretical power of a test is sometimes not a trival solution, but it
-can be estimated if the data generating process is known. We use a
+calculation of effect size are provided in Section \ref{effect-size}
+after the introduction of the simulation model used in our experiment.
+The theoretical power of a test is sometimes not a trival solution, but
+it can be estimated if the data generating process is known. We use a
 predefined model to generate a large set of simulated data under
@@
 
-\hypertarget{controlling-the-signal-strength-of-the-signal}{%
-\subsubsection{Controlling the signal strength of the
-signal}\label{controlling-the-signal-strength-of-the-signal}}
+\hypertarget{controlling-the-signal-strength}{%
+\subsubsection{Controlling the signal
+strength}\label{controlling-the-signal-strength}}
 
@@
 Effect size in statistics measures the strength of the signal relative
-to the noise. It is surprisingly difficult to quantify in general, even
-for simulated data as used in this experiment.
+to the noise. It is surprisingly difficult to quantify, even for
+simulated data as used in this experiment.
 
@@
 
-\[E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}\sigma^2))^{-1}\boldsymbol{\mu}_z\right)\]
+\begin{equation}
+E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}\sigma^2))^{-1}\boldsymbol{\mu}_z\right)
+\label{eq:nonlin-effect-size}
+\end{equation}
+
 \noindent where \(diag(.)\) is the diagonal matrix constructed from the
@@
 \(\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z\)
-is the expected values of residuals with \(\boldsymbol{Z}\) be any
-higher order terms of \(\boldsymbol{X}\) leave out by the regression
-equation and \(\boldsymbol{\beta}_z\) be the corresponding coefficients,
-and \(\sigma^2\boldsymbol{I}\) is the assumed covariance matrix of the
-error term when \(H_0\) is true.
+is the expected values of residuals where \(\boldsymbol{Z}\) contain any
+higher order terms of \(\boldsymbol{X}\) left out of the regression
+equation, \(\boldsymbol{\beta}_z\) contains the corresponding
+coefficients, and \(\sigma^2\boldsymbol{I}\) is the assumed covariance
+matrix of the error term when \(H_0\) is true.
 
@@
 
-\[E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|diag(\boldsymbol{R})|} - n + tr(diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}diag(\boldsymbol{R}))\right)\]
+\begin{equation}
+E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|diag(\boldsymbol{R})|} - n + tr(diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}diag(\boldsymbol{R}))\right)
+\label{eq:hetero-effect-size}
+\end{equation}
+
 \noindent where \(\boldsymbol{V}\) is the actual covariance matrix of
-the error term. (Derivations for these equations are provided in the
-Appendix.)
+the error term. Derivations for these equations are provided in the
+Appendix.
 
@@
 
-As shown in Table \ref{tab:model-factor-table}, there are a total of
-\(4 \times 4 \times 3 \times 4 = 192\) and
-\(3 \times 5 \times 3 \times 4 = 180\) number of combinations of
-parameter values for non-linearity model and heteroskedasticity model
-respectively. Three replications are made for each of the combination
-results in \(192 \times 3 = 576\) and \(180 \times 3 = 540\) lineups. In
-addition, each lineup is designed to be evaluated by five different
-participants. After attempting some pilot studies internally, we decide
-to present a block of 20 lineups to every participant. And to ensure the
-quality of the survey data, two lineups with obvious visual patterns are
-included as attention checks. Thus, \(576 \times 5 / (20-2) = 160\) and
-\(540 \times 5 / (20-2) = 150\) participants are recruited to satisfy
-the design of the data collection period I and II respectively.
+From Table \ref{tab:model-factor-table} it can be computed that there
+are a total of \(4 \times 4 \times 3 \times 4 = 192\) and
+\(3 \times 5 \times 3 \times 4 = 180\) combinations of parameter values
+for the non-linearity model and heteroskedasticity models respectively.
+Three replications for each combination results in
+\(192 \times 3 = 576\) and \(180 \times 3 = 540\) lineups, respectively.
+
+Each lineup needs to be evaluated by at least five participants. From
+previous work, and additional pilot studies for this experiment, we
+decided to record evaluations from 20 lineups for each participant. Two
+of the 20 lineups with clear visual patterns were used as attention
+checks, to ensure quality data. Thus, \(576 \times 5 / (20-2) = 160\)
+and \(540 \times 5 / (20-2) = 150\) participants were needed to cover
+the experimental design for the data collection periods I and II,
+respectively. The factor levels and range of difficulty was assigned
+relatively equally among participants. The Appendix contains graphical
+summaries of these assignments for each subject.
 

### 2023-04-23 paper.Rmd
done with 4.2.2

@@
 
-As mentioned in Section \ref{power-of-the-tests}, $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using null lineups.
+Data collection period III was primarily to obtain Rorschach lineup evaluations to estimate $\alpha$ from Equation \ref{eq:pvalue-beta-binomial}, and also to obtain additional evaluations of lineups made with uniform fitted value distribution to ensure consistency in the results.
+To construct a Rorschach lineup, the data is generated from a model with zero effect size, while the null data are generated using the same simulation method discussed in Section \ref{visual-test-procedure-based-on-lineups}.
+This procedure differs from that of the canonical Rorschach lineup, where all 20 plots are generated directly from the null model, so the method suggested in @vanderplas2021statistical for typical lineups containing a data plot is used to estimate $\alpha$.
+(The Appendix contains a sensitivity analysis on the effect of uncertainty in the estimation of $\alpha$.) The $3 \times 4 = 12$ treatment levels of the common factors, replicated three times results in 36 lineups.
+And 6 more evaluations on the 279 lineups with uniform fitted value distribution, results in at least $(36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133$ participants needed.
+
+<!--As mentioned in Section \ref{power-of-the-tests}, $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using null lineups.
 Three replications are made for $3 \times 4 = 12$ combinations of common factors $n$ and fitted value distribution, results in $12 \times 3 = 36$ lineups included in data collection period III.
@@
 Therefore, we plan to collect 6 more evaluations on the 279 lineups with uniform fitted value distribution, result in $(36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133$ participants recruited for data collection period III.
+-->
 

### 2023-04-23 paper.tex
done with 4.2.2

@@
 
-As mentioned in Section \ref{power-of-the-tests}, \(\alpha\) used in
-Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using null
-lineups. Three replications are made for \(3 \times 4 = 12\)
-combinations of common factors \(n\) and fitted value distribution,
-results in \(12 \times 3 = 36\) lineups included in data collection
-period III. In these lineups, the data of the data plot is generated
-from a model with zero effect size, while the data of the 19 null plots
-are generated using the same simulation method discussed in Section
-\ref{visual-test-procedure-based-on-lineups}. This generation procedure
-differs from the canonical Rorschach lineup procedure, which requires
-that all 20 plots are generated directly from the null model. However,
-these lineups serve the same fundamental purpose: to assess the number
-of visually interesting plots generated under \(H_0\).
-
-To account for the fact that our simulation method for these lineups is
-not the Rorschach procedure, we use the method suggested in
+Data collection period III was primarily to obtain Rorschach lineup
+evaluations to estimate \(\alpha\) from Equation
+\ref{eq:pvalue-beta-binomial}, and also to obtain additional evaluations
+of lineups made with uniform fitted value distribution to ensure
+consistency in the results. To construct a Rorschach lineup, the data is
+generated from a model with zero effect size, while the null data are
+generated using the same simulation method discussed in Section
+\ref{visual-test-procedure-based-on-lineups}. This procedure differs
+from that of the canonical Rorschach lineup, where all 20 plots are
+generated directly from the null model, so the method suggested in
 \citet{vanderplas2021statistical} for typical lineups containing a data
-plot to estimate \(\alpha\). (We have included a sensitivity analysis in
-the Appendix to examine the impact of the variance of the \(\alpha\)
-estimate on our findings.)
-
-All lineups consist of only null plots are planned to be evaluated by 20
-participants. However, presenting only these lineups to participants are
-considered to be bad practices as participants will lose interest
-quickly. Therefore, we plan to collect 6 more evaluations on the 279
-lineups with uniform fitted value distribution, result in
+plot is used to estimate \(\alpha\). (The Appendix contains a
+sensitivity analysis on the effect of uncertainty in the estimation of
+\(\alpha\).) The \(3 \times 4 = 12\) treatment levels of the common
+factors, replicated three times results in 36 lineups. And 6 more
+evaluations on the 279 lineups with uniform fitted value distribution,
+results in at least
 \((36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133\) participants
-recruited for data collection period III.
+needed.
 

### 2023-04-23 paper.Rmd
done with 4.2.3

@@
 
-Participants for all three data collection periods are recruited from an crowdsourcing platform called Prolific [@palan2018prolific].
-Prescreening procedure is applied during the recruitment, participants are required to be fluent in English, with $98\%$ minimum approval rate and 10 minimum submissions in other studies.
+Participants for all three data collection periods were recruited from the Prolific crowd-sourcing platform [@palan2018prolific].
+Pre-screening procedures were applied during the recruitment: participants were required to be fluent in English, with $98\%$ minimum approval rate and at least 10 submissions in other studies.
 
@@
 A lineup consists of a randomly placed data plot and 19 null plots, which are all residual plots drawn with raw residuals on the y-axis and fitted values on the x-axis.
-An additional horizontal red line is added at $y = 0$ as a helping line.
-
-The data of the data plot is simulated from one of two models described in Section \ref{simulating-departures-from-good-residuals}, while the data of the remaining 19 null plots are generated by the residual rotation technique discussed in Section \ref{visual-test-procedure-based-on-lineups}.
-
-In every lineup evaluation, the participant is asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots are from others.
-If there is no noticeable difference between plots in a lineup, participants are permitted to select zero plots without providing the reason.
-No participant are shown the same lineup twice.
-Information about preferred pronoun, age group, education, and previous experience in visual experiments are also collected.
-A participant's submission is only accepted if the data plot is identified for at least one attention check.
-Data of rejected submissions are discarded automatically to maintain the overall data quality.
+An additional horizontal red line is added at $y = 0$ as a visual reference.
+The data in the data plot is simulated from one of two models described in Section \ref{simulating-departures-from-good-residuals}, while the data of the remaining 19 null plots are generated by the residual rotation technique discussed in Section \ref{visual-test-procedure-based-on-lineups}.
+
+In each lineup evaluation, the participant was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots are from others.
+If there is no noticeable difference between plots in a lineup, participants were permitted to select zero plots without providing the reason.
+No participant was shown the same lineup twice.
+Information about preferred pronouns, age group, education, and previous experience in visual experiments were also collected.
+A participant's submission was only included in the analysis if the data plot is identified for at least one attention check.
+<!--Data of rejected submissions are discarded automatically to maintain the overall data quality.-->
 

### 2023-04-23 paper.tex
done with 4.2.3

@@
 
-Participants for all three data collection periods are recruited from an
-crowdsourcing platform called Prolific \citep{palan2018prolific}.
-Prescreening procedure is applied during the recruitment, participants
-are required to be fluent in English, with \(98\%\) minimum approval
-rate and 10 minimum submissions in other studies.
+Participants for all three data collection periods were recruited from
+the Prolific crowd-sourcing platform \citep{palan2018prolific}.
+Pre-screening procedures were applied during the recruitment:
+participants were required to be fluent in English, with \(98\%\)
+minimum approval rate and at least 10 submissions in other studies.
 
@@
 y-axis and fitted values on the x-axis. An additional horizontal red
-line is added at \(y = 0\) as a helping line.
-
-The data of the data plot is simulated from one of two models described
-in Section \ref{simulating-departures-from-good-residuals}, while the
-data of the remaining 19 null plots are generated by the residual
-rotation technique discussed in Section
-\ref{visual-test-procedure-based-on-lineups}.
-
-In every lineup evaluation, the participant is asked to select one or
+line is added at \(y = 0\) as a visual reference. The data in the data
+plot is simulated from one of two models described in Section
+\ref{simulating-departures-from-good-residuals}, while the data of the
+remaining 19 null plots are generated by the residual rotation technique
+discussed in Section \ref{visual-test-procedure-based-on-lineups}.
+
+In each lineup evaluation, the participant was asked to select one or
 more plots that are most different from others, provide a reason for
@@
 plots are from others. If there is no noticeable difference between
-plots in a lineup, participants are permitted to select zero plots
-without providing the reason. No participant are shown the same lineup
-twice. Information about preferred pronoun, age group, education, and
-previous experience in visual experiments are also collected. A
-participant's submission is only accepted if the data plot is identified
-for at least one attention check. Data of rejected submissions are
-discarded automatically to maintain the overall data quality.
+plots in a lineup, participants were permitted to select zero plots
+without providing the reason. No participant was shown the same lineup
+twice. Information about preferred pronouns, age group, education, and
+previous experience in visual experiments were also collected. A
+participant's submission was only included in the analysis if the data
+plot is identified for at least one attention check.
 

### 2023-04-23 paper.Rmd
done with 5.1

@@
 Data collection used a total of `r filter(vi_survey, !attention_check) %>% count(unique_lineup_id) %>% nrow()` lineups, and resulted in a total of `r nrow(filter(vi_survey, !attention_check))` evaluations from `r vi_survey %>% count(set, exp) %>% nrow()` participants.
-Roughly half corresponded to the two models, non-linearity and heteroskedasticiy, and the three collection periods had similar numbers in each.
+Roughly half corresponded to the two models, non-linearity and heteroskedasticiy, and the three collection periods had similar numbers of evaluations.
 Each participant received two of the `r vi_survey %>% filter(attention_check) %>% count(unique_lineup_id) %>% nrow()` attention check lineups which were used to filter results of participants who were clearly not making an honest effort (only 11 of `r vi_survey %>% count(set, exp) %>% nrow() + 11`).
 To estimate $\alpha$ for calculating statistical significance (see Section \ref{sig}) there were `r nrow(filter(vi_survey, null_lineup, !attention_check))` evaluations of `r vi_survey %>% filter(null_lineup, !attention_check) %>% count(unique_lineup_id) %>% nrow()` null lineups.
-Neither the attention checks nor null lineups were used for the subsequent analysis.
+Neither the attention checks nor null lineups were used in the subsequent analysis.
 The de-identified data, `vi_survey`, is made available in the R package, `visage`.
@@
 The data was collected on lineups constructed from four different fitted value distributions: uniform, normal, skewed and discrete.
-More data was collected on the uniform (each evaluated by 11 participants) than the others (each evaluated by 5 participants).
-The analysis in Sections \ref{power-analysis}-\ref{hetero-analysis}, uses only results from lineups generated with uniform fitted values, a total `r nrow(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"))` lineup evaluations.
-This was decided in order to compare the conventional and visual test performance for an optimal scenario.
+More data was collected on the uniform distribution (each evaluated by 11 participants) than the others (each evaluated by 5 participants).
+The analysis in Sections \ref{power-analysis}-\ref{hetero-analysis} uses only results from lineups generated with uniform fitted values, for a total `r nrow(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"))` lineup evaluations.
+This allows us to compare the conventional and visual test performance in an optimal scenario.
 Section \ref{effect-of-fitted-value-distributions} examines how the results may be affected if the fitted value distribution was different.
@@
 
-Figures \ref{fig:polypower} and \ref{fig:heterpower} present the power curves of various tests plotted against the effect size in the residuals, for non-linearity and heteroskedasticity, respectively.
+Figures \ref{fig:polypower} and \ref{fig:heterpower} present the power curves of various tests plotted against the effect size in the residuals for non-linearity and heteroskedasticity respectively.
 In each case the power of visual test is calculated for multiple bootstrap samples leading to the many (orange) curves.
@@
 These plots serve as a visual aid to help readers understand how different effect size values translate to changes in the diagnostic plots.
-The horizontal lines of dots at 0 and 1 represent the non-reject or reject made by visual tests for each lineup.
+The horizontal lines of dots at 0 and 1 represent the non-rejection or rejection decisions made by visual tests for each lineup.
 
@@
 The test with the uniformly higher power is the RESET test, one that specifically tests for non-linearity.
-(Note that the BP and SW tests have much lower power, which is expected because they are not designed to detect non-linearity.) The bootstrapped power curves for the visual test are effectively a shift right from that of the RESET test.
+Note that the BP and SW tests have much lower power, which is expected because they are not designed to detect non-linearity.
+The bootstrapped power curves for the visual test are effectively a shift right from that of the RESET test.
 This means that the RESET test will reject at a lower effect size (less structure) than the visual test, but otherwise the performance will be similar.
 In other words, the RESET test is more sensitive than the visual test.
-This is not necessarily a good feature for the purposes of diagnosing model defects.
-If one scans the residual plot examples at the bottom, we might argue that the non-linearity is not sufficiently problematic until an effect size of around 3 or 3.5.
+This is not necessarily a good feature for the purposes of diagnosing model defects: If we scan the residual plot examples at the bottom, we might argue that the non-linearity is not sufficiently problematic until an effect size of around 3 or 3.5.
 The RESET test would reject closer to an effect size of 2, but the visual test would reject closer 3.25, for a significance level of 0.05.
+The visual test matches the robustness of the model to (minor) violations of assumptions much better.
 
@@
 The visual test power curve is a right shift.
-This shows a similar story to the power curves for non-linearity pattern - the conventional test is more sensitive than the visual test.
+This shows a similar story to the power curves for non-linearity pattern: the conventional test is more sensitive than the visual test.
 From the example residual plots at the bottom we might argue that the heteroskedasticity becomes noticeably visible around an effect size of 3 or 3.5.
@@
 What we gain from using a visual test for this purpose is the removal of any subjective arguments about whether a pattern is visible or not.
-The lineup protocol provides the calibration for detecting patterns: that if the pattern in the data plot cannot be distinguished from patterns in good residual plots, then no discernible problem with the model exists.
+The lineup protocol provides the calibration for detecting patterns: if the pattern in the data plot cannot be distinguished from the patterns in good residual plots, then no discernible problem with the model exists.
 

### 2023-04-23 paper.tex
done with 5.1

@@
 two models, non-linearity and heteroskedasticiy, and the three
-collection periods had similar numbers in each. Each participant
+collection periods had similar numbers of evaluations. Each participant
 received two of the 24 attention check lineups which were used to filter
@@
 significance (see Section \ref{sig}) there were 720 evaluations of 36
-null lineups. Neither the attention checks nor null lineups were used
-for the subsequent analysis. The de-identified data,
-\texttt{vi\_survey}, is made available in the R package,
-\texttt{visage}.
+null lineups. Neither the attention checks nor null lineups were used in
+the subsequent analysis. The de-identified data, \texttt{vi\_survey}, is
+made available in the R package, \texttt{visage}.
 
@@
 value distributions: uniform, normal, skewed and discrete. More data was
-collected on the uniform (each evaluated by 11 participants) than the
-others (each evaluated by 5 participants). The analysis in Sections
-\ref{power-analysis}-\ref{hetero-analysis}, uses only results from
-lineups generated with uniform fitted values, a total 3069 lineup
-evaluations. This was decided in order to compare the conventional and
-visual test performance for an optimal scenario. Section
+collected on the uniform distribution (each evaluated by 11
+participants) than the others (each evaluated by 5 participants). The
+analysis in Sections \ref{power-analysis}-\ref{hetero-analysis} uses
+only results from lineups generated with uniform fitted values, for a
+total 3069 lineup evaluations. This allows us to compare the
+conventional and visual test performance in an optimal scenario. Section
 \ref{effect-of-fitted-value-distributions} examines how the results may
@@
 Figures \ref{fig:polypower} and \ref{fig:heterpower} present the power
-curves of various tests plotted against the effect size in the
-residuals, for non-linearity and heteroskedasticity, respectively. In
-each case the power of visual test is calculated for multiple bootstrap
-samples leading to the many (orange) curves. The effect size was
-computed at a 5\% significance level and plotted on a natural
-logarithmic scale. To facilitate visual calibration of effect size
-values with the corresponding diagnostic plots, a sequence of example
-residual plots with increasing effect sizes is provided at the bottom of
-these figures. These plots serve as a visual aid to help readers
-understand how different effect size values translate to changes in the
-diagnostic plots. The horizontal lines of dots at 0 and 1 represent the
-non-reject or reject made by visual tests for each lineup.
+curves of various tests plotted against the effect size in the residuals
+for non-linearity and heteroskedasticity respectively. In each case the
+power of visual test is calculated for multiple bootstrap samples
+leading to the many (orange) curves. The effect size was computed at a
+5\% significance level and plotted on a natural logarithmic scale. To
+facilitate visual calibration of effect size values with the
+corresponding diagnostic plots, a sequence of example residual plots
+with increasing effect sizes is provided at the bottom of these figures.
+These plots serve as a visual aid to help readers understand how
+different effect size values translate to changes in the diagnostic
+plots. The horizontal lines of dots at 0 and 1 represent the
+non-rejection or rejection decisions made by visual tests for each
+lineup.
 
@@
 higher power is the RESET test, one that specifically tests for
-non-linearity. (Note that the BP and SW tests have much lower power,
-which is expected because they are not designed to detect
-non-linearity.) The bootstrapped power curves for the visual test are
-effectively a shift right from that of the RESET test. This means that
-the RESET test will reject at a lower effect size (less structure) than
-the visual test, but otherwise the performance will be similar. In other
-words, the RESET test is more sensitive than the visual test. This is
-not necessarily a good feature for the purposes of diagnosing model
-defects. If one scans the residual plot examples at the bottom, we might
-argue that the non-linearity is not sufficiently problematic until an
-effect size of around 3 or 3.5. The RESET test would reject closer to an
-effect size of 2, but the visual test would reject closer 3.25, for a
-significance level of 0.05.
+non-linearity. Note that the BP and SW tests have much lower power,
+which is expected because they are not designed to detect non-linearity.
+The bootstrapped power curves for the visual test are effectively a
+shift right from that of the RESET test. This means that the RESET test
+will reject at a lower effect size (less structure) than the visual
+test, but otherwise the performance will be similar. In other words, the
+RESET test is more sensitive than the visual test. This is not
+necessarily a good feature for the purposes of diagnosing model defects:
+If we scan the residual plot examples at the bottom, we might argue that
+the non-linearity is not sufficiently problematic until an effect size
+of around 3 or 3.5. The RESET test would reject closer to an effect size
+of 2, but the visual test would reject closer 3.25, for a significance
+level of 0.05. The visual test matches the robustness of the model to
+(minor) violations of assumptions much better.
 
@@
 The visual test power curve is a right shift. This shows a similar story
-to the power curves for non-linearity pattern - the conventional test is
+to the power curves for non-linearity pattern: the conventional test is
 more sensitive than the visual test. From the example residual plots at
@@
 any subjective arguments about whether a pattern is visible or not. The
-lineup protocol provides the calibration for detecting patterns: that if
-the pattern in the data plot cannot be distinguished from patterns in
+lineup protocol provides the calibration for detecting patterns: if the
+pattern in the data plot cannot be distinguished from the patterns in
 good residual plots, then no discernible problem with the model exists.

### 2023-04-23 paper.Rmd
done with 5.2

@@
 
-The power comparison demonstrated that the appropriate conventional tests will reject more aggressively than visual tests, but we don't know how the decisions for each lineup would agree or disagree.
+The power comparison demonstrates that the appropriate conventional tests will reject more aggressively than visual tests, but we don't know how the decisions for each lineup would agree or disagree.
 Here we compare the reject or fail to reject decisions of these tests, across all the lineups.
 Figure \ref{fig:p-value-comparison} shows the agreement of the conventional and visual tests using a mosaic plot for both non-linearity patterns and heteroskedasticity patterns.
-
 For both patterns the lineups resulting in a rejection by the visual test are *all* also rejected by the conventional test, except for one from the heteroskedasticity model.
@@
 
-For lineups containing non-linearity patterns, conventional tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% and visual tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time.
+For non-linearity lineups, conventional tests and visual tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% and `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time, respectively.
 Of the lineups rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by the visual test, that is, approximately half as many as the conventional test.
@@
 
-In terms of lineups containing heteroskedasticity patterns, `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by conventional tests, while `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by visual tests.
-The visual test rejects `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the lineups that the conventional visual test reject.
+In heteroskedasticity lineups, `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by conventional tests, while `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by visual tests.
+Of the lineups rejected by the conventional test, the visual test rejects more than two-thirds of them, too.
 
@@
 Figure \ref{fig:heter-example} shows this lineup.
-The data plot in position seventeen displays a relatively strong heteroskedasticity pattern, and has a strong effect size ($log_e(E)=4.02$).
-This is reflected by the visual test $p\text{-value} = 0.026$.
+The data plot in position seventeen displays a relatively strong heteroskedasticity pattern, and has a strong effect size ($log_e(E)=4.02$), which is reflected by the visual test $p\text{-value} = 0.026$.
 But the BP test $p\text{-value} = 0.056$, is slightly above the significance cutoff of $0.05$.
@@
 
-Because the power curve of the visual tests are a shift to the right of the conventional test (Figures \ref{fig:polypower} and \ref{fig:heter-example}) it would be curious to know whether adjusting the significance level (to .001, .0001, .00001, ...) of the conventional test would generate similar decisions to that of the visual test.
-Interestingly, it doesn't.
-Despite resulting less rejections, neither the RESET or BP tests come to complete agreement with the visual test (See Appendix).
+Because the power curve of the visual tests are a shift to the right of the conventional test (Figures \ref{fig:polypower} and \ref{fig:heter-example}) we examined whether adjusting the significance level (to .001, .0001, .00001, ...) of the conventional test would generate similar decisions to that of the visual test.
+Interestingly, it doesn't: despite resulting less rejections, neither the RESET or BP tests come to complete agreement with the visual test (see Appendix).
 

### 2023-04-23 paper.tex
done with 5.2

@@
 
-The power comparison demonstrated that the appropriate conventional
+The power comparison demonstrates that the appropriate conventional
 tests will reject more aggressively than visual tests, but we don't know
@@
 of the conventional and visual tests using a mosaic plot for both
-non-linearity patterns and heteroskedasticity patterns.
-
-For both patterns the lineups resulting in a rejection by the visual
-test are \emph{all} also rejected by the conventional test, except for
-one from the heteroskedasticity model. This reflects exactly the story
-from the previous section, that the conventional tests reject more
-aggressively than the visual test.
-
-For lineups containing non-linearity patterns, conventional tests reject
-69\% and visual tests reject 32\% of the time. Of the lineups rejected
-by the conventional test, 46\% are rejected by the visual test, that is,
+non-linearity patterns and heteroskedasticity patterns. For both
+patterns the lineups resulting in a rejection by the visual test are
+\emph{all} also rejected by the conventional test, except for one from
+the heteroskedasticity model. This reflects exactly the story from the
+previous section, that the conventional tests reject more aggressively
+than the visual test.
+
+For non-linearity lineups, conventional tests and visual tests reject
+69\% and 32\% of the time, respectively. Of the lineups rejected by the
+conventional test, 46\% are rejected by the visual test, that is,
 approximately half as many as the conventional test. There are no
@@
 
-In terms of lineups containing heteroskedasticity patterns, 76\% are
-rejected by conventional tests, while 56\% are rejected by visual tests.
-The visual test rejects 73\% of the lineups that the conventional visual
-test reject.
+In heteroskedasticity lineups, 76\% are rejected by conventional tests,
+while 56\% are rejected by visual tests. Of the lineups rejected by the
+conventional test, the visual test rejects more than two-thirds of them,
+too.
 
@@
 relatively strong heteroskedasticity pattern, and has a strong effect
-size (\(log_e(E)=4.02\)). This is reflected by the visual test
+size (\(log_e(E)=4.02\)), which is reflected by the visual test
 \(p\text{-value} = 0.026\). But the BP test \(p\text{-value} = 0.056\),
@@
 the conventional test (Figures \ref{fig:polypower} and
-\ref{fig:heter-example}) it would be curious to know whether adjusting
-the significance level (to .001, .0001, .00001, \ldots) of the
-conventional test would generate similar decisions to that of the visual
-test. Interestingly, it doesn't. Despite resulting less rejections,
-neither the RESET or BP tests come to complete agreement with the visual
-test (See Appendix).
+\ref{fig:heter-example}) we examined whether adjusting the significance
+level (to .001, .0001, .00001, \ldots) of the conventional test would
+generate similar decisions to that of the visual test. Interestingly, it
+doesn't: despite resulting less rejections, neither the RESET or BP
+tests come to complete agreement with the visual test (See Appendix).
 

### 2023-04-23 paper.Rmd
done with section 5

@@
 Figure \ref{fig:poly-power-uniform-j} explores the relationship between polynomial order and power of the tests.
-The conventional tests have higher power for lower orders of Hermite polynomials, and it drops substantially for the "triple-U" shape.
-To understand why this is, one needs to return to the way the RESET test is applied.
-It requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four [@ramsey_tests_1969].
+The conventional tests have higher power for lower orders of Hermite polynomials, and the power drops substantially for the "triple-U" shape.
+To understand why this is, we return to the application of the RESET test, which requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four [@ramsey_tests_1969].
 However, the "triple-U" shape is constructed from the Hermite polynomials using power up to 18.
-If the RESET test had been applied using a higher power no less than six, the power curve of "triple-U" shape will be closer to other power curves.
-This illustrates the sensitivity of the conventional test to the parameter choice, and highlights a limitation that it helps to know the data generating process to set the parameters for the test, which is unrealistic.
-However, we examined this in more detail (see Appendix) and found that there is no harm for setting the parameter higher than four on the tests' operation for lower order polynomial shapes.
-Using a parameter value of six, instead of four, yields higher power regardless of generating process, and would be recommended.
+If the RESET test had been applied using a higher power of no less than six, the power curve of "triple-U" shape will be closer to other power curves.
+This illustrates the sensitivity of the conventional test to the parameter choice, and highlights a limitation: it helps to know the data generating process to set the parameters for the test, which is unrealistic in practice.
+However, we examined this in more detail (see Appendix) and found that there is no harm in setting the parameter higher than four on the tests' operation for lower order polynomial shapes.
+Using a parameter value of six, instead of four, yields higher power regardless of generating process, and is recommended.
 
 For visual tests, we expect the "U" shape to be detected more readily, followed by the "S", "M" and "triple-U" shape.
-From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves mostly align with these expectations, except for the "M" shape, which is as easy to be detected as the "S" shape.
+From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves mostly align with these expectations, except for the "M" shape, which is as easily detected as the "S" shape.
 This suggests a benefit of the visual test: knowing the shape ahead of time is *not* needed for its application.
@@
 The "left-triangle" and the "right-triangle" shapes are functionally identical, and this is observed for the conventional test, where the power curves are identical.
-Interestingly there is a difference for the visual test, where the power curve of the "left-triangle" shape is slightly higher than that of the "right-triangle" shape.
-This indicates a bias in perceiving heteroskedasticity depending on the direction.
-This would be worth investigating further.
+Interestingly there is a difference for the visual test: the power curve of the "left-triangle" shape is slightly higher than that of the "right-triangle" shape.
+This indicates a bias in perceiving heteroskedasticity depending on the direction, and may be worth investigating further.
 
@@
 This is an often forgotten element of regression analysis but it is important.
-Where $X$ is observed, particularly the distribution of the $X$ values in the sample, or consequently $\hat{Y}$, may affect the ability to read any patterns in the residual plots.
-This experiment was constructed to assess this, based on four different distributions of fitted values: uniform, normal, discrete and lognormal (skewed).
-We would expect that if $\hat{Y}$ has a uniform distribution, this would make it easier to read the relationship with the residuals.
+Where $X$ is observed, the distribution of the $X$ values in the sample, or consequently $\hat{Y}$, may affect the ability to read any patterns in the residual plots.
+The effect of fitted value distribution on test performance is assess using four different distributions of fitted values: uniform, normal, discrete and lognormal (skewed).
+We expect that if $\hat{Y}$ has a uniform distribution, it is easier to read the relationship with the residuals.
 
@@
 Figure \ref{fig:different-x-dist-poly-power} examines the impact of the fitted value distribution on the power of conventional (left) and visual (right) tests for both the non-linearity (top) and heteroskedasticity (bottom) patterns.
-For conventional tests, only the power curves of appropriate tests which are RESET tests and BP tests for non-linearity pattern and heteroskedasticity pattern respectively are shown.
-For visual tests, note that, more evaluations on lineup with uniform fitted value distribution were collected, so to have a fair comparison, we randomly sample five from the total eleven evaluations to estimate the power curves.
-Re-sampling produces the multiple curves for the uniform, and provides an indication of the variation of the power.
-
-Perhaps surprisingly, the visual tests have more consistent power across the different fitted value distributions.
-For the non-linear pattern, there is almost no power difference.
-For the heteroskedastic pattern, uniform and discrete have higher power than normal and lognormal.
-The likely reason is that these latter two have less observations in the tails where the heteroskedastic pattern needs to be detected.
+For conventional tests, only the power curves of appropriate tests are shown:  RESET tests for non-linearity and BP tests for heteroskedasticity.
+For visual tests, more evaluations on lineups with uniform fitted value distribution were collected, so to have a fair comparison, we randomly sample five from the 11 total evaluations to estimate the power curves,  producing the multiple curves for the uniform condition, and providing an indication of the variability in the power estimates.
+
+Perhaps surprisingly, the visual tests have more consistent power across the different fitted value distributions: for the non-linear pattern, there is almost no power difference, and for the heteroskedastic pattern, uniform and discrete have higher power than normal and lognormal.
+The likely reason is that these latter two have fewer observations in the tails where the heteroskedastic pattern needs to be detected.
 
@@
 See, for example, @jamshidian2007study, @olvera2019relationship and  @zhang2018practical which show derivations and use simulation to assess the effect of the observed distribution of the predictors on test power.
-The big differences in power curves that are seen in Figure \ref{fig:different-x-dist-poly-power} is echoed in the results reported in these articles.
+The big differences in the power curves seen in Figure \ref{fig:different-x-dist-poly-power} is echoed in the results reported in these articles.
 

### 2023-04-23 paper.tex
done with section 5

@@
 doesn't: despite resulting less rejections, neither the RESET or BP
-tests come to complete agreement with the visual test (See Appendix).
+tests come to complete agreement with the visual test (see Appendix).
 
@@
 polynomial order and power of the tests. The conventional tests have
-higher power for lower orders of Hermite polynomials, and it drops
-substantially for the ``triple-U'' shape. To understand why this is, one
-needs to return to the way the RESET test is applied. It requires a
+higher power for lower orders of Hermite polynomials, and the power
+drops substantially for the ``triple-U'' shape. To understand why this
+is, we return to the application of the RESET test, which requires a
 parameter indicating degree of fitted values to test for, and the
@@
 polynomials using power up to 18. If the RESET test had been applied
-using a higher power no less than six, the power curve of ``triple-U''
-shape will be closer to other power curves. This illustrates the
-sensitivity of the conventional test to the parameter choice, and
-highlights a limitation that it helps to know the data generating
-process to set the parameters for the test, which is unrealistic.
-However, we examined this in more detail (see Appendix) and found that
-there is no harm for setting the parameter higher than four on the
-tests' operation for lower order polynomial shapes. Using a parameter
-value of six, instead of four, yields higher power regardless of
-generating process, and would be recommended.
+using a higher power of no less than six, the power curve of
+``triple-U'' shape will be closer to other power curves. This
+illustrates the sensitivity of the conventional test to the parameter
+choice, and highlights a limitation: it helps to know the data
+generating process to set the parameters for the test, which is
+unrealistic in practice. However, we examined this in more detail (see
+Appendix) and found that there is no harm in setting the parameter
+higher than four on the tests' operation for lower order polynomial
+shapes. Using a parameter value of six, instead of four, yields higher
+power regardless of generating process, and is recommended.
 
@@
 mostly align with these expectations, except for the ``M'' shape, which
-is as easy to be detected as the ``S'' shape. This suggests a benefit of
-the visual test: knowing the shape ahead of time is \emph{not} needed
-for its application.
+is as easily detected as the ``S'' shape. This suggests a benefit of the
+visual test: knowing the shape ahead of time is \emph{not} needed for
+its application.
 
@@
 observed for the conventional test, where the power curves are
-identical. Interestingly there is a difference for the visual test,
-where the power curve of the ``left-triangle'' shape is slightly higher
-than that of the ``right-triangle'' shape. This indicates a bias in
-perceiving heteroskedasticity depending on the direction. This would be
-worth investigating further.
+identical. Interestingly there is a difference for the visual test: the
+power curve of the ``left-triangle'' shape is slightly higher than that
+of the ``right-triangle'' shape. This indicates a bias in perceiving
+heteroskedasticity depending on the direction, and may be worth
+investigating further.
 
@@
 \(E(Y|X)\). This is an often forgotten element of regression analysis
-but it is important. Where \(X\) is observed, particularly the
-distribution of the \(X\) values in the sample, or consequently
-\(\hat{Y}\), may affect the ability to read any patterns in the residual
-plots. This experiment was constructed to assess this, based on four
-different distributions of fitted values: uniform, normal, discrete and
-lognormal (skewed). We would expect that if \(\hat{Y}\) has a uniform
-distribution, this would make it easier to read the relationship with
-the residuals.
+but it is important. Where \(X\) is observed, the distribution of the
+\(X\) values in the sample, or consequently \(\hat{Y}\), may affect the
+ability to read any patterns in the residual plots. The effect of fitted
+value distribution on test performance is assess using four different
+distributions of fitted values: uniform, normal, discrete and lognormal
+(skewed). We expect that if \(\hat{Y}\) has a uniform distribution, it
+is easier to read the relationship with the residuals.
 
@@
 (bottom) patterns. For conventional tests, only the power curves of
-appropriate tests which are RESET tests and BP tests for non-linearity
-pattern and heteroskedasticity pattern respectively are shown. For
-visual tests, note that, more evaluations on lineup with uniform fitted
-value distribution were collected, so to have a fair comparison, we
-randomly sample five from the total eleven evaluations to estimate the
-power curves. Re-sampling produces the multiple curves for the uniform,
-and provides an indication of the variation of the power.
+appropriate tests are shown: RESET tests for non-linearity and BP tests
+for heteroskedasticity. For visual tests, more evaluations on lineups
+with uniform fitted value distribution were collected, so to have a fair
+comparison, we randomly sample five from the 11 total evaluations to
+estimate the power curves, producing the multiple curves for the uniform
+condition, and providing an indication of the variability in the power
+estimates.
 
 Perhaps surprisingly, the visual tests have more consistent power across
-the different fitted value distributions. For the non-linear pattern,
-there is almost no power difference. For the heteroskedastic pattern,
-uniform and discrete have higher power than normal and lognormal. The
-likely reason is that these latter two have less observations in the
-tails where the heteroskedastic pattern needs to be detected.
+the different fitted value distributions: for the non-linear pattern,
+there is almost no power difference, and for the heteroskedastic
+pattern, uniform and discrete have higher power than normal and
+lognormal. The likely reason is that these latter two have fewer
+observations in the tails where the heteroskedastic pattern needs to be
+detected.
 
@@
 derivations and use simulation to assess the effect of the observed
-distribution of the predictors on test power. The big differences in
-power curves that are seen in Figure
-\ref{fig:different-x-dist-poly-power} is echoed in the results reported
-in these articles.
+distribution of the predictors on test power. The big differences in the
+power curves seen in Figure \ref{fig:different-x-dist-poly-power} is
+echoed in the results reported in these articles.
 

### 2023-04-23 paper.Rmd
done with susns suggestions

@@
 
-Motivated by the advice of regression analysis experts, that residual plots as opposed to conventional tests are indispensible methods for assessing model fit, a human subjects experiment was conducted using visual inference.
+Regression analysis experts suggest that residual plots are indispensible methods for assessing model fit.
+We conducted a perceptual experiment using visual inference to assess this oft-repeated advice.
 The experiment tested two primary departures from good residuals: non-linearity and heteroskedasticity.
 
-The experiment found that conventional residual-based statistical tests are more sensitive to weak departures from model assumptions than visual tests as would be evaluated by humans.
-That is, a conventional test concludes there are problems with the model fit almost twice as often as a human would.
-They often reject when departures in the form of non-linearity and heteroskedasticity are not visibly different from null residual plots.
-
-While it might be argued that the conventional tests are correctly detecting the small but real effects, it can also be considered that the conventional tests are rejecting when it is not necessary.
+We found that conventional residual-based statistical tests are more sensitive to weak departures from model assumptions than visual tests.
+That is, a conventional test concludes there are problems with the model fit almost twice as often as a human.
+Conventional tests often reject when departures in the form of non-linearity and heteroskedasticity are not visibly different from null residual plots.
+
+While it might be argued that the conventional tests are correctly detecting small but real effects, this can also be seen as the conventional tests are rejecting unnecessarily.
 Many of these rejections happen even when downstream analysis and results would not be significantly affected by the small departures from a good fit.
@@
 
-Now it is important to note that residual plots need to be delivered as a lineup, where it is embedded in a field of null plots.
+It is important to note that residual plots need to be delivered as a lineup, embedded in a field of null plots.
 A residual plot may contain many visual features, but some are caused by the characteristics of the predictors and the randomness of the error, not by the violation of the model assumptions.
@@
 
-Human evaluation of residuals is expensive, though.
-It is time-consuming and laborious.
-This is possibly why residual plot analysis is often not done, in practice.
-However, with the emergence of effective computer vision this work could form the basis to provide automated residual plot reading.
+Human evaluation of residuals is expensive, time-consuming and laborious.
+This is possibly why residual plot analysis is often not done in practice.
+However, with the emergence of effective computer vision, our work could lay a foundation for automated residual plot assessment.
 <!--The findings suggest the strong demand of graphical inspection in regression diagnostics, so developing an automatic visual inference system to evaluate lineups of residual plots is valuable.
@@
 The experiment also revealed some interesting results.
-For the most part, the visual test performed very similarly to the appropriate conventional test except that it has a power curve shifted in the less sensitive direction.
-Unlike the conventional tests, where one needs to specifically test for non-linearity or heteroskedasticity the visual test operated effectively across the range of departures from good residuals.
+For the most part, the visual test performed similarly to the appropriate conventional test with a shift in the power curve.
+Unlike conventional tests, where one needs to specifically test for non-linearity or heteroskedasticity the visual test operated effectively across the range of departures from good residuals.
 If the fitted value distribution is not uniform, there is a small loss of power in the visual test.
-But quite a surprise to us was the big difference in power of the conventional test, which appear to be quite sensitive to how the predictors are sampled.
+Surprisingly there is a big difference in power of the conventional test across fitted value distributions.
 Another surprising finding was that the direction of heteroskedasticity appears to affect the ability to visually detect it: both triangles being more difficult to detect than the butterfly, and a small difference in detection between left- and right-triangle.

### 2023-04-23 paper.tex
done with susns suggestions

@@
 
-Motivated by the advice of regression analysis experts, that residual
-plots as opposed to conventional tests are indispensible methods for
-assessing model fit, a human subjects experiment was conducted using
-visual inference. The experiment tested two primary departures from good
-residuals: non-linearity and heteroskedasticity.
-
-The experiment found that conventional residual-based statistical tests
-are more sensitive to weak departures from model assumptions than visual
-tests as would be evaluated by humans. That is, a conventional test
-concludes there are problems with the model fit almost twice as often as
-a human would. They often reject when departures in the form of
-non-linearity and heteroskedasticity are not visibly different from null
-residual plots.
+Regression analysis experts suggest that residual plots are
+indispensible methods for assessing model fit. We conducted a perceptual
+experiment using visual inference to assess this oft-repeated advice.
+The experiment tested two primary departures from good residuals:
+non-linearity and heteroskedasticity.
+
+We found that conventional residual-based statistical tests are more
+sensitive to weak departures from model assumptions than visual tests.
+That is, a conventional test concludes there are problems with the model
+fit almost twice as often as a human. Conventional tests often reject
+when departures in the form of non-linearity and heteroskedasticity are
+not visibly different from null residual plots.
 
 While it might be argued that the conventional tests are correctly
-detecting the small but real effects, it can also be considered that the
-conventional tests are rejecting when it is not necessary. Many of these
-rejections happen even when downstream analysis and results would not be
+detecting small but real effects, this can also be seen as the
+conventional tests are rejecting unnecessarily. Many of these rejections
+happen even when downstream analysis and results would not be
 significantly affected by the small departures from a good fit. The
@@
 
-Now it is important to note that residual plots need to be delivered as
-a lineup, where it is embedded in a field of null plots. A residual plot
-may contain many visual features, but some are caused by the
-characteristics of the predictors and the randomness of the error, not
-by the violation of the model assumptions. These irrelevant visual
-features have a chance to be filtered out by participants with a
-comparison to null plots, resulting in more accurate reading. The lineup
-enables a careful calibration for reading structure in residual plots.
-
-Human evaluation of residuals is expensive, though. It is time-consuming
-and laborious. This is possibly why residual plot analysis is often not
-done, in practice. However, with the emergence of effective computer
-vision this work could form the basis to provide automated residual plot
-reading.
+It is important to note that residual plots need to be delivered as a
+lineup, embedded in a field of null plots. A residual plot may contain
+many visual features, but some are caused by the characteristics of the
+predictors and the randomness of the error, not by the violation of the
+model assumptions. These irrelevant visual features have a chance to be
+filtered out by participants with a comparison to null plots, resulting
+in more accurate reading. The lineup enables a careful calibration for
+reading structure in residual plots.
+
+Human evaluation of residuals is expensive, time-consuming and
+laborious. This is possibly why residual plot analysis is often not done
+in practice. However, with the emergence of effective computer vision,
+our work could lay a foundation for automated residual plot assessment.
 
 The experiment also revealed some interesting results. For the most
-part, the visual test performed very similarly to the appropriate
-conventional test except that it has a power curve shifted in the less
-sensitive direction. Unlike the conventional tests, where one needs to
-specifically test for non-linearity or heteroskedasticity the visual
-test operated effectively across the range of departures from good
-residuals. If the fitted value distribution is not uniform, there is a
-small loss of power in the visual test. But quite a surprise to us was
-the big difference in power of the conventional test, which appear to be
-quite sensitive to how the predictors are sampled. Another surprising
-finding was that the direction of heteroskedasticity appears to affect
-the ability to visually detect it: both triangles being more difficult
-to detect than the butterfly, and a small difference in detection
-between left- and right-triangle.
+part, the visual test performed similarly to the appropriate
+conventional test with a shift in the power curve. Unlike conventional
+tests, where one needs to specifically test for non-linearity or
+heteroskedasticity the visual test operated effectively across the range
+of departures from good residuals. If the fitted value distribution is
+not uniform, there is a small loss of power in the visual test.
+Surprisingly there is a big difference in power of the conventional test
+across fitted value distributions. Another surprising finding was that
+the direction of heteroskedasticity appears to affect the ability to
+visually detect it: both triangles being more difficult to detect than
+the butterfly, and a small difference in detection between left- and
+right-triangle.
 

### 2023-04-25 paper.Rmd

@@
 
-Data used in the $m - 1$ null plots is typically simulated.
+Data used in the $m - 1$ null plots needs to be simulated.
 In regression diagnostics, sampling data consistent with $H_0$ is equivalent to sampling data from the assumed model.

### 2023-04-28 appendix.Rmd
format code and theme

@@
 library(ggpubr)
+
 nl_rates <- NULL
 h_rates <- NULL
+
 for (i in c(0.1, 0.05, 0.01, 0.001, 0.0001, 0.00001, 0.000001, 0.0000001, 0.00000001, 0.000000001,
             0.0000000001, 0.00000000001)) {
+
 p_value_cmp_dat <- vi_survey %>%
@@
   select(type, conv_reject, reject)
+
 nl <- p_value_cmp_dat %>% 
@@
   count(conv_reject, reject, .drop=FALSE) %>%
-  summarise(rej_pct=round(n[1]/(n[1]+n[2])*100, 0),
-            nrej_pct=round(n[3]/(n[3]+n[4])*100, 0)) %>%
-  bind_cols(alpha=i)
+  summarise(rej_pct = round(n[1]/(n[1]+n[2])*100, 0),
+            nrej_pct = round(n[3]/(n[3]+n[4])*100, 0)) %>%
+  bind_cols(alpha = i)
+
 nl_rates <- bind_rows(nl_rates, nl)
+
 # Heteroskedasticity 
@@
   filter(type == "heteroskedasticity") %>% 
-  count(conv_reject, reject, .drop=FALSE) %>%
-  summarise(rej_pct=round(n[1]/(n[1]+n[2])*100, 0),
-            nrej_pct=round(n[3]/(n[3]+n[4])*100, 0)) %>%
-  bind_cols(alpha=i)
+  count(conv_reject, reject, .drop = FALSE) %>%
+  summarise(rej_pct = round(n[1]/(n[1]+n[2])*100, 0),
+            nrej_pct = round(n[3]/(n[3]+n[4])*100, 0)) %>%
+  bind_cols(alpha = i)
+
 h_rates <- bind_rows(h_rates, h)
 }
+
 nl_rates <- nl_rates %>% mutate(type = "non-linearity")
@@
   mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity")))
+
 rej_by_sig <- ggplot(rates) +
-  geom_vline(xintercept=0.05, colour="grey50") +
-  geom_line(aes(x=alpha, y=rej_pct, colour=type), alpha=0.8) +
-  geom_line(aes(x=alpha, y=nrej_pct, colour=type), linetype=2) +
-  scale_x_log10(breaks=c(0.1, 0.001, 0.00001, 0.0000001, 0.000000001, 0.00000000001)) + 
+  geom_vline(xintercept = 0.05, colour = "grey50") +
+  geom_line(aes(x = alpha, y = rej_pct, colour = type), alpha = 0.8) +
+  geom_line(aes(x = alpha, y = nrej_pct, colour = type), linetype = 2) +
+  scale_x_log10(breaks = c(0.1, 0.001, 0.00001, 0.0000001, 0.000000001, 0.00000000001)) + 
   ylab("Visual test rejects (%)") + xlab("Significance level") +
   scale_colour_brewer("", palette = "Dark2") +
-  ggtitle("A")
+  ggtitle("A") +
+  theme_light()
+
 roc <- ggplot(rates) +
-  geom_path(aes(x=nrej_pct, y=rej_pct, colour=type)) +
+  geom_path(aes(x = nrej_pct, y = rej_pct, colour=type)) +
   ylab("Reject | Reject (%)") + xlab("Reject | Fail to reject (%)") +
@@
   coord_equal() +
-  ggtitle("B")
-ggarrange(rej_by_sig, roc, ncol=2, common.legend=TRUE, legend = "bottom")
+  ggtitle("B") +
+  theme_light()
+
+ggarrange(rej_by_sig, roc, ncol = 2, common.legend = TRUE, legend = "bottom")
 ```
@@
 Thus, the algorithm will try to allocate the most different lineup to a participant at each step.
+
+XXX PLACEHOLDER FOR CODE FOR FIGURES
+
+```{r}
+knitr::include_graphics("figures/check_allocation_for_subject_nonlin.png")
+knitr::include_graphics("figures/check_allocation_for_subject_hetero.png")
+```
 

### 2023-04-29 paper.Rmd
found an incorrect figure reference, and difference between quantities in caption and text

@@
 The data plot (position 17) contains a ``butterfly" shape.
-It has a log effect size of $ 3.76$, and visibly displays heteroskedasticity, making it somewhat surprising that it is not detected by the BP test.'}
+It visibly displays heteroskedasticity, making it somewhat surprising that it is not detected by the BP test.'}
 # plot heter-331
@@
 
-Because the power curve of the visual tests are a shift to the right of the conventional test (Figures \ref{fig:polypower} and \ref{fig:heter-example}) we examined whether adjusting the significance level (to .001, .0001, .00001, ...) of the conventional test would generate similar decisions to that of the visual test.
+Because the power curve of the visual tests are a shift to the right of the conventional test (Figures \ref{fig:polypower} and \ref{fig:heterpower}) we examined whether adjusting the significance level (to .001, .0001, .00001, ...) of the conventional test would generate similar decisions to that of the visual test.
 Interestingly, it doesn't: despite resulting less rejections, neither the RESET or BP tests come to complete agreement with the visual test (see Appendix).

### 2023-04-29 paper.tex
found an incorrect figure reference, and difference between quantities in caption and text

@@
 
-Data used in the \(m - 1\) null plots is typically simulated. In
+Data used in the \(m - 1\) null plots needs to be simulated. In
 regression diagnostics, sampling data consistent with \(H_0\) is
@@
 simulated by regressing \(N\) i.i.d standard normal random draws on the
-predictors, then rescaling it by the ratio of residual sum of square in
+predictors, then re-scaling it by the ratio of residual sum of square in
 two regressions.
@@
 The data plot (position 17) contains a ``butterfly" shape.
-It has a log effect size of $ 3.76$, and visibly displays heteroskedasticity, making it somewhat surprising that it is not detected by the BP test.}\label{fig:heter-example}
+It visibly displays heteroskedasticity, making it somewhat surprising that it is not detected by the BP test.}\label{fig:heter-example}
 \end{figure}
@@
 the conventional test (Figures \ref{fig:polypower} and
-\ref{fig:heter-example}) we examined whether adjusting the significance
+\ref{fig:heterpower}) we examined whether adjusting the significance
 level (to .001, .0001, .00001, \ldots) of the conventional test would

### 2023-04-29 coverletter/coverletter.Rmd
draft cover letter added

---
from: Weihao Li
fromqualifications: B.Comm (Hons)
fromtitle: PhD student
fromWWW: https://github.com/TengMCing
fromemail: patrick.li@monash.edu
fromphone: +61 3 9905 2608
fromsignature: sigfile.png
toname: Editors
toaddress: Journal of Computational and Graphical Statistics
opening: Dear Dr Jones and Dr Liang,
closing: Sincerely
spacing: 1.4
output: monash::letter
---

\bigskip
Please consider our manuscript "A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol" for publication in the Journal of Computational and Graphical Statistics.
\bigskip

The paper describes a problem with conventional testing for doing residual diagnostics. The work arose from preliminary research into building a computer vision model to process residual plots. When collecting the training data, from crowd-sourcing evaluations of lineups on simulated residuals, we discovered that the power of the conventional tests are much higher than the visual test. The difference was even bigger than reported in the validation of visual inference study conducted by Majumder, Hofmann and Cook (2013) in JASA, where the comparison was on regression coefficients. So we initially thought, oh this is bad for visual inference, but actually it's good for visual evaluation of residual plots and bad for conventional testing. The paper explains why this is. It should be of interest to readers, epecially for analysts interested in diagnosing their models.

\bigskip
We believe that it is a good fit for the Journal of Computational and Graphical Statistics.

\bigskip
Thank you for the consideration of this manuscript.

### 2023-04-30 paper.Rmd
include reference for type-III error
Fix ref and add footnote

@@
 Because we focus on scatterplots, non-normality is not one of the departures examined in depth in this paper.
-(loy_variations_2016 discuss related work on non-normality checking.)
+(@loy_variations_2016 discuss related work on non-normality checking.)
 
@@
 The SW test was computed using the `shapiro.test` from the core R package `stats`.
-(Although we didn't use it, it's useful to know that the R package `skedastic` [@skedastic] also contains a large collection of functions to test for heteroskedasticity.) Although, the RESET test is exact, it requires the selection of a power parameter.
+^[Although we didn't use it, it's useful to know that the R package `skedastic` [@skedastic] also contains a large collection of functions to test for heteroskedasticity.] Although, the RESET test is exact, it requires the selection of a power parameter.
 @ramsey_tests_1969 recommends a power of four, which we adopted in our analysis.
@@
 @cook1982residuals explain that most residual-based tests for particular types of departure from model assumptions are also sensitive to other types of departures.
-This could be considered a Type III error (XXXREF), where the null hypothesis of good residuals is correctly rejected but for the wrong reason.
+This could be considered a Type III error [@kimball1957errors], where the null hypothesis of good residuals is correctly rejected but for the wrong reason.
 Also, some types of departure can have elements of other types of departure, for example, non-linearity could be viewed as heteroskedasticity.

### 2023-04-30 paper.Rmd
modify power analysis

@@
 Overall, the results show that the conventional tests are more sensitive than the visual test.
-The conventional tests do have higher power for the patterns they are designed to detect, and are generally unable to detect other patterns.
+The conventional tests do have higher power for the patterns they are designed to detect, but they typically fail to detect other patterns unless those patterns are particularly strong.
 The visual test doesn't require specifying the pattern ahead of time, relying purely on whether the observed residual plot is detectably different from "good" residual plots.

### 2023-04-30 paper.Rmd
average out subject ability

@@
 Therefore, we directly estimate the probability of a lineup being rejected by assuming that individual skill has negligible effect on the variation of the power.
-This assumption is not necessary true, but it helps to simplify the model structure, thereby obviating a costly large-scale experiment to estimate complex covariance matrices.
+This assumption essentially averages out the subject ability and helps to simplify the model structure, thereby obviating a costly large-scale experiment to estimate complex covariance matrices.
 The same model given in Equation \ref{eq:logistic-regression-1-1} is applied to model the power of a visual test.

### 2023-05-01 paper.Rmd
0 -> 20 selections

@@
 In each lineup evaluation, the participant was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots are from others.
-If there is no noticeable difference between plots in a lineup, participants were permitted to select zero plots without providing the reason.
+If there is no noticeable difference between plots in a lineup, participants had the option to select zero plots without the need to provide a reason.
+During the process of recording the responses, a zero selection was considered to be equivalent to selecting all 20 plots.
 No participant was shown the same lineup twice.

### 2023-05-01 appendix.Rmd
use google ref for all citations

@@
       Department of Statistics, University of Nebraska, Lincoln, Nebraska, USA
-bibliography: paper.bib
+bibliography: ref.bib
 header-includes: |
@@
 With a greater value of $\hat{\alpha}$, the $p$-value will be smaller, resulting in more lineups being rejected.
-However, The way we generate Rorschach lineup is not strictly the same as what suggested in @vanderplas2021statistical and @buja_statistical_2009.
+However, The way we generate Rorschach lineup is not strictly the same as what suggested in @vanderplas2021statistical and @buja2009statistical.
 Therefore, we conduct a sensitivity analysis in this section to examine the impact of the variation of the estimator $\alpha$ on our primary findings.

### 2023-05-01 paper.Rmd
use google ref for all citations

@@
       Department of Statistics, University of Nebraska, Lincoln, Nebraska, USA
-bibliography: paper.bib
+bibliography: ref.bib
 abstract: |
@@
 To address this, statistical tests are available to check for non-linear patterns [e.g.
-@ramsey_tests_1969], and heteroskedasticity  [e.g.
-@breusch_simple_1979].
+@ramsey1969tests], and heteroskedasticity  [e.g.
+@breusch1979simple].
 
 Linear regression is a well-established procedure, and there is considerable literature describing diagnostic procedures, e.g.
-@draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals.
+@draper1998applied, @montgomery1982introduction, @belsley1980regression, @cook1999applied and @cook1982residuals.
 Interestingly, despite the abundance of conventional tests, ALL of these writings advise that plotting residuals is an essential tool for diagnosing regression model problems:
@@
 
-Graphical summaries where residuals are plotted against fitted values, or other functions of the predictors (expected to be approximately orthogonal to the residuals) are considered to be the most important residual plots by @cook_applied_1999.
+Graphical summaries where residuals are plotted against fitted values, or other functions of the predictors (expected to be approximately orthogonal to the residuals) are considered to be the most important residual plots by @cook1999applied.
 Figure \ref{fig:residual-plot-common-departures}A shows an example of an ideal residual plot where points are symmetrically distributed around the horizontal zero line (red), with no discernible patterns.
@@
 Because we focus on scatterplots, non-normality is not one of the departures examined in depth in this paper.
-(@loy_variations_2016 discuss related work on non-normality checking.)
+(@loy2016variations discuss related work on non-normality checking.)
 
@@
 Many different hypothesis tests are available to detect specific model defects.
-For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan (BP) test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution.
-To test specific forms of non-linearity, one may apply the F-test as a model structural test to examine the significance of specific polynomial and non-linear forms of the predictors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].
+For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white1980heteroskedasticity] or the Breusch-Pagan (BP) test [@breusch1979simple], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution.
+To test specific forms of non-linearity, one may apply the F-test as a model structural test to examine the significance of specific polynomial and non-linear forms of the predictors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey1969tests].
 The Shapiro-Wilk (SW) normality test [@shapiro1965analysis] is the most widely used test of non-normality included by many of the statistical software programs.
@@
 ^[Although we didn't use it, it's useful to know that the R package `skedastic` [@skedastic] also contains a large collection of functions to test for heteroskedasticity.] Although, the RESET test is exact, it requires the selection of a power parameter.
-@ramsey_tests_1969 recommends a power of four, which we adopted in our analysis.
+@ramsey1969tests recommends a power of four, which we adopted in our analysis.
 The BP and SW tests are approximate.
@@
 Also, some types of departure can have elements of other types of departure, for example, non-linearity could be viewed as heteroskedasticity.
-Additionally, other data problems such as outliers can trigger rejection (or not) of the null hypothesis [@cook_applied_1999].
+Additionally, other data problems such as outliers can trigger rejection (or not) of the null hypothesis [@cook1999applied].
 
@@
 To set up the RESET test, we include different powers of fitted values as proxies.
-According to @ramsey_tests_1969, there are no general rules for the power of the fitted values needed by the RESET test, but it finds power up to four is usually sufficient.
+According to @ramsey1969tests, there are no general rules for the power of the fitted values needed by the RESET test, but it finds power up to four is usually sufficient.
 Thus, we follow this guideline to conduct the RESET test.
-For the BP test, the choice of predictors in the auxiliary regression is left to the user [@breusch_simple_1979].
+For the BP test, the choice of predictors in the auxiliary regression is left to the user [@breusch1979simple].
 But as @waldman1983note suggested, it is a good choice for the set of auxiliary predictors in the BP test be the same as the White test.
@@
 
-The examination of data plots to infer signal or patterns (or lack thereof) is fraught with variation in the human ability to interpret and decode the information embedded in a graph [@cleveland_graphical_1984].
+The examination of data plots to infer signal or patterns (or lack thereof) is fraught with variation in the human ability to interpret and decode the information embedded in a graph [@cleveland1984graphical].
 Human examination of diagnostic plots can feel subjective.
@@
 In practice, over-interpretation of a single plot is common.
-For instance, @roy_chowdhury_using_2015 described a published example where authors over-interpreted separation between gene groups from a two-dimensional projection of a linear discriminant analysis.
-@roy_chowdhury_using_2015 showed that there were no differences in the expression levels between the gene groups.
-One solution to over-interpretation is to examine the plot in the context of natural sampling variability assumed by the model, called the lineup protocol, as proposed in @buja_statistical_2009.
-@majumder_validation_2013 showed that the lineup protocol is analogous to the null hypothesis significance testing framework.
+For instance, @roy2015using described a published example where authors over-interpreted separation between gene groups from a two-dimensional projection of a linear discriminant analysis.
+@roy2015using showed that there were no differences in the expression levels between the gene groups.
+One solution to over-interpretation is to examine the plot in the context of natural sampling variability assumed by the model, called the lineup protocol, as proposed in @buja2009statistical.
+@majumder2013validation showed that the lineup protocol is analogous to the null hypothesis significance testing framework.
 The protocol consists of $m$ randomly placed plots, where one plot is the data plot, and the remaining $m - 1$ plots, referred to as the _null plots_, are constructed using the same graphical procedure as the data plot but the data is replaced with null data that is generated in a manner consistent with the null hypothesis, $H_0$.
@@
 In regression diagnostics, sampling data consistent with $H_0$ is equivalent to sampling data from the assumed model.
-As @buja_statistical_2009 suggested, $H_0$ is usually a composite hypothesis controlled by nuisance parameters.
+As @buja2009statistical suggested, $H_0$ is usually a composite hypothesis controlled by nuisance parameters.
 Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to a so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling.
-The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009].
+The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja2009statistical].
 Essentially, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the predictors, then re-scaling it by the ratio of residual sum of square in two regressions.
 
-The effectiveness of lineup protocol for regression analysis has been validated by @majumder_validation_2013 under relatively simple settings with up to two predictors.
+The effectiveness of lineup protocol for regression analysis has been validated by @majumder2013validation under relatively simple settings with up to two predictors.
 Their results suggest that visual tests are capable of testing the significance of a single predictor with a similar power to a t-test, though they express that in general it is unnecessary to use visual inference if there exists a conventional test, and they do not expect the visual test to perform equally well as the conventional test.
@@
 
-The approach used in @majumder_validation_2013 is as follows.
+The approach used in @majumder2013validation is as follows.
 Define $X_j = \{0,1\}$ to be a Bernoulli random variable measuring whether participant $j$ detected the data plot, and $X = \sum_{j=1}^{K}X_j$ be the total number of observers who detected the data plot.
@@
 
-The power of a visual test on the other hand, may additionally depend on the ability of the particular participant, as the skill of each individual may affect the number of observers who identify the data plot from the lineup [@majumder_validation_2013].
-To address this issue, @majumder_validation_2013 models the probability of participant $j$ correctly picking the data plot from lineup $l$ using a mixed-effect logistic regression, with participants treated as random effect.
+The power of a visual test on the other hand, may additionally depend on the ability of the particular participant, as the skill of each individual may affect the number of observers who identify the data plot from the lineup [@majumder2013validation].
+To address this issue, @majumder2013validation models the probability of participant $j$ correctly picking the data plot from lineup $l$ using a mixed-effect logistic regression, with participants treated as random effect.
 Then, the estimated power of a visual test evaluated by a single participant is the predicted value obtained from the mixed effects model.
@@
 Neither the attention checks nor null lineups were used in the subsequent analysis.
-The de-identified data, `vi_survey`, is made available in the R package, `visage`.
+The de-identified data, `vi_survey`, is made available in the R package, `visage`.^[The R package `visage` is currently available on GitHub (https://github.com/TengMCing/visage).]
 
@@
 The conventional tests have higher power for lower orders of Hermite polynomials, and the power drops substantially for the "triple-U" shape.
-To understand why this is, we return to the application of the RESET test, which requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four [@ramsey_tests_1969].
+To understand why this is, we return to the application of the RESET test, which requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four [@ramsey1969tests].
 However, the "triple-U" shape is constructed from the Hermite polynomials using power up to 18.
@@
 The article was created with R packages `rticles` [@rticles], `knitr` [@knitr] and `rmarkdown` [@rmarkdown].
-The project's Github repository (https://github.com/TengMCing/lineup_residual_diagnostics) contains all materials required to reproduce this article.
+The project's GitHub repository (https://github.com/TengMCing/lineup_residual_diagnostics) contains all materials required to reproduce this article.
 

### 2023-05-01 paper.tex
use google ref for all citations

@@
 statistical tests are available to check for non-linear patterns
-\citep[e.g.][]{ramsey_tests_1969}, and heteroskedasticity
-\citep[e.g.][]{breusch_simple_1979}.
+\citep[e.g.][]{ramsey1969tests}, and heteroskedasticity
+\citep[e.g.][]{breusch1979simple}.
 
@@
 \citet{draper1998applied}, \citet{montgomery1982introduction},
-\citet{belsley_regression_1980}, \citet{cook_applied_1999} and
+\citet{belsley1980regression}, \citet{cook1999applied} and
 \citet{cook1982residuals}. Interestingly, despite the abundance of
@@
 orthogonal to the residuals) are considered to be the most important
-residual plots by \citet{cook_applied_1999}. Figure
+residual plots by \citet{cook1999applied}. Figure
 \ref{fig:residual-plot-common-departures}A shows an example of an ideal
@@
 scatterplots, non-normality is not one of the departures examined in
-depth in this paper. (loy\_variations\_2016 discuss related work on
+depth in this paper. (\citet{loy2016variations} discuss related work on
 non-normality checking.)
@@
 defects. For example, the presence of heteroskedasticity can usually be
-tested by applying the White test
-\citep{white_heteroskedasticity-consistent_1980} or the Breusch-Pagan
-(BP) test \citep{breusch_simple_1979}, which are both derived from the
-Lagrange multiplier test \citep{silvey1959lagrangian} principle that
-relies on the asymptotic properties of the null distribution. To test
-specific forms of non-linearity, one may apply the F-test as a model
-structural test to examine the significance of specific polynomial and
-non-linear forms of the predictors, or the significance of proxy
-variables as in the Ramsey Regression Equation Specification Error Test
-(RESET) \citep{ramsey_tests_1969}. The Shapiro-Wilk (SW) normality test
-\citep{shapiro1965analysis} is the most widely used test of
-non-normality included by many of the statistical software programs. The
-Jarque-Bera test \citep{jarque1980efficient} is also used to directly
-check whether the sample skewness and kurtosis match a normal
-distribution.
+tested by applying the White test \citep{white1980heteroskedasticity} or
+the Breusch-Pagan (BP) test \citep{breusch1979simple}, which are both
+derived from the Lagrange multiplier test \citep{silvey1959lagrangian}
+principle that relies on the asymptotic properties of the null
+distribution. To test specific forms of non-linearity, one may apply the
+F-test as a model structural test to examine the significance of
+specific polynomial and non-linear forms of the predictors, or the
+significance of proxy variables as in the Ramsey Regression Equation
+Specification Error Test (RESET) \citep{ramsey1969tests}. The
+Shapiro-Wilk (SW) normality test \citep{shapiro1965analysis} is the most
+widely used test of non-normality included by many of the statistical
+software programs. The Jarque-Bera test \citep{jarque1980efficient} is
+also used to directly check whether the sample skewness and kurtosis
+match a normal distribution.
 
@@
 computed using the \texttt{shapiro.test} from the core R package
-\texttt{stats}. (Although we didn't use it, it's useful to know that the
-R package \texttt{skedastic} \citep{skedastic} also contains a large
-collection of functions to test for heteroskedasticity.) Although, the
-RESET test is exact, it requires the selection of a power parameter.
-\citet{ramsey_tests_1969} recommends a power of four, which we adopted
-in our analysis. The BP and SW tests are approximate.
+\texttt{stats}. \footnote{Although we didn't use it, it's useful to know
+  that the R package \texttt{skedastic} \citep{skedastic} also contains
+  a large collection of functions to test for heteroskedasticity.}
+Although, the RESET test is exact, it requires the selection of a power
+parameter. \citet{ramsey1969tests} recommends a power of four, which we
+adopted in our analysis. The BP and SW tests are approximate.
 
@@
 to other types of departures. This could be considered a Type III error
-(XXXREF), where the null hypothesis of good residuals is correctly
-rejected but for the wrong reason. Also, some types of departure can
-have elements of other types of departure, for example, non-linearity
-could be viewed as heteroskedasticity. Additionally, other data problems
-such as outliers can trigger rejection (or not) of the null hypothesis
-\citep{cook_applied_1999}.
+\citep{kimball1957errors}, where the null hypothesis of good residuals
+is correctly rejected but for the wrong reason. Also, some types of
+departure can have elements of other types of departure, for example,
+non-linearity could be viewed as heteroskedasticity. Additionally, other
+data problems such as outliers can trigger rejection (or not) of the
+null hypothesis \citep{cook1999applied}.
 
@@
 decode the information embedded in a graph
-\citep{cleveland_graphical_1984}. Human examination of diagnostic plots
+\citep{cleveland1984graphical}. Human examination of diagnostic plots
 can feel subjective. Ideally we would expect visual evaluation to be
@@
 In practice, over-interpretation of a single plot is common. For
-instance, \citet{roy_chowdhury_using_2015} described a published example
-where authors over-interpreted separation between gene groups from a
+instance, \citet{roy2015using} described a published example where
+authors over-interpreted separation between gene groups from a
 two-dimensional projection of a linear discriminant analysis.
-\citet{roy_chowdhury_using_2015} showed that there were no differences
-in the expression levels between the gene groups. One solution to
+\citet{roy2015using} showed that there were no differences in the
+expression levels between the gene groups. One solution to
 over-interpretation is to examine the plot in the context of natural
 sampling variability assumed by the model, called the lineup protocol,
-as proposed in \citet{buja_statistical_2009}.
-\citet{majumder_validation_2013} showed that the lineup protocol is
+as proposed in \citet{buja2009statistical}.
+\citet{majumder2013validation} showed that the lineup protocol is
 analogous to the null hypothesis significance testing framework. The
@@
 equivalent to sampling data from the assumed model. As
-\citet{buja_statistical_2009} suggested, \(H_0\) is usually a composite
+\citet{buja2009statistical} suggested, \(H_0\) is usually a composite
 hypothesis controlled by nuisance parameters. Since regression models
@@
 statistic is the best justified reference distribution among the three
-\citep{buja_statistical_2009}. Essentially, null residuals can be
+\citep{buja2009statistical}. Essentially, null residuals can be
 simulated by regressing \(N\) i.i.d standard normal random draws on the
@@
 The effectiveness of lineup protocol for regression analysis has been
-validated by \citet{majumder_validation_2013} under relatively simple
+validated by \citet{majumder2013validation} under relatively simple
 settings with up to two predictors. Their results suggest that visual
@@
 
-The approach used in \citet{majumder_validation_2013} is as follows.
+The approach used in \citet{majumder2013validation} is as follows.
 Define \(X_j = \{0,1\}\) to be a Bernoulli random variable measuring
@@
 individual may affect the number of observers who identify the data plot
-from the lineup \citep{majumder_validation_2013}. To address this issue,
-\citet{majumder_validation_2013} models the probability of participant
+from the lineup \citep{majumder2013validation}. To address this issue,
+\citet{majumder2013validation} models the probability of participant
 \(j\) correctly picking the data plot from lineup \(l\) using a
@@
 skill has negligible effect on the variation of the power. This
-assumption is not necessary true, but it helps to simplify the model
-structure, thereby obviating a costly large-scale experiment to estimate
-complex covariance matrices. The same model given in Equation
-\ref{eq:logistic-regression-1-1} is applied to model the power of a
-visual test.
+assumption essentially averages out the subject ability and helps to
+simplify the model structure, thereby obviating a costly large-scale
+experiment to estimate complex covariance matrices. The same model given
+in Equation \ref{eq:logistic-regression-1-1} is applied to model the
+power of a visual test.
 
@@
 plots are from others. If there is no noticeable difference between
-plots in a lineup, participants were permitted to select zero plots
-without providing the reason. No participant was shown the same lineup
-twice. Information about preferred pronouns, age group, education, and
-previous experience in visual experiments were also collected. A
-participant's submission was only included in the analysis if the data
-plot is identified for at least one attention check.
+plots in a lineup, participants had the option to select zero plots
+without the need to provide a reason. During the process of recording
+the responses, a zero selection was considered to be equivalent to
+selecting all 20 plots. No participant was shown the same lineup twice.
+Information about preferred pronouns, age group, education, and previous
+experience in visual experiments were also collected. A participant's
+submission was only included in the analysis if the data plot is
+identified for at least one attention check.
 
@@
 the subsequent analysis. The de-identified data, \texttt{vi\_survey}, is
-made available in the R package, \texttt{visage}.
+made available in the R package, \texttt{visage}.\footnote{The package
+  \texttt{visag} is currently available on GitHub
+  (\url{https://github.com/TengMCing/visage})}
 
@@
 than the visual test. The conventional tests do have higher power for
-the patterns they are designed to detect, and are generally unable to
-detect other patterns. The visual test doesn't require specifying the
-pattern ahead of time, relying purely on whether the observed residual
-plot is detectably different from ``good'' residual plots. They will
-perform equally well regardless of the type of model defect. This aligns
-with the advice of experts on residual analysis, who consider residual
-plot analysis to be an indispensable tool for diagnosing model problems.
-What we gain from using a visual test for this purpose is the removal of
-any subjective arguments about whether a pattern is visible or not. The
-lineup protocol provides the calibration for detecting patterns: if the
-pattern in the data plot cannot be distinguished from the patterns in
-good residual plots, then no discernible problem with the model exists.
+the patterns they are designed to detect, but they typically fail to
+detect other patterns unless those patterns are particularly strong. The
+visual test doesn't require specifying the pattern ahead of time,
+relying purely on whether the observed residual plot is detectably
+different from ``good'' residual plots. They will perform equally well
+regardless of the type of model defect. This aligns with the advice of
+experts on residual analysis, who consider residual plot analysis to be
+an indispensable tool for diagnosing model problems. What we gain from
+using a visual test for this purpose is the removal of any subjective
+arguments about whether a pattern is visible or not. The lineup protocol
+provides the calibration for detecting patterns: if the pattern in the
+data plot cannot be distinguished from the patterns in good residual
+plots, then no discernible problem with the model exists.
 
@@
 parameter indicating degree of fitted values to test for, and the
-recommendation is to generically use four \citep{ramsey_tests_1969}.
+recommendation is to generically use four \citep{ramsey1969tests}.
 However, the ``triple-U'' shape is constructed from the Hermite
@@
 \citep{rticles}, \texttt{knitr} \citep{knitr} and \texttt{rmarkdown}
-\citep{rmarkdown}. The project's Github repository
+\citep{rmarkdown}. The project's GitHub repository
 (\url{https://github.com/TengMCing/lineup_residual_diagnostics})
@@
 \bibliographystyle{tfcad}
-\bibliography{paper.bib}
+\bibliography{ref.bib}
 

### 2023-05-01 paper.tex
Double check the reference

@@
 the subsequent analysis. The de-identified data, \texttt{vi\_survey}, is
-made available in the R package, \texttt{visage}.\footnote{The package
-  \texttt{visag} is currently available on GitHub
-  (\url{https://github.com/TengMCing/visage})}
+made available in the R package, \texttt{visage}.\footnote{The R package
+  \texttt{visage} is currently available on GitHub
+  (\url{https://github.com/TengMCing/visage}).}
 

### 2023-05-01 paper.tex
Fix remaining issues with bib file

@@
 difficult task for humans. Humans will almost always will see a pattern
-(see \citet{kahneman}), so the question that really needs answering is
-whether any pattern perceived is consistent with randomness, purely
-sampling variability, or noise. It is especially difficult to teach this
-to new analysts and students \citep{loy2021bringing}. To address this,
-statistical tests are available to check for non-linear patterns
+(see \citet{kahneman2011thinking}), so the question that really needs
+answering is whether any pattern perceived is consistent with
+randomness, purely sampling variability, or noise. It is especially
+difficult to teach this to new analysts and students
+\citep{loy2021bringing}. To address this, statistical tests are
+available to check for non-linear patterns
 \citep[e.g.][]{ramsey1969tests}, and heteroskedasticity
@@
 when there is only a small effect. (A good discussion can be found in
-\citet{kirk1996}.) While such rejections may be statistically correct,
-their sensitivity may render the results impractical. A key goal of
-residual plot diagnostics is to identify potential issues that could
-lead to incorrect conclusions or errors in subsequent analyses, but
-minor defects in the model are unlikely to have a significant impact and
-may be best disregarded for practical purposes. The experiment discussed
-in this paper specifically addresses this tension between statistical
-and practical statistics.
+\citet{kirk1996practical}.) While such rejections may be statistically
+correct, their sensitivity may render the results impractical. A key
+goal of residual plot diagnostics is to identify potential issues that
+could lead to incorrect conclusions or errors in subsequent analyses,
+but minor defects in the model are unlikely to have a significant impact
+and may be best disregarded for practical purposes. The experiment
+discussed in this paper specifically addresses this tension between
+statistical and practical statistics.
 

### 2023-05-02 paper.Rmd

@@
 However, correctly judging whether NO pattern exists in a residual plot is a difficult task for humans.
-Humans will almost always will see a pattern (see @kahneman), so the question that really needs answering is whether any pattern perceived is consistent with randomness, purely sampling variability, or noise.
+Humans will almost always will see a pattern (see @kahneman2011thinking), so the question that really needs answering is whether any pattern perceived is consistent with randomness, purely sampling variability, or noise.
 It is especially difficult to teach this to new analysts and students [@loy2021bringing].
@@
 With large sample sizes, hypothesis tests may reject the null hypothesis when there is only a small effect.
-(A good discussion can be found in @kirk1996.) While such rejections may be statistically correct, their sensitivity may render the results impractical.
+(A good discussion can be found in @kirk1996practical.) While such rejections may be statistically correct, their sensitivity may render the results impractical.
 A key goal of residual plot diagnostics is to identify potential issues that could lead to incorrect conclusions or errors in subsequent analyses, but minor defects in the model are unlikely to have a significant impact and may be best disregarded for practical purposes.

### 2023-05-02 appendix.Rmd
add allocation plot

@@
 Since the Kullback-Leibler divergence requires a proper density function, we need to simplify the covariance matrix of $\boldsymbol{e}$ by setting all the off-diagonal elements to 0.
-Then, the residuals will assumed to follow $N(\boldsymbol{0}, diag(\boldsymbol{R}\sigma^2))$ under the null hypothesis that the model is correctly specified.
+Then, the residuals will be assumed to follow $N(\boldsymbol{0}, diag(\boldsymbol{R}\sigma^2))$ under the null hypothesis that the model is correctly specified.
 If the model is however misspecified due to omitted variables $\boldsymbol{Z}$, or a non-constant variance $\boldsymbol{V}$, the distribution of residuals can be derived as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}\sigma^2))$ and $N(\boldsymbol{0}, diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'))$ respectively.
@@
 
-XXX PLACEHOLDER FOR CODE FOR FIGURES
-
-```{r}
-knitr::include_graphics("figures/check_allocation_for_subject_nonlin.png")
-knitr::include_graphics("figures/check_allocation_for_subject_hetero.png")
+In Figure \ref{fig:poly-allocate} and Figure \ref{fig:heter-allocate}, we present examples of lineup allocations generated by our algorithm for data collection periods I and II, respectively.
+These figures illustrate the factor values presented to the first 20 participants recruited during each period.
+It can be observed that the algorithm effectively distributed almost all possible one-way and two-way factor combinations among the participants, thereby ensuring a diverse set of lineups for each participant.
+
+```{r poly-allocate, fig.cap = 'Factor values assigned to the first 20 participants recruited during data collection period I, with data points slightly jittered to prevent overlap.
+Each participant have been exposed to all one-way and two-way factor combinations.'}
+vi_survey %>%
+  filter(exp == 3) %>%
+  filter(!attention_check) %>%
+  filter(set <= 20) %>%
+  mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
+  mutate(shape = factor(shape, levels = c("U", "S", "M", "Triple-U"))) %>%
+  mutate(e_sigma = factor(e_sigma)) %>%
+  mutate(n = factor(n)) %>%
+  ggplot() +
+  geom_jitter(aes(e_sigma, shape, col = n),
+             width = 0.2,
+             height = 0.2,
+             alpha = 0.5,
+             size = 3) +
+  facet_wrap(~set) +
+  scale_color_brewer(palette = "Dark2") +
+  theme_light() +
+  ylab("Non-linearity pattern") +
+  xlab(expression(sigma)) +
+  labs(col = "Sample size")
+```
+
+```{r heter-allocate, fig.cap = 'Factor values assigned to the first 20 participants recruited during data collection period II, with data points slightly jittered to prevent overlap.
+Each participant appears to have been exposed to almost all one-way and two-way factor combinations.
+However, two participants, namely participant 16 and participant 18, were not presented with lineups containing a butterfly shape with $b = 4$.
+Additionally, participant 7 did not encounter a lineup featuring a butterfly shape with $b = 1$.'}
+vi_survey %>%
+  filter(exp == 4) %>%
+  filter(!attention_check, !null_lineup) %>%
+  filter(set <= 20) %>%
+  mutate(a = c("Left-triangle", "Butterfly", "Right-triangle")[a + 2]) %>%
+  mutate(a = factor(a, levels = c("Left-triangle", "Butterfly", "Right-triangle"))) %>%
+  mutate(b = factor(b)) %>%
+  mutate(n = factor(n)) %>%
+  ggplot() +
+  geom_jitter(aes(b, a, col = n),
+             width = 0.2,
+             height = 0.2,
+             alpha = 0.5,
+             size = 3) +
+  facet_wrap(~set) +
+  scale_color_brewer(palette = "Dark2") +
+  theme_light() +
+  ylab("Heteroskedasticity pattern") +
+  xlab("b") +
+  labs(col = "Sample size")
 ```

### 2023-05-02 appendix.tex
use google ref for all citations
add allocation plot

@@
 \(\boldsymbol{e}\) by setting all the off-diagonal elements to 0. Then,
-the residuals will assumed to follow
+the residuals will be assumed to follow
 \(N(\boldsymbol{0}, diag(\boldsymbol{R}\sigma^2))\) under the null
@@
 not strictly the same as what suggested in
-\citet{vanderplas2021statistical} and \citet{buja_statistical_2009}.
+\citet{vanderplas2021statistical} and \citet{buja2009statistical}.
 Therefore, we conduct a sensitivity analysis in this section to examine
@@
 participant at each step.
+
+In Figure \ref{fig:poly-allocate} and Figure \ref{fig:heter-allocate},
+we present examples of lineup allocations generated by our algorithm for
+data collection periods I and II, respectively. These figures illustrate
+the factor values presented to the first 20 participants recruited
+during each period. It can be observed that the algorithm effectively
+distributed almost all possible one-way and two-way factor combinations
+among the participants, thereby ensuring a diverse set of lineups for
+each participant.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{appendix_files/figure-latex/poly-allocate-1}
+
+}
+
+\caption{Factor values assigned to the first 20 participants recruited during data collection period I, with data points slightly jittered to prevent overlap.
+Each participant have been exposed to all one-way and two-way factor combinations.}\label{fig:poly-allocate}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{appendix_files/figure-latex/heter-allocate-1}
+
+}
+
+\caption{Factor values assigned to the first 20 participants recruited during data collection period II, with data points slightly jittered to prevent overlap.
+Each participant appears to have been exposed to almost all one-way and two-way factor combinations.
+However, two participants, namely participant 16 and participant 18, were not presented with lineups containing a butterfly shape with $b = 4$.
+Additionally, participant 7 did not encounter a lineup featuring a butterfly shape with $b = 1$.}\label{fig:heter-allocate}
+\end{figure}
 

### 2023-05-02 paper.Rmd
spell check

@@
 Scenario 3 is the most feasible to apply, but has the most dependencies to accommodate for the $p$-value calculation.
-For Scenario 3, VanderPlas et al propose modeling the probability of plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$.
+For Scenario 3, VanderPlas et al propose modelling the probability of plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$.
 The number of times plot $i$ being selected in $K$ evaluations is denoted as $c_i$.
@@
 Details about the calculation of effect size are provided in Section \ref{effect-size} after the introduction of the simulation model used in our experiment.
-The theoretical power of a test is sometimes not a trival solution, but it can be estimated if the data generating process is known.
+The theoretical power of a test is sometimes not a trivial solution, but it can be estimated if the data generating process is known.
 We use a predefined model to generate a large set of simulated data under different effect sizes, and record if the conventional test rejects $H_0$.
@@
 We expect signal strength to decline in the simulated data plots with smaller $n$.
-A sample size of 300 is typically enough for structure to be visible in a scatter plot reliably.
+A sample size of 300 is typically enough for structure to be visible in a scatterplot reliably.
 Beyond 300, structure can be obscured by over-plotting, and a different type of display that included transparency, binning or density mapping, might be needed.
@@
 Interestingly, the power curve for the SW test (for non-normality) is only slightly different to that of the visual test, suggesting that it performs reasonably for detecting heteroskedasticity, too.
-The power curve for the BP test suggests it is not useful for detecting heteroskeadsticity, as expected.
+The power curve for the BP test suggests it is not useful for detecting heteroskedasticity, as expected.
 
@@
 
-Regression analysis experts suggest that residual plots are indispensible methods for assessing model fit.
+Regression analysis experts suggest that residual plots are indispensable methods for assessing model fit.
 We conducted a perceptual experiment using visual inference to assess this oft-repeated advice.
@@
 Many of these rejections happen even when downstream analysis and results would not be significantly affected by the small departures from a good fit.
-The results from human evaluations provide a more practical solution, which reinforces the statements from regression experts that residual plots are an indispensible method for model diagnostics.
+The results from human evaluations provide a more practical solution, which reinforces the statements from regression experts that residual plots are an indispensable method for model diagnostics.
 
@@
 The study website is powered by `PythonAnywhere` [@pythonanywhere] and the `Flask` web framework [@flask].
-The `jsPsych` framework [@jspsych] is used to create behavioral experiments that run in our study website.
+The `jsPsych` framework [@jspsych] is used to create behavioural experiments that run in our study website.
 

### 2023-05-02 appendix.Rmd
spell check

@@
 The vertical grey line marks significance level 0.05.
-When the signicance level gets smaller, the it is possible te see that the visual tests reject (nearly) 100% of the time that the conventional test rejects.
-However, there is not agreement, because the visual tests also increasingly reject residual plots where the conventional test failed to reject.
+When the significance level gets smaller, it is possible to see that the visual tests reject (nearly) 100% of the time that the conventional test rejects.
+However, there is no agreement, because the visual tests also increasingly reject residual plots where the conventional test failed to reject.
 Plot B is comparable to an ROC curve, where the percentage visual test rejection conditional on conventional test decision is plotted: Reject conditional on reject is plotted against reject conditional on fail to reject, for different significance levels.
@@
 It is configured to handle HTTP requests such that participants can correctly receive webpages and submit responses.
-Embedded in the resources sent to participants, the `jsPsych` front-end framework [@jspsych] instructs participants' browsers to render an environment for running behavioral experiments.
-During the experiment, this framework will automatically collect common behavioral data such as response time and clicks on buttons.
+Embedded in the resources sent to participants, the `jsPsych` front-end framework [@jspsych] instructs participants' browsers to render an environment for running behavioural experiments.
+During the experiment, this framework will automatically collect common behavioural data such as response time and clicks on buttons.
 participants' responses are first validated by a scheduled `Python` script run on the server, then push to a Github repository.

### 2023-05-02 paper.Rmd
abstract revised

@@
 abstract: |
-  Regression experts consistently recommend plotting residuals for model diagnosis, despite the existence of numerous hypothesis test procedures.
-This paper provides evidence for why this is good advice, using data from a visual inference experiment.
-We show how conventional tests are too sensitive to be practically useful, and how the lineup protocol can be used to yield reliable and consistent reading of residual plots for better model diagnosis.
+  Regression experts consistently recommend plotting residuals for model diagnosis, despite the availability of many numerical hypothesis test procedures designed to use residuals to assess problems with a model fit.
+Here we provide evidence for why this is good advice using data from a visual inference experiment.
+We show how conventional tests are too sensitive, which means that too often the conclusion would be that the model fit is inadequate.
+The experiment uses the lineup protocol which puts a residual plot in the context of null plots.
+This helps generate reliable and consistent reading of residual plots for better model diagnosis.
+It can also help in an obverse situation where a conventional test would fail to detect a problem with a model due to contaminated data.
+The lineup protocol also serves as an omnibus test, capable of detecting a range of departures from good residuals simultaneously.
 keywords: |

### 2023-05-02 paper.tex
abstract revised

@@
 Regression experts consistently recommend plotting residuals for model
-diagnosis, despite the existence of numerous hypothesis test procedures.
-This paper provides evidence for why this is good advice, using data
+diagnosis, despite the availability of many numerical hypothesis test
+procedures designed to use residuals to assess problems with a model
+fit. Here we provide evidence for why this is good advice using data
 from a visual inference experiment. We show how conventional tests are
-too sensitive to be practically useful, and how the lineup protocol can
-be used to yield reliable and consistent reading of residual plots for
-better model diagnosis.
+too sensitive, which means that too often the conclusion would be that
+the model fit is inadequate. The experiment uses the lineup protocol
+which puts a residual plot in the context of null plots. This helps
+generate reliable and consistent reading of residual plots for better
+model diagnosis. It can also help in an obverse situation where a
+conventional test would fail to detect a problem with a model due to
+contaminated data. The lineup protocol also serves as an omnibus test,
+capable of detecting a range of departures from good residuals
+simultaneously.
 \end{abstract}
@@
 \(p\)-value calculation. For Scenario 3, VanderPlas et al propose
-modeling the probability of plot \(i\) being selected from a lineup as
+modelling the probability of plot \(i\) being selected from a lineup as
 \(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
@@
 after the introduction of the simulation model used in our experiment.
-The theoretical power of a test is sometimes not a trival solution, but
+The theoretical power of a test is sometimes not a trivial solution, but
 it can be estimated if the data generating process is known. We use a
@@
 decline in the simulated data plots with smaller \(n\). A sample size of
-300 is typically enough for structure to be visible in a scatter plot
+300 is typically enough for structure to be visible in a scatterplot
 reliably. Beyond 300, structure can be obscured by over-plotting, and a
@@
 heteroskedasticity, too. The power curve for the BP test suggests it is
-not useful for detecting heteroskeadsticity, as expected.
+not useful for detecting heteroskedasticity, as expected.
 
@@
 Regression analysis experts suggest that residual plots are
-indispensible methods for assessing model fit. We conducted a perceptual
+indispensable methods for assessing model fit. We conducted a perceptual
 experiment using visual inference to assess this oft-repeated advice.
@@
 reinforces the statements from regression experts that residual plots
-are an indispensible method for model diagnostics.
+are an indispensable method for model diagnostics.
 
@@
 the \texttt{Flask} web framework \citep{flask}. The \texttt{jsPsych}
-framework \citep{jspsych} is used to create behavioral experiments that
+framework \citep{jspsych} is used to create behavioural experiments that
 run in our study website.

### 2023-05-02 paper.Rmd
update so all sections are mentioned in the last paragraph of intro

@@
 The paper is structured as follows.
-The next section describes the background on the types of departures that one expects to detect, and outlines a formal statistical process for reading residual plots, called visual inference.
+Section \ref{background} describes the background on the types of departures that one expects to detect, and outlines a formal statistical process for reading residual plots, called visual inference.
+Section \ref{significance-calculation} describes the calculation of the statistical significance and power of the test.
 Section \ref{experimental-design} details the experimental design to compare the decisions made by formal hypothesis testing, and how humans would read diagnostic plots.
@@
 
-# Calculation of statistical significance and test power
+# Calculation of statistical significance and test power {#significance-calculation}
 

### 2023-05-03 paper.Rmd
model mispecifcation description not quite right -> reworked

@@
 
-Model misspecification occurs if functions of predictors that needed to accurately describe the relationship between response and predictors are omitted.
+Model misspecification occurs if functions of predictors that needed to accurately describe the relationship with the response are incorrectly specified.
+This includes instances where a higher-order polynomial term of a predictor is wrongfully omitted.
 Any non-linear pattern visible in the residual plot could be indicative of this problem.

### 2023-05-03 paper.Rmd
reworked 2.2

@@
 For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white1980heteroskedasticity] or the Breusch-Pagan (BP) test [@breusch1979simple], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution.
-To test specific forms of non-linearity, one may apply the F-test as a model structural test to examine the significance of specific polynomial and non-linear forms of the predictors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey1969tests].
+To test specific forms of non-linearity, one may apply the F-test as a model structural test to examine the significance of a specific polynomial and non-linear forms of the predictors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey1969tests].
 The Shapiro-Wilk (SW) normality test [@shapiro1965analysis] is the most widely used test of non-normality included by many of the statistical software programs.
@@
 
-We would expect the RESET test for non-linearity to reject residual plot B, the BP test for heteroskedasticity to reject the residual plot C, and the SW test for non-normality to reject residual plot D, which they all do and all tests also correctly fail to reject residual plot A.
+For residual plots in Figure \ref{fig:residual-plot-common-departures}, we would expect the RESET test for non-linearity to reject residual plot B, the BP test for heteroskedasticity to reject the residual plot C, and the SW test for non-normality to reject residual plot D, which they all do and all tests also correctly fail to reject residual plot A.
 Interestingly, the BP and SW tests also reject the residual plots exhibiting structure that they weren't designed for.
-@cook1982residuals explain that most residual-based tests for particular types of departure from model assumptions are also sensitive to other types of departures.
+@cook1982residuals explain that most residual-based tests for a particular departure from the model assumptions are also sensitive to other types of departures.
 This could be considered a Type III error [@kimball1957errors], where the null hypothesis of good residuals is correctly rejected but for the wrong reason.
-Also, some types of departure can have elements of other types of departure, for example, non-linearity could be viewed as heteroskedasticity.
+Also, some types of departure can have elements of other types of departure, for example, non-linearity can appear like heteroskedasticity.
 Additionally, other data problems such as outliers can trigger rejection (or not) of the null hypothesis [@cook1999applied].

### 2023-05-03 paper.Rmd
fix some minor issues

@@
 We can assess residuals in multiple ways.
-To examine the their univariate distribution, residuals may be plotted as a histogram or normal probability plot.
+To examine the univariate distribution, residuals may be plotted as a histogram or normal probability plot.
 Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, we would consider it to be well-behaved.
@@
 However, correctly judging whether NO pattern exists in a residual plot is a difficult task for humans.
-Humans will almost always will see a pattern (see @kahneman2011thinking), so the question that really needs answering is whether any pattern perceived is consistent with randomness, purely sampling variability, or noise.
+Humans will almost always see a pattern (see @kahneman2011thinking), so the question that really needs answering is whether any pattern perceived is consistent with randomness, purely sampling variability, or noise.
 It is especially difficult to teach this to new analysts and students [@loy2021bringing].
@@
 A greater value of $j$ will result in a curve with more turning points.
-It is expected that the "U" shape will be the easiest detect, as the shape gets more complex it will be harder to perceive in a scatterplot, particularly when there is noise.
+It is expected that the "U" shape will be the easiest to detect, as the shape gets more complex it will be harder to perceive in a scatterplot, particularly when there is noise.
 Figure \ref{fig:different-sigma} shows the "U" shape for different amounts of noise ($\sigma$).

### 2023-05-03 paper.Rmd
rework 2.3

--- 2023-05-03 paper.Rmd
@@
 Human examination of diagnostic plots can feel subjective.
-Ideally we would expect visual evaluation to be done objectively with small amounts of human-to-human variation in reading data plots, but we cannot assume this is the case.
+Ideally we would expect visual evaluation to be done objectively with low human-to-human variation in reading data plots, but we cannot assume this is the case.
 
 In practice, over-interpretation of a single plot is common.
-For instance, @roy2015using described a published example where authors over-interpreted separation between gene groups from a two-dimensional projection of a linear discriminant analysis.
-@roy2015using showed that there were no differences in the expression levels between the gene groups.
+For instance, @roy2015using described a published example where authors over-interpreted separation between gene groups from a two-dimensional projection of a linear discriminant analysis even when there were no differences in the expression levels between the gene groups.
 One solution to over-interpretation is to examine the plot in the context of natural sampling variability assumed by the model, called the lineup protocol, as proposed in @buja2009statistical.
@@
 The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja2009statistical].
-Essentially, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the predictors, then re-scaling it by the ratio of residual sum of square in two regressions.
+Under this method, the null residuals can essentially be simulated by independent drawing from  a standard normal random distribution, then regressing the draws on the predictors, and then re-scaling it by the ratio of the residual sum of square in two regressions.
 
 The effectiveness of lineup protocol for regression analysis has been validated by @majumder2013validation under relatively simple settings with up to two predictors.
-Their results suggest that visual tests are capable of testing the significance of a single predictor with a similar power to a t-test, though they express that in general it is unnecessary to use visual inference if there exists a conventional test, and they do not expect the visual test to perform equally well as the conventional test.
-In their third experiment, where there is not a conventional test, visual test outperforms the conventional test by a large margin.
-This is encouraging, as it supports the use of visual inference in situations where there are no existing statistical testing procedures.
+Their results suggest that visual tests are capable of testing the significance of a single predictor with a similar power to a t-test, though they express that in general it is unnecessary to use visual inference if there exists a corresponding conventional test, and they do not expect the visual test to perform equally well as the conventional test.
+In their third experiment, where the contamination of the data violate the assumptions of the convetional test, visual test outperforms the conventional test by a large margin.
+This is encouraging, as it supports the use of visual inference in situations where there are no existing numerical testing procedures.
 Visual inference has also been integrated into diagnostics for hierarchical linear models where the lineup protocol is used to judge the assumptions of linearity, normality and constant error variance for both the level-1 and level-2 residuals [@loy2013diagnostic, @loy2014hlmdiag and @loy2015you].

### 2023-05-03 coverletter/coverletter.Rmd
tweak of cover letter

@@
 The paper explains why this is.
-It should be of interest to readers, epecially for analysts interested in diagnosing their models.
+It should be of interest to readers, especially for analysts interested in diagnosing their models.
 
 \bigskip
-We believe that it is a good fit for the Journal of Computational and Graphical Statistics.
+Thank you for the consideration of this manuscript. We believe that it is a good fit for the Journal of Computational and Graphical Statistics.
 
 \bigskip
-Thank you for the consideration of this manuscript.

### 2023-05-04 appendix.Rmd
final changes

@@
 Following the terminology introduced by @kullback1951information, $P$ represents the measured probability distribution, and $Q$ represents the assumed probability distribution.
-The Kullback-Leibler divergence is defined as $\int_{-\infty}^{\infty}log(p(x)/q(x))p(x)dx$, where $p(.)$ and $q(.)$ denote probability densities of $P$ and $Q$.
+The Kullback-Leibler divergence is defined as $\int_{-\infty}^{\infty}\log(p(x)/q(x))p(x)dx$, where $p(.)$ and $q(.)$ denote probability densities of $P$ and $Q$.
 
@@
 
-Because $rank(\boldsymbol{R}) = n - p - 1 < n$, $\boldsymbol{e}$ follows a degenerate multivariate normal distribution and does not have a density.
+Because $\text{rank}(\boldsymbol{R}) = n - p - 1 < n$, $\boldsymbol{e}$ follows a degenerate multivariate normal distribution and does not have a density.
 Since the Kullback-Leibler divergence requires a proper density function, we need to simplify the covariance matrix of $\boldsymbol{e}$ by setting all the off-diagonal elements to 0.
 Then, the residuals will be assumed to follow $N(\boldsymbol{0}, diag(\boldsymbol{R}\sigma^2))$ under the null hypothesis that the model is correctly specified.
-If the model is however misspecified due to omitted variables $\boldsymbol{Z}$, or a non-constant variance $\boldsymbol{V}$, the distribution of residuals can be derived as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}\sigma^2))$ and $N(\boldsymbol{0}, diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'))$ respectively.
+If the model is however misspecified due to omitted variables $\boldsymbol{Z}$, or a non-constant variance $\boldsymbol{V}$, the distribution of residuals can be derived as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, \text{diag}(\boldsymbol{R}\sigma^2))$ and $N(\boldsymbol{0}, \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'))$ respectively.
 
 By assuming both $P$ and $Q$ are multivariate normal density functions, the Kullback-Leibler divergence can be rewritten as 
-$$KL = \frac{1}{2}\left(log\frac{|\Sigma_p|}{|\Sigma_q|} - n + tr(\Sigma_p^{-1}\Sigma_q) + (\mu_p - \mu_q)'\Sigma_p^{-1}(\mu_p - \mu_q)\right).$$
+$$KL = \frac{1}{2}\left(\log\frac{|\Sigma_p|}{|\Sigma_q|} - n + \text{tr}(\Sigma_p^{-1}\Sigma_q) + (\mu_p - \mu_q)'\Sigma_p^{-1}(\mu_p - \mu_q)\right).$$
 
@@
 \label{eq:effect-size}
-KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|diag(\boldsymbol{R}\sigma^2)|} - n + tr(diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}diag(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\boldsymbol{\mu}_z\right).
+KL = \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\boldsymbol{\mu}_z\right).
 \end{equation*}
@@
 
-$$KL = \frac{1}{2}\left(\boldsymbol{\mu}_z^{T}(diag(\boldsymbol{R}\sigma^2))^{-1}\boldsymbol{\mu}_z\right).$$
+$$KL = \frac{1}{2}\left(\boldsymbol{\mu}_z'(\text{diag}(\boldsymbol{R}\sigma^2))^{-1}\boldsymbol{\mu}_z\right).$$
 
@@
 
-$$KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|diag(\boldsymbol{R}\sigma^2)|} - n + tr(diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}diag(\boldsymbol{R}\sigma^2))\right).$$
+$$KL = \frac{1}{2}\left(log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}
+\text{diag}(\boldsymbol{R}\sigma^2))\right).$$
 Since we assume $\sigma = 1$ for the heteroskedasticity model, the final form of the formula is
 
-$$KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|diag(\boldsymbol{R})|} - n + tr(diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}diag(\boldsymbol{R}))\right).$$
+$$KL = \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R})|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\text{diag}(\boldsymbol{R}))\right).$$
 
@@
 It is found in the result that the power of a RESET test will be affected by the highest order of fitted values included in the auxiliary formula.
-And we suspect that the current recommendation of the highest order - four, is insufficient to test complex non-linear structures such as the "Triple-U" shape designed in this paper.
-Figure \ref{fig:reset} illustrates the change of power of RESET test while testing the "U" shape and the "Triple-U" shape with different highest orders.
+And we suspect that the current recommendation of the highest order - four, is insufficient to test complex non-linear structures such as the "triple-U" shape designed in this paper.
+Figure \ref{fig:reset} illustrates the change of power of RESET test while testing the "U" shape and the "triple-U" shape with different highest orders.
 Clearly, when testing a simple shape like the "U" shape, the highest order has very little impact on the power.
-But for testing the "Triple-U" shape, there will be a loss of power if the recommended order is used.
+But for testing the "triple-U" shape, there will be a loss of power if the recommended order is used.
 To avoid the loss of power, the highest order needs to be set to at least six.
@@
 ```{r reset, fig.cap = 'Change of power of RESET tests for different orders of fitted values included in the auxiliary formula.
-The left panel is the power of testing the "U" shape and the right panel is the power of testing the "Triple-U" shape.
+The left panel is the power of testing the "U" shape and the right panel is the power of testing the "triple-U" shape.
 The power will not be greatly affected by the highest order in the case of testing the "U" shape.
-In the case of testing the "Triple-U" shape, the highest order needs to be set to at least six to avoid the loss of power.', fig.height = 4}
+In the case of testing the "triple-U" shape, the highest order needs to be set to at least six to avoid the loss of power.', fig.height = 4}
 reset_diff_dat %>%
   bind_rows(reset_diff_dat2) %>%
-  mutate(shape = factor(ifelse(shape == 1, "U", "Triple-U"), levels = c("U", "Triple-U"))) %>%
+  mutate(shape = factor(ifelse(shape == 1, "U", "triple-U"), levels = c("U", "triple-U"))) %>%
   ggplot() +
@@
   scale_color_brewer(palette = "Reds") +
-  facet_wrap(~shape)
+  facet_wrap(~shape) +
+  ylab("Power")
 ```
@@
   filter(set <= 20) %>%
-  mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
-  mutate(shape = factor(shape, levels = c("U", "S", "M", "Triple-U"))) %>%
+  mutate(shape = c("U", "S", "M", "triple-U")[shape]) %>%
+  mutate(shape = factor(shape, levels = c("U", "S", "M", "triple-U"))) %>%
   mutate(e_sigma = factor(e_sigma)) %>%
@@
   filter(set <= 20) %>%
-  mutate(a = c("Left-triangle", "Butterfly", "Right-triangle")[a + 2]) %>%
-  mutate(a = factor(a, levels = c("Left-triangle", "Butterfly", "Right-triangle"))) %>%
+  mutate(a = c("left-triangle", "butterfly", "right-triangle")[a + 2]) %>%
+  mutate(a = factor(a, levels = c("left-triangle", "butterfly", "right-triangle"))) %>%
   mutate(b = factor(b)) %>%

### 2023-05-04 appendix.tex
final changes

@@
 divergence is defined as
-\(\int_{-\infty}^{\infty}log(p(x)/q(x))p(x)dx\), where \(p(.)\) and
+\(\int_{-\infty}^{\infty}\log(p(x)/q(x))p(x)dx\), where \(p(.)\) and
 \(q(.)\) denote probability densities of \(P\) and \(Q\).
@@
 
-Because \(rank(\boldsymbol{R}) = n - p - 1 < n\), \(\boldsymbol{e}\)
-follows a degenerate multivariate normal distribution and does not have
-a density. Since the Kullback-Leibler divergence requires a proper
-density function, we need to simplify the covariance matrix of
-\(\boldsymbol{e}\) by setting all the off-diagonal elements to 0. Then,
-the residuals will be assumed to follow
+Because \(\text{rank}(\boldsymbol{R}) = n - p - 1 < n\),
+\(\boldsymbol{e}\) follows a degenerate multivariate normal distribution
+and does not have a density. Since the Kullback-Leibler divergence
+requires a proper density function, we need to simplify the covariance
+matrix of \(\boldsymbol{e}\) by setting all the off-diagonal elements to
+0. Then, the residuals will be assumed to follow
 \(N(\boldsymbol{0}, diag(\boldsymbol{R}\sigma^2))\) under the null
@@
 can be derived as
-\(N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}\sigma^2))\)
+\(N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, \text{diag}(\boldsymbol{R}\sigma^2))\)
 and
-\(N(\boldsymbol{0}, diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'))\)
+\(N(\boldsymbol{0}, \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'))\)
 respectively.
@@
 functions, the Kullback-Leibler divergence can be rewritten as
-\[KL = \frac{1}{2}\left(log\frac{|\Sigma_p|}{|\Sigma_q|} - n + tr(\Sigma_p^{-1}\Sigma_q) + (\mu_p - \mu_q)'\Sigma_p^{-1}(\mu_p - \mu_q)\right).\]
+\[KL = \frac{1}{2}\left(\log\frac{|\Sigma_p|}{|\Sigma_q|} - n + \text{tr}(\Sigma_p^{-1}\Sigma_q) + (\mu_p - \mu_q)'\Sigma_p^{-1}(\mu_p - \mu_q)\right).\]
 
@@
 
-\[KL = \frac{1}{2}\left(\boldsymbol{\mu}_z^{T}(diag(\boldsymbol{R}\sigma^2))^{-1}\boldsymbol{\mu}_z\right).\]
+\[KL = \frac{1}{2}\left(\boldsymbol{\mu}_z'(\text{diag}(\boldsymbol{R}\sigma^2))^{-1}\boldsymbol{\mu}_z\right).\]
 
@@
 
-\[KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|diag(\boldsymbol{R}\sigma^2)|} - n + tr(diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}diag(\boldsymbol{R}\sigma^2))\right).\]
-Since we assume \(\sigma = 1\) for the heteroskedasticity model, the
-final form of the formula is
+\[KL = \frac{1}{2}\left(log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}
+\text{diag}(\boldsymbol{R}\sigma^2))\right).\] Since we assume
+\(\sigma = 1\) for the heteroskedasticity model, the final form of the
+formula is
 
-\[KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|diag(\boldsymbol{R})|} - n + tr(diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}diag(\boldsymbol{R}))\right).\]
+\[KL = \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R})|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\text{diag}(\boldsymbol{R}))\right).\]
 
@@
 order - four, is insufficient to test complex non-linear structures such
-as the ``Triple-U'' shape designed in this paper. Figure \ref{fig:reset}
+as the ``triple-U'' shape designed in this paper. Figure \ref{fig:reset}
 illustrates the change of power of RESET test while testing the ``U''
-shape and the ``Triple-U'' shape with different highest orders. Clearly,
+shape and the ``triple-U'' shape with different highest orders. Clearly,
 when testing a simple shape like the ``U'' shape, the highest order has
-very little impact on the power. But for testing the ``Triple-U'' shape,
+very little impact on the power. But for testing the ``triple-U'' shape,
 there will be a loss of power if the recommended order is used. To avoid
@@
 \caption{Change of power of RESET tests for different orders of fitted values included in the auxiliary formula.
-The left panel is the power of testing the "U" shape and the right panel is the power of testing the "Triple-U" shape.
+The left panel is the power of testing the "U" shape and the right panel is the power of testing the "triple-U" shape.
 The power will not be greatly affected by the highest order in the case of testing the "U" shape.
-In the case of testing the "Triple-U" shape, the highest order needs to be set to at least six to avoid the loss of power.}\label{fig:reset}
+In the case of testing the "triple-U" shape, the highest order needs to be set to at least six to avoid the loss of power.}\label{fig:reset}
 \end{figure}
@@
 rejected (solid lines) or failed to reject (dashed lines). The vertical
-grey line marks significance level 0.05. When the signicance level gets
-smaller, the it is possible te see that the visual tests reject (nearly)
-100\% of the time that the conventional test rejects. However, there is
-not agreement, because the visual tests also increasingly reject
+grey line marks significance level 0.05. When the significance level
+gets smaller, it is possible to see that the visual tests reject
+(nearly) 100\% of the time that the conventional test rejects. However,
+there is no agreement, because the visual tests also increasingly reject
 residual plots where the conventional test failed to reject. Plot B is
@@
 instructs participants' browsers to render an environment for running
-behavioral experiments. During the experiment, this framework will
-automatically collect common behavioral data such as response time and
+behavioural experiments. During the experiment, this framework will
+automatically collect common behavioural data such as response time and
 clicks on buttons. participants' responses are first validated by a

### 2023-05-04 coverletter/coverletter.Rmd
final changes

@@
 fromphone: +61 3 9905 2608
-fromsignature: sigfile.png
+fromsignature: signature.png
 toname: Editors
@@
 The difference was even bigger than reported in the validation of visual inference study conducted by Majumder, Hofmann and Cook (2013) in JASA, where the comparison was on regression coefficients.
-So we initially thought, oh this is bad for visual inference, but actually it's good for visual evaluation of residual plots and bad for conventional testing.
-The paper explains why this is.
-It should be of interest to readers, especially for analysts interested in diagnosing their models.
+While we initially thought this is bad for visual inference, on further investigation we realised this provides empirical support for the wide-spread recommendation of visual evaluation of residual plots.
+The paper should be of interest to readers, especially useful for model diagnostics.
 

### 2023-05-04 paper.Rmd
final changes

@@
 It can also help in an obverse situation where a conventional test would fail to detect a problem with a model due to contaminated data.
-The lineup protocol also serves as an omnibus test, capable of detecting a range of departures from good residuals simultaneously.
+The lineup protocol also detects a range of departures from good residuals simultaneously.
 keywords: |
@@
 The SW test was computed using the `shapiro.test` from the core R package `stats`.
-^[Although we didn't use it, it's useful to know that the R package `skedastic` [@skedastic] also contains a large collection of functions to test for heteroskedasticity.] Although, the RESET test is exact, it requires the selection of a power parameter.
+^[Although we didn't use it, it's useful to know that the R package `skedastic` [@skedastic] also contains a large collection of functions to test for heteroskedasticity.] The RESET test requires the selection of a power parameter.
 @ramsey1969tests recommends a power of four, which we adopted in our analysis.
-The BP and SW tests are approximate.
 
@@
 Scenario 3 is the most feasible to apply, but has the most dependencies to accommodate for the $p$-value calculation.
-For Scenario 3, VanderPlas et al propose modelling the probability of plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$.
+For Scenario 3, VanderPlas et al propose modelling the probability of plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim \text{Dirichlet}(\alpha)$ for $i=1,...,m$ and $\alpha > 0$.
 The number of times plot $i$ being selected in $K$ evaluations is denoted as $c_i$.
@@
 VanderPlas et al define $c$-interesting plot to be if $c$ or more participants select the plot as the most different.
-The expected number of plots selected at least $c$ times, $E[Z_c]$, is then calculated as
+The expected number of plots selected at least $c$ times, $\mathrm{E}[Z_c]$, is then calculated as
 
@@
 \begin{equation} \label{eq:c-interesting-expectation}
-E[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}
+\mathrm{E}[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}
 ```
@@
 
-Since $supp(X) = [-1, 1]$, choosing $a$ to be $-1$, $0$ and $1$ can generate "left-triangle", "butterfly" and "right-triangle" shapes as displayed in Figure \ref{fig:different-shape-of-heter}.
+Since $\text{supp}(X) = [-1, 1]$, choosing $a$ to be $-1$, $0$ and $1$ can generate "left-triangle", "butterfly" and "right-triangle" shapes as displayed in Figure \ref{fig:different-shape-of-heter}.
 The term $(2 - |a|)$ maintains the magnitude of residuals across different values of $a$.
@@
 
-a_labels <- c("left-triangle", "butterfly", "right-triangle")
+a_labels <- c("Left-triangle", "Butterfly", "Right-triangle")
 
@@
 - normal, $N(0, 0.3^2)$, 
-- skewed, $lognormal(0, 0.6^2)/3$, and 
+- skewed, $\text{lognormal}(0, 0.6^2)/3$, and 
 - discrete, $U\{1, 5\}$.
@@
 \begin{equation}
-E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}\sigma^2))^{-1}\boldsymbol{\mu}_z\right)
+E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(\text{diag}(\boldsymbol{R}\sigma^2))^{-1}\boldsymbol{\mu}_z\right)
 \label{eq:nonlin-effect-size}
@@
 
-\noindent where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,
+\noindent where $\text{diag}(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,
 $\boldsymbol{R} = \boldsymbol{I}_n - \boldsymbol{H}$ is the residual operator, $\boldsymbol{H} = \boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the hat matrix, $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals where $\boldsymbol{Z}$ contain any higher order terms of $\boldsymbol{X}$ left out of the regression equation, $\boldsymbol{\beta}_z$ contains the corresponding coefficients, and $\sigma^2\boldsymbol{I}$ is the assumed covariance matrix of the error term when $H_0$ is true.
@@
 \begin{equation}
-E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|diag(\boldsymbol{R})|} - n + tr(diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}diag(\boldsymbol{R}))\right)
+E = \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R})|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\text{diag}(\boldsymbol{R}))\right)
 \label{eq:hetero-effect-size}
@@
 Neither the attention checks nor null lineups were used in the subsequent analysis.
-The de-identified data, `vi_survey`, is made available in the R package, `visage`.^[The R package `visage` is currently available on GitHub (https://github.com/TengMCing/visage).]
+The de-identified data, `vi_survey`, is made available in the R package, `visage`.
 
@@
   geom_line(data = pred_visual_poly_boot,
-            aes(log_effect_size, power, col = "Visual test", group = boot_id),
+            aes(log_effect_size, power, col = "visual test", group = boot_id),
             size = 1,
@@
   geom_line(data = pred_visual_poly,
-            aes(log_effect_size, power, col = "Visual test"), 
+            aes(log_effect_size, power, col = "visual test"), 
             size = 1) +
@@
   geom_line(data = pred_visual_heter_boot,
-            aes(log_effect_size, power, col = "Visual test", group = boot_id),
+            aes(log_effect_size, power, col = "visual test", group = boot_id),
             size = 1,
@@
   geom_line(data = pred_visual_heter,
-            aes(log_effect_size, power, col = "Visual test"), 
+            aes(log_effect_size, power, col = "visual test"), 
             size = 1) +
@@
 p_value_cmp_dat %>%
+  mutate(type = ifelse(type == "non-linearity", "Non-linearity", "Heteroskedasticity")) %>%
+  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity"))) %>%
   ggplot() +
@@
 
-In regression analysis, predictions are conditional on the observed values of the predictors, that is, the conditional mean of the dependent variable $Y$ given the value of the independent variable $X$, $E(Y|X)$.
+In regression analysis, predictions are conditional on the observed values of the predictors, that is, the conditional mean of the dependent variable $Y$ given the value of the independent variable $X$, $\text{E}(Y|X)$.
 This is an often forgotten element of regression analysis but it is important.
@@
   mutate(log_effect_size = log(effect_size)) %>%
-  mutate(type = factor(stringr::str_to_title(type), 
-                       levels = c("Non-Linearity", "Heteroskedasticity")))
+  mutate(type = factor(ifelse(type == "non-linearity", "Non-linearity", "Heteroskedasticity"), 
+                       levels = c("Non-linearity", "Heteroskedasticity")))
 ```
@@
   mutate(x_dist = ifelse(x_dist == "even_discrete", 
-                         "Discrete", 
-                         stringr::str_to_title(x_dist))) %>%
-  mutate(type = factor(stringr::str_to_title(type), 
-                       levels = c("Non-Linearity", 
-                                  "Heteroskedasticity"))) %>%
+                         "discrete", 
+                         x_dist)) %>%
+  mutate(type = factor(ifelse(type == "non-linearity", "Non-linearity", "Heteroskedasticity"), 
+                       levels = c("Non-linearity", "Heteroskedasticity"))) %>%
   mutate(x_dist = factor(x_dist, 
-                         levels = c("Uniform", 
-                                    "Normal", 
-                                    "Lognormal", 
-                                    "Discrete"))) %>%
+                         levels = c("uniform", 
+                                    "normal", 
+                                    "lognormal", 
+                                    "discrete"))) %>%
 
@@
   geom_line(data = dat_uniform_5_pred,
-            aes(log_effect_size, power, col = "Uniform", group = boot_id),
+            aes(log_effect_size, power, col = "uniform", group = boot_id),
             size = 1,
@@
 It includes more details about the experimental setup, the derivation of the effect size, the effect of data collection period, and the estimate of $\alpha$.
+
+The R package `visage` is currently available at https://github.com/TengMCing/visage.

### 2023-05-04 paper.tex
final changes

@@
 conventional test would fail to detect a problem with a model due to
-contaminated data. The lineup protocol also serves as an omnibus test,
-capable of detecting a range of departures from good residuals
-simultaneously.
+contaminated data. The lineup protocol also detects a range of
+departures from good residuals simultaneously.
 \end{abstract}
@@
 
-We can assess residuals in multiple ways. To examine the their
-univariate distribution, residuals may be plotted as a histogram or
-normal probability plot. Using the classical normal linear regression
-model as an example, if the distribution is symmetric and unimodal, we
-would consider it to be well-behaved. However, if the distribution is
-skewed, bimodal, multimodal, or contains outliers, there would be cause
-for concern. We can also inspect the distribution by conducting a
+We can assess residuals in multiple ways. To examine the univariate
+distribution, residuals may be plotted as a histogram or normal
+probability plot. Using the classical normal linear regression model as
+an example, if the distribution is symmetric and unimodal, we would
+consider it to be well-behaved. However, if the distribution is skewed,
+bimodal, multimodal, or contains outliers, there would be cause for
+concern. We can also inspect the distribution by conducting a
 goodness-of-fit test, such as the Shapiro-Wilk normality test
@@
 correctly judging whether NO pattern exists in a residual plot is a
-difficult task for humans. Humans will almost always will see a pattern
-(see \citet{kahneman2011thinking}), so the question that really needs
+difficult task for humans. Humans will almost always see a pattern (see
+\citet{kahneman2011thinking}), so the question that really needs
 answering is whether any pattern perceived is consistent with
@@
 
-The paper is structured as follows. The next section describes the
-background on the types of departures that one expects to detect, and
-outlines a formal statistical process for reading residual plots, called
-visual inference. Section \ref{experimental-design} details the
-experimental design to compare the decisions made by formal hypothesis
-testing, and how humans would read diagnostic plots. The results are
-reported in Section \ref{results}. We conclude with a discussion of the
-presented work, and ideas for future directions.
+The paper is structured as follows. Section \ref{background} describes
+the background on the types of departures that one expects to detect,
+and outlines a formal statistical process for reading residual plots,
+called visual inference. Section \ref{significance-calculation}
+describes the calculation of the statistical significance and power of
+the test. Section \ref{experimental-design} details the experimental
+design to compare the decisions made by formal hypothesis testing, and
+how humans would read diagnostic plots. The results are reported in
+Section \ref{results}. We conclude with a discussion of the presented
+work, and ideas for future directions.
 
@@
 Model misspecification occurs if functions of predictors that needed to
-accurately describe the relationship between response and predictors are
-omitted. Any non-linear pattern visible in the residual plot could be
-indicative of this problem. An example residual plot containing visual
-pattern of non-linearity is shown in Figure
-\ref{fig:residual-plot-common-departures}B. One can clearly observe the
-``S-shape'' from the residual plot, which corresponds to cubic term that
-should have been included in the model.
+accurately describe the relationship with the response are incorrectly
+specified. This includes instances where a higher-order polynomial term
+of a predictor is wrongfully omitted. Any non-linear pattern visible in
+the residual plot could be indicative of this problem. An example
+residual plot containing visual pattern of non-linearity is shown in
+Figure \ref{fig:residual-plot-common-departures}B. One can clearly
+observe the ``S-shape'' from the residual plot, which corresponds to
+cubic term that should have been included in the model.
 
@@
 distribution. To test specific forms of non-linearity, one may apply the
-F-test as a model structural test to examine the significance of
+F-test as a model structural test to examine the significance of a
 specific polynomial and non-linear forms of the predictors, or the
@@
   that the R package \texttt{skedastic} \citep{skedastic} also contains
-  a large collection of functions to test for heteroskedasticity.}
-Although, the RESET test is exact, it requires the selection of a power
-parameter. \citet{ramsey1969tests} recommends a power of four, which we
-adopted in our analysis. The BP and SW tests are approximate.
-
-We would expect the RESET test for non-linearity to reject residual plot
+  a large collection of functions to test for heteroskedasticity.} The
+RESET test requires the selection of a power parameter.
+\citet{ramsey1969tests} recommends a power of four, which we adopted in
+our analysis.
+
+For residual plots in Figure \ref{fig:residual-plot-common-departures},
+we would expect the RESET test for non-linearity to reject residual plot
 B, the BP test for heteroskedasticity to reject the residual plot C, and
@@
 exhibiting structure that they weren't designed for.
-\citet{cook1982residuals} explain that most residual-based tests for
-particular types of departure from model assumptions are also sensitive
-to other types of departures. This could be considered a Type III error
+\citet{cook1982residuals} explain that most residual-based tests for a
+particular departure from the model assumptions are also sensitive to
+other types of departures. This could be considered a Type III error
 \citep{kimball1957errors}, where the null hypothesis of good residuals
@@
 departure can have elements of other types of departure, for example,
-non-linearity could be viewed as heteroskedasticity. Additionally, other
+non-linearity can appear like heteroskedasticity. Additionally, other
 data problems such as outliers can trigger rejection (or not) of the
@@
 can feel subjective. Ideally we would expect visual evaluation to be
-done objectively with small amounts of human-to-human variation in
-reading data plots, but we cannot assume this is the case.
+done objectively with low human-to-human variation in reading data
+plots, but we cannot assume this is the case.
 
@@
 authors over-interpreted separation between gene groups from a
-two-dimensional projection of a linear discriminant analysis.
-\citet{roy2015using} showed that there were no differences in the
-expression levels between the gene groups. One solution to
-over-interpretation is to examine the plot in the context of natural
-sampling variability assumed by the model, called the lineup protocol,
-as proposed in \citet{buja2009statistical}.
+two-dimensional projection of a linear discriminant analysis even when
+there were no differences in the expression levels between the gene
+groups. One solution to over-interpretation is to examine the plot in
+the context of natural sampling variability assumed by the model, called
+the lineup protocol, as proposed in \citet{buja2009statistical}.
 \citet{majumder2013validation} showed that the lineup protocol is
@@
 statistic is the best justified reference distribution among the three
-\citep{buja2009statistical}. Essentially, null residuals can be
-simulated by regressing \(N\) i.i.d standard normal random draws on the
-predictors, then re-scaling it by the ratio of residual sum of square in
-two regressions.
+\citep{buja2009statistical}. Under this method, the null residuals can
+essentially be simulated by independent drawing from a standard normal
+random distribution, then regressing the draws on the predictors, and
+then re-scaling it by the ratio of the residual sum of square in two
+regressions.
 
@@
 a similar power to a t-test, though they express that in general it is
-unnecessary to use visual inference if there exists a conventional test,
-and they do not expect the visual test to perform equally well as the
-conventional test. In their third experiment, where there is not a
-conventional test, visual test outperforms the conventional test by a
-large margin. This is encouraging, as it supports the use of visual
-inference in situations where there are no existing statistical testing
-procedures. Visual inference has also been integrated into diagnostics
-for hierarchical linear models where the lineup protocol is used to
-judge the assumptions of linearity, normality and constant error
-variance for both the level-1 and level-2 residuals
-\citep[\citet{loy2014hlmdiag} and
+unnecessary to use visual inference if there exists a corresponding
+conventional test, and they do not expect the visual test to perform
+equally well as the conventional test. In their third experiment, where
+the contamination of the data violate the assumptions of the convetional
+test, visual test outperforms the conventional test by a large margin.
+This is encouraging, as it supports the use of visual inference in
+situations where there are no existing numerical testing procedures.
+Visual inference has also been integrated into diagnostics for
+hierarchical linear models where the lineup protocol is used to judge
+the assumptions of linearity, normality and constant error variance for
+both the level-1 and level-2 residuals \citep[\citet{loy2014hlmdiag} and
 \citet{loy2015you}]{loy2013diagnostic}.
 
-\hypertarget{calculation-of-statistical-significance-and-test-power}{%
+\hypertarget{significance-calculation}{%
 \section{Calculation of statistical significance and test
-power}\label{calculation-of-statistical-significance-and-test-power}}
+power}\label{significance-calculation}}
 
@@
 modelling the probability of plot \(i\) being selected from a lineup as
-\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
+\(\theta_i\), where \(\theta_i \sim \text{Dirichlet}(\alpha)\) for
 \(i=1,...,m\) and \(\alpha > 0\). The number of times plot \(i\) being
@@
 more participants select the plot as the most different. The expected
-number of plots selected at least \(c\) times, \(E[Z_c]\), is then
-calculated as
+number of plots selected at least \(c\) times, \(\mathrm{E}[Z_c]\), is
+then calculated as
 
 \begin{equation} \label{eq:c-interesting-expectation}
-E[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}
+\mathrm{E}[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}
 
@@
 result in a curve with more turning points. It is expected that the
-``U'' shape will be the easiest detect, as the shape gets more complex
-it will be harder to perceive in a scatterplot, particularly when there
-is noise. Figure \ref{fig:different-sigma} shows the ``U'' shape for
-different amounts of noise (\(\sigma\)).
+``U'' shape will be the easiest to detect, as the shape gets more
+complex it will be harder to perceive in a scatterplot, particularly
+when there is noise. Figure \ref{fig:different-sigma} shows the ``U''
+shape for different amounts of noise (\(\sigma\)).
 
@@
 
-Since \(supp(X) = [-1, 1]\), choosing \(a\) to be \(-1\), \(0\) and
-\(1\) can generate ``left-triangle'', ``butterfly'' and
+Since \(\text{supp}(X) = [-1, 1]\), choosing \(a\) to be \(-1\), \(0\)
+and \(1\) can generate ``left-triangle'', ``butterfly'' and
 ``right-triangle'' shapes as displayed in Figure
@@
 \item
-  skewed, \(lognormal(0, 0.6^2)/3\), and
+  skewed, \(\text{lognormal}(0, 0.6^2)/3\), and
 \item
@@
 
-\noindent where \(diag(.)\) is the diagonal matrix constructed from the
-diagonal elements of a matrix,
+\noindent where \(\text{diag}(.)\) is the diagonal matrix constructed
+from the diagonal elements of a matrix,
 \(\boldsymbol{R} = \boldsymbol{I}_n - \boldsymbol{H}\) is the residual
@@
 the subsequent analysis. The de-identified data, \texttt{vi\_survey}, is
-made available in the R package, \texttt{visage}.\footnote{The R package
-  \texttt{visage} is currently available on GitHub
-  (\url{https://github.com/TengMCing/visage}).}
+made available in the R package, \texttt{visage}.
 
@@
 variable \(Y\) given the value of the independent variable \(X\),
-\(E(Y|X)\). This is an often forgotten element of regression analysis
-but it is important. Where \(X\) is observed, the distribution of the
-\(X\) values in the sample, or consequently \(\hat{Y}\), may affect the
-ability to read any patterns in the residual plots. The effect of fitted
-value distribution on test performance is assess using four different
-distributions of fitted values: uniform, normal, discrete and lognormal
-(skewed). We expect that if \(\hat{Y}\) has a uniform distribution, it
-is easier to read the relationship with the residuals.
+\(\text{E}(Y|X)\). This is an often forgotten element of regression
+analysis but it is important. Where \(X\) is observed, the distribution
+of the \(X\) values in the sample, or consequently \(\hat{Y}\), may
+affect the ability to read any patterns in the residual plots. The
+effect of fitted value distribution on test performance is assess using
+four different distributions of fitted values: uniform, normal, discrete
+and lognormal (skewed). We expect that if \(\hat{Y}\) has a uniform
+distribution, it is easier to read the relationship with the residuals.
 
@@
 
+The R package \texttt{visage} is currently available at
+\url{https://github.com/TengMCing/visage}.
+
 \bibliographystyle{tfcad}

### 2023-08-07 JCGS_response_to_reviews/response_to_review.Rmd
init response to reviews

---
title: "Response to Reviews: JCGS-23-139"
author: Weihao Li, Emi Tanaka, Dianne Cook, Susan VanderPlas
output: pdf_document
date: "2023-08-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Thank you for the careful reviews. What follows is our point-by-point response to reviewer comments.

Note that *The original reviewer comments are in italic* and our response is normal text.

# Reviewer 1

*Comments to the Author*

*I did not need any convincing that residual diagnostic plots are superior to tests but this paper presents a narrower argument to support this claim than many analysts would employ. For example, diagnostic plots may reveal features which are not directly connected to a linear model assumption and yet are interesting and important to identify. Of course, it is difficult to justify a journal article on the basis of such observations. In this respect, I am impressed that you have succeeded in demonstrating the superiority within the more formal framework of the lineup protocol. You have restricted attention to the residual-fitted plot and a particular selection of formal tests. I accept that this was necessary to make definite progress. Nevertheless, it would be good to add some discussion of the wider picture.*

*1. "The ubiquity of this advice is curious: investigating why, is the subject of this paper. " I don't find this advice curious - it seems almost universal. I have taught this topic for many years and give the same advice. I explain why I give this advice and I believe several other authors do the same. Their reasons are wider in scope than you express in this article.*

*2. All your examples are simulated data and very obviously so. Real data shows much greater irregularity - the points will usually not be uniformly spaced along the x-axis. The response may be somewhat discrete or have other features that will distinguish it from bland randomness. This would seem to offer greater challenges to your lineup idea as generating data under the null will be much more challenging. If we simulate from the null model, the real data will be very obvious. You have done an experiment but the reality would be more like an observational study. This limits the scope of your conclusions which you should clearly acknowledge.*

*3. It is true that formal diagnostic tests tend to be too sensitive and yet we don't have a definite sense of how much divergence from the assumed model would cause a significant problem. How much non-normality in the residuals is acceptable? We don't really know. Formal tests give us definite answers about violations of assumptions which we might find hard to ignore but diagnostic plots give us license to sweep inconvenient features under the carpet. Is there really an advantage to the diagnostic plot in this case other than the opportunity to hand-wave about our answer?*

*4. I use something like the lineup protocol in teaching to help students calibrate their assessment of diagnostic plots. But I doubt very much experienced analysts use them in practice. I do not think it is realistic or really that helpful to recommend their everyday use.*

# Reviewer 2

*Summary*

*This article uses the lineup protocol in a visual inference experiment, which puts a residual plot in the context of null plots. Focus is mostly on detecting nonlinearities and heteroskedasticity. Comparisons with some classic tests are given.*

*Overall Comments*

*Overall, this paper is very well written and interesting. However, I have some serious reservations about the practicality of the procedure, which I articulate in greater detail below. I also think the paper is currently too long, and that a significant chunk of material should be shifted to the Appendix.*

*1. As mentioned above, the paper is too long. While the onus should be on the authors for getting the page number down, some possible suggestions could be combining Sections 1 and 2 while moving some of the historical context (which I fully acknowledge is quite interesting) to the Appendix, and moving some of the technical details of Section 3.2 to the Appendix.*

*2. The practical implementation of this as a procedure is, in my opinion, the biggest weakness. The authors acknowledge the costliness of human evaluation of residuals, but not until Page 33, Line 47. I find this a tad disingenuous as this will likely jump out to the reader as being a problem right from the beginning of reading this paper (it did for me). So I think being upfront about this limitation, and providing some practical considerations to address this concern will assuage my concerns. There might be some concepts from human-computer interaction that could be effective for operationalizing and improving the efficiency of the present work.*

*3. I think a better connection needs to be made with where this computationally expensive process could potentially pay off. For example, applying the lineup protocol to, say, any of the data examples in the texts cited by the authors would amount to nothing more than an academic exercise. But, what sort of data problems can the authors point to and definitively claim that the lineup protocol will be invaluable at ensuring we visually capture possible nonlinearities or heteroskedasticity? I am not doubting the efficacy of the approach, but I’d like to see a stronger case made with its practical application.*

*4. What should be claimed about the sample size n as a limitation? Take in point the comments at the bottom of page 19. Am I to understand that the lineup protocol will not be appropriate for big data problems?*

*5. At the beginning of Section 4.2.3, what cultural biases do the authors anticipate by restricting the participants to be “fluent in English”? Granted this is likely done as a matter of convenience, but it would be good to articulate how cultural biases may be present in interpreting a lineup protocol.*

*6. I would like the authors to be a little clearer about their specific contributions. I felt as if the biggest contribution in this work was the experiments that they did. A lot of references are included for work that is applied in this manuscript, so at times I got lost as to what already existed in the literature and what “new” material the authors were giving the reader.*

*Minor Comments*

*1. Page 9, Line 13: Since the three references are by the same authors, use \\cite{Ref1,Ref2,Ref3}.*

*2. Page 12, Line 7: $\alpha$.*

*3. Why not write the power formula in terms of a generic significance level? Granted, one should be mindful of an abuse of notation if using $\alpha$ as it has been introduced in the Rorschach setup of the previous subsection.*

*4. Page 13, Line 14: "...the scenario..."*

*5. Page 20, Line 36: "...variety of difficulty..."*

*6. In Figures 11 and 12, I am not sure the example residual plots beneath the x-axis are all that terribly helpful. In fact, I think they all look mostly similar, and am not entirely convinced that they add any value. Can the authors convince me otherwise?*

### 2023-08-18 JCGS_response_to_reviews/response_to_review.Rmd
paper reviewer 1 (1) update

@@
 
+Thank for this feedback.
+We are happy to see that, like many people, you don't need convincing and that you are glad to see that we have succeeded in demonstrating the superiority of residual diagnostic plots.
+
+We have added a broader discussion on the use of residual plots in Section XXX, beyond those tested in this work.
+(Need to include: (1) finding interesting feature that was not intended to test for; (2) other residual plots - need to make a comprehensive list of common residual plots.)
+
 *1.
@@
 Their reasons are wider in scope than you express in this article.*
+
+We have replaced the "is curious" to better acknowledge that many practitioners already know this. (lack of empirical evidence for the advice is curious)
 
@@
 
+The work very specifically uses simulated data because we need to study a problem in controlled conditions.
+You are right, that data will pose more interesting challenges, and this can be in two directions.
+One, being that the lineup protocol will allow interesting and unexpected patterns in the residuals to be detected (see Buja et al and Wickham et al for examples).
+The other is that the current process for generating null data uses residual rotation and will likely maintain irrelevant quirks of the data in the null plots.
+These should not inadvertently allow the reader to detect the data plot based on quirks.
+
+We have addressed this criticism in the broader discussion of the use of residual plots and lineups for residual diagnostics added to Section XXX (see issue #11).
+
+An observational study will hopefully be possible once a computer vision model is deployed that will allow analysts to upload their residuals for automated testing.
+
 *3.
@@
 
+When the diagnostic plots are embedded in a lineup of nulls, it does have the advantage of providing a definitive answer, and protects against the subjectiveness of hand-waving.
+We are advocating that you should always calibrate the diagnostic plot with a comparison with null plots.
+We have strengthened this point of view by adding a sentence in the introduction, in addition to what is provided in the Conclusions.
+
+(Is there a paper to refer the reader to?)
+
 *4.
@@
 I do not think it is realistic or really that helpful to recommend their everyday use.*
+
+It can be argued that it is being used by practitioners based on download rates of the nullabor package, and applications articles that report usage.
+We agree that it is not widespread, and think that this is because it's a little unwieldy.
+The goal is to make it easier, particularly by employing an automated testing tool using computer vision, but this is not the subject of the current paper.
+
+We have added a sentence to Section XXX to make this clearer.
 

### 2023-08-18 paper.Rmd
double space
paper reviewer 1 (1) update

@@
   \def\tightlist{}
+  \usepackage{setspace}
+  \doublespacing
 output: rticles::tf_article
@@
 If there are any visually discoverable patterns, the model is potentially inadequate or incorrectly specified.
-In general, one looks for noticeable departures from the model such as non-linear pattern or heteroskedasticity.
+**Interesting features that are not directly connected to a linear model assumption can also be found from residual plots, such as the discreteness of the fitted values, outliers and influential points.** In general, one looks for noticeable departures from the model such as non-linear pattern or heteroskedasticity.
 A non-linear pattern would suggest that the model needs to have some additional non-linear terms.
@@
 @breusch1979simple].
+
+**There are other types of residual plot to extract the available information and identify inappropriate assumptions, like the added variable plot for studying the role of an external variable if it enters linearly into a model.
+Partial residual plots have been suggested as cornputationally convenient substitutes for the added variable plots.
+For use in diagnostic procedures, several transformations of the ordinary residuals have been suggested to overcome partially some of their shortcomings.
+Studentized residuals have null distributions that are independent of the scale parameters.
+These residuals are shown to be closely related to a mean shift model for outliers and to the residuals obtained when each case in turn is left out of the data.
+Alternatively.
+the residuals can be transformed to have a selected covariance structure: The usual suggestion is to obtain residuals with uncorrelated elements.**
 

### 2023-08-25 JCGS_response_to_reviews/response_to_review.Rmd
update responses

@@
 
-We have added a broader discussion on the use of residual plots in Section XXX, beyond those tested in this work.
-(Need to include: (1) finding interesting feature that was not intended to test for; (2) other residual plots - need to make a comprehensive list of common residual plots.)
+We have added more explanation of the use of residual plots in the Introduction and we have added a section Limitations and practicality where the broader picture is discussed.
 
@@
 
-We have replaced the "is curious" to better acknowledge that many practitioners already know this. (lack of empirical evidence for the advice is curious)
+We have changed the language to focus on that is the advice is given despite the lack of empirical evidence.
 
@@
 
+Thanks for the positive feedback. We have shortened the paper by XXX pages. We have expanded the explanations of why the practicality of the procedure should not be questioned in Section XXX.
+
 *1.
@@
 While the onus should be on the authors for getting the page number down, some possible suggestions could be combining Sections 1 and 2 while moving some of the historical context (which I fully acknowledge is quite interesting) to the Appendix, and moving some of the technical details of Section 3.2 to the Appendix.*
+
+See the response above, thanks for your suggestions. We have moved most of Section 3.2 Statistical significance to the Appendix. (how we shorten the rest of the paper XXX)
+
+- Part of Experimental design -> appendix
+- Part of Explanations of the lineup (p-value) -> appendix
+- Shorten the Section 1 and 2 (more precise)
+- Modify the discussion and conclusion to address the issues raised by reviews
 
@@
 
+The new Section Limitations and practicality addresses these concerns and we have added a sentence in Introduction so that the reader will not be concerned upfront.
+
 *3.
@@
 
+As mentioned above,
+
+This is not computationally expensive. The conventional test also make mistakes. You need something more than the conventional test. And you also need something more than the residual plot.
+
+check earlier paper.
+
 *4.
@@
 
+This is a display issue not a lineup protocol issue.
+
 *5.
@@
 Granted this is likely done as a matter of convenience, but it would be good to articulate how cultural biases may be present in interpreting a lineup protocol.*
+
+The instruction of this study was in English.
+very unlikely there will be a difference between English speaker and non-English speaker.
 
@@
 A lot of references are included for work that is applied in this manuscript, so at times I got lost as to what already existed in the literature and what “new” material the authors were giving the reader.*
+
+We have clarified the language of the Introduction and the Conclusion.
 

### 2023-08-25 paper.Rmd
update responses

@@
 Heteroskedasticity suggests that the error is dependent on the predictors, and hence violates the independence assumption.
+
 However, correctly judging whether NO pattern exists in a residual plot is a difficult task for humans.
@@
 
-**There are other types of residual plot to extract the available information and identify inappropriate assumptions, like the added variable plot for studying the role of an external variable if it enters linearly into a model.
-Partial residual plots have been suggested as cornputationally convenient substitutes for the added variable plots.
+**There are other types of residual plot to extract the available information and identify inappropriate assumptions, like the added variable plot (cite here) for studying the role of an external variable if it enters linearly into a model.
+Partial residual plots (cite here) have been suggested as computationally convenient substitutes for the added variable plots.
 For use in diagnostic procedures, several transformations of the ordinary residuals have been suggested to overcome partially some of their shortcomings.
-Studentized residuals have null distributions that are independent of the scale parameters.
+Studentized residuals (cite here) have null distributions that are independent of the scale parameters.
 These residuals are shown to be closely related to a mean shift model for outliers and to the residuals obtained when each case in turn is left out of the data.
@@
 
-Linear regression is a well-established procedure, and there is considerable literature describing diagnostic procedures, e.g.
+There is sustainable literature describing diagnostic procedures for linear models, e.g.
 @draper1998applied, @montgomery1982introduction, @belsley1980regression, @cook1999applied and @cook1982residuals.
@@
 \noindent The common wisdom of experts is that plotting the residuals is indispensable for diagnosing model fits.
-The ubiquity of this advice is *curious*: investigating why, is the subject of this paper.
+The lack of empirical evidence for the ubiquitous advice is *curious*.
+The research study discussed in this paper fills this gap.
 
@@
 
+# Limitations and practicality
+
+- explain how the lineup protocol can be cheaply done with the R package "nullabor".
+everyone can do it easily.
+The cost of using lineup in stat education under classroom setting is cheap (adam loy teaching paper)
+- Formal tests.
+reliable results.
+crowd sourcing is not expensive compared to the cost other devices in the lab and is not entirely unreasonable.
+Only if there is grey area (you can not confidently detect the data).
+- The possiblity of computer vision in the future.
+
+- repeat the scope of the study
+- list other things we could have tested
+- explain why the result should apply to other residual diagnostics settings 
+- finding interesting feature that was not intended to test for
+
 # Conclusions

### 2023-08-25 paper.Rmd
formatting the paper

@@
   \usepackage{setspace}
-  \doublespacing
+  % \doublespacing
 output: rticles::tf_article
@@
 If there are any visually discoverable patterns, the model is potentially inadequate or incorrectly specified.
-**Interesting features that are not directly connected to a linear model assumption can also be found from residual plots, such as the discreteness of the fitted values, outliers and influential points.** In general, one looks for noticeable departures from the model such as non-linear pattern or heteroskedasticity.
+\textcolor{red}{Interesting features that are not directly connected to a linear model assumption can also be found from residual plots, such as the discreteness of the fitted values, outliers and influential points.} In general, one looks for noticeable departures from the model such as non-linear pattern or heteroskedasticity.
 A non-linear pattern would suggest that the model needs to have some additional non-linear terms.
@@
 
-**There are other types of residual plot to extract the available information and identify inappropriate assumptions, like the added variable plot (cite here) for studying the role of an external variable if it enters linearly into a model.
+\textcolor{red}{There are other types of residual plot to extract the available information and identify inappropriate assumptions, like the added variable plot (cite here) for studying the role of an external variable if it enters linearly into a model.
 Partial residual plots (cite here) have been suggested as computationally convenient substitutes for the added variable plots.
@@
 Alternatively.
-the residuals can be transformed to have a selected covariance structure: The usual suggestion is to obtain residuals with uncorrelated elements.**
+the residuals can be transformed to have a selected covariance structure: The usual suggestion is to obtain residuals with uncorrelated elements.}
 

### 2023-08-25 paper.tex
double space
formatting the paper

@@
 \def\tightlist{}
+\usepackage{setspace}
+% \doublespacing
 
@@
 relationships. If there are any visually discoverable patterns, the
-model is potentially inadequate or incorrectly specified. In general,
-one looks for noticeable departures from the model such as non-linear
-pattern or heteroskedasticity. A non-linear pattern would suggest that
-the model needs to have some additional non-linear terms.
+model is potentially inadequate or incorrectly specified.
+\textcolor{red}{Interesting features that are not directly connected to a linear model assumption can also be found from residual plots, such as the discreteness of the fitted values, outliers and influential points.}
+In general, one looks for noticeable departures from the model such as
+non-linear pattern or heteroskedasticity. A non-linear pattern would
+suggest that the model needs to have some additional non-linear terms.
 Heteroskedasticity suggests that the error is dependent on the
-predictors, and hence violates the independence assumption. However,
-correctly judging whether NO pattern exists in a residual plot is a
-difficult task for humans. Humans will almost always see a pattern (see
-\citet{kahneman2011thinking}), so the question that really needs
+predictors, and hence violates the independence assumption.
+
+However, correctly judging whether NO pattern exists in a residual plot
+is a difficult task for humans. Humans will almost always see a pattern
+(see \citet{kahneman2011thinking}), so the question that really needs
 answering is whether any pattern perceived is consistent with
@@
 
-Linear regression is a well-established procedure, and there is
-considerable literature describing diagnostic procedures, e.g.
-\citet{draper1998applied}, \citet{montgomery1982introduction},
-\citet{belsley1980regression}, \citet{cook1999applied} and
-\citet{cook1982residuals}. Interestingly, despite the abundance of
-conventional tests, ALL of these writings advise that plotting residuals
-is an essential tool for diagnosing regression model problems:
+\textcolor{red}{There are other types of residual plot to extract the available information and identify inappropriate assumptions, like the added variable plot (cite here) for studying the role of an external variable if it enters linearly into a model.
+Partial residual plots (cite here) have been suggested as computationally convenient substitutes for the added variable plots.
+For use in diagnostic procedures, several transformations of the ordinary residuals have been suggested to overcome partially some of their shortcomings.
+Studentized residuals (cite here) have null distributions that are independent of the scale parameters.
+These residuals are shown to be closely related to a mean shift model for outliers and to the residuals obtained when each case in turn is left out of the data.
+Alternatively.
+the residuals can be transformed to have a selected covariance structure: The usual suggestion is to obtain residuals with uncorrelated elements.}
+
+There is sustainable literature describing diagnostic procedures for
+linear models, e.g. \citet{draper1998applied},
+\citet{montgomery1982introduction}, \citet{belsley1980regression},
+\citet{cook1999applied} and \citet{cook1982residuals}. Interestingly,
+despite the abundance of conventional tests, ALL of these writings
+advise that plotting residuals is an essential tool for diagnosing
+regression model problems:
 
@@
 \noindent The common wisdom of experts is that plotting the residuals is
-indispensable for diagnosing model fits. The ubiquity of this advice is
-\emph{curious}: investigating why, is the subject of this paper.
+indispensable for diagnosing model fits. The lack of empirical evidence
+for the ubiquitous advice is \emph{curious}. The research study
+discussed in this paper fills this gap.
 
@@
 
+\hypertarget{limitations-and-practicality}{%
+\section{Limitations and
+practicality}\label{limitations-and-practicality}}
+
+\begin{itemize}
+\item
+  explain how the lineup protocol can be cheaply done with the R package
+  ``nullabor''. everyone can do it easily. The cost of using lineup in
+  stat education under classroom setting is cheap (adam loy teaching
+  paper)
+\item
+  Formal tests. reliable results. crowd sourcing is not expensive
+  compared to the cost other devices in the lab and is not entirely
+  unreasonable. Only if there is grey area (you can not confidently
+  detect the data).
+\item
+  The possiblity of computer vision in the future.
+\item
+  repeat the scope of the study
+\item
+  list other things we could have tested
+\item
+  explain why the result should apply to other residual diagnostics
+  settings
+\item
+  finding interesting feature that was not intended to test for
+\end{itemize}
+
 \hypertarget{conclusions}{%

### 2023-08-28 paper.Rmd
upload limitation part

@@
 
-- explain how the lineup protocol can be cheaply done with the R package "nullabor".
-everyone can do it easily.
-The cost of using lineup in stat education under classroom setting is cheap (adam loy teaching paper)
-- Formal tests.
-reliable results.
-crowd sourcing is not expensive compared to the cost other devices in the lab and is not entirely unreasonable.
-Only if there is grey area (you can not confidently detect the data).
+One of the notable limitations of the lineup protocol lies in its reliance on human judgements.
+In this context, the effectiveness of the lineup protocol is closely connected with the perceptive ability and subjective interpretation of individual evaluators.
+This reliance on human judgment introduces an element of variability and subjectivity that can potentially impact the robustness of the results obtained.
+To ensure reliable outcomes, it's often necessary to include a diverse group of participants capable of offering unbiased judgements.
+This requirement has led researchers, including those in our study, to turn to crowd-sourcing platforms that offer access to a random selection of participants, which can help mitigate potential bias.
+
+It is worth noting that despite the need for participants with unbiased perspectives, the lineup protocol can be implemented at a relatively low cost.
+Leveraging tools like the R package `nullabor`, the process of constructing and evaluating lineups becomes accessible and straightforward.
+This accessibility empowers a wider range of researchers, even those without extensive statistical backgrounds, to employ the lineup protocol in their studies.
+Moreover, the cost-effectiveness of lineup-based analyses makes them particularly appealing for integration into statistical education within classroom settings (cite adam loy paper here).
+The practicality of lineup demonstrations can facilitate a deeper understanding of the concept of randomness and statistical testing among students.
+
+However, it's important to understand that if the goal is to attain results similar to those of formal hypothesis testing in rigor and reliability, using crowd-sourcing becomes necessary.
+The expense associated with crowd-sourcing is relatively reasonable when  compared to the costs of other laboratory devices, making it a sensible investment for research that requires clear and valuable findings.
+Importantly, this higher level of scrutiny is typically justified only when the data plot can not be confidently identified and requires a thorough investigation.
+
+Human evaluation of residuals is expensive, time-consuming and laborious.
+This is possibly why residual plot analysis is often not done in practice.
+However, with the emergence of effective computer vision, our work could lay a foundation for automated residual plot assessment.
+
 - The possiblity of computer vision in the future.

### 2023-08-28 paper.Rmd
update the limitation part

@@
 
-Human evaluation of residuals is expensive, time-consuming and laborious.
-This is possibly why residual plot analysis is often not done in practice.
-However, with the emergence of effective computer vision, our work could lay a foundation for automated residual plot assessment.
-
-- The possiblity of computer vision in the future.
-
-- repeat the scope of the study
-- list other things we could have tested
-- explain why the result should apply to other residual diagnostics settings 
-- finding interesting feature that was not intended to test for
+The study aims to showcase the utility of residual plots in model diagnostics while comparing them with conventional tests.
+Our primary focus is on employing residual-fitted plots within a controlled model environment to facilitate a human subject experiment.
+It's worth mentioning that there are other captivating aspects of residual diagnostics that are not covered within the scope of this paper.
+We consider the findings of this paper, can be applied to other types of statistical models.
+
+<!-- - The possiblity of computer vision in the future.  -->
+
+<!-- - repeat the scope of the study -->
+<!-- - list other things we could have tested -->
+<!-- - explain why the result should apply to other residual diagnostics settings  -->
+<!-- - finding interesting feature that was not intended to test for -->
 

### 2023-09-04 paper.Rmd
update limitation

--- 2023-08-28 paper.Rmd
@@
 
-Human evaluation of residuals is expensive, time-consuming and laborious.
-This is possibly why residual plot analysis is often not done in practice.
-However, with the emergence of effective computer vision, our work could lay a foundation for automated residual plot assessment.
-
-- The possiblity of computer vision in the future.
-
-- repeat the scope of the study
-- list other things we could have tested
-- explain why the result should apply to other residual diagnostics settings 
-- finding interesting feature that was not intended to test for
+Additionally, there is the exciting possibility of utilizing computer vision technology to automate the analysis of lineup of residual plots.
+This innovation has the potential to significantly reduce the expenses associated with recruiting human judges.
+By incorporating computer vision models, researchers can simplify the evaluation process, making it more efficient and cost-effective.
+
+The scope of this study is to highlight practical value of residual plots in model diagnostics, drawing comparisons with conventional tests.
+Our primary focus is on the use of residual-fitted plots in a controlled model environment for the purpose of a human subject experiment.
+It's important to note that there are other intriguing aspects of residual diagnostics that fall beyond the boundaries of this paper.
+For example, as mentioned in Introduction, residual plots can be used to explore interesting data characteristics while conducting model diagnostics.
+Furthermore, some model violations, such as serial correlation, necessitate the use of alternative residual plots like residual against row number plots.
+Nevertheless, we believe that the insights derived from this study have the potential to apply to other residual diagnostics setting, because the fundamental principles and statistical concepts underlying the lineup protocol and residual plots remain consistent across different contexts.
+In real-world situations, researchers often rely on visual assessment to make informed decisions about model validity.
+The lineup protocol's capability to identify deviations from randomness serves as a valuable aid in this scenario.
 

### 2023-09-04 paper.tex
update limitation

@@
 
-\begin{itemize}
-\item
-  explain how the lineup protocol can be cheaply done with the R package
-  ``nullabor''. everyone can do it easily. The cost of using lineup in
-  stat education under classroom setting is cheap (adam loy teaching
-  paper)
-\item
-  Formal tests. reliable results. crowd sourcing is not expensive
-  compared to the cost other devices in the lab and is not entirely
-  unreasonable. Only if there is grey area (you can not confidently
-  detect the data).
-\item
-  The possiblity of computer vision in the future.
-\item
-  repeat the scope of the study
-\item
-  list other things we could have tested
-\item
-  explain why the result should apply to other residual diagnostics
-  settings
-\item
-  finding interesting feature that was not intended to test for
-\end{itemize}
+One of the notable limitations of the lineup protocol lies in its
+reliance on human judgements. In this context, the effectiveness of the
+lineup protocol is closely connected with the perceptive ability and
+subjective interpretation of individual evaluators. This reliance on
+human judgment introduces an element of variability and subjectivity
+that can potentially impact the robustness of the results obtained. To
+ensure reliable outcomes, it's often necessary to include a diverse
+group of participants capable of offering unbiased judgements. This
+requirement has led researchers, including those in our study, to turn
+to crowd-sourcing platforms that offer access to a random selection of
+participants, which can help mitigate potential bias.
+
+It is worth noting that despite the need for participants with unbiased
+perspectives, the lineup protocol can be implemented at a relatively low
+cost. Leveraging tools like the R package \texttt{nullabor}, the process
+of constructing and evaluating lineups becomes accessible and
+straightforward. This accessibility empowers a wider range of
+researchers, even those without extensive statistical backgrounds, to
+employ the lineup protocol in their studies. Moreover, the
+cost-effectiveness of lineup-based analyses makes them particularly
+appealing for integration into statistical education within classroom
+settings (cite adam loy paper here). The practicality of lineup
+demonstrations can facilitate a deeper understanding of the concept of
+randomness and statistical testing among students.
+
+However, it's important to understand that if the goal is to attain
+results similar to those of formal hypothesis testing in rigor and
+reliability, using crowd-sourcing becomes necessary. The expense
+associated with crowd-sourcing is relatively reasonable when compared to
+the costs of other laboratory devices, making it a sensible investment
+for research that requires clear and valuable findings. Importantly,
+this higher level of scrutiny is typically justified only when the data
+plot can not be confidently identified and requires a thorough
+investigation.
+
+Additionally, there is the exciting possibility of utilizing computer
+vision technology to automate the analysis of lineup of residual plots.
+This innovation has the potential to significantly reduce the expenses
+associated with recruiting human judges. By incorporating computer
+vision models, researchers can simplify the evaluation process, making
+it more efficient and cost-effective.
+
+The scope of this study is to highlight practical value of residual
+plots in model diagnostics, drawing comparisons with conventional tests.
+Our primary focus is on the use of residual-fitted plots in a controlled
+model environment for the purpose of a human subject experiment. It's
+important to note that there are other intriguing aspects of residual
+diagnostics that fall beyond the boundaries of this paper. For example,
+as mentioned in Introduction, residual plots can be used to explore
+interesting data characteristics while conducting model diagnostics.
+Furthermore, some model violations, such as serial correlation,
+necessitate the use of alternative residual plots like residual against
+row number plots. Nevertheless, we believe that the insights derived
+from this study have the potential to apply to other residual
+diagnostics setting, because the fundamental principles and statistical
+concepts underlying the lineup protocol and residual plots remain
+consistent across different contexts. In real-world situations,
+researchers often rely on visual assessment to make informed decisions
+about model validity. The lineup protocol's capability to identify
+deviations from randomness serves as a valuable aid in this scenario.
 

### 2023-09-06 JCGS_response_to_reviews/response_to_review.Rmd
formatting the paper
add important notes
finished revising the paper

--- 2023-08-07 JCGS_response_to_reviews/response_to_review.Rmd
@@
 
+Thank for this feedback.
+We are happy to see that, like many people, you don't need convincing and that you are glad to see that we have succeeded in demonstrating the superiority of residual diagnostic plots.
+
+We have added more explanation about the use of residual plots in the Introduction (highlighted in red) and we have added a new section "Limitations and practicality" that provides some insight into the wider picture.
+
+<!-- (diagnostic plots may reveal features which are not directly connected to a linear model assumption; You have restricted attention to the residual-fitted plot and a particular selection of formal tests; Nevertheless, it would be good to add some discussion of the wider picture.)-->
+
 *1.
@@
 Their reasons are wider in scope than you express in this article.*
+
+We have changed the language.
 
@@
 
+The work very specifically uses simulated data because we need to study a problem in controlled conditions.
+You are right, though, that data will pose more interesting challenges, and this can be in two directions.
+One, being that the lineup protocol will allow interesting and unexpected patterns in the residuals to be detected (see Buja et al and Wickham et al for examples).
+The other is that the current process for generating null data uses residual rotation and will likely maintain irrelevant quirks of the data in the null plots.
+These should not inadvertently allow the reader to detect the data plot based on quirks.
+
+We have addressed this criticism in the broader discussion of the use of residual plots and lineups for residual diagnostics added to the new Section "Limitations and practicality".
+<!-- (see issue #11).-->
+
+An observational study involving actual residual plots will hopefully be possible once a computer vision model is deployed that will allow analysts to upload their residuals for automated testing.
+
 *3.
@@
 
+When the diagnostic plots are embedded in a lineup of nulls, it does have the advantage of providing a calibrated answer, and protects against the subjectiveness of hand-waving.
+We are advocating that you should always calibrate the diagnostic plot with a comparison with null plots.
+We have strengthened this point of view by adding more explanation in the Introduction, in addition to what is provided in the Conclusions.
+
+Having a definitive (numerical) answer to "how much is too much" would require more and different work.
+For our purposes, it was that the true residual plot could be detected from the null plots when the lineup was evaluated by K people.
+We've added a sentence to the Conclusion mentioning this.
+
+<!--(Is there a paper to refer the reader to?)-->
+
 *4.
@@
 I do not think it is realistic or really that helpful to recommend their everyday use.*
+
+It can be argued that it is being used by practitioners based on download rates of the nullabor package, and applications articles that report usage.
+Using it formally, we agree is cumbersome and would require engaging a crowd-sourcing service, that involves a small expense.
+Science labs typically have a budget for equipment and experimental devices and we would advocate this is a useful addition.
+We have definitely used it ourselves in 10.1007/s00180-014-0534-x for ecological data and  10.4172/2153-0602.1000139 for gene expression analysis.
+We have added some details in the "Limitations and practicality" section.
+<!--We agree that it is not widespread, and think that this is because it's a little unwieldy.
+The goal is to make it easier, particularly by employing an automated testing tool using computer vision, but this is not the subject of the current paper.-->
 
@@
 
+Thanks for the positive feedback. We have shortened the paper by XXX pages. We have expanded the explanations on the practicality in the Section "Limitations and practicality".
+
 *1.
@@
 While the onus should be on the authors for getting the page number down, some possible suggestions could be combining Sections 1 and 2 while moving some of the historical context (which I fully acknowledge is quite interesting) to the Appendix, and moving some of the technical details of Section 3.2 to the Appendix.*
+
+Thanks for your suggestions. This is how we have shortened the paper:
+
+- Details of Section 3.2 Statistical significance moved to the Appendix. 
+- Sections 1 and 2 have been made more precise, while still covering the requested additional topics.
+- The explanation of power calculation was left in the main paper because this is new and of primary importance for understanding the results.
+
+- Part of Experimental design -> appendix
 
@@
 
+The new Section "Limitations and practicality" addresses these concerns and we have added two sentences to the Introduction.
+
 *3.
@@
 
+See the new Section "Limitations and practicality".
+Some additional text is provided in various places.
+It's very practical to employ the lineup protocol using the `nullabor` package, and the (growing) number of citations from publications in a variety of areas show that it is being used.
+
+<!-- As mentioned above,
+
+This is not computationally expensive. The conventional test also make mistakes. You need something more than the conventional test. And you also need something more than the residual plot.
+
+check earlier paper.-->
+
 *4.
@@
 
+The lineup protocol is very suitable for big data.
+A plot other than a scatterplot is probably needed to better display a large number of observations.
+We have changed the language so that this is clearer.
+
 *5.
@@
 Granted this is likely done as a matter of convenience, but it would be good to articulate how cultural biases may be present in interpreting a lineup protocol.*
+
+The instructions for this study were in English, which motivated the constraint applied in the crowd-sourcing service.
+However many participants were likely multi-lingual, and they joined from a variety of countries.
+We don't anticipate that the results would be different with a different sample.
 
@@
 
+We have clarified the language of the Introduction and the Conclusion.
+
 *Minor Comments*
 
-*1. Page 9, Line 13: Since the three references are by the same authors, use \\cite{Ref1,Ref2,Ref3}.*
+*1. Page 9, Line 13: Since the three references are by the same authors, use \\cite{Ref1,Ref2,Ref3}.* DONE
 
-*2. Page 12, Line 7: $\alpha$.*
+*2. Page 12, Line 7: $\alpha$.* DONE
 
@@
 Why not write the power formula in terms of a generic significance level?
-Granted, one should be mindful of an abuse of notation if using $\alpha$ as it has been introduced in the Rorschach setup of the previous subsection.*
+Granted, one should be mindful of an abuse of notation if using $\alpha$ as it has been introduced in the Rorschach setup of the previous subsection.* We have kept the current explanation of power.
 
-*4. Page 13, Line 14: "...the scenario..."*
+*4. Page 13, Line 14: "...the scenario..."* DONE
 
-*5. Page 20, Line 36: "...variety of difficulty..."*
+*5. Page 20, Line 36: "...variety of difficulty..."* DONE
 
@@
 Can the authors convince me otherwise?*
+
+The small residual plots help to understand the change in effect size represented by the numerical values on the x-axis.
+They are indispensable, as reported by numerous readers during the proofing stage of the paper.

### 2023-09-06 appendix.Rmd
double space
finished revising the paper

@@
 # Additional details of testing procedures
+
+## Statistical significance
+
+Within the context of visual inference, with $K$ independent observers, the visual $p$-value can be seen as the probability of having as many or more participants detect the data plot than the observed result.
+
+The approach used in @majumder2013validation is as follows.
+Define $X_j = \{0,1\}$ to be a Bernoulli random variable measuring whether participant $j$ detected the data plot, and $X = \sum_{j=1}^{K}X_j$ be the total number of observers who detected the data plot.
+Then, by imposing a relatively strong assumption that all $K$ evaluations are fully independent, under $H_0$ $X \sim \mathrm{Binom}_{K,1/m}$.
+Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is estimated with $P(X \geq x) = 1 - F(x) + f(x)$, where $F(.)$ is the binomial cumulative distribution function, $f(.)$ is the binomial probability mass function and $x$ is the realization of number of observers choosing the data plot.
+
+As pointed out by @vanderplas2021statistical, this basic binomial model is deficient.
+It does not take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup, or account for when participants are offered the option to select one or more "most different" plots, or none, from a lineup.
+They suggest three common lineup scenarios: (1) $K$ different lineups are shown to $K$ participants, (2) $K$ lineups with different null plots but the same data plot are shown to $K$ participants, and (3) the same lineup is shown to $K$ participants.
+Scenario 3 is the most feasible to apply, but has the most dependencies to accommodate for the $p$-value calculation.
+For Scenario 3, VanderPlas et al propose modelling the probability of plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim \text{Dirichlet}(\alpha)$ for $i=1,...,m$ and $\alpha > 0$.
+The number of times plot $i$ being selected in $K$ evaluations is denoted as $c_i$.
+In case participant $j$ makes multiple selections, $1/s_j$ will be added to $c_i$ instead of one, where $s_j$ is the number of plots participant $j$ selected for $j=1,...K$.
+This ensures $\sum_{i}c_i=K$.
+Since we are only interested in the selections of the data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual $p$-value is given as
+
+```{=tex}
+\begin{equation} \label{eq:pvalue-beta-binomial}
+P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+
+\end{equation}
+```
+
+\noindent where $B(.)$ is the beta function defined as
+
+```{=tex}
+\begin{equation} \label{eq:betafunction}
+B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0.
+\end{equation}
+```
+
+\noindent We extend the equation to non-negative real number $c_i$ by applying a linear approximation
+
+```{=tex}
+\begin{equation} \label{eq:pvalue-beta-binomial-approx}
+P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+,
+\end{equation}
+```
+
+\noindent where $P(C \geq \lceil c_i \rceil)$ is calculated using Equation \ref{eq:pvalue-beta-binomial} and $P(C = \lfloor c_i \rfloor)$ is calculated by
+
+\begin{equation} \label{eq:pmf-beta-binomial}
+P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+.
+\end{equation}
+
+The parameter $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} and \ref{eq:pmf-beta-binomial} is usually unknown and will need to be estimated from the survey data.
+An interpretation of $\alpha$ is that when it is low only a few plots are attractive to the observers and tend to be selected, and when high, most plots are equally likely to be chosen.
+VanderPlas et al define $c$-interesting plot to be if $c$ or more participants select the plot as the most different.
+The expected number of plots selected at least $c$ times, $\mathrm{E}[Z_c]$, is then calculated as
+
+```{=tex}
+\begin{equation} \label{eq:c-interesting-expectation}
+\mathrm{E}[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}
+```
+
+With Equation \ref{eq:c-interesting-expectation}, $\alpha$ can be estimated using maximum likelihood estimation.
+Precise estimation of $\alpha$, is aided by evaluation of Rorschach lineups, where all plots are null plots.
+In a Rorschach, in theory all plots should be equally likely, but in practice some (irrelevant) visual elements may be more eye-catching than others.
+This is what $alpha$ captures, the capacity for extraneous features to distract the observer for a particular type of plot display.
+
 ## Effect size derivation

### 2023-09-06 appendix.tex
double space
finished revising the paper

@@
 \def\tightlist{}
+\usepackage{setspace}
+\doublespacing
 
@@
 procedures}\label{additional-details-of-testing-procedures}}
+
+\hypertarget{statistical-significance}{%
+\subsection{Statistical significance}\label{statistical-significance}}
+
+Within the context of visual inference, with \(K\) independent
+observers, the visual \(p\)-value can be seen as the probability of
+having as many or more participants detect the data plot than the
+observed result.
+
+The approach used in \citet{majumder2013validation} is as follows.
+Define \(X_j = \{0,1\}\) to be a Bernoulli random variable measuring
+whether participant \(j\) detected the data plot, and
+\(X = \sum_{j=1}^{K}X_j\) be the total number of observers who detected
+the data plot. Then, by imposing a relatively strong assumption that all
+\(K\) evaluations are fully independent, under \(H_0\)
+\(X \sim \mathrm{Binom}_{K,1/m}\). Therefore, the \(p\)-value of a
+lineup of size \(m\) evaluated by \(K\) observer is estimated with
+\(P(X \geq x) = 1 - F(x) + f(x)\), where \(F(.)\) is the binomial
+cumulative distribution function, \(f(.)\) is the binomial probability
+mass function and \(x\) is the realization of number of observers
+choosing the data plot.
+
+As pointed out by \citet{vanderplas2021statistical}, this basic binomial
+model is deficient. It does not take into account the possible
+dependencies in the visual test due to repeated evaluations of the same
+lineup, or account for when participants are offered the option to
+select one or more ``most different'' plots, or none, from a lineup.
+They suggest three common lineup scenarios: (1) \(K\) different lineups
+are shown to \(K\) participants, (2) \(K\) lineups with different null
+plots but the same data plot are shown to \(K\) participants, and (3)
+the same lineup is shown to \(K\) participants. Scenario 3 is the most
+feasible to apply, but has the most dependencies to accommodate for the
+\(p\)-value calculation. For Scenario 3, VanderPlas et al propose
+modelling the probability of plot \(i\) being selected from a lineup as
+\(\theta_i\), where \(\theta_i \sim \text{Dirichlet}(\alpha)\) for
+\(i=1,...,m\) and \(\alpha > 0\). The number of times plot \(i\) being
+selected in \(K\) evaluations is denoted as \(c_i\). In case participant
+\(j\) makes multiple selections, \(1/s_j\) will be added to \(c_i\)
+instead of one, where \(s_j\) is the number of plots participant \(j\)
+selected for \(j=1,...K\). This ensures \(\sum_{i}c_i=K\). Since we are
+only interested in the selections of the data plot \(i\), the marginal
+model can be simplified to a beta-binomial model and thus the visual
+\(p\)-value is given as
+
+\begin{equation} \label{eq:pvalue-beta-binomial}
+P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+
+\end{equation}
+
+\noindent where \(B(.)\) is the beta function defined as
+
+\begin{equation} \label{eq:betafunction}
+B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0.
+\end{equation}
+
+\noindent We extend the equation to non-negative real number \(c_i\) by
+applying a linear approximation
+
+\begin{equation} \label{eq:pvalue-beta-binomial-approx}
+P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+,
+\end{equation}
+
+\noindent where \(P(C \geq \lceil c_i \rceil)\) is calculated using
+Equation \ref{eq:pvalue-beta-binomial} and
+\(P(C = \lfloor c_i \rfloor)\) is calculated by
+
+\begin{equation} \label{eq:pmf-beta-binomial}
+P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+.
+\end{equation}
+
+The parameter \(\alpha\) used in Equation \ref{eq:pvalue-beta-binomial}
+and \ref{eq:pmf-beta-binomial} is usually unknown and will need to be
+estimated from the survey data. An interpretation of \(\alpha\) is that
+when it is low only a few plots are attractive to the observers and tend
+to be selected, and when high, most plots are equally likely to be
+chosen. VanderPlas et al define \(c\)-interesting plot to be if \(c\) or
+more participants select the plot as the most different. The expected
+number of plots selected at least \(c\) times, \(\mathrm{E}[Z_c]\), is
+then calculated as
+
+\begin{equation} \label{eq:c-interesting-expectation}
+\mathrm{E}[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}
+
+With Equation \ref{eq:c-interesting-expectation}, \(\alpha\) can be
+estimated using maximum likelihood estimation. Precise estimation of
+\(\alpha\), is aided by evaluation of Rorschach lineups, where all plots
+are null plots. In a Rorschach, in theory all plots should be equally
+likely, but in practice some (irrelevant) visual elements may be more
+eye-catching than others. This is what \(alpha\) captures, the capacity
+for extraneous features to distract the observer for a particular type
+of plot display.
 

### 2023-09-06 paper.Rmd
finished revising the paper

@@
   \usepackage{setspace}
-  % \doublespacing
+  \doublespacing
 output: rticles::tf_article
@@
 
-In addition, scatterplots of residuals plotted against the fitted values and each of the explanatory variables, can be used to scrutinize their relationships.
-If there are any visually discoverable patterns, the model is potentially inadequate or incorrectly specified.
-\textcolor{red}{Interesting features that are not directly connected to a linear model assumption can also be found from residual plots, such as the discreteness of the fitted values, outliers and influential points.} In general, one looks for noticeable departures from the model such as non-linear pattern or heteroskedasticity.
+It is also common to make scatterplots of residuals against the fitted values, and each of the explanatory variables, can be used to scrutinize their relationships.
+If there are any visually discoverable associations, the model is potentially inadequate or incorrectly specified.
+\textcolor{red}{It is also possible to discover patterns not directly connected to a linear model assumption from these residual plots, such as the discreteness or skewness of the fitted values, and outliers.} To read residual plots, one looks for noticeable departures from the model such as non-linear pattern or heteroskedasticity.
 A non-linear pattern would suggest that the model needs to have some additional non-linear terms.
 Heteroskedasticity suggests that the error is dependent on the predictors, and hence violates the independence assumption.
-
-However, correctly judging whether NO pattern exists in a residual plot is a difficult task for humans.
-Humans will almost always see a pattern (see @kahneman2011thinking), so the question that really needs answering is whether any pattern perceived is consistent with randomness, purely sampling variability, or noise.
-It is especially difficult to teach this to new analysts and students [@loy2021bringing].
-To address this, statistical tests are available to check for non-linear patterns [e.g.
+Statistical tests were developed to provide objective assessment, for example, of non-linear patterns [e.g.
 @ramsey1969tests], and heteroskedasticity  [e.g.
@@
 
-\textcolor{red}{There are other types of residual plot to extract the available information and identify inappropriate assumptions, like the added variable plot (cite here) for studying the role of an external variable if it enters linearly into a model.
+\textcolor{red}{There are also many variations on residual plots that are important to consider making.
+The normal probability plot is used to assess a normality assumption on the error distribution.
+Residuals are often plotted against  order of observations to check for collection bias, batch effects or autocorrelation.
+Partial residual plots assess the importance of an independent variable to a model, conditional on other variables, by plotting partial residuals against this variable.
+Partial residuals are computed from the difference between the response value and the fitted model excluding term of the independent variable of interest.
+A partial regression plot assesses the unique effect of an independent variable in a multiple regression model, by plotting residuals from the model with and without the this variable against each other.}
+
+<!--other types of residual plot to extract the available information and identify inappropriate assumptions, like the added variable plot (cite here) for studying the role of an external variable if it enters linearly into a model.
 Partial residual plots (cite here) have been suggested as computationally convenient substitutes for the added variable plots.
@@
 Alternatively.
-the residuals can be transformed to have a selected covariance structure: The usual suggestion is to obtain residuals with uncorrelated elements.}
-
-There is sustainable literature describing diagnostic procedures for linear models, e.g.
+the residuals can be transformed to have a selected covariance structure: The usual suggestion is to obtain residuals with uncorrelated elements.}-->
+
+These are all topics explained thoroughly in the sizeable literature on diagnostic procedures for linear models, e.g.
 @draper1998applied, @montgomery1982introduction, @belsley1980regression, @cook1999applied and @cook1982residuals.
@@
 \noindent The common wisdom of experts is that plotting the residuals is indispensable for diagnosing model fits.
-The lack of empirical evidence for the ubiquitous advice is *curious*.
-The research study discussed in this paper fills this gap.
+The lack of empirical evidence for the ubiquitous advice is *curious*, and is what this article tackles.
+
+\textcolor{red}{However, it needs to be noted that relying purely on subjective assessment of plots can be a source of anxiety.
+Indeed, correctly judging whether NO pattern exists in a residual plot is a really difficult task.
+People will almost always see a pattern (see @kahneman2011thinking), so the question that really needs answering is whether any pattern perceived is consistent with randomness, or sampling variability, or noise.
+@loy2021bringing emphasizes that this is especially difficult to teach to new analysts and students, and advocates to the broader use of the lineup protocol [@bujastatistical2009].}
+
+\textcolor{red}{The lineup protocol places a data plot in a field of null plots, allowing for a comparison of patterns due purely by chance to what is perceived in the data plot.
+For residual analysis this is especially helpful for gauging whether there is NO pattern.
+In its strict use, one would insist that the data plot is not seen before seeing the lineup, so that the observer doesn't know which is the true plot.
+When used this way, it provides an objective test for data plots.
+@majumdervalidation2013 validated that results from lineups assessed by human observers performed similarly to conventional tests.
+One wouldn't use a lineup when a conventional test exists and is adequate because it is more manually expensive to conduct.
+However, where no adequate conventional test exists, it is invaluable, as shown by @loy2013diagnostic.
+Here we use the lineup as a vehicle to rigorously explore why experts advise that residual plots are indispensable despite the prevalence of numerical tests.}
 
@@
 Conventional hypothesis tests usually have an existing method to derive or compute the $p$-value based on the null distribution.
-Here we establish the method to estimate a $p$-value for a visual test.
-
-Within the context of visual inference, with $K$ independent observers, the visual $p$-value can be seen as the probability of having as many or more participants detect the data plot than the observed result.
-
-The approach used in @majumder2013validation is as follows.
-Define $X_j = \{0,1\}$ to be a Bernoulli random variable measuring whether participant $j$ detected the data plot, and $X = \sum_{j=1}^{K}X_j$ be the total number of observers who detected the data plot.
-Then, by imposing a relatively strong assumption that all $K$ evaluations are fully independent, under $H_0$ $X \sim \mathrm{Binom}_{K,1/m}$.
-Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is estimated with $P(X \geq x) = 1 - F(x) + f(x)$, where $F(.)$ is the binomial cumulative distribution function, $f(.)$ is the binomial probability mass function and $x$ is the realization of number of observers choosing the data plot.
-
-As pointed out by @vanderplas2021statistical, this basic binomial model is deficient.
-It does not take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup, or account for when participants are offered the option to select one or more "most different" plots, or none, from a lineup.
-They suggest three common lineup scenarios: (1) $K$ different lineups are shown to $K$ participants, (2) $K$ lineups with different null plots but the same data plot are shown to $K$ participants, and (3) the same lineup is shown to $K$ participants.
-Scenario 3 is the most feasible to apply, but has the most dependencies to accommodate for the $p$-value calculation.
-For Scenario 3, VanderPlas et al propose modelling the probability of plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim \text{Dirichlet}(\alpha)$ for $i=1,...,m$ and $\alpha > 0$.
-The number of times plot $i$ being selected in $K$ evaluations is denoted as $c_i$.
-In case participant $j$ makes multiple selections, $1/s_j$ will be added to $c_i$ instead of one, where $s_j$ is the number of plots participant $j$ selected for $j=1,...K$.
-This ensures $\sum_{i}c_i=K$.
-Since we are only interested in the selections of the data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual $p$-value is given as
-
-```{=tex}
-\begin{equation} \label{eq:pvalue-beta-binomial}
-P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+
-\end{equation}
-```
-
-\noindent where $B(.)$ is the beta function defined as
-
-```{=tex}
-\begin{equation} \label{eq:betafunction}
-B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0.
-\end{equation}
-```
-
-\noindent We extend the equation to non-negative real number $c_i$ by applying a linear approximation
-
-```{=tex}
-\begin{equation} \label{eq:pvalue-beta-binomial-approx}
-P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+,
-\end{equation}
-```
-
-\noindent where $P(C \geq \lceil c_i \rceil)$ is calculated using Equation \ref{eq:pvalue-beta-binomial} and $P(C = \lfloor c_i \rfloor)$ is calculated by
-
-\begin{equation} \label{eq:pmf-beta-binomial}
-P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+.
-\end{equation}
-
-The parameter $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} and \ref{eq:pmf-beta-binomial} is usually unknown and will need to be estimated from the survey data.
-An interpretation of $\alpha$ is that when it is low only a few plots are attractive to the observers and tend to be selected, and when high, most plots are equally likely to be chosen.
-VanderPlas et al define $c$-interesting plot to be if $c$ or more participants select the plot as the most different.
-The expected number of plots selected at least $c$ times, $\mathrm{E}[Z_c]$, is then calculated as
-
-```{=tex}
-\begin{equation} \label{eq:c-interesting-expectation}
-\mathrm{E}[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}
-```
-
-With Equation \ref{eq:c-interesting-expectation}, $\alpha$ can be estimated using maximum likelihood estimation.
-Precise estimation of $\alpha$, is aided by evaluation of Rorschach lineups, where all plots are null plots.
-In a Rorschach, in theory all plots should be equally likely, but in practice some (irrelevant) visual elements may be more eye-catching than others.
-This is what $alpha$ captures, the capacity for extraneous features to distract the observer for a particular type of plot display.
+The method to estimate a $p$-value for a visual test essentially follows the process detailed by @vanderplas2021statistical.
+Details are given in Appendix A.
 
@@
 However, this mixed effects model does not work with scenario where participants are asked to select one or more most different plots.
-In this scenario, having the probability of a participant $j$ correctly picking the data plot from a lineup $l$ is insufficient to determine the power of a visual test because it does not provide information about the number of selections made by the participant for the calculation of the $p$-value (See Equation \ref{eq:pvalue-beta-binomial-approx}).
+In this scenario, having the probability of a participant $j$ correctly picking the data plot from a lineup $l$ is insufficient to determine the power of a visual test because it does not provide information about the number of selections made by the participant for the calculation of the $p$-value.
 Therefore, we directly estimate the probability of a lineup being rejected by assuming that individual skill has negligible effect on the variation of the power.
@@
 Two types of departures, non-linearity and heteroskedasticity, were collected during data collection periods I and II.
-The data collection period III was designed primarily to measure human responses to null lineups so that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated.
+The data collection period III was designed primarily to measure human responses to null lineups so that the visual $p$-values can be estimated.
 Additional lineups for both non-linearity and heteroskedasticity, using uniform fitted value distributions, were included for additional data, and to avoid participant frustration of too many difficult tasks.
@@
 We expect signal strength to decline in the simulated data plots with smaller $n$.
-A sample size of 300 is typically enough for structure to be visible in a scatterplot reliably.
-Beyond 300, structure can be obscured by over-plotting, and a different type of display that included transparency, binning or density mapping, might be needed.
+We chose 300 as the upper limit, because it is typically enough for structure to be visible in a scatterplot reliably.
+\textcolor{red}{Beyond 300, the scatterplot should probably be used with transparency or replaced with a density or binned plot.
+Scatterplots suffer from over-plotting.}
 
@@
 
-One of the notable limitations of the lineup protocol lies in its reliance on human judgements.
-In this context, the effectiveness of the lineup protocol is closely connected with the perceptive ability and subjective interpretation of individual evaluators.
-This reliance on human judgment introduces an element of variability and subjectivity that can potentially impact the robustness of the results obtained.
+One of the primary limitations of the lineup protocol lies in its reliance on human judgements.
+In this context, the effectiveness of a single lineup evaluation is closely connected with the perceptual ability and visual skills of the individual.
+When results from multiple individuals are combined the outcome is encouragingly robust.
+For simple plots and strong patterns just a few individuals are needed to arrive at a clear answer, but more individuals will be needed when the plot design is complex, or the signal strength is weak.
+<!--This reliance on human judgment introduces an element of variability and subjectivity that can potentially impact the robustness of the results obtained.
 To ensure reliable outcomes, it's often necessary to include a diverse group of participants capable of offering unbiased judgements.
-This requirement has led researchers, including those in our study, to turn to crowd-sourcing platforms that offer access to a random selection of participants, which can help mitigate potential bias.
-
-It is worth noting that despite the need for participants with unbiased perspectives, the lineup protocol can be implemented at a relatively low cost.
-Leveraging tools like the R package `nullabor`, the process of constructing and evaluating lineups becomes accessible and straightforward.
+This requirement has led researchers, including those in our study, to turn to crowd-sourcing platforms that offer access to a random selection of participants, which can help mitigate potential bias.-->
+
+Using a lineup protocol removes subjectiveness in interpreting patterns in plots.
+A plot is compared with draws from a null model, in much the same way as a test statistic is compared to its sampling distribution.
+It is important to define the data plot using the language provided by a grammar of graphics (e.g.
+using @ggplot2) because this provides the functional mapping of abstract variables to plot elements.
+It is also important to remove plot elements that might introduce bias, such as axis labels, text and legends, or to make them generic.
+
+The lineup protocol can be used cheaply and informally with the R package `nullabor`.
+There is evidence that it is being used fairly broadly, based on software download rates and citations of the original papers.
+For residual plot analysis we recommend that the lineup be the default first plot so that the data plot is only seen in the context of null plots.
+When a rigorous test is needed, we recommend using a crowd-sourcing service, as done in gene expression experiment described in @RNAseq2013.
+While it takes extra effort it is not difficult today, and costs are tiny compared to the overall costs of conducting a scientific experiment.
+We do also expect that at some point a computer vision model can be developed to take over the task of employing people to evaluate residual plots.
+
+<!--the process of constructing and evaluating lineups becomes accessible and straightforward.
 This accessibility empowers a wider range of researchers, even those without extensive statistical backgrounds, to employ the lineup protocol in their studies.
 Moreover, the cost-effectiveness of lineup-based analyses makes them particularly appealing for integration into statistical education within classroom settings (cite adam loy paper here).
-The practicality of lineup demonstrations can facilitate a deeper understanding of the concept of randomness and statistical testing among students.
-
-However, it's important to understand that if the goal is to attain results similar to those of formal hypothesis testing in rigor and reliability, using crowd-sourcing becomes necessary.
+The practicality of lineup demonstrations can facilitate a deeper understanding of the concept of randomness and statistical testing among students.-->
+
+<!--However, it's important to understand that if the goal is to attain results similar to those of formal hypothesis testing in rigor and reliability, using crowd-sourcing becomes necessary.
 The expense associated with crowd-sourcing is relatively reasonable when compared to the costs of other laboratory devices, making it a sensible investment for research that requires clear and valuable findings.
-Importantly, this higher level of scrutiny is typically justified only when the data plot can not be confidently identified and requires a thorough investigation.
-
-Additionally, there is the exciting possibility of utilizing computer vision technology to automate the analysis of lineup of residual plots.
+Importantly, this higher level of scrutiny is typically justified only when the data plot can not be confidently identified and requires a thorough investigation.-->
+
+For this study, simulated data was used to provide a precisely controlled environment within which to compare results from conventional testing to those from visual testing.
+We also explored only the most commonly used, the residual vs fitted value plots.
+However, we expect the behavior of the conventional test and the visual test to be similar when observed residuals are diagnosed with this type of plot or other residual plots.
+The conventional tests will be more sensitive to small departures from the null.
+They will also fail to detect departures when delivered residuals that have some contamination, like outliers or anomalies, as is often encountered when working with data.
+The lineup approach is well-suited for generally interpreting data plots, and also detecting unexpected patterns not related to the model.
+This is supported by earlier research (e.g.
+@wickham2010, @roy2015using, @loy2015you, @VanderPlas2015, @loy2016variations).
+
+<!-- Additionally, there is the exciting possibility of utilizing computer vision technology to automate the analysis of lineup of residual plots.
 This innovation has the potential to significantly reduce the expenses associated with recruiting human judges.
@@
 In real-world situations, researchers often rely on visual assessment to make informed decisions about model validity.
-The lineup protocol's capability to identify deviations from randomness serves as a valuable aid in this scenario.
+The lineup protocol's capability to identify deviations from randomness serves as a valuable aid in this scenario.-->
 
@@
 
-Regression analysis experts suggest that residual plots are indispensable methods for assessing model fit.
-We conducted a perceptual experiment using visual inference to assess this oft-repeated advice.
-The experiment tested two primary departures from good residuals: non-linearity and heteroskedasticity.
-
+This paper has described experimental evidence providing support for the advice of regression analysis experts *that residual plots are indispensable methods for assessing model fit*, using the formal framework of the lineup protocol.
+We conducted a perceptual experiment on scatterplots of residuals vs fitted values, with two primary departures from good residuals: non-linearity and heteroskedasticity.
 We found that conventional residual-based statistical tests are more sensitive to weak departures from model assumptions than visual tests.
@@
 The results from human evaluations provide a more practical solution, which reinforces the statements from regression experts that residual plots are an indispensable method for model diagnostics.
-
-It is important to note that residual plots need to be delivered as a lineup, embedded in a field of null plots.
+Further work would be needed to *quantify* how much departure from good residuals is too much.
+
+It is important to emphasize that this work also supports a change in common practice, which is to deliver residual plots as a lineup, embedded in a field of null plots, rather than be viewed out of context.
 A residual plot may contain many visual features, but some are caused by the characteristics of the predictors and the randomness of the error, not by the violation of the model assumptions.
 These irrelevant visual features have a chance to be filtered out by participants with a comparison to null plots, resulting in more accurate reading.
-The lineup enables a careful calibration for reading structure in residual plots.
+The lineup enables a careful calibration for reading structure in residual plots, and also provides the potential to discover interesting and important features in the data not directly connected to linear model assumptions.
 
@@
 This is possibly why residual plot analysis is often not done in practice.
-However, with the emergence of effective computer vision, our work could lay a foundation for automated residual plot assessment.
+However, with the emergence of effective computer vision, it is hoped this work helps to lay the foundation for automated residual plot assessment.
 <!--The findings suggest the strong demand of graphical inspection in regression diagnostics, so developing an automatic visual inference system to evaluate lineups of residual plots is valuable.

### 2023-09-06 paper.tex
finished revising the paper

--- 2023-05-04 paper.tex
@@
 
-In addition, scatterplots of residuals plotted against the fitted values
-and each of the explanatory variables, can be used to scrutinize their
-relationships. If there are any visually discoverable patterns, the
-model is potentially inadequate or incorrectly specified. In general,
-one looks for noticeable departures from the model such as non-linear
-pattern or heteroskedasticity. A non-linear pattern would suggest that
-the model needs to have some additional non-linear terms.
-Heteroskedasticity suggests that the error is dependent on the
-predictors, and hence violates the independence assumption. However,
-correctly judging whether NO pattern exists in a residual plot is a
-difficult task for humans. Humans will almost always see a pattern (see
-\citet{kahneman2011thinking}), so the question that really needs
-answering is whether any pattern perceived is consistent with
-randomness, purely sampling variability, or noise. It is especially
-difficult to teach this to new analysts and students
-\citep{loy2021bringing}. To address this, statistical tests are
-available to check for non-linear patterns
+It is also common to make scatterplots of residuals against the fitted
+values, and each of the explanatory variables, can be used to scrutinize
+their relationships. If there are any visually discoverable
+associations, the model is potentially inadequate or incorrectly
+specified.
+\textcolor{red}{It is also possible to discover patterns not directly connected to a linear model assumption from these residual plots, such as the discreteness or skewness of the fitted values, and outliers.}
+To read residual plots, one looks for noticeable departures from the
+model such as non-linear pattern or heteroskedasticity. A non-linear
+pattern would suggest that the model needs to have some additional
+non-linear terms. Heteroskedasticity suggests that the error is
+dependent on the predictors, and hence violates the independence
+assumption. Statistical tests were developed to provide objective
+assessment, for example, of non-linear patterns
 \citep[e.g.][]{ramsey1969tests}, and heteroskedasticity
@@
 
-Linear regression is a well-established procedure, and there is
-considerable literature describing diagnostic procedures, e.g.
-\citet{draper1998applied}, \citet{montgomery1982introduction},
-\citet{belsley1980regression}, \citet{cook1999applied} and
-\citet{cook1982residuals}. Interestingly, despite the abundance of
-conventional tests, ALL of these writings advise that plotting residuals
-is an essential tool for diagnosing regression model problems:
+\textcolor{red}{There are also many variations on residual plots that are important to consider making.
+The normal probability plot is used to assess a normality assumption on the error distribution.
+Residuals are often plotted against  order of observations to check for collection bias, batch effects or autocorrelation.
+Partial residual plots assess the importance of an independent variable to a model, conditional on other variables, by plotting partial residuals against this variable.
+Partial residuals are computed from the difference between the response value and the fitted model excluding term of the independent variable of interest.
+A partial regression plot assesses the unique effect of an independent variable in a multiple regression model, by plotting residuals from the model with and without the this variable against each other.}
+
+These are all topics explained thoroughly in the sizeable literature on
+diagnostic procedures for linear models, e.g. \citet{draper1998applied},
+\citet{montgomery1982introduction}, \citet{belsley1980regression},
+\citet{cook1999applied} and \citet{cook1982residuals}. Interestingly,
+despite the abundance of conventional tests, ALL of these writings
+advise that plotting residuals is an essential tool for diagnosing
+regression model problems:
 
@@
 \noindent The common wisdom of experts is that plotting the residuals is
-indispensable for diagnosing model fits. The ubiquity of this advice is
-\emph{curious}: investigating why, is the subject of this paper.
+indispensable for diagnosing model fits. The lack of empirical evidence
+for the ubiquitous advice is \emph{curious}, and is what this article
+tackles.
+
+\textcolor{red}{However, it needs to be noted that relying purely on subjective assessment of plots can be a source of anxiety.
+Indeed, correctly judging whether NO pattern exists in a residual plot is a really difficult task.
+People will almost always see a pattern (see @kahneman2011thinking), so the question that really needs answering is whether any pattern perceived is consistent with randomness, or sampling variability, or noise.
+@loy2021bringing emphasizes that this is especially difficult to teach to new analysts and students, and advocates to the broader use of the lineup protocol [@bujastatistical2009].}
+
+\textcolor{red}{The lineup protocol places a data plot in a field of null plots, allowing for a comparison of patterns due purely by chance to what is perceived in the data plot.
+For residual analysis this is especially helpful for gauging whether there is NO pattern.
+In its strict use, one would insist that the data plot is not seen before seeing the lineup, so that the observer doesn't know which is the true plot.
+When used this way, it provides an objective test for data plots.
+@majumdervalidation2013 validated that results from lineups assessed by human observers performed similarly to conventional tests.
+One wouldn't use a lineup when a conventional test exists and is adequate because it is more manually expensive to conduct.
+However, where no adequate conventional test exists, it is invaluable, as shown by @loy2013diagnostic.
+Here we use the lineup as a vehicle to rigorously explore why experts advise that residual plots are indispensable despite the prevalence of numerical tests.}
 
@@
 existing method to derive or compute the \(p\)-value based on the null
-distribution. Here we establish the method to estimate a \(p\)-value for
-a visual test.
-
-Within the context of visual inference, with \(K\) independent
-observers, the visual \(p\)-value can be seen as the probability of
-having as many or more participants detect the data plot than the
-observed result.
-
-The approach used in \citet{majumder2013validation} is as follows.
-Define \(X_j = \{0,1\}\) to be a Bernoulli random variable measuring
-whether participant \(j\) detected the data plot, and
-\(X = \sum_{j=1}^{K}X_j\) be the total number of observers who detected
-the data plot. Then, by imposing a relatively strong assumption that all
-\(K\) evaluations are fully independent, under \(H_0\)
-\(X \sim \mathrm{Binom}_{K,1/m}\). Therefore, the \(p\)-value of a
-lineup of size \(m\) evaluated by \(K\) observer is estimated with
-\(P(X \geq x) = 1 - F(x) + f(x)\), where \(F(.)\) is the binomial
-cumulative distribution function, \(f(.)\) is the binomial probability
-mass function and \(x\) is the realization of number of observers
-choosing the data plot.
-
-As pointed out by \citet{vanderplas2021statistical}, this basic binomial
-model is deficient. It does not take into account the possible
-dependencies in the visual test due to repeated evaluations of the same
-lineup, or account for when participants are offered the option to
-select one or more ``most different'' plots, or none, from a lineup.
-They suggest three common lineup scenarios: (1) \(K\) different lineups
-are shown to \(K\) participants, (2) \(K\) lineups with different null
-plots but the same data plot are shown to \(K\) participants, and (3)
-the same lineup is shown to \(K\) participants. Scenario 3 is the most
-feasible to apply, but has the most dependencies to accommodate for the
-\(p\)-value calculation. For Scenario 3, VanderPlas et al propose
-modelling the probability of plot \(i\) being selected from a lineup as
-\(\theta_i\), where \(\theta_i \sim \text{Dirichlet}(\alpha)\) for
-\(i=1,...,m\) and \(\alpha > 0\). The number of times plot \(i\) being
-selected in \(K\) evaluations is denoted as \(c_i\). In case participant
-\(j\) makes multiple selections, \(1/s_j\) will be added to \(c_i\)
-instead of one, where \(s_j\) is the number of plots participant \(j\)
-selected for \(j=1,...K\). This ensures \(\sum_{i}c_i=K\). Since we are
-only interested in the selections of the data plot \(i\), the marginal
-model can be simplified to a beta-binomial model and thus the visual
-\(p\)-value is given as
-
-\begin{equation} \label{eq:pvalue-beta-binomial}
-P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+
-\end{equation}
-
-\noindent where \(B(.)\) is the beta function defined as
-
-\begin{equation} \label{eq:betafunction}
-B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0.
-\end{equation}
-
-\noindent We extend the equation to non-negative real number \(c_i\) by
-applying a linear approximation
-
-\begin{equation} \label{eq:pvalue-beta-binomial-approx}
-P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+,
-\end{equation}
-
-\noindent where \(P(C \geq \lceil c_i \rceil)\) is calculated using
-Equation \ref{eq:pvalue-beta-binomial} and
-\(P(C = \lfloor c_i \rfloor)\) is calculated by
-
-\begin{equation} \label{eq:pmf-beta-binomial}
-P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+.
-\end{equation}
-
-The parameter \(\alpha\) used in Equation \ref{eq:pvalue-beta-binomial}
-and \ref{eq:pmf-beta-binomial} is usually unknown and will need to be
-estimated from the survey data. An interpretation of \(\alpha\) is that
-when it is low only a few plots are attractive to the observers and tend
-to be selected, and when high, most plots are equally likely to be
-chosen. VanderPlas et al define \(c\)-interesting plot to be if \(c\) or
-more participants select the plot as the most different. The expected
-number of plots selected at least \(c\) times, \(\mathrm{E}[Z_c]\), is
-then calculated as
-
-\begin{equation} \label{eq:c-interesting-expectation}
-\mathrm{E}[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}
-
-With Equation \ref{eq:c-interesting-expectation}, \(\alpha\) can be
-estimated using maximum likelihood estimation. Precise estimation of
-\(\alpha\), is aided by evaluation of Rorschach lineups, where all plots
-are null plots. In a Rorschach, in theory all plots should be equally
-likely, but in practice some (irrelevant) visual elements may be more
-eye-catching than others. This is what \(alpha\) captures, the capacity
-for extraneous features to distract the observer for a particular type
-of plot display.
+distribution. The method to estimate a \(p\)-value for a visual test
+essentially follows the process detailed by
+\citet{vanderplas2021statistical}. Details are given in Appendix A.
 
@@
 information about the number of selections made by the participant for
-the calculation of the \(p\)-value (See Equation
-\ref{eq:pvalue-beta-binomial-approx}). Therefore, we directly estimate
-the probability of a lineup being rejected by assuming that individual
-skill has negligible effect on the variation of the power. This
-assumption essentially averages out the subject ability and helps to
-simplify the model structure, thereby obviating a costly large-scale
-experiment to estimate complex covariance matrices. The same model given
-in Equation \ref{eq:logistic-regression-1-1} is applied to model the
-power of a visual test.
+the calculation of the \(p\)-value. Therefore, we directly estimate the
+probability of a lineup being rejected by assuming that individual skill
+has negligible effect on the variation of the power. This assumption
+essentially averages out the subject ability and helps to simplify the
+model structure, thereby obviating a costly large-scale experiment to
+estimate complex covariance matrices. The same model given in Equation
+\ref{eq:logistic-regression-1-1} is applied to model the power of a
+visual test.
 
@@
 period III was designed primarily to measure human responses to null
-lineups so that the parameter \(\alpha\) in Equation
-\ref{eq:pvalue-beta-binomial} can be estimated. Additional lineups for
-both non-linearity and heteroskedasticity, using uniform fitted value
-distributions, were included for additional data, and to avoid
-participant frustration of too many difficult tasks. Overall, we
+lineups so that the visual \(p\)-values can be estimated. Additional
+lineups for both non-linearity and heteroskedasticity, using uniform
+fitted value distributions, were included for additional data, and to
+avoid participant frustration of too many difficult tasks. Overall, we
 collected 7974 evaluations on 1152 unique lineups performed by 443
@@
 ``S'' shape for different sample sizes. We expect signal strength to
-decline in the simulated data plots with smaller \(n\). A sample size of
-300 is typically enough for structure to be visible in a scatterplot
-reliably. Beyond 300, structure can be obscured by over-plotting, and a
-different type of display that included transparency, binning or density
-mapping, might be needed.
+decline in the simulated data plots with smaller \(n\). We chose 300 as
+the upper limit, because it is typically enough for structure to be
+visible in a scatterplot reliably.
+\textcolor{red}{Beyond 300, the scatterplot should probably be used with transparency or replaced with a density or binned plot. Scatterplots suffer from over-plotting.}
 
@@
 
+\hypertarget{limitations-and-practicality}{%
+\section{Limitations and
+practicality}\label{limitations-and-practicality}}
+
+One of the primary limitations of the lineup protocol lies in its
+reliance on human judgements. In this context, the effectiveness of a
+single lineup evaluation is closely connected with the perceptual
+ability and visual skills of the individual. When results from multiple
+individuals are combined the outcome is encouragingly robust. For simple
+plots and strong patterns just a few individuals are needed to arrive at
+a clear answer, but more individuals will be needed when the plot design
+is complex, or the signal strength is weak.
+
+Using a lineup protocol removes subjectiveness in interpreting patterns
+in plots. A plot is compared with draws from a null model, in much the
+same way as a test statistic is compared to its sampling distribution.
+It is important to define the data plot using the language provided by a
+grammar of graphics (e.g.~using \citet{ggplot2}) because this provides
+the functional mapping of abstract variables to plot elements. It is
+also important to remove plot elements that might introduce bias, such
+as axis labels, text and legends, or to make them generic.
+
+The lineup protocol can be used cheaply and informally with the R
+package \texttt{nullabor}. There is evidence that it is being used
+fairly broadly, based on software download rates and citations of the
+original papers. For residual plot analysis we recommend that the lineup
+be the default first plot so that the data plot is only seen in the
+context of null plots. When a rigorous test is needed, we recommend
+using a crowd-sourcing service, as done in gene expression experiment
+described in \citet{RNAseq2013}. While it takes extra effort it is not
+difficult today, and costs are tiny compared to the overall costs of
+conducting a scientific experiment. We do also expect that at some point
+a computer vision model can be developed to take over the task of
+employing people to evaluate residual plots.
+
+For this study, simulated data was used to provide a precisely
+controlled environment within which to compare results from conventional
+testing to those from visual testing. We also explored only the most
+commonly used, the residual vs fitted value plots. However, we expect
+the behavior of the conventional test and the visual test to be similar
+when observed residuals are diagnosed with this type of plot or other
+residual plots. The conventional tests will be more sensitive to small
+departures from the null. They will also fail to detect departures when
+delivered residuals that have some contamination, like outliers or
+anomalies, as is often encountered when working with data. The lineup
+approach is well-suited for generally interpreting data plots, and also
+detecting unexpected patterns not related to the model. This is
+supported by earlier research (e.g. \citet{wickham2010},
+\citet{roy2015using}, \citet{loy2015you}, \citet{VanderPlas2015},
+\citet{loy2016variations}).
+
 \hypertarget{conclusions}{%
@@
 
-Regression analysis experts suggest that residual plots are
-indispensable methods for assessing model fit. We conducted a perceptual
-experiment using visual inference to assess this oft-repeated advice.
-The experiment tested two primary departures from good residuals:
-non-linearity and heteroskedasticity.
-
-We found that conventional residual-based statistical tests are more
+This paper has described experimental evidence providing support for the
+advice of regression analysis experts \emph{that residual plots are
+indispensable methods for assessing model fit}, using the formal
+framework of the lineup protocol. We conducted a perceptual experiment
+on scatterplots of residuals vs fitted values, with two primary
+departures from good residuals: non-linearity and heteroskedasticity. We
+found that conventional residual-based statistical tests are more
 sensitive to weak departures from model assumptions than visual tests.
@@
 reinforces the statements from regression experts that residual plots
-are an indispensable method for model diagnostics.
-
-It is important to note that residual plots need to be delivered as a
-lineup, embedded in a field of null plots. A residual plot may contain
-many visual features, but some are caused by the characteristics of the
-predictors and the randomness of the error, not by the violation of the
-model assumptions. These irrelevant visual features have a chance to be
-filtered out by participants with a comparison to null plots, resulting
-in more accurate reading. The lineup enables a careful calibration for
-reading structure in residual plots.
+are an indispensable method for model diagnostics. Further work would be
+needed to \emph{quantify} how much departure from good residuals is too
+much.
+
+It is important to emphasize that this work also supports a change in
+common practice, which is to deliver residual plots as a lineup,
+embedded in a field of null plots, rather than be viewed out of context.
+A residual plot may contain many visual features, but some are caused by
+the characteristics of the predictors and the randomness of the error,
+not by the violation of the model assumptions. These irrelevant visual
+features have a chance to be filtered out by participants with a
+comparison to null plots, resulting in more accurate reading. The lineup
+enables a careful calibration for reading structure in residual plots,
+and also provides the potential to discover interesting and important
+features in the data not directly connected to linear model assumptions.
 
@@
 in practice. However, with the emergence of effective computer vision,
-our work could lay a foundation for automated residual plot assessment.
+it is hoped this work helps to lay the foundation for automated residual
+plot assessment.
 

### 2023-09-18 JCGS_response_to_reviews/response_to_review.Rmd
updated response to reviewers

@@
 output: pdf_document
-date: "2023-08-07"
+date: "`r Sys.Date()`"
+bibliography: ../ref.bib
 ---
@@
 
-Thank you for the careful reviews. What follows is our point-by-point response to reviewer comments.
+We thank the anonymous reviewers for their careful reviews. What follows is our point-by-point response to reviewer comments.
 
-Note that *The original reviewer comments are in italic* and our response is normal text.
+Note that *the original reviewer comments are in italic* and our response is in normal text.
 
 # Reviewer 1
-
-*Comments to the Author*
 
@@
 
-Thank for this feedback.
-We are happy to see that, like many people, you don't need convincing and that you are glad to see that we have succeeded in demonstrating the superiority of residual diagnostic plots.
+Thank you for this feedback.
+We are glad to see that the reviewer agrees with the superiority of the residual diagnostic plots and is impressed with our formalised approach to demonstrate this superiority.
 
@@
 
-The work very specifically uses simulated data because we need to study a problem in controlled conditions.
-You are right, though, that data will pose more interesting challenges, and this can be in two directions.
-One, being that the lineup protocol will allow interesting and unexpected patterns in the residuals to be detected (see Buja et al and Wickham et al for examples).
-The other is that the current process for generating null data uses residual rotation and will likely maintain irrelevant quirks of the data in the null plots.
-These should not inadvertently allow the reader to detect the data plot based on quirks.
-
-We have addressed this criticism in the broader discussion of the use of residual plots and lineups for residual diagnostics added to the new Section "Limitations and practicality".
+The work specifically uses simulated data because we need to study the problem in controlled conditions.
+The reviewer is correct, though, that real data will pose more interesting challenges.
+It should be noted that our simulation setup do include discrete independent variables, and consequently somewhat discretized response.
+The other point to note is that our null generating process uses residual rotation and will likely maintain irrelevant quirks of the data in the null plots as well.
+In that sense, the real data plot is not necessary always very obvious (see @buja2009statistical and @wickham2010 for examples).
+Nonetheless, our simulation does only test handful of conditions.
+We have addressed this limited scope in the broader discussion of the use of residual plots and lineups for residual diagnostics added to the new Section "Limitations and practicality".
 <!-- (see issue #11).-->
 
+<!--
+ET: I don't think talking about computer vision model makes sense here as the reviewer would not know about our other work?
+
 An observational study involving actual residual plots will hopefully be possible once a computer vision model is deployed that will allow analysts to upload their residuals for automated testing.
+-->
 
@@
 Having a definitive (numerical) answer to "how much is too much" would require more and different work.
-For our purposes, it was that the true residual plot could be detected from the null plots when the lineup was evaluated by K people.
+For our purposes, it was that the true residual plot could be detected from the null plots when the lineup was evaluated by people.
 We've added a sentence to the Conclusion mentioning this.
@@
 
-It can be argued that it is being used by practitioners based on download rates of the nullabor package, and applications articles that report usage.
+It can be argued that it is being used by practitioners based on download rates of the `nullabor` package (>19,000 downloads in the past year) which implements a process for creating lineups, and application articles that report usage.
 Using it formally, we agree is cumbersome and would require engaging a crowd-sourcing service, that involves a small expense.
 Science labs typically have a budget for equipment and experimental devices and we would advocate this is a useful addition.
-We have definitely used it ourselves in 10.1007/s00180-014-0534-x for ecological data and  10.4172/2153-0602.1000139 for gene expression analysis.
+Where a practitioner lacks the budget for running crowd-sourcing services, the protocol can be used with their team or network and can be an excellent way to engage their team/network with the data.
+We have definitely used it ourselves in @roychowdhuryUsingVisualStatistical2015a for ecological data and @mahbubulmajumderVisualMiningMethods2013 for gene expression analysis.
 We have added some details in the "Limitations and practicality" section.
@@
 
-Thanks for the positive feedback. We have shortened the paper by XXX pages. We have expanded the explanations on the practicality in the Section "Limitations and practicality".
+Thank you for the positive feedback. We have shortened the paper by XXX pages. We have expanded the explanations on the practicality in the Section "Limitations and practicality".
 
@@
 
-Thanks for your suggestions. This is how we have shortened the paper:
+Thank you for your suggestions. This is how we have shortened the paper:
 
@@
 - The explanation of power calculation was left in the main paper because this is new and of primary importance for understanding the results.
-
-- Part of Experimental design -> appendix
+- Part of the Experimental design has been moved into the Appendix.
 
@@
 
-The new Section "Limitations and practicality" addresses these concerns and we have added two sentences to the Introduction.
+The new Section "Limitations and practicality" addresses these concerns and we have added a couple of sentences to the Introduction.
 
@@
 However many participants were likely multi-lingual, and they joined from a variety of countries.
-We don't anticipate that the results would be different with a different sample.
+We don't anticipate that the results would significantly differ with a different sample.
 
@@
 
-*1. Page 9, Line 13: Since the three references are by the same authors, use \\cite{Ref1,Ref2,Ref3}.* DONE
+*1. Page 9, Line 13: Since the three references are by the same authors, use \\cite{Ref1,Ref2,Ref3}.* Done.
 
-*2. Page 12, Line 7: $\alpha$.* DONE
+*2. Page 12, Line 7: $\alpha$.* Done.
 
@@
 
-*4. Page 13, Line 14: "...the scenario..."* DONE
+*4. Page 13, Line 14: "...the scenario..."* Done.
 
-*5. Page 20, Line 36: "...variety of difficulty..."* DONE
+*5. Page 20, Line 36: "...variety of difficulty..."* Done.
 

### 2023-09-18 README.Rmd
update citation to arxiv version
update README with reproducibility section

@@
 
-Li et al. (2023) `r metadata$title`. URL:  https://github.com/TengMCing/lineup_residual_diagnostics
+Li et al. (2023) `r metadata$title`. arXiv preprint. URL: http://arxiv.org/abs/2308.05964
+
+## Reproducibility
+
+The main paper is written using R Markdown and contained in "paper.Rmd". To render this, you need to install all relevant packages. Two custom packages are:
+
+```{r, eval = FALSE}
+# OOP supports needed by `visage`
+remotes::install_github("TengMCing/bandicoot")
+# 
+# Visual inference models and p-value calculation
+remotes::install_url("https://github.com/TengMCing/visage/raw/master/built/visage_0.1.0.tar.gz")
+```

### 2023-09-18 paper.Rmd
fix citation so that it is working within \textcolor{red} which renders markdown syntax void

@@
 
-Diagnosing a model is an important part of analysing data.
-In linear regression analysis, studying the residuals from a model fit is a common diagnostic activity.
+\textcolor{red}{Diagnosing a model is an important part of building an appropriate model.} In linear regression analysis, studying the residuals from a model fit is a common diagnostic activity.
 Residuals summarise what is not captured by the model, and thus provide the capacity to identify what might be wrong.
@@
 
-\textcolor{red}{There are also many variations on residual plots that are important to consider making.
+<!--- ET: I suggest removing this to save space and we don't use this plot type.
+
+\textcolor{red}{There are also other types of plots that can be used for assessing residuals.
 The normal probability plot is used to assess a normality assumption on the error distribution.
@@
 A partial regression plot assesses the unique effect of an independent variable in a multiple regression model, by plotting residuals from the model with and without the this variable against each other.}
+
+--->
 
@@
 
+<!--ET: I think we need to remove this to reduce pages.
+
 These are all topics explained thoroughly in the sizeable literature on diagnostic procedures for linear models, e.g.
 @draper1998applied, @montgomery1982introduction, @belsley1980regression, @cook1999applied and @cook1982residuals.
-Interestingly, despite the abundance of conventional tests, ALL of these writings advise that plotting residuals is an essential tool for diagnosing regression model problems:
+Interestingly, despite the abundance of conventional tests, _all_ of these writings advise that plotting residuals is an essential tool for diagnosing regression model problems:
 
@@
 
-\noindent The common wisdom of experts is that plotting the residuals is indispensable for diagnosing model fits.
+-->
+
+The common wisdom of experts is that plotting the residuals is indispensable for diagnosing model fits [@draper1998applied; @cook1982residuals; @montgomery1982introduction].
 The lack of empirical evidence for the ubiquitous advice is *curious*, and is what this article tackles.
@@
 \textcolor{red}{However, it needs to be noted that relying purely on subjective assessment of plots can be a source of anxiety.
-Indeed, correctly judging whether NO pattern exists in a residual plot is a really difficult task.
-People will almost always see a pattern (see @kahneman2011thinking), so the question that really needs answering is whether any pattern perceived is consistent with randomness, or sampling variability, or noise.
-@loy2021bringing emphasizes that this is especially difficult to teach to new analysts and students, and advocates to the broader use of the lineup protocol [@bujastatistical2009].}
+Indeed, correctly judging whether \emph{no} pattern exists in a residual plot is a really difficult task.
+People will almost always see a pattern \citep[see ][]{kahneman2011thinking}, so the question that really needs answering is whether any pattern perceived is consistent with randomness, or sampling variability, or noise.
+\cite{loy2021bringing} emphasizes that this is especially difficult to teach to new analysts and students, and advocates to the broader use of the lineup protocol \citep{bujastatistical2009}.}
 
 \textcolor{red}{The lineup protocol places a data plot in a field of null plots, allowing for a comparison of patterns due purely by chance to what is perceived in the data plot.
-For residual analysis this is especially helpful for gauging whether there is NO pattern.
+For residual analysis this is especially helpful for gauging whether there is \emph{no} pattern.
 In its strict use, one would insist that the data plot is not seen before seeing the lineup, so that the observer doesn't know which is the true plot.
 When used this way, it provides an objective test for data plots.
-@majumdervalidation2013 validated that results from lineups assessed by human observers performed similarly to conventional tests.
+\cite{majumdervalidation2013} validated that results from lineups assessed by human observers performed similarly to conventional tests.
 One wouldn't use a lineup when a conventional test exists and is adequate because it is more manually expensive to conduct.
-However, where no adequate conventional test exists, it is invaluable, as shown by @loy2013diagnostic.
+However, where no adequate conventional test exists, it is invaluable, as shown by \cite{loy2013diagnostic}.
 Here we use the lineup as a vehicle to rigorously explore why experts advise that residual plots are indispensable despite the prevalence of numerical tests.}
@@
 
-# Limitations and practicality
+# \textcolor{red}{Limitations and practicality}
 

### 2023-09-18 paper.tex
fix citation so that it is working within \textcolor{red} which renders markdown syntax void

@@
 
-Diagnosing a model is an important part of analysing data. In linear
-regression analysis, studying the residuals from a model fit is a common
-diagnostic activity. Residuals summarise what is not captured by the
-model, and thus provide the capacity to identify what might be wrong.
+\textcolor{red}{Diagnosing a model is an important part of building an appropriate model.}
+In linear regression analysis, studying the residuals from a model fit
+is a common diagnostic activity. Residuals summarise what is not
+captured by the model, and thus provide the capacity to identify what
+might be wrong.
 
@@
 
-\textcolor{red}{There are also many variations on residual plots that are important to consider making.
-The normal probability plot is used to assess a normality assumption on the error distribution.
-Residuals are often plotted against  order of observations to check for collection bias, batch effects or autocorrelation.
-Partial residual plots assess the importance of an independent variable to a model, conditional on other variables, by plotting partial residuals against this variable.
-Partial residuals are computed from the difference between the response value and the fitted model excluding term of the independent variable of interest.
-A partial regression plot assesses the unique effect of an independent variable in a multiple regression model, by plotting residuals from the model with and without the this variable against each other.}
-
-These are all topics explained thoroughly in the sizeable literature on
-diagnostic procedures for linear models, e.g. \citet{draper1998applied},
-\citet{montgomery1982introduction}, \citet{belsley1980regression},
-\citet{cook1999applied} and \citet{cook1982residuals}. Interestingly,
-despite the abundance of conventional tests, ALL of these writings
-advise that plotting residuals is an essential tool for diagnosing
-regression model problems:
-
-\begin{quote}
-\emph{Some most useful checks allowed by data plots should be done on a
-routine basis for every regression.} \citep{draper1998applied}
-\end{quote}
-
-\begin{quote}
-\emph{Such formal and informal procedures are complementary, and both
-have a place in residual analysis.} \citep{cook1982residuals}
-\end{quote}
-
-\begin{quote}
-\emph{In our experience, statistical tests on regression model residuals
-are not widely used. In most practical situations the residual plots are
-more informative than the corresponding tests. However, since residual
-plots do require skill and experience to interpret, the statistical
-tests may occasionally prove useful.} \citep{montgomery1982introduction}
-\end{quote}
-
-\noindent The common wisdom of experts is that plotting the residuals is
-indispensable for diagnosing model fits. The lack of empirical evidence
-for the ubiquitous advice is \emph{curious}, and is what this article
-tackles.
+The common wisdom of experts is that plotting the residuals is
+indispensable for diagnosing model fits
+\citep{draper1998applied, cook1982residuals, montgomery1982introduction}.
+The lack of empirical evidence for the ubiquitous advice is
+\emph{curious}, and is what this article tackles.
 
 \textcolor{red}{However, it needs to be noted that relying purely on subjective assessment of plots can be a source of anxiety.
-Indeed, correctly judging whether NO pattern exists in a residual plot is a really difficult task.
-People will almost always see a pattern (see @kahneman2011thinking), so the question that really needs answering is whether any pattern perceived is consistent with randomness, or sampling variability, or noise.
-@loy2021bringing emphasizes that this is especially difficult to teach to new analysts and students, and advocates to the broader use of the lineup protocol [@bujastatistical2009].}
+Indeed, correctly judging whether \emph{no} pattern exists in a residual plot is a really difficult task.
+People will almost always see a pattern \citep[see ][]{kahneman2011thinking}, so the question that really needs answering is whether any pattern perceived is consistent with randomness, or sampling variability, or noise.
+\cite{loy2021bringing} emphasizes that this is especially difficult to teach to new analysts and students, and advocates to the broader use of the lineup protocol \citep{bujastatistical2009}.}
 
 \textcolor{red}{The lineup protocol places a data plot in a field of null plots, allowing for a comparison of patterns due purely by chance to what is perceived in the data plot.
-For residual analysis this is especially helpful for gauging whether there is NO pattern.
+For residual analysis this is especially helpful for gauging whether there is \emph{no} pattern.
 In its strict use, one would insist that the data plot is not seen before seeing the lineup, so that the observer doesn't know which is the true plot.
 When used this way, it provides an objective test for data plots.
-@majumdervalidation2013 validated that results from lineups assessed by human observers performed similarly to conventional tests.
+\cite{majumdervalidation2013} validated that results from lineups assessed by human observers performed similarly to conventional tests.
 One wouldn't use a lineup when a conventional test exists and is adequate because it is more manually expensive to conduct.
-However, where no adequate conventional test exists, it is invaluable, as shown by @loy2013diagnostic.
+However, where no adequate conventional test exists, it is invaluable, as shown by \cite{loy2013diagnostic}.
 Here we use the lineup as a vehicle to rigorously explore why experts advise that residual plots are indispensable despite the prevalence of numerical tests.}
@@
 
-\hypertarget{limitations-and-practicality}{%
-\section{Limitations and
-practicality}\label{limitations-and-practicality}}
+\hypertarget{section}{%
+\section{\texorpdfstring{\textcolor{red}{Limitations and practicality}}{}}\label{section}}
 

### 2023-09-18 paper.Rmd
intro done.

@@
 
-It is also common to make scatterplots of residuals against the fitted values, and each of the explanatory variables, can be used to scrutinize their relationships.
+\textcolor{red}{Scatterplots of residuals against the fitted values, and each of the explanatory variables, are commonly used to scrutinize their relationships}.
 If there are any visually discoverable associations, the model is potentially inadequate or incorrectly specified.
-\textcolor{red}{It is also possible to discover patterns not directly connected to a linear model assumption from these residual plots, such as the discreteness or skewness of the fitted values, and outliers.} To read residual plots, one looks for noticeable departures from the model such as non-linear pattern or heteroskedasticity.
+\textcolor{red}{We can also potentially discover patterns not directly connected to a linear model assumption from these residual plots, such as the discreteness or skewness of the fitted values, and outliers.} To read residual plots, one looks for noticeable departures from the model such as non-linear pattern or heteroskedasticity.
 A non-linear pattern would suggest that the model needs to have some additional non-linear terms.
@@
 
-\textcolor{red}{However, it needs to be noted that relying purely on subjective assessment of plots can be a source of anxiety.
-Indeed, correctly judging whether \emph{no} pattern exists in a residual plot is a really difficult task.
+\textcolor{red}{However, relying purely on subjective assessment of a single plot can be deceiving.
+Indeed, correctly judging whether \emph{no} pattern exists in a residual plot is a difficult task.
 People will almost always see a pattern \citep[see ][]{kahneman2011thinking}, so the question that really needs answering is whether any pattern perceived is consistent with randomness, or sampling variability, or noise.

### 2023-09-18 paper.tex
intro done.

@@
 
-It is also common to make scatterplots of residuals against the fitted
-values, and each of the explanatory variables, can be used to scrutinize
-their relationships. If there are any visually discoverable
-associations, the model is potentially inadequate or incorrectly
-specified.
-\textcolor{red}{It is also possible to discover patterns not directly connected to a linear model assumption from these residual plots, such as the discreteness or skewness of the fitted values, and outliers.}
+\textcolor{red}{Scatterplots of residuals against the fitted values, and each of the explanatory variables, are commonly used to scrutinize their relationships}.
+If there are any visually discoverable associations, the model is
+potentially inadequate or incorrectly specified.
+\textcolor{red}{We can also potentially discover patterns not directly connected to a linear model assumption from these residual plots, such as the discreteness or skewness of the fitted values, and outliers.}
 To read residual plots, one looks for noticeable departures from the

### 2023-09-20 paper.Rmd
background done

@@
 There can be various types of departures from this ideal pattern.
-Non-linearity, heteroskedasticity and non-normality, shown in Figure \ref{fig:residual-plot-common-departures}B,C,D are three commonly checked departures.
+Non-linearity, heteroskedasticity and non-normality, shown in Figures \ref{fig:residual-plot-common-departures}B, \ref{fig:residual-plot-common-departures}C, and \ref{fig:residual-plot-common-departures}D are three commonly checked departures.
 
@@
 Figure \ref{fig:residual-plot-common-departures}D shows a scatterplot where the residuals have a skewed distribution, as seen by the uneven vertical spread.
-Unlike non-linearity and heteroskedasticity, non-normality is usually detected with a different type of residual plot: a histogram or normal probability plot.
+Unlike non-linearity and heteroskedasticity, non-normality is usually detected with a different type of residual plot: a histogram or a normal probability plot.
 Because we focus on scatterplots, non-normality is not one of the departures examined in depth in this paper.
@@
 This is encouraging, as it supports the use of visual inference in situations where there are no existing numerical testing procedures.
-Visual inference has also been integrated into diagnostics for hierarchical linear models where the lineup protocol is used to judge the assumptions of linearity, normality and constant error variance for both the level-1 and level-2 residuals [@loy2013diagnostic, @loy2014hlmdiag and @loy2015you].
+Visual inference has also been integrated into diagnostics for hierarchical linear models where the lineup protocol is used to judge the assumptions of linearity, normality and constant error variance for both the level-1 and level-2 residuals \citep{loy2013diagnostic, loy2014hlmdiag, loy2015you}.
 

### 2023-09-20 paper.Rmd
clean up section 3

@@
 \begin{equation} \label{eq:logistic-regression-1-1}
-Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),
+Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(\log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),
 \end{equation}
@@
 
-\noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
-The effect size $E$ is the only predictor and the intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$, the desired significance level.
+\noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = \exp(z)(1+\exp(z))^{-1}$.
+The effect size $E$ is the only predictor and the intercept is fixed to $\log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$, the desired significance level.
 
@@
                booktabs = TRUE, 
-               caption  = 'Levels of the factors used in data collection periods I, II, III.', 
+               caption  = 'Levels of the factors used in data collection periods I, II, and III.', 
                label = "model-factor-table" ) %>%

### 2023-09-20 paper.Rmd
remove two lineup plots
remove frivolous mentions of packages and replace tidyverse package with tidyverse paper

@@
 
+<!----
+
 Figure \ref{fig:example-poly-lineup} is one of the non-linearity lineups used in experiment.
@@
 ```
+
+--->
 
@@
 
+<!---
+
 ```{r example-heter-lineup, fig.pos="t!", fig.height = 7, fig.width = 7, fig.cap = "One of the lineups containing heteroskedasticity pattern used in data collection period II.
@@
 Nine out of 11 participants chose this plot from the lineup.
+
+--->
 
@@
 
-These `R` packages are used for the work: 
-`cli` [@cli], `curl` [@curl], `dplyr` [@dplyr], `ggplot2` [@ggplot2], `jsonlite` [@jsonlite], `lmtest` [@lmtest], `mpoly` [@mpoly], `progress` [@progress], `tibble` [@tibble], `ggmosaic` [@ggmosaic], `purrr` [@purrr], `tidyr` [@tidyr], `readr` [@readr], `stringr` [@stringr], `here` [@here], `kableExtra` [@kableextra], `patchwork` [@patchwork], `rcartocolor` [@rcartocolor].
+These `R` packages are used for the work: `tidyverse` [@tidyverse], `lmtest` [@lmtest], `mpoly` [@mpoly], `ggmosaic` [@ggmosaic], `kableExtra` [@kableextra], `patchwork` [@patchwork], `rcartocolor` [@rcartocolor].
 The study website is powered by `PythonAnywhere` [@pythonanywhere] and the `Flask` web framework [@flask].

### 2023-09-20 paper.Rmd
tighten up the maths notation

@@
 \begin{align*} \label{eq:nonlinearity-model}
-\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
+\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
 \boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
@@
 
-\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
+\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
 
@@
 \begin{equation} \label{eq:scaling-function}
-g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x}))2k - k, \quad \text{for} \quad k > 0. 
+g(\boldsymbol{x}, k) = 2k\cdot\frac{\boldsymbol{x} - x_{min}\boldsymbol{1}_n}{x_{max} - x_{min}} - k, \quad \text{for} \quad k > 0. 
 \end{equation}
 ```
+
+\noindent where $\displaystyle x_{min} = \min_{i\in \{1, ..., n\}} x_i$, $\displaystyle x_{max} = \max_{i\in \{1, ..., n\}} x_i$ and $x_i$ is the $i$-th entry of $\boldsymbol{x}$.
 
@@
 \begin{equation} \label{eq:null-model}
-\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
+\boldsymbol{y} = \beta_0\boldsymbol{1}_n + \beta_1 \boldsymbol{x} + \boldsymbol{u},
 \end{equation}
@@
 \begin{align*} \label{eq:heter-model}
-\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
+\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
 \boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
-\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}), 
+\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}_n, \boldsymbol{1}_n + (2 - |a|)(\boldsymbol{x} - a\boldsymbol{1}_n)^\top(\boldsymbol{x} - a\boldsymbol{1}_n) b \boldsymbol{I}_n), 
 \end{align*}

### 2023-09-20 paper.Rmd
tighten language + maths + combine power figures to one

@@
 
-\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
+\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials [@hermite1864nouveau; originally by @de1820theorie], $\varepsilon \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
 
@@
 \noindent where $\displaystyle x_{min} = \min_{i\in \{1, ..., n\}} x_i$, $\displaystyle x_{max} = \max_{i\in \{1, ..., n\}} x_i$ and $x_i$ is the $i$-th entry of $\boldsymbol{x}$.
+The function `hermite` from the R package `mpoly` [@mpoly] is used to  simulate $\boldsymbol{z}_{raw}$ to generate Hermite polynomials.
+
+<!--- ET: interesting history, but not pertinent to the story
 
 According to @abramowitz1964handbook, Hermite polynomials were initially defined by @de1820theorie, but named after @hermite1864nouveau because of the unrecognisable form of Laplace's work. 
-The function `hermite` from the R package `mpoly` [@mpoly] is used to  simulate $\boldsymbol{z}_{raw}$ to generate Hermite polynomials.
+--->
 
@@
 \noindent where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
-Since $z = O(x^j)$, for $j > 1$, $z$ is a higher order term left out of the null regression, which will lead to model misspecification.
+Here $\boldsymbol{z}$ is a higher order term of $\boldsymbol{x}$ and leaving it out in the null regression results in model misspecification.
+
+<!--- ET: let's keep the language simpler.
+
+Since $z = O(x^j)$, for $j > 1$, $z$ is a higher order term left out of the null regression, it leads to model misspecification. 
+--->
 
@@
 A greater value of $j$ will result in a curve with more turning points.
-It is expected that the "U" shape will be the easiest to detect, as the shape gets more complex it will be harder to perceive in a scatterplot, particularly when there is noise.
+It is expected that the "U" shape will be the easiest to detect, and as the shape gets more complex it will be harder to perceive in a scatterplot, particularly when there is noise.
 Figure \ref{fig:different-sigma} shows the "U" shape for different amounts of noise ($\sigma$).
@@
 \boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
-\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}_n, \boldsymbol{1}_n + (2 - |a|)(\boldsymbol{x} - a\boldsymbol{1}_n)^\top(\boldsymbol{x} - a\boldsymbol{1}_n) b \boldsymbol{I}_n), 
+\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}_n, \boldsymbol{1}_n + (2 - |a|)(\boldsymbol{x} - a\boldsymbol{1}_n)'(\boldsymbol{x} - a\boldsymbol{1}_n) b \boldsymbol{I}_n), 
 \end{align*}
@@
 Sometimes the fitted values are discrete because one or more predictors were measured discretely.
-The distribution may be relatively Gaussian, reflecting a linear combination of many predictors, adhering to the Central Limit Theorem.
 It is also common to see a skewed distribution of fitted values if one or more of the predictors has a skewed distribution.
@@
 
+<!--- ET: previous reference was "normal" but now Gaussian. This doesn't add to the story so I suggest just remove.
+
+The distribution may be relatively Gaussian, reflecting a linear combination of many predictors, adhering to the Central Limit Theorem.
+
+--->
+
 \noindent Figure \ref{fig:different-dist} shows the non-linear pattern, a "U" shape, with the different fitted value distributions.
@@
 We chose 300 as the upper limit, because it is typically enough for structure to be visible in a scatterplot reliably.
-\textcolor{red}{Beyond 300, the scatterplot should probably be used with transparency or replaced with a density or binned plot.
-Scatterplots suffer from over-plotting.}
+\textcolor{red}{Beyond 300, the scatterplot should probably be used with transparency or replaced with a density or binned plot as scatterplots suffer from over-plotting.}
 
@@
 \noindent where $\text{diag}(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,
-$\boldsymbol{R} = \boldsymbol{I}_n - \boldsymbol{H}$ is the residual operator, $\boldsymbol{H} = \boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the hat matrix, $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals where $\boldsymbol{Z}$ contain any higher order terms of $\boldsymbol{X}$ left out of the regression equation, $\boldsymbol{\beta}_z$ contains the corresponding coefficients, and $\sigma^2\boldsymbol{I}$ is the assumed covariance matrix of the error term when $H_0$ is true.
+$\boldsymbol{R} = \boldsymbol{I}_n - \boldsymbol{H}$ is the residual operator, $\boldsymbol{H} = \boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the hat matrix, $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals where $\boldsymbol{Z}$ contain any higher order terms of $\boldsymbol{X}$ left out of the regression equation, $\boldsymbol{\beta}_z$ contains the corresponding coefficients, and $\sigma^2\boldsymbol{I}_n$ is the assumed covariance matrix of the error term when $H_0$ is true.
 
@@
 
-To compute the effect size for each lineup we simulate a sufficient large number of samples from the same model, in each sample, the number of observations $n$ is fixed.
+To compute the effect size for each lineup we simulate a sufficient large number of samples from the same model with the number of observations $n$ fixed for each sample.
 We then compute the effect size for each sample and take the average as the final value.
@@
 
-Data collection period III was primarily to obtain Rorschach lineup evaluations to estimate $\alpha$ from Equation \ref{eq:pvalue-beta-binomial}, and also to obtain additional evaluations of lineups made with uniform fitted value distribution to ensure consistency in the results.
+Data collection period III was primarily to obtain Rorschach lineup evaluations to estimate $\alpha$ (see Section A.1 in the Appendix), and also to obtain additional evaluations of lineups made with uniform fitted value distribution to ensure consistency in the results.
 To construct a Rorschach lineup, the data is generated from a model with zero effect size, while the null data are generated using the same simulation method discussed in Section \ref{visual-test-procedure-based-on-lineups}.
@@
 Each participant received two of the `r vi_survey %>% filter(attention_check) %>% count(unique_lineup_id) %>% nrow()` attention check lineups which were used to filter results of participants who were clearly not making an honest effort (only 11 of `r vi_survey %>% count(set, exp) %>% nrow() + 11`).
-To estimate $\alpha$ for calculating statistical significance (see Section \ref{sig}) there were `r nrow(filter(vi_survey, null_lineup, !attention_check))` evaluations of `r vi_survey %>% filter(null_lineup, !attention_check) %>% count(unique_lineup_id) %>% nrow()` null lineups.
+To estimate $\alpha$ for calculating statistical significance (see Section A.1 in the Appendix) there were `r nrow(filter(vi_survey, null_lineup, !attention_check))` evaluations of `r vi_survey %>% filter(null_lineup, !attention_check) %>% count(unique_lineup_id) %>% nrow()` null lineups.
 Neither the attention checks nor null lineups were used in the subsequent analysis.
@@
 More data was collected on the uniform distribution (each evaluated by 11 participants) than the others (each evaluated by 5 participants).
-The analysis in Sections \ref{power-analysis}-\ref{hetero-analysis} uses only results from lineups generated with uniform fitted values, for a total `r nrow(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"))` lineup evaluations.
+The analysis in Sections \ref{power-analysis}--\ref{hetero-analysis} uses only results from lineups generated with uniform fitted values, for a total `r nrow(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"))` lineup evaluations.
 This allows us to compare the conventional and visual test performance in an optimal scenario.
@@
 
-Figures \ref{fig:polypower} and \ref{fig:heterpower} present the power curves of various tests plotted against the effect size in the residuals for non-linearity and heteroskedasticity respectively.
+Figure \ref{fig:nonlinearheterpower} present the power curves of various tests plotted against the effect size in the residuals for non-linearity and heteroskedasticity.
 In each case the power of visual test is calculated for multiple bootstrap samples leading to the many (orange) curves.
@@
 # Use area() to design layout
-map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
+A <- map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
   reduce(c) %>%
   c(patchwork::area(3, 1, 20, 11), .) %>%
-  patchwork::wrap_plots(append(plist, list(p), after = 0), 
+  patchwork::wrap_plots(append(plist, list(p + labs(title = "(A) Non-linearity") + theme(plot.title = element_text(size = 10))), after = 0), 
                         design = .)
@@
 
-```{r heterpower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only).
+```{r nonlinearheterpower, fig.cap = "Comparison of power between different tests for (A) non-linear and (B) heteroskedasticity patterns (uniform fitted values only).
 Main plot shows the power curves, with dots indicating non-reject and reject in visual testing of lineups.
 The multiple lines for the visual test arise from estimating the power on many bootstrap samples.
-The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.", dev = 'png', dpi = 450}
+The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.", fig.height = 6.5, fig.width = 5, dev = 'png', dpi = 450}
 
@@
 # Use area() to design layout
-map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
+B <- map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
   reduce(c) %>%
   c(patchwork::area(3, 1, 20, 11), .) %>%
-patchwork::wrap_plots(append(plist, list(p), after = 0), 
+patchwork::wrap_plots(append(plist, list(p + labs(title = "(B) Heteroskedasticity") + theme(plot.title = element_text(size = 10))), after = 0), 
                       design = .)
-```
-
-Figure \ref{fig:polypower} compares the power for the different tests for non-linear structure in the residuals.
+
+A / B + patchwork::plot_layout(guides = "collect")
+```
+
+Figure \ref{fig:nonlinearheterpower}A compares the power for the different tests for non-linear structure in the residuals.
 The test with the uniformly higher power is the RESET test, one that specifically tests for non-linearity.
@@
 In other words, the RESET test is more sensitive than the visual test.
-This is not necessarily a good feature for the purposes of diagnosing model defects: If we scan the residual plot examples at the bottom, we might argue that the non-linearity is not sufficiently problematic until an effect size of around 3 or 3.5.
-The RESET test would reject closer to an effect size of 2, but the visual test would reject closer 3.25, for a significance level of 0.05.
+This is not necessarily a good feature for the purposes of diagnosing model defects: if we scan the residual plot examples at the bottom, we might argue that the non-linearity is not sufficiently problematic until an effect size of around 3 or 3.5.
+The RESET test would reject closer to an effect size of 2, but the visual test would reject closer to 3.25, for a significance level of 0.05.
 The visual test matches the robustness of the model to (minor) violations of assumptions much better.
@@
 For the heteroskedasticity pattern, the power of BP test, designed for detecting heteroskedasticity, is uniformly higher than the other tests.
-The visual test power curve is a right shift.
+The visual test power curve shifts to the right.
 This shows a similar story to the power curves for non-linearity pattern: the conventional test is more sensitive than the visual test.
@@
 Figure \ref{fig:heter-example} shows this lineup.
-The data plot in position seventeen displays a relatively strong heteroskedasticity pattern, and has a strong effect size ($log_e(E)=4.02$), which is reflected by the visual test $p\text{-value} = 0.026$.
+The data plot in position seventeen displays a relatively strong heteroskedasticity pattern, and has a strong effect size ($\log_e(E)=4.02$), which is reflected by the visual test $p\text{-value} = 0.026$.
 But the BP test $p\text{-value} = 0.056$, is slightly above the significance cutoff of $0.05$.
@@
 
-Because the power curve of the visual tests are a shift to the right of the conventional test (Figures \ref{fig:polypower} and \ref{fig:heterpower}) we examined whether adjusting the significance level (to .001, .0001, .00001, ...) of the conventional test would generate similar decisions to that of the visual test.
-Interestingly, it doesn't: despite resulting less rejections, neither the RESET or BP tests come to complete agreement with the visual test (see Appendix).
+Because the power curve of the visual tests are a shift to the right of the conventional test (Figure \ref{fig:nonlinearheterpower}) we examined whether adjusting the significance level (to .001, .0001, .00001, ...) of the conventional test would generate similar decisions to that of the visual test.
+Interestingly, it doesn't: despite resulting in less rejections, neither the RESET or BP tests come to complete agreement with the visual test (see Appendix).
 
@@
 The butterfly has higher power in both tests.
-Curiously, the visual test has slightly higher power for the "left-triangle" than the "right-triangle" shape, when it would be expected that they should be identical, which is observed in conventional testing.'}
+Curiously, the visual test has a slightly higher power for the "left-triangle" than the "right-triangle" shape, when it would be expected that they should be similar, which is observed in conventional testing.'}
 

### 2023-09-20 paper.Rmd
done

@@
 A plot is compared with draws from a null model, in much the same way as a test statistic is compared to its sampling distribution.
-It is important to define the data plot using the language provided by a grammar of graphics (e.g.
-using @ggplot2) because this provides the functional mapping of abstract variables to plot elements.
-It is also important to remove plot elements that might introduce bias, such as axis labels, text and legends, or to make them generic.
+It is important to remove plot elements that might introduce bias, such as axis labels, text and legends, or to make them generic.
 
@@
 The conventional tests will be more sensitive to small departures from the null.
-They will also fail to detect departures when delivered residuals that have some contamination, like outliers or anomalies, as is often encountered when working with data.
+They will also fail to detect departures when residuals have some contamination, like outliers or anomalies, as is often encountered when working with data.
 The lineup approach is well-suited for generally interpreting data plots, and also detecting unexpected patterns not related to the model.
-This is supported by earlier research (e.g.
-@wickham2010, @roy2015using, @loy2015you, @VanderPlas2015, @loy2016variations).
+This is supported by earlier research [e.g.
+@wickham2010; @roy2015using; @loy2015you; @VanderPlas2015; @loy2016variations].
 
@@
 That is, a conventional test concludes there are problems with the model fit almost twice as often as a human.
-Conventional tests often reject when departures in the form of non-linearity and heteroskedasticity are not visibly different from null residual plots.
+Conventional tests often reject the null hypothesis when departures in the form of non-linearity and heteroskedasticity are not visibly different from null residual plots.
 

### 2023-09-20 paper.tex
done

@@
 various types of departures from this ideal pattern. Non-linearity,
-heteroskedasticity and non-normality, shown in Figure
-\ref{fig:residual-plot-common-departures}B,C,D are three commonly
-checked departures.
+heteroskedasticity and non-normality, shown in Figures
+\ref{fig:residual-plot-common-departures}B,
+\ref{fig:residual-plot-common-departures}C, and
+\ref{fig:residual-plot-common-departures}D are three commonly checked
+departures.
 
@@
 non-normality is usually detected with a different type of residual
-plot: a histogram or normal probability plot. Because we focus on
+plot: a histogram or a normal probability plot. Because we focus on
 scatterplots, non-normality is not one of the departures examined in
@@
 the assumptions of linearity, normality and constant error variance for
-both the level-1 and level-2 residuals \citep[\citet{loy2014hlmdiag} and
-\citet{loy2015you}]{loy2013diagnostic}.
+both the level-1 and level-2 residuals
+\citep{loy2013diagnostic, loy2014hlmdiag, loy2015you}.
 
@@
 \noindent where \(\Lambda(.)\) is the standard logistic function given
-as \(\Lambda(z) = exp(z)/(1+exp(z))\). The effect size \(E\) is the only
-predictor and the intercept is fixed to \(log(0.05/0.95)\) so that
-\(\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05\), the desired
+as \(\Lambda(z) = \exp(z)(1+\exp(z))^{-1}\). The effect size \(E\) is
+the only predictor and the intercept is fixed to \(\log(0.05/0.95)\) so
+that \(\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05\), the desired
 significance level.
@@
 
-\caption{\label{tab:model-factor-table}Levels of the factors used in data collection periods I, II, III.}
+\caption{\label{tab:model-factor-table}Levels of the factors used in data collection periods I, II, and III.}
 \centering
@@
 \(\boldsymbol{\varepsilon}\), \(\boldsymbol{x}_{raw}\),
-\(\boldsymbol{z}_{raw}\) are vectors of size \(n\), \(He_{j}(.)\) is the
-\(j\)th-order probabilist's Hermite polynomials,
-\(\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\), and
+\(\boldsymbol{z}_{raw}\) are vectors of size \(n\), \(\boldsymbol{1}_n\)
+is a vector of ones of size \(n\), \(He_{j}(.)\) is the \(j\)th-order
+probabilist's Hermite polynomials
+\citetext{\citealp{hermite1864nouveau}; \citealp[originally
+by][]{de1820theorie}},
+\(\varepsilon \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)\), and
 \(g(\boldsymbol{x}, k)\) is a scaling function to enforce the support of
@@
 \begin{equation} \label{eq:scaling-function}
-g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x}))2k - k, \quad \text{for} \quad k > 0. 
+g(\boldsymbol{x}, k) = 2k\cdot\frac{\boldsymbol{x} - x_{min}\boldsymbol{1}_n}{x_{max} - x_{min}} - k, \quad \text{for} \quad k > 0. 
 \end{equation}
 
-According to \citet{abramowitz1964handbook}, Hermite polynomials were
-initially defined by \citet{de1820theorie}, but named after
-\citet{hermite1864nouveau} because of the unrecognisable form of
-Laplace's work. The function \texttt{hermite} from the R package
-\texttt{mpoly} \citep{mpoly} is used to simulate
+\noindent where
+\(\displaystyle x_{min} = \min_{i\in \{1, ..., n\}} x_i\),
+\(\displaystyle x_{max} = \max_{i\in \{1, ..., n\}} x_i\) and \(x_i\) is
+the \(i\)-th entry of \(\boldsymbol{x}\). The function \texttt{hermite}
+from the R package \texttt{mpoly} \citep{mpoly} is used to simulate
 \(\boldsymbol{z}_{raw}\) to generate Hermite polynomials.
@@
 \(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\).
-Since \(z = O(x^j)\), for \(j > 1\), \(z\) is a higher order term left
-out of the null regression, which will lead to model misspecification.
+Here \(\boldsymbol{z}\) is a higher order term of \(\boldsymbol{x}\) and
+leaving it out in the null regression results in model misspecification.
 
@@
 result in a curve with more turning points. It is expected that the
-``U'' shape will be the easiest to detect, as the shape gets more
+``U'' shape will be the easiest to detect, and as the shape gets more
 complex it will be harder to perceive in a scatterplot, particularly
@@
 
-Figure \ref{fig:example-poly-lineup} is one of the non-linearity lineups
-used in experiment. This data plot is produced by the non-linearity
-model with \(j = 6\), and is in location \(2^3+1\). All five
-participants who viewed the lineup chose this plot as most different,
-thus all detected the data plot.
-
-\begin{figure}[t!]
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/example-poly-lineup-1}
-
-}
-
-\caption{One of the lineups containing non-linearity patterns used in data collection period I.
-Can you spot the most different plot?
-The data plot is positioned at $2^3 + 1$.}\label{fig:example-poly-lineup}
-\end{figure}
-
 \hypertarget{heteroskedasticity}{%
@@
 \begin{align*} \label{eq:heter-model}
-\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
+\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
 \boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
-\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}), 
+\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}_n, \boldsymbol{1}_n + (2 - |a|)(\boldsymbol{x} - a\boldsymbol{1}_n)'(\boldsymbol{x} - a\boldsymbol{1}_n) b \boldsymbol{I}_n), 
 \end{align*}
@@
 
-\begin{figure}[t!]
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/example-heter-lineup-1}
-
-}
-
-\caption{One of the lineups containing heteroskedasticity pattern used in data collection period II.
-Can you spot the most different plot?
-The data plot is positioned at $3^3 - 3^2$.}\label{fig:example-heter-lineup}
-\end{figure}
-
-An example lineup of this model used in data collection period II is
-shown in Figure \ref{fig:example-heter-lineup} with \(a = -1\). The data
-plot location is in position \(2^4 + 2\). Nine out of 11 participants
-chose this plot from the lineup.
-
 \hypertarget{factors-common-to-both-data-collection-periods}{%
@@
 always present in the collected data. Sometimes the fitted values are
-discrete because one or more predictors were measured discretely. The
-distribution may be relatively Gaussian, reflecting a linear combination
-of many predictors, adhering to the Central Limit Theorem. It is also
-common to see a skewed distribution of fitted values if one or more of
-the predictors has a skewed distribution. This latter problem is usually
-corrected before modelling, using a variable transformation. Our
+discrete because one or more predictors were measured discretely. It is
+also common to see a skewed distribution of fitted values if one or more
+of the predictors has a skewed distribution. This latter problem is
+usually corrected before modelling, using a variable transformation. Our
 simulation assess this by using four different distributions to
@@
 visible in a scatterplot reliably.
-\textcolor{red}{Beyond 300, the scatterplot should probably be used with transparency or replaced with a density or binned plot. Scatterplots suffer from over-plotting.}
+\textcolor{red}{Beyond 300, the scatterplot should probably be used with transparency or replaced with a density or binned plot as scatterplots suffer from over-plotting.}
 
@@
 equation, \(\boldsymbol{\beta}_z\) contains the corresponding
-coefficients, and \(\sigma^2\boldsymbol{I}\) is the assumed covariance
+coefficients, and \(\sigma^2\boldsymbol{I}_n\) is the assumed covariance
 matrix of the error term when \(H_0\) is true.
@@
 To compute the effect size for each lineup we simulate a sufficient
-large number of samples from the same model, in each sample, the number
-of observations \(n\) is fixed. We then compute the effect size for each
-sample and take the average as the final value. This ensures lineups
-constructed with the same experimental factors will share the same
-effect size.
+large number of samples from the same model with the number of
+observations \(n\) fixed for each sample. We then compute the effect
+size for each sample and take the average as the final value. This
+ensures lineups constructed with the same experimental factors will
+share the same effect size.
 
@@
 Data collection period III was primarily to obtain Rorschach lineup
-evaluations to estimate \(\alpha\) from Equation
-\ref{eq:pvalue-beta-binomial}, and also to obtain additional evaluations
-of lineups made with uniform fitted value distribution to ensure
-consistency in the results. To construct a Rorschach lineup, the data is
-generated from a model with zero effect size, while the null data are
-generated using the same simulation method discussed in Section
+evaluations to estimate \(\alpha\) (see Section A.1 in the Appendix),
+and also to obtain additional evaluations of lineups made with uniform
+fitted value distribution to ensure consistency in the results. To
+construct a Rorschach lineup, the data is generated from a model with
+zero effect size, while the null data are generated using the same
+simulation method discussed in Section
 \ref{visual-test-procedure-based-on-lineups}. This procedure differs
@@
 (only 11 of 454). To estimate \(\alpha\) for calculating statistical
-significance (see Section \ref{sig}) there were 720 evaluations of 36
-null lineups. Neither the attention checks nor null lineups were used in
-the subsequent analysis. The de-identified data, \texttt{vi\_survey}, is
-made available in the R package, \texttt{visage}.
+significance (see Section A.1 in the Appendix) there were 720
+evaluations of 36 null lineups. Neither the attention checks nor null
+lineups were used in the subsequent analysis. The de-identified data,
+\texttt{vi\_survey}, is made available in the R package,
+\texttt{visage}.
 
@@
 participants) than the others (each evaluated by 5 participants). The
-analysis in Sections \ref{power-analysis}-\ref{hetero-analysis} uses
+analysis in Sections \ref{power-analysis}--\ref{hetero-analysis} uses
 only results from lineups generated with uniform fitted values, for a
@@
 
-Figures \ref{fig:polypower} and \ref{fig:heterpower} present the power
-curves of various tests plotted against the effect size in the residuals
-for non-linearity and heteroskedasticity respectively. In each case the
-power of visual test is calculated for multiple bootstrap samples
-leading to the many (orange) curves. The effect size was computed at a
-5\% significance level and plotted on a natural logarithmic scale. To
-facilitate visual calibration of effect size values with the
-corresponding diagnostic plots, a sequence of example residual plots
-with increasing effect sizes is provided at the bottom of these figures.
-These plots serve as a visual aid to help readers understand how
-different effect size values translate to changes in the diagnostic
-plots. The horizontal lines of dots at 0 and 1 represent the
-non-rejection or rejection decisions made by visual tests for each
-lineup.
+Figure \ref{fig:nonlinearheterpower} present the power curves of various
+tests plotted against the effect size in the residuals for non-linearity
+and heteroskedasticity. In each case the power of visual test is
+calculated for multiple bootstrap samples leading to the many (orange)
+curves. The effect size was computed at a 5\% significance level and
+plotted on a natural logarithmic scale. To facilitate visual calibration
+of effect size values with the corresponding diagnostic plots, a
+sequence of example residual plots with increasing effect sizes is
+provided at the bottom of these figures. These plots serve as a visual
+aid to help readers understand how different effect size values
+translate to changes in the diagnostic plots. The horizontal lines of
+dots at 0 and 1 represent the non-rejection or rejection decisions made
+by visual tests for each lineup.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/polypower-1}
-
-}
-
-\caption{Comparison of power between different tests for non-linear patterns (uniform fitted values only).
-The power curves are estimated using logistic regression, and the horizontal lines of dots represent non-reject and reject results from visual tests for each lineup.
-The visual test has multiple power curves estimated from bootstrap samples.
-The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}\label{fig:polypower}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/heterpower-1}
-
-}
-
-\caption{Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only).
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/nonlinearheterpower-1}
+
+}
+
+\caption{Comparison of power between different tests for (A) non-linear and (B) heteroskedasticity patterns (uniform fitted values only).
 Main plot shows the power curves, with dots indicating non-reject and reject in visual testing of lineups.
 The multiple lines for the visual test arise from estimating the power on many bootstrap samples.
-The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}\label{fig:heterpower}
+The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}\label{fig:nonlinearheterpower}
 \end{figure}
 
-Figure \ref{fig:polypower} compares the power for the different tests
-for non-linear structure in the residuals. The test with the uniformly
-higher power is the RESET test, one that specifically tests for
-non-linearity. Note that the BP and SW tests have much lower power,
-which is expected because they are not designed to detect non-linearity.
-The bootstrapped power curves for the visual test are effectively a
-shift right from that of the RESET test. This means that the RESET test
-will reject at a lower effect size (less structure) than the visual
-test, but otherwise the performance will be similar. In other words, the
-RESET test is more sensitive than the visual test. This is not
-necessarily a good feature for the purposes of diagnosing model defects:
-If we scan the residual plot examples at the bottom, we might argue that
-the non-linearity is not sufficiently problematic until an effect size
-of around 3 or 3.5. The RESET test would reject closer to an effect size
-of 2, but the visual test would reject closer 3.25, for a significance
-level of 0.05. The visual test matches the robustness of the model to
-(minor) violations of assumptions much better.
+Figure \ref{fig:nonlinearheterpower}A compares the power for the
+different tests for non-linear structure in the residuals. The test with
+the uniformly higher power is the RESET test, one that specifically
+tests for non-linearity. Note that the BP and SW tests have much lower
+power, which is expected because they are not designed to detect
+non-linearity. The bootstrapped power curves for the visual test are
+effectively a shift right from that of the RESET test. This means that
+the RESET test will reject at a lower effect size (less structure) than
+the visual test, but otherwise the performance will be similar. In other
+words, the RESET test is more sensitive than the visual test. This is
+not necessarily a good feature for the purposes of diagnosing model
+defects: if we scan the residual plot examples at the bottom, we might
+argue that the non-linearity is not sufficiently problematic until an
+effect size of around 3 or 3.5. The RESET test would reject closer to an
+effect size of 2, but the visual test would reject closer to 3.25, for a
+significance level of 0.05. The visual test matches the robustness of
+the model to (minor) violations of assumptions much better.
 
@@
 detecting heteroskedasticity, is uniformly higher than the other tests.
-The visual test power curve is a right shift. This shows a similar story
-to the power curves for non-linearity pattern: the conventional test is
-more sensitive than the visual test. From the example residual plots at
-the bottom we might argue that the heteroskedasticity becomes noticeably
-visible around an effect size of 3 or 3.5. However the BP test would
-reject at around effect size 2.5. Interestingly, the power curve for the
-SW test (for non-normality) is only slightly different to that of the
-visual test, suggesting that it performs reasonably for detecting
-heteroskedasticity, too. The power curve for the BP test suggests it is
-not useful for detecting heteroskedasticity, as expected.
+The visual test power curve shifts to the right. This shows a similar
+story to the power curves for non-linearity pattern: the conventional
+test is more sensitive than the visual test. From the example residual
+plots at the bottom we might argue that the heteroskedasticity becomes
+noticeably visible around an effect size of 3 or 3.5. However the BP
+test would reject at around effect size 2.5. Interestingly, the power
+curve for the SW test (for non-normality) is only slightly different to
+that of the visual test, suggesting that it performs reasonably for
+detecting heteroskedasticity, too. The power curve for the BP test
+suggests it is not useful for detecting heteroskedasticity, as expected.
 
@@
 relatively strong heteroskedasticity pattern, and has a strong effect
-size (\(log_e(E)=4.02\)), which is reflected by the visual test
+size (\(\log_e(E)=4.02\)), which is reflected by the visual test
 \(p\text{-value} = 0.026\). But the BP test \(p\text{-value} = 0.056\),
@@
 Because the power curve of the visual tests are a shift to the right of
-the conventional test (Figures \ref{fig:polypower} and
-\ref{fig:heterpower}) we examined whether adjusting the significance
-level (to .001, .0001, .00001, \ldots) of the conventional test would
-generate similar decisions to that of the visual test. Interestingly, it
-doesn't: despite resulting less rejections, neither the RESET or BP
-tests come to complete agreement with the visual test (see Appendix).
+the conventional test (Figure \ref{fig:nonlinearheterpower}) we examined
+whether adjusting the significance level (to .001, .0001, .00001,
+\ldots) of the conventional test would generate similar decisions to
+that of the visual test. Interestingly, it doesn't: despite resulting in
+less rejections, neither the RESET or BP tests come to complete
+agreement with the visual test (see Appendix).
 
@@
 The butterfly has higher power in both tests.
-Curiously, the visual test has slightly higher power for the "left-triangle" than the "right-triangle" shape, when it would be expected that they should be identical, which is observed in conventional testing.}\label{fig:heter-power-uniform-a}
+Curiously, the visual test has a slightly higher power for the "left-triangle" than the "right-triangle" shape, when it would be expected that they should be similar, which is observed in conventional testing.}\label{fig:heter-power-uniform-a}
 \end{figure}
@@
 same way as a test statistic is compared to its sampling distribution.
-It is important to define the data plot using the language provided by a
-grammar of graphics (e.g.~using \citet{ggplot2}) because this provides
-the functional mapping of abstract variables to plot elements. It is
-also important to remove plot elements that might introduce bias, such
+It is important to remove plot elements that might introduce bias, such
 as axis labels, text and legends, or to make them generic.
@@
 departures from the null. They will also fail to detect departures when
-delivered residuals that have some contamination, like outliers or
-anomalies, as is often encountered when working with data. The lineup
-approach is well-suited for generally interpreting data plots, and also
-detecting unexpected patterns not related to the model. This is
-supported by earlier research (e.g. \citet{wickham2010},
-\citet{roy2015using}, \citet{loy2015you}, \citet{VanderPlas2015},
-\citet{loy2016variations}).
+residuals have some contamination, like outliers or anomalies, as is
+often encountered when working with data. The lineup approach is
+well-suited for generally interpreting data plots, and also detecting
+unexpected patterns not related to the model. This is supported by
+earlier research
+\citep[e.g.][]{wickham2010, roy2015using, loy2015you, VanderPlas2015, loy2016variations}.
 
@@
 fit almost twice as often as a human. Conventional tests often reject
-when departures in the form of non-linearity and heteroskedasticity are
-not visibly different from null residual plots.
+the null hypothesis when departures in the form of non-linearity and
+heteroskedasticity are not visibly different from null residual plots.
 
@@
 
-These \texttt{R} packages are used for the work: \texttt{cli}
-\citep{cli}, \texttt{curl} \citep{curl}, \texttt{dplyr} \citep{dplyr},
-\texttt{ggplot2} \citep{ggplot2}, \texttt{jsonlite} \citep{jsonlite},
-\texttt{lmtest} \citep{lmtest}, \texttt{mpoly} \citep{mpoly},
-\texttt{progress} \citep{progress}, \texttt{tibble} \citep{tibble},
-\texttt{ggmosaic} \citep{ggmosaic}, \texttt{purrr} \citep{purrr},
-\texttt{tidyr} \citep{tidyr}, \texttt{readr} \citep{readr},
-\texttt{stringr} \citep{stringr}, \texttt{here} \citep{here},
-\texttt{kableExtra} \citep{kableextra}, \texttt{patchwork}
-\citep{patchwork}, \texttt{rcartocolor} \citep{rcartocolor}. The study
-website is powered by \texttt{PythonAnywhere} \citep{pythonanywhere} and
-the \texttt{Flask} web framework \citep{flask}. The \texttt{jsPsych}
-framework \citep{jspsych} is used to create behavioural experiments that
-run in our study website.
+These \texttt{R} packages are used for the work: \texttt{tidyverse}
+\citep{tidyverse}, \texttt{lmtest} \citep{lmtest}, \texttt{mpoly}
+\citep{mpoly}, \texttt{ggmosaic} \citep{ggmosaic}, \texttt{kableExtra}
+\citep{kableextra}, \texttt{patchwork} \citep{patchwork},
+\texttt{rcartocolor} \citep{rcartocolor}. The study website is powered
+by \texttt{PythonAnywhere} \citep{pythonanywhere} and the \texttt{Flask}
+web framework \citep{flask}. The \texttt{jsPsych} framework
+\citep{jspsych} is used to create behavioural experiments that run in
+our study website.
 

### 2023-09-22 JCGS_response_to_reviews/response_to_review.Rmd
mention red text in response to reviews
response tweaked

@@
 
-We thank the anonymous reviewers for their careful reviews. What follows is our point-by-point response to reviewer comments.
+We thank the anonymous associate editor and reviewers for their careful reviews. What follows is our point-by-point response to reviewer comments.
 
 Note that *the original reviewer comments are in italic* and our response is in normal text.
+
+Text in the paper corresponding to the changes as a response to reviewers' comments are highlighted in red.
+We have also made a lot of minor changes smoothing the language and fixing small issues in the paper which are not highlighted.
 
@@
 
-The work specifically uses simulated data because we need to study the problem in controlled conditions.
+The work intentionally uses simulated data because we need to study the problem in controlled conditions.
 The reviewer is correct, though, that real data will pose more interesting challenges.
-It should be noted that our simulation setup do include discrete independent variables, and consequently somewhat discretized response.
+It should be noted that our simulation setup does include discrete independent variables, and consequently somewhat discretized response.
 The other point to note is that our null generating process uses residual rotation and will likely maintain irrelevant quirks of the data in the null plots as well.
-In that sense, the real data plot is not necessary always very obvious (see @buja2009statistical and @wickham2010 for examples).
+In that sense, the real data plot is not necessarily always obvious (see @buja2009statistical and @wickham2010 for examples).
 Nonetheless, our simulation does only test handful of conditions.
@@
 When the diagnostic plots are embedded in a lineup of nulls, it does have the advantage of providing a calibrated answer, and protects against the subjectiveness of hand-waving.
-We are advocating that you should always calibrate the diagnostic plot with a comparison with null plots.
+We are advocating that you should always calibrate the diagnostic plot by a comparison with null plots.
 We have strengthened this point of view by adding more explanation in the Introduction, in addition to what is provided in the Conclusions.
@@
 For our purposes, it was that the true residual plot could be detected from the null plots when the lineup was evaluated by people.
-We've added a sentence to the Conclusion mentioning this.
+We've added a sentence to the Conclusion clarifying this.
 
@@
 It can be argued that it is being used by practitioners based on download rates of the `nullabor` package (>19,000 downloads in the past year) which implements a process for creating lineups, and application articles that report usage.
-Using it formally, we agree is cumbersome and would require engaging a crowd-sourcing service, that involves a small expense.
-Science labs typically have a budget for equipment and experimental devices and we would advocate this is a useful addition.
+Using it formally, we agree requires the effort of engaging a crowd-sourcing service, involving a small expense.
+Science labs typically have a budget for equipment and experimental devices and we would advocate that this is a useful addition.
 Where a practitioner lacks the budget for running crowd-sourcing services, the protocol can be used with their team or network and can be an excellent way to engage their team/network with the data.
-We have definitely used it ourselves in @roychowdhuryUsingVisualStatistical2015a for ecological data and @mahbubulmajumderVisualMiningMethods2013 for gene expression analysis.
+We have formally used crowd-sourcing ourselves for application purposes in @roychowdhuryUsingVisualStatistical2015a for ecological data and @mahbubulmajumderVisualMiningMethods2013 for gene expression analysis.
 We have added some details in the "Limitations and practicality" section.
@@
 
-Thank you for the positive feedback. We have shortened the paper by XXX pages. We have expanded the explanations on the practicality in the Section "Limitations and practicality".
+Thank you for the positive feedback. We have shortened the paper by about five pages, while also adding a discussion on the practicality in the Section "Limitations and practicality".
 
@@
 
-The new Section "Limitations and practicality" addresses these concerns and we have added a couple of sentences to the Introduction.
+The new Section "Limitations and practicality" addresses these concerns and we have added a couple of sentences to the Introduction (highlighted in red).
 
@@
 They are indispensable, as reported by numerous readers during the proofing stage of the paper.
+
+# Additional references, also cited in main paper

### 2023-09-22 JCGS_response_to_reviews/response_to_review.Rmd
changes reviewed, and small fixes made

@@
 
-Thank you for the positive feedback. We have shortened the paper by about five pages, while also adding a discussion on the practicality in the Section "Limitations and practicality".
+Thank you for the positive feedback. We have shortened the paper by about four pages, while also adding a discussion on the practicality in the Section "Limitations and practicality".
 
@@
 - Part of the Experimental design has been moved into the Appendix.
+- Two of the full lineup plots were removed.
 
@@
 A plot other than a scatterplot is probably needed to better display a large number of observations.
-We have changed the language so that this is clearer.
+We have changed the language so that this is clearer (highlighted in red).
 

### 2023-09-22 paper.Rmd
changes reviewed, and small fixes made

@@
 
-\textcolor{red}{However, relying purely on subjective assessment of a single plot can be deceiving.
-Indeed, correctly judging whether \emph{no} pattern exists in a residual plot is a difficult task.
-People will almost always see a pattern \citep[see ][]{kahneman2011thinking}, so the question that really needs answering is whether any pattern perceived is consistent with randomness, or sampling variability, or noise.
+\textcolor{red}{Additionally, relying solely on the subjective assessment of a single plot can be problematic.
+People will almost always see a pattern \citep[see][]{kahneman2011thinking}, so the question that really needs answering is whether any pattern perceived is consistent with randomness, or sampling variability, or noise.
+Correctly judging whether \emph{no} pattern exists in a residual plot is a difficult task.
 \cite{loy2021bringing} emphasizes that this is especially difficult to teach to new analysts and students, and advocates to the broader use of the lineup protocol \citep{bujastatistical2009}.}
@@
 For residual analysis this is especially helpful for gauging whether there is \emph{no} pattern.
-In its strict use, one would insist that the data plot is not seen before seeing the lineup, so that the observer doesn't know which is the true plot.
+(Figure \ref{fig:first-example-lineup} shows an example of a lineup of residual plots.) In its strict use, one would insist that the data plot is not seen before seeing the lineup, so that the observer doesn't know which is the true plot.
 When used this way, it provides an objective test for data plots.
@@
 We conclude with a discussion of the presented work, and ideas for future directions.
+
+```{r first-example-lineup, fig.pos="t!", fig.height = 7, fig.width = 7, fig.cap = "Visual testing is conducted using a lineup, as in the example here.
+The residual plot computed from the observed data is embedded among 19 null plots, where the residuals are simulated from a standard error model.
+Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot.
+A small $p$-value would result from a substantial number selecting the data plot (at position $2^2 + 2$, exhibiting non-linearity)."}
+
+vi_lineup$poly_300$data %>%
+  VI_MODEL$plot_lineup(theme = theme_light(), remove_axis = TRUE, remove_grid_line = TRUE)
+```
 
@@
 This protocol requires a priori specification of $H_0$ (or at least a null data generating mechanism), much like the requirement of knowing the sampling distribution of the test statistic in null hypothesis significance testing framework.
-
-```{r first-example-lineup, fig.pos="t!", fig.height = 7, fig.width = 7, fig.cap = "Visual testing is conducted using a lineup, as in the example here.
-The residual plot computed from the observed data (plot $2^2 + 2$, exhibiting non-linearity) is embedded among 19 null plots, where the residuals are simulated from a standard error model.
-Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot.
-A small $p$-value would result from a substantial number selecting plot $2^2 + 2$."}
-
-vi_lineup$poly_300$data %>%
-  VI_MODEL$plot_lineup(theme = theme_light(), remove_axis = TRUE, remove_grid_line = TRUE)
-```
 
@@
   c(patchwork::area(3, 1, 20, 11), .) %>%
-  patchwork::wrap_plots(append(plist, list(p + labs(title = "(A) Non-linearity") + theme(plot.title = element_text(size = 10))), after = 0), 
-                        design = .)
+  patchwork::wrap_plots(append(plist, list(p + labs(title = "(A) Non-linearity") + theme(plot.title = element_text(size = 10))), after = 0), design = .)
 ```
@@
   c(patchwork::area(3, 1, 20, 11), .) %>%
-patchwork::wrap_plots(append(plist, list(p + labs(title = "(B) Heteroskedasticity") + theme(plot.title = element_text(size = 10))), after = 0), 
+patchwork::wrap_plots(append(plist, list(p + labs(title = "(B) Heteroskedasticity") + 
+  theme(plot.title = element_text(size = 10))), after = 0),
     design = .)
@@
 One of the primary limitations of the lineup protocol lies in its reliance on human judgements.
-In this context, the effectiveness of a single lineup evaluation is closely connected with the perceptual ability and visual skills of the individual.
-When results from multiple individuals are combined the outcome is encouragingly robust.
+In this context, the effectiveness of a single lineup evaluation can be dependent on the perceptual ability and visual skills of the individual.
+However, when results from multiple individuals are combined the outcome is encouragingly high-quality and robust.
 For simple plots and strong patterns just a few individuals are needed to arrive at a clear answer, but more individuals will be needed when the plot design is complex, or the signal strength is weak.
@@
 If the fitted value distribution is not uniform, there is a small loss of power in the visual test.
-Surprisingly there is a big difference in power of the conventional test across fitted value distributions.
-Another surprising finding was that the direction of heteroskedasticity appears to affect the ability to visually detect it: both triangles being more difficult to detect than the butterfly, and a small difference in detection between left- and right-triangle.
+Surprisingly, there is a big difference in power of the conventional test across fitted value distributions.
+Another unexpected finding was that the direction of heteroskedasticity appears to affect the ability to visually detect it: both triangles being more difficult to detect than the butterfly, and a small difference in detection between left- and right-triangle.
 

### 2023-09-25 JCGS_response_to_reviews/response_to_review.Rmd
remove unused/commented parts

@@
 
-<!-- (diagnostic plots may reveal features which are not directly connected to a linear model assumption; You have restricted attention to the residual-fitted plot and a particular selection of formal tests; Nevertheless, it would be good to add some discussion of the wider picture.)-->
-
 *1.
@@
 We have addressed this limited scope in the broader discussion of the use of residual plots and lineups for residual diagnostics added to the new Section "Limitations and practicality".
-<!-- (see issue #11).-->
-
-<!--
-ET: I don't think talking about computer vision model makes sense here as the reviewer would not know about our other work?
-
-An observational study involving actual residual plots will hopefully be possible once a computer vision model is deployed that will allow analysts to upload their residuals for automated testing.
--->
 
@@
 
-<!--(Is there a paper to refer the reader to?)-->
-
 *4.
@@
 We have added some details in the "Limitations and practicality" section.
-<!--We agree that it is not widespread, and think that this is because it's a little unwieldy.
-The goal is to make it easier, particularly by employing an automated testing tool using computer vision, but this is not the subject of the current paper.-->
 
@@
 
-<!-- As mentioned above,
-
-This is not computationally expensive. The conventional test also make mistakes. You need something more than the conventional test. And you also need something more than the residual plot.
-
-check earlier paper.-->
-
 *4.

### 2023-09-25 appendix.Rmd
remove unused/commented parts

@@
 ```{r}
+if (!requireNamespace("remotes", quietly = TRUE)) {
+  install.packages("remotes")
+}
+
 # OOP supports needed by `visage`
-# remotes::install_github("TengMCing/bandicoot")
-# 
+if (!requireNamespace("bandicoot", quietly = TRUE)) {
+  remotes::install_github("TengMCing/bandicoot")
+}
+
 # Visual inference models and p-value calculation
-# remotes::install_github("TengMCing/visage")
+if (!requireNamespace("visage", quietly = TRUE)) {
+  remotes::install_url("https://github.com/TengMCing/visage/raw/master/built/visage_0.1.2.tar.gz")
+} else {
+  if (packageVersion("visage") < "0.1.2") {
+    remotes::install_url("https://github.com/TengMCing/visage/raw/master/built/visage_0.1.2.tar.gz")
+  }
+}
+
+if (!requireNamespace("tidyverse", quietly = TRUE)) {
+  install.packages("tidyverse")
+}
+
+if (!requireNamespace("here", quietly = TRUE)) {
+  install.packages("here")
+}
+
+if (!requireNamespace("kableExtra", quietly = TRUE)) {
+  install.packages("kableExtra")
+}
+
+if (!requireNamespace("ggpubr", quietly = TRUE)) {
+  install.packages("ggpubr")
+}
+
+if (!requireNamespace("lvplot", quietly = TRUE)) {
+  install.packages("lvplot")
+}
 

### 2023-09-25 paper.Rmd
remove unused/commented parts

@@
 ```{r}
+if (!requireNamespace("remotes", quietly = TRUE)) {
+  install.packages("remotes")
+}
+
 # OOP supports needed by `visage`
-# remotes::install_github("TengMCing/bandicoot")
-# 
+if (!requireNamespace("bandicoot", quietly = TRUE)) {
+  remotes::install_github("TengMCing/bandicoot")
+}
+
 # Visual inference models and p-value calculation
-# remotes::install_url("https://github.com/TengMCing/visage/raw/master/built/visage_0.1.0.tar.gz")
+if (!requireNamespace("visage", quietly = TRUE)) {
+  remotes::install_url("https://github.com/TengMCing/visage/raw/master/built/visage_0.1.2.tar.gz")
+} else {
+  if (packageVersion("visage") < "0.1.2") {
+    remotes::install_url("https://github.com/TengMCing/visage/raw/master/built/visage_0.1.2.tar.gz")
+  }
+}
+
+if (!requireNamespace("tidyverse", quietly = TRUE)) {
+  install.packages("tidyverse")
+}
+
+if (!requireNamespace("here", quietly = TRUE)) {
+  install.packages("here")
+}
+
+if (!requireNamespace("ggmosaic", quietly = TRUE)) {
+  install.packages("ggmosaic")
+}
+
+if (!requireNamespace("rcartocolor", quietly = TRUE)) {
+  install.packages("rcartocolor")
+}
+
+if (!requireNamespace("kableExtra", quietly = TRUE)) {
+  install.packages("kableExtra")
+}
+
+if (!requireNamespace("patchwork", quietly = TRUE)) {
+  install.packages("patchwork")
+}
 
@@
 
-<!--- ET: I suggest removing this to save space and we don't use this plot type.
-
-\textcolor{red}{There are also other types of plots that can be used for assessing residuals.
-The normal probability plot is used to assess a normality assumption on the error distribution.
-Residuals are often plotted against  order of observations to check for collection bias, batch effects or autocorrelation.
-Partial residual plots assess the importance of an independent variable to a model, conditional on other variables, by plotting partial residuals against this variable.
-Partial residuals are computed from the difference between the response value and the fitted model excluding term of the independent variable of interest.
-A partial regression plot assesses the unique effect of an independent variable in a multiple regression model, by plotting residuals from the model with and without the this variable against each other.}
-
---->
-
-<!--other types of residual plot to extract the available information and identify inappropriate assumptions, like the added variable plot (cite here) for studying the role of an external variable if it enters linearly into a model.
-Partial residual plots (cite here) have been suggested as computationally convenient substitutes for the added variable plots.
-For use in diagnostic procedures, several transformations of the ordinary residuals have been suggested to overcome partially some of their shortcomings.
-Studentized residuals (cite here) have null distributions that are independent of the scale parameters.
-These residuals are shown to be closely related to a mean shift model for outliers and to the residuals obtained when each case in turn is left out of the data.
-Alternatively.
-the residuals can be transformed to have a selected covariance structure: The usual suggestion is to obtain residuals with uncorrelated elements.}-->
-
-<!--ET: I think we need to remove this to reduce pages.
-
-These are all topics explained thoroughly in the sizeable literature on diagnostic procedures for linear models, e.g.
-@draper1998applied, @montgomery1982introduction, @belsley1980regression, @cook1999applied and @cook1982residuals.
-Interestingly, despite the abundance of conventional tests, _all_ of these writings advise that plotting residuals is an essential tool for diagnosing regression model problems:
-
-> *Some most useful checks allowed by data plots should be done on a routine basis for every regression.* [@draper1998applied]
-
-> *Such formal and informal procedures are complementary, and both have a place in residual analysis.* [@cook1982residuals]
-
-> *In our experience, statistical tests on regression model residuals are not widely used.
-In most practical situations the residual plots are more informative than the corresponding tests.
-However, since residual plots do require skill and experience to interpret, the statistical tests may occasionally prove useful.* [@montgomery1982introduction]
-
--->
-
 The common wisdom of experts is that plotting the residuals is indispensable for diagnosing model fits [@draper1998applied; @cook1982residuals; @montgomery1982introduction].
@@
 The experiment discussed in this paper specifically addresses this tension between statistical and practical statistics.
-
-<!--that is addressed by the results in this paper because from a practical perspective we would only want to flag the model as problematic when ther is a XXX diagnostic  practical significance.
-These can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers and slight departures.
--->
-
-<!--
-As discussed in Section \ref{conventionally-testing-for-departures}, many conventional tests are available for detecting residual departures.
-Implementation-wise, the built-in R package `stats` provides some commonly used residual-based tests, such as SW test.
-A more comprehensive collection of regression diagnostics tests can be found in the R package `lmtest` [@lmtest].
-In terms of heteroskedasticity diagnostics, the R package `skedastic` [@skedastic] collects and implements 25 existing conventional tests published since 1961.
-
-We pick RESET test (`resettest`) and BP test (`bptest`) from the R package `lmtest`, and SW test (`shapiro.test`) from the built-in R package `stats`.
-Among them, RESET test is the only exact and appropriate test in this scenario.
-Both the BP test and the SW test are approximate and inappropriate tests.
-Their estimated power is shown in Figure \ref{fig:polypower}.
-To set up the RESET test, we include different powers of fitted values as proxies.
-According to @ramsey1969tests, there are no general rules for the power of the fitted values needed by the RESET test, but it finds power up to four is usually sufficient.
-Thus, we follow this guideline to conduct the RESET test.
-For the BP test, the choice of predictors in the auxiliary regression is left to the user [@breusch1979simple].
-But as @waldman1983note suggested, it is a good choice for the set of auxiliary predictors in the BP test be the same as the White test.
-Thus, we include both $x$ and $x^2$ in the auxiliary regression.
--->
 
@@
 
-<!--- ET: interesting history, but not pertinent to the story
-
-According to @abramowitz1964handbook, Hermite polynomials were initially defined by @de1820theorie, but named after @hermite1864nouveau because of the unrecognisable form of Laplace's work. 
---->
-
 The null regression model used to fit the realizations generated by the above model is formulated as
@@
 Here $\boldsymbol{z}$ is a higher order term of $\boldsymbol{x}$ and leaving it out in the null regression results in model misspecification.
-
-<!--- ET: let's keep the language simpler.
-
-Since $z = O(x^j)$, for $j > 1$, $z$ is a higher order term left out of the null regression, it leads to model misspecification. 
---->
 
@@
 
-<!----
-
-Figure \ref{fig:example-poly-lineup} is one of the non-linearity lineups used in experiment.
-This data plot is produced by the non-linearity model with $j = 6$, and is in location $2^3+1$.
-All five participants who viewed the lineup chose this plot as most different, thus all detected the data plot.
-
-```{r example-poly-lineup, fig.pos="t!", fig.height = 7, fig.width = 7, fig.cap = "One of the lineups containing non-linearity patterns used in data collection period I.
-Can you spot the most different plot?
-The data plot is positioned at $2^3 + 1$."}
-VI_MODEL$plot_lineup(vi_lineup$poly_24$data, 
-                     remove_grid_line = TRUE, 
-                     theme = theme_light(),
-                     remove_axis = TRUE)
-```
-
---->
-
 ### Heteroskedasticity
@@
 
-<!---
-
-```{r example-heter-lineup, fig.pos="t!", fig.height = 7, fig.width = 7, fig.cap = "One of the lineups containing heteroskedasticity pattern used in data collection period II.
-Can you spot the most different plot?
-The data plot is positioned at $3^3 - 3^2$."}
-
-VI_MODEL$plot_lineup(vi_lineup$heter_471$data, 
-                     remove_grid_line = TRUE, 
-                     theme = theme_light(),
-                     remove_axis = TRUE)
-```
-
-An example lineup of this model used in data collection period II is shown in Figure \ref{fig:example-heter-lineup} with $a = -1$.
-The data plot location is in position $2^4 + 2$.
-Nine out of 11 participants chose this plot from the lineup.
-
---->
-
 ### Factors common to both data collection periods
@@
 - discrete, $U\{1, 5\}$.
-
-<!--- ET: previous reference was "normal" but now Gaussian. This doesn't add to the story so I suggest just remove.
-
-The distribution may be relatively Gaussian, reflecting a linear combination of many predictors, adhering to the Central Limit Theorem.
-
---->
 
@@
 
-<!--As mentioned in Section \ref{power-of-the-tests}, $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using null lineups.
-Three replications are made for $3 \times 4 = 12$ combinations of common factors $n$ and fitted value distribution, results in $12 \times 3 = 36$ lineups included in data collection period III.
-In these lineups, the data of the data plot is generated from a model with zero effect size, while the data of the 19 null plots are generated using the same simulation method discussed in Section \ref{visual-test-procedure-based-on-lineups}.
-This generation procedure differs from the canonical Rorschach lineup procedure, which requires that all 20 plots are generated directly from the null model.
-However, these lineups serve the same fundamental purpose: to assess the number of visually interesting plots generated under $H_0$.
-
-To account for the fact that our simulation method for these lineups is not the Rorschach procedure, we use the method suggested in @vanderplas2021statistical for typical lineups containing a data plot to estimate $\alpha$.
-(We have included a sensitivity analysis in the Appendix to examine the impact of the variance of the $\alpha$ estimate on our findings.)
-
-All lineups consist of only null plots are planned to be evaluated by 20 participants. 
-However, presenting only these lineups to participants are considered to be bad practices as participants will lose interest quickly. 
-Therefore, we plan to collect 6 more evaluations on the 279 lineups with uniform fitted value distribution, result in $(36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133$ participants recruited for data collection period III.
--->
-
 ### Collecting results
@@
 A participant's submission was only included in the analysis if the data plot is identified for at least one attention check.
-<!--Data of rejected submissions are discarded automatically to maintain the overall data quality.-->
 
@@
 
-# 106
-# 112
-# 113
-# 115
-# 132
-# 133
-# 144!
-# 147
 set.seed(132)
@@
 
-# 10091
-# 101
-# 105
-# 106
-# 112
-# 119
-# 133
 set.seed(135)
@@
 
-<!-- In terms of visual tests, we do not observe significant difference between the power curve of the uniform distribution and the power curve of the discrete uniform distribution.
-Normal and lognormal distributions produce lower power than uniform and discrete uniform distributions in regards of residual departure patterns, but the gap is larger when the residual plot contains heteroskedasticity pattern.
-This suggests that residual departures displayed with evenly distributed fitted values are visually more attractive and effective.
-Besides, the skewness of the fitted value distribution plays a role in the degree of presence of visual features as lognormal distribution produce the lowest power for both non-linearity and heteroskedasticity patterns.
-A symmetric fitted value distribution can better reveal the underlying shape.-->
-
 The variation in power in the conventional tests is at first sight, shocking.
@@
 
-<!-- Although it is unusual to consider fitted value distribution or predictor distribution in power analysis, it is not so surprised that the power of conventional tests vary because of fitted value distribution.
-For instance, the power of the RESET test which is effectively a F test depends on the realizations of the additional predictors included in the auxiliary regression equation [@jamshidian2007study; @olvera2019relationship; @zhang2018practical].
-And those additional predictors are linear transformation of random vectors following certain predictor distributions.
-What is truly unexpected is the huge variability in power of conventional tests compared to visual tests.
-The discrete uniform distribution constantly produce the highest power while the normal distribution and the lognormal distribution produce the lowest power for non-linearity and heteroskedasticity pattern respectively.
-The difference in power corresponding to different distributions peak at around 90% when the effect size is at a moderate level.
-This huge difference is undesirable as is not uncommon for analysts to omit the factor of predictor distribution in power calculation and sample size calculation.
--->
-
 # \textcolor{red}{Limitations and practicality}
@@
 For simple plots and strong patterns just a few individuals are needed to arrive at a clear answer, but more individuals will be needed when the plot design is complex, or the signal strength is weak.
-<!--This reliance on human judgment introduces an element of variability and subjectivity that can potentially impact the robustness of the results obtained.
-To ensure reliable outcomes, it's often necessary to include a diverse group of participants capable of offering unbiased judgements.
-This requirement has led researchers, including those in our study, to turn to crowd-sourcing platforms that offer access to a random selection of participants, which can help mitigate potential bias.-->
 
@@
 We do also expect that at some point a computer vision model can be developed to take over the task of employing people to evaluate residual plots.
-
-<!--the process of constructing and evaluating lineups becomes accessible and straightforward.
-This accessibility empowers a wider range of researchers, even those without extensive statistical backgrounds, to employ the lineup protocol in their studies.
-Moreover, the cost-effectiveness of lineup-based analyses makes them particularly appealing for integration into statistical education within classroom settings (cite adam loy paper here).
-The practicality of lineup demonstrations can facilitate a deeper understanding of the concept of randomness and statistical testing among students.-->
-
-<!--However, it's important to understand that if the goal is to attain results similar to those of formal hypothesis testing in rigor and reliability, using crowd-sourcing becomes necessary.
-The expense associated with crowd-sourcing is relatively reasonable when compared to the costs of other laboratory devices, making it a sensible investment for research that requires clear and valuable findings.
-Importantly, this higher level of scrutiny is typically justified only when the data plot can not be confidently identified and requires a thorough investigation.-->
 
@@
 
-<!-- Additionally, there is the exciting possibility of utilizing computer vision technology to automate the analysis of lineup of residual plots.
-This innovation has the potential to significantly reduce the expenses associated with recruiting human judges.
-By incorporating computer vision models, researchers can simplify the evaluation process, making it more efficient and cost-effective.
-
-The scope of this study is to highlight practical value of residual plots in model diagnostics, drawing comparisons with conventional tests.
-Our primary focus is on the use of residual-fitted plots in a controlled model environment for the purpose of a human subject experiment.
-It's important to note that there are other intriguing aspects of residual diagnostics that fall beyond the boundaries of this paper.
-For example, as mentioned in Introduction, residual plots can be used to explore interesting data characteristics while conducting model diagnostics.
-Furthermore, some model violations, such as serial correlation, necessitate the use of alternative residual plots like residual against row number plots.
-Nevertheless, we believe that the insights derived from this study have the potential to apply to other residual diagnostics setting, because the fundamental principles and statistical concepts underlying the lineup protocol and residual plots remain consistent across different contexts.
-In real-world situations, researchers often rely on visual assessment to make informed decisions about model validity.
-The lineup protocol's capability to identify deviations from randomness serves as a valuable aid in this scenario.-->
-
 # Conclusions
-
-<!-- Make a note about "arguing" that there is no problem apparent in residual plot - reading a single plot -  vs the visual test which removes the argument, and provides objectivity.-->
 
@@
 However, with the emergence of effective computer vision, it is hoped this work helps to lay the foundation for automated residual plot assessment.
-<!--The findings suggest the strong demand of graphical inspection in regression diagnostics, so developing an automatic visual inference system to evaluate lineups of residual plots is valuable.
-We plan to build a completed open-source system in an `R` package and provide a web interface such as a website for public to interact with.
-Details about this system will be discussed in our next paper.-->
 

### 2023-09-25 paper.tex
remove unused/commented parts

@@
 work, and ideas for future directions.
+
+\begin{figure}[t!]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/first-example-lineup-1}
+
+}
+
+\caption{Visual testing is conducted using a lineup, as in the example here.
+The residual plot computed from the observed data is embedded among 19 null plots, where the residuals are simulated from a standard error model.
+Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot.
+A small $p$-value would result from a substantial number selecting the data plot (at position $2^2 + 2$, exhibiting non-linearity).}\label{fig:first-example-lineup}
+\end{figure}
 
@@
 
-\begin{figure}[t!]
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/first-example-lineup-1}
-
-}
-
-\caption{Visual testing is conducted using a lineup, as in the example here.
-The residual plot computed from the observed data (plot $2^2 + 2$, exhibiting non-linearity) is embedded among 19 null plots, where the residuals are simulated from a standard error model.
-Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot.
-A small $p$-value would result from a substantial number selecting plot $2^2 + 2$.}\label{fig:first-example-lineup}
-\end{figure}
-
 Figure \ref{fig:first-example-lineup} is an example of a lineup
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/nonlinearheterpower-1}
-
-}
-
-\caption{Comparison of power between different tests for (A) non-linear and (B) heteroskedasticity patterns (uniform fitted values only).
-Main plot shows the power curves, with dots indicating non-reject and reject in visual testing of lineups.
-The multiple lines for the visual test arise from estimating the power on many bootstrap samples.
-The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}\label{fig:nonlinearheterpower}
-\end{figure}
-
 Figure \ref{fig:nonlinearheterpower}A compares the power for the
@@
 reliance on human judgements. In this context, the effectiveness of a
-single lineup evaluation is closely connected with the perceptual
-ability and visual skills of the individual. When results from multiple
-individuals are combined the outcome is encouragingly robust. For simple
-plots and strong patterns just a few individuals are needed to arrive at
-a clear answer, but more individuals will be needed when the plot design
-is complex, or the signal strength is weak.
+single lineup evaluation can be dependent on the perceptual ability and
+visual skills of the individual. However, when results from multiple
+individuals are combined the outcome is encouragingly high-quality and
+robust. For simple plots and strong patterns just a few individuals are
+needed to arrive at a clear answer, but more individuals will be needed
+when the plot design is complex, or the signal strength is weak.
 
@@
 not uniform, there is a small loss of power in the visual test.
-Surprisingly there is a big difference in power of the conventional test
-across fitted value distributions. Another surprising finding was that
-the direction of heteroskedasticity appears to affect the ability to
-visually detect it: both triangles being more difficult to detect than
-the butterfly, and a small difference in detection between left- and
-right-triangle.
+Surprisingly, there is a big difference in power of the conventional
+test across fitted value distributions. Another unexpected finding was
+that the direction of heteroskedasticity appears to affect the ability
+to visually detect it: both triangles being more difficult to detect
+than the butterfly, and a small difference in detection between left-
+and right-triangle.
 

### 2023-10-25 paper.tex
reknit the pdf file

@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/nonlinearheterpower-1}
+
+}
+
+\caption{Comparison of power between different tests for (A) non-linear and (B) heteroskedasticity patterns (uniform fitted values only).
+Main plot shows the power curves, with dots indicating non-reject and reject in visual testing of lineups.
+The multiple lines for the visual test arise from estimating the power on many bootstrap samples.
+The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}\label{fig:nonlinearheterpower}
+\end{figure}
+
 Figure \ref{fig:nonlinearheterpower}A compares the power for the

### 2024-01-01 JCGS_response_to_reviews/response_to_review_R1.Rmd
reduce 2 pages

---
title: "Response to Reviews: JCGS-23-139.R1"
author: Weihao Li, Emi Tanaka, Dianne Cook, Susan VanderPlas
output: pdf_document
date: "`r Sys.Date()`"
bibliography: ../ref.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We thank the anonymous associate editor and reviewers for their careful reviews. What follows is our point-by-point response to reviewer comments.

Note that *the original reviewer comments are in italic* and our response is in normal text.

# Reviewer 1

*Good response.*

Thank you. (...)

# Reviewer 2

*I appreciate the effort provided by the authors for both my comments and the other reviewer’s comments. The revision is much improved. I have a few remaining comments.*

Thank you. (...)

*1. I still believe the paper is too long. Starting with Section 4, I believe the paper starts to become a little too cumbersome. I suggest the authors look for ways to condense down Sections 4 and 5 a bit as I think these are the biggest contributors to the length.*

*2. In the Figure 1 caption, clarify what is meant by the "at position $2^2 + 2$".*

*3. In the footnote on the bottom of page6, replace "didn't" and "it's" by "did not" and "it is", respectively.*

*4. In the last sentence of Section 2.2, the final segment "statistical and practical statistics" sounds awkward because "statistical" should be serving as an adjective to another word, and the segment as written makes it sound like it is also an adjective to "statistics". The authors might consider replacing "statistical" with "classic inference" as the paper is addressing the discrepancies between inferential results and what may be practically done.*

*5. Page 21 onward, there are many instances where the figure references did not compile correctly. These need to be corrected.*

### 2024-01-01 appendix.Rmd
reduce 2 pages

@@
 \appendix
+
+# Additional details of experimental design
+
+## Non-linearity
+
+The non-linearity model used in the experiment includes a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials on random vector $\boldsymbol{x}$ formulated as
+
+\begin{align*} \label{eq:nonlinearity-model}
+\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
+\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
+\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
+\end{align*}
+
+\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials [@hermite1864nouveau; originally by @de1820theorie], $\varepsilon \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
+
+```{=tex}
+\begin{equation} \label{eq:scaling-function}
+g(\boldsymbol{x}, k) = 2k\cdot\frac{\boldsymbol{x} - x_{min}\boldsymbol{1}_n}{x_{max} - x_{min}} - k, \quad \text{for} \quad k > 0. 
+\end{equation}
+```
+
+\noindent where $\displaystyle x_{min} = \min_{i\in \{1, ..., n\}} x_i$, $\displaystyle x_{max} = \max_{i\in \{1, ..., n\}} x_i$ and $x_i$ is the $i$-th entry of $\boldsymbol{x}$.
+The function `hermite` from the R package `mpoly` [@mpoly] is used to simulate $\boldsymbol{z}_{raw}$ to generate Hermite polynomials.
+
+The null regression model used to fit the realizations generated by the above model is formulated as
+
+```{=tex}
+\begin{equation} \label{eq:null-model}
+\boldsymbol{y} = \beta_0\boldsymbol{1}_n + \beta_1 \boldsymbol{x} + \boldsymbol{u},
+\end{equation}
+```
+\noindent where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
+Here $\boldsymbol{z}$ is a higher order term of $\boldsymbol{x}$ and leaving it out in the null regression results in model misspecification.
+
+## Heteroskedasticity
+
+The heteroskedasticity model used in the experiment is formulated as
+
+\begin{align*} \label{eq:heter-model}
+\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
+\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}_n, \boldsymbol{1}_n + (2 - |a|)(\boldsymbol{x} - a\boldsymbol{1}_n)'(\boldsymbol{x} - a\boldsymbol{1}_n) b \boldsymbol{I}_n), 
+\end{align*}
+
+\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vectors of size $n$ and $g(.)$ is the scaling function defined in Equation \ref{eq:scaling-function}.
+The null regression model used to fit the realizations generated by the above model is formulated exactly the same as Equation \ref{eq:null-model}.
+
+For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the predictor $\boldsymbol{x}$, which will lead to the presence of heteroskedasticity.
+
+Since $\text{supp}(X) = [-1, 1]$, choosing $a$ to be $-1$, $0$ and $1$ can generate "left-triangle", "butterfly" and "right-triangle" shapes.
+The term $(2 - |a|)$ maintains the magnitude of residuals across different values of $a$.
+
+## Controlling the signal strength
+
+The three parameters $n$, $\sigma$ and $b$ are important for controlling the strength of the signal to generate lineups with a variety difficulty levels.
+This will ensure that  estimated power curves will be smooth and continuous, and that participants are allocated a set of lineups with a range of difficulty.
+
+Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ are used in data collection periods I and II respectively.
+A large value of $\sigma$ will increase the variation of the error of the non-linearity model and decrease the visibility of the visual pattern.
+The parameter $b$ controls the variation in the standard deviation of the error across the support of the predictor.
+A larger value of $b$ generates a larger ratio between it's smallest and highest values, making the visual pattern more obvious.
+The sample size $n$ sharpens (or blurs) a pattern when it is larger (or smaller).
+
+<!-- Figures \ref{fig:different-sigma}, \ref{fig:different-b} and \ref{fig:different-n} demonstrate the impact of these parameters on signal strength.  -->
+
+### Lineup allocation to participants
+
+From Table \ref{tab:model-factor-table} it can be computed that there are a total of $4 \times 4 \times 3 \times 4 = 192$ and $3 \times 5 \times 3 \times 4 = 180$ combinations of parameter values for the non-linearity model and heteroskedasticity models respectively.
+Three replications for each combination results in $192 \times 3 = 576$ and $180 \times 3 = 540$ lineups, respectively.
+
+Each lineup needs to be evaluated by at least five participants.
+From previous work, and additional pilot studies for this experiment, we decided to record evaluations from 20 lineups for each participant.
+Two of the 20 lineups with clear visual patterns were used as attention checks, to ensure quality data.
+Thus, $576 \times 5 / (20-2) = 160$ and $540 \times 5 / (20-2) = 150$ participants were needed to cover the experimental design for the data collection periods I and II, respectively.
+The factor levels and range of difficulty was assigned relatively equally among participants.
+The Appendix contains graphical summaries of these assignments for each subject.
+
+Data collection period III was primarily to obtain Rorschach lineup evaluations to estimate $\alpha$ (see Section A.1 in the Appendix), and also to obtain additional evaluations of lineups made with uniform fitted value distribution to ensure consistency in the results.
+To construct a Rorschach lineup, the data is generated from a model with zero effect size, while the null data are generated using the same simulation method discussed in Section \ref{visual-test-procedure-based-on-lineups}.
+This procedure differs from that of the canonical Rorschach lineup, where all 20 plots are generated directly from the null model, so the method suggested in @vanderplas2021statistical for typical lineups containing a data plot is used to estimate $\alpha$.
+(The Appendix contains a sensitivity analysis on the effect of uncertainty in the estimation of $\alpha$.) The $3 \times 4 = 12$ treatment levels of the common factors, replicated three times results in 36 lineups.
+And 6 more evaluations on the 279 lineups with uniform fitted value distribution, results in at least $(36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133$ participants needed.
+
+### Collecting results
+
+Participants for all three data collection periods were recruited from the Prolific crowd-sourcing platform [@palan2018prolific].
+Pre-screening procedures were applied during the recruitment: participants were required to be fluent in English, with $98\%$ minimum approval rate and at least 10 submissions in other studies.
+
+In each lineup evaluation, the participant was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots are from others.
+If there is no noticeable difference between plots in a lineup, participants had the option to select zero plots without the need to provide a reason.
+During the process of recording the responses, a zero selection was considered to be equivalent to selecting all 20 plots.
+No participant was shown the same lineup twice.
+Information about preferred pronouns, age group, education, and previous experience in visual experiments were also collected.
+A participant's submission was only included in the analysis if the data plot is identified for at least one attention check.
 

### 2024-01-01 paper.Rmd
reduce 2 pages

@@
 Additional lineups for both non-linearity and heteroskedasticity, using uniform fitted value distributions, were included for additional data, and to avoid participant frustration of too many difficult tasks.
+
+During the experiment, every participant recruited from the Prolific crowd-sourcing platform [@palan2018prolific] is presented with a block of 20 lineups.
+A lineup consists of a randomly placed data plot and 19 null plots, which are all residual plots drawn with raw residuals on the y-axis and fitted values on the x-axis.
+An additional horizontal red line is added at $y = 0$ as a visual reference.
+The data in the data plot is simulated from one of two models described in Section \ref{simulating-departures-from-good-residuals}, while the data of the remaining 19 null plots are generated by the residual rotation technique discussed in Section \ref{visual-test-procedure-based-on-lineups}.
+
+In each lineup evaluation, the participant was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots are from others.
+If there is no noticeable difference between plots in a lineup, participants had the option to select zero plots without the need to provide a reason.
+During the process of recording the responses, a zero selection was considered to be equivalent to selecting all 20 plots.
+No participant was shown the same lineup twice.
+Information about preferred pronouns, age group, education, and previous experience in visual experiments were also collected.
+A participant's submission was only included in the analysis if the data plot is identified for at least one attention check.
+
 Overall, we collected `r vi_survey  %>% filter(!attention_check) %>% nrow()` evaluations on `r vi_survey  %>% filter(!attention_check) %>% pull(unique_lineup_id) %>% unique() %>% length()` unique lineups performed by `r vi_survey  %>% filter(!attention_check) %>% count(exp, set) %>% nrow()` participants throughout three data collection periods.
@@
 Additionally, three levels of sample size ($n$) and four different fitted value distributions were incorporated.
-
-## Simulating departures from good residuals
-
-### Non-linearity
 
@@
 
+## Simulating departures from good residuals
+
+### Non-linearity and Heteroskedasticity
+
 Data collection period I was designed to study the ability of participants to detect the effect of a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials on random vector $\boldsymbol{x}$ formulated as
 
-\begin{align*} \label{eq:nonlinearity-model}
-\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
-\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
-\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
-\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
-\end{align*}
-
-\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials [@hermite1864nouveau; originally by @de1820theorie], $\varepsilon \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
-
-```{=tex}
-\begin{equation} \label{eq:scaling-function}
-g(\boldsymbol{x}, k) = 2k\cdot\frac{\boldsymbol{x} - x_{min}\boldsymbol{1}_n}{x_{max} - x_{min}} - k, \quad \text{for} \quad k > 0. 
+$$ \boldsymbol{y} = \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},~\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)$$
+
+\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{z}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $\boldsymbol{z} \propto He_j(\boldsymbol{x})$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials [@hermite1864nouveau; originally by @de1820theorie].
+
+To fit a misspecified model, the $\boldsymbol{z}$ term, which is a higher order term of $\boldsymbol{x}$, was omitted from the null regression formula
+
+\begin{equation} \label{eq:null-model}
+\boldsymbol{y} = \beta_0\boldsymbol{1}_n + \beta_1 \boldsymbol{x} + \boldsymbol{u},~ \boldsymbol{u} \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n).
 \end{equation}
-```
-
-\noindent where $\displaystyle x_{min} = \min_{i\in \{1, ..., n\}} x_i$, $\displaystyle x_{max} = \max_{i\in \{1, ..., n\}} x_i$ and $x_i$ is the $i$-th entry of $\boldsymbol{x}$.
-The function `hermite` from the R package `mpoly` [@mpoly] is used to  simulate $\boldsymbol{z}_{raw}$ to generate Hermite polynomials.
-
-The null regression model used to fit the realizations generated by the above model is formulated as
-
-```{=tex}
-\begin{equation} \label{eq:null-model}
-\boldsymbol{y} = \beta_0\boldsymbol{1}_n + \beta_1 \boldsymbol{x} + \boldsymbol{u},
-\end{equation}
-```
-\noindent where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
-Here $\boldsymbol{z}$ is a higher order term of $\boldsymbol{x}$ and leaving it out in the null regression results in model misspecification.
+
+Four different values of $j = 2, 3, 6, 18$ were chosen so that distinct shapes of non-linearity were included in the residual plot.
+These include "U", "S", "M" and "triple-U" shape as shown in Figure \ref{fig:different-shape-of-herimite}.
+A greater value of $j$ will result in a curve with more turning points.
+It is expected that the "U" shape will be the easiest to detect, and as the shape gets more complex it will be harder to perceive in a scatterplot, particularly when there is noise.
+Figure \ref{fig:different-sigma} shows the "U" shape for different amounts of noise ($\sigma$).
+
+Data collection period II was similar to Data collection I but focuses on heteroskedasticity visual patterns formulated as
+
+$$\boldsymbol{y} = \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{\varepsilon},~\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{1}_n + (2 - |a|)(\boldsymbol{x} - a\boldsymbol{1}_n)'(\boldsymbol{x} - a\boldsymbol{1}_n) b \boldsymbol{I}_n),$$
+\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vectors of size $n$.
+
+The null regression model used to fit the realizations generated by the above model is formulated exactly the same as Equation \ref{eq:null-model}.
+For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the predictor $\boldsymbol{x}$, which will lead to the presence of heteroskedasticity.
+
+Visual patterns of heteroskedasticity are simulated using three different shapes ($a$ = -1, 0, 1) including "left-triangle", "butterfly" and "right-triangle" shapes as displayed in Figure \ref{fig:different-shape-of-heter}.
+Figure \ref{fig:different-b} shows the butterfly shape as the ratio parameter ($b$) is changed.
+
+More details about the design of the data generating process are provided in the Appendix.
 
@@
 
-Visual patterns of non-linearity are simulated using four different orders of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$).
-The values of $j$ were chosen so that distinct shapes of non-linearity were included in the residual plot.
-These include "U", "S", "M" and "triple-U" shape as shown in Figure \ref{fig:different-shape-of-herimite}.
-A greater value of $j$ will result in a curve with more turning points.
-It is expected that the "U" shape will be the easiest to detect, and as the shape gets more complex it will be harder to perceive in a scatterplot, particularly when there is noise.
-Figure \ref{fig:different-sigma} shows the "U" shape for different amounts of noise ($\sigma$).
-
-### Heteroskedasticity
-
-Data collection period II was designed to study the ability of participants to detect the heteroskedasticity under a simple linear regression model setting:
-
-\begin{align*} \label{eq:heter-model}
-\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
-\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
-\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}_n, \boldsymbol{1}_n + (2 - |a|)(\boldsymbol{x} - a\boldsymbol{1}_n)'(\boldsymbol{x} - a\boldsymbol{1}_n) b \boldsymbol{I}_n), 
-\end{align*}
-
-\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vectors of size $n$ and $g(.)$ is the scaling function defined in Equation \ref{eq:scaling-function}.
-The null regression model used to fit the realizations generated by the above model is formulated exactly the same as Equation \ref{eq:null-model}.
-
-For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the predictor $\boldsymbol{x}$, which will lead to the presence of heteroskedasticity.
-Visual patterns of heteroskedasticity are simulated using three different shapes ($a$ = -1, 0, 1).
-
-Since $\text{supp}(X) = [-1, 1]$, choosing $a$ to be $-1$, $0$ and $1$ can generate "left-triangle", "butterfly" and "right-triangle" shapes as displayed in Figure \ref{fig:different-shape-of-heter}.
-The term $(2 - |a|)$ maintains the magnitude of residuals across different values of $a$.
-Figure \ref{fig:different-b} shows the butterfly shape as the ratio parameter ($b$) is changed.
-
 ```{r different-shape-of-heter, fig.pos="!h", fig.height = 2.67, fig.cap = 'Heteroskedasticity forms used in the experiment.
@@
 ```{r different-b, fig.pos="!h", fig.height = 2, fig.cap = "Five different values of $b$ are used in heteroskedasticity simulation to control the strength of the signal.
-Larger values of $b$ yield a bigger difference in variation, and stus stronger heteroskedasticity signal."}
+Larger values of $b$ yield a bigger difference in variation, and thus stronger heteroskedasticity signal."}
 
@@
 This latter problem is usually corrected before modelling, using a variable transformation.
-Our simulation assess this by using four different distributions to represent fitted values, constructed by different sampling of the raw predictor $X_{raw}$:
-
-- uniform, $U(-1, 1)$, 
-- normal, $N(0, 0.3^2)$, 
-- skewed, $\text{lognormal}(0, 0.6^2)/3$, and 
-- discrete, $U\{1, 5\}$.
-
-\noindent Figure \ref{fig:different-dist} shows the non-linear pattern, a "U" shape, with the different fitted value distributions.
+Our simulation assess this by using four different distributions to represent fitted values, constructed by different sampling of the raw predictor $X_{raw}$, namely, uniform ($U(-1, 1)$), normal ($N(0, 0.3^2)$), skewed ($\text{lognormal}(0, 0.6^2)/3$) and discrete ($U\{1, 5\}$).
+
+Figure \ref{fig:different-dist} shows the non-linear pattern, a "U" shape, with the different fitted value distributions.
 We would expect that structure in residual plots would be easier to perceive when the fitted values are uniformly distributed.
@@
 
-### Controlling the signal strength
-
-The three parameters $n$, $\sigma$ and $b$ are important for controlling the strength of the signal to generate lineups with a variety difficulty levels.
-This will ensure that  estimated power curves will be smooth and continuous, and that participants are allocated a set of lineups with a range of difficulty.
-
-Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ are used in data collection periods I and II respectively.
-A large value of $\sigma$ will increase the variation of the error of the non-linearity model and decrease the visibility of the visual pattern.
-The parameter $b$ controls the variation in the standard deviation of the error across the support of the predictor.
-A larger value of $b$ generates a larger ratio between it's smallest and highest values, making the visual pattern more obvious.
-The sample size $n$ sharpens (or blurs) a pattern when it is larger (or smaller).
-Figures \ref{fig:different-sigma}, \ref{fig:different-b} and \ref{fig:different-n} demonstrate the impact of these parameters on signal strength.
-
-## Experimental setup
+## Effect size
 
@@
 
-### Effect size
-
 Effect size in statistics measures the strength of the signal relative to the noise. It is surprisingly difficult to quantify, even for simulated data as used in this experiment.
@@
 As sample size increases the effect size would increase, but as variance increases the effect size decreases.
-However, it is not clear how the additional parameter for the model polynomial order, $k$, should be incorporated.
-Intuitively, the large $k$ means more complex pattern, which likely means effect size would decrease.
+However, it is not clear how the additional parameter for the model polynomial order, $j$, should be incorporated.
+Intuitively, the large $j$ means more complex pattern, which likely means effect size would decrease.
 For the purposes of our calculations we have chosen to use an approach based on Kullback-Leibler divergence [@kullback1951information], coupled with simulation.
@@
 This ensures lineups constructed with the same experimental factors will share the same effect size.
-
-### Lineup allocation to participants
-
-From Table \ref{tab:model-factor-table} it can be computed that there are a total of $4 \times 4 \times 3 \times 4 = 192$ and $3 \times 5 \times 3 \times 4 = 180$ combinations of parameter values for the non-linearity model and heteroskedasticity models respectively.
-Three replications for each combination results in $192 \times 3 = 576$ and $180 \times 3 = 540$ lineups, respectively.
-
-Each lineup needs to be evaluated by at least five participants.
-From previous work, and additional pilot studies for this experiment, we decided to record evaluations from 20 lineups for each participant.
-Two of the 20 lineups with clear visual patterns were used as attention checks, to ensure quality data.
-Thus, $576 \times 5 / (20-2) = 160$ and $540 \times 5 / (20-2) = 150$ participants were needed to cover the experimental design for the data collection periods I and II, respectively.
-The factor levels and range of difficulty was assigned relatively equally among participants.
-The Appendix contains graphical summaries of these assignments for each subject.
-
-Data collection period III was primarily to obtain Rorschach lineup evaluations to estimate $\alpha$ (see Section A.1 in the Appendix), and also to obtain additional evaluations of lineups made with uniform fitted value distribution to ensure consistency in the results.
-To construct a Rorschach lineup, the data is generated from a model with zero effect size, while the null data are generated using the same simulation method discussed in Section \ref{visual-test-procedure-based-on-lineups}.
-This procedure differs from that of the canonical Rorschach lineup, where all 20 plots are generated directly from the null model, so the method suggested in @vanderplas2021statistical for typical lineups containing a data plot is used to estimate $\alpha$.
-(The Appendix contains a sensitivity analysis on the effect of uncertainty in the estimation of $\alpha$.) The $3 \times 4 = 12$ treatment levels of the common factors, replicated three times results in 36 lineups.
-And 6 more evaluations on the 279 lineups with uniform fitted value distribution, results in at least $(36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133$ participants needed.
-
-### Collecting results
-
-Participants for all three data collection periods were recruited from the Prolific crowd-sourcing platform [@palan2018prolific].
-Pre-screening procedures were applied during the recruitment: participants were required to be fluent in English, with $98\%$ minimum approval rate and at least 10 submissions in other studies.
-
-During the experiment, every participant is presented with a block of 20 lineups.
-A lineup consists of a randomly placed data plot and 19 null plots, which are all residual plots drawn with raw residuals on the y-axis and fitted values on the x-axis.
-An additional horizontal red line is added at $y = 0$ as a visual reference.
-The data in the data plot is simulated from one of two models described in Section \ref{simulating-departures-from-good-residuals}, while the data of the remaining 19 null plots are generated by the residual rotation technique discussed in Section \ref{visual-test-procedure-based-on-lineups}.
-
-In each lineup evaluation, the participant was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots are from others.
-If there is no noticeable difference between plots in a lineup, participants had the option to select zero plots without the need to provide a reason.
-During the process of recording the responses, a zero selection was considered to be equivalent to selecting all 20 plots.
-No participant was shown the same lineup twice.
-Information about preferred pronouns, age group, education, and previous experience in visual experiments were also collected.
-A participant's submission was only included in the analysis if the data plot is identified for at least one attention check.
 

### 2024-01-01 paper.tex
reduce 2 pages

@@
 fitted value distributions, were included for additional data, and to
-avoid participant frustration of too many difficult tasks. Overall, we
-collected 7974 evaluations on 1152 unique lineups performed by 443
-participants throughout three data collection periods.
+avoid participant frustration of too many difficult tasks.
+
+During the experiment, every participant recruited from the Prolific
+crowd-sourcing platform \citep{palan2018prolific} is presented with a
+block of 20 lineups. A lineup consists of a randomly placed data plot
+and 19 null plots, which are all residual plots drawn with raw residuals
+on the y-axis and fitted values on the x-axis. An additional horizontal
+red line is added at \(y = 0\) as a visual reference. The data in the
+data plot is simulated from one of two models described in Section
+\ref{simulating-departures-from-good-residuals}, while the data of the
+remaining 19 null plots are generated by the residual rotation technique
+discussed in Section \ref{visual-test-procedure-based-on-lineups}.
+
+In each lineup evaluation, the participant was asked to select one or
+more plots that are most different from others, provide a reason for
+their selections, and evaluate how different they think the selected
+plots are from others. If there is no noticeable difference between
+plots in a lineup, participants had the option to select zero plots
+without the need to provide a reason. During the process of recording
+the responses, a zero selection was considered to be equivalent to
+selecting all 20 plots. No participant was shown the same lineup twice.
+Information about preferred pronouns, age group, education, and previous
+experience in visual experiments were also collected. A participant's
+submission was only included in the analysis if the data plot is
+identified for at least one attention check.
+
+Overall, we collected 7974 evaluations on 1152 unique lineups performed
+by 443 participants throughout three data collection periods. Details
+about lineup allocation and data collection are provided in the
+Appendix.
 
@@
 and four different fitted value distributions were incorporated.
-
-\hypertarget{simulating-departures-from-good-residuals}{%
-\subsection{Simulating departures from good
-residuals}\label{simulating-departures-from-good-residuals}}
-
-\hypertarget{non-linearity}{%
-\subsubsection{Non-linearity}\label{non-linearity}}
 
@@
 
+\hypertarget{simulating-departures-from-good-residuals}{%
+\subsection{Simulating departures from good
+residuals}\label{simulating-departures-from-good-residuals}}
+
+\hypertarget{non-linearity-and-heteroskedasticity}{%
+\subsubsection{Non-linearity and
+Heteroskedasticity}\label{non-linearity-and-heteroskedasticity}}
+
 Data collection period I was designed to study the ability of
@@
 
-\begin{align*} \label{eq:nonlinearity-model}
-\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
-\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
-\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
-\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
-\end{align*}
+\[ \boldsymbol{y} = \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},~\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)\]
 
 \noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
-\(\boldsymbol{\varepsilon}\), \(\boldsymbol{x}_{raw}\),
-\(\boldsymbol{z}_{raw}\) are vectors of size \(n\), \(\boldsymbol{1}_n\)
-is a vector of ones of size \(n\), \(He_{j}(.)\) is the \(j\)th-order
-probabilist's Hermite polynomials
+\(\boldsymbol{\varepsilon}\), \(\boldsymbol{z}\) are vectors of size
+\(n\), \(\boldsymbol{1}_n\) is a vector of ones of size \(n\),
+\(\boldsymbol{z} \propto He_j(\boldsymbol{x})\), \(He_{j}(.)\) is the
+\(j\)th-order probabilist's Hermite polynomials
 \citetext{\citealp{hermite1864nouveau}; \citealp[originally
-by][]{de1820theorie}},
-\(\varepsilon \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)\), and
-\(g(\boldsymbol{x}, k)\) is a scaling function to enforce the support of
-the random vector to be \([-k, k]^n\) defined as
-
-\begin{equation} \label{eq:scaling-function}
-g(\boldsymbol{x}, k) = 2k\cdot\frac{\boldsymbol{x} - x_{min}\boldsymbol{1}_n}{x_{max} - x_{min}} - k, \quad \text{for} \quad k > 0. 
+by][]{de1820theorie}}.
+
+To fit a misspecified model, the \(\boldsymbol{z}\) term, which is a
+higher order term of \(\boldsymbol{x}\), was omitted from the null
+regression formula
+
+\begin{equation} \label{eq:null-model}
+\boldsymbol{y} = \beta_0\boldsymbol{1}_n + \beta_1 \boldsymbol{x} + \boldsymbol{u}, \boldsymbol{u} \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n).
 \end{equation}
 
-\noindent where
-\(\displaystyle x_{min} = \min_{i\in \{1, ..., n\}} x_i\),
-\(\displaystyle x_{max} = \max_{i\in \{1, ..., n\}} x_i\) and \(x_i\) is
-the \(i\)-th entry of \(\boldsymbol{x}\). The function \texttt{hermite}
-from the R package \texttt{mpoly} \citep{mpoly} is used to simulate
-\(\boldsymbol{z}_{raw}\) to generate Hermite polynomials.
-
-The null regression model used to fit the realizations generated by the
-above model is formulated as
-
-\begin{equation} \label{eq:null-model}
-\boldsymbol{y} = \beta_0\boldsymbol{1}_n + \beta_1 \boldsymbol{x} + \boldsymbol{u},
-\end{equation}
-
-\noindent where
-\(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\).
-Here \(\boldsymbol{z}\) is a higher order term of \(\boldsymbol{x}\) and
-leaving it out in the null regression results in model misspecification.
-
-\begin{figure}[!h]
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-shape-of-herimite-1}
-
-}
-
-\caption{Polynomial forms generated for the residual plots used to assess detecting non-linearity.
-The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$.}\label{fig:different-shape-of-herimite}
-\end{figure}
-
-\begin{figure}[!h]
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-sigma-1}
-
-}
-
-\caption{Examining the effect of $\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform fitted value distribution and the "U" shape.
-As $\sigma$ increases the signal strength decreases, to the point that the "U" is almost unrecognisable when $\sigma=4$.}\label{fig:different-sigma}
-\end{figure}
-
-Visual patterns of non-linearity are simulated using four different
-orders of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)). The
-values of \(j\) were chosen so that distinct shapes of non-linearity
-were included in the residual plot. These include ``U'', ``S'', ``M''
-and ``triple-U'' shape as shown in Figure
-\ref{fig:different-shape-of-herimite}. A greater value of \(j\) will
-result in a curve with more turning points. It is expected that the
+Four different values of \(j = 2, 3, 6, 18\) were chosen so that
+distinct shapes of non-linearity were included in the residual plot.
+These include ``U'', ``S'', ``M'' and ``triple-U'' shape as shown in
+Figure \ref{fig:different-shape-of-herimite}. A greater value of \(j\)
+will result in a curve with more turning points. It is expected that the
 ``U'' shape will be the easiest to detect, and as the shape gets more
@@
 
-\hypertarget{heteroskedasticity}{%
-\subsubsection{Heteroskedasticity}\label{heteroskedasticity}}
-
-Data collection period II was designed to study the ability of
-participants to detect the heteroskedasticity under a simple linear
-regression model setting:
-
-\begin{align*} \label{eq:heter-model}
-\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
-\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
-\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}_n, \boldsymbol{1}_n + (2 - |a|)(\boldsymbol{x} - a\boldsymbol{1}_n)'(\boldsymbol{x} - a\boldsymbol{1}_n) b \boldsymbol{I}_n), 
-\end{align*}
-
+Data collection period II was similar to Data collection I but focuses
+on heteroskedasticity visual patterns formulated as
+
+\[\boldsymbol{y} = \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{\varepsilon},~\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{1}_n + (2 - |a|)(\boldsymbol{x} - a\boldsymbol{1}_n)'(\boldsymbol{x} - a\boldsymbol{1}_n) b \boldsymbol{I}_n),\]
 \noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
-\(\boldsymbol{\varepsilon}\) are vectors of size \(n\) and \(g(.)\) is
-the scaling function defined in Equation \ref{eq:scaling-function}. The
-null regression model used to fit the realizations generated by the
+\(\boldsymbol{\varepsilon}\) are vectors of size \(n\).
+
+The null regression model used to fit the realizations generated by the
 above model is formulated exactly the same as Equation
-\ref{eq:null-model}.
-
-For \(b \neq 0\), the variance-covariance matrix of the error term
-\(\boldsymbol{\varepsilon}\) is correlated with the predictor
-\(\boldsymbol{x}\), which will lead to the presence of
-heteroskedasticity. Visual patterns of heteroskedasticity are simulated
-using three different shapes (\(a\) = -1, 0, 1).
-
-Since \(\text{supp}(X) = [-1, 1]\), choosing \(a\) to be \(-1\), \(0\)
-and \(1\) can generate ``left-triangle'', ``butterfly'' and
-``right-triangle'' shapes as displayed in Figure
-\ref{fig:different-shape-of-heter}. The term \((2 - |a|)\) maintains the
-magnitude of residuals across different values of \(a\). Figure
-\ref{fig:different-b} shows the butterfly shape as the ratio parameter
-(\(b\)) is changed.
+\ref{eq:null-model}. For \(b \neq 0\), the variance-covariance matrix of
+the error term \(\boldsymbol{\varepsilon}\) is correlated with the
+predictor \(\boldsymbol{x}\), which will lead to the presence of
+heteroskedasticity.
+
+Visual patterns of heteroskedasticity are simulated using three
+different shapes (\(a\) = -1, 0, 1) including ``left-triangle'',
+``butterfly'' and ``right-triangle'' shapes as displayed in Figure
+\ref{fig:different-shape-of-heter}. Figure \ref{fig:different-b} shows
+the butterfly shape as the ratio parameter (\(b\)) is changed.
+
+More details about the design of the data generating process are
+provided in the Appendix.
+
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-shape-of-herimite-1}
+
+}
+
+\caption{Polynomial forms generated for the residual plots used to assess detecting non-linearity.
+The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$.}\label{fig:different-shape-of-herimite}
+\end{figure}
+
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-sigma-1}
+
+}
+
+\caption{Examining the effect of $\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform fitted value distribution and the "U" shape.
+As $\sigma$ increases the signal strength decreases, to the point that the "U" is almost unrecognisable when $\sigma=4$.}\label{fig:different-sigma}
+\end{figure}
 
@@
 \caption{Five different values of $b$ are used in heteroskedasticity simulation to control the strength of the signal.
-Larger values of $b$ yield a bigger difference in variation, and stus stronger heteroskedasticity signal.}\label{fig:different-b}
+Larger values of $b$ yield a bigger difference in variation, and thus stronger heteroskedasticity signal.}\label{fig:different-b}
 \end{figure}
@@
 represent fitted values, constructed by different sampling of the raw
-predictor \(X_{raw}\):
-
-\begin{itemize}
-\tightlist
-\item
-  uniform, \(U(-1, 1)\),
-\item
-  normal, \(N(0, 0.3^2)\),
-\item
-  skewed, \(\text{lognormal}(0, 0.6^2)/3\), and
-\item
-  discrete, \(U\{1, 5\}\).
-\end{itemize}
-
-\noindent Figure \ref{fig:different-dist} shows the non-linear pattern,
-a ``U'' shape, with the different fitted value distributions. We would
-expect that structure in residual plots would be easier to perceive when
-the fitted values are uniformly distributed.
+predictor \(X_{raw}\), namely, uniform (\(U(-1, 1)\)), normal
+(\(N(0, 0.3^2)\)), skewed (\(\text{lognormal}(0, 0.6^2)/3\)) and
+discrete (\(U\{1, 5\}\)).
+
+Figure \ref{fig:different-dist} shows the non-linear pattern, a ``U''
+shape, with the different fitted value distributions. We would expect
+that structure in residual plots would be easier to perceive when the
+fitted values are uniformly distributed.
 
@@
 
-\hypertarget{controlling-the-signal-strength}{%
-\subsubsection{Controlling the signal
-strength}\label{controlling-the-signal-strength}}
-
-The three parameters \(n\), \(\sigma\) and \(b\) are important for
-controlling the strength of the signal to generate lineups with a
-variety difficulty levels. This will ensure that estimated power curves
-will be smooth and continuous, and that participants are allocated a set
-of lineups with a range of difficulty.
-
-Parameter \(\sigma \in \{0.5, 1, 2, 4\}\) and
-\(b \in \{0.25, 1, 4, 16, 64\}\) are used in data collection periods I
-and II respectively. A large value of \(\sigma\) will increase the
-variation of the error of the non-linearity model and decrease the
-visibility of the visual pattern. The parameter \(b\) controls the
-variation in the standard deviation of the error across the support of
-the predictor. A larger value of \(b\) generates a larger ratio between
-it's smallest and highest values, making the visual pattern more
-obvious. The sample size \(n\) sharpens (or blurs) a pattern when it is
-larger (or smaller). Figures \ref{fig:different-sigma},
-\ref{fig:different-b} and \ref{fig:different-n} demonstrate the impact
-of these parameters on signal strength.
-
-\hypertarget{experimental-setup}{%
-\subsection{Experimental setup}\label{experimental-setup}}
+\hypertarget{effect-size}{%
+\subsection{Effect size}\label{effect-size}}
 
@@
 set of lineups with a range of difficulties to each participant.
-
-\hypertarget{effect-size}{%
-\subsubsection{Effect size}\label{effect-size}}
 
@@
 increases the effect size decreases. However, it is not clear how the
-additional parameter for the model polynomial order, \(k\), should be
-incorporated. Intuitively, the large \(k\) means more complex pattern,
+additional parameter for the model polynomial order, \(j\), should be
+incorporated. Intuitively, the large \(j\) means more complex pattern,
 which likely means effect size would decrease. For the purposes of our
@@
 share the same effect size.
-
-\hypertarget{lineup-allocation-to-participants}{%
-\subsubsection{Lineup allocation to
-participants}\label{lineup-allocation-to-participants}}
-
-From Table \ref{tab:model-factor-table} it can be computed that there
-are a total of \(4 \times 4 \times 3 \times 4 = 192\) and
-\(3 \times 5 \times 3 \times 4 = 180\) combinations of parameter values
-for the non-linearity model and heteroskedasticity models respectively.
-Three replications for each combination results in
-\(192 \times 3 = 576\) and \(180 \times 3 = 540\) lineups, respectively.
-
-Each lineup needs to be evaluated by at least five participants. From
-previous work, and additional pilot studies for this experiment, we
-decided to record evaluations from 20 lineups for each participant. Two
-of the 20 lineups with clear visual patterns were used as attention
-checks, to ensure quality data. Thus, \(576 \times 5 / (20-2) = 160\)
-and \(540 \times 5 / (20-2) = 150\) participants were needed to cover
-the experimental design for the data collection periods I and II,
-respectively. The factor levels and range of difficulty was assigned
-relatively equally among participants. The Appendix contains graphical
-summaries of these assignments for each subject.
-
-Data collection period III was primarily to obtain Rorschach lineup
-evaluations to estimate \(\alpha\) (see Section A.1 in the Appendix),
-and also to obtain additional evaluations of lineups made with uniform
-fitted value distribution to ensure consistency in the results. To
-construct a Rorschach lineup, the data is generated from a model with
-zero effect size, while the null data are generated using the same
-simulation method discussed in Section
-\ref{visual-test-procedure-based-on-lineups}. This procedure differs
-from that of the canonical Rorschach lineup, where all 20 plots are
-generated directly from the null model, so the method suggested in
-\citet{vanderplas2021statistical} for typical lineups containing a data
-plot is used to estimate \(\alpha\). (The Appendix contains a
-sensitivity analysis on the effect of uncertainty in the estimation of
-\(\alpha\).) The \(3 \times 4 = 12\) treatment levels of the common
-factors, replicated three times results in 36 lineups. And 6 more
-evaluations on the 279 lineups with uniform fitted value distribution,
-results in at least
-\((36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133\) participants
-needed.
-
-\hypertarget{collecting-results}{%
-\subsubsection{Collecting results}\label{collecting-results}}
-
-Participants for all three data collection periods were recruited from
-the Prolific crowd-sourcing platform \citep{palan2018prolific}.
-Pre-screening procedures were applied during the recruitment:
-participants were required to be fluent in English, with \(98\%\)
-minimum approval rate and at least 10 submissions in other studies.
-
-During the experiment, every participant is presented with a block of 20
-lineups. A lineup consists of a randomly placed data plot and 19 null
-plots, which are all residual plots drawn with raw residuals on the
-y-axis and fitted values on the x-axis. An additional horizontal red
-line is added at \(y = 0\) as a visual reference. The data in the data
-plot is simulated from one of two models described in Section
-\ref{simulating-departures-from-good-residuals}, while the data of the
-remaining 19 null plots are generated by the residual rotation technique
-discussed in Section \ref{visual-test-procedure-based-on-lineups}.
-
-In each lineup evaluation, the participant was asked to select one or
-more plots that are most different from others, provide a reason for
-their selections, and evaluate how different they think the selected
-plots are from others. If there is no noticeable difference between
-plots in a lineup, participants had the option to select zero plots
-without the need to provide a reason. During the process of recording
-the responses, a zero selection was considered to be equivalent to
-selecting all 20 plots. No participant was shown the same lineup twice.
-Information about preferred pronouns, age group, education, and previous
-experience in visual experiments were also collected. A participant's
-submission was only included in the analysis if the data plot is
-identified for at least one attention check.
 

### 2024-01-07 paper.Rmd
reduce to 32 + 3 references (almost there)

@@
 
-\textcolor{red}{Diagnosing a model is an important part of building an appropriate model.} In linear regression analysis, studying the residuals from a model fit is a common diagnostic activity.
+Diagnosing a model is an important part of building an appropriate model.
+In linear regression analysis, studying the residuals from a model fit is a common diagnostic activity.
 Residuals summarise what is not captured by the model, and thus provide the capacity to identify what might be wrong.
@@
 
-\textcolor{red}{Scatterplots of residuals against the fitted values, and each of the explanatory variables, are commonly used to scrutinize their relationships}.
+Scatterplots of residuals against the fitted values, and each of the explanatory variables, are commonly used to scrutinize their relationships.
 If there are any visually discoverable associations, the model is potentially inadequate or incorrectly specified.
-\textcolor{red}{We can also potentially discover patterns not directly connected to a linear model assumption from these residual plots, such as the discreteness or skewness of the fitted values, and outliers.} To read residual plots, one looks for noticeable departures from the model such as non-linear pattern or heteroskedasticity.
+We can also potentially discover patterns not directly connected to a linear model assumption from these residual plots, such as the discreteness or skewness of the fitted values, and outliers.
+To read residual plots, one looks for noticeable departures from the model such as non-linear pattern or heteroskedasticity.
 A non-linear pattern would suggest that the model needs to have some additional non-linear terms.
@@
 
-\textcolor{red}{Additionally, relying solely on the subjective assessment of a single plot can be problematic.
+Additionally, relying solely on the subjective assessment of a single plot can be problematic.
 People will almost always see a pattern \citep[see][]{kahneman2011thinking}, so the question that really needs answering is whether any pattern perceived is consistent with randomness, or sampling variability, or noise.
 Correctly judging whether \emph{no} pattern exists in a residual plot is a difficult task.
-\cite{loy2021bringing} emphasizes that this is especially difficult to teach to new analysts and students, and advocates to the broader use of the lineup protocol \citep{bujastatistical2009}.}
-
-\textcolor{red}{The lineup protocol places a data plot in a field of null plots, allowing for a comparison of patterns due purely by chance to what is perceived in the data plot.
+\cite{loy2021bringing} emphasizes that this is especially difficult to teach to new analysts and students, and advocates to the broader use of the lineup protocol \citep{bujastatistical2009}.
+
+The lineup protocol places a data plot in a field of null plots, allowing for a comparison of patterns due purely by chance to what is perceived in the data plot.
 For residual analysis this is especially helpful for gauging whether there is \emph{no} pattern.
@@
 However, where no adequate conventional test exists, it is invaluable, as shown by \cite{loy2013diagnostic}.
-Here we use the lineup as a vehicle to rigorously explore why experts advise that residual plots are indispensable despite the prevalence of numerical tests.}
+Here we use the lineup as a vehicle to rigorously explore why experts advise that residual plots are indispensable despite the prevalence of numerical tests.
 
@@
 
+\vspace{-\baselineskip}
+
 ```{=tex}
@@
 
-During the experiment, every participant recruited from the Prolific crowd-sourcing platform [@palan2018prolific] is presented with a block of 20 lineups.
-A lineup consists of a randomly placed data plot and 19 null plots, which are all residual plots drawn with raw residuals on the y-axis and fitted values on the x-axis.
-An additional horizontal red line is added at $y = 0$ as a visual reference.
-The data in the data plot is simulated from one of two models described in Section \ref{simulating-departures-from-good-residuals}, while the data of the remaining 19 null plots are generated by the residual rotation technique discussed in Section \ref{visual-test-procedure-based-on-lineups}.
+During the experiment, every participant recruited from the Prolific crowd-sourcing platform [@palan2018prolific] was presented with a block of 20 lineups.
+A lineup consisted of a randomly placed data plot and 19 null plots, which were all residual plots drawn with raw residuals on the y-axis and fitted values on the x-axis.
+An additional horizontal red line was added at $y = 0$ as a visual reference.
+The data in the data plot was simulated from one of two models described in Section \ref{simulating-departures-from-good-residuals}, while the data of the remaining 19 null plots were generated by the residual rotation technique discussed in Section \ref{visual-test-procedure-based-on-lineups}.
 
@@
 
-Overall, we collected `r vi_survey  %>% filter(!attention_check) %>% nrow()` evaluations on `r vi_survey  %>% filter(!attention_check) %>% pull(unique_lineup_id) %>% unique() %>% length()` unique lineups performed by `r vi_survey  %>% filter(!attention_check) %>% count(exp, set) %>% nrow()` participants throughout three data collection periods.
-
+Overall, we collected `r vi_survey  %>% filter(!attention_check) %>% nrow()` evaluations on `r vi_survey  %>% filter(!attention_check) %>% pull(unique_lineup_id) %>% unique() %>% length()` unique lineups performed by `r vi_survey  %>% filter(!attention_check) %>% count(exp, set) %>% nrow()` participants.
 A summary of the factors used in the experiment can be found in Table \ref{tab:model-factor-table}.
@@
 
-Data collection period I was designed to study the ability of participants to detect the effect of a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials on random vector $\boldsymbol{x}$ formulated as
-
-$$ \boldsymbol{y} = \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},~\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)$$
-
-\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{z}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $\boldsymbol{z} \propto He_j(\boldsymbol{x})$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials [@hermite1864nouveau; originally by @de1820theorie].
-
-To fit a misspecified model, the $\boldsymbol{z}$ term, which is a higher order term of $\boldsymbol{x}$, was omitted from the null regression formula
-
-\begin{equation} \label{eq:null-model}
-\boldsymbol{y} = \beta_0\boldsymbol{1}_n + \beta_1 \boldsymbol{x} + \boldsymbol{u},~ \boldsymbol{u} \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n).
-\end{equation}
-
-Four different values of $j = 2, 3, 6, 18$ were chosen so that distinct shapes of non-linearity were included in the residual plot.
+Data collection period I was designed to study the ability of participants to detect non-linearity departures from residual plots.
+The non-linearity departure was constructed by omitting a $j$th order Hermite polynomial [@hermite1864nouveau; originally by @de1820theorie] term of the predictor from the simple linear regression equation.
+Four different values of $j = 2, 3, 6, 18$ were chosen so that distinct shapes of non-linearity were included in the residual plots.
 These include "U", "S", "M" and "triple-U" shape as shown in Figure \ref{fig:different-shape-of-herimite}.
@@
 
-Data collection period II was similar to Data collection I but focuses on heteroskedasticity visual patterns formulated as
-
-$$\boldsymbol{y} = \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{\varepsilon},~\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{1}_n + (2 - |a|)(\boldsymbol{x} - a\boldsymbol{1}_n)'(\boldsymbol{x} - a\boldsymbol{1}_n) b \boldsymbol{I}_n),$$
-\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vectors of size $n$.
-
-The null regression model used to fit the realizations generated by the above model is formulated exactly the same as Equation \ref{eq:null-model}.
-For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the predictor $\boldsymbol{x}$, which will lead to the presence of heteroskedasticity.
-
+Data collection period II was similar to period I but focuses on heteroskedasticity departures.
+We generated the heteroskedasticity departures by setting the variance-covariance matrix of the error term as a function of the predictor, but fitted the data with the simple linear regression model, intentionally violated the constant variance assumption.
 Visual patterns of heteroskedasticity are simulated using three different shapes ($a$ = -1, 0, 1) including "left-triangle", "butterfly" and "right-triangle" shapes as displayed in Figure \ref{fig:different-shape-of-heter}.
 Figure \ref{fig:different-b} shows the butterfly shape as the ratio parameter ($b$) is changed.
-
-More details about the design of the data generating process are provided in the Appendix.
+More details about the simulation process are provided in the Appendix.
 
@@
 This latter problem is usually corrected before modelling, using a variable transformation.
-Our simulation assess this by using four different distributions to represent fitted values, constructed by different sampling of the raw predictor $X_{raw}$, namely, uniform ($U(-1, 1)$), normal ($N(0, 0.3^2)$), skewed ($\text{lognormal}(0, 0.6^2)/3$) and discrete ($U\{1, 5\}$).
+Our simulation assess this by using four different distributions to represent fitted values, constructed by different sampling of the predictor, including $U(-1, 1)$ (uniform), $N(0, 0.3^2)$ (normal), $\text{lognormal}(0, 0.6^2)/3$ (skewed) and $U\{-1, 1\}$ (discrete).
 
@@
 We chose 300 as the upper limit, because it is typically enough for structure to be visible in a scatterplot reliably.
-\textcolor{red}{Beyond 300, the scatterplot should probably be used with transparency or replaced with a density or binned plot as scatterplots suffer from over-plotting.}
+Beyond 300, the scatterplot should probably be used with transparency or replaced with a density or binned plot as scatterplots suffer from over-plotting.
 
@@
 
-For the non-linearity model, the key items defining effect size are sample size ($n$) and variance of the error term ($\sigma^2$), and so effect size would be roughly calculated as $\sqrt{n}/{\sigma}$.
-As sample size increases the effect size would increase, but as variance increases the effect size decreases.
+For the non-linearity model, the key items defining effect size are sample size ($n$) and the noise level ($\sigma^2$), and so effect size would be roughly calculated as $\sqrt{n}/{\sigma}$.
+Increasing sample size tends to boost the effect size, while heightened noise diminishes it.
 However, it is not clear how the additional parameter for the model polynomial order, $j$, should be incorporated.
 Intuitively, the large $j$ means more complex pattern, which likely means effect size would decrease.
+Similarly, in the heteroskedasticity model, effect size relies on sample size ($n$) and the ratio of the largest to smallest variance, $b$.
+Larger values of both would produce higher effect size, but the role of the additional shape parameter, $a$, in this context is unclear.
+
 For the purposes of our calculations we have chosen to use an approach based on Kullback-Leibler divergence [@kullback1951information], coupled with simulation.
-This formulation defines effect size to be:
-
-\begin{equation}
-E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(\text{diag}(\boldsymbol{R}\sigma^2))^{-1}\boldsymbol{\mu}_z\right)
-\label{eq:nonlin-effect-size}
-\end{equation}
-
-\noindent where $\text{diag}(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,
-$\boldsymbol{R} = \boldsymbol{I}_n - \boldsymbol{H}$ is the residual operator, $\boldsymbol{H} = \boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the hat matrix, $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals where $\boldsymbol{Z}$ contain any higher order terms of $\boldsymbol{X}$ left out of the regression equation, $\boldsymbol{\beta}_z$ contains the corresponding coefficients, and $\sigma^2\boldsymbol{I}_n$ is the assumed covariance matrix of the error term when $H_0$ is true.
-
-In the heteroskedasticity model, the key elements for measuring effect size are sample size, $n$, and the ratio of the biggest variance to smallest variance, $b$.
-Larger values of both would produce higher effect size.
-However, it is not clear how to incorporate the additional shape parameter, $a$.
-Thus the same approach is used here, where the formula can be written as:
-
-\begin{equation}
-E = \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R})|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\text{diag}(\boldsymbol{R}))\right)
-\label{eq:hetero-effect-size}
-\end{equation}
-
-\noindent where $\boldsymbol{V}$ is the actual covariance matrix of the error term. Derivations for these equations are provided in the Appendix.
+This formulation defines effect size to be
+
+\vspace{-\baselineskip}
+
+$$E = \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\boldsymbol{\mu}_z\right),$$
+
+\noindent where $\text{diag}(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix, $\boldsymbol{X}$ is the design matrix, $\boldsymbol{V}$ is the actual covariance matrix of the error term,
+$\boldsymbol{R} = \boldsymbol{I}_n - \boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the residual operator, $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals where $\boldsymbol{Z}$ contains any higher order terms of $\boldsymbol{X}$ left out of the regression equation, $\boldsymbol{\beta}_z$ contains the corresponding coefficients, and $\sigma^2\boldsymbol{I}_n$ is the assumed covariance matrix of the error term when $H_0$ is true.
 
@@
 
-# \textcolor{red}{Limitations and practicality}
+# Limitations and practicality
 
@@
 
-These `R` packages are used for the work: `tidyverse` [@tidyverse], `lmtest` [@lmtest], `mpoly` [@mpoly], `ggmosaic` [@ggmosaic], `kableExtra` [@kableextra], `patchwork` [@patchwork], `rcartocolor` [@rcartocolor].
-The study website is powered by `PythonAnywhere` [@pythonanywhere] and the `Flask` web framework [@flask].
-The `jsPsych` framework [@jspsych] is used to create behavioural experiments that run in our study website.
+These `R` packages were used for the work: `tidyverse` [@tidyverse], `lmtest` [@lmtest], `mpoly` [@mpoly], `ggmosaic` [@ggmosaic], `kableExtra` [@kableextra], `patchwork` [@patchwork], `rcartocolor` [@rcartocolor].
+The study website was powered by `PythonAnywhere` [@pythonanywhere] and `Flask` web framework [@flask].
+The `jsPsych` framework [@jspsych] was used to create behavioural experiments that run in our study website.
 

### 2024-01-07 paper.tex
reduce to 32 + 3 references (almost there)

@@
 
-\textcolor{red}{Diagnosing a model is an important part of building an appropriate model.}
-In linear regression analysis, studying the residuals from a model fit
-is a common diagnostic activity. Residuals summarise what is not
-captured by the model, and thus provide the capacity to identify what
-might be wrong.
+Diagnosing a model is an important part of building an appropriate
+model. In linear regression analysis, studying the residuals from a
+model fit is a common diagnostic activity. Residuals summarise what is
+not captured by the model, and thus provide the capacity to identify
+what might be wrong.
 
@@
 
-\textcolor{red}{Scatterplots of residuals against the fitted values, and each of the explanatory variables, are commonly used to scrutinize their relationships}.
-If there are any visually discoverable associations, the model is
-potentially inadequate or incorrectly specified.
-\textcolor{red}{We can also potentially discover patterns not directly connected to a linear model assumption from these residual plots, such as the discreteness or skewness of the fitted values, and outliers.}
-To read residual plots, one looks for noticeable departures from the
-model such as non-linear pattern or heteroskedasticity. A non-linear
-pattern would suggest that the model needs to have some additional
-non-linear terms. Heteroskedasticity suggests that the error is
-dependent on the predictors, and hence violates the independence
-assumption. Statistical tests were developed to provide objective
-assessment, for example, of non-linear patterns
-\citep[e.g.][]{ramsey1969tests}, and heteroskedasticity
-\citep[e.g.][]{breusch1979simple}.
+Scatterplots of residuals against the fitted values, and each of the
+explanatory variables, are commonly used to scrutinize their
+relationships. If there are any visually discoverable associations, the
+model is potentially inadequate or incorrectly specified. We can also
+potentially discover patterns not directly connected to a linear model
+assumption from these residual plots, such as the discreteness or
+skewness of the fitted values, and outliers. To read residual plots, one
+looks for noticeable departures from the model such as non-linear
+pattern or heteroskedasticity. A non-linear pattern would suggest that
+the model needs to have some additional non-linear terms.
+Heteroskedasticity suggests that the error is dependent on the
+predictors, and hence violates the independence assumption. Statistical
+tests were developed to provide objective assessment, for example, of
+non-linear patterns \citep[e.g.][]{ramsey1969tests}, and
+heteroskedasticity \citep[e.g.][]{breusch1979simple}.
 
@@
 
-\textcolor{red}{Additionally, relying solely on the subjective assessment of a single plot can be problematic.
-People will almost always see a pattern \citep[see][]{kahneman2011thinking}, so the question that really needs answering is whether any pattern perceived is consistent with randomness, or sampling variability, or noise.
-Correctly judging whether \emph{no} pattern exists in a residual plot is a difficult task.
-\cite{loy2021bringing} emphasizes that this is especially difficult to teach to new analysts and students, and advocates to the broader use of the lineup protocol \citep{bujastatistical2009}.}
-
-\textcolor{red}{The lineup protocol places a data plot in a field of null plots, allowing for a comparison of patterns due purely by chance to what is perceived in the data plot.
-For residual analysis this is especially helpful for gauging whether there is \emph{no} pattern.
-(Figure \ref{fig:first-example-lineup} shows an example of a lineup of residual plots.) In its strict use, one would insist that the data plot is not seen before seeing the lineup, so that the observer doesn't know which is the true plot.
-When used this way, it provides an objective test for data plots.
-\cite{majumdervalidation2013} validated that results from lineups assessed by human observers performed similarly to conventional tests.
-One wouldn't use a lineup when a conventional test exists and is adequate because it is more manually expensive to conduct.
-However, where no adequate conventional test exists, it is invaluable, as shown by \cite{loy2013diagnostic}.
-Here we use the lineup as a vehicle to rigorously explore why experts advise that residual plots are indispensable despite the prevalence of numerical tests.}
+Additionally, relying solely on the subjective assessment of a single
+plot can be problematic. People will almost always see a pattern
+\citep[see][]{kahneman2011thinking}, so the question that really needs
+answering is whether any pattern perceived is consistent with
+randomness, or sampling variability, or noise. Correctly judging whether
+\emph{no} pattern exists in a residual plot is a difficult task.
+\cite{loy2021bringing} emphasizes that this is especially difficult to
+teach to new analysts and students, and advocates to the broader use of
+the lineup protocol \citep{bujastatistical2009}.
+
+The lineup protocol places a data plot in a field of null plots,
+allowing for a comparison of patterns due purely by chance to what is
+perceived in the data plot. For residual analysis this is especially
+helpful for gauging whether there is \emph{no} pattern. (Figure
+\ref{fig:first-example-lineup} shows an example of a lineup of residual
+plots.) In its strict use, one would insist that the data plot is not
+seen before seeing the lineup, so that the observer doesn't know which
+is the true plot. When used this way, it provides an objective test for
+data plots. \cite{majumdervalidation2013} validated that results from
+lineups assessed by human observers performed similarly to conventional
+tests. One wouldn't use a lineup when a conventional test exists and is
+adequate because it is more manually expensive to conduct. However,
+where no adequate conventional test exists, it is invaluable, as shown
+by \cite{loy2013diagnostic}. Here we use the lineup as a vehicle to
+rigorously explore why experts advise that residual plots are
+indispensable despite the prevalence of numerical tests.
 
@@
 
+\vspace{-\baselineskip}
+
 \begin{equation} \label{eq:logistic-regression-1-1}
@@
 During the experiment, every participant recruited from the Prolific
-crowd-sourcing platform \citep{palan2018prolific} is presented with a
-block of 20 lineups. A lineup consists of a randomly placed data plot
-and 19 null plots, which are all residual plots drawn with raw residuals
-on the y-axis and fitted values on the x-axis. An additional horizontal
-red line is added at \(y = 0\) as a visual reference. The data in the
-data plot is simulated from one of two models described in Section
-\ref{simulating-departures-from-good-residuals}, while the data of the
-remaining 19 null plots are generated by the residual rotation technique
-discussed in Section \ref{visual-test-procedure-based-on-lineups}.
+crowd-sourcing platform \citep{palan2018prolific} was presented with a
+block of 20 lineups. A lineup consisted of a randomly placed data plot
+and 19 null plots, which were all residual plots drawn with raw
+residuals on the y-axis and fitted values on the x-axis. An additional
+horizontal red line was added at \(y = 0\) as a visual reference. The
+data in the data plot was simulated from one of two models described in
+Section \ref{simulating-departures-from-good-residuals}, while the data
+of the remaining 19 null plots were generated by the residual rotation
+technique discussed in Section
+\ref{visual-test-procedure-based-on-lineups}.
 
@@
 Overall, we collected 7974 evaluations on 1152 unique lineups performed
-by 443 participants throughout three data collection periods. Details
-about lineup allocation and data collection are provided in the
-Appendix.
-
-A summary of the factors used in the experiment can be found in Table
-\ref{tab:model-factor-table}. There were four levels of the non-linear
-structure, and three levels of heteroskedastic structure. The signal
-strength was controlled by error variance (\(\sigma\)) for the
-non-linear pattern, and by a ratio (\(b\)) parameter for the
-heteroskedasticity. Additionally, three levels of sample size (\(n\))
-and four different fitted value distributions were incorporated.
+by 443 participants. A summary of the factors used in the experiment can
+be found in Table \ref{tab:model-factor-table}. There were four levels
+of the non-linear structure, and three levels of heteroskedastic
+structure. The signal strength was controlled by error variance
+(\(\sigma\)) for the non-linear pattern, and by a ratio (\(b\))
+parameter for the heteroskedasticity. Additionally, three levels of
+sample size (\(n\)) and four different fitted value distributions were
+incorporated.
 
@@
 Data collection period I was designed to study the ability of
-participants to detect the effect of a non-linear term
-\(\boldsymbol{z}\) constructed using Hermite polynomials on random
-vector \(\boldsymbol{x}\) formulated as
-
-\[ \boldsymbol{y} = \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},~\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)\]
-
-\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
-\(\boldsymbol{\varepsilon}\), \(\boldsymbol{z}\) are vectors of size
-\(n\), \(\boldsymbol{1}_n\) is a vector of ones of size \(n\),
-\(\boldsymbol{z} \propto He_j(\boldsymbol{x})\), \(He_{j}(.)\) is the
-\(j\)th-order probabilist's Hermite polynomials
+participants to detect non-linearity departures from residual plots. The
+non-linearity departure was constructed by omitting a \(j\)th order
+Hermite polynomial
 \citetext{\citealp{hermite1864nouveau}; \citealp[originally
-by][]{de1820theorie}}.
-
-To fit a misspecified model, the \(\boldsymbol{z}\) term, which is a
-higher order term of \(\boldsymbol{x}\), was omitted from the null
-regression formula
-
-\begin{equation} \label{eq:null-model}
-\boldsymbol{y} = \beta_0\boldsymbol{1}_n + \beta_1 \boldsymbol{x} + \boldsymbol{u}, \boldsymbol{u} \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n).
-\end{equation}
-
-Four different values of \(j = 2, 3, 6, 18\) were chosen so that
-distinct shapes of non-linearity were included in the residual plot.
-These include ``U'', ``S'', ``M'' and ``triple-U'' shape as shown in
-Figure \ref{fig:different-shape-of-herimite}. A greater value of \(j\)
-will result in a curve with more turning points. It is expected that the
-``U'' shape will be the easiest to detect, and as the shape gets more
-complex it will be harder to perceive in a scatterplot, particularly
-when there is noise. Figure \ref{fig:different-sigma} shows the ``U''
-shape for different amounts of noise (\(\sigma\)).
-
-Data collection period II was similar to Data collection I but focuses
-on heteroskedasticity visual patterns formulated as
-
-\[\boldsymbol{y} = \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{\varepsilon},~\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{1}_n + (2 - |a|)(\boldsymbol{x} - a\boldsymbol{1}_n)'(\boldsymbol{x} - a\boldsymbol{1}_n) b \boldsymbol{I}_n),\]
-\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
-\(\boldsymbol{\varepsilon}\) are vectors of size \(n\).
-
-The null regression model used to fit the realizations generated by the
-above model is formulated exactly the same as Equation
-\ref{eq:null-model}. For \(b \neq 0\), the variance-covariance matrix of
-the error term \(\boldsymbol{\varepsilon}\) is correlated with the
-predictor \(\boldsymbol{x}\), which will lead to the presence of
-heteroskedasticity.
-
-Visual patterns of heteroskedasticity are simulated using three
-different shapes (\(a\) = -1, 0, 1) including ``left-triangle'',
+by][]{de1820theorie}} term of the predictor from the simple linear
+regression equation. Four different values of \(j = 2, 3, 6, 18\) were
+chosen so that distinct shapes of non-linearity were included in the
+residual plots. These include ``U'', ``S'', ``M'' and ``triple-U'' shape
+as shown in Figure \ref{fig:different-shape-of-herimite}. A greater
+value of \(j\) will result in a curve with more turning points. It is
+expected that the ``U'' shape will be the easiest to detect, and as the
+shape gets more complex it will be harder to perceive in a scatterplot,
+particularly when there is noise. Figure \ref{fig:different-sigma} shows
+the ``U'' shape for different amounts of noise (\(\sigma\)).
+
+Data collection period II was similar to period I but focuses on
+heteroskedasticity departures. We generated the heteroskedasticity
+departures by setting the variance-covariance matrix of the error term
+as a function of the predictor, but fitted the data with the simple
+linear regression model, intentionally violated the constant variance
+assumption. Visual patterns of heteroskedasticity are simulated using
+three different shapes (\(a\) = -1, 0, 1) including ``left-triangle'',
 ``butterfly'' and ``right-triangle'' shapes as displayed in Figure
 \ref{fig:different-shape-of-heter}. Figure \ref{fig:different-b} shows
-the butterfly shape as the ratio parameter (\(b\)) is changed.
-
-More details about the design of the data generating process are
-provided in the Appendix.
+the butterfly shape as the ratio parameter (\(b\)) is changed. More
+details about the simulation process are provided in the Appendix.
 
@@
 simulation assess this by using four different distributions to
-represent fitted values, constructed by different sampling of the raw
-predictor \(X_{raw}\), namely, uniform (\(U(-1, 1)\)), normal
-(\(N(0, 0.3^2)\)), skewed (\(\text{lognormal}(0, 0.6^2)/3\)) and
-discrete (\(U\{1, 5\}\)).
+represent fitted values, constructed by different sampling of the
+predictor, including \(U(-1, 1)\) (uniform), \(N(0, 0.3^2)\) (normal),
+\(\text{lognormal}(0, 0.6^2)/3\) (skewed) and \(U\{-1, 1\}\) (discrete).
 
@@
 the upper limit, because it is typically enough for structure to be
-visible in a scatterplot reliably.
-\textcolor{red}{Beyond 300, the scatterplot should probably be used with transparency or replaced with a density or binned plot as scatterplots suffer from over-plotting.}
+visible in a scatterplot reliably. Beyond 300, the scatterplot should
+probably be used with transparency or replaced with a density or binned
+plot as scatterplots suffer from over-plotting.
 
@@
 For the non-linearity model, the key items defining effect size are
-sample size (\(n\)) and variance of the error term (\(\sigma^2\)), and
-so effect size would be roughly calculated as \(\sqrt{n}/{\sigma}\). As
-sample size increases the effect size would increase, but as variance
-increases the effect size decreases. However, it is not clear how the
-additional parameter for the model polynomial order, \(j\), should be
-incorporated. Intuitively, the large \(j\) means more complex pattern,
-which likely means effect size would decrease. For the purposes of our
-calculations we have chosen to use an approach based on Kullback-Leibler
-divergence \citep{kullback1951information}, coupled with simulation.
-This formulation defines effect size to be:
-
-\begin{equation}
-E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(\text{diag}(\boldsymbol{R}\sigma^2))^{-1}\boldsymbol{\mu}_z\right)
-\label{eq:nonlin-effect-size}
-\end{equation}
+sample size (\(n\)) and the noise level (\(\sigma^2\)), and so effect
+size would be roughly calculated as \(\sqrt{n}/{\sigma}\). Increasing
+sample size tends to boost the effect size, while heightened noise
+diminishes it. However, it is not clear how the additional parameter for
+the model polynomial order, \(j\), should be incorporated. Intuitively,
+the large \(j\) means more complex pattern, which likely means effect
+size would decrease. Similarly, in the heteroskedasticity model, effect
+size relies on sample size (\(n\)) and the ratio of the largest to
+smallest variance, \(b\). Larger values of both would produce higher
+effect size, but the role of the additional shape parameter, \(a\), in
+this context is unclear.
+
+For the purposes of our calculations we have chosen to use an approach
+based on Kullback-Leibler divergence \citep{kullback1951information},
+coupled with simulation. This formulation defines effect size to be
+
+\vspace{-\baselineskip}
+
+\[E = \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\boldsymbol{\mu}_z\right),\]
 
 \noindent where \(\text{diag}(.)\) is the diagonal matrix constructed
-from the diagonal elements of a matrix,
-\(\boldsymbol{R} = \boldsymbol{I}_n - \boldsymbol{H}\) is the residual
-operator,
-\(\boldsymbol{H} = \boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\)
-is the hat matrix,
+from the diagonal elements of a matrix, \(\boldsymbol{X}\) is the design
+matrix, \(\boldsymbol{V}\) is the actual covariance matrix of the error
+term,
+\(\boldsymbol{R} = \boldsymbol{I}_n - \boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\)
+is the residual operator,
 \(\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z\)
-is the expected values of residuals where \(\boldsymbol{Z}\) contain any
-higher order terms of \(\boldsymbol{X}\) left out of the regression
+is the expected values of residuals where \(\boldsymbol{Z}\) contains
+any higher order terms of \(\boldsymbol{X}\) left out of the regression
 equation, \(\boldsymbol{\beta}_z\) contains the corresponding
@@
 matrix of the error term when \(H_0\) is true.
-
-In the heteroskedasticity model, the key elements for measuring effect
-size are sample size, \(n\), and the ratio of the biggest variance to
-smallest variance, \(b\). Larger values of both would produce higher
-effect size. However, it is not clear how to incorporate the additional
-shape parameter, \(a\). Thus the same approach is used here, where the
-formula can be written as:
-
-\begin{equation}
-E = \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R})|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\text{diag}(\boldsymbol{R}))\right)
-\label{eq:hetero-effect-size}
-\end{equation}
-
-\noindent where \(\boldsymbol{V}\) is the actual covariance matrix of
-the error term. Derivations for these equations are provided in the
-Appendix.
 
@@
 
-\hypertarget{section}{%
-\section{\texorpdfstring{\textcolor{red}{Limitations and practicality}}{}}\label{section}}
+\hypertarget{limitations-and-practicality}{%
+\section{Limitations and
+practicality}\label{limitations-and-practicality}}
 
@@
 
-These \texttt{R} packages are used for the work: \texttt{tidyverse}
+These \texttt{R} packages were used for the work: \texttt{tidyverse}
 \citep{tidyverse}, \texttt{lmtest} \citep{lmtest}, \texttt{mpoly}
@@
 \citep{kableextra}, \texttt{patchwork} \citep{patchwork},
-\texttt{rcartocolor} \citep{rcartocolor}. The study website is powered
-by \texttt{PythonAnywhere} \citep{pythonanywhere} and the \texttt{Flask}
-web framework \citep{flask}. The \texttt{jsPsych} framework
-\citep{jspsych} is used to create behavioural experiments that run in
-our study website.
+\texttt{rcartocolor} \citep{rcartocolor}. The study website was powered
+by \texttt{PythonAnywhere} \citep{pythonanywhere} and \texttt{Flask} web
+framework \citep{flask}. The \texttt{jsPsych} framework \citep{jspsych}
+was used to create behavioural experiments that run in our study
+website.
 

### 2024-01-15 paper.Rmd
sec 1-2 update

@@
 \cite{majumdervalidation2013} validated that results from lineups assessed by human observers performed similarly to conventional tests.
-One wouldn't use a lineup when a conventional test exists and is adequate because it is more manually expensive to conduct.
+One would not use a lineup when a conventional test exists and is adequate because it is more manually expensive to conduct.
 However, where no adequate conventional test exists, it is invaluable, as shown by \cite{loy2013diagnostic}.
@@
 There can be various types of departures from this ideal pattern.
-Non-linearity, heteroskedasticity and non-normality, shown in Figures \ref{fig:residual-plot-common-departures}B, \ref{fig:residual-plot-common-departures}C, and \ref{fig:residual-plot-common-departures}D are three commonly checked departures.
+Non-linearity, heteroskedasticity and non-normality, shown in Figures \ref{fig:residual-plot-common-departures}B, \ref{fig:residual-plot-common-departures}C, and \ref{fig:residual-plot-common-departures}D, respectively, are three commonly checked departures.
 
@@
 An example residual plot containing visual pattern of non-linearity is shown in Figure \ref{fig:residual-plot-common-departures}B.
-One can clearly observe the "S-shape" from the residual plot, which corresponds to cubic term that should have been included in the model.
+One can clearly observe the "S-shape" from the residual plot, which corresponds to the cubic term that should have been included in the model.
 
@@
 Because we focus on scatterplots, non-normality is not one of the departures examined in depth in this paper.
-(@loy2016variations discuss related work on non-normality checking.)
+[@loy2016variations discuss related work on non-normality checking.]
 
@@
 A key goal of residual plot diagnostics is to identify potential issues that could lead to incorrect conclusions or errors in subsequent analyses, but minor defects in the model are unlikely to have a significant impact and may be best disregarded for practical purposes.
-The experiment discussed in this paper specifically addresses this tension between statistical and practical statistics.
+The experiment discussed in this paper specifically addresses this tension between statistical and practical significance.
 
@@
 
-The examination of data plots to infer signal or patterns (or lack thereof) is fraught with variation in the human ability to interpret and decode the information embedded in a graph [@cleveland1984graphical].
-Human examination of diagnostic plots can feel subjective.
-Ideally we would expect visual evaluation to be done objectively with low human-to-human variation in reading data plots, but we cannot assume this is the case.
-
+The examination of data plots to infer signals or patterns (or lack thereof) is fraught with variation in the human ability to interpret and decode the information embedded in a graph [@cleveland1984graphical].
 In practice, over-interpretation of a single plot is common.
 For instance, @roy2015using described a published example where authors over-interpreted separation between gene groups from a two-dimensional projection of a linear discriminant analysis even when there were no differences in the expression levels between the gene groups.
+
 One solution to over-interpretation is to examine the plot in the context of natural sampling variability assumed by the model, called the lineup protocol, as proposed in @buja2009statistical.
@@
 Their results suggest that visual tests are capable of testing the significance of a single predictor with a similar power to a t-test, though they express that in general it is unnecessary to use visual inference if there exists a corresponding conventional test, and they do not expect the visual test to perform equally well as the conventional test.
-In their third experiment, where the contamination of the data violate the assumptions of the convetional test, visual test outperforms the conventional test by a large margin.
-This is encouraging, as it supports the use of visual inference in situations where there are no existing numerical testing procedures.
+In their third experiment, where the contamination of the data violate the assumptions of the conventional test, visual test outperforms the conventional test by a large margin.
+This supports the use of visual inference in situations where there are no existing numerical testing procedures.
 Visual inference has also been integrated into diagnostics for hierarchical linear models where the lineup protocol is used to judge the assumptions of linearity, normality and constant error variance for both the level-1 and level-2 residuals \citep{loy2013diagnostic, loy2014hlmdiag, loy2015you}.
@@
 It includes more details about the experimental setup, the derivation of the effect size, the effect of data collection period, and the estimate of $\alpha$.
-
 The R package `visage` is currently available at https://github.com/TengMCing/visage.

### 2024-01-15 paper.tex
sec 1-2 update

@@
 lineups assessed by human observers performed similarly to conventional
-tests. One wouldn't use a lineup when a conventional test exists and is
+tests. One would not use a lineup when a conventional test exists and is
 adequate because it is more manually expensive to conduct. However,
@@
 \ref{fig:residual-plot-common-departures}C, and
-\ref{fig:residual-plot-common-departures}D are three commonly checked
-departures.
+\ref{fig:residual-plot-common-departures}D, respectively, are three
+commonly checked departures.
 
@@
 Figure \ref{fig:residual-plot-common-departures}B. One can clearly
-observe the ``S-shape'' from the residual plot, which corresponds to
+observe the ``S-shape'' from the residual plot, which corresponds to the
 cubic term that should have been included in the model.
@@
 scatterplots, non-normality is not one of the departures examined in
-depth in this paper. (\citet{loy2016variations} discuss related work on
-non-normality checking.)
+depth in this paper. \citep[ discuss related work on non-normality
+checking.]{loy2016variations}
 
@@
 discussed in this paper specifically addresses this tension between
-statistical and practical statistics.
+statistical and practical significance.
 
@@
 
-The examination of data plots to infer signal or patterns (or lack
+The examination of data plots to infer signals or patterns (or lack
 thereof) is fraught with variation in the human ability to interpret and
 decode the information embedded in a graph
-\citep{cleveland1984graphical}. Human examination of diagnostic plots
-can feel subjective. Ideally we would expect visual evaluation to be
-done objectively with low human-to-human variation in reading data
-plots, but we cannot assume this is the case.
-
-In practice, over-interpretation of a single plot is common. For
-instance, \citet{roy2015using} described a published example where
-authors over-interpreted separation between gene groups from a
-two-dimensional projection of a linear discriminant analysis even when
-there were no differences in the expression levels between the gene
-groups. One solution to over-interpretation is to examine the plot in
-the context of natural sampling variability assumed by the model, called
-the lineup protocol, as proposed in \citet{buja2009statistical}.
+\citep{cleveland1984graphical}. In practice, over-interpretation of a
+single plot is common. For instance, \citet{roy2015using} described a
+published example where authors over-interpreted separation between gene
+groups from a two-dimensional projection of a linear discriminant
+analysis even when there were no differences in the expression levels
+between the gene groups.
+
+One solution to over-interpretation is to examine the plot in the
+context of natural sampling variability assumed by the model, called the
+lineup protocol, as proposed in \citet{buja2009statistical}.
 \citet{majumder2013validation} showed that the lineup protocol is
@@
 equally well as the conventional test. In their third experiment, where
-the contamination of the data violate the assumptions of the convetional
-test, visual test outperforms the conventional test by a large margin.
-This is encouraging, as it supports the use of visual inference in
-situations where there are no existing numerical testing procedures.
-Visual inference has also been integrated into diagnostics for
-hierarchical linear models where the lineup protocol is used to judge
-the assumptions of linearity, normality and constant error variance for
-both the level-1 and level-2 residuals
+the contamination of the data violate the assumptions of the
+conventional test, visual test outperforms the conventional test by a
+large margin. This supports the use of visual inference in situations
+where there are no existing numerical testing procedures. Visual
+inference has also been integrated into diagnostics for hierarchical
+linear models where the lineup protocol is used to judge the assumptions
+of linearity, normality and constant error variance for both the level-1
+and level-2 residuals
 \citep{loy2013diagnostic, loy2014hlmdiag, loy2015you}.
@@
 the effect size, the effect of data collection period, and the estimate
-of \(\alpha\).
-
-The R package \texttt{visage} is currently available at
+of \(\alpha\). The R package \texttt{visage} is currently available at
 \url{https://github.com/TengMCing/visage}.

### 2024-01-15 paper.Rmd
sec 4 done

@@
 In diagnosing a model fit using the residuals, we are generally interested in testing whether "*the regression model is correctly specified*" ($H_0$) against the broad alternative  "*the regression model is misspecified*" ($H_a$).
-
 However, it is practically impossible to test this broad $H_0$ with conventional tests, because they need specific structure causing the departure to be quantifiable in order to be computable.
@@
 
-The power of a conventional hypothesis test is affected by both the true parameter $\boldsymbol{\theta}$ and the sample size $n$.
+The power of a conventional hypothesis test is affected by both the true parameters $\boldsymbol{\theta}$ and the sample size $n$.
 These two can be quantified in terms of effect size $E$ to measure the strength of the residual departures from the model assumptions.
@@
 The power of a visual test on the other hand, may additionally depend on the ability of the particular participant, as the skill of each individual may affect the number of observers who identify the data plot from the lineup [@majumder2013validation].
-To address this issue, @majumder2013validation models the probability of participant $j$ correctly picking the data plot from lineup $l$ using a mixed-effect logistic regression, with participants treated as random effect.
+To address this issue, @majumder2013validation models the probability of participant $j$ correctly picking the data plot from lineup $l$ using a mixed-effect logistic regression, with participants treated as random effects.
 Then, the estimated power of a visual test evaluated by a single participant is the predicted value obtained from the mixed effects model.
@@
 ```{r different-shape-of-heter, fig.pos="!h", fig.height = 2.67, fig.cap = 'Heteroskedasticity forms used in the experiment.
-Three different shapes ($a = -1, 0, 1$) are used in the experiment to create left-triangle, "butterfly" and "right-triangle" shapes, respectively.'}
+Three different shapes ($a = -1, 0, 1$) are used in the experiment to create "left-triangle", "butterfly" and "right-triangle" shapes, respectively.'}
 
@@
 
-Fitted values are a function of the independent variables, and the distribution of the observed values affects the distribution of the fitted values.
-In the best case scenario the fitted values will have a uniform distribution, which means that there is even coverage of possible observed values across all of the predictors.
+Fitted values are a function of the independent variables (or predictors), and the distribution of the observed values affects the distribution of the fitted values.
+Ideally, we would want the fitted values to have a uniform coverage across the range of observed values or have a uniform distribution across all of the predictors.
 This is not always present in the collected data.
@@
 Figure \ref{fig:different-dist} shows the non-linear pattern, a "U" shape, with the different fitted value distributions.
-We would expect that structure in residual plots would be easier to perceive when the fitted values are uniformly distributed.
+We would expect that structure in residual plots would be easier to perceive when the predictors are uniformly distributed.
 
@@
 
-The lineups are allocated to participants in a manner that covers the experimental design and is relatively uniform.
-To manage this, we use effect size to measure the signal strength, which helps in assigning a set of lineups with a range of difficulties to each participant.
+The lineups are allocated to participants in a manner that uniformly covers the combination of experimental factors in Table \ref{tab:model-factor-table}.
+In addition, we use effect size to measure the signal strength, which helps in assigning a set of lineups with a range of difficulties to each participant.
 
@@
 
-To compute the effect size for each lineup we simulate a sufficient large number of samples from the same model with the number of observations $n$ fixed for each sample.
+To compute the effect size for each lineup we simulate a sufficiently large number of samples from the same model with the number of observations $n$ fixed for each sample.
 We then compute the effect size for each sample and take the average as the final value.

### 2024-01-15 paper.tex
sec 4 done

@@
 specified}'' (\(H_0\)) against the broad alternative ``\emph{the
-regression model is misspecified}'' (\(H_a\)).
-
-However, it is practically impossible to test this broad \(H_0\) with
-conventional tests, because they need specific structure causing the
-departure to be quantifiable in order to be computable. For example, the
-RESET test for detecting non-linear departures is formulated by fitting
+regression model is misspecified}'' (\(H_a\)). However, it is
+practically impossible to test this broad \(H_0\) with conventional
+tests, because they need specific structure causing the departure to be
+quantifiable in order to be computable. For example, the RESET test for
+detecting non-linear departures is formulated by fitting
 \(y = \tau_0 + \sum_{i=1}^{p}\tau_px_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)\)
@@
 The power of a conventional hypothesis test is affected by both the true
-parameter \(\boldsymbol{\theta}\) and the sample size \(n\). These two
+parameters \(\boldsymbol{\theta}\) and the sample size \(n\). These two
 can be quantified in terms of effect size \(E\) to measure the strength
@@
 mixed-effect logistic regression, with participants treated as random
-effect. Then, the estimated power of a visual test evaluated by a single
-participant is the predicted value obtained from the mixed effects
-model. However, this mixed effects model does not work with scenario
-where participants are asked to select one or more most different plots.
-In this scenario, having the probability of a participant \(j\)
-correctly picking the data plot from a lineup \(l\) is insufficient to
-determine the power of a visual test because it does not provide
-information about the number of selections made by the participant for
-the calculation of the \(p\)-value. Therefore, we directly estimate the
-probability of a lineup being rejected by assuming that individual skill
-has negligible effect on the variation of the power. This assumption
-essentially averages out the subject ability and helps to simplify the
-model structure, thereby obviating a costly large-scale experiment to
-estimate complex covariance matrices. The same model given in Equation
-\ref{eq:logistic-regression-1-1} is applied to model the power of a
-visual test.
+effects. Then, the estimated power of a visual test evaluated by a
+single participant is the predicted value obtained from the mixed
+effects model. However, this mixed effects model does not work with
+scenario where participants are asked to select one or more most
+different plots. In this scenario, having the probability of a
+participant \(j\) correctly picking the data plot from a lineup \(l\) is
+insufficient to determine the power of a visual test because it does not
+provide information about the number of selections made by the
+participant for the calculation of the \(p\)-value. Therefore, we
+directly estimate the probability of a lineup being rejected by assuming
+that individual skill has negligible effect on the variation of the
+power. This assumption essentially averages out the subject ability and
+helps to simplify the model structure, thereby obviating a costly
+large-scale experiment to estimate complex covariance matrices. The same
+model given in Equation \ref{eq:logistic-regression-1-1} is applied to
+model the power of a visual test.
 
@@
 \caption{Heteroskedasticity forms used in the experiment.
-Three different shapes ($a = -1, 0, 1$) are used in the experiment to create left-triangle, "butterfly" and "right-triangle" shapes, respectively.}\label{fig:different-shape-of-heter}
+Three different shapes ($a = -1, 0, 1$) are used in the experiment to create "left-triangle", "butterfly" and "right-triangle" shapes, respectively.}\label{fig:different-shape-of-heter}
 \end{figure}
@@
 
-Fitted values are a function of the independent variables, and the
-distribution of the observed values affects the distribution of the
-fitted values. In the best case scenario the fitted values will have a
-uniform distribution, which means that there is even coverage of
-possible observed values across all of the predictors. This is not
+Fitted values are a function of the independent variables (or
+predictors), and the distribution of the observed values affects the
+distribution of the fitted values. Ideally, we would want the fitted
+values to have a uniform coverage across the range of observed values or
+have a uniform distribution across all of the predictors. This is not
 always present in the collected data. Sometimes the fitted values are
@@
 that structure in residual plots would be easier to perceive when the
-fitted values are uniformly distributed.
+predictors are uniformly distributed.
 
@@
 
-The lineups are allocated to participants in a manner that covers the
-experimental design and is relatively uniform. To manage this, we use
-effect size to measure the signal strength, which helps in assigning a
-set of lineups with a range of difficulties to each participant.
+The lineups are allocated to participants in a manner that uniformly
+covers the combination of experimental factors in Table
+\ref{tab:model-factor-table}. In addition, we use effect size to measure
+the signal strength, which helps in assigning a set of lineups with a
+range of difficulties to each participant.
 
@@
 
-To compute the effect size for each lineup we simulate a sufficient
+To compute the effect size for each lineup we simulate a sufficiently
 large number of samples from the same model with the number of

### 2024-01-16 paper.Rmd
fixes on technicality

@@
                                  "Size (sym6)", 
-                                 "Distribution of fitted values")) %>% #"Sample size\n(sym6)")) %>%
+                                 "Distribution of the predictor")) %>% #"Sample size\n(sym6)")) %>%
   kableExtra::add_header_above(c("Non-linearity" = 2, 
@@
 
-```{r different-sigma, fig.pos="!h", fig.height = 2, fig.cap = 'Examining the effect of $\\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform fitted value distribution and the "U" shape.
-As $\\sigma$ increases the signal strength decreases, to the point that the "U" is almost unrecognisable when $\\sigma=4$.'}
+```{r different-sigma, fig.pos="!h", fig.height = 2, fig.cap = 'Examining the effect of $\\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform fitted value distribution and the ``U" shape.
+As $\\sigma$ increases the signal strength decreases, to the point that the ``U" is almost unrecognisable when $\\sigma=4$.'}
 
@@
 ```{r different-shape-of-heter, fig.pos="!h", fig.height = 2.67, fig.cap = 'Heteroskedasticity forms used in the experiment.
-Three different shapes ($a = -1, 0, 1$) are used in the experiment to create "left-triangle", "butterfly" and "right-triangle" shapes, respectively.'}
+Three different shapes ($a = -1, 0, 1$) are used in the experiment to create ``left-triangle", ``butterfly" and ``right-triangle" shapes, respectively.'}
 
@@
 
-The data was collected on lineups constructed from four different fitted value distributions: uniform, normal, skewed and discrete.
+The data was collected on lineups constructed from four different fitted value distributions that stem from the corresponding predictor distribution: uniform, normal, skewed and discrete.
+Henceforth, we refer to these four different fitted value distributions with respect to their predictor distribution.
 More data was collected on the uniform distribution (each evaluated by 11 participants) than the others (each evaluated by 5 participants).
-The analysis in Sections \ref{power-analysis}--\ref{hetero-analysis} uses only results from lineups generated with uniform fitted values, for a total `r nrow(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"))` lineup evaluations.
+The analysis in Sections \ref{power-analysis}--\ref{hetero-analysis} uses only results from lineups with uniform distribution, for a total `r nrow(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"))` lineup evaluations.
 This allows us to compare the conventional and visual test performance in an optimal scenario.
@@
 
-```{r polypower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for non-linear patterns (uniform fitted values only).
+```{r polypower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for non-linear patterns (uniform distribution only).
 The power curves are estimated using logistic regression, and the horizontal lines of dots represent non-reject and reject results from visual tests for each lineup.
@@
 Note that the BP and SW tests have much lower power, which is expected because they are not designed to detect non-linearity.
-The bootstrapped power curves for the visual test are effectively a shift right from that of the RESET test.
+The bootstrapped power curves for the visual test are effectively a right shift from that of the RESET test.
 This means that the RESET test will reject at a lower effect size (less structure) than the visual test, but otherwise the performance will be similar.
@@
 However the BP test would reject at around effect size 2.5.
-Interestingly, the power curve for the SW test (for non-normality) is only slightly different to that of the visual test, suggesting that it performs reasonably for detecting heteroskedasticity, too.
+Interestingly, the power curve for the SW test (for non-normality) is only slightly different to that of the visual test, suggesting that it performs reasonably well for detecting heteroskedasticity, too.
 The power curve for the BP test suggests it is not useful for detecting heteroskedasticity, as expected.
@@
 But the BP test $p\text{-value} = 0.056$, is slightly above the significance cutoff of $0.05$.
-This lineup was evaluated by 11 participants, it has experimental factors $a = 0$ ("butterfly" shape), $b = 64$ (large variance ratio), $n = 50$ (small sample size), and a uniform distribution for the fitted values.
+This lineup was evaluated by 11 participants, it has experimental factors $a = 0$ ("butterfly" shape), $b = 64$ (large variance ratio), $n = 50$ (small sample size), and a uniform distribution for the predictor.
 It may have been the small sample size and the presence of a few outliers that may have resulted in the lack of detection by the conventional test.
@@
 The butterfly has higher power in both tests.
-Curiously, the visual test has a slightly higher power for the "left-triangle" than the "right-triangle" shape, when it would be expected that they should be similar, which is observed in conventional testing.'}
+Curiously, the visual test has a slightly higher power for the ``left-triangle" than the ``right-triangle" shape, when it would be expected that they should be similar, which is observed in conventional testing.'}
 
@@
 Where $X$ is observed, the distribution of the $X$ values in the sample, or consequently $\hat{Y}$, may affect the ability to read any patterns in the residual plots.
-The effect of fitted value distribution on test performance is assess using four different distributions of fitted values: uniform, normal, discrete and lognormal (skewed).
-We expect that if $\hat{Y}$ has a uniform distribution, it is easier to read the relationship with the residuals.
+The effect of fitted value distribution on test performance is assess using four different distributions of fitted values stemming from the predictor distributions: uniform, normal, discrete and lognormal (skewed).
+We expect that if all predictors have a uniform distribution, it is easier to read the relationship with the residuals.
 

### 2024-01-16 paper.tex
fixes on technicality

@@
 \cmidrule(l{3pt}r{3pt}){1-2} \cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6}
-\multicolumn{1}{c}{Poly Order ($j$)} & \multicolumn{1}{c}{SD ($\sigma$)} & \multicolumn{1}{c}{Shape ($a$)} & \multicolumn{1}{c}{Ratio ($b$)} & \multicolumn{1}{c}{Size ($n$)} & \multicolumn{1}{c}{Distribution of fitted values} \\
+\multicolumn{1}{c}{Poly Order ($j$)} & \multicolumn{1}{c}{SD ($\sigma$)} & \multicolumn{1}{c}{Shape ($a$)} & \multicolumn{1}{c}{Ratio ($b$)} & \multicolumn{1}{c}{Size ($n$)} & \multicolumn{1}{c}{Distribution of the predictor} \\
 \cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
@@
 
-\caption{Examining the effect of $\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform fitted value distribution and the "U" shape.
-As $\sigma$ increases the signal strength decreases, to the point that the "U" is almost unrecognisable when $\sigma=4$.}\label{fig:different-sigma}
+\caption{Examining the effect of $\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform fitted value distribution and the ``U" shape.
+As $\sigma$ increases the signal strength decreases, to the point that the ``U" is almost unrecognisable when $\sigma=4$.}\label{fig:different-sigma}
 \end{figure}
@@
 \caption{Heteroskedasticity forms used in the experiment.
-Three different shapes ($a = -1, 0, 1$) are used in the experiment to create "left-triangle", "butterfly" and "right-triangle" shapes, respectively.}\label{fig:different-shape-of-heter}
+Three different shapes ($a = -1, 0, 1$) are used in the experiment to create ``left-triangle", ``butterfly" and ``right-triangle" shapes, respectively.}\label{fig:different-shape-of-heter}
 \end{figure}
@@
 The data was collected on lineups constructed from four different fitted
-value distributions: uniform, normal, skewed and discrete. More data was
-collected on the uniform distribution (each evaluated by 11
-participants) than the others (each evaluated by 5 participants). The
-analysis in Sections \ref{power-analysis}--\ref{hetero-analysis} uses
-only results from lineups generated with uniform fitted values, for a
-total 3069 lineup evaluations. This allows us to compare the
-conventional and visual test performance in an optimal scenario. Section
+value distributions that stem from the corresponding predictor
+distribution: uniform, normal, skewed and discrete. Henceforth, we refer
+to these four different fitted value distributions with respect to their
+predictor distribution. More data was collected on the uniform
+distribution (each evaluated by 11 participants) than the others (each
+evaluated by 5 participants). The analysis in Sections
+\ref{power-analysis}--\ref{hetero-analysis} uses only results from
+lineups with uniform distribution, for a total 3069 lineup evaluations.
+This allows us to compare the conventional and visual test performance
+in an optimal scenario. Section
 \ref{effect-of-fitted-value-distributions} examines how the results may
@@
 non-linearity. The bootstrapped power curves for the visual test are
-effectively a shift right from that of the RESET test. This means that
+effectively a right shift from that of the RESET test. This means that
 the RESET test will reject at a lower effect size (less structure) than
@@
 curve for the SW test (for non-normality) is only slightly different to
-that of the visual test, suggesting that it performs reasonably for
+that of the visual test, suggesting that it performs reasonably well for
 detecting heteroskedasticity, too. The power curve for the BP test
@@
 (``butterfly'' shape), \(b = 64\) (large variance ratio), \(n = 50\)
-(small sample size), and a uniform distribution for the fitted values.
-It may have been the small sample size and the presence of a few
-outliers that may have resulted in the lack of detection by the
-conventional test.
+(small sample size), and a uniform distribution for the predictor. It
+may have been the small sample size and the presence of a few outliers
+that may have resulted in the lack of detection by the conventional
+test.
 
@@
 The butterfly has higher power in both tests.
-Curiously, the visual test has a slightly higher power for the "left-triangle" than the "right-triangle" shape, when it would be expected that they should be similar, which is observed in conventional testing.}\label{fig:heter-power-uniform-a}
+Curiously, the visual test has a slightly higher power for the ``left-triangle" than the ``right-triangle" shape, when it would be expected that they should be similar, which is observed in conventional testing.}\label{fig:heter-power-uniform-a}
 \end{figure}
@@
 effect of fitted value distribution on test performance is assess using
-four different distributions of fitted values: uniform, normal, discrete
-and lognormal (skewed). We expect that if \(\hat{Y}\) has a uniform
-distribution, it is easier to read the relationship with the residuals.
+four different distributions of fitted values stemming from the
+predictor distributions: uniform, normal, discrete and lognormal
+(skewed). We expect that if all predictors have a uniform distribution,
+it is easier to read the relationship with the residuals.
 

### 2024-03-24 JCGS_response_to_reviews/response_to_review_R1.Rmd
R2 submitted

@@
 title: "Response to Reviews: JCGS-23-139.R1"
-author: Weihao Li, Emi Tanaka, Dianne Cook, Susan VanderPlas
 output: pdf_document
 date: "`r Sys.Date()`"
-bibliography: ../ref.bib
 ---
@@
 
-Thank you. (...)
+Thank you for the positive feedback.
 
@@
 
-Thank you. (...)
+We appreciate the reviewer's acknowledgment of our efforts in addressing the comments and improving the manuscript.
+Thank you for your remaining comments, which we have carefully considered and addressed in the revised version.
 
@@
 
+Thank you for the suggestion to condense Sections 4 and 5.
+We agree that the length of these sections contributed to the overall length of the paper.
+In the revised version, we have streamlined these sections, reducing the main text by 4 pages.
+Some of the details have been moved to the appendix to improve readability while retaining the necessary information.
+
 *2. In the Figure 1 caption, clarify what is meant by the "at position $2^2 + 2$".*
 
+Thank you for seeking clarification on the caption of Figure 1.
+The expression $2^2 + 2$ was intended to introduce a brief pause before revealing the data plot, allowing the reader time to examine the lineup first.
+However, we agree that this notation may be unclear.
+In the revised version, we have replaced it with the numerical value $6$, which represents the position where the data plot is presented.
+
 *3. In the footnote on the bottom of page6, replace "didn't" and "it's" by "did not" and "it is", respectively.*
+
+Thank you for pointing out the use of contractions in the footnote on page 6.
+We have replaced "didn't" with "did not" and "it's" with "it is" as suggested.
+Additionally, we have carefully reviewed the manuscript and addressed any other inappropriate use of contractions to ensure appropriate academic writing style throughout the text.
 
@@
 
+Thank you for the insightful feedback on the language used in the last sentence of Section 2.2.
+We appreciate you pointing out the awkward phrasing of "statistical and practical statistics." As you suggested, the use of "statistical" as an adjective modifying "statistics" is redundant and may cause confusion.
+In the revised version, we have replaced that phrase with "statistical significance and practical significance" to better reflect the contrast between inferential results and practical implications, which is the focus of that discussion.
+
 *5. Page 21 onward, there are many instances where the figure references did not compile correctly. These need to be corrected.*
+
+Thank you for bringing to our attention the issue with the figure references not compiling correctly from page 21 onward.
+We have carefully reviewed and recompiled the manuscript, ensuring that all figure references are now accurately represented and correctly linked.

### 2024-03-24 appendix.Rmd
R2 submitted

@@
 title: |
-  Appendix: A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol
+  Appendix to "A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol" published in the Journal of Computational and Graphical Statistics
 author:
@@
 
-# Additional details of experimental design
-
-## Non-linearity
-
-The non-linearity model used in the experiment includes a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials on random vector $\boldsymbol{x}$ formulated as
-
-\begin{align*} \label{eq:nonlinearity-model}
-\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
-\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
-\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
-\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
-\end{align*}
-
-\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials [@hermite1864nouveau; originally by @de1820theorie], $\varepsilon \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
-
-```{=tex}
-\begin{equation} \label{eq:scaling-function}
-g(\boldsymbol{x}, k) = 2k\cdot\frac{\boldsymbol{x} - x_{min}\boldsymbol{1}_n}{x_{max} - x_{min}} - k, \quad \text{for} \quad k > 0. 
-\end{equation}
-```
-
-\noindent where $\displaystyle x_{min} = \min_{i\in \{1, ..., n\}} x_i$, $\displaystyle x_{max} = \max_{i\in \{1, ..., n\}} x_i$ and $x_i$ is the $i$-th entry of $\boldsymbol{x}$.
-The function `hermite` from the R package `mpoly` [@mpoly] is used to simulate $\boldsymbol{z}_{raw}$ to generate Hermite polynomials.
-
-The null regression model used to fit the realizations generated by the above model is formulated as
-
-```{=tex}
-\begin{equation} \label{eq:null-model}
-\boldsymbol{y} = \beta_0\boldsymbol{1}_n + \beta_1 \boldsymbol{x} + \boldsymbol{u},
-\end{equation}
-```
-\noindent where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
-Here $\boldsymbol{z}$ is a higher order term of $\boldsymbol{x}$ and leaving it out in the null regression results in model misspecification.
-
-## Heteroskedasticity
-
-The heteroskedasticity model used in the experiment is formulated as
-
-\begin{align*} \label{eq:heter-model}
-\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
-\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
-\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}_n, \boldsymbol{1}_n + (2 - |a|)(\boldsymbol{x} - a\boldsymbol{1}_n)'(\boldsymbol{x} - a\boldsymbol{1}_n) b \boldsymbol{I}_n), 
-\end{align*}
-
-\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vectors of size $n$ and $g(.)$ is the scaling function defined in Equation \ref{eq:scaling-function}.
-The null regression model used to fit the realizations generated by the above model is formulated exactly the same as Equation \ref{eq:null-model}.
-
-For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the predictor $\boldsymbol{x}$, which will lead to the presence of heteroskedasticity.
-
-Since $\text{supp}(X) = [-1, 1]$, choosing $a$ to be $-1$, $0$ and $1$ can generate "left-triangle", "butterfly" and "right-triangle" shapes.
-The term $(2 - |a|)$ maintains the magnitude of residuals across different values of $a$.
-
-## Controlling the signal strength
-
-The three parameters $n$, $\sigma$ and $b$ are important for controlling the strength of the signal to generate lineups with a variety difficulty levels.
-This will ensure that  estimated power curves will be smooth and continuous, and that participants are allocated a set of lineups with a range of difficulty.
-
-Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ are used in data collection periods I and II respectively.
-A large value of $\sigma$ will increase the variation of the error of the non-linearity model and decrease the visibility of the visual pattern.
-The parameter $b$ controls the variation in the standard deviation of the error across the support of the predictor.
-A larger value of $b$ generates a larger ratio between it's smallest and highest values, making the visual pattern more obvious.
-The sample size $n$ sharpens (or blurs) a pattern when it is larger (or smaller).
-
-<!-- Figures \ref{fig:different-sigma}, \ref{fig:different-b} and \ref{fig:different-n} demonstrate the impact of these parameters on signal strength.  -->
-
-### Lineup allocation to participants
-
-From Table \ref{tab:model-factor-table} it can be computed that there are a total of $4 \times 4 \times 3 \times 4 = 192$ and $3 \times 5 \times 3 \times 4 = 180$ combinations of parameter values for the non-linearity model and heteroskedasticity models respectively.
-Three replications for each combination results in $192 \times 3 = 576$ and $180 \times 3 = 540$ lineups, respectively.
-
-Each lineup needs to be evaluated by at least five participants.
-From previous work, and additional pilot studies for this experiment, we decided to record evaluations from 20 lineups for each participant.
-Two of the 20 lineups with clear visual patterns were used as attention checks, to ensure quality data.
-Thus, $576 \times 5 / (20-2) = 160$ and $540 \times 5 / (20-2) = 150$ participants were needed to cover the experimental design for the data collection periods I and II, respectively.
-The factor levels and range of difficulty was assigned relatively equally among participants.
-The Appendix contains graphical summaries of these assignments for each subject.
-
-Data collection period III was primarily to obtain Rorschach lineup evaluations to estimate $\alpha$ (see Section A.1 in the Appendix), and also to obtain additional evaluations of lineups made with uniform fitted value distribution to ensure consistency in the results.
-To construct a Rorschach lineup, the data is generated from a model with zero effect size, while the null data are generated using the same simulation method discussed in Section \ref{visual-test-procedure-based-on-lineups}.
-This procedure differs from that of the canonical Rorschach lineup, where all 20 plots are generated directly from the null model, so the method suggested in @vanderplas2021statistical for typical lineups containing a data plot is used to estimate $\alpha$.
-(The Appendix contains a sensitivity analysis on the effect of uncertainty in the estimation of $\alpha$.) The $3 \times 4 = 12$ treatment levels of the common factors, replicated three times results in 36 lineups.
-And 6 more evaluations on the 279 lineups with uniform fitted value distribution, results in at least $(36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133$ participants needed.
-
-### Collecting results
-
-Participants for all three data collection periods were recruited from the Prolific crowd-sourcing platform [@palan2018prolific].
-Pre-screening procedures were applied during the recruitment: participants were required to be fluent in English, with $98\%$ minimum approval rate and at least 10 submissions in other studies.
-
-In each lineup evaluation, the participant was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots are from others.
-If there is no noticeable difference between plots in a lineup, participants had the option to select zero plots without the need to provide a reason.
-During the process of recording the responses, a zero selection was considered to be equivalent to selecting all 20 plots.
-No participant was shown the same lineup twice.
-Information about preferred pronouns, age group, education, and previous experience in visual experiments were also collected.
-A participant's submission was only included in the analysis if the data plot is identified for at least one attention check.
-
 # Additional details of testing procedures
@@
 In a Rorschach, in theory all plots should be equally likely, but in practice some (irrelevant) visual elements may be more eye-catching than others.
-This is what $alpha$ captures, the capacity for extraneous features to distract the observer for a particular type of plot display.
+This is what $\alpha$ captures, the capacity for extraneous features to distract the observer for a particular type of plot display.
 
@@
 
-Let $\boldsymbol{X}$ denotes the $p + 1$ regressors with $n$ observations, $\boldsymbol{b} = (\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\boldsymbol{y}$ denotes the OLS solution, $\boldsymbol{R} = \boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ denotes the residual operator, and let $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0},\sigma^2\boldsymbol{I})$ denotes the error.
+Let $\boldsymbol{X}$ denotes the $p + 1$ predictors with $n$ observations, $\boldsymbol{b} = (\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\boldsymbol{y}$ denotes the OLS solution, $\boldsymbol{R} = \boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ denotes the residual operator, and let $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0},\sigma^2\boldsymbol{I})$ denotes the error.
 The residual vector
@@
 \text{diag}(\boldsymbol{R}\sigma^2))\right).$$
-Since we assume $\sigma = 1$ for the heteroskedasticity model, the final form of the formula is
-
-$$KL = \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R})|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\text{diag}(\boldsymbol{R}))\right).$$
+
+To compute the effect size for each lineup we simulate a sufficiently large number of samples from the same model with the number of observations $n$ fixed for each sample.
+We then compute the effect size for each sample and take the average as the final value.
+This ensures lineups constructed with the same experimental factors will share the same effect size.
 
@@
 Many of the residual plots not rejected by the conventional test are rejected by the visual test.
-It isn't possible to vary the significance level of the conventional test to match the decisions made by the visual test."}
+It is not possible to vary the significance level of the conventional test to match the decisions made by the visual test."}
 # compute rejection rate for different significance levels, holding visual test constant
@@
 
-# Experiment setup
+# Additional details of experiment setup
+
+## Non-linearity Model
+
+The non-linearity model used in the experiment includes a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials on random vector $\boldsymbol{x}$ formulated as
+
+\begin{align*} \label{eq:nonlinearity-model}
+\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
+\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
+\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
+\end{align*}
+
+\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials [@hermite1864nouveau; originally by @de1820theorie], $\varepsilon \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
+
+```{=tex}
+\begin{equation} \label{eq:scaling-function}
+g(\boldsymbol{x}, k) = 2k\cdot\frac{\boldsymbol{x} - x_{min}\boldsymbol{1}_n}{x_{max} - x_{min}} - k, \quad \text{for} \quad k > 0. 
+\end{equation}
+```
+
+\noindent where $\displaystyle x_{min} = \min_{i\in \{1, ..., n\}} x_i$, $\displaystyle x_{max} = \max_{i\in \{1, ..., n\}} x_i$ and $x_i$ is the $i$-th entry of $\boldsymbol{x}$.
+The function `hermite` from the R package `mpoly` [@mpoly] is used to simulate $\boldsymbol{z}_{raw}$ to generate Hermite polynomials.
+
+The null regression model used to fit the realizations generated by the above model is formulated as
+
+```{=tex}
+\begin{equation} \label{eq:null-model}
+\boldsymbol{y} = \beta_0\boldsymbol{1}_n + \beta_1 \boldsymbol{x} + \boldsymbol{u},
+\end{equation}
+```
+\noindent where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
+Here $\boldsymbol{z}$ is a higher order term of $\boldsymbol{x}$ and leaving it out in the null regression results in model misspecification.
+
+## Heteroskedasticity Model
+
+The heteroskedasticity model used in the experiment is formulated as
+
+\begin{align*} \label{eq:heter-model}
+\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
+\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}_n, \boldsymbol{1}_n + (2 - |a|)(\boldsymbol{x} - a\boldsymbol{1}_n)'(\boldsymbol{x} - a\boldsymbol{1}_n) b \boldsymbol{I}_n), 
+\end{align*}
+
+\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vectors of size $n$ and $g(.)$ is the scaling function defined in Equation \ref{eq:scaling-function}.
+The null regression model used to fit the realizations generated by the above model is formulated exactly the same as Equation \ref{eq:null-model}.
+
+For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the predictor $\boldsymbol{x}$, which will lead to the presence of heteroskedasticity.
+
+Since $\text{supp}(X) = [-1, 1]$, choosing $a$ to be $-1$, $0$ and $1$ can generate "left-triangle", "butterfly" and "right-triangle" shapes.
+The term $(2 - |a|)$ maintains the magnitude of residuals across different values of $a$.
+
+## Controlling the signal strength
+
+The three parameters $n$, $\sigma$ and $b$ are important for controlling the strength of the signal to generate lineups with a variety difficulty levels.
+This will ensure that  estimated power curves will be smooth and continuous, and that participants are allocated a set of lineups with a range of difficulty.
+
+Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ are used in data collection periods I and II respectively.
+A large value of $\sigma$ will increase the variation of the error of the non-linearity model and decrease the visibility of the visual pattern.
+The parameter $b$ controls the variation in the standard deviation of the error across the support of the predictor.
+A larger value of $b$ generates a larger ratio between it is smallest and highest values, making the visual pattern more obvious.
+The sample size $n$ sharpens (or blurs) a pattern when it is larger (or smaller).
+
+## Lineup allocation to participants
+
+There are a total of $4 \times 4 \times 3 \times 4 = 192$ and $3 \times 5 \times 3 \times 4 = 180$ combinations of parameter values for the non-linearity model and heteroskedasticity models respectively.
+Three replications for each combination results in $192 \times 3 = 576$ and $180 \times 3 = 540$ lineups, respectively.
+
+Each lineup needs to be evaluated by at least five participants.
+From previous work, and additional pilot studies for this experiment, we decided to record evaluations from 20 lineups for each participant.
+Two of the 20 lineups with clear visual patterns were used as attention checks, to ensure quality data.
+Thus, $576 \times 5 / (20-2) = 160$ and $540 \times 5 / (20-2) = 150$ participants were needed to cover the experimental design for the data collection periods I and II, respectively.
+The factor levels and range of difficulty was assigned relatively equally among participants.
+
+Data collection period III was primarily to obtain Rorschach lineup evaluations to estimate $\alpha$, and also to obtain additional evaluations of lineups made with uniform fitted value distribution to ensure consistency in the results.
+To construct a Rorschach lineup, the data is generated from a model with zero effect size, while the null data are generated using the residual rotation technique.
+This procedure differs from that of the canonical Rorschach lineup, where all 20 plots are generated directly from the null model, so the method suggested in @vanderplas2021statistical for typical lineups containing a data plot is used to estimate $\alpha$.
+The $3 \times 4 = 12$ treatment levels of the common factors, replicated three times results in 36 lineups.
+And 6 more evaluations on the 279 lineups with uniform fitted value distribution, results in at least $(36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133$ participants needed.
 

### 2024-03-24 appendix.tex
R2 submitted

@@
 
-\title{Appendix: A Plot is Worth a Thousand Tests: Assessing Residual
-Diagnostics with the Lineup Protocol}
+\title{Appendix to ``A Plot is Worth a Thousand Tests: Assessing
+Residual Diagnostics with the Lineup Protocol'' published in the Journal
+of Computational and Graphical Statistics}
 
-\author{\name{Weihao Li$^{a}$, Dianne Cook$^{a}$, Emi
-Tanaka$^{a}$, Susan VanderPlas$^{b}$}
+\author{\name{Weihao Li$^{a}$, Dianne Cook$^{a}$, Emi Tanaka$^{a, b,
+c}$, Susan VanderPlas$^{d}$}
 \affil{$^{a}$Department of Econometrics and Business Statistics, Monash
-University, Clayton, VIC, Australia; $^{b}$Department of Statistics,
-University of Nebraska, Lincoln, Nebraska, USA}
+University, Clayton, VIC, Australia; $^{b}$Biological Data Science
+Institute, Australian National University, Acton, ACT,
+Australia; $^{c}$Research School of Finance, Actuarial Studies and
+Statistics, Australian National University, Acton, ACT,
+Australia; $^{d}$Department of Statistics, University of Nebraska,
+Lincoln, Nebraska, USA}
 }
@@
 
-\hypertarget{additional-details-of-testing-procedures}{%
 \section{Additional details of testing
-procedures}\label{additional-details-of-testing-procedures}}
+procedures}\label{additional-details-of-testing-procedures}
 
-\hypertarget{statistical-significance}{%
-\subsection{Statistical significance}\label{statistical-significance}}
+\subsection{Statistical significance}\label{statistical-significance}
 
@@
 likely, but in practice some (irrelevant) visual elements may be more
-eye-catching than others. This is what \(alpha\) captures, the capacity
+eye-catching than others. This is what \(\alpha\) captures, the capacity
 for extraneous features to distract the observer for a particular type
@@
 
-\hypertarget{effect-size-derivation}{%
-\subsection{Effect size derivation}\label{effect-size-derivation}}
+\subsection{Effect size derivation}\label{effect-size-derivation}
 
@@
 
-Let \(\boldsymbol{X}\) denotes the \(p + 1\) regressors with \(n\)
+Let \(\boldsymbol{X}\) denotes the \(p + 1\) predictors with \(n\)
 observations,
@@
 \[KL = \frac{1}{2}\left(log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}
-\text{diag}(\boldsymbol{R}\sigma^2))\right).\] Since we assume
-\(\sigma = 1\) for the heteroskedasticity model, the final form of the
-formula is
+\text{diag}(\boldsymbol{R}\sigma^2))\right).\]
 
-\[KL = \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R})|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\text{diag}(\boldsymbol{R}))\right).\]
+To compute the effect size for each lineup we simulate a sufficiently
+large number of samples from the same model with the number of
+observations \(n\) fixed for each sample. We then compute the effect
+size for each sample and take the average as the final value. This
+ensures lineups constructed with the same experimental factors will
+share the same effect size.
 
-\hypertarget{sensitivity-analysis-for-alpha}{%
 \subsection{\texorpdfstring{Sensitivity analysis for
-\(\alpha\)}{Sensitivity analysis for \textbackslash alpha}}\label{sensitivity-analysis-for-alpha}}
+\(\alpha\)}{Sensitivity analysis for \textbackslash alpha}}\label{sensitivity-analysis-for-alpha}
 
@@
 
-\hypertarget{effect-of-number-of-evaluations-on-the-power-of-a-visual-test}{%
 \subsection{Effect of number of evaluations on the power of a visual
-test}\label{effect-of-number-of-evaluations-on-the-power-of-a-visual-test}}
+test}\label{effect-of-number-of-evaluations-on-the-power-of-a-visual-test}
 
@@
 
-\hypertarget{power-of-a-reset-test-under-different-auxiliary-regression-formulas}{%
 \subsection{Power of a RESET test under different auxiliary regression
-formulas}\label{power-of-a-reset-test-under-different-auxiliary-regression-formulas}}
+formulas}\label{power-of-a-reset-test-under-different-auxiliary-regression-formulas}
 
@@
 
-\hypertarget{conventional-test-rejection-rate-for-varying-significance-levels}{%
 \subsection{Conventional test rejection rate for varying significance
-levels}\label{conventional-test-rejection-rate-for-varying-significance-levels}}
+levels}\label{conventional-test-rejection-rate-for-varying-significance-levels}
 
@@
 Many of the residual plots not rejected by the conventional test are rejected by the visual test.
-It isn't possible to vary the significance level of the conventional test to match the decisions made by the visual test.}\label{fig:rej-rates}
+It is not possible to vary the significance level of the conventional test to match the decisions made by the visual test.}\label{fig:rej-rates}
 \end{figure}
 
-\hypertarget{experiment-setup}{%
-\section{Experiment setup}\label{experiment-setup}}
+\section{Additional details of experiment
+setup}\label{additional-details-of-experiment-setup}
 
-\hypertarget{mapping-of-participants-to-experimental-factors}{%
+\subsection{Non-linearity Model}\label{non-linearity-model}
+
+The non-linearity model used in the experiment includes a non-linear
+term \(\boldsymbol{z}\) constructed using Hermite polynomials on random
+vector \(\boldsymbol{x}\) formulated as
+
+\begin{align*} \label{eq:nonlinearity-model}
+\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
+\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
+\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
+\end{align*}
+
+\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
+\(\boldsymbol{\varepsilon}\), \(\boldsymbol{x}_{raw}\),
+\(\boldsymbol{z}_{raw}\) are vectors of size \(n\), \(\boldsymbol{1}_n\)
+is a vector of ones of size \(n\), \(He_{j}(.)\) is the \(j\)th-order
+probabilist's Hermite polynomials
+\citetext{\citealp{hermite1864nouveau}; \citealp[originally
+by][]{de1820theorie}},
+\(\varepsilon \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)\), and
+\(g(\boldsymbol{x}, k)\) is a scaling function to enforce the support of
+the random vector to be \([-k, k]^n\) defined as
+
+\begin{equation} \label{eq:scaling-function}
+g(\boldsymbol{x}, k) = 2k\cdot\frac{\boldsymbol{x} - x_{min}\boldsymbol{1}_n}{x_{max} - x_{min}} - k, \quad \text{for} \quad k > 0. 
+\end{equation}
+
+\noindent where
+\(\displaystyle x_{min} = \min_{i\in \{1, ..., n\}} x_i\),
+\(\displaystyle x_{max} = \max_{i\in \{1, ..., n\}} x_i\) and \(x_i\) is
+the \(i\)-th entry of \(\boldsymbol{x}\). The function \texttt{hermite}
+from the R package \texttt{mpoly} \citep{mpoly} is used to simulate
+\(\boldsymbol{z}_{raw}\) to generate Hermite polynomials.
+
+The null regression model used to fit the realizations generated by the
+above model is formulated as
+
+\begin{equation} \label{eq:null-model}
+\boldsymbol{y} = \beta_0\boldsymbol{1}_n + \beta_1 \boldsymbol{x} + \boldsymbol{u},
+\end{equation}
+
+\noindent where
+\(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\).
+Here \(\boldsymbol{z}\) is a higher order term of \(\boldsymbol{x}\) and
+leaving it out in the null regression results in model misspecification.
+
+\subsection{Heteroskedasticity Model}\label{heteroskedasticity-model}
+
+The heteroskedasticity model used in the experiment is formulated as
+
+\begin{align*} \label{eq:heter-model}
+\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
+\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}_n, \boldsymbol{1}_n + (2 - |a|)(\boldsymbol{x} - a\boldsymbol{1}_n)'(\boldsymbol{x} - a\boldsymbol{1}_n) b \boldsymbol{I}_n), 
+\end{align*}
+
+\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
+\(\boldsymbol{\varepsilon}\) are vectors of size \(n\) and \(g(.)\) is
+the scaling function defined in Equation \ref{eq:scaling-function}. The
+null regression model used to fit the realizations generated by the
+above model is formulated exactly the same as Equation
+\ref{eq:null-model}.
+
+For \(b \neq 0\), the variance-covariance matrix of the error term
+\(\boldsymbol{\varepsilon}\) is correlated with the predictor
+\(\boldsymbol{x}\), which will lead to the presence of
+heteroskedasticity.
+
+Since \(\text{supp}(X) = [-1, 1]\), choosing \(a\) to be \(-1\), \(0\)
+and \(1\) can generate ``left-triangle'', ``butterfly'' and
+``right-triangle'' shapes. The term \((2 - |a|)\) maintains the
+magnitude of residuals across different values of \(a\).
+
+\subsection{Controlling the signal
+strength}\label{controlling-the-signal-strength}
+
+The three parameters \(n\), \(\sigma\) and \(b\) are important for
+controlling the strength of the signal to generate lineups with a
+variety difficulty levels. This will ensure that estimated power curves
+will be smooth and continuous, and that participants are allocated a set
+of lineups with a range of difficulty.
+
+Parameter \(\sigma \in \{0.5, 1, 2, 4\}\) and
+\(b \in \{0.25, 1, 4, 16, 64\}\) are used in data collection periods I
+and II respectively. A large value of \(\sigma\) will increase the
+variation of the error of the non-linearity model and decrease the
+visibility of the visual pattern. The parameter \(b\) controls the
+variation in the standard deviation of the error across the support of
+the predictor. A larger value of \(b\) generates a larger ratio between
+it is smallest and highest values, making the visual pattern more
+obvious. The sample size \(n\) sharpens (or blurs) a pattern when it is
+larger (or smaller).
+
+\subsection{Lineup allocation to
+participants}\label{lineup-allocation-to-participants}
+
+There are a total of \(4 \times 4 \times 3 \times 4 = 192\) and
+\(3 \times 5 \times 3 \times 4 = 180\) combinations of parameter values
+for the non-linearity model and heteroskedasticity models respectively.
+Three replications for each combination results in
+\(192 \times 3 = 576\) and \(180 \times 3 = 540\) lineups, respectively.
+
+Each lineup needs to be evaluated by at least five participants. From
+previous work, and additional pilot studies for this experiment, we
+decided to record evaluations from 20 lineups for each participant. Two
+of the 20 lineups with clear visual patterns were used as attention
+checks, to ensure quality data. Thus, \(576 \times 5 / (20-2) = 160\)
+and \(540 \times 5 / (20-2) = 150\) participants were needed to cover
+the experimental design for the data collection periods I and II,
+respectively. The factor levels and range of difficulty was assigned
+relatively equally among participants.
+
+Data collection period III was primarily to obtain Rorschach lineup
+evaluations to estimate \(\alpha\), and also to obtain additional
+evaluations of lineups made with uniform fitted value distribution to
+ensure consistency in the results. To construct a Rorschach lineup, the
+data is generated from a model with zero effect size, while the null
+data are generated using the residual rotation technique. This procedure
+differs from that of the canonical Rorschach lineup, where all 20 plots
+are generated directly from the null model, so the method suggested in
+\citet{vanderplas2021statistical} for typical lineups containing a data
+plot is used to estimate \(\alpha\). The \(3 \times 4 = 12\) treatment
+levels of the common factors, replicated three times results in 36
+lineups. And 6 more evaluations on the 279 lineups with uniform fitted
+value distribution, results in at least
+\((36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133\) participants
+needed.
+
 \subsection{Mapping of participants to experimental
-factors}\label{mapping-of-participants-to-experimental-factors}}
+factors}\label{mapping-of-participants-to-experimental-factors}
 
@@
 
-\hypertarget{data-collection-process}{%
-\subsection{Data collection process}\label{data-collection-process}}
+\subsection{Data collection process}\label{data-collection-process}
 
@@
 
-\hypertarget{analysis-of-results-relative-to-data-collection-process}{%
 \section{Analysis of results relative to data collection
-process}\label{analysis-of-results-relative-to-data-collection-process}}
+process}\label{analysis-of-results-relative-to-data-collection-process}
 
-\hypertarget{demographics}{%
-\subsection{Demographics}\label{demographics}}
+\subsection{Demographics}\label{demographics}
 
@@
 
-\hypertarget{data-collection-periods}{%
-\subsection{Data collection periods}\label{data-collection-periods}}
+\subsection{Data collection periods}\label{data-collection-periods}
 

### 2024-03-24 paper.Rmd
R2 submitted

@@
   A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol
-type: ARTICLE TEMPLATE
 author:
@@
 The lineup protocol also detects a range of departures from good residuals simultaneously.
+Supplemental materials for the article are available online.
 keywords: |
@@
 For residual analysis this is especially helpful for gauging whether there is \emph{no} pattern.
-(Figure \ref{fig:first-example-lineup} shows an example of a lineup of residual plots.) In its strict use, one would insist that the data plot is not seen before seeing the lineup, so that the observer doesn't know which is the true plot.
+(Figure \ref{fig:first-example-lineup} shows an example of a lineup of residual plots.) In its strict use, one would insist that the data plot is not seen before seeing the lineup, so that the observer does not know which is the true plot.
 When used this way, it provides an objective test for data plots.
@@
 Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot.
-A small $p$-value would result from a substantial number selecting the data plot (at position $2^2 + 2$, exhibiting non-linearity)."}
+A small $p$-value would result from a substantial number selecting the data plot (at position $6$, exhibiting non-linearity)."}
 
@@
 
-```{r residual-plot-common-departures, fig.pos="t!", fig.width = 10, fig.height = 2.5, fig.cap = "Example residual vs fitted value plots (red line indicates 0): (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
+```{r residual-plot-common-departures, fig.pos="t!", fig.width = 10, fig.height = 2.5, fig.cap = "Example residual vs fitted value plots (horizontal line indicates 0): (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
 The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted value plot."}
@@
 The SW test was computed using the `shapiro.test` from the core R package `stats`.
-^[Although we didn't use it, it's useful to know that the R package `skedastic` [@skedastic] also contains a large collection of functions to test for heteroskedasticity.] The RESET test requires the selection of a power parameter.
+^[Although we did not use it, it is useful to know that the R package `skedastic` [@skedastic] also contains a large collection of functions to test for heteroskedasticity.] The RESET test requires the selection of a power parameter.
 @ramsey1969tests recommends a power of four, which we adopted in our analysis.
@@
 For residual plots in Figure \ref{fig:residual-plot-common-departures}, we would expect the RESET test for non-linearity to reject residual plot B, the BP test for heteroskedasticity to reject the residual plot C, and the SW test for non-normality to reject residual plot D, which they all do and all tests also correctly fail to reject residual plot A.
-Interestingly, the BP and SW tests also reject the residual plots exhibiting structure that they weren't designed for.
+Interestingly, the BP and SW tests also reject the residual plots exhibiting structure that they were not designed for.
 @cook1982residuals explain that most residual-based tests for a particular departure from the model assumptions are also sensitive to other types of departures.
@@
 A key goal of residual plot diagnostics is to identify potential issues that could lead to incorrect conclusions or errors in subsequent analyses, but minor defects in the model are unlikely to have a significant impact and may be best disregarded for practical purposes.
-The experiment discussed in this paper specifically addresses this tension between statistical and practical significance.
+The experiment discussed in this paper specifically addresses this tension between statistical significance and practical significance.
 
@@
 Figure \ref{fig:first-example-lineup} is an example of a lineup protocol.
-If the data plot at position $2^2 + 2$ is identifiable, then it is evidence for the rejection of $H_0$.
+If the data plot at position $6$ is identifiable, then it is evidence for the rejection of $H_0$.
 In fact, the actual residual plot is obtained from a misspecified regression model with missing non-linear terms.
@@
                                  "Size (sym6)", 
-                                 "Distribution of the predictor")) %>% #"Sample size\n(sym6)")) %>%
+                                 "Distribution of the fitted values")) %>% #"Sample size\n(sym6)")) %>%
   kableExtra::add_header_above(c("Non-linearity" = 2, 
@@
 Figure \ref{fig:different-b} shows the butterfly shape as the ratio parameter ($b$) is changed.
-More details about the simulation process are provided in the Appendix.
+More details about the simulation process are provided in Appendix B.
 
@@
 Figure \ref{fig:different-dist} shows the non-linear pattern, a "U" shape, with the different fitted value distributions.
-We would expect that structure in residual plots would be easier to perceive when the predictors are uniformly distributed.
+We would expect that structure in residual plots would be easier to perceive when the fitted values are uniformly distributed.
 
@@
 
-For the purposes of our calculations we have chosen to use an approach based on Kullback-Leibler divergence [@kullback1951information], coupled with simulation.
-This formulation defines effect size to be
+For the purposes of our calculations we have chosen to use an approach based on Kullback-Leibler divergence [@kullback1951information]. This formulation defines effect size to be
 
@@
 $\boldsymbol{R} = \boldsymbol{I}_n - \boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the residual operator, $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals where $\boldsymbol{Z}$ contains any higher order terms of $\boldsymbol{X}$ left out of the regression equation, $\boldsymbol{\beta}_z$ contains the corresponding coefficients, and $\sigma^2\boldsymbol{I}_n$ is the assumed covariance matrix of the error term when $H_0$ is true.
-
-To compute the effect size for each lineup we simulate a sufficiently large number of samples from the same model with the number of observations $n$ fixed for each sample.
-We then compute the effect size for each sample and take the average as the final value.
-This ensures lineups constructed with the same experimental factors will share the same effect size.
+More details about the effect size derivation are provided in Appendix A.
 
@@
 Figure \ref{fig:nonlinearheterpower} present the power curves of various tests plotted against the effect size in the residuals for non-linearity and heteroskedasticity.
-In each case the power of visual test is calculated for multiple bootstrap samples leading to the many (orange) curves.
+In each case the power of visual test is calculated for multiple bootstrap samples leading to the many (solid orange) curves.
 The effect size was computed at a 5% significance level and plotted on a natural logarithmic scale.
@@
   mutate(power = map(mod, function(mod) {
-    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
+    result <- data.frame(effect_size = exp(seq(log(min_poly_es), log(max_poly_es), 0.1)),
                          offset0 = log(0.05/0.95))
@@
 # The prediction of the visual test
-pred_visual_poly <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
+pred_visual_poly <- data.frame(effect_size = exp(seq(log(min_poly_es), log(max_poly_es), 0.1)),
                                offset0 = log(0.05/0.95)) %>%
@@
     mutate(power = map(mod, function(mod) {
-      result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
+      result <- data.frame(effect_size = exp(seq(log(min_poly_es), log(max_poly_es), 0.1)),
                            offset0 = log(0.05/0.95))
@@
   geom_line(data = pred_conv_poly,
-            aes(log_effect_size, power, col = name),
-            size = 1) +
+            aes(log_effect_size, power, col = name, linetype = name)) +
 
@@
   geom_line(data = pred_visual_poly_boot,
-            aes(log_effect_size, power, col = "visual test", group = boot_id),
-            size = 1,
+            aes(log_effect_size, power, col = "visual test", group = boot_id, linetype = "visual test"),
             alpha = 0.01) +
@@
   geom_line(data = pred_visual_poly,
-            aes(log_effect_size, power, col = "visual test"), 
-            size = 1) +
+            aes(log_effect_size, power, col = "visual test", linetype = "visual test")) +
 
@@
   theme_light(base_size = 5) +
-  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
+  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.3)) +
   scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
   scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
+  scale_linetype_manual(values = c("visual test" = 1, "RESET test" = 2, "BP test" = 3, "SW test" = 4)) +
   xlab(expression(log[e] (Effect_size))) +
   ylab("Power") +
-  labs(col = "", size = "# lineups")
+  labs(col = "", linetype = "")
 ```
@@
   mutate(power = map(mod, function(mod) {
-    result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
+    result <- data.frame(effect_size = exp(seq(log(min_heter_es), log(max_heter_es), 0.1)),
                          offset0 = log(0.05/0.95))
@@
 # The prediction of the visual test
-pred_visual_heter <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
+pred_visual_heter <- data.frame(effect_size = exp(seq(log(min_heter_es), log(max_heter_es), 0.1)),
                                 offset0 = log(0.05/0.95)) %>%
@@
     mutate(power = map(mod, function(mod) {
-      result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
+      result <- data.frame(effect_size = exp(seq(log(min_heter_es), log(max_heter_es), 0.1)),
                            offset0 = log(0.05/0.95))
@@
   geom_line(data = pred_conv_heter,
-            aes(log_effect_size, power, col = name),
-            size = 1) +
+            aes(log_effect_size, power, col = name, linetype = name)) +
 
@@
   geom_line(data = pred_visual_heter_boot,
-            aes(log_effect_size, power, col = "visual test", group = boot_id),
-            size = 1,
+            aes(log_effect_size, power, col = "visual test", group = boot_id, linetype = "visual test"),
             alpha = 0.01) +
@@
   geom_line(data = pred_visual_heter,
-            aes(log_effect_size, power, col = "visual test"), 
-            size = 1) +
+            aes(log_effect_size, power, col = "visual test", linetype = "visual test")) +
 
@@
   theme_light(base_size = 5) +
-  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
+  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.3)) +
   scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
-  # scale_size_manual(values = seq(1, 9)) +
+  scale_linetype_manual(values = c("visual test" = 1, "RESET test" = 2, "BP test" = 3, "SW test" = 4)) +
   scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
@@
   ylab("Power") +
-  labs(col = "", size = "# lineups")
+  labs(col = "", linetype = "")
 ```
@@
 The conventional tests do have higher power for the patterns they are designed to detect, but they typically fail to detect other patterns unless those patterns are particularly strong.
-The visual test doesn't require specifying the pattern ahead of time, relying purely on whether the observed residual plot is detectably different from "good" residual plots.
+The visual test does not require specifying the pattern ahead of time, relying purely on whether the observed residual plot is detectably different from "good" residual plots.
 They will perform equally well regardless of the type of model defect.
@@
 
-The power comparison demonstrates that the appropriate conventional tests will reject more aggressively than visual tests, but we don't know how the decisions for each lineup would agree or disagree.
+The power comparison demonstrates that the appropriate conventional tests will reject more aggressively than visual tests, but we do not know how the decisions for each lineup would agree or disagree.
 Here we compare the reject or fail to reject decisions of these tests, across all the lineups.
@@
 Because the power curve of the visual tests are a shift to the right of the conventional test (Figure \ref{fig:nonlinearheterpower}) we examined whether adjusting the significance level (to .001, .0001, .00001, ...) of the conventional test would generate similar decisions to that of the visual test.
-Interestingly, it doesn't: despite resulting in less rejections, neither the RESET or BP tests come to complete agreement with the visual test (see Appendix).
+Interestingly, it does not: despite resulting in less rejections, neither the RESET or BP tests come to complete agreement with the visual test (see Appendix A).
 
@@
 This illustrates the sensitivity of the conventional test to the parameter choice, and highlights a limitation: it helps to know the data generating process to set the parameters for the test, which is unrealistic in practice.
-However, we examined this in more detail (see Appendix) and found that there is no harm in setting the parameter higher than four on the tests' operation for lower order polynomial shapes.
+However, we examined this in more detail (see Appendix A) and found that there is no harm in setting the parameter higher than four on the tests' operation for lower order polynomial shapes.
 Using a parameter value of six, instead of four, yields higher power regardless of generating process, and is recommended.
@@
   ggplot() +
-  geom_line(aes(log_effect_size, power, col = shape), size = 1) +
+  geom_line(aes(log_effect_size, power, col = shape, linetype = shape), size = 1) +
   xlab(expression(log[e] (Effect_size))) +
@@
   scale_color_manual(values = rcartocolor::carto_pal(4, "TealGrn")) +
+  scale_linetype_manual(values = c("U" = 1, "S" = 2, "M" = 3, "triple-U" = 4)) +
   theme_light() +
-  theme(legend.position = "bottom") +
+  theme(legend.position = "bottom", legend.key.width = unit(2.5, "line")) +
   facet_wrap(~conv_or_not) +
-  labs(col = "Non-linearity shape")
+  labs(col = "Non-linearity shape", linetype = "Non-linearity shape")
 ```
@@
   ggplot() +
-  geom_line(aes(log_effect_size, power, col = a), size = 1) +
+  geom_line(aes(log_effect_size, power, col = a, linetype = a), size = 1) +
   xlab(expression(log[e] (Effect_size))) +
@@
   scale_color_manual(values = c(rcartocolor::carto_pal(3, "Earth")[1], "grey", rcartocolor::carto_pal(3, "Earth")[3])) +
+  scale_linetype_manual(values = c("left-triangle" = 1, "butterfly" = 2, "right-triangle" = 3)) +
   theme_light() +
-  theme(legend.position = "bottom") +
+  theme(legend.position = "bottom", legend.key.width = unit(2.5, "line")) +
   facet_wrap(~conv_or_not) +
-  labs(col = "Heteroskedasticity shape")
+  labs(col = "Heteroskedasticity shape", linetype = "Heteroskedasticity shape")
 ```
@@
     this_max_es <- ifelse(type == "heteroskedasticity", max_heter_es, max_poly_es)
-    result <- data.frame(effect_size = seq(this_min_es[1], this_max_es[1], 0.1),
+    result <- data.frame(effect_size = exp(seq(log(this_min_es[1]), log(this_max_es[1]), 0.1)),
                          offset0 = log(0.05/0.95))
@@
 Power curves of visual tests are estimated using five evaluations on each lineup.
-For lineups with a uniform fitted value distribution, the five evaluations are repeatedly sampled from the total eleven evaluations to give multiple power curves (grey).
+For lineups with a uniform fitted value distribution, the five evaluations are repeatedly sampled from the total eleven evaluations to give multiple power curves (solid grey).
 Surprisingly, the fitted value distribution has produces more variability in the power of conventional tests than visual tests.
-Uneven distributions, normal and lognormal distributions, tend to yield lower power.", dev = 'png', dpi = 450}
+Uneven distributions, normal and skewed distributions, tend to yield lower power.", dev = 'png', dpi = 450}
 
@@
                        levels = c("Non-linearity", "Heteroskedasticity"))) %>%
+  mutate(x_dist = ifelse(x_dist == "lognormal", "skewed", x_dist)) %>%
   mutate(x_dist = factor(x_dist, 
@@
                                     "normal", 
-                                    "lognormal", 
+                                    "skewed", 
                                     "discrete"))) %>%
@@
   # Draw all smooth lines
-  geom_line(aes(log_effect_size, power, col = x_dist),
+  geom_line(aes(log_effect_size, power, col = x_dist, linetype = x_dist),
             size = 1) +
@@
   geom_line(data = dat_uniform_5_pred,
-            aes(log_effect_size, power, col = "uniform", group = boot_id),
-            size = 1,
+            aes(log_effect_size, power, col = "uniform", group = boot_id, linetype = "uniform"),
+            size = 0.5,
             alpha = 0.01) +
@@
   scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(4,1,3,2)]) +
+  scale_linetype_manual(values = c("uniform" = 1, "normal" = 2, "skewed" = 3, "discrete" = 4)) + 
   xlab(expression(log[e] (Effect_size))) +
   ylab("Power") +
-  labs(col = "", linetype = "Test") +
+  labs(col = "Fitted value distribution", linetype = "Fitted value distribution") +
   facet_grid(type~conv_or_not, scales = "free_x") +
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom", legend.key.width = unit(2.5, "line"))
 ```
@@
 
-Perhaps surprisingly, the visual tests have more consistent power across the different fitted value distributions: for the non-linear pattern, there is almost no power difference, and for the heteroskedastic pattern, uniform and discrete have higher power than normal and lognormal.
+Perhaps surprisingly, the visual tests have more consistent power across the different fitted value distributions: for the non-linear pattern, there is almost no power difference, and for the heteroskedastic pattern, uniform and discrete have higher power than normal and skewed.
 The likely reason is that these latter two have fewer observations in the tails where the heteroskedastic pattern needs to be detected.
@@
 
-# Supplementary material {-}
-
-The supplementary material is available at https://github.com/TengMCing/lineup_residual_diagnostics/blob/master/appendix.pdf.
-It includes more details about the experimental setup, the derivation of the effect size, the effect of data collection period, and the estimate of $\alpha$.
-The R package `visage` is currently available at https://github.com/TengMCing/visage.
+# Supplementary materials {-}
+
+\begin{description}
+\item{Appendix:} The appendix includes more details about the experiment setup, the derivation of the effect size, the effect of data collection period, and the estimate of $\alpha$.
+(appendix.pdf, Portable Document Format file)
+\item{R package ``visage":} The R package ``visage" containing code to simulate lineups used in the experiment.
+The package also contains de-identified data collected from the experiment.
+(visage\_0.1.2.tar.gz, GNU zipped tar file)
+\end{description}

### 2024-03-24 paper.tex
R2 submitted

@@
 
-\articletype{ARTICLE TEMPLATE}
+\articletype{}
 
@@
 contaminated data. The lineup protocol also detects a range of
-departures from good residuals simultaneously.
+departures from good residuals simultaneously. Supplemental materials
+for the article are available online.
 \end{abstract}
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
@@
 plots.) In its strict use, one would insist that the data plot is not
-seen before seeing the lineup, so that the observer doesn't know which
+seen before seeing the lineup, so that the observer does not know which
 is the true plot. When used this way, it provides an objective test for
@@
 Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot.
-A small $p$-value would result from a substantial number selecting the data plot (at position $2^2 + 2$, exhibiting non-linearity).}\label{fig:first-example-lineup}
-\end{figure}
-
-\hypertarget{background}{%
-\section{Background}\label{background}}
-
-\hypertarget{departures-from-good-residual-plots}{%
+A small $p$-value would result from a substantial number selecting the data plot (at position $6$, exhibiting non-linearity).}\label{fig:first-example-lineup}
+\end{figure}
+
+\section{Background}\label{background}
+
 \subsection{Departures from good residual
-plots}\label{departures-from-good-residual-plots}}
+plots}\label{departures-from-good-residual-plots}
 
@@
 
-\caption{Example residual vs fitted value plots (red line indicates 0): (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
+\caption{Example residual vs fitted value plots (horizontal line indicates 0): (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
 The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted value plot.}\label{fig:residual-plot-common-departures}
@@
 
-\hypertarget{conventionally-testing-for-departures}{%
 \subsection{Conventionally testing for
-departures}\label{conventionally-testing-for-departures}}
+departures}\label{conventionally-testing-for-departures}
 
@@
 computed using the \texttt{shapiro.test} from the core R package
-\texttt{stats}. \footnote{Although we didn't use it, it's useful to know
-  that the R package \texttt{skedastic} \citep{skedastic} also contains
-  a large collection of functions to test for heteroskedasticity.} The
-RESET test requires the selection of a power parameter.
-\citet{ramsey1969tests} recommends a power of four, which we adopted in
-our analysis.
+\texttt{stats}. \footnote{Although we did not use it, it is useful to
+  know that the R package \texttt{skedastic} \citep{skedastic} also
+  contains a large collection of functions to test for
+  heteroskedasticity.} The RESET test requires the selection of a power
+parameter. \citet{ramsey1969tests} recommends a power of four, which we
+adopted in our analysis.
 
@@
 Interestingly, the BP and SW tests also reject the residual plots
-exhibiting structure that they weren't designed for.
+exhibiting structure that they were not designed for.
 \citet{cook1982residuals} explain that most residual-based tests for a
@@
 discussed in this paper specifically addresses this tension between
-statistical and practical significance.
+statistical significance and practical significance.
 
@@
 
-\hypertarget{visual-test-procedure-based-on-lineups}{%
 \subsection{Visual test procedure based on
-lineups}\label{visual-test-procedure-based-on-lineups}}
+lineups}\label{visual-test-procedure-based-on-lineups}
 
@@
 Figure \ref{fig:first-example-lineup} is an example of a lineup
-protocol. If the data plot at position \(2^2 + 2\) is identifiable, then
-it is evidence for the rejection of \(H_0\). In fact, the actual
-residual plot is obtained from a misspecified regression model with
-missing non-linear terms.
+protocol. If the data plot at position \(6\) is identifiable, then it is
+evidence for the rejection of \(H_0\). In fact, the actual residual plot
+is obtained from a misspecified regression model with missing non-linear
+terms.
 
@@
 
-\hypertarget{significance-calculation}{%
 \section{Calculation of statistical significance and test
-power}\label{significance-calculation}}
-
-\hypertarget{what-is-being-tested}{%
-\subsection{What is being tested?}\label{what-is-being-tested}}
+power}\label{significance-calculation}
+
+\subsection{What is being tested?}\label{what-is-being-tested}
 
@@
 
-\hypertarget{statistical-significance}{%
 \subsection{\texorpdfstring{Statistical
-significance\label{sig}}{Statistical significance}}\label{statistical-significance}}
+significance\label{sig}}{Statistical significance}}\label{statistical-significance}
 
@@
 
-\hypertarget{power-of-the-tests}{%
-\subsection{Power of the tests}\label{power-of-the-tests}}
+\subsection{Power of the tests}\label{power-of-the-tests}
 
@@
 
-\hypertarget{experimental-design}{%
-\section{Experimental design}\label{experimental-design}}
+\section{Experimental design}\label{experimental-design}
 
@@
 \cmidrule(l{3pt}r{3pt}){1-2} \cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6}
-\multicolumn{1}{c}{Poly Order ($j$)} & \multicolumn{1}{c}{SD ($\sigma$)} & \multicolumn{1}{c}{Shape ($a$)} & \multicolumn{1}{c}{Ratio ($b$)} & \multicolumn{1}{c}{Size ($n$)} & \multicolumn{1}{c}{Distribution of the predictor} \\
+\multicolumn{1}{c}{Poly Order ($j$)} & \multicolumn{1}{c}{SD ($\sigma$)} & \multicolumn{1}{c}{Shape ($a$)} & \multicolumn{1}{c}{Ratio ($b$)} & \multicolumn{1}{c}{Size ($n$)} & \multicolumn{1}{c}{Distribution of the fitted values} \\
 \cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
@@
 
-\hypertarget{simulating-departures-from-good-residuals}{%
 \subsection{Simulating departures from good
-residuals}\label{simulating-departures-from-good-residuals}}
-
-\hypertarget{non-linearity-and-heteroskedasticity}{%
+residuals}\label{simulating-departures-from-good-residuals}
+
 \subsubsection{Non-linearity and
-Heteroskedasticity}\label{non-linearity-and-heteroskedasticity}}
+Heteroskedasticity}\label{non-linearity-and-heteroskedasticity}
 
@@
 the butterfly shape as the ratio parameter (\(b\)) is changed. More
-details about the simulation process are provided in the Appendix.
+details about the simulation process are provided in Appendix B.
 
@@
 
-\hypertarget{factors-common-to-both-data-collection-periods}{%
 \subsubsection{Factors common to both data collection
-periods}\label{factors-common-to-both-data-collection-periods}}
+periods}\label{factors-common-to-both-data-collection-periods}
 
@@
 that structure in residual plots would be easier to perceive when the
-predictors are uniformly distributed.
+fitted values are uniformly distributed.
 
@@
 
-\hypertarget{effect-size}{%
-\subsection{Effect size}\label{effect-size}}
+\subsection{Effect size}\label{effect-size}
 
@@
 For the purposes of our calculations we have chosen to use an approach
-based on Kullback-Leibler divergence \citep{kullback1951information},
-coupled with simulation. This formulation defines effect size to be
+based on Kullback-Leibler divergence \citep{kullback1951information}.
+This formulation defines effect size to be
 
@@
 coefficients, and \(\sigma^2\boldsymbol{I}_n\) is the assumed covariance
-matrix of the error term when \(H_0\) is true.
-
-To compute the effect size for each lineup we simulate a sufficiently
-large number of samples from the same model with the number of
-observations \(n\) fixed for each sample. We then compute the effect
-size for each sample and take the average as the final value. This
-ensures lineups constructed with the same experimental factors will
-share the same effect size.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
+matrix of the error term when \(H_0\) is true. More details about the
+effect size derivation are provided in Appendix A.
+
+\section{Results}\label{results}
 
@@
 
-\hypertarget{power-comparison-of-the-tests}{%
 \subsection{\texorpdfstring{Power comparison of the
-tests\label{power-analysis}}{Power comparison of the tests}}\label{power-comparison-of-the-tests}}
+tests\label{power-analysis}}{Power comparison of the tests}}\label{power-comparison-of-the-tests}
 
@@
 and heteroskedasticity. In each case the power of visual test is
-calculated for multiple bootstrap samples leading to the many (orange)
-curves. The effect size was computed at a 5\% significance level and
-plotted on a natural logarithmic scale. To facilitate visual calibration
-of effect size values with the corresponding diagnostic plots, a
-sequence of example residual plots with increasing effect sizes is
-provided at the bottom of these figures. These plots serve as a visual
-aid to help readers understand how different effect size values
+calculated for multiple bootstrap samples leading to the many (solid
+orange) curves. The effect size was computed at a 5\% significance level
+and plotted on a natural logarithmic scale. To facilitate visual
+calibration of effect size values with the corresponding diagnostic
+plots, a sequence of example residual plots with increasing effect sizes
+is provided at the bottom of these figures. These plots serve as a
+visual aid to help readers understand how different effect size values
 translate to changes in the diagnostic plots. The horizontal lines of
@@
 detect other patterns unless those patterns are particularly strong. The
-visual test doesn't require specifying the pattern ahead of time,
+visual test does not require specifying the pattern ahead of time,
 relying purely on whether the observed residual plot is detectably
@@
 
-\hypertarget{comparison-of-test-decisions-based-on-p-values}{%
 \subsection{\texorpdfstring{Comparison of test decisions based on
-\(p\)-values\label{p-value}}{Comparison of test decisions based on p-values}}\label{comparison-of-test-decisions-based-on-p-values}}
+\(p\)-values\label{p-value}}{Comparison of test decisions based on p-values}}\label{comparison-of-test-decisions-based-on-p-values}
 
 The power comparison demonstrates that the appropriate conventional
-tests will reject more aggressively than visual tests, but we don't know
-how the decisions for each lineup would agree or disagree. Here we
+tests will reject more aggressively than visual tests, but we do not
+know how the decisions for each lineup would agree or disagree. Here we
 compare the reject or fail to reject decisions of these tests, across
@@
 \ldots) of the conventional test would generate similar decisions to
-that of the visual test. Interestingly, it doesn't: despite resulting in
-less rejections, neither the RESET or BP tests come to complete
-agreement with the visual test (see Appendix).
-
-\hypertarget{effect-of-amount-of-non-linearity}{%
+that of the visual test. Interestingly, it does not: despite resulting
+in less rejections, neither the RESET or BP tests come to complete
+agreement with the visual test (see Appendix A).
+
 \subsection{\texorpdfstring{Effect of amount of
-non-linearity\label{nonlin-analysis}}{Effect of amount of non-linearity}}\label{effect-of-amount-of-non-linearity}}
+non-linearity\label{nonlin-analysis}}{Effect of amount of non-linearity}}\label{effect-of-amount-of-non-linearity}
 
@@
 unrealistic in practice. However, we examined this in more detail (see
-Appendix) and found that there is no harm in setting the parameter
+Appendix A) and found that there is no harm in setting the parameter
 higher than four on the tests' operation for lower order polynomial
@@
 
-\hypertarget{effect-of-shape-of-heteroskedasticity}{%
 \subsection{\texorpdfstring{Effect of shape of
-heteroskedasticity\label{hetero-analysis}}{Effect of shape of heteroskedasticity}}\label{effect-of-shape-of-heteroskedasticity}}
+heteroskedasticity\label{hetero-analysis}}{Effect of shape of heteroskedasticity}}\label{effect-of-shape-of-heteroskedasticity}
 
@@
 
-\hypertarget{effect-of-fitted-value-distributions}{%
 \subsection{Effect of fitted value
-distributions}\label{effect-of-fitted-value-distributions}}
+distributions}\label{effect-of-fitted-value-distributions}
 
@@
 Power curves of visual tests are estimated using five evaluations on each lineup.
-For lineups with a uniform fitted value distribution, the five evaluations are repeatedly sampled from the total eleven evaluations to give multiple power curves (grey).
+For lineups with a uniform fitted value distribution, the five evaluations are repeatedly sampled from the total eleven evaluations to give multiple power curves (solid grey).
 Surprisingly, the fitted value distribution has produces more variability in the power of conventional tests than visual tests.
-Uneven distributions, normal and lognormal distributions, tend to yield lower power.}\label{fig:different-x-dist-poly-power}
+Uneven distributions, normal and skewed distributions, tend to yield lower power.}\label{fig:different-x-dist-poly-power}
 \end{figure}
@@
 there is almost no power difference, and for the heteroskedastic
-pattern, uniform and discrete have higher power than normal and
-lognormal. The likely reason is that these latter two have fewer
-observations in the tails where the heteroskedastic pattern needs to be
-detected.
+pattern, uniform and discrete have higher power than normal and skewed.
+The likely reason is that these latter two have fewer observations in
+the tails where the heteroskedastic pattern needs to be detected.
 
@@
 
-\hypertarget{limitations-and-practicality}{%
 \section{Limitations and
-practicality}\label{limitations-and-practicality}}
+practicality}\label{limitations-and-practicality}
 
@@
 
-\hypertarget{conclusions}{%
-\section{Conclusions}\label{conclusions}}
+\section{Conclusions}\label{conclusions}
 
@@
 
-\hypertarget{acknowledgements}{%
-\section*{Acknowledgements}\label{acknowledgements}}
+\section*{Acknowledgements}\label{acknowledgements}
 \addcontentsline{toc}{section}{Acknowledgements}
@@
 
-\hypertarget{supplementary-material}{%
-\section*{Supplementary material}\label{supplementary-material}}
-\addcontentsline{toc}{section}{Supplementary material}
-
-The supplementary material is available at
-\url{https://github.com/TengMCing/lineup_residual_diagnostics/blob/master/appendix.pdf}.
-It includes more details about the experimental setup, the derivation of
-the effect size, the effect of data collection period, and the estimate
-of \(\alpha\). The R package \texttt{visage} is currently available at
-\url{https://github.com/TengMCing/visage}.
+\section*{Supplementary materials}\label{supplementary-materials}
+\addcontentsline{toc}{section}{Supplementary materials}
+
+\begin{description}
+\item{Appendix:} The appendix includes more details about the experiment setup, the derivation of the effect size, the effect of data collection period, and the estimate of $\alpha$.
+(appendix.pdf, Portable Document Format file)
+\item{R package ``visage":} The R package ``visage" containing code to simulate lineups used in the experiment.
+The package also contains de-identified data collected from the experiment.
+(visage\_0.1.2.tar.gz, GNU zipped tar file)
+\end{description}
 
